### 2018-07-09 src/EC50aggregation.Rmd
publishing orphan commit

---
title: "EC50 calculation + calculation and upload"
author: "Andreas Scharmüller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
editor_options:
  chunk_output_type: console
---

The following script collects different ecotoxicological tests and extracts EC50 values for multiple organisms for different pesticide compounds. The compounds found in the __psm-data.frame__ are compiled in the EC50sensitivity_data.Rmd script and loaded here in a first step. As a first source for ecotoxicological test data the EPA ECOTOX data base is queried (data-epa). The data itself is stored in a PostgreSQL data base built accoring to [this guide](http://edild.github.io/localecotox/). Secondly the bfg_monitoring data-base is queried specifically for values of __Daphnia magna__ tests. The vast amount of test therein was collected by Eduard Szöcs. In a third step data from the [PPDB](https://sitem.herts.ac.uk/aeru/ppdb/en/search.htm) is queried. This is not in accordance to their data policy, but it works. After quering the different data bases, habitat information on family level which was compiled by Verena, Gesa and me is added to the test data set (epa1). The acquire test data is then aggregated (min, percentile, mean) for single substances (aggregate-ec50-all) over all organisms and over freshwater organisms only (aggregate-ec50-freshwater). In a last step the aggregated test data is uploaded to a PostgreSQL data base.

# Setup

```{r setup}
#rm(list = ls())
# switches
online = FALSE
local = FALSE # have you built the EPA ECOTOX DB locally?
plots = FALSE

# packages
require(data.table)
require(RPostgreSQL)
require(taxize)
require(webchem)
require(ggplot2)
require(ggrepel)
require(vegan)

# variables
cachedir = '/home/andreas/Documents/Projects/species-sensitivity/cache'
fundir = '/home/andreas/Documents/Projects/species-sensitivity/functions'
plotdir = '/home/andreas/Documents/Projects/species-sensitivity/plots'
srcdir = '/home/andreas/Documents/Projects/species-sensitivity/src'

# source
source('/home/andreas/Documents/UBA/Project/R/src/DB_access.R')
source(file.path(fundir, 'ppdb-0-0-5.R'))
source(file.path(fundir, 'ppdb_own_functions.R'))
source(file.path(fundir, 'casconv.R'))
source(file.path(fundir, 'functions.R')) # contains gm_mean(), ksource()
source(file.path(fundir, 'agg_group.R'))
source(file.path(fundir, 'plot_outlier.R'))
```

# Load substances
```{r subst1}
## load data
psm = readRDS(file.path(cachedir, 'psm1.rds')) # created in EC50sensitivity_data.Rmd
psm = psm[ !is.na(casnr) ] # only metabolite and sumed pesticides
psm = psm[ !subst_name %in% c('Desethylsimazin') ] # casnr: 1007289; duplicated with Desethylsimazin

## for Nadin and verena
# psm = fread('/home/andreas/Documents/data/romania/det_compounds_with_CAS.csv')
# psm[ , casnr := gsub('-', '', cas) ]
# psm[ , `:=`
#      (variable_id = NA,
#       subst_name = NA,
#       psm_type = NA)]

## CHECKs
cas_check = 
  psm[ is.na(casnr) | casnr == '' | 
       is.na(cas) | cas == '' ]

if (nrow(cas_check) != 0) {
  warning(nrow(cas_check), 'missing CAS or CASNR entries.')
}

rm(cas_check)
```

# PubChem
```{r cas-resolver}
## pubchem
cas = psm$cas

if (online) {

  cid = get_cid(cas)

  pc_l = list()
  for (i in 1:length(cid)) {
    cas1 = names(cid[i])
    cid1 = cid[[i]]
    message('Querying: CAS:', cas1, '; CID:', cid1)

    pc1 = pc_prop(cid1)

    pc_l[[i]] = pc1
    names(pc_l)[i] = cas1
  }

  saveRDS(pc_l, file.path(cachedir, 'pc_l.rds'))
  saveRDS(cid, file.path(cachedir, 'cid.rds'))

} else {
  pc_l = readRDS(file.path(cachedir, 'pc_l.rds'))
  cir = readRDS(file.path(cachedir, 'cid.rds'))
}

# convert NAs from negative results to data.tables to be able to bind 'em
pc_l[ is.na(pc_l) ] = lapply(pc_l[ is.na(pc_l) ], data.table)

pubchem = rbindlist(pc_l, fill = TRUE, idcol = 'cas')

# cleaning
rm(cir)

# how to handle duplicated CAS
# test = pubchem[ , .(N = .N,
#              InChI = paste0(InChIKey),
#              IUPACName), by = cas][ N > 1]
```

As some CAS are duplicated the Pubchem data base is queried to get InChikeys and IUPAC names to be able to identify unique substances.
```{r subst2}
psm[pubchem, on = 'cas', iupac_name := i.IUPACName ]
# replacing missing subst_name with IUPAC names
psm[ , subst_name := ifelse(is.na(subst_name), iupac_name, subst_name)]
saveRDS(psm, file.path(cachedir, 'psm2.rds'))
```

# Ecotox test data

## Source EPA ECOTOX data
```{r SOURCE-epa}
ksource(file.path(srcdir, 'source_epa.Rmd'))
```

## Source bfg_monitoring data
Not needed anymore.

## Source PPDB data
```{r SOURCE-ppdb}
ksource(file.path(srcdir, 'source_ppdb.Rmd'))
# data cleaning
ppdb_dt[ latin_BIname %in% c('A4', 'A5', 'Cyanobacteria', 'Cyprinidae', 'Green algae', 'P4'),
         latin_BIname := 'Unknown species' ]
```

## Source Malaj data
```{r SOURCE-mlaj}
ksource(file.path(srcdir, 'source_malaj.Rmd'))
```

## Bind test reults
```{r tests-bind}
# bind the results from different sources
tests = rbindlist(list(epa2, ppdb_dt, malaj_m), fill = TRUE) # dm, 
```

## Errata
```{r tests-errata}
tests[ latin_BIname == 'Notropis ludibundis', `:=`
       (latin_BIname = 'Notropis ludibundus',
        latin_short = 'N. ludibundus')]

## remove hybrid X marking
tests[ , latin_BIname := gsub('\\sX\\s', ' ', latin_BIname) ]
## remove sp.
tests[ , latin_BIname := gsub('\\ssp\\.', '', latin_BIname) ]
## change species names
tests[ latin_BIname == 'Paratelphusa masoniana', latin_BIname := 'Parathelphusa masoniana' ]
tests[ latin_BIname == 'Tresus nuttalli', latin_BIname := 'Tresus nuttallii' ]
tests[ latin_BIname == 'Mysidopsis bahia', latin_BIname := 'Americamysis bahia' ]
tests[ latin_BIname == 'Planaria gonocephala', latin_BIname := 'Dugesia gonocephala' ]

## change family names
tests[ latin_BIname == 'Tubifex', family_epa := 'Naididae' ]
## change species and family names
tests[ latin_BIname == 'Chlorella fusca', `:=`
       (latin_BIname = 'Scenedesmus fuscus',
         family_epa = 'Scenedesmaceae ') ] # according to worms

## save the results locally
saveRDS(tests, file.path(cachedir, 'tests1.rds'))
```

# Taxonomic data

```{r MERGE-taxonomic-data}
#### Taxonomic data ----
ksource(file.path(srcdir, 'speciesinfo.Rmd'))

# Merge
tests = merge(tests, tx_dt, on = 'latin_BIname', all = TRUE)

#### Check taxize family classification ----
# tests[ source == 'epa_ecotox' & family != family_epa,
#        c('latin_BIname', 'family', 'family_epa', 'source_tax') ][ , .N, .(latin_BIname, family, family_epa, source_tax)][order(-N)]
use_epa_fam = c('Lemna minor', 'Lemna gibba', 'Lemna aequinoctialis', 'Viviparus bengalensis', 'Rattus rattus', 'Phytophthora capsici', 'Ganaspidium utilis', 'Indonaia caerulea', 'Phytophthora citricola', 'Rutilus rutilus', 'Sigara substriata', 'Stenonema candidum', 'Eriophyes datura', 'Hyalella curvispina', 'Neopanope texana', 'Paecilomyces farinosus', 'Melanopsis dufouri', 'Thiara lineata', 'Phytophthora cactorum', 'Waitea circinata', 'Culex melanurus', 'Lithodes antarcticus', 'Meloidogyne incognita', 'Paratelphusa masoniana', 'Physalaemus biligonigerus', 'Stethorus japonicus', 'Aphanizomenon flos-aquae', 'Caridina rajadhari', 'Palaemonetes africanus', 'Plea atomaria', 'Anacystis alpicola', 'Anisops bouvieri', 'Clibanarius africanus', 'Katelysia opima', 'Palustra laboulbeni', 'Physalaemus centralis', 'Tresus nuttalli', 'Anticarsia gemmatalis', 'Elasmopus bampo', 'Physalaemus santafecinus', 'Planaria gonocephala', 'Podisus nigrispinus', 'Ranatra elongata', 'Zonotrichia albicollis', 'Oziotelphusa senex',
'Perca', 'Nitzschia', 'Lemna', 'Libellula', 'Opius', 'Hydra', 'Asellus', 'Cyclops', 'Cypria', 'Sphaerium', 'Crangon', 'Danio', 'Lucifer', 'Micronecta', 'Sagitta', 'Alona', 'Ameletus', 'Amphora', 'Colpoda', 'Lestes', 'Micropterus', 'Physella', 'Ulnaria', 'Choristoneura rosaceana', 'Cydia pomonella', 'Epiphyas postvittana', 'Helicoverpa armigera', 'Pholetesor ornigis', 'Phthorimaea operculella', 'Plodia interpunctella', 'Plutella xylostella', 'Sclerotinia sclerotiorum', 'Spodoptera littoralis')

#### final family classification ----
## by taxize wrongly classified columns
tests[ , family := ifelse(latin_BIname %in% use_epa_fam, family_epa, family_tax) ]

## family errata
tests[ latin_BIname == 'Pseudokirchneriella subcapitata',
       family := 'Selenastraceae']

#! INTERMEDIATE: They are not yet classified in speciesinfo.org
#! were added when time span was expanded to 360h
tests[ source == 'epa_ecotox' & is.na(family),
       family := 'family_epa' ]

# save the results locally
saveRDS(tests, file.path(cachedir, 'tests2.rds'))
```

# Habitat data
```{r MERGE-habitat-data}
#### Habitat data ----
# self compiled list
ksource(file.path(srcdir, 'familylookup.Rmd'))
# worms
ksource(file.path(srcdir, 'habitat_classification.Rmd'))
# freshwaterecology.info
#! todo
# algaebase
#! scraping is against their terms of use

## Merge
tests = merge(tests, lookup, by = 'family', all.x = TRUE)
tests = merge(tests, lookup_worms_sp, by = 'latin_BIname', all.x = TRUE)

#### Create isFre final column ----
# Freshwater
tests[ , 
       isFre_sum := rowSums(sapply(.SD, as.numeric), na.rm = TRUE),
       .SDcols = c('isFre', 'isFre_w') ]
tests[ , isFre_fin := ifelse(isFre_sum >= 1, '1', '0') ][ , isFre_sum := NULL ]
# Marine
tests[ , 
       isMar_sum := rowSums(sapply(.SD, as.numeric), na.rm = TRUE),
       .SDcols = c('isMar', 'isMar_w') ]
tests[ , isMar_fin := ifelse(isMar_sum >= 1, '1', '0') ][ , isMar_sum := NULL ]
# Terrestrial
tests[ , 
       isTer_sum := rowSums(sapply(.SD, as.numeric), na.rm = TRUE),
       .SDcols = c('isTer', 'isTer_w') ]
tests[ , isTer_fin := ifelse(isTer_sum >= 1, '1', '0') ][ , isTer_sum := NULL ]

#### Classify Invertebrates ----
#invertebrates_phylum = c("Annelida", "Bryozoa", "Chaetognatha", "Ciliophora", "Cnidaria", "Gastrotricha", "Echinodermata", "Mollusca", "Nematoda", "Nemertea", "Platyhelminthes", "Porifera", "Rotifera")
inv_makro_phylum = c('Annelida', 'Echinodermata', 'Mollusca', 'Nemertea', 'Platyhelminthes', 'Porifera')

inv_mikro_phylum = c('Bryozoa', 'Chaetognatha', 'Ciliophora', 'Cnidaria', 'Gastrotricha', 'Nematoda', 'Rotifera')

#invertebrates_subphylum = c('Crustacea') # phylum: Arthropoda, Chordata
inv_makro_subphylum = c('Crustacea')
inv_makro_class = c('Arachnida', 'Diplopoda', 'Entognatha', 'Insecta') # phylum: Arthropoda

invertebrates_makro = c(inv_makro_phylum, inv_makro_subphylum, inv_makro_class)
invertebrates_mikro = inv_mikro_phylum

tests[ , supgroup2 := ifelse(supgroup %in% invertebrates_makro, 'Makro_Inv',
                      ifelse(supgroup %in% invertebrates_mikro, 'Mikro_Inv', supgroup)) ]

#### Classify trophic level
autotrophs = c('Plants', 'Algae', 'Bryophyta')
tests[ , trophic_lvl := ifelse(supgroup2 %in% autotrophs,
                               'autotrophic', 'heterotrophic') ]
# cleaning
rm(list = grep('inv_', ls(), value = TRUE))
```

# Checks and save

Check whether al entries have CAS, CASNR and substance and species names.
```{r checks}
subst_check = 
  tests[ is.na(casnr) | casnr == '' |
         is.na(cas) | cas == '' |
         is.na(subst_name) | subst_name == '' ]
tax_check = 
  tests[ is.na(latin_BIname) | latin_BIname == '' |
         is.na(group) | group == '' |
         is.na(supgroup) | supgroup == '' |
         is.na(supgroup2) | supgroup == '' ]
value_check = 
  tests[ is.na(value) | value == '' ]
fresh_check =
  tests[ is.na(isFre_fin) | isFre_fin == '' ]
marin_check = 
  tests[ is.na(isMar_fin) | isMar_fin == '' ]
terre_check = 
  tests[ is.na(isTer_fin) | isTer_fin == '' ]

if (nrow(subst_check) != 0) {
  warning(nrow(subst_check), ' missing CAS, CASNR or substance names.')
}
if (nrow(tax_check) != 0)  {
  warning(nrow(tax_check), ' missing latin_BIname, group, supgroup or supgroup2 entry.')
}
if (nrow(value_check) != 0) {
  warning(nrow(value_check), ' missing value entries.')
}
if (nrow(fresh_check) != 0) {
  warning(nrow(value_check), ' missing freshwater classifications.')
}
if (nrow(marin_check) != 0) {
  warning(nrow(marin_check), ' missing marine classifications.')
}
if (nrow(terre_check) != 0) {
  warning(nrow(terre_check), ' missing terrestrial classifications.')
}

# cleaning
rm(list = grep('check', ls(), value = TRUE))

# write to cache
saveRDS(tests, file.path(cachedir, 'tests3.rds'))
```

# Overview table

```{r overview-table}
taxa_fresh = tests[ isFre_fin == '1' , .N, .(supgroup2, supgroup, group) ][order(supgroup2, supgroup, group)]
taxa_marin = tests[ isMar_fin == '1' , .N, .(supgroup2, supgroup, group) ][order(supgroup2, supgroup, group)]
taxa_terre = tests[ isTer_fin == '1' , .N, .(supgroup2, supgroup, group) ][order(supgroup2, supgroup, group)]

taxa_fresh
taxa_marin
taxa_terre

# cleaning
rm(list = grep('taxa', ls(), value = TRUE))
```

# EC50 aggregation

```{r read-tests-data}
tests3 = readRDS(file.path(cachedir, 'tests3.rds'))
tests3[ str_count(tests3$latin_BIname, '\\S+') < 2, 
        latin_BIname := paste0(latin_BIname, ' sp.') ]
```

## Malaj aggregate
```{r AGGREGATE-malaj}
# aggregating by casnr
malaj_agg =
  tests3[ source == 'malaj',
          j = .(mj50f_al4896_vl = min(.SD[latin_BIname == 'Pseudokirchneriella subcapitata']$value),
                mj50f_da48_vl = min(.SD[latin_BIname == 'Daphnia magna']$value),
                mj50f_fi96_vl = min(.SD[latin_BIname == 'Pimephales promelas']$value)),
          by = .(casnr) ]

# Inf to NA
for (i in names(malaj_agg)) {
  malaj_agg[get(i) == Inf, (i) := NA]
}

# add info column
malaj_agg[ , mj50f_al4896_tax := ifelse(!is.na(mj50f_al4896_vl), 'Pseudokirchneriella subcapitata', NA) ]
malaj_agg[ , mj50f_da48_tax := ifelse(!is.na(mj50f_da48_vl), 'Daphnia magna', NA) ]
malaj_agg[ , mj50f_fi96_tax := ifelse(!is.na(mj50f_fi96_vl), 'Pimephales promelas', NA) ]

# add test count column (careful, some are estimated!)
malaj_agg[ , mj50f_al4896_n := ifelse(!is.na(mj50f_al4896_vl), 1, NA) ]
malaj_agg[ , mj50f_da48_n := ifelse(!is.na(mj50f_da48_vl), 1, NA) ]
malaj_agg[ , mj50f_fi96_n := ifelse(!is.na(mj50f_fi96_vl), 1, NA) ]

# colorder
setcolorder(malaj_agg, c('casnr',
                         'mj50f_al4896_vl', 'mj50f_al4896_tax', 'mj50f_al4896_n',
                         'mj50f_da48_vl', 'mj50f_da48_tax', 'mj50f_da48_n',
                         'mj50f_fi96_vl', 'mj50f_fi96_tax', 'mj50f_fi96_n'))
```

## PPDB aggregate
```{r AGGREGATE-ppdb}
ppdb_agg = 
  tests3[ source == 'ppdb',
          j = .(# algae
                pp50f_al72_vl = min(.SD[ppdb_group == 'algae']$value),
                pp50f_al72_tax = .SD[ppdb_group == 'algae' & which.min(value)]$latin_BIname,
                # aquatic plants
                pp50f_pl168_vl = min(.SD[ppdb_group == 'aquatic plants']$value),
                pp50f_pl168_tax = .SD[ppdb_group == 'aquatic plants' & which.min(value)]$latin_BIname,
                # Daphnia
                pp50f_da48_vl = min(.SD[latin_BIname %like% '(?i)daphnia']$value),
                pp50f_da48_tax = .SD[latin_BIname %like% '(?i)daphnia' & which.min(value)]$latin_BIname,
                # crustaceans
                pp50f_cr96_vl = min(.SD[ppdb_group == 'aquatic crustaceans']$value),
                pp50f_cr96_tax = .SD[ppdb_group == 'aquatic crustaceans' & which.min(value)]$latin_BIname,
                # sediment dwelling organisms (mostly Chironomidae)
                pp50f_ch96_vl = min(.SD[ppdb_group == 'sediment dwelling organisms']$value),
                pp50f_ch96_tax = .SD[ppdb_group == 'sediment dwelling organisms' & which.min(value)]$latin_BIname,
                # fish
                pp50f_fi96_vl = min(.SD[ppdb_group == 'fish']$value),
                pp50f_fi96_tax = .SD[ppdb_group == 'fish' & which.min(value)]$latin_BIname),
          by = .(casnr) ]

# cleaning
for (i in names(ppdb_agg)) {
  ppdb_agg[get(i) == Inf, (i) := NA]
}

# add count column
ppdb_agg[ , pp50f_al72_n := ifelse(!is.na(pp50f_al72_vl), 1, NA) ]
ppdb_agg[ , pp50f_pl168_n := ifelse(!is.na(pp50f_pl168_vl), 1, NA) ]
ppdb_agg[ , pp50f_da48_n := ifelse(!is.na(pp50f_da48_vl), 1, NA) ]
ppdb_agg[ , pp50f_cr96_n := ifelse(!is.na(pp50f_cr96_vl), 1, NA) ]
ppdb_agg[ , pp50f_ch96_n := ifelse(!is.na(pp50f_ch96_vl), 1, NA) ]
ppdb_agg[ , pp50f_fi96_n := ifelse(!is.na(pp50f_fi96_vl), 1, NA) ]

#### Combine Invertebrates ----
ppdb_agg_m = 
  melt(ppdb_agg, id.vars = 'casnr',
  measure.vars = list(c('pp50f_da48_vl', 'pp50f_cr96_vl', 'pp50f_ch96_vl'),
                      c('pp50f_da48_tax', 'pp50f_cr96_tax', 'pp50f_ch96_tax'),
                      c('pp50f_da48_n', 'pp50f_cr96_n', 'pp50f_ch96_n')))
ppdb_agg_m_agg =
  ppdb_agg_m[ !is.na(value1) & !is.infinite(value1),
              .(pp50f_iv4896_min = min(value1),
                pp50f_iv4896_tax = .SD[which.min(value1)]$value2,
                pp50f_iv4896_info = paste0(.SD[order(value1)]$value2, ' (',
                                           .SD[order(value1)]$value1, 'ug/L)',
                                           collapse = ' - '),
                pp50f_iv4896_n = sum(.SD[!is.na(value1)]$value3)),
              casnr ]

ppdb_agg = ppdb_agg[ppdb_agg_m_agg, on = 'casnr']
# removing Inf and (µg/L)
for (i in names(ppdb_agg)) {
  ppdb_agg[get(i) == ' (ug/L)', (i) := NA]
}

# col order
names(ppdb_agg)
setcolorder(ppdb_agg, c('casnr',
                        'pp50f_al72_vl', 'pp50f_al72_tax', 'pp50f_al72_n',
                        'pp50f_pl168_vl', 'pp50f_pl168_tax', 'pp50f_pl168_n',
                        'pp50f_da48_vl', 'pp50f_da48_tax', 'pp50f_da48_n',
                        'pp50f_cr96_vl', 'pp50f_cr96_tax', 'pp50f_cr96_n',
                        'pp50f_ch96_vl', 'pp50f_ch96_tax', 'pp50f_ch96_n',
                        'pp50f_fi96_vl', 'pp50f_fi96_tax', 'pp50f_fi96_n',
                        'pp50f_iv4896_min', 'pp50f_iv4896_tax', 'pp50f_iv4896_info', 'pp50f_iv4896_n'))
```

### Malaj-PPDB check
```{r malaj-ppdb-check}
# check
malj_check = length(which(duplicated(malaj_agg$casnr)))
ppdb_check = length(which(duplicated(ppdb_agg$casnr)))

if (malj_check != 0) {
  stop(malj_check, ' duplicated CAS entries.')
}
if (ppdb_check != 0) {
  stop(ppdb_check, ' duplicated CAS entries.')
}

# cleaning
rm(malj_check, ppdb_check)
```

## EPA aggregate
```{r AGGREGATE-epa1, eval=FALSE}
# ksource(file.path(srcdir, 'EC50aggregation_epa1.Rmd'))
```

## EPA aggregate 2
```{r AGGREGATE-epa2}
ksource(file.path(srcdir, 'EC50aggregation_epa2.Rmd'))
```

# Merge
```{r MERGE-ec50}
source_l = list(malaj_agg, ppdb_agg, ep50f_agg)
ec50 = Reduce(function(...) merge(..., on = 'casnr', all = TRUE), source_l)
ec50[tests3, on = 'casnr', subst_name := i.subst_name]
setcolorder(ec50, c(1, ncol(ec50), 3:ncol(ec50)-1))
```

## Check
```{r ec50-check}
na_check = ec50[ is.na(casnr) | is.na(subst_name) ]
if (nrow(na_check) != 0) {
  warning(nrow(na_check), ' missing CAS, CASNR or substance names.')
}
rm(na_check)
```

# Write to database
```{r disclaimer}
su = length(unique(tests3$casnr)) # substances
sp = length(as.character(unique(tests3$latin_BIname))) # species
fa = length(unique(tests3$family)) # families
ha = nrow(lookup)
```

```{r WRITE-postgres, eval=FALSE}
drv = dbDriver("PostgreSQL")
con = dbConnect(drv, user = DBuser, dbname = DBname_1, host = DBhost, port = DBport, password = DBpassword)

schema = 'phch'
table = 'phch_ec50'

## upload
dbSendQuery(con, paste0("DROP TABLE IF EXISTS  ", schema, ".", table, ";"))
dbWriteTable(con, ec50,
             name = c(schema, table), row.names = FALSE)

## User rights
dbSendQuery(con, paste0("GRANT ALL ON TABLE ", schema, ".", table, " TO scharmueller;
                  GRANT ALL ON TABLE ", schema, ".", table, " TO bfg;
                  GRANT SELECT ON TABLE ", schema, ".", table, " TO bfg_read;"))

## Primary Key
# dbSendQuery(con, paste0("ALTER TABLE ", schema, ".", table, " ADD PRIMARY KEY (variable_id);"))

## COMMENT ON
dbSendQuery(con, paste0("COMMENT ON TABLE  ", schema, ".", table, "  IS '
pesticide ec50 table fo specific organism groups per CASNR:

#### Malaj et al. (2014) data ----
(1) mj50_<ps|dm|pp> = EC50 values from Malaj for dm-Daphnia magna, pp-Pimephales promelas, ps-Pseudokirchneriella subcapitata (Already aggregated).
(2) mj50_<dm_type|pp_type|ps_type> = value type used by Malaj et al.:
(B-Baseline toxicity, E-Empirical, P1-4-Predictet(3,4-trustworthy;1,2-not))

#### PPDB data ----
(1) pp50f_<al|pl|da|cr|ch|fi>
EC50 vlaues from PPDB for iv-Invertebrates, al-Algae and fi-Fish (the minimum per CASNR and tax. group was taken).
(2) pp50f_<al|pl|da|cr|ch|fi>_info
string containing all EC50 values per CASNR and tax. group.

#### EPA data ----
(1) n<f|m|t|n>_<al|pl|iv|da|cr|ch|in|fi|am><48|72|96|168>
number of tests for specific substance and organism group (here: algae).
(2) ep50<f|m|t|n>_<al|pl|iv|da|cr|ch|in|fi|am><48|72|96|168>_<min|gm|mn>
e.g:
ep50_al72_min = ec50_algae72h_minimum_value
ep50_da48_gm = ec50_Daphniidae48h_geometric mean (with deleted ouliers)
ep50f_al4896_mn = ec50freshwater_algae48h-96h_mean
! For those w/o time specification tests with a duration from 24 to 120h (for plants: 360h) were used.
(3) ep50<f|m|t|n>_<al|pl|iv|da|cr|ch|in|fi|am><48|72|96|168>_info
info column which taxa wered used (orderd by value), the EPA ecotox ref. number and the value (separated by comma)
e.g Daphnia magna-(984)-50ugL,Daphnia magna-(984)-50ugL
(4) ep50<f|m|t|n>_<al|pl|iv|da|cr|ch|in|fi|am><48|72|96|168>_out
colum which denotes test results that were excluded based on outlier detection

Aggregated EPA taxa:
", freshw_grp[ ,1], "
", freshw_grp[ ,2], "
", freshw_grp[ ,3], "
", freshw_grp[ ,4], "

---
Abbreviations:
value types:
vl = value (there was only one value, hence no aggregation)
min = minimum
gm = geometric mean
mn = mean
md = median

habitat types:
f = freshwater
m = marine
t = terrestrial
n = non

taxa groups:
pl = Plants
al = Algae
fu = Fungi
ma = Makro-Invertebrates (", paste0(invertebrates_makro, collapse = ' + '),")
mi = Mikro-Invertebrates (", paste0(invertebrates_mikro, collapse = ' + '),")
da = Daphniidae (Fam.)
cr = Crustacea
ch = Chrionomidae (Fam.)
in = Insects
mo = Molluscs
fi = Fish
ar = Araneae
au = autotroph organisms (", paste0(autotrophs, collapse = ' + '),")
he = heterotroph organisms

You can look up the reference number, casnr, etc. here:
https://cfpub.epa.gov/ecotox/quick_query.htm

---
All values are in ug/L.
---

Note:
In order to build this table, ecological test data on ", su, " substances and ", sp, " species (from ", fa, " families) were processed.\nThe test data were mainly obtained through the EPA ECOTOX data base and some scientific articles and the PPDB.\nIn order to properly classify taxonomic ranks and habitat preferences, the ITIS, NCBI, COL, EOL, NBN and WORMS data bases were queried.\nFor ", ha, " families, additional habitat information (freshwater, marine, terretrial) was collected by hand. No guarantee for perfectness.\nEspecially habitat information can still have wrong classifications.

---
Uploaded: ", Sys.Date(), "';"))

## Maintainance
dbSendQuery(con, paste0("VACUUM ANALYZE  ", schema, ".", table, ";"))

dbDisconnect(con)
dbUnloadDriver(drv)

message('Written to DB:'); freshw_grp
```

```{r cleaning}
rm(su, sp, fa, ha)
```

# Problems
```{r problems. eval=FALSE}
ppdb_epoxi = '135319732' # Kein EPA-Eintrag
ppdb_epoxi_cas = '135319-73-2' # Kein EPA-Eintrag
ppdb_epoxi_cas_old = '106325-08-0'
epa_epoxi = '133855988'

tests[ casnr %in% c(ppdb_epoxi, epa_epoxi) ]
ppdb_l[[ppdb_epoxi_cas]]$etox
```

# Misc
```{r misc, eval=FALSE}
daphnia = tests[ latin_BIname %like% '(?i)Daphnia', .N, latin_BIname ][order(-N)]
fwrite(daphnia, file.path(tempdir(), 'daphnia.csv'))
```

### 2018-07-09 src/EC50aggregation_fin.Rmd
publishing orphan commit

---
title: "Untitled"
author: "AS"
date: "March 29, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

This script creates the table phch_ec50_fin on the basis of phch_ec50 and loads it into the data base.

# Setup
```{r setup}
# packages
require(RPostgreSQL)
require(data.table)
require(stringr)

# variables
cachedir = '/home/andreas/Documents/Projects/species-sensitivity/cache'
plotdir = '/home/andreas/Documents/Projects/species-sensitivity'

# source
source('/home/andreas/Documents/UBA/Project/R/src/DB_access.R')
```

# Load ec50-table
```{r load}
drv = dbDriver("PostgreSQL")
con = dbConnect(drv, user = DBuser, dbname = DBname_1, host = DBhost, port = DBport, password = DBpassword)

ec50 = dbGetQuery(con, "SELECT * FROM phch.phch_ec50")
setDT(ec50)

dbDisconnect(con)
dbUnloadDriver(drv)
```

# Final EC50 columns

## Hierachical partition

### Daphniidae
```{r daphniidae}
ec50_da48_fin = 
  ec50[ ,
        .(casnr = casnr,
          ec50f_da48_fin = ifelse(!is.na(mj50f_da48_vl), mj50f_da48_vl,
                           ifelse(!is.na(pp50f_da48_vl), pp50f_da48_vl,
                           ifelse(!is.na(ep50f_da48_md), ep50f_da48_md, NA))),
          ec50f_da48_tax = ifelse(!is.na(mj50f_da48_vl), mj50f_da48_tax,
                           ifelse(!is.na(pp50f_da48_vl), pp50f_da48_tax,
                           ifelse(!is.na(ep50f_da48_md), ep50f_da48_tax, NA))),
          ec50f_da48_n = ifelse(!is.na(mj50f_da48_vl), mj50f_da48_n,
                         ifelse(!is.na(pp50f_da48_vl), pp50f_da48_n,
                         ifelse(!is.na(ep50f_da48_md), ep50f_da48_n, NA))),
          ec50f_da48_src = ifelse(!is.na(mj50f_da48_vl), 'malaj',
                           ifelse(!is.na(pp50f_da48_vl), 'ppdb',
                           ifelse(!is.na(ep50f_da48_md), 'epa', NA))))]

# keep only the first part of the string (until '-'): This is the minimum value
ec50_da48_fin[ , ec50f_da48_tax := sapply(strsplit(ec50f_da48_tax, '_'), '[', 1)]
```

## Minimum partition

### Algae
```{r algae}
ec50_al_m = 
  melt(ec50, id.vars = 'casnr',
       measure.vars = list(c('mj50f_al4896_vl', 'pp50f_al72_vl', 'ep50f_al4896_min'),
                           c('mj50f_al4896_tax', 'pp50f_al72_tax', 'ep50f_al4896_tax'),
                           c('mj50f_al4896_n', 'pp50f_al72_n', 'ep50f_al4896_n')))

ec50_al_m[ , value4 := ifelse(variable == '1', 'malaj',
                       ifelse(variable == '2', 'ppdb',
                       ifelse(variable == '3', 'epa', NA))) ] #! careful with the order

ec50_al4896_fin = 
  ec50_al_m[ ,
             .(min(value1, na.rm = TRUE),
               .SD[which.min(value1)]$value2,
               .SD[which.min(value1)]$value3,
               .SD[which.min(value1)]$value4),
             casnr  ]

setnames(ec50_al4896_fin, c('casnr', 'ec50f_al4896_fin', 'ec50f_al4896_tax', 'ec50f_al4896_n', 'ec50f_al4896_src'))

# keep only the first part of the string (until '-'): This is the minimum value
ec50_al4896_fin[ , ec50f_al4896_tax := sapply(strsplit(ec50f_al4896_tax, ' - '), '[', 1)]
#ec50_al4896_fin[ , ec50f_al4896_dur := gsub('[^0-9]', '', ec50_al4896_fin$ec50f_al4896_tax) ]
ec50_al4896_fin[ , ec50f_al4896_tax := sapply(strsplit(ec50f_al4896_tax, '_'), '[', 1)]
ec50_al4896_fin[ , ec50f_al4896_tax := sapply(strsplit(ec50f_al4896_tax, '_'), '[', 1)]
ec50_al4896_fin[ , ec50f_al4896_tax := ifelse(str_count(ec50f_al4896_tax, '\\S+') == 1,
                                              paste0(ec50f_al4896_tax, ' sp.'),
                                              ec50f_al4896_tax)]

for (i in names(ec50_al4896_fin)) {
  ec50_al4896_fin[ get(i) == Inf, (i) := NA ]
  ec50_al4896_fin[ get(i) == '', (i) := NA ]
}
rm(ec50_al_m)
```

### Invertebrates
```{r makro-invertebrates}
ec50_ma_m = 
  melt(ec50, id.vars = 'casnr',
       measure.vars = list(c('mj50f_da48_vl', 'pp50f_iv4896_min', 'ep50f_ma4896_min'),
                           c('mj50f_da48_tax', 'pp50f_iv4896_tax', 'ep50f_ma4896_tax'),
                           c('mj50f_da48_n', 'pp50f_iv4896_n', 'ep50f_ma4896_n')))

ec50_ma_m[ , value4 := ifelse(variable == '1', 'malaj',
                       ifelse(variable == '2', 'ppdb',
                       ifelse(variable == '3', 'epa', NA))) ] #! careful with the order

ec50_ma4896_fin = 
  ec50_ma_m[ ,
             .(min(value1, na.rm = TRUE),
               .SD[which.min(value1)]$value2,
               .SD[which.min(value1)]$value3,
               .SD[which.min(value1)]$value4),
             casnr  ]

setnames(ec50_ma4896_fin, c('casnr', 'ec50f_ma4896_fin', 'ec50f_ma4896_tax', 'ec50f_ma4896_n', 'ec50f_ma4896_src'))

# keep only the first part of the string (until '-'): This is the minimum value
ec50_ma4896_fin[ , ec50f_ma4896_tax := sapply(strsplit(ec50f_ma4896_tax, ' - '), '[', 1)]
#ec50_ma4896_fin[ , ec50f_ma4896_dur := gsub('[^0-9]', '', ec50_ma4896_fin$ec50f_ma4896_tax) ]
ec50_ma4896_fin[ , ec50f_ma4896_tax := sapply(strsplit(ec50f_ma4896_tax, '_'), '[', 1)]
ec50_ma4896_fin[ , ec50f_ma4896_tax := ifelse(str_count(ec50f_ma4896_tax, '\\S+') == 1,
                                              paste0(ec50f_ma4896_tax, ' sp.'),
                                              ec50f_ma4896_tax)]

for (i in names(ec50_ma4896_fin)) {
  ec50_ma4896_fin[ get(i) == Inf, (i) := NA ]
  ec50_ma4896_fin[ get(i) == '', (i) := NA ]
}
rm(ec50_ma_m)
```

### Fish
```{r fish}
ec50_fi_m = 
  melt(ec50, id.vars = 'casnr',
       measure.vars = list(c('mj50f_fi96_vl', 'pp50f_fi96_vl', 'ep50f_fi96_min'),
                           c('mj50f_fi96_tax', 'pp50f_fi96_tax', 'ep50f_fi96_tax'),
                           c('mj50f_fi96_n', 'pp50f_fi96_n', 'ep50f_fi96_n')))

ec50_fi_m[ , value4 := ifelse(variable == '1', 'malaj',
                       ifelse(variable == '2', 'ppdb',
                       ifelse(variable == '3', 'epa', NA))) ] #! careful with the order

ec50_fi96_fin = 
  ec50_fi_m[ ,
             .(min(value1, na.rm = TRUE),
               .SD[which.min(value1)]$value2,
               .SD[which.min(value1)]$value3,
               .SD[which.min(value1)]$value4),
             casnr  ]

setnames(ec50_fi96_fin, c('casnr', 'ec50f_fi96_fin', 'ec50f_fi96_tax', 'ec50f_fi96_n', 'ec50f_fi96_src'))

# keep only the first part of the string (until '-'): This is the minimum value
ec50_fi96_fin[ , ec50f_fi96_tax := sapply(strsplit(ec50f_fi96_tax, ' - '), '[', 1)]
#ec50_fi96_fin[ , ec50f_fi96_dur := gsub('[^0-9]', '', ec50_fi96_fin$ec50f_fi96_tax) ]
ec50_fi96_fin[ , ec50f_fi96_tax := sapply(strsplit(ec50f_fi96_tax, '_'), '[', 1)]
ec50_fi96_fin[ , ec50f_fi96_tax := ifelse(str_count(ec50f_fi96_tax, '\\S+') == 1,
                                              paste0(ec50f_fi96_tax, ' sp.'),
                                              ec50f_fi96_tax)]

for (i in names(ec50_fi96_fin)) {
  ec50_fi96_fin[ get(i) == Inf, (i) := NA ]
  ec50_fi96_fin[ get(i) == '', (i) := NA ]
}
rm(ec50_fi_m)
```

## Merge 1
```{r merge1}
fin_dts = grep('_fin', ls(), value = TRUE, ignore.case = TRUE)
fin_l = lapply(fin_dts, get)
names(fin_l) = fin_dts

ec50f_grp = Reduce(function(...) merge(..., on = 'casnr', all = TRUE), fin_l)
ec50f_grp[ec50, on = 'casnr', subst_name := i.subst_name]
setcolorder(ec50f_grp, c(1, ncol(ec50f_grp), 3:ncol(ec50f_grp)-1))
```

# Write to data base

```{r write-ec50f_fin}
drv = dbDriver("PostgreSQL")
con = dbConnect(drv, user = DBuser, dbname = DBname_1, host = DBhost, port = DBport, password = DBpassword)

schema = 'phch'
table = 'phch_ec50_fin'

## upload
dbSendQuery(con, paste0("DROP TABLE IF EXISTS  ", schema, ".", table, ";"))
dbWriteTable(con, ec50f_grp,
             name = c(schema, table), row.names = FALSE)

## User rights
dbSendQuery(con, paste0("GRANT ALL ON TABLE ", schema, ".", table, " TO scharmueller;
                  GRANT ALL ON TABLE ", schema, ".", table, " TO bfg;
                  GRANT SELECT ON TABLE ", schema, ".", table, " TO bfg_read;"))

## Primary Key
# dbSendQuery(con, paste0("ALTER TABLE ", schema, ".", table, " ADD PRIMARY KEY (variable_id);"))

## COMMENT ON
dbSendQuery(con, paste0("COMMENT ON TABLE  ", schema, ".", table, "  IS '
final ec50 table with aggregated Invertebrate values based on phch.phch_ec50

(1) ec50f_al4896_fin = minimum from mj50f_al4896_vl, pp50f_al72_vl, ep50f_al4896_min
(2) ec50f_da48_fin = if present: mj50f_al4896_vl, pp50f_al72_vl, ep50f_al4896_md
(3) ec50f_fi96_fin = minimum from mj50f_da48_vl, pp50f_ma48_min, ep50f_ma4896_min
(4) ec50f_ma4896_fin = minimum from mj50f_da48_vl, pp50f_ma48_min, ep50f_ma4896_min

---
Abbreviations:
suffixes:
_fin = final aggregated value
_tax = Taxon corresponding to _fin value
_n = number of vlaues/tests that were used in the aggregation
_src = source (malaj, ppdb or epa)

taxa groups:
pl = Plants
al = Algae
fu = Fungi
ma = Makro-Invertebrates (Annelida + Echinodermata + Mollusca + Nemertea + Platyhelminthes + Porifera + Crustacea + Arachnida + Diplopoda + Entognatha + Insecta)
mi = Mikro-Invertebrates (Bryozoa + Chaetognatha + Ciliophora + Cnidaria + Gastrotricha + Nematoda + Rotifera)
da = Daphniidae (Fam.)
cr = Crustacea
ch = Chrionomidae (Fam.)
in = Insects
mo = Molluscs
fi = Fish
ar = Araneae
au = autotroph organisms (Plants + Algae + Bryophyta)
he = heterotroph organisms

---
All values are in ug/L.

Uploaded: ", Sys.Date(), "';"))

## Maintainance
dbSendQuery(con, paste0("VACUUM ANALYZE  ", schema, ".", table, ";"))

dbDisconnect(con)
dbUnloadDriver(drv)
```

### 2018-07-09 src/EC50comparison_plots.Rmd
publishing orphan commit

---
title: "Untitled"
author: "AS"
date: "April 2, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup}
# packages
require(data.table)
require(RPostgreSQL)
require(ggplot2)

# variables
cachedir = '/home/andreas/Documents/Projects/species-sensitivity/cache'
plotdir = '/home/andreas/Documents/Projects/species-sensitivity/plots'

# source
source('/home/andreas/Documents/UBA/Project/R/src/DB_access.R')
```

# Load test results
```{r load-tests}
tests3 = readRDS(file.path(cachedir, 'tests3.rds'))
psm = readRDS(file.path(cachedir, 'psm2.rds'))
```

```{r load-ec50s}
drv = dbDriver("PostgreSQL")
con = dbConnect(drv, user = DBuser, dbname = DBname_1, host = DBhost, port = DBport, password = DBpassword)

ec50 = dbGetQuery(con, "SELECT * FROM phch.phch_ec50")
setDT(ec50)

dbDisconnect(con)
dbUnloadDriver(drv)

ec50[psm, on = 'casnr', psm_type := i.psm_type]
ec50 = ec50[psm_type %in% c('fungicide', 'herbicide', 'insecticide')]
```

# Test durations per organism groups
```{r duration-taxon}
tests_epa = tests3[ source == 'epa_ecotox' & isFre_fin == '1' ]

# merging taxa that are lower than supgroup2:
dt_ch = tests_epa[ family == 'Chironomidae' , .N, duration2][, grp := 'Chironomidae']
setnames(dt_ch, c('dur', 'N', 'grp'))
dt_da = tests_epa[ family == 'Daphniidae' , .N, duration2][, grp := 'Daphniidae']
setnames(dt_da, c('dur', 'N', 'grp'))
dt_cr = tests_epa[ supgroup == 'Crustacea' , .N, duration2][, grp := 'Crustacea']
setnames(dt_cr, c('dur', 'N', 'grp'))
dt_mo = tests_epa[ supgroup == 'Mollusca' , .N, duration2][, grp := 'Mollusca']
setnames(dt_mo, c('dur', 'N', 'grp'))
dt_in = tests_epa[ supgroup == 'Insecta' , .N, duration2][, grp := 'Insecta']
setnames(dt_in, c('dur', 'N', 'grp'))

dt_dur = tests_epa[ , .N, .(supgroup2, duration2)]
setnames(dt_dur, c('grp', 'dur', 'N'))

dt_dur = rbindlist(list(dt_dur, dt_ch, dt_da, dt_cr, dt_mo, dt_in), use.names = TRUE)

# plot
gg_dur = ggplot(dt_dur[ !grp %in% c('Aves', 'Mammalia', NA)], aes(y = N, x = dur)) +
  geom_label(aes(label = dur)) +
  facet_wrap( ~ grp, scales = 'free') +
  labs(title = 'Duration of tests per taxon',
       subtitle = 'freshwater organisms') +
  theme_bw()

ggsave(gg_dur, filename = file.path(plotdir, 'duration-taxon.png'),
       width = 10, height = 7)

# Why are there NAs???!!
# 'cause it's not yet classified in the lookup table
tests3[ source == 'epa_ecotox' & is.na(supgroup2)][ , .N, latin_BIname]
```

# Effect endpoint per taxon
```{r effect-taxon}
dt_eff = tests_epa[ , .N, .(supgroup2, effect)]
setnames(dt_eff, c('grp', 'effect', 'N'))

gg_eff = ggplot(dt_eff[ !grp %in% c('Mammalia', 'Aves', NA) ], aes(y = N, x = effect)) +
  geom_point() +
  facet_wrap( ~ grp, scales = 'free') +
  labs(title = 'Effect numbers per taxon',
       subtitle = 'freshwater organisms') +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1.1))

ggsave(gg_eff, filename = file.path(plotdir, 'effect-taxon.png'),
       width = 10, height = 7)
```

# Compare 24-120 tests to specific times
```{r comparison-plots2}
## Algae
# data
al_cols = grep('al.+(vl|gm)', names(ec50), value = TRUE)

ec50al = ec50[ , .SD, .SDcols = c('casnr', 'psm_type', al_cols)]
ec50al = ec50al[ psm_type %in% c('fungicide', 'herbicide', 'insecticide')]
ec50al_m = melt(ec50al, id.var = c('casnr', 'psm_type')) # melt
ec50al_m72 = melt(ec50al[ ,!c('psm_type')], id.vars = c('casnr', 'pp50f_al72_vl')) # range per duratiion group

# plot 1 [doesn't help much]
ggplot(ec50al_m, aes(y = reorder(casnr, value), x = value, col = variable)) +
  geom_point() +
  scale_x_log10() +
  facet_wrap( ~ variable, scales = 'free_x', ncol = 3)

# plpt 2 [doesn't help much]
ec50al_m_rng = ec50al_m[ , .(rng = range(value, na.rm = TRUE)[2] -
                               range(value, na.rm = TRUE)[1]),
                         casnr]
ec50al_m_rng[ is.infinite(rng), rng := NA ]

ggplot(ec50al_m_rng, aes(y = reorder(casnr, rng), x = rng)) +
  geom_point() +
  coord_cartesian(xlim = c(0,1000))
  #scale_x_log10()

# plot 3 [doesn't help much]
# sum per psm_type and variable
ec50al_m_sum = ec50al_m[ , .(mn = mean(value, na.rm = TRUE),
                             sd = sd(value, na.rm = TRUE)), .(variable, psm_type) ]

ggplot(ec50al_m_sum, aes(y = mn, variable)) +
  geom_point() +
  geom_errorbar(aes(ymin = mn-sd, ymax = mn+sd)) +
  facet_wrap( ~ psm_type) +
  coord_flip() +
  theme_bw()

# plot 4 [doesn't help much]
cols = grep('casnr|al.+(vl|gm)', names(ec50), value = TRUE)#, 'pp50f_al72_vl')
ec50al = ec50[ , .SD, .SDcols = cols]
cols = setdiff(cols, c('pp50f_al72_vl', 'casnr'))
ec50al_diff = copy(ec50al)[ , (cols) := abs(pp50f_al72_vl - .SD), .SDcols = cols]

ec50al_diff_m = melt(ec50al_diff, id.vars = 'casnr')

ggplot(ec50al_diff_m, aes(y = casnr, x = value)) +
  geom_point() +
  facet_wrap( ~ variable)

```

### 2018-07-09 src/habitat_classification.Rmd
publishing orphan commit

---
title: "Untitled"
author: "AS"
date: "March 5, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Setup
```{r setup}
## packages
require(data.table)
# devtools::install_github("ropensci/taxizesoap") # install
require(taxizesoap)

## switches
online = F

## variables
cachedir = '/home/andreas/Documents/Projects/species-sensitivity/cache'
```

# Load data
```{r data}
tests2 = readRDS(file.path(cachedir, 'tests2.rds'))
```

## WORMS

In order to test whether the retrieved EPA ECOTOX test results comprise only limnic species multiple databases are used for cross-checking. The [WORMS](http://www.marinespecies.org/index.php) database lists taxonomic information mainly on marine taxa. However as taxa not always make only use of marine habitats (e.g. eels which live both in marine and limnic habitats) the database also comprises taxa from limnic and estuarine environments. This is coded in the varaiable __Environment__ in the WORMS database. To query the database the [taxizesoap](https://github.com/ropensci/taxizesoap) package and its function `worms_records()` is used.

### Family

```{r worms-family, echo=TRUE}
# installation of taxizesoap: https://github.com/ropensci/taxizesoap
# install.packages(c("XMLSchema", "SSOAP"), repos = c("http://packages.ropensci.org", "http://cran.rstudio.com"))
# devtools::install_github("ropensci/taxizesoap")
if (online) {
  message('online = TRUE')
  family_todo = sort(unique(tests2$family))
  # family_todo = family_todo[1:2]

  worms_family_l = list()
  for (i in 1:length(family_todo)) {
    family = family_todo[i]

    message('Querying (', i, '/', length(family_todo), '): ', family)
    time = Sys.time()
    worms_data = taxizesoap::worms_records(scientific = family, marine_only = FALSE)
    Sys.time() - time

    if (nrow(worms_data) > 0) {
      worms_data = cbind(worms_data, latin_BIname = family)
    } else {
      worms_data = worms_data
    }

    worms_family_l[[i]] = worms_data
  }
  saveRDS(worms_family_l, file.path(cachedir, 'worms_family_list.rds'))

} else {
  message('online = FALSE')
  worms_family_l = readRDS(file.path(cachedir, 'worms_family_list.rds'))
}

worms_fam = rbindlist(worms_family_l)
worms_fam = worms_fam[ , c('family', 'isFreshwater', 'isMarine', 'isTerrestrial')]
# Take minimum 'cause some fmailies have mulitple entries (e.g Bopyridae)
lookup_worms_fam = worms_fam[ , lapply(.SD, function(x) min(as.numeric(x))),
                              by = family, .SDcols = c('isFreshwater', 'isMarine', 'isTerrestrial')]
setnames(lookup_worms_fam, c('family', 'isFre_w', 'isMar_w', 'isTer_w'))
```

### Species

```{r worms-species, echo=TRUE}
# installation of taxizesoap: https://github.com/ropensci/taxizesoap
# install.packages(c("XMLSchema", "SSOAP"), repos = c("http://packages.ropensci.org", "http://cran.rstudio.com"))
# devtools::install_github("ropensci/taxizesoap")
if (online) {
  message('online = TRUE')
  species_todo = sort(unique(tests2$latin_BIname))
  # species_todo = species_todo[1:2] # debug me!

  worms_species_l = list()
  for (i in 1:length(species_todo)) {
    species = species_todo[i]

    message('Querying (', i, '/', length(species_todo), '): ', species)
    time = Sys.time()
    worms_data = taxizesoap::worms_records(scientific = species, marine_only = FALSE)
    Sys.time() - time

    if (nrow(worms_data) > 0) {
      worms_data = cbind(worms_data, latin_BIname = species)
    } else {
      worms_data = worms_data
    }

    worms_species_l[[i]] = worms_data
  }
  saveRDS(worms_species_l, file.path(cachedir, 'worms_species_list.rds'))

} else {
  message('online = FALSE')
  worms_species_l = readRDS(file.path(cachedir, 'worms_species_list.rds'))
}

worms_sp = rbindlist(worms_species_l)
# Sometimes there are more results for the input (lain_BIname) that's why 
lookup_worms_sp = worms_sp[ , lapply(.SD, function(x) min(as.numeric(x))),
                            by = latin_BIname, .SDcols = c('isFreshwater', 'isMarine', 'isTerrestrial')]
setnames(lookup_worms_sp, c('latin_BIname', 'isFre_w', 'isMar_w', 'isTer_w'))
```

## Freshwaterecology.info

```{r cleaning}
rm(worms_sp, worms_species_l)
rm(worms_fam, worms_family_l)
rm(tests2)
```

### 2018-07-09 src/paper.Rmd
publishing orphan commit

---
title: "Untitled"
author: "AS"
date: "December 4, 2017"
output: pdf_document
---

# from the deleted file article hints (somewhen created in winter 2016/17)
INTRODUCTION:

(Lorenz et al., 2016)
Lack of information on small water bodies (SWB)
Golla & Horney (unpublished data):
	97% of the 321,128 German lentic WB are smaller than 0.05km² surface area
	91.3% of the 544,967 German streams <3m width

Questions: lotic SWB: 	How did they find out?
				How does this translate into catchment area ? 
				Will their paper be finished? How to cite them?

Include (Biggs et al., 2016)

Have a look through (Ulrich et al., 2015). Why do they have so much information on Bavaria?

Agricultural land use effects ecological conditions of aquatic systems (Feld, 2013)

METHODS:

Account for geographical effects in (Feld, 2013):
Ecoregion,
longitude/latitude (decimal degrees) - https://en.wikipedia.org/wiki/Decimal_degrees
altitude (m.a.s.l.)
country (Germany, France)

Boosted Regression Trees – vignette? (Elith and Leathwick, 2016)
allows for correlated predictor variables

DISCUSSION:

data originates from governmental monitoring programs, grab samples  → underestimating the real effects probably (Malaj et al., 2014)

Anthropogenic factors like application regime and intensity are hard to address

Criticize (Van Dijk et al., 2013) because of the usage of lm  (+ high variability + justification), made up groups,

### 2018-07-09 src/source_bfg_monitoring.Rmd
publishing orphan commit

---
title: "Untitled"
author: "AS"
date: "March 26, 2018"
output: html_document
---

## Query bfg_monitoring data base [PUT AWAY COMPLETELY IN THE FUTURE!]
```{r DATA-daphnia-from-bfg-monitoring, eval = FALSE}
if (online) {
  #! FROM BFG MONITORING DATA BASE (not very elegant)!
  drv = dbDriver("PostgreSQL")
  con = dbConnect(drv, user = DBuser, dbname = DBname_1, host = DBhost, port = DBport, password = DBpassword)

  dm = dbGetQuery(con, "SELECT var.variable_id, var.name AS subst_name, var.casnr, prop.lc50_dm_fin AS value, prop.lc50_dm_fin_source AS source, var.psm_type
                        FROM phch.phch_variables var
                        JOIN phch.phch_variables_prop prop ON var.variable_id = prop.variable_id
                        WHERE var.psm_type IN ('fungicide', 'herbicide', 'insecticide', 'metabolite');")
  setDT(dm)

  dbDisconnect(con)
  dbUnloadDriver(drv)

  fwrite(dm, file.path(cachedir, 'bfg_dm_data.csv'))

} else {
  dm = fread(file.path(cachedir, 'bfg_dm_data.csv'))
}

# Biology
dm[ , `:=`
    (latin_BIname = 'Daphnia magna')]
     #genus = 'Daphnia',
     #family = 'Daphniidae') ]
# Other
dm[ , cas := casconv(casnr) ]
dm[ , unit := 'µg/L' ]
dm[ , source := paste0('bfg_', source) ]
dm[ , duration := NA ]
dm[ , endpoint := 'ec50' ]

dm = dm[ !is.na(value) ]

# Column order
setcolorder(dm, c('variable_id', 'subst_name', 'casnr', 'cas', 'value', 'unit', 'psm_type', 'source', 'duration', 'endpoint', 'latin_BIname'))
```

### 2018-07-09 src/source_epa.Rmd
publishing orphan commit

---
title: "Untitled"
author: "Andreas Scharmüller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
editor_options:
  chunk_output_type: console
---

# Setup
```{r setup}
#rm(list = ls())
# switches
online = FALSE
local = FALSE # have you built the EPA ECOTOX DB locally?

# packages
require(data.table)
require(RPostgreSQL)

# variables
cachedir = '/home/andreas/Documents/Projects/species-sensitivity/cache'
fundir = '/home/andreas/Documents/Projects/species-sensitivity/functions'

# source
source(file.path(fundir, 'casconv.R'))
source('/home/andreas/Documents/UBA/Project/R/src/DB_access.R')
```

# Load PSM data
```{r load-psm}
psm = readRDS(file.path(cachedir, 'psm2.rds'))
```

# Query EPA data
```{r query-epa, eval=TRUE, include = FALSE}
if (online) {
  todo = unique(psm$casnr) # remove duplicates
  todo = todo[which(nchar(todo) > 0)] # 495 substances!
  # todo = '122349' # debuging!
  # todo = '15972608' # debuging
  # todo = '15972608' # debug me
  drv = dbDriver("PostgreSQL")
  if (local) {
    con = dbConnect(drv, user = DBuserL, dbname = DBnameL, host = DBhostL, port = DBportL, password = DBpasswordL) # local  
  } else {
    con = dbConnect(drv, user = DBuser, dbname = DBetox, host = DBhost, port = DBport, password = DBpassword) # on server  
  }

  #epa1 <- data.table()
  epa1_list <- list()
  for (i in seq_along(todo)) {
    casnr <- todo[i]
    d <- dbGetQuery(con, paste0("
  SELECT
  	tests.test_cas::varchar AS casnr,
    chemicals.chemical_name,
    tests.test_characteristics, -- how stupid to hide names here
    chemicals.ecotox_group AS chemical_group,
    species.latin_name,
    --species.variety,
    --species.subspecies,
    --species.species,
    species.genus,
    species.family,
    --species.tax_order,
    --species.class,
    --species.superclass,
    --species.subphylum_div,
    --species.phylum_division,
    --species.kingdom,
    tests.organism_habitat,
    tests.subhabitat,
    -- concentration + unit conversion	
	CASE
	  WHEN unit_convert.conv = 'yes'
	    THEN cast_to_num(results.conc1_mean) * unit_convert.multiplier
	  WHEN unit_convert.conv = 'no'
	    THEN cast_to_num(results.conc1_mean) -- same cast as above
	  ELSE cast_to_num(results.conc1_mean) 
	END AS conc1_conv,
	CASE
	  WHEN unit_convert.conv = 'yes'
	    THEN unit_convert.unit_conv
	  WHEN unit_convert.conv = 'no'
	    THEN results.conc1_unit
	  ELSE results.conc1_unit
	END AS conc1_unit_conv,
  results.conc1_mean, -- fo safety reasons to compare to converted
  results.conc1_unit, -- fo safety reasons to compare to converted
	-- test duration
	CASE
	  WHEN duration_convert_as.conv = 'yes'
	    THEN cast_to_num(results.obs_duration_mean) * duration_convert_as.multiplier
	  WHEN duration_convert_as.conv = 'no'
	    THEN cast_to_num(results.obs_duration_mean)
	  ELSE cast_to_num(results.obs_duration_mean)
	END AS obs_duration_conv,
	CASE
	  WHEN duration_convert_as.conv = 'yes'
	    THEN duration_convert_as.unit_conv
	  WHEN duration_convert_as.conv = 'no'
	    THEN results.obs_duration_unit
	  ELSE results.obs_duration_unit
	END AS obs_duration_unit_conv,
    results.endpoint,
    results.effect,
    -- type of concentration (e.g. ai, formulation etc.)
    results.conc1_type,
    tests.reference_number,
    refs.author,
    refs.title
  FROM ecotox.tests
    LEFT JOIN ecotox.results ON tests.test_id = results.test_id
    RIGHT JOIN ecotox.species ON tests.species_number = species.species_number
    LEFT JOIN ecotox.chemicals ON tests.test_cas = chemicals.cas_number
    LEFT JOIN ecotox.refs ON tests.reference_number = refs.reference_number
    LEFT JOIN lookup.unit_convert ON results.conc1_unit = unit_convert.unit
    LEFT JOIN lookup.duration_convert_as ON results.obs_duration_unit = duration_convert_as.unit
    --LEFT JOIN lookup.ecotox_group_convert ON species.ecotox_group = ecotox_group_convert.ecotox_group
  WHERE tests.test_cas = ", casnr, "
    -- 15972608 -- casnr for debuging
    -- endpoints:
    AND results.endpoint IN ('EC50', 'EC50/', 'EC50*', 'EC50*/', 'LC50', 'LC50/', 'LC50*', 'LC50*/')
    -- duration:
    AND (results.obs_duration_unit = 'h' AND cast_to_num(results.obs_duration_mean) BETWEEN 24 AND 360
      OR results.obs_duration_unit = 'd' AND cast_to_num(results.obs_duration_mean) BETWEEN 0 AND 15)
    -- empty result cell? 
    AND results.conc1_mean != 'NR'
    -- organism habitat:
    AND tests.organism_habitat NOT ILIKE 'soil'
    -- subhabitat:
    --AND tests.subhabitat IN ('P', 'R', 'L', 'E') -- Palustrine, Riverine, Lacustrine, Estuarine (not enough data provided!)
    AND tests.subhabitat NOT IN ('D', 'F', 'G', 'M')
    --AND results.effect IN ('MOR', 'ITX') --?
    AND coalesce(species.genus, '') <> '' -- same as !=
  ;"))

    ## DEBUGINGN:
    # epa1[ endpoint %in% c('EC50', 'EC50/', 'EC50*', 'EC50*/', 'LC50', 'LC50/', 'LC50*', 'LC50*/'), .N, endpoint ][order(-N)]
    # epa1[ value == 'NR' ]
    # range(epa1$obs_duration_conv)
    # epa1[ duration <= 120 ]
    # epa1[ , .N, duration_unit]
    # epa1[ , .N, subhabitat ]
    # epa1[ ]$
    # epa1[ , .N, genus]
    # epa1[ nchar(genus) < 3 ]
    ##

    if (nrow(d) > 0) {
      d = merge(d, psm[ ,c('casnr', 'variable_id', 'subst_name', 'psm_type')],
                by.x = 'casnr', by.y = 'casnr',
                all.x = TRUE)
    }

  message('testing casnr: ', casnr, '\nname: ', d$chemical_name,
          '\npsm_type: ', d$psm_type, '\n nobs: ', nrow(d)) # timestamp(quiet = TRUE)

  #epa1 <- plyr::rbind.fill(epa1, d)
  epa1_list[[i]] <- d
  names(epa1_list)[i] <- casnr

}

  dbDisconnect(con)
  dbUnloadDriver(drv)

  ## save
  saveRDS(epa1_list, file.path(cachedir, 'source_epa1_list.rds'))

} else {
  epa1_list = readRDS(file.path(cachedir, 'source_epa1_list.rds'))
}

epa1 = rbindlist(epa1_list)

## refine effect indices (could also be done in the query)
effect = epa1[ , .N, effect][order(-N)]
effect[ , effect := gsub('~|/', '', effect) ]
effect = effect[ , .(N = sum(N)), effect]
fwrite(effect, file.path(tempdir(), 'effect_endpoints.csv'))

#epa1 = epa1[ effect %like% '(?i)MOR|POP|ITX|GRO' ]
epa1[ , effect := gsub('~|/', '', effect) ]
# https://cfpub.epa.gov/ecotox/help.cfm?help_id=CONTENTFAQ&help_type=define&help_back=1#asterisk
# ~ (leading tilde) - 
# / (trailing slash) - denotes that a comment is associated with the data
# MOR - mortality; POP - population; ITX - Intoxication; GRO - Growth; BEH - Behaviour
# NOC - Multiple or undefined; MPH - Morphology; PHY - Physiology; DVP - Development
# REP - Reproduction; BCM - Biochemistry; ENZ - Enzyme; FDB - Feeding behaviour;
# AVO - Avoidance; HIS - Histology; INJ - Injury; PRS - Ecosystem Process; CEL- Cell
# Gen - Genetics; IMM - Immunological; NR - Not reported

## resolve substance names manualy
# mostly taken from Sigmar Aldrich - careful!
miss = data.table(casnr = c("26002802", "26172554", "26530201", "2814202", "39515407", "39515418", "51036", "54406483", "584792", "72963725", "962583", "134623", "3380345"),
                  subst_name = c('Phenothrin', 'Methylchloroisothiazolinone', 'Octhilinone', 'Pyrimidinol', 'Cyphenothrin', 'Fenpropathrin', 'Piperonylbutoxide', 'Empethrin', 'Allethrin', 'Imiprothrin', 'Diazoxon', 'DEET', 'Triclosan'))

epa1[miss, on = 'casnr',
     subst_name := ifelse(is.na(subst_name), i.subst_name, subst_name)]

# adding qualifier column
# * means that this value is calculated 'cause it didn't meet epa standards (22 cases). This is not done anymore recently. see: https://cfpub.epa.gov/ecotox/help.cfm?help_id=CONTENTFAQ&help_type=define&help_back=1#asterisk
# + probably stands for more than (16 cases)
epa1[ , qualifier := trimws(gsub('[0-9\\.]+|E\\+|E\\-', '', conc1_mean)) ]
epa1[ qualifier == '', qualifier := '=' ]
epa1 = epa1[ qualifier != '+' ] # removing + entries

# delete conc1_mean and conc1_conv as they are only for comparing
epa1[ , c('test_characteristics', 'conc1_mean', 'conc1_unit', 'qualifier') := NULL ]

# delete reference information author and title:
epa1[ ,c('author', 'title') := NULL ]

#### Restrictions ----
#rm(d, epa1_list)
epa1 = epa1[ conc1_unit_conv %in% c('ug/L', 'ul/L') ]

#! check this in future. Which effect types should be included?
# epa1[ , .N, effect]

## CONTINUE HERE!!!!!
# 1) What is fd in ul/L fd??
# 2) Check conversions I added?
# 3) Should the changes be included to Edi's github? No, public
# 4) run all the script! Check every part
# 5) build WORMS query for family AND Species!
# 6) continue at sort(todo_habitat[69:80])

#### Additions ----
# harmonize NAs
epa1[ subhabitat %in% c('NC', '--', 'NR'), subhabitat := NA ]

# transform all Hybrid X to capital letters => they aren't seen as species-names in the next REGEX
epa1[ , latin_BIname := gsub('\\sx\\s', ' X ', latin_name, ignore.case = TRUE) ]
epa1[ , latin_BIname := gsub('([A-z]+)\\s([a-z]+\\s)(.+)*', '\\1 \\2', latin_BIname) ]
epa1[ , latin_BIname := sub('\\s+$', '', latin_BIname) ] # delete trailing \\s
epa1[ , latin_short := paste0(substr(latin_name,1,1),
                                     '. ',
                                     gsub('([a-z]+)\\s([a-z]+)', '\\2', latin_BIname, ignore.case = TRUE)) ]

# add source column
epa1[ , source := 'epa_ecotox' ]

# cas
epa1[ , cas := casconv(casnr) ]

# set all "" to NA
for (i in names(epa1)) {
  epa1[get(i) == "", (i) := NA]
}

# delete columnms & change order
setcolorder(epa1, c('variable_id', 'casnr', 'cas', 'subst_name', 'chemical_name', 'psm_type', 'chemical_group', 'conc1_conv', 'conc1_unit_conv', 'obs_duration_conv', 'obs_duration_unit_conv', 'endpoint', 'effect', 'conc1_type', 'organism_habitat', 'subhabitat',  'latin_name', 'latin_BIname', 'latin_short', 'genus', 'family', 'source', 'reference_number'))

# change names
setnames(epa1, c('variable_id', 'casnr', 'cas', 'subst_name', 'chemical_name', 'psm_type','chemical_group', 'value', 'unit', 'duration', 'duration_unit', 'endpoint', 'effect', 'conc_type', 'habitat', 'subhabitat',  'latin_name', 'latin_BIname', 'latin_short', 'genus', 'family_epa', 'source', 'ref_num'))

# Reduce columns to match PPDB and bfg_monitoring cols:
epa2 = epa1[ , c('variable_id', 'subst_name', 'casnr', 'cas', 'value', 'unit', 'psm_type', 'source', 'ref_num', 'duration', 'endpoint', 'effect', 'latin_BIname', 'family_epa') ]

# checks
subst_check = 
  epa2[ is.na(casnr) | casnr == '' |
        is.na(cas) | cas == '' |
        is.na(subst_name) | subst_name == '' ]

if (nrow(subst_check) != 0) {
  warning(nrow(subst_check), ' missing CAS, CASNR or substance names.')
}
```

# Durations
```{r duration}
epa2[ , duration2 := ifelse(duration >= 21 & duration <= 27, 24,
                     ifelse(duration >= 46 & duration <= 50, 48,
                     duration))]
# epa2[ , .N, duration][N > 10][order(duration)] # debuging
```

# Cleaning
```{r cleaning}
rm(d, drv, con, epa1, miss, subst_check, psm)
```

### 2018-07-09 src/source_malaj.Rmd
publishing orphan commit

---
title: "Source Malaj data"
author: "Andreas Scharmüller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
editor_options:
  chunk_output_type: console
---

# Setup
```{r setup}
# packages
require(data.table)

# variables
cachedir = '/home/andreas/Documents/Projects/species-sensitivity/cache'
fundir = '/home/andreas/Documents/Projects/species-sensitivity/functions'

# source
source(file.path(fundir, 'casconv.R'))
```

# Load PSM data
```{r load-psm}
psm = readRDS(file.path(cachedir, 'psm2.rds'))
```

# Load Malaj data
```{r load-malaj}
malaj = fread('/home/andreas/Documents/data/malaj2014/Toxicity_data_malaj_2014.csv')
```

## Data cleaning
```{r malaj-cleaning}
names(malaj) = tolower(names(malaj))
malaj = malaj[ ,c('casnumber', 'cas', 'lc50 dm', 'source details_dm', 'type of data_dm', 'lc50 pim', 'source details_pim', 'type of data_pim', 'lc50 sele', 'source details_sele', 'type of data_sele', 'code', 'determinand_hazsubs')]
setnames(malaj, c('casnr', 'cas', 'lc50_dm', 'lc50_dm_source', 'lc50_dm_type', 'lc50_pim', 'lc50_pim_source', 'lc50_pim_type', 'lc50_sel', 'lc50_sel_source', 'lc50_sel_type', 'psm_type', 'subst_name'))

# cas
malaj = malaj[ !is.na(casnr) ] # no single pesticides
malaj[ , cas := casconv(casnr) ]

# replace '' with NA
for(i in names(malaj)) {
  malaj[get(i) == '', (i) := NA ]
}

# taxa
malaj[ , `:=` (dm = 'Daphnia magna',
               pp = 'Pimephales promelas',
               ps = 'Pseudokirchneriella subcapitata')]

# melt
malaj_m = melt(malaj,
               id.vars = c('casnr', 'cas', 'psm_type', 'subst_name'),
               measure.vars = list(
                 c('dm', 'pp', 'ps'),
                 c('lc50_dm', 'lc50_pim', 'lc50_sel'),
                 c('lc50_dm_source', 'lc50_pim_source', 'lc50_sel_source'),
                 c('lc50_dm_type', 'lc50_pim_type', 'lc50_sel_type')),
               value.name = c('latin_BIname', 'value', 'source', 'type'))

# other variables
malaj_m[ , `:=` (endpoint = 'lc50',
                 variable_id = NA,
                 unit = 'ug/L',
                 duration = ifelse(latin_BIname == 'Daphnia magna', 48,
                            ifelse(latin_BIname == 'Pimephales promelas', 96,
                            ifelse(latin_BIname == 'Pseudokirchneriella subcapitata', 4896, NA))),
                 source = 'malaj')]
malaj_m[ , variable := NULL]

malaj_m = malaj_m[ !is.na(value) ]

# column order
setcolorder(malaj_m, c('variable_id', 'subst_name', 'casnr', 'cas', 'value', 'unit', 'psm_type', 'source', 'type', 'duration', 'endpoint', 'latin_BIname'))
```

## Filter
```{r filter}
malaj_m = malaj_m[ !type %in% c('B', 'P/1', 'P/2') ] 
```

# Cleaning
```{r cleaning}
rm(malaj, psm)
```

### 2018-07-09 src/source_ppdb.Rmd
publishing orphan commit

---
title: "Source PPDB data"
author: "Andreas Scharmüller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
editor_options:
  chunk_output_type: console
---

# Setup
```{r setup}
# packages
require(data.table)
require(webchem)

# switches
online = FALSE

# variables
cachedir = '/home/andreas/Documents/Projects/species-sensitivity/cache'
fundir = '/home/andreas/Documents/Projects/species-sensitivity/functions'

# source
source('/home/andreas/Documents/UBA/Project/R/src/DB_access.R')
source(file.path(fundir, 'ppdb-0-0-5.R'))
source(file.path(fundir, 'ppdb_own_functions.R'))
```

# Load PSM data
```{r load-psm}
psm = readRDS(file.path(cachedir, 'psm2.rds'))
```

# Query PPDB data
```{r query, eval = TRUE}
# don't switch online on too often!
if (FALSE) {
  # create data.frame of indices before the loop - otherwise it's done every time!
  # ppdb_query() from version 0.05 only excepts cas (separated with hyphen)!
  idx = ppdb_buildidx()

  psm_fin = psm[ nchar(cas) > 0 ]
  # 21 entries which don't have a CAS: Mostly like: 'Azoxystrobin-CA'.  Azoxystronin itself has a CAS
  psm_fin = psm_fin[ !duplicated(cas) ]
  todo = psm_fin$cas

  # todo = todo[1:3] # debug me!

  ppdb_l = list()
  for (i in 1:length(todo)) {
    cas = todo[i]
    subst = psm_fin$subst_name[i]

    message(i, ') Querying: ', cas, ': ', subst)
    res = ppdb_query(cas, verbose = FALSE, index = idx)

    ppdb_l[[i]] = res
    names(ppdb_l)[i] = cas
  }

  saveRDS(ppdb_l, file.path(cachedir, 'ppdb_l.rds'))
  saveRDS(idx, file.path(cachedir, 'ppdb_idx.rds'))

} else {
  ppdb_l = readRDS(file.path(cachedir, 'ppdb_l.rds'))
  idx = readRDS(file.path(cachedir, 'ppdb_idx.rds'))
}

rm(idx)
```

### Extracting
```{r extracting}
## PPDB-scrap ----
# etox information
ppdb_dt = ppdb_etox(ppdb_l)
ppdb_dt[psm[, c('cas', 'subst_name')],
        subst_name := i.subst_name, on = 'cas' ]
# psm_type
ppdb_psm_dt = ppdb_psm_type(ppdb_l)
ppdb_dt = merge(ppdb_dt, ppdb_psm_dt, on = 'cas')
# PPDB name
ppdb_subst = ppdb_name(ppdb_l)
```

### Extra
For CAS that have more than one PPDB entry. In the case of Propiconazole there are __cis-propiconazole__, __propiconazole__, __trans-propiconazole__. The first one is taken by default, which lacks ecotoxicological data. However we want the second entry as this is the main compound with the most data. This should be included in ppdb_query() in the future. For now I substitute the values manualy.
```{r extra}
# looked up by hand
## Propiconazole
propiconazole = data.table(variable_id = NA,
                           subst_name = 'Propiconazole',
                           casnr = '60207901',
                           cas = '60207-90-1',
                           value = c(2.6, 0.068, 10.2, 0.31, 0.37, 8.0, 25.0, 4.9, 0.093),
                           unit = c('mg/L','mg/L','mg/L','mg/L','mg/L','mg/L','mg/kg','mg/L','mg/L'),
                           psm_type = 'fungicide',
                           source = 'ppdb',
                           duration = c(96,504,48,504,96,672,672,168,72),
                           duration_unit = 'h',
                           endpoint = c('lc50', 'noec', 'lc50', 'noec', 'lc50', 'noec', 'NOEC', 'EC50', 'EC50'),
                           latin_BIname = c('Leiostomus xanthurus', 'Oncorhynchus mykiss', 'Daphnia magna', 'Daphnia magna', 'Americamysis bahia', 'Chironomus riparius', 'Chironomus riparius', 'Lemna gibba', 'Navicula seminulum'),
                           ppdb_group = c('fish', 'fish', 'aquatic invertebrates', 'aquatic invertebrates', 'aquatic crustaceans', 'sediment dwelling organisms', 'sediment dwelling organisms', 'aquatic plants', 'algae'),
                           quality = c('A5', 'A5', 'A5', 'A5', 'F3', 'A5', 'A5', 'F3', 'A4'))

ppdb_dt_extr = rbindlist(list(propiconazole))

## PPDB-extra ----
ppdb_dt = rbindlist(list(ppdb_dt, ppdb_dt_extr), fill = TRUE)
```

### Cleaning
```{r cleaning}
# remove qualifier entiries > and <
ppdb_dt = ppdb_dt[ is.na(qualifier) | qualifier == '=' ]
ppdb_dt[ is.na(qualifier)]
# multiply value by 1000 to get µg/L
ppdb_dt[ unit == 'mg/L', `:=`
                         (value = value * 1000,
                          unit = 'ug/L')]
# only some Chiro NOEC values!
ppdb_dt[ unit == 'mg/kg', `:=`
                         (value = value * 1000,
                          unit = 'ug/kg')]
# change everything to hour (could already be done in the function - change this in the future!)
ppdb_dt[ , duration := ifelse(duration_unit == 'day', duration*24, duration) ]

ppdb_dt[ , duration_unit := 'hour' ]
ppdb_dt[ , casnr := gsub('-', '', cas) ]
ppdb_dt[ , variable_id := NA ]

# PPDB name
ppdb_dt[ppdb_subst, on = 'cas',
        subst_name := ifelse(is.na(subst_name), i.subst_name, subst_name)]
```

## Filter
```{r filter}
# EC50 tests
ppdb_dt = ppdb_dt[ tolower(endpoint) %in% c('lc50', 'ec50') ]
ppdb_dt = ppdb_dt[ substr(quality,2,2) %in% c('5', '4') ]
```

## Errata
```{r errata}
#### O. mykiss ----
ppdb_dt[ latin_BIname == 'Oncorhychus mykiss',
         latin_BIname := 'Oncorhynchus mykiss' ]
ppdb_dt[ latin_BIname %like% '(?i)Salmo gairdneri|Salmo gardneri|Salmo gairdneri|Salmo gardneri|Trout',
         latin_BIname := 'Oncorhynchus mykiss' ]
# from Salmonidae to O. mykiss
message('The taxon fot the following substances was set from Salmonidae to O. mykiss (like all other fish entries in PPDB - this seems to be a mistake)\n',
        paste0(ppdb_dt[ latin_BIname == 'Salmonidae' ,c('casnr', 'subst_name')],
               sep = '\n'))
ppdb_dt[ latin_BIname == 'Salmonidae',
         latin_BIname := 'Oncorhynchus mykiss' ]

#### P. promelas ----
ppdb_dt[ latin_BIname == 'Pimephalis promelas',
         latin_BIname := 'Pimephales promelas' ]
ppdb_dt[ latin_BIname == 'Pimehales',
         latin_BIname := 'Pimephales' ]

#### D. rerio ----
ppdb_dt[ latin_BIname == 'Brachydanio reri',
         latin_BIname := 'Danio rerio' ]

#### Lempomis ----
ppdb_dt[ latin_BIname == 'Lepomis machrochinus',
         latin_BIname := 'Lepomis macrochirus' ]

#### P. subcapitata ----
# a lot of spelling mistakes!
ppdb_dt[ latin_BIname %like% '(?i)Pseudokirchneriella subcaptata|Psuedokirchneriella subcapitata|Pseudokirchnella subcapitata|Pseudokirchneriella subcaptitata|Pseudokirch subcap|Pseudokircheriella subcapitata|Selenastrum capricornutum|Selenestrum capricornutum|Selenastrum|Rhapidocelis subcapitata|Raphidocelis subcapitata',
         latin_BIname := 'Pseudokirchneriella subcapitata' ]

#### Scenedesmus subspicatus ----
ppdb_dt[ latin_BIname %like% '(?i)Scenedesmus subspicutus|Scenedemus subspicatus|Scenedesmus subcapitata|Scenedesmus subspicutus',
         latin_BIname := 'Scenedesmus subspicatus' ]

#### Lemna ----
ppdb_dt[ latin_BIname == 'Lemna spp',
         latin_BIname := 'Lemna sp.']
```

## Grouping
```{r grouping }
ppdb_dt[ , .N, .(ppdb_group, duration) ]
```

## Cleaning2
```{r cleaning2}
# subset
ppdb_dt = ppdb_dt[ , c('variable_id', 'subst_name', 'casnr', 'cas', 'value', 'unit', 'psm_type', 'ppdb_group', 'source', 'duration', 'duration_unit', 'endpoint', 'latin_BIname') ]

# check
subst_check = 
  ppdb_dt[ is.na(casnr) | casnr == '' |
           is.na(cas) | cas == '' |
           is.na(subst_name) | subst_name == '' ]

if (nrow(subst_check) != 0) {
  warning(nrow(subst_check), ' missing CAS, CASNR or substance names.')
}
```

# Cleaning
```{r env-cleaning}
rm(ppdb_dt_extr, ppdb_psm_dt, propiconazole, subst_check, ppdb_subst, psm)
```

### 2018-07-09 functions/agg_group.R

@@ -1,0 +1,258 @@
+# This function filters for tests on where the same compound, taxon and
+# optionally the same duration was used.
+# It calculates the range of EC50 values if the is more than one of these
+# combinations.
+# Where the difference of max and min EC50 values surpasses a factor 10,
+# an outlier detection method is applied.
+# The data.table is then otionally filtered for non-outliers and aggregated
+# by casnr.
+
+require(data.table)
+require(outliers)
+
+agg_group = function (dt, habi = NULL, grp_par = NULL, grp_arg = NULL,
+                      dur_par = NULL, dur_arg = c(24,120),
+                      outl_rm = FALSE,  stat = 'iqr', prob = NA, lim = 1.5,
+                      plot = FALSE, chunk = 50, n_test = 5, ...) {
+
+  # debuging dur_par = 'duration2'; dur_arg = 72;
+  # plot='dot'; dt = epa3; habi = 'fresh';  outl_rm = TRUE; grp_par = 'supgroup2'; grp_arg = 'Makro_Inv'; stat = 'iqr'; prob = NA; lim = 1.5; n_test = 5; chunk = 50
+  
+  #### functions ----
+  gm_mean = function(x) exp(mean(log(x)))
+  
+  #### stops ----
+  if (is.data.frame(dt)) {
+    dt = as.data.table(dt)
+  }
+  if (!is.data.table(dt)) {
+    stop('dt is not a data.table.')
+  }
+  
+  #### pre-filtering ----
+  # habitat filter
+  if (!is.null(habi)) {
+    if (habi %like% '(?i)fresh') {
+      dt = dt[ isFre_fin == '1' ]
+      hab = 'f'
+    } else if (habi %like% '(?i)marin') {
+      dt = dt[ isMar_fin == '1' ]
+      hab = 'm'
+    } else if (habi %like% '(?i)terr') {
+      dt = dt[ isTer_fin == '1' ]
+      hab = 't'
+    } 
+  } else {
+      hab = 'n'
+      message('Not filtering for habitat (e.g fresh, marine or terrestrial).')
+  }
+  
+  # taxon filter
+  dt = dt[ get(grp_par) == grp_arg ]
+  message('Compiling ', grp_arg, ' data:')
+  
+  # duration filter
+  if (missing(dur_par) | is.null(dur_par)) {
+    dt1 = dt
+    message('No duration filter applied.')
+    dur_id = ''
[…]

### 2018-07-09 functions/agg_group2.R

@@ -1,0 +1,113 @@
+# This function filters for tests on where the same compound, taxon and
+# optionally the same duration was used.
+# It calculates the range of EC50 values if the is more than one of these
+# combinations.
+# Where the difference of max and min EC50 values surpasses a factor 10,
+# an outlier detection method is applied.
+# The data.table is then otionally filtered for non-outliers and aggregated
+# by casnr.
+
+# This functions builds on agg_group.R
+# changes: Don't use range and don't remove outliers
+
+require(data.table)
+require(outliers)
+
+agg_group2 = function (dt, habi = NULL, grp_par = NULL, grp_arg = NULL,
+                       dur_par = NULL, dur_arg = c(24,120),
+                       ...) {
+  
+  # debuging dur_par = 'duration2'; dur_arg = 72;
+  # dt = epa3; habi = 'fresh';  grp_par = 'supgroup2'; grp_arg = 'Makro_Inv'; dur_par = 'duration'; dur_arg = c(48,96)
+  
+  #### functions ----
+  gm_mean = function(x) exp(mean(log(x)))
+  
+  #### stops ----
+  if (is.data.frame(dt)) {
+    dt = as.data.table(dt)
+  }
+  if (!is.data.table(dt)) {
+    stop('dt is not a data.table.')
+  }
+  
+  #### pre-filtering ----
+  # habitat filter
+  if (!is.null(habi)) {
+    if (habi %like% '(?i)fresh') {
+      dt = dt[ isFre_fin == '1' ]
+      hab = 'f'
+    } else if (habi %like% '(?i)marin') {
+      dt = dt[ isMar_fin == '1' ]
+      hab = 'm'
+    } else if (habi %like% '(?i)terr') {
+      dt = dt[ isTer_fin == '1' ]
+      hab = 't'
+    } 
+  } else {
+    hab = 'n'
+    message('Not filtering for habitat (e.g fresh, marine or terrestrial).')
+  }
+  
+  # taxon filter
+  dt = dt[ get(grp_par) == grp_arg ]
+  message('Compiling ', grp_arg, ' data:')
+  
+  # duration filter
+  if (missing(dur_par) | is.null(dur_par)) {
+    dt1 = dt
+    message('No duration filter applied.')
+    dur_id = ''
+  } else if (is.numeric(dur_arg) & length(dur_arg) == 1) {
[…]

### 2018-07-09 functions/casconv.R

@@ -1,0 +1,24 @@
+# function to convert CAS (e.g. 82-68-8) to CASNR (e.g. 82688) and vice versa
+# debuging: tes = c('82688', '76578148', '1918167')
+
+# note: same as in /home/andreas/Documents/UBA/Project/R/functions
+
+tocas = function (x) {
+  paste(substr(x, 1, nchar(x)-3),
+        substr(x, nchar(x)-2, nchar(x)-1),
+        substr(x, nchar(x), nchar(x)),
+        sep = '-')
+}
+
+casconv = function (x) {
+  if (anyNA(x)) {
+    warning('Careful with NAs.')
+  }
+  if (all(grepl('-', x))) {
+    message('Converting from CAS to CASNR.')
+    gsub('-', '', x)
+  } else {
+    message('Converting from CASNR to CAS.')
+    tocas(x)
+  }
+}

### 2018-07-09 functions/functions.R

@@ -1,0 +1,12 @@
+## functions written during the species sensitivity project:
+
+#### 1 geometric mean --------
+gm_mean <- function(x) exp(mean(log(x)))
+
+#### 2 ksource --------
+# https://stackoverflow.com/questions/10966109/how-to-source-r-markdown-file-like-sourcemyfile-r
+# doesn't load eval=FALSE
+ksource = function(x, ...) {
+  library(knitr)
+  source(purl(x, output = tempfile()), ...)
+}

### 2018-07-09 functions/gam_quick_summary.R

@@ -1,0 +1,29 @@
+# function to extract R2 from gam and gamm objects
+# model_list = readRDS('/home/andreas/Documents/Projects/species-sensitivity/cache/moTU_sit_l.rds') # debuging
+quick_summary = function(model_list) {
+  
+  quick_summary_l = list()
+  for (i in 1:length(model_list)) {
+    mo = model_list[[i]]# model output
+    
+    if (is.null(mo)) {
+      rsq = NA_character_
+    } else if (anyNA(mo)) {
+      rsq = NA_character_
+    } else {
+      if (inherits(mo, 'gam')) {
+        rsq = summary(mo)$r.sq
+        rsq = round(rsq,2)  
+      } else if (inherits(mo, 'gamm')) {
+        rsq = summary(mo$gam)$r.sq
+        rsq = round(rsq,2)
+      }
+    }
+    quick_summary_l[[i]] = data.frame(name = names(model_list)[i],
+                                      rsq = rsq,
+                                      stringsAsFactors = FALSE)
+  }
+  out = do.call(rbind, quick_summary_l)
+  out = out[with(out, order(rsq, decreasing = TRUE)), ]
+  return(out)
+}

### 2018-07-09 functions/panel_cor.R

@@ -1,0 +1,19 @@
+panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor)
+{
+  usr <- par("usr"); on.exit(par(usr))
+  par(usr = c(0, 1, 0, 1))
+  r <- abs(cor(x, y, use = "complete.obs"))
+  txt <- format(c(r, 0.123456789), digits = digits)[1]
+  txt <- paste(prefix, txt, sep = "")
+  if (missing(cex.cor)) 
+    cex <- 0.8/strwidth(txt)
+  
+  test <- cor.test(x,y)
+  # borrowed from printCoefmat
+  Signif <- symnum(test$p.value, corr = FALSE, na = FALSE,
+                   cutpoints = c(0, 0.05, 0.1, 1),
+                   symbols = c("*", ".", " "))
+  
+  text(0.5, 0.5, txt, cex = cex * r)
+  text(.8, .8, Signif, cex = cex, col = 2)
+}

### 2018-07-09 functions/plot_outlier.R

@@ -1,0 +1,111 @@
+# packages
+require(data.table)
+require(ggplot2)
+
+# create plot function
+plot_outlier = function(dt = NULL, y = NULL, x = NULL, grp_par = 'supgroup2', grp_arg = 'Algae', chunk = 50, n_test = 5, dur_par = NULL, dur_arg = NULL, flip = TRUE, ...) {
+  
+  if (is.null(dt)) {
+    stop('Please supply a data.table.')
+  }
+  if (is.null(y)) {
+    stop('Please supply a valid y variable.')
+  }
+  if (is.null(x)) {
+    stop('Please supply a valid y variable.')
+  }
+  
+  # DEB
+  # dt = epa3f; y = 'value'; x = 'rng_id'; flip = TRUE; grp_par = 'supgroup2'; grp_arg = 'Algae';
+  # dur_par = 'duration2'; dur_arg = 72; chunk = 50; n_test = 1 # debuging
+  
+  # duration filter
+  if (is.null(dur_par)) {
+    dt1 = dt
+    message('No duration filter applied.')
+  } else if (is.numeric(dur_arg) & length(dur_arg) == 1) {
+    dt1 = dt[ get(dur_par) == dur_arg ]
+    message(dur_arg, 'h duration filter is applied.')
+  }
+  
+  ## determining range of tests per compound-taxon combination
+  dt1[ , `:=`
+        (rng_id = paste0(casnr, '_',
+                         vegan::make.cepnames(latin_BIname)),
+          rng = paste0(min(value), '-', max(value)),
+          rng_fac = round(max(value) / min(value),1),
+          N = .N),
+        by = .(casnr, latin_BIname) ]
+  
+  ## outlier detection if rng_fac >= 10
+  #stat = 'z'; prob = 0.95; lim = NA
+  stat = 'iqr'; prob = NA; lim = 1.5
+  
+  dt1[ rng_fac > 10, `:=`
+       (outliers = outliers::scores(value, type = stat, prob = prob, lim),
+        outl_type = stat,
+        outl_thresh = ifelse(stat %in% c('z', 't', 'chisq'), prob, lim)),
+       by = rng_id ]
+  dt1[ is.na(outliers), outliers := FALSE ] # otherwise in the next step NAs are lost
+  
+  ot = unique(dt1[!is.na(outl_type)]$outl_type)
+  othresh = unique(dt1[!is.na(outl_type)]$outl_thresh)
+  
[…]

### 2018-07-09 functions/ppdb-0-0-5.R

@@ -1,0 +1,227 @@
+## this function was defunct as of v0.1. In 0.0.5 it still worked ;)
+
+# same as in: /home/andreas/Documents/Projects/webchem
+
+require(stringr)
+require(xml2)
+require(rvest)
+
+ppdb_buildidx <- function(){
+  # query seach index
+  qurl <- 'http://sitem.herts.ac.uk/aeru/iupac/search.htm'
+  Sys.sleep(5)
+  tt <- read_html(qurl)
+  cont <- xml_text(xml_find_all(tt, "//script[contains(.,'# of titles present in the database')]"))
+  Encoding(cont) <- "latin1"
+  cont <- strsplit(cont, '\\n')
+  links <- cont[[1]][grepl('^links\\[', cont[[1]])]
+  links <- str_match(links, '\\\"(.*)\\"')[ , 2]
+  baseurl <- 'http://sitem.herts.ac.uk/aeru/iupac/'
+  links <- paste0(baseurl, links)
+
+  titles <- cont[[1]][grepl('^title\\[', cont[[1]])][-1]
+  # rm start and end of titles
+  titles <- str_match(titles, '\\\"(.*)\\"')[ , 2]
+  # possible titles
+  ptitles <- str_split(titles, ', ')
+  cas <- sapply(ptitles, function(x) x[[3]])
+  ppdb_idx <- data.frame(cas = iconv(cas, from = "UTF-8", to = "ASCII"),
+                         link = iconv(links, from = "UTF-8", to = "ASCII"),
+                         stringsAsFactors = FALSE)
+  # remove empty entries
+  ppdb_idx <- ppdb_idx[!(ppdb_idx$cas == '' | is.na(ppdb_idx$cas)), ]
+  # remove non-cas entries
+  trm <- c( "AE1277106", "AE1394083", "AE-F130619", "ASU 70 480 1", "D-3598" ,
+            "IN-EQW78", "IR5839" ,"methyl ester", "MON 0139", "None",  "sodium salt" )
+  ppdb_idx <- ppdb_idx[!ppdb_idx$cas %in% trm, ]
+  # remove duplicated cas entries
+  ppdb_idx <- ppdb_idx[!duplicated(ppdb_idx$cas), ]
+
+  # save(ppdb_idx, file = 'data/ppdb_idx.rda')
+  return(ppdb_idx)
+}
+
+
+ppdb_query <- function(cas, verbose = TRUE, index = NULL){
+  # cas <- '1071-83-6'
+  # cas <- '50-00-0'
[…]

### 2018-07-09 functions/ppdb_own_functions.R

@@ -1,0 +1,111 @@
+#### function to extract psm type form PPDB output ----
+ppdb_name = function(ppdb_list) {
+  names_list = lapply(ppdb_list, function(x) lapply(x, '[')$names)
+  cas = names(names_list)
+  nam = as.character(lapply(names_list, '[', 1, 2))
+  nam[ nam == 'NULL' ] = NA
+  dt = data.table(cas = cas,
+                  subst_name = nam)
+  return(dt)
+}
+  
+# from herem, same as in: /home/andreas/Documents/Projects/webchem
+ppdb_psm_type = function(ppdb_list) {
+  # extract the general list within the PPDB list for each cas
+  general_list = lapply(ppdb_list, function(x) lapply(x, '[')$general)
+  # extract the value for each cas
+  pest_list = lapply(general_list, '[', 1, 2)
+  # convert NULL to NA
+  psm_raw = unlist(lapply(pest_list, function(x) if (is.null(x)) NA else x))
+  psm_type = ifelse(grepl('inse', psm_raw, ignore.case = TRUE), 'insecticide',
+                    ifelse(grepl('herb', psm_raw, ignore.case = TRUE), 'herbicide',
+                           ifelse(grepl('fung', psm_raw, ignore.case = TRUE), 'fungicide', NA)))
+  dt = data.table(cas = names(ppdb_list),
+                  psm_type = psm_type)
+  return(dt)
+}
+
+## grep all numeric strings/vlaues
+# https://stackoverflow.com/questions/19252663/extracting-decimal-numbers-from-a-string
+# lapply(x, `[[`, 1) also takes 1st list element, hence ignores structures such as mg/L-1
+grep_dec = function(str) {
+  x = regmatches(str,gregexpr("(?>-)*[[:digit:]]+\\.*[[:digit:]]*",str, perl=TRUE))
+  as.numeric(unlist(lapply(x, `[[`, 1)))
+}
+
+ppdb_etox = function(ppdb_list) {
+  # ppdb_list = ppdb_l # debug me!
+  etox_list = lapply(ppdb_list, function(x) lapply(x, '[')$etox)
+  # etox_list = etox_list['2921-88-2'] # debug me! Chlorpyrifos
+  # etox_list = etox_list['135410-20-7'] # debug me! Acetamiprid
+  # etox_list = etox_list['60207-90-1'] # debug me! Propiconazole # get's cis-propiconazole instead of propic.
[…]

### 2018-07-09 src/fliessgewaesserbewertung.R

@@ -1,0 +1,27 @@
+#### Fliessgewaesserbewertung: Operationelle Taxaliste Mai 2011 ####
+# http://www.fliessgewaesserbewertung.de/download/bestimmung/
+require(readxl)
+xlsx1 <- '/home/andreas/Documents/UBA/data/Operationelle Taxaliste_Mai2011.xls'
+fl_bew = read_excel(xlsx1)
+
+fl_bew = as.data.frame(fl_bew)
+names(fl_bew)[1:9] <- c('group', 'family', 'dvnr', 'taxa_bliste', 'id_art', 'taxa_perlodesDB',
+                        'author', 'determination_lit', 'note')
+# fill Taxa columns with according values
+source('/home/andreas/Documents/UBA/Project/R/functions/repeat_last.R')
+fl_bew[ ,1:2] = lapply(fl_bew[ ,1:2], repeat_last)
+fl_bew = fl_bew[!is.na(fl_bew$taxa_bliste), ]
+
+## Species
+fl_bew_sp = fl_bew[-grep('Gen.', fl_bew$taxa_perlodesDB), ]$taxa_perlodesDB
+fl_bew_sp = fl_bew_sp[-grep('sp.', fl_bew_sp)]
+fl_bew_sp = fl_bew_sp[-grep('Gr.', fl_bew_sp)]
+fl_bew_sp = fl_bew_sp[-grep('/', fl_bew_sp)]
+
+## Genus
+fl_bew_sp = unique(gsub('([A-z]+)\\s([a-z]+)(\\s.+)*', '\\1 \\2', fl_bew_sp)) # due to reduction there's a duplicate
+fl_bew_gn = unique(gsub('([A-z]+)\\s(.+)*', '\\1', fl_bew_sp))
+
+## Check if Species name is logner than 2 words
+# https://stackoverflow.com/questions/8920145/count-the-number-of-words-in-a-string-in-r
+any(vapply(strsplit(fl_bew_sp, "\\W+"), length, integer(1)) != 2)

### 2018-07-09 src/freshwaterecology_data.R

@@ -1,0 +1,21 @@
+#### Species data from freshwatere cology.info ####
+## Andreas Scharmüller
+
+#! Maybe not the newest version
+fwecol = read.csv('/media/andreas/disk_AS2TB/data/koenig_NETSTORAGE_20170724/FreshwaterEcology/Freshwaterecology_komplett.csv',
+                  header = TRUE, sep = ',') 
+## Species
+fwecol_sp = unique(gsub('([a-z]+)\\s([a-z]+)\\s?.*', '\\1 \\2',
+                        fwecol$Taxon, ignore.case = TRUE))
+fwecol_sp = fwecol_sp[-grep('Micropsectra \"', fwecol_sp)]
+fwecol_sp = fwecol_sp[-grep('Parakiefferiella \"spec Kingbeek', fwecol_sp)]
+
+fwecol_sp[grep('(Holoconops)', fwecol_sp)] =
+  gsub('([A-z]+)\\s(.+)*\\s([a-z]+)', '\\1 \\3', fwecol_sp[grep('(Holoconops)', fwecol_sp)])
+
+## Genus
+fwecol_gn = unique(gsub('([A-z]+)\\s([a-z]+)', '\\1', fwecol_sp))
+
+## Check if Species name is logner than 2 words
+# https://stackoverflow.com/questions/8920145/count-the-number-of-words-in-a-string-in-r
+any(vapply(strsplit(fwecol_sp, "\\W+"), length, integer(1)) != 2)

### 2018-07-09 R/EC50aggregation_epa2.Rmd
renamed folder <src> to <R>

---
title: "Untitled"
author: "AS"
date: "April 5, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

This script is loaded by EC50aggregation.Rmd and aggregates the EPA data. There is also EC50aggregation_epa.Rmd which does the aggregation differently, by excluding outliers and not taking the median.

# Setup
```{r setup}
# packages
require(data.table)

# cachedir
cachedir = '/home/andreas/Documents/Projects/species-sensitivity/cache'
fundir = '/home/andreas/Documents/Projects/species-sensitivity/functions'
plotdir = '/home/andreas/Documents/Projects/species-sensitivity/plots'

# functions
source(file.path(fundir, 'agg_group2.R'))
```

# Load
```{r load}
tests3 = readRDS(file.path(cachedir, 'tests3.rds'))
```

# Filter
```{r epa}
epa3 = tests3[ source == 'epa_ecotox' ]
epa3 = epa3[ !supgroup %in% c('Entognatha', 'Arachnida') ]
```

# Aggregate

## Freshwater organism groups
### Input data
```{r AGGREGATE-epa-fresh-todo}
## create variable parameter list
tol_f =
  list(grp_names = c('Chironomidae96', 'Daphnia48', 'Insecta4896', 'Crustacea4896', 'Annelida4896', 'Platyhelminthes4896', 'Mollusca4896', 'Makro_Inv4896', 'Fish96', 'Algae4896', 'Plants168'),

       grp_par = c(rep('family', 2), rep('supgroup', 5), rep('supgroup2', 4)),

       grp_arg = c('Chironomidae', 'Daphniidae', 'Insecta', 'Crustacea', 'Annelida', 'Platyhelminthes', 'Mollusca', 'Makro_Inv', 'Fish', 'Algae', 'Plants'),

       dur_par = rep('duration', 11),

       dur_arg = list(96, 48, c(48,96), c(48,96), c(48,96), c(48,96), c(48,96), c(48,96), 96, c(48,96), 168))

freshw_grp = as.data.table(tol_f); freshw_grp
```

##### Output data
```{r AGGREGATE-epa-fresh-mapply-data}
ep50f_l =
  mapply(agg_group2,
         grp_par = tol_f$grp_par, grp_arg = tol_f$grp_arg, dur_arg = tol_f$dur_arg, # variable arguments
         MoreArgs = list(dt = epa3, habi = 'fresh', dur_par = 'duration'), # fixed arguments
         SIMPLIFY = FALSE)
names(ep50f_l) = tol_f$grp_names

ep50f_agg = Reduce(function(...) merge(..., by = 'casnr', all = TRUE), ep50f_l)
ep50f_agg[ ep50f_agg == Inf ] = NA; ep50f_agg[ ep50f_agg == NaN ] = NA;
fwrite(ep50f_agg, file.path(tempdir(), 'ep50f_agg.csv')) # debuging
```

### 2018-07-09 R/EC50aggregstion_epa1.Rmd
renamed folder <src> to <R>

---
title: "Untitled"
author: "AS"
date: "April 5, 2018"
output: html_document
---
# Setup
```{r setup}
# packages
require(data.table)

# cachedir
cachedir = '/home/andreas/Documents/Projects/species-sensitivity/cache'
fundir = '/home/andreas/Documents/Projects/species-sensitivity/functions'
plotdir = '/home/andreas/Documents/Projects/species-sensitivity/plots'

# functions
source(file.path(fundir, 'agg_group.R'))
```

```{r load}
tests3 = readRDS(file.path(cachedir, 'tests3.rds'))
```

```{r epa}
epa3 = tests3[ source == 'epa_ecotox' ]
```
Change: now the steps are merged
In a first step aggregate tests on the same species and compounds. Ocasion where only one such combination was found couldn't be aggregated. For ocasions, where the number of combinations was lower than 10 and their range (i.e. min to max) was above 20 the geometric mean was calculated. However as the values are verying strongly, they are difficult to assess and should be checked manually. For species compound combinations with 10 or more ocasions and a range below 20 the geometric mean was calculated. For ocasions where more than 10 tests could be found, [SSD](https://edild.github.io/ssd/) were built and an estimated vlaue derived. Though SSD is the wrong terminus in this case. It should rather be TSD - Tests sensivity distribution, as SSD would mean differen species. This was only done for EPA data as Malaj and PPDB are already aggregated.

#### Freshwater organism groups
##### Input data
```{r AGGREGATE-epa-fresh-todo}
## create variable parameter list
tol_f =
  list(grp_names = c('Plants', 'Plants168', 'Algae', 'Algae72', 'Algae4896', 'Fish96', 'Makro_Inv24', 'Makro_Inv48', 'Makro_Inv96', 'Makro_Inv', 'Mikro_Inv', 'Daphnia', 'Daphnia48', 'Chironomidae', 'Chironomidae96', 'Insecta24', 'Crustacea96', 'Mollusca4896', 'Amphibia'),

       grp_par = c(rep('supgroup2', 11), rep('family', 4), rep('supgroup', 3), 'supgroup2'),

       grp_arg = c(rep('Plants', 2), rep('Algae', 3), 'Fish', 'Makro_Inv', 'Makro_Inv', 'Makro_Inv', 'Makro_Inv', 'Mikro_Inv', rep('Daphniidae', 2), rep('Chironomidae', 2), 'Insecta24', 'Crustacea', 'Mollusca', 'Amphibia'),

       dur_arg = list(c(24,360), 168, c(24,120), 72, c(48,96), 96, 24, 48, 96, c(24,120), c(24,120), c(24,120), 48, c(24,120), 96, 24, 96, c(48,96), c(24,120)))

freshw_grp = as.data.table(tol_f); freshw_grp
fwrite(freshw_grp, file.path(tempdir(), 'EC50_freshwater_groups.csv'))
```

##### Output data
```{r AGGREGATE-epa-fresh-mapply-data}
ep50f_l =
  mapply(agg_group,
         grp_par = tol_f$grp_par, grp_arg = tol_f$grp_arg, dur_arg = tol_f$dur_arg, # variable arguments
         MoreArgs = list(dt = epa3, habi = 'fresh', dur_par = 'duration2',
                         outl_rm = TRUE, stat = 'iqr', lim = 1.5), # fixed arguments
         SIMPLIFY = FALSE)
names(ep50f_l) = tol_f$grp_names

ep50f_agg = Reduce(function(...) merge(..., by = 'casnr', all = T), ep50f_l)
ep50f_agg[ ep50f_agg == Inf ] = NA; ep50f_agg[ ep50f_agg == NaN ] = NA;
```

##### Plots
```{r AGGREGATE-epa-fresh-mapply-plots}
if (plots) {
  ep50f_plots =
    mapply(agg_group,
           grp_par = tol_f$grp_par, grp_arg = tol_f$grp_arg, dur_arg = tol_f$dur_arg, # variable arguments
           MoreArgs = list(dt = epa3, habi = 'fresh', dur_par = 'duration2',
                           outl_rm = TRUE, stat = 'iqr', lim = 1.5,
                           plot = 'all'), # fixed arguments
           SIMPLIFY = FALSE)
  names(ep50f_plots) = tol_f$grp_names

  ## save plots (takes some time!)
  subf = 'box' #! change this structures!
  plotdir2 = '/home/andreas/Documents/ownCloud'
  system(paste0('rm ', plotdir2, '/outl_plots/fresh/', subf, '/*'))
  for (i in names(ep50f_plots)) {
    plts = ep50f_plots[[i]]
    message('Saving ', i, ' plots:')
    for (j in names(plts)) {
      plt = plts[[j]]
      ggsave(plt, filename =
               file.path(plotdir2, 'outl_plots', 'fresh', subf, paste0(i, '_', j, '.png')),
                         width = 10, height = 7)
      message(j)
    }
  }
}
```

#### No habitat organism groups
##### Input data
```{r AGGREGATE-epa-fresh-todo}
## create variable parameter list
tol_n = 
  list(grp_names = c('Algae', 'Fungi'),
       grp_par = c(rep('supgroup2', 2)),
       grp_arg = c('Algae', 'Fungi'),
       dur_arg = list(c(24,120), c(24,120)))
```

##### Output data
```{r AGGREGATE-epa-fresh-mapply-data}
ep50n_l =
  mapply(agg_group,
         grp_par = tol_n$grp_par, grp_arg = tol_n$grp_arg, dur_arg = tol_n$dur_arg, # variable arguments
         MoreArgs = list(dt = epa3, habi = NULL, dur_par = 'duration2',
                         outl_rm = TRUE, stat = 'iqr', lim = 1.5), # fixed arguments
         SIMPLIFY = FALSE)
names(ep50n_l) = tol_n$grp_names

ep50n_agg = Reduce(function(...) merge(..., by = 'casnr', all = T), ep50n_l)
ep50n_agg[ ep50n_agg == Inf ] = NA; ep50n_agg[ ep50n_agg == NaN ] = NA;
```

##### Plots
```{r AGGREGATE-epa-fresh-mapply-plots}
if (plots) {
  ep50n_plots =
    mapply(agg_group,
           grp_par = tol_n$grp_par, grp_arg = tol_n$grp_arg, dur_arg = tol_n$dur_arg, # variable arguments
           MoreArgs = list(dt = epa3, habi = NULL, dur_par = 'duration2',
                           outl_rm = TRUE, stat = 'iqr', lim = 1.5,
                           plot = T), # fixed arguments
           SIMPLIFY = FALSE)
  names(ep50n_plots) = tol_n$grp_names

  ## save plots (takes some time!)
  plotdir2 = '/home/andreas/Documents/ownCloud'
  system(paste0('rm ', plotdir2, '/outl_plots/nohab/*'))
  for (i in names(ep50n_plots)) {
    plts = ep50n_plots[[i]]
    message('Saving ', i, ' plots:')
    for (j in names(plts)) {
      plt = plts[[j]]
      ggsave(plt, filename =
               file.path(plotdir2, 'outl_plots', 'nohab', paste0(i, '_', j, '.png')),
                         width = 10, height = 7)
      message(j)
    }
  }
}
```

### 2018-07-09 R/EC50sensitivity_data.Rmd
renamed folder <src> to <R>

---
title: "Species sensitivity against pesticides"
author: "Andreas Scharmüller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  # pdf_document: default
  html_document:
    toc: true
    number_sections: yes
---

# Packages and functions

In a first step we define some variables and load required R-packages.
```{r setup, eval=TRUE, include=FALSE, echo=TRUE}
## switches
online = FALSE
plots = FALSE

## variables
cachedir = '/home/andreas/Documents/Projects/species-sensitivity/cache'
srcdir = '/home/andreas/Documents/Projects/species-sensitivity/src'
datadir = '/home/andreas/Documents/Projects/species-sensitivity'
fundir = '/home/andreas/Documents/Projects/species-sensitivity/functions'

## R-Packages
require(RPostgreSQL)
require(plyr)
require(webchem)
require(ggplot2)
require(scales)
require(gridExtra)
require(zoo)
require(cowplot)
require(RCurl)
require(readxl)
require(data.table)
require(stringr)
require(knitr)
require(xtable)
require(DT)

## Source scripts
source('/home/andreas/Documents/UBA/Project/R/src/DB_access.R')
source('/home/andreas/Documents/UBA/Project/R/src/gg_theme_AS.R')
source(file.path(fundir, 'ppdb-0-0-5.R')) # sourcing deprecated ppdb-query (was still active in V 0.0.5)
source(file.path(fundir, 'functions.R')) # functions
source(file.path(fundir, 'casconv.R'))
```

# Load pesticide data

The data for our analyses is stored in the bfg_monitoring data base. We query information on the 484 substances that were found between 2005 and 2015 in sites in small streams by the governmental monitoring. Additionally data from a sampling campaign in Romania is loaded.

## PSM bfg_monitoring
```{r DATA-bfg, eval=TRUE, include=TRUE, echo=TRUE}
if (online) {
  drv <- dbDriver("PostgreSQL")
  con <- dbConnect(drv, user = DBuser, dbname = DBname_1, host = DBhost, port = DBport, password = DBpassword)

  psm_db = dbGetQuery(con, "SELECT DISTINCT ON (name) variable_id, name AS subst_name, cas, casnr, psm_type, pan_class
                             FROM phch.phch_variables;")
  setDT(psm_db)

  sam_ezg = dbGetQuery(con, "SELECT var.variable_id, var.cas, var.casnr, var.name AS subst_name, var.psm_type, var.rak_uba0317,
                             var.ger_auth_2015, 
                             prop.lc50_dm_fin,
                             samp.date, samp.value_fin, samp.site_id,
                             rain.rain_md, rain_db,
                             site.area_sqkm, site.atkis_perc
                            FROM views.psm_samples_2005 samp
                            JOIN phch.phch_variables var ON var.variable_id = samp.variable_id
                            JOIN phch.phch_variables_prop prop ON var.variable_id = prop.variable_id
                            JOIN phch.phch_ec50 ec50 ON var.variable_id = ec50.variable_id
                            JOIN spatial_derived.land_use_ezg5000 site ON samp.site_id = site.site_id
                            JOIN spatial_derived.psm_samples_2005_rain rain ON samp.sample_id = rain.sample_id
                            WHERE var.psm_type IN ('insecticide', 'herbicide', 'fungicide', 'metabolite')")
  setDT(sam_ezg)

  dbDisconnect(con)
  dbUnloadDriver(drv)

  saveRDS(psm_db, file.path(cachedir, 'psm_db.rds'))
  saveRDS(sam_ezg, file.path(cachedir, 'sam_ezg.rds'))

} else {
  psm_db = readRDS(file.path(cachedir, 'psm_db.rds'))
  sam_ezg = readRDS(file.path(cachedir, 'sam_ezg.rds'))
}

psm_db = psm_db[ psm_type %in% c('fungicide', 'herbicide', 'insecticide', 'metabolite')]
psm_db[ casnr == '', casnr := NA ]
psm_db[ , cas := ifelse(!is.na(casnr), casconv(casnr), NA) ]

## checks
if (any(duplicated(psm_db$casnr))) {
  warning('Duplicated CASNR')
  psm_db[ !is.na(casnr) & duplicated(casnr) ]
  message('Checked on 2018-03-15.\nDuplicate is Desehtylsimazin & Desethylatrazin.')
}
if (anyNA(psm_db$casnr)) {
  warning('CASNR with NA.')
  message('Checked on 2018-03-15.\nNAs are mostly metabolites and sumed pesticides.')
}
if(any(!is.na(psm_db$casnr) & psm_db$casnr == ''))
  warning('CASNR with \'\'.')
```

## PSM Romania
```{r DATA-romania, eval=TRUE, include=TRUE, echo=TRUE}
# files from Verena
# 0) new file:
psm_rom0 = fread('/home/andreas/Documents/data/romania/det_compounds_with_CAS.csv')
psm_rom0[ , casnr := casconv(cas) ]
names(psm_rom0)[1] = 'subst_name'
any(duplicated(psm_rom0$casnr))

#### old ----
# 1) first file:
psm_rom1 = fread('/home/andreas/Documents/UBA/data/PPP_without_EC_50_Verena_Romania.csv')
names(psm_rom1)[3] = 'subst_name'; names(psm_rom1)[6] = 'active_comp'
psm_rom1 = psm_rom1[ ,c('subst_name', 'active_comp', 'casnr')]
psm_rom1[ , casnr := as.character(casnr) ]
psm_rom1[ , cas := casconv(casnr) ]
psm_rom1 = psm_rom1[ !psm_rom1$casnr %in% psm_rom0$casnr ]

# 2) second file:
psm_rom2 = fread('/home/andreas/Documents/UBA/data/Missing_compounds.csv')
psm_rom2 = psm_rom2[ ,-c(1,3)]
psm_rom2[ , casnr := as.character(casnr) ]
psm_rom2[ , cas := casconv(casnr) ]
psm_rom2 = psm_rom2[ !psm_rom2$casnr %in% c(psm_rom0$casnr, psm_rom1$casnr)]
names(psm_rom2)[1] = 'active_comp'

# bind dts
psm_rom = rbindlist(list(psm_rom0, psm_rom1, psm_rom2), fill = TRUE)

# # Check whether PSM Romania are on bfg_monitoring
# drv = dbDriver("PostgreSQL")
# con = dbConnect(drv, user = DBuser, dbname = DBname_1, host = DBhost, port = DBport, password = DBpassword)
# 
# phch_var = dbGetQuery(con, "SELECT * FROM phch.phch_variables")
# setDT(phch_var)
# 
# dbDisconnect(con)
# dbUnloadDriver(drv)
# 
# psm_rom[ casnr %in% phch_var$casnr ]
# psm_rom

## check
if (any(duplicated(psm_rom$casnr)))
  warning('Duplicated CASNR')
if (anyNA(psm_rom$casnr))
  warning('CASN with NA')
```

As we don't know what types of compounds we have in Romania we query the PPDB.
```{r DATA-romania-query, eval=TRUE}
online2 = FALSE
if (online2) {
  todo = psm_rom$cas
  # todo = todo[1:3] # debug me!

  ppdb_l = list()
  for (i in 1:length(todo)) {
    subst = todo[i]
    res = ppdb_query(subst, verbose = FALSE)
    message(i, ') Result of ', subst, ': ', res)

    ppdb_l[[i]] = res
    names(ppdb_l)[i] = subst
  }
  saveRDS(ppdb_l, file.path(cachedir, 'ppdb_l_rom.rds'))

} else {
  ppdb_l = readRDS(file.path(cachedir, 'ppdb_l_rom.rds'))
}

# function to extract psm type form PPDB output
extr_psm_type = function(ppdb_list) {
  # extract the general list within the PPDB list for each cas
  general_list = lapply(ppdb_list, function(x) lapply(x, '[')$general)
  # extract the value for each cas
  pest_list = lapply(general_list, '[', 1, 2)
  # convert NULL to NA
  psm_raw = unlist(lapply(pest_list, function(x) if (is.null(x)) NA else x))
  psm_type = ifelse(grepl('inse', psm_raw, ignore.case = TRUE), 'insecticide',
             ifelse(grepl('herb', psm_raw, ignore.case = TRUE), 'herbicide',
             ifelse(grepl('fung', psm_raw, ignore.case = TRUE), 'fungicide', NA)))
  dt = data.table(cas = names(ppdb_list),
                  psm_type = psm_type)
  return(dt)
}

ppdb_dt = extr_psm_type(ppdb_l)

## update psm_type column on PPDB query
psm_rom[ppdb_dt, psm_type := psm_type, on = 'cas']
rm(ppdb_dt, ppdb_l)
```

```{r PSM-combine, eval=TRUE}
## combine pesticide data sets:
psm_rom = psm_rom[ !casnr %in% psm_db$casnr ]
psm = merge(psm_db, psm_rom, on = 'casnr', all.x = TRUE, all.y = TRUE)

## checks
if (any(duplicated(psm$casnr))) {
  warning('Duplicated CASNR')
  warning('Checked on 2018-03-15.\nDuplicate is Desehtylsimazin & Desethylatrazin.')
  duplcas = psm[ !is.na(casnr) & duplicated(casnr) ]
  fwrite(duplcas, file.path(tempdir(), 'duplcas.csv'))
  duplcas
}
if (anyNA(psm$casnr)) {
  warning('CASNR with NA.')
  warning('Checked on 2018-03-15.\nNAs are only metabolites and sumed pesticides.')
  nocas = psm[ is.na(casnr), c('subst_name', 'cas', 'casnr', 'psm_type') ]
  fwrite(nocas, file.path(tempdir(), 'nocas.csv'))
  nocas
}
if(any(!is.na(psm$casnr) & psm$casnr == '')) {
  warning('CASNR with \'\'.')
}

psm = psm[ !is.na(casnr) ]
psm[ , cas := ifelse(!is.na(casnr), casconv(casnr), NA) ]

saveRDS(psm, file.path(cachedir, 'psm1.rds'))
rm(psm_db, psm_rom0, psm_rom1, psm_rom2)
```

## Query for additional chemical properties

### Retrieve chemical group data from the Alan Wood pesticide compendium [NOT YET FINISHED!]
```{r QUERY-aw, eval=FALSE}
# todo = psm$cas
# # todo = todo[1:2] debug!
# aw_list = list()
# for (i in 1:length(todo)) {
#   cas = psm$cas[i]
#   aw_list[[i]] = aw_query(cas, type = 'cas') # Aland wood's compendium
#   result_list = aw_list[[i]][[cas]]
#   if (!is.null(result_list[[1]]) && !is.na(result_list[[1]])) {
#       psm$aw_activity[i] = result_list$activity
#       psm$aw_subactivity[i] = result_list$subactivity
#     } else {
#       psm$aw_activity[i] = NA
#       psm$aw_subactivity[i] = NA
#     }
# }
```

# Load organism data

Invertebrate Taxa found in Germany and Taxa listed in monitoring guidlines and data bases. In order to limit toxicity test reults to only European taxa a vector of Taxa names is created building upon different Taxa lists incorporating data from the [__freshwaterecology.info database__](http://freshwaterecology.info/), taxa included in the [__Fliessgewässerbewertung database__](http://www.fliessgewaesserbewertung.de/) and taxa that are listed in a databse on invertebrates in Germany: __bfg_monitoring__.

## MZB 
data from data base bfg_monitoring, website freshwaterecolog.info and fliessgewaesserbewertung
```{r DATA-MZB, eval=TRUE, echo=TRUE, include=FALSE, message=FALSE}
if (online) {
  message('online = TRUE')
  source(file.path(srcdir, 'mzb_data.R')) # bfg_monitoring
  mzb_sp
  mzb_gn
  source(file.path(srcdir, 'freshwaterecology_data.R')) # freshwaterecology.info
  fwecol_sp
  fwecol_gn
  source(file.path(srcdir, 'fliessgewaesserbewertung.R')) # Fliessgewässerbewertung
  fl_bew_sp
  fl_bew_gn

  ## combine all MZB Tax
  mzb_sp_fin = sort(unique(c(mzb_sp, fwecol_sp, fl_bew_sp)))
  mzb_gn_fin = sort(unique(c(mzb_gn, fwecol_gn, fl_bew_gn)))

  # remove not needed
  rm(mzb_sp, fwecol_sp, fl_bew_sp)

  fwrite(list(mzb_sp_fin), file.path(cachedir, 'mzb_sp_fin.csv'))
  fwrite(list(mzb_gn_fin), file.path(cachedir, 'mzb_gn_fin.csv'))

} else {
  message('online = FALSE')
  mzb_sp_fin = unlist(fread(file.path(cachedir, 'mzb_sp_fin.csv'), sep = ','))
  mzb_gn_fin = unlist(fread(file.path(cachedir, 'mzb_gn_fin.csv'), sep = ','))
}
```

```{r DT-MZB-taxa, eval=FALSE, echo=TRUE}
mzb_sp_finNA = c(mzb_sp_fin, rep(NA, 4)) # length(mzb_fin) / 5 # real number, no recycling when creating a matrix
tbl = matrix(mzb_sp_finNA, ncol = 5, byrow = TRUE)
DT::datatable(
  tbl, colnames = 'Species',
  width = 300, 
  class = 'cell-border',
  caption = 'Benthic organisms that are listed in the bfg_monitoring data base, in the freshwatercology.info data base')
rm(mzb_sp_finNA, tbl)
```

## Diatoms
```{r DATA-Diatoms, eval=TRUE, echo=TRUE}
if (online) {
  drv = dbDriver("PostgreSQL")
  con = dbConnect(drv, user = DBuser, dbname = DBname_1, host = DBhost, port = DBport, password = DBpassword)

  dia = dbGetQuery(con, "SELECT *
                         FROM dia.dia_samples")

  dbDisconnect(con)
  dbUnloadDriver(drv)

  dia_sp_fin = unique(dia$taxon)
  dia_gn_fin = unique(sapply(strsplit(dia_sp_fin, ' '), '[', 1))

  fwrite(list(dia_sp_fin), file.path(cachedir, 'dia_sp_fin.csv'))
  fwrite(list(dia_gn_fin), file.path(cachedir, 'dia_gn_fin.csv'))

} else {
  dia_sp_fin = unlist(fread(file.path(cachedir, 'dia_sp_fin.csv'), sep = ','))
  dia_gn_fin = unlist(fread(file.path(cachedir, 'dia_gn_fin.csv'), sep = ','))
}
```

# Ecotoxicological test data

In this section we query different data bases on ecotoxicological tests. The [ETOX data base](https://webetox.uba.de/webETOX/index.do?language=en) from the German Environmental Agency (UBA - Umweltbundesamt), the [US EPA ECOTOX data base](https://cfpub.epa.gov/ecotox/)

## UBA ETOX data database [NOT YET FINISHED]
```{r DATA-ETOX, eval=FALSE, include=FALSE}

# require(httr)
# require(xml2)
# query = '210880-92-5'
# source('src/etox_function_test.R')
# etox_id = get_etoxid2(query = query, match = 'all', type = 'cas')
# 
# webchem::etox_basic(etox_id[[1]][1])
# webchem::etox_tafrgets(etox_id[[1]][1])
# webchem::etox_tests(etox_id[[1]][1])

```

## EPA ECOTOX database

The data is queried/loaded in _EC50aggregation.Rmd_ mainly into the object epa1

```{r SOURCE-EC50-aggregation}
# also loads familylookup.Rmd
ksource(file.path(datadir, 'src', 'EC50aggregation.Rmd'))
```

### Number of organism groups in the whole EPA ECOTOX database  
```{r DATA-EPA-organism-overview, eval=TRUE, echo=TRUE}
drv = dbDriver("PostgreSQL")
con = dbConnect(drv, user = DBuserL, dbname = DBnameL, host = DBhostL, port = DBportL, password = DBpasswordL)

total_tests = dbGetQuery(con, '
SELECT DISTINCT ON (c.ecotox_group_convert) c.ecotox_group_convert AS organism_group, COUNT(c.ecotox_group_convert)
FROM ecotox.tests t
RIGHT JOIN ecotox.species s ON t.species_number = s.species_number
LEFT JOIN lookup.ecotox_group_convert c ON s.ecotox_group = c.ecotox_group
GROUP BY c.ecotox_group_convert')

total_tests = setorder(as.data.table(total_tests), -count)
names(total_tests) = c('Organism group', 'N')

invisible(dbDisconnect(con))
invisible(dbUnloadDriver(drv))

# kable(na.omit(total_tests)) %>%
#   kable_styling(full_width = FALSE) # doesn't work
```

```{r KABLE-EPA-orgnism-overview, eval=TRUE, echo=TRUE}
kable(na.omit(total_tests), format = 'html', table.attr = "style='width:30%;'",
      caption = 'Number of ecotoxicological tests in the EPA ECOTOX data base per organism group (classification by EPA)')

# DT::datatable(total_tests, 
#               rownames = FALSE, 
#               width = 300, 
#               class = 'cell-border',
#               caption = 'Test results on ecotoxicological tests ordered by number of tests')
rm(total_tests)
```
Fish are the most tested organism group, followed by plants and insects.

### Query for aquatic organism ecotoxicological tests {#QUERY-organisms}
In a next step we queried the EPA ECOTOX database for ecotoxicological test results limiting the query to tests with EC/LC50 results only and a test duration range from 23 to 120 hours. Furthermore we reduced the result to the following organism groups: Algae, plants, invertebrates, crustaceans, insects, molluscs and fish. Although most of the tested insects are living in a terrestrial habitat, some quatic inseccts are within this EPA group. We also excluded tests which were assigned to a soil habitat. Although there are column organism_habitat comprises _water_, _soil_ and _non-sloil_ categories we couldn't rely just on the category _water_ as most of the column had no entries. Likewise in the subhabitat column we were only able to exclude _D - _, _F - Forest_, _G - _ and _M - Marine_ and not rely only on the categories _P - Palustrine_, _R - Riverine_, _L - Lacustrine_ and _E - Estuarian_  due to the same reason. We downloaded the [US EPA ECOTOX database version (03_14_2017)](https://cfpub.epa.gov/ecotox/) and built a local version of it following [this guide](http://edild.github.io/localecotox/).

TUs are also calculated in TU_calculation.Rmd

### Categorize EPA taxa groups

#### EC50 values and test organisms
```{r DT-EPA-tests, eval=TRUE}
DT::datatable(epa1[ ,c('cas', 'subst_name', 'value', 'psm_type', 'pan_class', 'chemical_group',
                       'family', 'genus', 'latin_BIname', 'ecotox_group')],
              colnames = c('conc' = 3, 'species' = 9, 'organism_group' = 10),
              rownames = FALSE,
              class = 'cell-border')
```

#### Most tested substances:
The most tested substances accoring to the EPA ECOTOX data base are Atrazin, Methoxyfenozid, Tebufenozid, Chlorpyrifos and p,p-DDT.
```{r DT-EPA-tests-per-substance, eval=TRUE}
DT::datatable(
  epa1[ , .N, by = .(casnr, subst_name)][order(-N)],
  rownames = FALSE,
  class = 'cell-border'
  )
```

#### Most tested organisms
```{r DT-EPA-tests-per-organism, eval=TRUE}
DT::datatable(
  epa1[ , .N, by = .(latin_BIname, family, ecotox_group)][order(-N)],
  rownames = FALSE,
  class = 'cell-border'
  )
```

#### EPA ECOTOX results checks

In order to check whether categorical varaibles in the data bases help the analysis we analyse how well they are filled in.

```{r CHECK-epa-result, eval=TRUE}
## checks and balances
table(epa1$organism_habitat, exclude = NULL) # habitat categories

table(epa1$subhabitat, exclude = NULL) # subhabitat categories (E-Estuarian, L-Lacustrine, P -, R-Riverine)
table(epa1$ecotox_group, exclude = NULL) # EPA organism group classification
summary(epa1$obs_duration_conv, na.rm = TRUE) # ecotox test duration range (hrs)
```

## Match ecotox tests with freshwater species | genera

using Taxize amongst other stuff

### Filter genera and species level

The following code excludes tests for which only the family level of the tested taxa was reported. From 2896 test results 2877 remain after this step. In a second step test entries like _Apis sp._ are marked in the column spec_lvl (0).

```{r FILTER-genus-species, eval=TRUE, echo=TRUE, include=TRUE}
# exclude taxa which aren't determined until Genus level
epa1 = epa1[ genus == gsub('(\\w+).+', '\\1', latin_name) ]
# taxa identified 'till species level
epa1[ , spec_lvl := ifelse(grepl('\\.', epa1$latin_BIname), '0', '1') ]
# Count word in Species coĺumn (expected to be two - allows for hyphen)
# all(sapply(gregexpr("\\W[-]+", epa1sp$latin_BIname), length) + 1 == 2)
## count entries with hyphens
# epa1sp[sapply(gregexpr("\\W+", epa1sp$latin_BIname), length) + 1 != 2, ]
```

```{r DEPRECATED-SAVE-image, eval=TRUE, echo=FALSE}
# source('/home/andreas/Documents/UBA/Project/R/src/functions.R')
# paste0('Saving: ', workspace.size(object = ls())$total)
# save.image(file.path(cachedir, 'EC50-sensitivity.RData'))
```

### 2018-07-09 R/familylookup.Rmd
renamed folder <src> to <R>

---
title: "Family lookup"
author: "AS"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Setup
```{r SETUP-familylookup, eval=TRUE}
## packages
require(data.table)
require(knitr)

## switches
online = F

## variables
cachedir = '/home/andreas/Documents/Projects/species-sensitivity/cache'
```

# Load data
```{r LOAD-familylookup, eval=TRUE}
tests2 = readRDS(file.path(cachedir, 'tests2.rds'))
```

# Explanation

family - family you just searched
supgroup - refer to the groups that are already here (e.g. Plants, Algae, Fungi, Fish, etc.)
group - refer to the entries that are already here (e.g. Angiospermae, Ascomycota, Perciformes etc.)
group_rank - rank that relates to epa_taxon (puting epa_taxon & epa_tx_rank here like this is a little bad practice ;))
german_name
isFre - is freshwater yes/no (1/0)
isMar - is marine 
isTer - is Terrestrial

# Plants

```{r plants-list}
# Family lookup list
lookup = data.table(family = 'Poaceae', german_name = 'Suessgraeser',
                    supgroup = 'Plants', group = 'Spermatophyta', group_rank = 'subdivision', isFre = '0', isMar = '0', isTer = '1') # some Macrophytes!
lookup = rbindlist(list(lookup,

               # plants
                ## Spermatophyta (subdivision)
               as.list(c('Nothofagaceae', "Scheinbuchen", 'Plants', 'Magnoliopsida', 'class', '0', '0', '1')),
               as.list(c('Potamogetonaceae', "Laichkrautgewächse", 'Plants', 'Spermatophyta', 'subdivision', '1', '0', '0')),
               as.list(c('Fabaceae', 'Huelsenfruechtler', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Asteraceae', 'Korbbluetler', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Brassicaceae', 'Kreuzbluetler', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')), # some Macrophytes!
               as.list(c('Solanaceae', 'Nachtschattengewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Rosaceae', 'Rosengewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Malvaceae', 'Malvengewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Cucurbitaceae', 'Kuerbisgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Amaranthaceae', 'Fuchsschwanzgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Pinaceae', 'Kieferngewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Liliaceae', 'Liliengewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Convolvulaceae', 'Windengewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Salicaceae', 'Weidengewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Cyperaceae', 'Suessgraesser', 'Plants', 'Spermatophyta', 'subdivision', '1', '0', '1')),
               as.list(c('Apiaceae', 'Doldenbluetler', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Lamiaceae', 'Lippenbluetler', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Chenopodiaceae', 'Fuchsschwanzgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')), # auf wikipedia mit Amaranthaceae gemeinsam
               as.list(c('Polygonaceae', 'Knoeterichgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Cupressaceae', 'Zypressengewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Fagaceae', 'Buchengewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Caryophyllaceae', 'Nelkengewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Euphorbiaceae', 'Wolfsmilchgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Ericaceae', 'Heidekrautgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Lemnaceae', 'Wasserlinsengewaechse', 'Plants', 'Spermatophyta', 'subdivision', '1', '0', '0')),
               as.list(c('Portulacaceae', 'Portulakgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Oleaceae', 'Oelbaumgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Aquifoliaceae', 'Stechpalmengewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Scrophulariaceae', 'Braunwurzgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Rutaceae', 'Rautengewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Myrtaceae', 'Myrtengewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Theaceae', 'Teestrauchgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Rubiaceae', 'Raetegewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Haloragaceae', 'Tausendblattgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '1', '0', '1')),
               as.list(c('Aizoaceae', 'Mittagsblumengewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Vitaceae', 'Weinrebengewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Aceraceae', 'Rosskastaniengewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Celastraceae', 'Spindelbaumgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Plantaginaceae', 'Wegerichgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Arecaceae', 'Palmengewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Apocynaceae', 'Hundsgiftgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Commelinaceae', 'Commelinagewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Oxalidaceae', 'Sauerkleegewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Linaceae', 'Leingewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Iridaceae', 'Schwertliliengewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Taxaceae', 'Eibengewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Primulaceae', 'Primelgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Molluginaceae', 'Mollugogewaechse', 'Plants', 'Spermatophyta', 'subdivision', '1', '0', '1')), # some wetland plants ?
               as.list(c('Araceae', 'Aronstabgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '1', '0', '1')), # some macrophytes (also sub-family Lemnidae)
               as.list(c('Geraniaceae', 'Storchschnabelgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Rhamnaceae', 'Kreuzdorngewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Hydrocharitaceae', 'Froschbissgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '1', '1', '0')), # macrophytes only
               as.list(c('Cornaceae', 'Hartriegelgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Lythraceae', 'Blutweiderichgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Saxifragaceae', 'Steinbrechgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Crassulaceae', 'Dickblattgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')), #! relatively few aquatic plants
               as.list(c('Podocarpaceae', 'Steineibengewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Caprifoliaceae', 'Geißblattgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Begoniaceae', 'Schiefblattgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Anacardiaceae', 'Sumachgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Juglandaceae', 'Walnussgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Araliaceae', 'Walnussgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Buxaceae', 'Buchsbaumgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Verbenaceae', 'Eisenkrautgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Asclepiadaceae', 'Seidenpflanzengewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Asparagaceae', 'Spargelgewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Taxodiaceae', NA, 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')), # aktuell keine Familie mehr
                as.list(c('Droseraceae', 'Sonnentaugewaechse', 'Plants', 'Spermatophyta', 'subdivision', '0', '0', '1')),
               as.list(c('Acoraceae', 'Kalmusartige', 'Plants', 'Angiospermae', 'clade', '1', '0', '1')), # only one genera (Kalmus), Sumpfplanze
               as.list(c('Tiliaceae', 'Lindengewaechse', 'Plants', 'Angiospermae', 'clade', '0', '0', '1')),
               as.list(c('Xanthorrhoeaceae', 'Grasbaumgewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Papaveraceae', 'Mohngewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Tamaricaceae', 'Tamariskengewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Cannabaceae', 'Hanfgewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Cuscutaceae', 'Seide', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Hypericaceae', 'Johanneskrautgewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Balsaminaceae', 'Balsaminengewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Zingiberaceae', 'Ingergewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Valerianaceae', 'Baldriangewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Typhaceae', 'Rohkoblemgewaechse', 'Plants', 'Angiospermen', 'clade', '1', '0', '0')),
               as.list(c('Lauraceae', 'Lorbeergewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Melastomataceae', 'Schwarzmundgewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Sparganiaceae', 'Rohrkolbengewaechse', 'Plants', 'Angiospermen', 'clade', '1', '0', '0')), # Veraltet, heutzutage gemeinsame Familie mit Typhaceae
               as.list(c('Elaeagnaceae', 'oelweidengewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Pontederiaceae', 'Wasserhyazinthengewaechse', 'Plants', 'Angiospermen', 'clade', '1', '0', '0')),
               as.list(c('Cannaceae', 'Blumenrohr', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Sapindaceae', 'Seifenbaumgewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Magnoliaceae', 'Magnoliengewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Berberidaceae', 'Berberitzengewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Ulmaceae', 'Ulmengewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Ranunculaceae', 'Hahnenfussgewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Piperaceae', 'Pfeffergewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Pedaliaceae', 'Sesamgewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Musaceae', 'Bananengewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Onagraceae', 'Nachtkerzengewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Tropaeolaceae', 'Kapuzinerkressengewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Orobanchaceae', 'Sommerwurzgewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Capparaceae', 'Kaperngewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Ebenaceae', 'Ebenholzgewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Hydrangeaceae', 'Hortensiengewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Marantaceae', 'Pfeilwurzgewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Betulaceae', 'Birkengewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Pittosporaceae', 'Klebsamengewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Cactaceae', 'Kakteengewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Hamamelidaceae', 'Zaubernussgewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Boraginaceae', 'Raublattgewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Polemoniaceae', 'Sperrkrautgewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Urticaceae', 'Brennnesselgewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Nyssaceae', 'Tupelogewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')), #! Sumpfpflanzen
               as.list(c('Amaryllidaceae', 'Amaryllisgewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Montiaceae', 'Quellkrautgewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Cleomaceae', NA, 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Asphodelaceae', NA, 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Altingiaceae', NA, 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Nelumbonaceae', 'Lotosblumen', 'Plants', 'Angiospermen', 'clade', '1', '0', '1')),
               as.list(c('Paeoniaceae', 'Pfingstrosen', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Phyllanthaceae', NA, 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Ptiliaceae', NA, 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),
               as.list(c('Adoxaceae', 'Moschuskrautgewaechse', 'Plants', 'Angiospermen', 'clade', '0', '0', '1')),

                ## Polypodiopsida (echte Farne) (class)
               as.list(c('Dryopteridaceae', 'Wurmfarngewaechse', 'Plants', 'Polypodiopsida', 'class', '0', '0', '1')),
               as.list(c('Lomariopsidaceae', NA, 'Plants', 'Polypodiopsida', 'clade', '1', '0', '1')),
               as.list(c('Pteridaceae', 'Saumfarngewaechse', 'Plants', 'Polypodiopsida', 'class', '0', '0', '1')),
               as.list(c('Azollaceae', 'Schwimmfarngewaechse', 'Plants', 'Polypodiopsida', 'class', '1', '0', '0')),
               as.list(c('Characeae', NA, 'Plants', 'Charophyta', 'division', '1', '0', '0'))

               ))

```

# Bryophyta (division) - Mosses
```{r bryophyta-list}
lookup = rbindlist(list(lookup,
               as.list(c('Mielichhoferiaceae', NA, 'Bryophyta', 'Bryopsida', 'class', '0', '0', '1')),
               as.list(c('Dicranaceae', NA, 'Plants', 'Bryopsida', 'class', '0', '0', '1'))
))
```

# Algae (monophyletic group) - Algen

[Heterokonts](https://en.wikipedia.org/wiki/Heterokont) are distributed between Algae and Fungi chunks, although they aren't fungi but pseudofungi (usually water moulds). Generally Heterkonts seperate between Colored groups (alga-like):_Ochrophyta (phylum) and Colorless groups: Pseudofungi (unranked), Bigyra (unranked).

```{r algae-list}
lookup = rbindlist(list(lookup,
               #### Bacteria (domain): Eubacteria (kingdom): Cyanobacteria (phylum)
               as.list(c('Nostocaceae', NA, 'Algae', 'Cyanobacteria', 'phylum', '1', '0', '0')),
               as.list(c('Stigonemataceae', NA, 'Algae', 'Cyanobacteria', 'phylum', '1', '0', '1')),
               as.list(c('Naviculaceae', NA, 'Algae', 'Cyanobacteria', 'phylum', '1', '1', '0')),
               as.list(c('Fragilariaceae', NA, 'Algae', 'Cyanobacteria', 'phylum', '1', '0', '0')),
               as.list(c('Chroococcaceae', NA, 'Algae', 'Cyanobacteria', 'phylum', '0', '1', '0')),
               as.list(c('Oscillatoriaceae', NA, 'Algae', 'Cyanobacteria', 'phylum', '1', '1', '0')),
               as.list(c('Borziaceae', NA, 'Algae', 'Cyanobacteria', 'phylum', '1', '0', '0')),
               as.list(c('Aphanizomenonaceae', NA, 'Algae', 'Cyanobacteria', 'phylum', '1', '1', '0')),
               as.list(c('Rivulariaceae', NA, 'Algae', 'Cyanobacteria', 'phylum', '0', '1', '0')),
               as.list(c('Fortieaceae', NA, 'Algae', 'Cyanobacteria', 'phylum', '1', '1', '0')),
               as.list(c('Microcystaceae', NA, 'Algae', 'Cyanobacteria', 'phylum', '1', '1', '0')),
               as.list(c('Merismopediaceae', NA, 'Algae', 'Cyanobacteria', 'phylum', '0', '1', '0')),
               as.list(c('Hapalosiphonaceae', NA, 'Algae', 'Cyanobacteria', 'phylum', '0', '1', '0')),
               as.list(c('Scytonemataceae', NA, 'Algae', 'Cyanobacteria', 'phylum', '1', '1', '1')),

               #### Eukaryota (domain): Chlorophyta (division) ----
               as.list(c('Haematococcaceae', NA, 'Algae', 'Chlorophyta', 'phylum', '1', '1', '0')),
               as.list(c('Ankistrodesmaceae', NA, 'Algae', 'Chlorophyta', 'phylum', '1', '1', '0')),
               as.list(c('Neochloridaceae', NA, 'Algae', 'Chlorophyta', 'phylum', '1', '0', '1')),
               as.list(c('Prasiolaceae', NA, 'Algae', 'Chlorophyta', 'phylum', '0', '1', '0')),
               as.list(c('Hydrodictyaceae', NA, 'Algae', 'Chlorophyta','division', '1', '0', '0')),
               as.list(c('Scenedesmaceae', NA, 'Algae', 'Chlorophyta','division', '1', '0', '0')),
               as.list(c('Oocystaceae', NA, 'Algae', 'Chlorophyta','division', '1', '0', '0')),
               as.list(c('Chlorococcaceae', NA, 'Algae', 'Chlorophyta','division', '1', '1', '0')),
               as.list(c('Dictyosphaeriaceae', NA, 'Algae', 'Chlorophyta','division', '1', '1', '0')),
               as.list(c('Oedogoniaceae', NA, 'Algae', 'Chlorophyta','division', '1', '0', '0')),
               as.list(c('Dunaliellaceae', NA, 'Algae', 'Chlorophyta','division', '0', '1', '0')),#GA:wiki says nothing about habitat
               as.list(c('Chlamydomonadaceae', NA, 'Algae', 'Chlorophyta','division', '1', '1', '0')), ##GA: havent checked were they live yet. info not on wikipedia,
               as.list(c('Polyphysaceae', NA, 'Algae', 'Chlorophyta','division', '1', '1', '0')),
               as.list(c('Selenastraceae', NA, 'Algae', 'Chlorophyta','division', '1', '1', '0')),
               as.list(c('Cladophoraceae', NA, 'Algae', 'Chlorophyta','division', '0', '1', '0')),
               as.list(c('Chlorellaceae', NA, 'Algae', 'Chlorophyta','division', '1', '1', '0')),
               as.list(c('Volvocaceae', NA, 'Algae', 'Chlorophyta','division', '1', '0', '0')),  
               as.list(c('Coccomyxaceae', NA, 'Algae', 'Chlorophyta','division', '1', '1', '1')), # auch symbiotisch mit Algen 
               as.list(c('Ulvaceae', NA, 'Algae', 'Chlorophyta','division', '1', '1', '0')), #GA:wiki sagt nichts ueber habitat
               as.list(c('Micractiniaceae', NA, 'Algae', 'Chlorophyta','division', '1', '0', '0')),
               as.list(c('Dasycladaceae', NA, 'Algae', 'Chlorophyta', 'division', '1', '1', '0')), # schwierig heruazufinden wo vorkam, hauptsaechlich in Fossilien :(
               as.list(c('Radiococcaceae', NA, 'Algae', 'Chlorophyta', 'division', '1', '1', '0')),
               as.list(c('Chlorochytriaceae', NA, 'Algae', 'Chlorophyta', 'division', '1', '1', '0')),
               as.list(c('Halimedaceae', NA, 'Algae', 'Chlorophyta', 'division', '0', '1', '0')),
               as.list(c('Pterospermataceae', NA, 'Algae', 'Chlorophyta', 'division', '0', '1', '0')),
               as.list(c('Ulotrichaceae', NA, 'Algae', 'Chlorophyta', 'division', '0', '1', '0')),
               as.list(c('Chlorodendraceae', NA, 'Algae', 'Chlorophyta', 'division', '0', '1', '0')),
               as.list(c('Chaetophoraceae', NA, 'Algae', 'Chlorophyta', 'division', '0', '1', '0')),

               #### Eukaryota: Plantae: Charophyta: ----
                ## Zygnematophyceae (class)
               as.list(c('Zygnemataceae', NA, 'Algae', 'Zygnematophyceae', 'class', '1', '0', '0')),
               as.list(c('Desmidiaceae', NA, 'Algae', 'Zygnematophyceae', 'class', '1', '1', '0')),
               as.list(c('Closteriaceae', NA, 'Algae', 'Zygnematophyceae', 'class', '1', '1', '0')),

               #### Eukaryota: Hacrobia: Cryptista: Cryptophyta: ----
                ## Cryptophyceae (class)
               as.list(c('Cryptomonadaceae', NA, 'Algae', 'Cryptophyta', 'phylum', '1', '1', '0')),
               as.list(c('Pyrenomonadaceae', NA, 'Algae', 'Cryptophyta', 'phylum', '0', '1', '0')),

               #### Eukaryota: Hacrobia (unranked): Haptophyta (unranked): Prymnesiophyceae (class)
               as.list(c('Isochrysidaceae', NA, 'Algae', 'Prymnesiophyceae', 'class', '1', '1', '0')),
               as.list(c('Syracosphaeraceae', NA, 'Algae', 'Prymnesiophyceae', 'class', '1', '1', '0')),
               as.list(c('Coccolithaceae', NA, 'Algae', 'Prymnesiophyceae', 'class', '1', '1', '0')),

               #### Eukaryota (domain): SAR: Alveolata: Dinoflagellata (phylum) ----
               as.list(c('Gonyaulacaceae', NA, 'Algae', 'Dinoflagellata', 'infraphylum', '0', '1', '0')),
               as.list(c('Oxytoxaceae', NA, 'Algae', 'Dinoflagellata', 'infraphylum', '0', '1', '0')),
               as.list(c('Gymnodiniaceae', NA, 'Algae', 'Dinoflagellata', 'infraphylum', '0', '1', '0')),
               as.list(c('Pyrocystaceae', NA, 'Algae', 'Dinoflagellata', 'infraphylum', '0', '1', '0')),
               as.list(c('Heterodiniaceae', NA, 'Algae', 'Dinoflagellata', 'infraphylum', '0', '1', '0')),
               as.list(c('Peridiniaceae', NA, 'Algae', 'Dinoflagellata', 'infraphylum', '0', '1', '0')),

               #### Eukaryota (domain): Rhodophyta (division) - Red algae ----
                ## Florideophyceae (class)
               as.list(c('Gracilariaceae', NA, 'Algae', 'Rhodophyta', 'phylum', '0', '1', '0')),
               as.list(c('Rhodymeniaceae', NA, 'Algae', 'Rhodophyta', 'phylum', '0', '1', '0')),
                ## Bangiophyceae (class)
               as.list(c('Bangiaceae', NA, 'Algae', 'Rhodophyta', 'phylum', '0', '1', '0')),
                ## Porphyridiophyceae (class)
               as.list(c('Porphyridiaceae', NA, 'Algae', 'Rhodophyta', 'phylum', '0', '1', '0')),

               #### Eukaryota (domain): Chromista (kingdom): Heterokonts (infrakingdom) ----
               # https://en.wikipedia.org/wiki/Heterokont
                 ## Xantophyceae (class)
               as.list(c('Pleurochloridaceae', NA, 'Algae', 'Xantophyceae', 'class', '0', '1', '0')),
                ## Phaeophyceae (class) - Brown algae
               as.list(c('Chordariaceae', NA, 'Algae', 'Phaeophyceae', 'class', '0', '1', '0')),
               as.list(c('Lessoniaceae', NA, 'Algae', 'Phaeophyceae', 'class', '0', '1', '0')),
               as.list(c('Hormosiraceae', NA, 'Algae', 'Phaeophyceae', 'class', '0', '1', '0')),
                ## Bacillariophyceae (class)
               as.list(c('Pinnulariaceae', NA, 'Algae', 'Bacillariophyceae', 'class', '0', '1', '0')),
               as.list(c('Cocconeidaceae', NA, 'Algae', 'Bacillariophyceae', 'class', '0', '1', '0')),
               as.list(c('Entomoneidaceae', NA, 'Algae', 'Bacillariophyceae', 'class', '0', '1', '0')),
               as.list(c('Skeletonemaceae', NA, 'Algae', 'Bacillariophyceae', 'class', '1', '1', '0')),
               as.list(c('Skeletonemataceae', NA, 'Algae', 'Bacillariophyceae', 'class', '1', '1', '0')),
               as.list(c('Thalassiosiraceae', NA, 'Algae', 'Bacillariophyceae', 'class', '1', '1', '0')),
               as.list(c('Gomphonemataceae', NA, 'Algae', 'Bacillariophyceae', 'class', '1', '1', '0')),
               as.list(c('Stauroneidaceae', NA, 'Algae', 'Bacillariophyceae', 'class', '1', '1', '0')),
               as.list(c('Bacillariaceae', NA, 'Algae', 'Bacillariophyceae', 'class', '1', '1', '0')),
               as.list(c('Cymbellaceae', NA, 'Algae', 'Bacillariophyceae', 'class', '1', '1', '0')),
               as.list(c('Achnanthidiaceae', NA, 'Algae', 'Bacillariophyceae', 'class', '1', '1', '0')),
               as.list(c('Phaeodactylaceae', NA, 'Algae', 'Bacillariophyceae', 'class', '0', '1', '0')), 
               as.list(c('Amphipleuraceae', NA, 'Algae', 'Bacillariophyceae', 'class', '0', '1', '0')),
               as.list(c('Sellaphoraceae', NA, 'Algae', 'Bacillariophyceae', 'class', '0', '1', '0')),
               as.list(c('Stephanodiscaceae', NA, 'Algae', 'Bacillariophyceae', 'class', '1', '1', '0')),
               as.list(c('Bacillariophyceae', NA, 'Algae', 'Bacillariophyceae', 'class', '1', '1', '0')), #! wrong! but can't find family for Frustilia
               as.list(c('Tabellariaceae', NA, 'Algae', 'Bacillariophyceae', 'class', '0', '1', '0')),
               as.list(c('Nitzschiaceae', NA, 'Algae', 'Bacillariophyceae', 'class', '1', '1', '0')),
               as.list(c('Achnanthaceae', NA, 'Algae', 'Bacillariophyceae', 'class', '0', '1', '0')),
               as.list(c('Catenulaceae', NA, 'Algae', 'Bacillariophyceae', 'class', '0', '1', '0')),
               as.list(c('Aulacoseiraceae', NA, 'Algae', 'Bacillariophyceae', 'class', '0', '1', '0')),
               as.list(c('Bellerocheaceae', NA, 'Algae', 'Bacillariophyceae', 'class', '0', '1', '0')),
               as.list(c('Cymatosiraceae', NA, 'Algae', 'Bacillariophyceae', 'class', '0', '1', '0')),
               as.list(c('Chaetocerotaceae', NA, 'Algae', 'Bacillariophyceae', 'class', '0', '1', '0')),
               as.list(c('Eunotiaceae', NA, 'Algae', 'Bacillariophyceae', 'class', '1', '0', '0')), # https://biotaxa.org/Phytotaxa/article/view/phytotaxa.79.2.1/0
               as.list(c('Melosiraceae', NA, 'Algae', 'Bacillariophyceae', 'class', '0', '1', '0')),
               as.list(c('Ulnariaceae', NA, 'Algae', 'Bacillariophyceae', 'class', '1', '0', '0')),
                ## Chrysophyceae (class) - Golden algae
               as.list(c('Synuraceae', NA, 'Algae', 'Chrysophyceae', 'class', '1', '0', '0')),
               as.list(c('Chromulinales', NA, 'Algae', 'Chrysophyceae', 'class', '1', '0', '0')),
               as.list(c('Ochromonadaceae', 'Goldalgen', 'Algae', 'Chrysophyceae', 'class', '1','0', '0')),
               as.list(c('Chrysococcaceae', NA, 'Algae', 'Chrysophyceae', 'class', '0', '1', '0')),
                ## Synurophyceae (class)
               as.list(c('Mallomonadaceae', NA, 'Algae', 'Synurophyceae', 'class', '1', '0', '0')),
               as.list(c('Dinobryaceae', NA, 'Algae', 'Synurophyceae', 'class', '1', '0', '0')),
                ## Raphidophyceae
               as.list(c('Chattonellaceae', NA, 'Algae', 'Raphidophyceae', 'class', '0', '1', '0')),
                ## Eustigmatophyceae
               as.list(c('Monodopsidaceae', NA, 'Algae', 'Eustigmatophyceae', 'class', '0', '1', '0')),
                #### Eukaryota (domain): SAR: Rhizaria: Cercozoa (phylum) ----
               as.list(c('Cercomonadidae', NA, 'Algae', 'Cercozoa', 'phylum', '1', '1', '1'))
))
```

# Euglenozoa (phylum)
There are autotrophic, heterotrophic species and parasites.
```{r euglenozoa}
lookup = rbindlist(list(lookup,
               as.list(c('Euglenaceae', NA, 'Algae', 'Euglenozoa', 'phylum', '1', '1', '0')),
               as.list(c('Paranemataceae', NA, 'Algae', 'Euglenozoa', 'phylum', '1', '1', '0'))
))
```

# Ciliophora (phylum) - Wimpertierchen
Most Ciliates are heterotroph.

```{r ciliophora-list}               
lookup = rbindlist(list(lookup,
               #### Eukaryota (domain): SAR: Alveolata: Ciliophora (phylum) ----
               as.list(c('Tetrahymenidae', NA, 'Ciliophora', 'Ciliophora', 'phylum', '1', '0', '0')), # some are parasitic like Ichthyophthirius multifillis (SAR) # feeds on bacteria
               as.list(c('Parameciidae', NA, 'Ciliophora', 'Ciliophora', 'phylum', '1', '1', '1')),
               as.list(c('Euplotidae', NA, 'Ciliophora', 'Ciliophora', 'phylum', '1', '1', '1')),
               as.list(c('Urotrichidae', NA, 'Ciliophora', 'Ciliophora', 'phylum', '1', '1', '0')),
               as.list(c('Oxytrichidae', NA, 'Ciliophora', 'Ciliophora', 'phylum', '1', '1', '1')),
               as.list(c('Spirostomidae', NA, 'Ciliophora', 'Ciliophora', 'phylum', '0', '1', '0')),
               as.list(c('Colpodidae', NA, 'Ciliophora', 'Ciliophora', 'phylum', '1', '1', '1'))
))
```

# Rotifera (phylum) - Raedertierchen
Heterotrophic. Rotifers eat particulate organic detritus. Mostly in freshwater environmemnts.
```{r rotifera-list}
lookup = rbindlist(list(lookup,
               as.list(c('Proalidae', NA, 'Rotifera', 'Rotifera', 'phylum', '0', '1', '0')),
               as.list(c('Brachionidae', NA, 'Rotifera', 'Rotifera', 'phylum', '1', '1', '0')),
               as.list(c('Synchaetidae', NA, 'Rotifera', 'Rotifera', 'phylum', '0', '1', '0')),
               as.list(c('Collothecidae', NA, 'Rotifera', 'Rotifera', 'phylum', '0', '1', '0')),
               as.list(c('Asplanchnidae', NA, 'Rotifera', 'Rotifera', 'phylum', '1', '1', '0')),
               as.list(c('Trochosphaeridae', NA, 'Rotifera', 'Rotifera', 'phylum', '0', '1', '0')),
               as.list(c('Lecanidae', NA, 'Rotifera', 'Rotifera', 'phylum', '1', '1', '0')),
               as.list(c('Trichocercidae', NA, 'Rotifera', 'Rotifera', 'phylum', '0', '1', '0')),
               as.list(c('Philodinidae', NA, 'Rotifera', 'Rotifera', 'phylum', '1', '1', '1')),
               as.list(c('Notommatidae', NA, 'Rotifera', 'Rotifera', 'phylum', '0', '1', '0'))
))
```

# Fungi

```{r fungi-list}
 # fungi (#!all terrestrial? not clear!!)
lookup = rbindlist(list(lookup,
#### Heterokonts (infrakingdom): Colorless groups: Pseudofungi: Oomycetes: ----
#! not actually fungi but due to the lack of pigments also not photosysnthetically active
# https://en.wikipedia.org/wiki/Heterokont
               as.list(c('Saprolegniaceae', NA, 'Fungi', 'Oomycota', 'phylum', '0', '1', '0')),
               as.list(c('Peronosporaceae', NA, 'Fungi', 'Oomycota', 'phylum', '0', '1', '0')),
               as.list(c('Lagenidiaceae', NA, 'Fungi', 'Oomycota', 'phylum', '0', '0', '1')),
               as.list(c('Pythiaceae', NA, 'Fungi', 'Oomycota', 'phylum', '0', '0', '1')),
#### Real Fungi: ----                        
                ## Ascomycota (Schlauchpilze)
               as.list(c('Phlogicylindriaceae', NA, 'Fungi', 'Ascomycota', 'phylum', '0', '1', '0')),
               as.list(c('Pleosporaceae', NA, 'Fungi', 'Ascomycota', 'division', '0', '0', '1')),
               as.list(c('Sclerotiniaceae', NA, 'Fungi', 'Ascomycota', 'division', '0', '0', '1')),
               as.list(c('Trichocomaceae', NA, 'Fungi', 'Ascomycota', 'division', '0', '0', '1')),
               as.list(c('Nectriaceae', NA, 'Fungi', 'Ascomycota', 'division', '0', '0', '1')),
               as.list(c('Glomerellaceae', NA, 'Fungi', 'Ascomycota', 'division', '0', '0', '1')),
               as.list(c('Herpotrichiellaceae', NA, 'Fungi', 'Ascomycota', 'division', '0', '0', '1')),
               as.list(c('Togniniaceae', NA, 'Fungi', 'Ascomycota', 'division', '0', '0', '1')),
               as.list(c('Hypocreaceae', NA, 'Fungi', 'Ascomycota', 'division', '0', '0', '1')),
               as.list(c('Halosphaeriaceae', NA, 'Fungi', 'Ascomycota', 'division', '0', '0', '1')),
               as.list(c('Botryosphaeriaceae', NA, 'Fungi', 'Ascomycota', 'division', '0', '0', '1')),
               as.list(c('Erysiphaceae', NA, 'Fungi', 'Ascomycota', 'division', '0', '0', '1')),
               as.list(c('Mycosphaerellaceae', NA, 'Fungi', 'Ascomycota', 'division', '0', '0', '1')),
               as.list(c('Valsaceae', NA, 'Fungi', 'Ascomycota', 'division', '0', '0', '1')),
               as.list(c('Dothioraceae', NA, 'Fungi', 'Ascomycota', 'division', '0', '0', '1')),
               as.list(c('Aspergillaceae', NA, 'Fungi', 'Ascomycota', 'division', '0', '0', '1')),
               as.list(c('Saccotheciaceae', NA, 'Fungi', 'Ascomycota', 'division', '0', '1', '0')), # listed in WORMS
               as.list(c('Cordycipitaceae', NA, 'Fungi', 'Ascomycota', 'division', '0', '0', '1')),
               as.list(c('Leotiaceae', 'Gallertkaeppchenverwandte', 'Fungi', 'Ascomycota', 'division', '0', '0', '1')),
               as.list(c('Phyllostictaceae', NA, 'Fungi', 'Ascomycota', 'division', '0', '1', '0')),
               as.list(c('Phaeomoniellaceae', NA, 'Fungi', 'Ascomycota', 'division', '0', '0', '1')),
               as.list(c('Parmeliaceae', NA, 'Fungi', 'Ascomycota', 'division', '0', '0', '1')),
               as.list(c('Megalariaceae', NA, 'Fungi', 'Ascomycota', 'division', '0', '0', '1')),
               as.list(c('Acarosporaceae', NA, 'Fungi', 'Ascomycota', 'division', '0', '0', '1')),
               as.list(c('Discinaceae', NA, 'Fungi', 'Ascomycota', 'division', '0', '0', '1')),
               as.list(c('Laboulbeniaceae', NA, 'Fungi', 'Ascomycota', 'division', '0', '0', '1')),
               as.list(c('Dermateaceae', NA, 'Fungi', 'Ascomycota', 'division', '0', '0', '1')),
               as.list(c('Hyponectriaceae', NA, 'Fungi', 'Ascomycota', 'division', '0', '0', '1')),
               as.list(c('Rhizinaceae', NA, 'Fungi', 'Ascomycota', 'division', '0', '1', '0')),

                ## Basidiomycota (Staenderpilze)
               as.list(c('Bankeraceae', NA, 'Fungi', 'Basidiomycota', 'division', '0', '0', '1')),
               as.list(c('Entolomataceae', NA, 'Fungi', 'Basidiomycota', 'division', '0', '0', '1')),
               as.list(c('Pucciniaceae', NA, 'Fungi', 'Basidiomycota', 'division', '0', '0', '1')),
               as.list(c('Corticiaceae', NA, 'Fungi', 'Basidiomycota', 'division', '0', '0', '1')),
               as.list(c('Ceratobasidiaceae', NA, 'Fungi', 'Basidiomycota', 'division', '0', '0', '1')),
               as.list(c('Agaricaceae', 'Champignonverwandten', 'Fungi', 'Basidiomycota', 'division', '0', '0', '1')),
                as.list(c('Quambalariaceae', NA, 'Fungi', 'Basidiomycota', 'division', '0', '0', '1')),
               as.list(c('Atheliaceae', 'Gewebehautverwandte', 'Fungi', 'Basidiomycota', 'division', '0', '1', '1')), #eigentlich terretrisch, aber eine Art bevorzugt salzwasser habitate
               as.list(c('Phanerochaetaceae', NA, 'Fungi', 'Basidiomycota', 'division', '0', '1', '1')),
      	        ## Zygomycota
      	       as.list(c('Neozygitaceae', NA, 'Fungi',  'Zygomycota', 'division', '0', '0', '1')),  
               as.list(c('Ustilaginaceae', 'Brandpilzverwandte', 'Fungi',  'Zygomycota', 'division', '0', '0', '1')),  
               as.list(c('Suillaceae', 'Schmierroehrlingsverwandte', 'Fungi',  'Zygomycota', 'division', '0', '0', '1')),  
               as.list(c('Venturiaceae', NA, 'Fungi',  'Zygomycota', 'division', '0', '0', '1')),  
               as.list(c('Clavicipitaceae', 'Mutterkornpilzverwandte', 'Fungi',  'Zygomycota', 'division', '0', '0', '1')),
      	        ## Glomeromycota
      	       as.list(c('Glomeraceae', NA, 'Fungi',  'Glomeromycota', 'division', '0', '0', '1'))

	       ))
```

# Bryozoa (phylum)
```{r bryozoa-list}
lookup = rbindlist(list(lookup,
                        as.list(c('Electridae', NA, 'Bryozoa', 'Bryozoa', 'phylum', '0', '1', '0'))
))
```

# Amoebozoa (phylum)
```{r amoebozoa-list}
lookup = rbindlist(list(lookup,
                        as.list(c('Tubiferaceae', 'Blutmilchpilz', 'Amoebozoa', 'Amoebozoa', 'phylum', '0', '0', '1'))
))
```

# Nematoda (phylum) - Fadenwuermer
```{r nematoda-list}
lookup = rbindlist(list(lookup,
               as.list(c('Hoplolaimidae', NA, 'Nematoda', 'Secernentea', 'class', '0', '0', '1')),
               as.list(c('Mermithidae', NA, 'Nematoda', 'Adenophorea', 'class', '0', '0', '1')),
               as.list(c('Rhabditidae', NA, 'Nematoda', 'Adenophorea', 'class', '0', '1', '0')),
               as.list(c('Heteroderidae', 'Ruebenzystennematode', 'Nematoda', 'Adenophorea', 'class', '0', '0', '1')),
               as.list(c('Enoplidae', NA, 'Nematoda', 'Enoplea', 'class', '0', '1', '0')),
               as.list(c('Monhysteridae', NA, 'Nematoda', 'Chromadorea', 'class', '0', '1', '0')),
               as.list(c('Meloidogynidae', NA, 'Nematoda', 'Adenophorea', 'class', '0', '0', '1')),
               as.list(c('Panagrolaimidae', NA, 'Nematoda', 'Rhabditida', 'class', '1', '1', '1'))
))
```

# Nemertea (phylum) - Schnurwuermer
```{r nemertea-list}
lookup = rbindlist(list(lookup,
               as.list(c('Tubulanidae', NA, 'Nemertea', 'Anopla', 'class', '0', '1', '0'))
))
```

# Annelida (phylum) - Ringelwuermer
```{r annelida-list}
lookup = rbindlist(list(lookup,
               as.list(c('Arenicolidae', NA, 'Annelida', 'Polychaeta', 'class', '0', '1', '0')),
               as.list(c('Capitellidae', NA, 'Annelida', 'Polychaeta', 'class', '0', '1', '0')),
               as.list(c('Amphinomidae', NA, 'Annelida', 'Polychaeta', 'class', '0', '1', '0')),
               as.list(c('Dorvilleidae', NA, 'Annelida', 'Polychaeta', 'class', '0', '1', '0')),
               as.list(c('Nereididae', NA, 'Annelida', 'Polycheata', 'class', '1', '1', '0')),
               as.list(c('Serpulidae', 'Kalkroehrenwuermer', 'Annelida', 'Polychaeta', 'class', '0', '1', '0')),
               as.list(c('Syllidae', NA, 'Annelida', 'Polycheata', 'class', '0', '1', '0')),
                ## Clitellata (class)
               as.list(c('Naididae', NA, 'Annelida', 'Clitellata', 'class', '1', '1', '0')),
               as.list(c('Lumbricidae', NA, 'Annelida', 'Clitellata', 'class', '1', '1', '0')),
               as.list(c('Enchytraeidae', NA, 'Annelida', 'Clitellata', 'class', '0', '0', '1')),
               as.list(c('Lumbriculidae', NA, 'Annelida', 'Clitellata', 'class', '1', '1', '0')),
                ## Hirudinea (subclass)
               as.list(c('Glossiphoniidae', NA, 'Annelida', 'Clitellata', 'class', '1', '0', '0')),
               as.list(c('Erpobdellidae', NA, 'Annelida', 'Clitellata', 'class', '1', '1', '0')), ##! hard to find info on habitat
               as.list(c('Hirudinidae', NA, 'Annelida', 'Clitellata', 'class', '1', '0', '1')),
               as.list(c('Tubificidae', 'Schlammröhrenwürmer ', 'Annelida', 'Clitellata', 'class', '1', '1', '0')),
               as.list(c('Acoetidae', NA, 'Annelida', 'Clitellata', 'class', '0', '1', '0')) # class Gürtelwürmer (Clitellata)
))
```

# Fish

```{r fish-list}
# fish (#! extra: isBrackish column?)
lookup = rbindlist(list(lookup,
                ## Beryciformes
               as.list(c('Cetomimidae', "Walkoepfe", 'Fish', 'Beryciformes', 'order', '0', '1', '0')),
               ## Tetraodontiformes
               as.list(c('Tetraodontidae', "Kugelfische", 'Fish', 'Tetraodontiformes', 'order', '0', '1', '0')),
                ## Batrachoidiformes
               as.list(c('Batrachoididae', 'Froschfische', 'Fish', 'Batrachoidiformes', 'order', '0', '1', '0')),
               ## Carangiformes
               as.list(c('Carangidae', 'Stachelmakrelen', 'Fish', 'Carangiformes', 'order', '0', '1', '0')),
                ## Syngnathiformes
               as.list(c('Syngnathidae', 'Seenadel', 'Fish', 'Syngnathiformes', 'order', '0', '1', '0')),
                ## Labriformes
               as.list(c('Labridae', 'Lippfische', 'Fish', 'Labriformes', 'order', '0', '1', '0')),
                ## Acipenseriformes
               as.list(c('Acipenseridae', 'Stoere', 'Fish', 'Acipenseriformes', 'order', '1', '1', '0')),
                ## Cypriniformes
               as.list(c('Cyprinidae', 'Karpfenfische', 'Fish', 'Cypriniformes', 'order', '1', '1', '0')),
               as.list(c('Balitoridae', 'Flossensauger', 'Fish', 'Cypriniformes', 'order', '1', '1', '0')),
               as.list(c('Nemacheilidae', 'Bachschmerlen', 'Fish', 'Cypriniformes', 'order', '1', '0', '0')),
                ## Salmoniformes
               as.list(c('Salmonidae', 'Lachsfische', 'Fish', 'Salmoniformes', 'order', '1', '1', '0')),
                ## Centrachiformes
               as.list(c('Centrarchidae', 'Sonnenbarsche', 'Fish', 'Centrarchiformes', 'order', '1', '0', '0')), # Sonnenbarsche
               as.list(c('Kyphosidae', 'Steuerbarsche', 'Fish', 'Centrarchiformes', 'order', '0', '1', '0')),
               as.list(c('Terapontidae', 'Grunzbarsche', 'Fish', 'Centrarchiformes', 'order', '1', '0', '0')),
                ## Cyprinodontiformes
               as.list(c('Cyprinodontidae', 'Zahnkaerpflinge', 'Fish', 'Cyprinodontiformes', 'order', '1', '0', '0')),
               as.list(c('Poeciliidae', 'Zahnkaerpflinge', 'Fish', 'Cyprinodontiformes', 'order', '1', '0', '0')),
               as.list(c('Fundulidae', NA, 'Fish', 'Cyprinodontiformes', 'order', '1', '1', '0')),
               as.list(c('Cobitidae', 'Steinbeiszer', 'Fish', 'Cyprinodontiformes', 'order', '1', '0', '0')),
               as.list(c('Aplocheilidae', NA, 'Fish', 'Cyprinodontiformes', 'order', '1', '0', '0')),
               as.list(c('Anablepidae', NA, 'Fish', 'Cyprinodontiformes', 'order', '1', '0', '0')), # freshwater and brackish
                ## Siluriformes
               as.list(c('Ictaluridae', 'Katzenwelse', 'Fish', 'Siluriformes', 'order', '1', '0', '0')), # endemisch in Nordamerika
               as.list(c('Clariidae', 'Kiemensackwelse', 'Fish', 'Siluriformes', 'order', '1', '0', '0')),
               as.list(c('Bagridae', 'Stachelwelse', 'Fish', 'Siluriformes', 'order', '1', '0', '0')),
               as.list(c('Siluridae', "Echte Welse", 'Fish', 'Siluriformes', 'order', '1', '0', '0')),
	             as.list(c('Catostomidae', 'Saugkarpfen', 'Fish', 'Cyprinodontiformes', 'order', '1', '0', '0')),
               as.list(c('Callichthyidae', 'Panzer- und Schwielenwelse', 'Fish', 'Cyprinodontiformes', 'order', '1', '0', '0')),
               as.list(c('Heptapteridae', NA, 'Fish', 'Cyprinodontiformes', 'order', '1', '0', '0')),
               as.list(c('Loricariidae', 'Harnischwelse', 'Fish', 'Cyprinodontiformes', 'order', '1', '0', '0')),
               as.list(c('Pangasiidae', "Haiwelse", 'Fish', 'Siluriformes', 'order', '1', '1', '0')),
                ## Perciformes
               as.list(c('Nototheniidae', "Antaktisdorsche", 'Fish', 'Perciformes', 'order', '0', '1', '0')),
               as.list(c('Haemulidae', 'Süßlippen und Grunzer', 'Fish', 'Perciformes', 'order', '1', '1', '0')),
               as.list(c('Cichlidae', 'Buntbarsche', 'Fish', 'Perciformes', 'order', '1', '0', '0')), # Mittel u. Suedamerika, Afrika
               as.list(c('Percidae', 'Echte Barsche', 'Fish', 'Perciformes', 'order', '1', '0', '0')),
               as.list(c('Embiotocidae', 'Brandungsbarsche', 'Fish', 'Perciformes', 'order', '0', '1', '0')),
               as.list(c('Gobiidae', 'Grundeln', 'Fish', 'Perciformes', 'order', '1', '1', '0')), # das muesste man ggf. weiter aufteilen
               as.list(c('Sparidae', 'Meerbrassen', 'Fish', 'Perciformes', 'order', '0', '1', '0')), 
               as.list(c('Mugilidae', 'Meeraeschen', 'Fish', 'Perciformes', 'order', '1', '1', '0')),
               as.list(c('Moronidae', 'Wolfsbarsche', 'Fish', 'Perciformes', 'order', '1', '1', '0')),
               as.list(c('Kuhliidae', NA, 'Fish', 'Perciformes', 'order', '1', '1', '0')),
               as.list(c('Gasterosteidae', 'Stichlinge', 'Fish', 'Perciformes', 'order', '1', '1', '0')),
               as.list(c('Sciaenidae', 'Umberfische', 'Fish', 'Perciformes', 'order', '1', '1', '0')),
               as.list(c('Callionymidae', 'Umberfische', 'Fish', 'Perciformes', 'order', '0', '1', '0')),
               as.list(c('Zoarcidae', 'Aalmuttern', 'Fish', 'Perciformes', 'order', '0', '1', '0')),
               as.list(c('Agonidae', 'Panzergroppen', 'Fish', 'Perciformes', 'order', '0', '1', '0')),
               as.list(c('Centropomidae', 'Panzergroppen', 'Fish', 'Perciformes', 'order', '0', '1', '0')),
               as.list(c('Scorpaenidae', "Skorpionfische", 'Fish', 'Perciformes', 'order', '0', '1', '0')),

                ## Anabantiformes
               as.list(c('Channidae', 'Schlangenkopffische', 'Fish', 'Anabantiformes', 'order', '1', '0', '0')), # Afrika, Asien
               as.list(c('Heteropneustidae', 'Kiemenschlauchwelse', 'Fish', 'Anabantiformes', 'order', '1', '0', '0')), # Asien
               as.list(c('Anabantidae', 'Kletterfische und Buschfische', 'Fish', 'Anabantiformes', 'order', '1', '1', '0')),# primarily in fresh water but also in brackish water
               as.list(c('Osphronemidae', NA, 'Fish', 'Anabantiformes', 'order', '1', '0', '0')),
                ## Atheriniformes
               as.list(c('Atherinidae', 'Altweltlichen aehrenfische', 'Fish', 'Atheriniformes', 'order', '1', '1', '0')),
               as.list(c('Atherinopsidae', 'Neuweltliche aehrenfische', 'Fish', 'Atheriniformes', 'order', '1', '1', '0')),
               as.list(c('Melanotaeniidae', 'Regenbogenfische', 'Fish', 'Atheriniformes', 'order', '1', '0', '0')),
                ## Anguilliformes
               as.list(c('Anguillidae', 'Aale', 'Fish', 'Atheriniformes', 'order', '1', '1', '0')),
               as.list(c('Muraenidae', 'Muränen', 'Fish', 'Atheriniformes', 'order', '0', '1', '0')),
                ## Beloniformes
               as.list(c('Adrianichthyidae', 'Reisfische', 'Fish', 'Beloniformes', 'order', '1', '0', '0')), # auch Brackwasser (Indien, Indonesien, Japan)
	              ## Osteoglossiformes
               as.list(c('Notopteridae', 'Altwelt-Messerfische', 'Fish', 'Osteoglossiformes', 'order', '1', '0', '0')),
               as.list(c('Mormyridae', 'Nilhechte', 'Fish', 'Osteoglossiformes', 'order', '1', '0', '0')),

                ## Pleuronectiformes
               as.list(c('Pleuronectidae', 'Schollen', 'Fish', 'Pleuronectiformes', 'order', '1', '1', '0')),
               as.list(c('Soleidae', 'Seezungen', 'Fish', 'Pleuronectiformes', 'order', '1', '1', '0')),
               as.list(c('Scophthalmidae', 'Steinbutte', 'Fish', 'Pleuronectiformes', 'order', '0', '1', '0')),
               as.list(c('Paralichthyidae', 'Steinbutte', 'Fish', 'Pleuronectiformes', 'order', '0', '1', '0')),

                ## Osmeriformes
               as.list(c('Osmeridae', 'Stinte', 'Fish', 'Osmeriformes', 'order', '0', '1', '0')),
               as.list(c('Plecoglossidae', NA, 'Fish', 'Osmeriformes', 'order', '1', '1', '0')),

                ## Esociformes
               as.list(c('Esocidae', 'Hechtartige', 'Fish', 'Esociformes', 'order', '0', '1', '0')),
               as.list(c('Umbridae', 'Hundsfische', 'Fish', 'Esociformes', 'order', '1', '0', '0')),
                ## Galaxiiformes
               as.list(c('Galaxiidae', 'Galaxien', 'Fish', 'Galaxiiformes', 'order', '1', '1', '0')),
                ## Characiformes
               as.list(c('Curimatidae', 'Breitlingssalmler', 'Fish', 'Characiformes', 'order', '1', '0', '0')),
               as.list(c('Lebiasinidae', 'Schlanksalmler', 'Fish', 'Characiformes', 'order', '1', '0', '0')),
               as.list(c('Characidae', 'Echte Salmler', 'Fish', 'Characiformes', 'order', '1', '0', '0')),
                ## Petromyzontiformes
               as.list(c('Petromyzontidae', NA, 'Fish', 'Petromyzontiformes', 'order', '1', '1', '0')),
                ## Clupeiformes
               as.list(c('Clupeidae', NA, 'Fish', 'Clupeiformes', 'order', '1', '1', '0'))

))
```

# Amphibia
```{r amphibia-list}
lookup = rbindlist(list(lookup,
               as.list(c('Pelobatidae', NA, 'Amphibia', 'Urodela', 'order', '0', '0', '1')),
               as.list(c('Plethodontidae', 'Lungenlose Salamander', 'Amphibia', 'Urodela', 'order', '0', '0', '1')),
               as.list(c('Alytidae', NA, 'Amphibia', 'Anura', 'order', '1', '0', '1')),
               as.list(c('Hylidae', 'Laubfroesche', 'Amphibia', 'Anura', 'order', '1', '0', '1')),
               as.list(c('Limnodynastidae', 'Australische Suedfroesche', 'Amphibia', 'Anura', 'order', '1', '0', '1')),
               as.list(c('Myobatrachidae', 'Australische Suedfroesche', 'Amphibia', 'Anura', 'order', '1', '0', '1')),
               as.list(c('Ambystomatidae', NA, 'Amphibia', 'Anura', 'order', '1', '0', '1')),
               as.list(c('Bufonidae', 'Kroeten', 'Amphibia', 'Anura', 'order', '0', '0', '1')),
               as.list(c('Bombinatoridae', 'Unken und Barbourfroesche', 'Amphibia', 'Anura', 'order', '1', '0', '1')),
               as.list(c('Microhylidae', 'Engmaulfroesche', 'Amphibia', 'Anura', 'order', '1', '0', '1')),
               as.list(c('Dicroglossidae', NA, 'Amphibia', 'Anura', 'order', '1', '0', '1')),
               as.list(c('Pipidae', 'Zungenlose', 'Amphibia', 'Anura', 'order', '1', '0', '1')), # only adults are terrestrial
               as.list(c('Leiuperidae', NA, 'Amphibia', 'Anura', 'order', '1', '0', '1')),
               as.list(c('Leptodactylidae', 'Pfeiffrösche', 'Amphibia', 'Anura', 'order', '1', '0', '1')),
               as.list(c('Scaphiopodidae', "Amerikanische Schaufelfußkröte", 'Amphibia', 'Anura', 'order', '1', '0', '1')),
               as.list(c('Rhacophoridae', "Ruderfrösche", 'Amphibia', 'Anura', 'order', '1', '0', '1')),
               as.list(c('Ranidae', "Echte Frösche", 'Amphibia', 'Anura', 'order', '1', '0', '1')),

                ## Caudata (order)
               as.list(c('Salamandridae', 'Echte Samalander', 'Amphibia', 'Caudata', 'order', '1', '0', '1'))
               ))
```

# Reptilia
```{r reptilia-list}
lookup = rbindlist(list(lookup,
               as.list(c('Eublepharidae', NA, 'Reptilia', 'Squamata', 'order', '1', '0', '1'))
))
```

# Crustacea (subphylum)

```{r Crustacea-list}
lookup = rbindlist(list(lookup,
               as.list(c('Triopsidae', 'Rückenschaler', 'Crustacea', 'Notostraca', 'order', '1', '0', '0')),
               as.list(c('Thamnocephalidae', NA, 'Crustacea', 'Anostraca', 'order', '1', '0', '0')),
               as.list(c('Streptocephalidae', NA, 'Crustacea', 'Anostraca', 'order', '1', '0', '0')),
               # Crustacea (subphylum)
                ## Branchiopoda (class)
                  ### Cladocera (order)
               as.list(c('Daphniidae', 'Wasserfloehe', 'Crustacea', 'Cladocera', 'order', '1', '0', '0')),
               as.list(c('Artemiidae', 'Salzwasserkrebse', 'Crustacea', 'Cladocera', 'order', '1', '0', '0')), # Binnensalzgewaesser, nicht im Meer!
	             as.list(c('Macrothricidae', NA, 'Crustacea', 'Cladocera', 'order', '1', '0', '0')),
	             as.list(c('Moinidae', NA, 'Crustacea', 'Cladocera', 'order', '1', '1', '0')), #GA: fresh and brackish waters
               as.list(c('Chydoridae', NA, 'Crustacea', 'Cladocera', 'order', '1', '1', '0')),
               as.list(c('Bosminidae', NA, 'Crustacea', 'Cladocera', 'order', '1', '1', '0')),
               as.list(c('Sididae', NA, 'Crustacea', 'Cladocera', 'order', '1', '1', '0')),
               as.list(c('Ilyocryptidae', NA, 'Crustacea', 'Cladocera', 'order', '1', '0', '0')),
                ## Malacostraca (class)
                   ### Amphipoda (order)
               as.list(c('Niphargidae', NA, 'Crustacea', 'Amphipoda', 'order', '1', '0', '0')),
               as.list(c('Neoniphargidae', NA, 'Crustacea', 'Amphipoda', 'order', '1', '0', '0')),
               as.list(c('Gammaridae', 'Bachflohkrebse', 'Crustacea', 'Amphipoda', 'order', '1', '0', '0')),
               as.list(c('Hyalellidae', 'Bachflohkrebse', 'Crustacea', 'Amphipoda', 'order', '1', '0', '0')), # used to belong to Gammaridae
	             as.list(c('Parathelphusidae', NA, 'Crustacea', 'Amphipoda', 'order', '1', '0', '0')), # ASia
               as.list(c('Phoxocephalidae', NA, 'Crustacea', 'Amphipoda', 'order', '0', '1', '0')),
               as.list(c('Melitidae', NA, 'Crustacea', 'Amphipoda', 'order', '1', '1', '0')),
               as.list(c('Ampeliscidae', NA, 'Crustacea', 'Amphipoda', 'order', '0', '1', '0')),
               as.list(c('Pontoporeiidae', NA, 'Crustacea', 'Amphipoda', 'order', '1', '1', '0')),
               as.list(c('Haustoriidae', NA, 'Crustacea', 'Amphipoda', 'order', '0', '1', '0')),
               as.list(c('Talitridae', NA, 'Crustacea', 'Amphipoda', 'order', '1', '1', '1')),
               as.list(c('Paramelitidae', NA, 'Crustacea', 'Amphipoda', 'order', '1', '0', '0')),

                  ### Decapoda
               as.list(c('Palaemonidae', 'Felsen- und Partnergarnelen', 'Crustacea', 'Decapoda', 'order', '1', '1', '0')),
               as.list(c('Cancridae', 'Taschenkrebse', 'Crustacea', 'Decapoda', 'order', '0', '1', '0')),
               as.list(c('Portunidae', 'Schwimmkrabben', 'Crustacea', 'Decapoda', 'order', '1', '1', '0')),
               as.list(c('Crangonidae', NA, 'Crustacea', 'Decapoda', 'order', '0', '1', '0')),
               as.list(c('Cambaridae', NA, 'Crustacea', 'Decapoda', 'order', '1', '0', '0')), # ueberfam.: Flusskrebse
               as.list(c('Sesarmidae', NA, 'Crustacea', 'Decapoda', 'order', '0', '1', '0')),
               as.list(c('Grapsidae', 'Quadratkrabben', 'Crustacea', 'Decapoda', 'order', '1', '1', '1')), # hauptsaechlich in den Tropen
   	           as.list(c('Nephropidae', 'Hummerartige', 'Crustacea', 'Decapoda', 'order', '0', '1', '0')),
               as.list(c('Pandalidae', 'Tiefseegarnelen', 'Crustacea', 'Decapoda', 'order', '0', '1', '0')),
               as.list(c('Paguridae', 'Rechtshaendige Einsiedlerkrebse', 'Crustacea', 'Decapoda', 'order', '0', '1', '0')),
	             as.list(c('Ocypodidae', NA, 'Crustacea', 'Decapoda', 'order', '0', '1', '1')),
	             as.list(c('Xanthidae', NA, 'Crustacea', 'Decapoda', 'order', '0', '1', '0')), #lebensraum in wikipedia nicht beschrieben
	             as.list(c('Atyidae', 'Suesswassergarnelen', 'Crustacea', 'Decapoda', 'order', '1', '1', '0')), #GA:keine richtigen marinen arten, aber manche leben in brackwasser
               as.list(c('Alpheidae', 'Knallkrebse', 'Crustacea', 'Decapoda', 'order', '0', '1', '0')),
               as.list(c('Gecarcinucidae', NA, 'Crustacea', 'Decapoda', 'order', '1', '0', '0')),
               as.list(c('Parastacidae', NA, 'Crustacea', 'Decapoda', 'order', '1', '0', '0')),
               as.list(c('Astacidae', NA, 'Crustacea', 'Decapoda', 'order', '1', '0', '0')),
               as.list(c('Panopeidae', NA, 'Crustacea', 'Decapoda', 'order', '0', '1', '0')),
               as.list(c('Trichodactylidae', NA, 'Crustacea', 'Decapoda', 'order', '1', '0', '0')),
               as.list(c('Hippolytidae', 'Putzer- und Marmorgarnelen', 'Crustacea', 'Decapoda', 'order', '0', '1', '0')),
               as.list(c('Macrophthalmidae', NA, 'Crustacea', 'Decapoda', 'order', '0', '1', '0')),
               as.list(c('Majidae', 'Dreieckskrabben', 'Crustacea', 'Decapoda', 'order', '0', '1', '0')),
               as.list(c('Menippidae', NA, 'Crustacea', 'Decapoda', 'order', '0', '1', '0')),
               as.list(c('Diogenidae', NA, 'Crustacea', 'Decapoda', 'order', '0', '1', '0')),
               as.list(c('Lithodidae', NA, 'Crustacea', 'Decapoda', 'order', '0', '1', '0')),
               as.list(c('Luciferidae', NA, 'Crustacea', 'Decapoda', 'order', '0', '1', '0')),

                  ### Mysida (order)
               as.list(c('Mysidae', 'Wasserfloehe', 'Crustacea', 'Mysida', 'order', '1', '1', '0')), # Porter et al. 2008
               as.list(c('Varunidae', NA, 'Crustacea', 'Mysida', 'order', '0', '1', '0')),
                  ### Isopoda (order)
               as.list(c('Penaeidae', NA, 'Crustacea', 'Isopoda', 'order', '0', '1', '0')), # ev. genauere Differnzierung ueber Genera
               as.list(c('Asellidae', NA, 'Crustacea', 'Isopoda', 'order', '1', '0', '1')),
               as.list(c('Armadillidiidae', NA, 'Crustacea', 'Isopoda', 'order', '0', '0', '1')),
	             as.list(c('Porcellionidae', NA, 'Crustacea', 'Isopoda', 'order', '0', '0', '1')), # lebensraum in wikipedia nicht beschrieben
	             as.list(c('Oniscidae', NA, 'Crustacea', 'Isopoda', 'order', '0', '0', '1')),
               as.list(c('Phreatoicidae', NA, 'Crustacea', 'Isopoda', 'order', '1', '0', '0')),
               as.list(c('Bopyridae', NA, 'Crustacea', 'Isopoda', 'order', '1', '1', '0')),
               as.list(c('Sphaeromatidae', NA, 'Crustacea', 'Isopoda', 'order', '1', '1', '0')),

                ## Ostracoda (class)
               as.list(c('Cyprididae', NA, 'Crustacea', 'Ostracoda', 'class', '1', '0', '0')), # the most diverse group of freshwater ostracods
               as.list(c('Ilyocyprididae', NA, 'Crustacea', 'Ostracoda', 'class', '1', '0', '0')),
               as.list(c('Notodromadidae', NA, 'Crustacea', 'Ostracoda', 'class', '0', '1', '0')),

                ## Copepoda (subclass)
               as.list(c('Temoridae', NA, 'Crustacea', 'Copepoda', 'subclass', '1', '1', '0')),
	             as.list(c('Ameiridae', NA, 'Crustacea', 'Copepoda', 'subclass', '1', '1', '0')), 
	             as.list(c('Harpacticidae', NA, 'Crustacea', 'Copepoda', 'subclass', '1', '1', '0')),
               as.list(c('Acartiidae', NA, 'Crustacea', 'Copepoda', 'subclass', '0', '1', '0')),
               as.list(c('Diosaccidae', NA, 'Crustacea', 'Copepoda', 'subclass', '0', '1', '0')),
               as.list(c('Centropagidae', NA, 'Crustacea', 'Copepoda', 'subclass', '1', '1', '0')),
               as.list(c('Caligidae', NA, 'Crustacea', 'Copepoda', 'subclass', '0', '1', '0')),
	             as.list(c('Canthocamptidae', NA, 'Crustacea', 'Copepoda', 'subclass', '1', '1', '0')), #GA: only few marine species
               as.list(c('Cyclopidae', NA, 'Crustacea', 'Copepoda', 'subclass', '1', '1', '0')),
               as.list(c('Tisbidae', NA, 'Crustacea', 'Copepoda', 'subclass', '1', '1', '0')),
               as.list(c('Diaptomidae', NA, 'Crustacea', 'Copepoda', 'subclass', '1', '0', '0')),
               as.list(c('Eucalanidae', NA, 'Crustacea', 'Copepoda', 'subclass', '0', '1', '0')),
               as.list(c('Clausocalanidae', NA, 'Crustacea', 'Copepoda', 'subclass', '0', '1', '0')),
               as.list(c('Laophontidae', NA, 'Crustacea', 'Copepoda', 'subclass', '0', '1', '0')),
               as.list(c('Pseudodiaptomidae', NA, 'Crustacea', 'Calanoida', 'order', '1', '1', '0')),
               as.list(c('Oithonidae', NA, 'Crustacea', 'Copepoda', 'subclass', '1', '1', '0')),

                ## Thecostraca (subclass)
               as.list(c('Balanidae', 'Seepocken', 'Crustacea', 'Thecostraca', 'subclass', '0', '1', '0'))
	       ))
```

# Diplopoda (class)

```{r diplopoda-list}
lookup = rbindlist(list(lookup,
               as.list(c('Julidae', NA, 'Diplopoda', 'Julida', 'order', '0', '1', '1'))
))
```

# Insecta (class)

```{r insecta-list}
lookup = rbindlist(list(lookup,
                ## Dermaptera (order)
               as.list(c('Forficulidae', NA, 'Insecta', 'Dermaptera', 'order', '0', '0', '1')),
                ## Megaloptera (order)
               as.list(c('Corydalidae', NA, 'Insecta', 'Megaloptera', 'order', '1', '0', '1')),
               as.list(c('Sialidae', "Schlammfliegen", 'Insecta', 'Megaloptera', 'order', '1', '0', '1')),
                ## Diptera (order)
               as.list(c('Mycetophilidae', 'Pilzmücken', 'Insecta', 'Diptera', 'order', '0', '0', '1')),
               as.list(c('Culicidae', 'Stechmuecken', 'Insecta', 'Diptera', 'order', '1', '0', '1')),
               as.list(c('Muscidae', 'Echte Fliegen', 'Insecta', 'Diptera', 'order', '0', '0', '1')),
               as.list(c('Chironomidae', 'Zuckmuecken', 'Insecta', 'Diptera', 'order', '1', '1', '0')),
               as.list(c('Agromyzidae', 'Minierfliegen', 'Insecta', 'Diptera', 'order', '0', '0', '1')),
               as.list(c('Simuliidae', 'Kriebelmuecken', 'Insecta', 'Diptera', 'order', '1', '0', '1')),
               as.list(c('Cecidomyiidae', 'Gallmuecken', 'Insecta', 'Diptera', 'order', '0', '0', '1')),
               as.list(c('Anthomyiidae', 'Blumenfliegen', 'Insecta', 'Diptera', 'order', '0', '0', '1')),
               as.list(c('Syrphidae', 'Schwebfliegen', 'Insecta', 'Diptera', 'order', '0', '0', '1')),
               as.list(c('Drosophilidae', 'Taufliegen', 'Insecta', 'Diptera', 'order', '0', '0', '1')),
               as.list(c('Ceratopogonidae', 'Gnitzen', 'Insecta', 'Diptera', 'order', '1', '0', '1')),
               as.list(c('Chaoboridae', 'Bueschelmuecken', 'Insecta', 'Diptera', 'order', '1', '0', '1')),
               as.list(c('Tephritidae', 'Bohrfliegen', 'Insecta', 'Diptera', 'order', '0', '0', '1')),
               as.list(c('Chloropidae', 'Halmfliegen', 'Insecta', 'Diptera', 'order', '0', '0', '1')),
               as.list(c('Limoniidae', 'Stelzmuecken', 'Insecta', 'Diptera', 'order', '0', '0', '1')),
               as.list(c('Athericidae', 'Ibisfliegen', 'Insecta', 'Diptera', 'order', '0', '0', '1')),
               as.list(c('Tipulidae', 'Ibisfliegen', 'Insecta', 'Diptera', 'order', '0', '0', '1')),
               as.list(c('Calliphoridae', 'Schmeißfliegen', 'Insecta', 'Diptera', 'order', '0', '0', '1')),
               as.list(c('Sciaridae', "Trauermücke", 'Insecta', 'Diptera', 'order', '0', '0', '1')),

                ## Lepidoptera
               as.list(c('Noctuidae', 'Eulenfalter', 'Insecta', 'Lepidoptera', 'order', '0', '0', '1')),
               as.list(c('Tortricidae', 'Wickler', 'Insecta', 'Lepidoptera', 'order', '0', '0', '1')),
               as.list(c('Pyralidae', 'Zuensler', 'Insecta', 'Lepidoptera', 'order', '0', '0', '1')),
               as.list(c('Crambidae', 'Eulenfalter', 'Insecta', 'Lepidoptera', 'order', '0', '0', '1')),
               as.list(c('Aphididae', 'Roehrenblattlaeuse', 'Insecta', 'Lepidoptera', 'order', '0', '0', '1')),
               as.list(c('Plutellidae', 'Schleier- und Halbmotten', 'Insecta', 'Lepidoptera', 'order', '0', '0', '1')),
               as.list(c('Sphingidae', 'Schwaermer', 'Insecta', 'Lepidoptera', 'order', '0', '0', '1')),
               as.list(c('Gelechiidae', 'Palpenmotten', 'Insecta', 'Lepidoptera', 'order', '0', '0', '1')),
               as.list(c('Papilionidae', 'Ritterfalter', 'Insecta', 'Lepidoptera', 'order', '0', '0', '1')),
               as.list(c('Lycaenidae', 'Blaeulinge', 'Insecta', 'Lepidoptera', 'order', '0', '0', '1')),
               as.list(c('Nepticulidae', 'Zwergminiermotten', 'Insecta', 'Lepidoptera', 'order', '0', '0', '1')),
               as.list(c('Pieridae', 'Weißlinge', 'Insecta', 'Lepidoptera', 'order', '0', '0', '1')),
               as.list(c('Blastobasidae', NA, 'Insecta', 'Lepidoptera', 'order', '0', '0', '1')),
               as.list(c('Arctiidae', 'Baerenspinner', 'Insecta', 'Lepidoptera', 'order', '0', '0', '1')),
               as.list(c('Psychidae', 'Echte Sacktraeger', 'Insecta', 'Lepidoptera', 'order', '0', '0', '1')),
               as.list(c('Gracillariidae', 'Miniermotten', 'Insecta', 'Lepidoptera', 'order', '0', '0', '1')),
               as.list(c('Lasiocampidae', 'Wollraupenspinner', 'Insecta', 'Lepidoptera', 'order', '0', '0', '1')),
               as.list(c('Geometridae', 'Spanner', 'Insecta', 'Lepidoptera', 'order', '0', '0', '1')),
               as.list(c('Erebidae', NA, 'Insecta', 'Lepidoptera', 'order', '0', '0', '1')),
               as.list(c('Nolidae', 'Kahneulchen', 'Insecta', 'Lepidoptera', 'order', '0', '0', '1')),
               as.list(c('Bombycidae', 'Echte Spinner', 'Insecta', 'Lepidoptera', 'order', '0', '0', '1')),
               as.list(c('Saturniidae', 'Pfauenspinner', 'Insecta', 'Lepidoptera', 'order', '0', '0', '1')),
                ## Hemiptera
               as.list(c('Nepidae', "Skorpionswanzen", 'Insecta', 'Hemiptera', 'order', '1', '0', '0')),
               as.list(c('Aleyrodidae', NA, 'Insecta', 'Hemiptera', 'order', '0', '0', '1')),
               as.list(c('Delphacidae', NA, 'Insecta', 'Hemiptera', 'order', '0', '0', '1')),
               as.list(c('Diaspididae', 'Deckelschildlaeuse', 'Insecta', 'Hemiptera', 'order', '0', '0', '1')),
               as.list(c('Pentatomidae', 'Baumwanzen', 'Insecta', 'Hemiptera', 'order', '0', '0', '1')), # superfamily: Pentatomoidea
               as.list(c('Lygaeidae', 'Bodenwanzen', 'Insecta', 'Hemiptera', 'order', '0', '0', '1')), #! were changed to Geocoridae?
               as.list(c('Geocoridae', 'Bodenwanzen', 'Insecta', 'Hemiptera', 'order', '0', '0', '1')),
               as.list(c('Cicadellidae', 'Zwergzikaden', 'Insecta', 'Hemiptera', 'order', '0', '0', '1')),
               as.list(c('Anthocoridae', 'Blumenwanzen', 'Insecta', 'Hemiptera', 'order', '0', '0', '1')),
               as.list(c('Miridae', 'Weichwanzen', 'Insecta', 'Hemiptera', 'order', '0', '0', '1')),
               as.list(c('Coccidae', 'Napfschilzwanze', 'Insecta', 'Hemiptera', 'order', '0', '0', '1')),
               as.list(c('Cercopidae', 'Blutzikade', 'Insecta', 'Hemiptera', 'order', '0', '0', '1')),
               as.list(c('Pseudococcidae', 'Schmierlaeuse', 'Insecta', 'Hemiptera', 'order', '0', '0', '1')),
               as.list(c('Psyllidae', NA, 'Insecta', 'Hemiptera', 'order', '0', '0', '1')),
               as.list(c('Aphrophoridae', 'Schaumzikaden', 'Insecta', 'Hemiptera', 'order', '0', '0', '1')),
               as.list(c('Aphalaridae', NA, 'Insecta', 'Hemiptera', 'order', '0', '0', '1')),
               as.list(c('Belostomatidae', 'Riesenwanzen', 'Insecta', 'Hemiptera', 'order', '0', '0', '1')),
               as.list(c('Gerridae', 'Wasserlaeufer', 'Insecta', 'Hemiptera', 'order', '1', '0', '0')),
               as.list(c('Cicadidae', 'Singzikade', 'Insecta', 'Hemiptera', 'order', '0', '0', '1')),
               as.list(c('Nabidae', 'Sichelwanzen', 'Insecta', 'Hemiptera', 'order', '0', '0', '1')),
               as.list(c('Pleidae', "Zwergrückenschwimmer", 'Insecta', 'Hemiptera', 'order', '1', '0', '0')),

                #### Nepomorpha - Wasserwanzen
               as.list(c('Notonectidae', 'Rueckenschwimmer', 'Insecta', 'Hemiptera', 'order', '1', '0', '1')), 
               as.list(c('Corixidae', 'Ruderwanzen', 'Insecta', 'Hemiptera', 'order', '1', '0', '1')),
               as.list(c('Micronectidae', 'Singzikade', 'Insecta', 'Hemiptera', 'order', '1', '0', '0')),

                ## Hymenoptera
               as.list(c('Eumenidae', 'Solitäre Faltenwespen', 'Insecta', 'Hymenoptera', 'order', '0', '0', '1')),
               as.list(c('Apidae', 'Bienen', 'Insecta', 'Hymenoptera', 'order', '0', '0', '1')),
               as.list(c('Trichogrammatidae', NA, 'Insecta', 'Hymenoptera', 'order', '0', '0', '1')),
               as.list(c('Braconidae', 'Brackwespen', 'Insecta', 'Hymenoptera', 'order', '0', '0', '1')),
               as.list(c('Curculionidae', 'Ruesselkaefer', 'Insecta', 'Hymenoptera', 'order', '0', '0', '1')),
               as.list(c('Eulophidae', 'Erzwespen', 'Insecta', 'Hymenoptera', 'order', '0', '0', '1')),
               as.list(c('Ichneumonidae', 'Schlupfwespen', 'Insecta', 'Hymenoptera', 'order', '0', '0', '1')),
               as.list(c('Scelionidae', NA, 'Insecta', 'Hymenoptera', 'order', '1', '0', '1')), # few attack aqu Insecta under water
               as.list(c('Aphelinidae', NA, 'Insecta', 'Hymenoptera', 'order', '0', '0', '1')),
               as.list(c('Vespidae', 'Faltenwespen', 'Insecta', 'Hymenoptera', 'order', '0', '0', '1')),
               as.list(c('Formicidae', 'Ameisen', 'Insecta', 'Hymenoptera', 'order', '0', '0', '1')),
               as.list(c('Encyrtidae', NA, 'Insecta', 'Hymenoptera', 'order', '0', '0', '1')), # parasites
               as.list(c('Tenthredinidae', 'Echte Blattwespen', 'Insecta', 'Hymenoptera', 'order', '0', '0', '1')),
               as.list(c('Eupelmidae', NA, 'Insecta', 'Hymenoptera', 'order', '0', '0', '1')), # parasisch
               as.list(c('Megachilidae', NA, 'Insecta', 'Hymenoptera', 'order', '0', '0', '1')), 
               as.list(c('Cephidae', 'Halmwespen', 'Insecta', 'Hymenoptera', 'order', '0', '0', '1')),
               as.list(c('Cynipidae', 'Gallwespen', 'Insecta', 'Hymenoptera', 'order', '0', '0', '1')),
               as.list(c('Mymaridae', 'Zwergwespen', 'Insecta', 'Hymenoptera', 'order', '1', '0', '1')),
               as.list(c('Figitidae', NA, 'Insecta', 'Hymenoptera', 'order', '0', '0', '1')),
               as.list(c('Dryinidae', NA, 'Insecta', 'Hymenoptera', 'order', '0', '0', '1')),
               as.list(c('Pteromalidae', NA, 'Insecta', 'Hymenoptera', 'order', '0', '0', '1')),

                ## Coleoptera (Kaefer)
               as.list(c('Cetoniidae', 'Rosenkaefer', 'Insecta', 'Coleoptera', 'order', '0', '0', '1')),
               as.list(c('Coccinellidae', 'Marienkaefer', 'Insecta', 'Coleoptera', 'order', '0', '0', '1')),
               as.list(c('Chrysomelidae', 'Blattkaefer', 'Insecta', 'Coleoptera', 'order', '0', '0', '1')),
               as.list(c('Scarabaeidae', 'Blattkaefer', 'Insecta', 'Coleoptera', 'order', '0', '0', '1')),
               as.list(c('Tenebrionidae', 'Schwarzkaefer', 'Insecta', 'Coleoptera', 'order', '0', '0', '1')),
               as.list(c('Carabidae', 'Laufkaefer', 'Insecta', 'Coleoptera', 'order', '0', '0', '1')),
               as.list(c('Nitidulidae', 'Glanzkaefer', 'Insecta', 'Coleoptera', 'order', '0', '0', '1')),
               # Amphicrossus japonicus (einziger Glanzkaefer der adult ins Wasser geht)
               as.list(c('Bostrichidae', 'Bohrkaefer', 'Insecta', 'Coleoptera', 'order', '0', '0', '1')),
               as.list(c('Elateridae', 'Schnellkaefer', 'Insecta', 'Coleoptera', 'order', '0', '0', '1')),
               as.list(c('Trogidae', 'Erdkaefer', 'Insecta', 'Coleoptera', 'order', '0', '0', '1')),
               as.list(c('Buprestidae', 'Prachtkaefer', 'Insecta', 'Coleoptera', 'order', '0', '0', '1')),
               as.list(c('Erirhinidae', NA, 'Insecta', 'Coleoptera', 'order', '0', '0', '1')), # ist eine Familie der Ruesselkaefer
               as.list(c('Dytiscidae', 'Schwimmkaefer', 'Insecta', 'Coleoptera', 'order', '1', '0', '0')),
               as.list(c('Cantharidae', 'Weichkaefer', 'Insecta', 'Coleoptera', 'order', '0', '0', '1')), 
               as.list(c('Cerambycidae', 'Bockkaefer', 'Insecta', 'Coleoptera', 'order', '0', '0', '1')), 
               as.list(c('Staphylinidae', 'Kurzfluegler', 'Insecta', 'Coleoptera', 'order', '0', '0', '1')), 
               as.list(c('Anthicidae', 'Bluetenmulmkaefer', 'Insecta', 'Coleoptera', 'order', '0', '0', '1')),
               as.list(c('Hydrophilidae', 'Wasserkaefer', 'Insecta', 'Coleoptera', 'order', '1', '0', '1')),
               as.list(c('Laemophloeidae', NA, 'Insecta', 'Coleoptera', 'order', '0', '0', '1')),
               as.list(c('Histeridae', 'Stutzkaefer', 'Insecta', 'Coleoptera', 'order', '0', '0', '1')),
               as.list(c('Anobiidae', 'Nagekäfer', 'Insecta', 'Coleoptera', 'order', '0', '0', '1')),
               as.list(c('Cryptophagidae', 'Schimmelkäfer', 'Insecta', 'Coleoptera', 'order', '0', '0', '1')),
               as.list(c('Silvanidae', NA, 'Insecta', 'Coleoptera', 'order', '0', '0', '1')),

                #### Coleoptera in aquatic habitats
               as.list(c('Gyrinidae', 'Taumelkäfer', 'Insecta', 'Coleoptera', 'order', '1', '0', '0')), # water beetles
               as.list(c('Brentidae', 'Taumelkäfer', 'Insecta', 'Coleoptera', 'order', '1', '0', '0')),
               as.list(c('Haliplidae', 'Wassertreter', 'Insecta', 'Coleoptera', 'order', '1', '0', '1')),

                ## Neuroptera (Netzfluegler)
               as.list(c('Sisyridae', NA, 'Insecta', 'Neuroptera', 'order', '1', '0', '0')),
               as.list(c('Nevrorthidae', NA, 'Insecta', 'Neuroptera', 'order', '1', '0', '0')),
               as.list(c('Osmylidae', NA, 'Insecta', 'Neuroptera', 'order', '1', '0', '0')),
               as.list(c('Chrysopidae', 'Florfliegen', 'Insecta', 'Neuroptera', 'order', '0', '0', '1')),

                ## Othoptera
               as.list(c('Gryllidae', NA, 'Insecta', 'Othoptera', 'order', '0', '0', '1')),
               as.list(c('Acrididae', NA, 'Insecta', 'Othoptera', 'order', '0', '0', '1')),
                ## Thripidae
               as.list(c('Thripidae', NA, 'Insecta', 'Thysanoptera', 'order', '0', '0', '1')), # probably paraphyletic group
                ## Blattodea (Schaben)
               as.list(c('Blattellidae', NA, 'Insecta', 'Blattodea', 'order', '0', '0', '1')),
               as.list(c('Ectobiidae', NA, 'Insecta', 'Blattodea', 'order', '0', '0', '1')), # new name of Blattellidae
               as.list(c('Termitidae', NA, 'Insecta', 'Blattodea', 'order', '0', '0', '1')), # Termiten  (infraorder, Isoptera)
               as.list(c('Blattidae', 'Schaben', 'Insecta', 'Blattodea', 'order', '0', '0', '1')),
                ## Ephemeroptera (Eintagsfliegen) larval aquatisch, adult terrestrisch
               as.list(c('Baetidae', NA, 'Insecta', 'Ephemeroptera', 'order', '1', '0', '1')),
               as.list(c('Heptageniidae', NA, 'Insecta', 'Ephemeroptera', 'order', '1', '0', '1')),
               as.list(c('Ephemerellidae', NA, 'Insecta', 'Ephemeroptera', 'order', '1', '0', '1')),
               as.list(c('Ephemeridae', NA, 'Insecta', 'Ephemeroptera', 'order', '1', '0', '1')),
               as.list(c('Siphlonuridae', NA, 'Insecta', 'Ephemeroptera', 'order', '1', '0', '1')),
               as.list(c('Caenidae', NA, 'Insecta', 'Ephemeroptera', 'order', '1', '0', '1')),
               as.list(c('Leptophlebiidae', NA, 'Insecta', 'Ephemeroptera', 'order', '1', '0', '1')),
               as.list(c('Polymitarcyidae', NA, 'Insecta', 'Ephemeroptera', 'order', '1', '0', '1')),
               as.list(c('Isonychiidae', NA, 'Insecta', 'Ephemeroptera', 'order', '1', '0', '1')),
               as.list(c('Leptohyphidae', NA, 'Insecta', 'Ephemeroptera', 'order', '1', '0', '1')),
               as.list(c('Ameletidae', NA, 'Insecta', 'Ephemeroptera', 'order', '1', '0', '1')),
                ## Trichoptera (Koecherfliegen) i.d.R. larval aquatisch (in den Tropen ein paar Ausnahmen), Adult terrestisch
               as.list(c('Hydropsychidae', NA, 'Insecta', 'Trichoptera', 'order', '1', '0', '1')),
               as.list(c('Limnephilidae', NA, 'Insecta', 'Trichoptera', 'order', '1', '0', '1')),
               as.list(c('Brachycentridae', NA, 'Insecta', 'Trichoptera', 'order', '1', '0', '1')),
               as.list(c('Philopotamidae', NA, 'Insecta', 'Trichoptera', 'order', '1', '0', '1')),
               as.list(c('Polycentropodidae', NA, 'Insecta', 'Trichoptera', 'order', '1', '0', '1')),
               as.list(c('Helicopsychidae', NA, 'Insecta', 'Trichoptera', 'order', '1', '0', '1')),
               as.list(c('Goeridae', NA, 'Insecta', 'Trichoptera', 'order', '1', '0', '1')),
               as.list(c('Lepidostomatidae', NA, 'Insecta', 'Trichoptera', 'order', '1', '0', '1')),
               as.list(c('Leptoceridae', NA, 'Insecta', 'Trichoptera', 'order', '1', '0', '1')),
               as.list(c('Molannidae', NA, 'Insecta', 'Trichoptera', 'order', '1', '0', '1')),
               as.list(c('Odontoceridae', NA, 'Insecta', 'Trichoptera', 'order', '1', '0', '1')),
                ## Plecoptera (Steinfliegen) - larval aquatisch, adult terrestrisch
               as.list(c('Pteronarcyidae', NA, 'Insecta', 'Plecoptera', 'order', '1', '0', '1')),
               as.list(c('Perlidae', NA, 'Insecta', 'Plecoptera', 'order', '1', '0', '1')),
               as.list(c('Perlodidae', NA, 'Insecta', 'Plecoptera', 'order', '1', '0', '1')),
               as.list(c('Capniidae', NA, 'Insecta', 'Plecoptera', 'order', '1', '0', '1')),
               as.list(c('Peltoperlidae', NA, 'Insecta', 'Plecoptera', 'order', '1', '0', '1')),
               as.list(c('Leuctridae', NA, 'Insecta', 'Plecoptera', 'order', '1', '0', '1')),
               as.list(c('Nemouridae', NA, 'Insecta', 'Plecoptera', 'order', '1', '0', '1')),
               as.list(c('Taeniopterygidae', NA, 'Insecta', 'Plectoptera', 'order', '1', '0', '1')),

		            ## Siphonaptera
               as.list(c('Pulicidae', NA, 'Insecta', 'Siphonaptera', 'order', '0', '0', '1')),# parasites
            		## Odonata
            	 as.list(c('Libellulidae', 'Segellibellen', 'Insecta', 'Odonata', 'order', '1', '0', '1')),
               as.list(c('Aeshnidae', 'Edellibellen', 'Insecta', 'Odonata', 'order', '1', '0', '1')),
               as.list(c('Lestidae', 'Teichjungfern', 'Insecta', 'Odonata', 'order', '1', '0', '1')),
               as.list(c('Coenagrionidae', 'Schlanklibellen', 'Insecta', 'Odonata', 'order', '1', '0', '1')),
               as.list(c('Corduliidae', 'Falkenlibellen', 'Insecta', 'Odonata', 'order', '1', '0', '1')),
               as.list(c('Gomphidae', 'Flussjungfern', 'Insecta', 'Odonata', 'order', '1', '0', '1'))
               ))

```

```{r entognatha-list}
# class: Entognatha, sunclass: Collembola
lookup = rbindlist(list(lookup,
               as.list(c('Isotomidae', NA, 'Entognatha', 'Collembola', 'subclass', '0', '0', '1')),
               as.list(c('Onychiuridae', NA, 'Entognatha', 'Collembola', 'subclass', '0', '0', '1')),
               as.list(c('Sminthuridae', NA, 'Entognatha', 'Collembola', 'subclass', '0', '0', '1')),
               as.list(c('Entomobryidae', NA, 'Entognatha', 'Collembola', 'subclass', '0', '0', '1')),
               as.list(c('Poduromorpha', NA, 'Entognatha', 'Collembola', 'subclass', '0', '0', '1')),
               as.list(c('Tullbergiidae', NA, 'Entognatha', 'Collembola', 'subclass', '0', '0', '1')),
               as.list(c('Dermestidae', 'Speckkäfer', 'Entognatha', 'Collembola', 'subclass', '0', '0', '1')),
                ## im Wasser lebend:
               as.list(c('Poduridae', 'Wasserspringer', 'Entognatha', 'Collembola', 'subclass', '1', '0', '0'))

))
```

# Arachnida (class)

```{r arachnida-list}
lookup = rbindlist(list(lookup,
                ## Acari (Milben) (subclass)
               as.list(c('Suctobelbidae', NA, 'Arachnida', 'Acari', 'subclass', '0', '0', '1')),
               as.list(c('Polyaspididae', NA, 'Arachnida', 'Acari', 'subclass', '0', '0', '1')),
               as.list(c('Trachytidae', NA, 'Arachnida', 'Acari', 'subclass', '0', '0', '1')),
               as.list(c('Perlohmanniidae', NA, 'Arachnida', 'Acari', 'subclass', '0', '0', '1')),
               as.list(c('Tenuipalpidae', NA, 'Arachnida', 'Acari', 'subclass', '0', '0', '1')),
               as.list(c('Phytoseiidae', NA, 'Arachnida', 'Acari', 'subclass', '0', '0', '1')),
               as.list(c('Ascidae', NA, 'Arachnida', 'Acari', 'subclass', '0', '0', '1')),
               as.list(c('Varroidae', 'Schnabelmilben', 'Arachnida', 'Acari', 'subclass', '0', '0', '1')),
               as.list(c('Hemisarcoptidae', NA, 'Arachnida', 'Acari', 'subclass', '0', '0', '1')),
               as.list(c('Tetranychidae', NA, 'Arachnida', 'Acari', 'subclass', '0', '0', '1')),
               as.list(c('Bdellidae', 'Schnabelmilben', 'Arachnida', 'Acari', 'subclass', '0', '0', '1')),
               as.list(c('Eriophyidae', 'Gallmilben', 'Arachnida', 'Acari', 'subclass', '0', '0', '1')),   
	             as.list(c('Penthaleidae', NA, 'Arachnida', 'Acari', 'subclass', '0', '0', '1')),
               as.list(c('Pionidae', NA, 'Arachnida', 'Acari', 'subclass', '0', '0', '1')),
               as.list(c('Anystidae', NA, 'Arachnida', 'Acari', 'subclass', '0', '0', '1')),
                    #### Hydrachnidiae (Unterkohorte) - Süßwassermilben
               as.list(c('Arrenuridae', NA, 'Arachnida', 'Acari', 'subclass', '1', '0', '0')),
                  ### Sarcoptiformes (order)
	             as.list(c('Acaridae', NA, 'Arachnida', 'Acari', 'subclass', '0', '0', '1')),
               ## Ixodida (=Metastigmata) (Zecken)
               as.list(c('Argasidae', 'Lederzecken', 'Arachnida', 'Ixodida', 'order', '0', '0', '1')),

                ## Aranae (order) - Echte Spinnen
               as.list(c('Lycosidae', 'Wolfspinnen', 'Arachnida', 'Araneae', 'order', '0', '0', '1')),
      	       as.list(c('Linyphiidae', 'Baldachinspinnen', 'Arachnida', 'Araneae', 'order', '0', '0', '1')),
      	       as.list(c('Philodromidae', 'Laufspinnen', 'Arachnida', 'Araneae', 'order', '0', '0', '1')),
               as.list(c('Eresidae', 'Röhrenspinnen', 'Arachnida', 'Araneae', 'order', '0', '0', '1')),
               as.list(c('Tetragnathidae', 'Dickkieferspinnen', 'Arachnida', 'Araneae', 'order', '0', '0', '1'))
               ))
```

# Mollusca

```{r mollusca-list}
lookup = rbindlist(list(lookup,
                ## Bivalvia (Muscheln)
               as.list(c('Ostreidae', 'Austern', 'Mollusca', 'bivalvia', 'class', '0', '1', '0')),
               as.list(c('Veneridae', 'Venusmuscheln', 'Mollusca', 'bivalvia', 'class', '0', '1', '0')),
               as.list(c('Unionidae', 'Fluss- und Teichmuscheln', 'Mollusca', 'bivalvia', 'class', '0', '1', '0')),
               as.list(c('Mytilidae', 'Miesmuscheln', 'Mollusca', 'bivalvia', 'class', '0', '1', '0')),
               as.list(c('Dreissenidae', 'Dreikantmuscheln', 'Mollusca', 'bivalvia', 'class', '1', '1', '0')),
               as.list(c('Pectinidae', 'Kammmuscheln', 'Mollusca', 'bivalvia', 'class', '0', '1', '0')),
               as.list(c('Corbiculidae', 'Koerbchenmuscheln', 'Mollusca', 'bivalvia', 'class', '1', '1', '0')),
               as.list(c('Astartidae', 'Astarten', 'Mollusca', 'bivalvia', 'class', '1', '1', '0')),
               as.list(c('Arcidae', 'Archenmuscheln', 'Mollusca', 'bivalvia', 'class', '0', '1', '0')),
               as.list(c('Cardiidae', 'Herzmuscheln', 'Mollusca', 'bivalvia', 'class', '0', '1', '0')),
               as.list(c('Donacidae', 'Koffermuscheln', 'Mollusca', 'bivalvia', 'class', '0', '1', '0')),
               as.list(c('Mactridae', 'Trogmuscheln', 'Mollusca', 'bivalvia', 'class', '0', '1', '0')),
               as.list(c('Myidae', 'Klaffmuscheln', 'Mollusca', 'bivalvia', 'class', '0', '1', '0')),
               as.list(c('Pisidiidae', 'Erbsenmuscheln', 'Mollusca', 'bivalvia', 'class', '1', '0', '0')),
               ## Polyplacophora (class) - 
               as.list(c('Leptochitonidae', NA, 'Mollusca', 'Polyplacophora', 'class', '0', '1', '0')),
		            ## Gastropoda (Schnecken)
               as.list(c('Naticidae', NA, 'Mollusca', 'gastropoda', 'class', '0', '1', '0')),
               as.list(c('Bulinidae', NA, 'Mollusca', 'gastropoda', 'class', '1', '0', '0')),
               as.list(c('Pleuroceridae', NA, 'Mollusca', 'gastropoda', 'class', '1', '0', '0')),
               as.list(c('Hydrobiidae', 'Wasserdeckelschnecken', 'Mollusca', 'gastropoda', 'class', '1', '1', '0')),
               as.list(c('Haliotidae', 'Seeohren', 'Mollusca', 'gastropoda', 'class', '0', '1', '0')),
               as.list(c('Ellobiidae', NA, 'Mollusca', 'gastropoda', 'class', '0', '0', '1')),
               as.list(c('Acteonidae', NA, 'Mollusca', 'gastropoda', 'class', '0', '1', '0')),
               as.list(c('Semisulcospiridae', NA, 'Mollusca', 'Gastropoda', 'class', '1', '0', '0')),
               as.list(c('Melanopsidae', NA, 'Mollusca', 'Gastropoda', 'class', '1', '0', '0')),

                  ### Pulmonata (Lungenschnecken) (order)
                    #### Basommatophora (Wasserlungenschnecken) (suborder)
               as.list(c('Planorbidae', 'Tellerschnecken', 'Mollusca', 'gastropoda', 'class', '1', '0', '0')),
               as.list(c('Physidae', 'Blasenschnecken', 'Mollusca', 'gastropoda', 'class', '1', '0', '0')),
               as.list(c('Lymnaeidae', 'Schlammschnecken', 'Mollusca', 'gastropoda', 'class', '1', '0', '0')),
                   ### Panpulmonata (order)
               as.list(c('Pyramidellidae', NA, 'Mollusca', 'gastropoda', 'class', '0', '1', '0')),
                    #### Stylommatophora - Landlungenschnecken (suborder)
               as.list(c('Limacidae', 'Schnegel', 'Mollusca', 'gastropoda', 'class', '0', '0', '1')),
	           	 as.list(c('Achatinidae', 'Afrikanischen Riesenschnecken', 'Mollusca', 'gastropoda', 'class', '0', '0', '1')),
               as.list(c('Vertiginidae', 'Windelschnecken', 'Mollusca', 'gastropoda', 'class', '0', '0', '1')),
                  ### Sorbeoconcha - Sauggehaeuseschnecken
               as.list(c('Thiaridae', 'Kronenschnecke', 'Mollusca', 'gastropoda', 'class', '1', '0', '0')),
               as.list(c('Ampullariidae', 'Apfelschnecken', 'Mollusca', 'gastropoda', 'class', '1', '0', '0')),
	             as.list(c('Bullinidae', NA, 'Mollusca', 'gastropoda', 'class', '0', '1', '0')),
	             as.list(c('Nassariidae', NA, 'Mollusca', 'gastropoda', 'class', '0', '1', '0')),
		            ### Architaenioglossa (order)
	           	 as.list(c('Viviparidae', 'Sumpfdeckelschnecken', 'Mollusca', 'gastropoda', 'class', '1', '0', '0')),
                ## Cephalopoda (class)
               as.list(c('Onychoteuthidae', NA, 'Mollusca', 'Cephalopoda', 'class', '0', '1', '0'))
		))
```

# Aves (class) - Voegel
```{r aves-list}
lookup = rbindlist(list(lookup,
               as.list(c('Anatidae', 'Entenvoegel', 'Aves', 'Anseriformes', 'order', '1', '0', '1')),
               as.list(c('Odontophoridae', 'Zahnwachteln', 'Aves', 'Galliformes', 'order', '1', '0', '1')),
               as.list(c('Phasianidae', 'Fasanenartige', 'Aves', 'Galliformes', 'order', '1', '0', '1')),
               as.list(c('Emberizidae', 'Ammern', 'Aves', 'Passeriformes', 'order', '0', '0', '1')),
               as.list(c('Icteridae', 'Stärlinge', 'Aves', 'Passeriformes', 'order', '0', '0', '1')),
               as.list(c('Rallidae', "Rallen", 'Aves', 'Gruiformes', 'order', '0', '0', '1')),
               as.list(c('Fringillidae', "Finken", 'Aves', 'Gruiformes', 'order', '0', '0', '1'))
               ))
```

# Tunicata - Manteltiere (subphylum)
```{r tunicata-list}
lookup = rbindlist(list(lookup,
               as.list(c('Cionidae', NA, 'Tunicata', 'Tunicata', 'subphylum', '0', '1', '0')),
               as.list(c('Doliolidae', NA, 'Tunicata', 'Tunicata', 'subphylum', '0', '1', '0'))
               ))
```

# Echinodermata - Stachehaeuter (phylum)

```{r echinodermata-list}
lookup = rbindlist(list(lookup,
                ## Crinoidea (class) - Seelilien / Haarsterne
               as.list(c('Antedonidae', NA, 'Echinodermata', 'Crinoidea', 'class', '0', '1', '0')),
                ## Ophiuroidea (class) - Schlangensterne
               as.list(c('Amphiuridae', NA, 'Echinodermata', 'Ophiuroidea', 'class', '0', '1', '0')),
                ## Echinoidea (class) - Seeigel
               as.list(c('Echinidae', NA, 'Echinodermata', 'Echinoidea', 'class', '0', '1', '0')),
               as.list(c('Arbaciidae', NA, 'Echinodermata', 'Echinoidea', 'class', '0', '1', '0')),
               as.list(c('Dendrasteridae', NA, 'Echinodermata', 'Echinoidea', 'class', '0', '1', '0')),
               as.list(c('Echinometridae', NA, 'Echinodermata', 'Echinoidea', 'class', '0', '1', '0')),
               as.list(c('Strongylocentrotidae', NA, 'Echinodermata', 'Echinoidea', 'phylum', '0', '1', '0')),
               as.list(c('Temnopleuridae', NA, 'Echinodermata', 'Echinoidea', 'class', '0', '1', '0'))
               ))
```

# Platyhelminthes (phylum) - Plattwuermer

```{r  platyhelminthes-list}
lookup = rbindlist(list(lookup,
              as.list(c('Apocreadiidae', NA, 'Platyhelminthes', 'Platyhelminthes', 'phylum', '1', '1', '0')),
              as.list(c('Dugesiidae', NA, 'Platyhelminthes', 'Platyhelminthes', 'phylum', '1', '0', '0')),
              as.list(c('Schistosomatidae', NA, 'Platyhelminthes', 'Platyhelminthes', 'order', '1', '1', '1')), # parasite
              as.list(c('Planariidae', NA, 'Platyhelminthes', 'Platyhelminthes', 'phylum', '1', '1', '0'))
               ))
```

# Gastrotricha (phylum) - Bauchhaerlinge

```{r gastrotricha-list}
lookup = rbindlist(list(lookup,
               as.list(c('Chaetonotidae', NA , 'Gastrotricha', 'Gastrotricha', 'phylum', '1', '1', '0'))
))
```

# Chaetognatha (phylum)
```{r chaetognatha-list}
lookup = rbindlist(list(lookup,
               as.list(c('Sagittidae', NA , 'Chaetognatha', 'Chaetognatha', 'phylum', '1', '1', '0'))
))
```

# Cnidaria - Nesseltiere (phylum)
Heterotrophi. Mostly in marine envrionments. Rarely by use of endosymbionts.
```{r cnidaria-list}
lookup = rbindlist(list(lookup,
               as.list(c('Antipathidae', NA, 'Cnidaria', 'Hydrozoa', 'class', '0', '1', '0')),
               as.list(c('Hydridae', 'Suesswasserpolypen', 'Cnidaria', 'Hydrozoa', 'class', '1', '0', '0')),
               as.list(c('Dendrophylliidae', NA, 'Cnidaria', 'Hydrozoa', 'class', '0', '1', '0')),
               as.list(c('Acroporidae', NA, 'Cnidaria', 'Hydrozoa', 'class', '0', '1', '0')),
               as.list(c('Aiptasiidae', NA, 'Cnidaria', 'Hydrozoa', 'class', '0', '1', '0')),
               as.list(c('Campanulinidae', NA, 'Cnidaria', 'Hydrozoa', 'class', '0', '1', '0'))
))
```

# Porifera (phylum) - Schwaemme
```{r porifera-list}
lookup = rbindlist(list(lookup,
               as.list(c('Halichondriidae', NA, 'Porifera', 'Porifera', 'phylum', '0', '1', '0')),
               as.list(c('Raspailiidae', NA, 'Porifera', 'Porifera', 'phylum', '0', '1', '0'))
))
```

# Viruses
```{r virus-list}
lookup = rbindlist(list(lookup,
               as.list(c('Polydnaviridae', NA, 'Viruses', 'Viruses', 'viruses', '0', '0', '1')),
               as.list(c('Baculoviridae', NA, 'Viruses', 'Viruses', 'viruses', '0', '0', '1')),
               as.list(c('Alphaflexiviridae', NA, 'Viruses', 'Viruses', 'viruses', '0', '0', '1'))
))
```

# Mammalia (class)
```{r mammalia-list}
lookup = rbindlist(list(lookup,
               as.list(c('Caviidae', 'Meerschweinchen', 'Mammalia', 'Rodentia', 'order', '0', '0', '1')),
               as.list(c('Muridae', 'Langschwanzmäuse', 'Mammalia', 'Rodentia', 'order', '0', '0', '1'))
))
```

# Checks
```{r checks}
if (length(which(isTRUE(duplicated(lookup$family)))) < 1) {
  message('No duplicates.')
} else {
  message('Duplicates!')
}
```

# Todo
```{r todo}
todo_habitat =
  unique(tests2[ !family %in% lookup$family, ]$latin_BIname)
if (length(todo_habitat) == 0) {
  rm(tests2, todo_habitat)
  message('All entries classified.')
} else {
  message('The following taxa are still to be classified:\n', paste0(todo_habitat, '\n'))
}
```

# Macrophytes
Maybe create an own list as they often don't form own families.

```{r lookup-macrophytes, eval=FALSE}
# necessary 'cause in Poaceae and Brassicaceae are some Macrophytes
# genus level
#! not finished yet!
lookup_macrophyte = data.table(genus = 'Limnosipanea', german_name = NA, family = 'Rubiaceae', epa_supgroup = 'Plants', epa_taxon = 'Angiospermae', epa_tx_rank = 'clade', epa_isFre = '1', isMacPhy = '1', epa_isMar = '0', epa_isTer = '0')

lookup_macrophyte = rbindlist(list(lookup_macrophyte,

               # plants
               as.list(c('Hippuris', NA, 'Plantaginaceae', 'Plants', 'Angiospermae', 'clade', '1', '1', '0', '0')),
               as.list(c('Callitriche', NA, 'Plantaginaceae', 'Plants', 'Angiospermae', 'clade', '1', '1', '0', '0')),
               as.list(c('Anubias', 'Speerblaetter', 'Araceae', 'Plants', 'Angiospermae', 'clade', '1', '1', '0', '0')),
               as.list(c('Cryptocoryne', 'Wasserkelche', 'Araceae', 'Plants', 'Angiospermae', 'clade', '1', '1', '0', '0')),
               as.list(c('Orontium', NA, 'Araceae', 'Plants', 'Angiospermae', 'clade', '1', '1', '0', '0')),
               as.list(c('Pistia', 'Wassersalat', 'Araceae', 'Plants', 'Angiospermae', 'clade', '1', '1', '0', '0')),
               as.list(c('Wolffia', NA, 'Araceae', 'Plants', 'Angiospermae', 'clade', '1', '1', '0', '0')),
               as.list(c('Spirodela', NA, 'Araceae', 'Plants', 'Angiospermae', 'clade', '1', '1', '0', '0')),
               as.list(c('Lemna', NA, 'Araceae', 'Plants', 'Angiospermae', 'clade', '1', '1', '0', '0')),
               as.list(c('Trapa', NA, 'Lythraceae', 'Plants', 'Angiospermae', 'clade', '1', '1', '0', '0'))
               as.list(c('Nelumbonaceae', 'Lotosblumen', 'Plants', 'Angiospermen', 'clade', '1', '0', '1', '1')),
               as.list(c('Azollaceae', 'Schwimmfarngewaechse', 'Plants', 'Polypodiopsida', 'class', '1', '1', '0', '0')),
               as.list(c('Characeae', NA, 'Plants', 'Charophyta', 'division', '1', '1', '0', '0'))
               )),
               as.list(c('Nelumbonaceae', 'Lotosblumen', 'Plants', 'Angiospermen', 'clade', '1', '0', '1', '1')),
'Azolla pinnata'

as.list(c('Potamogetonaceae', "Laichkrautgewächse", 'Plants', 'Spermatophyta', 'subdivision', '1', '1', '0', '0'))
))
```

# Write to file
```{r write-to-file-rm-overhead}
fwrite(lookup, file.path(tempdir(), 'family_lookup_list.csv'))
```

### 2018-07-09 R/speciesinfo.Rmd
renamed folder <src> to <R>

---
title: "speciesinfo"
author: "AS"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Setup
```{r setup}
## packages
require(data.table)
require(taxize)

online = F

cachedir = '/home/andreas/Documents/Projects/species-sensitivity/cache'
```

# TODO
```{r TODO}
tests_todo = readRDS(file.path(cachedir, 'tests1.rds'))
todo_taxa = sort(unique(tests_todo$latin_BIname))
rm(tests_todo)
```

# QUERIES

## ITIS, NCBI, COL

As the EPA ECOTOX organism groups (algae, plants, crustaceans, fish, insects, invertebrates and molluscs) are not concise we establish new groups querying taxa information using taxize.

```{r QUERY-ITIS-COL-NBN-taxa-groups, eval=TRUE, echo=TRUE}
time_tot = Sys.time()
#### Query ITIS ----
if (online) {
  itis_todo = todo_taxa
  # itis_todo = itis_todo[1:5] # debug me!

  time = Sys.time()
  result_itis = classification(itis_todo, db = 'itis', rows = 1)
  message('Query took: ', format(Sys.time() - time, format = '%M')) # 8 min

  saveRDS(result_itis, file.path(cachedir, 'ITIS_taxa_result.rds'))

} else {
  result_itis = readRDS(file.path(cachedir, 'ITIS_taxa_result.rds'))
}

# add itis identifier (to distinguish later between ITIS and NCBI)
result_itis[!is.na(result_itis)] = mapply(cbind, result_itis[!is.na(result_itis)], source = 'itis', SIMPLIFY = FALSE) # don't spoil the NAs

# Entries that are not found in ITIS
todo2_1 = sort(names(result_itis[ is.na(result_itis) ]))

# Matches that don't have a family entry
todo2_2 = names(which(unlist(lapply(
  lapply(result_itis[ !is.na(result_itis) ],
         function(x) x[ x$rank == 'family', ]$name),
  length)) == 0))

# Only keep valid results
result_itis = result_itis[ !is.na(result_itis) & !names(result_itis) %in% todo2_2 ]

# todo NCBI
todo2 = c(todo2_1, todo2_2)

#### Query COL (Catalogue of Life) - seems to be the fastest ----
if (online) {
  time = Sys.time()
  result_col = classification(todo2, db = 'col', rows = 1)  
  message('Query took: ', format(Sys.time() - time, format = '%M')) # 6 sec

  saveRDS(result_col, file.path(cachedir, 'COL_taxa_result_NOT_IN_ITIS.rds'))

} else {
  result_col = readRDS(file.path(cachedir, 'COL_taxa_result_NOT_IN_ITIS.rds'))
}

# add identifier col
result_col[!is.na(result_col)] = lapply(result_col[!is.na(result_col)], cbind, source = 'col')

# Entries that are not found
todo3_1 = sort(names(result_col[ is.na(result_col) ]))

# Matchin entries that don't have a family entry
todo3_2 = names(which(unlist(lapply(
  lapply(result_col[ !is.na(result_col) ],
         function(x) x[ x$rank == 'family', ]$name),
  length)) == 0))

todo3 = c(todo3_1, todo3_2)

# Only keep valid results
result_col = result_col[ !is.na(result_col) & !names(result_col) %in% todo3_2 ]

#### Query NBN (National Biodiversity Network UK) ----
if (online) {
  time = Sys.time()
  result_nbn = classification(todo3, db = 'nbn', rows = 1)
  message('Query took: ', format(Sys.time() - time, format = '%M'))

  saveRDS(result_nbn, file.path(cachedir, 'NBN_taxa_result_NOT_IN_ITIS_COL.rds'))

} else {
  result_nbn = readRDS(file.path(cachedir, 'NBN_taxa_result_NOT_IN_ITIS_COL.rds'))
}

# add ifentifier nbn
result_nbn[ !is.na(result_nbn) ] = lapply(result_nbn[ !is.na(result_nbn) ], cbind, source = 'nbn')

# Entries that are not found
todo4_1 = sort(names(result_nbn[ is.na(result_nbn) ]))

# Matchin entries that don't have a family entry
todo4_2 = names(which(unlist(lapply(
  lapply(result_nbn[ !is.na(result_nbn) ],
         function(x) x[ x$rank == 'family', ]$name),
  length)) == 0))

todo4 = c(todo4_1, todo4_2)

# Only keep valid results
result_nbn = result_nbn[ !is.na(result_nbn) & !names(result_nbn) %in% todo4_2 ]

#### TOL (Open Tree of Life) ----
if (online) {
  time = Sys.time()
  result_tol = classification(todo4, db = 'tol', rows = 1)
  message('Query took: ', format(Sys.time() - time, format = '%M')) # 2.8 min

  saveRDS(result_tol, file.path(cachedir, 'TOL_taxa_result_NOT_IN_ITIS_COL_NBN.rds'))

} else {
  result_tol = readRDS(file.path(cachedir, 'TOL_taxa_result_NOT_IN_ITIS_COL_NBN.rds'))
}

# add identifier: ncbi
result_tol[!is.na(result_tol)] = mapply(cbind, result_tol[!is.na(result_tol)], source = 'tol', SIMPLIFY = FALSE)

# Entries that are not found
todo5_1 = sort(names(result_tol[ is.na(result_tol) ]))

# Matches that don't have a family entry
todo5_2 = names(which(unlist(lapply(
  lapply(result_tol[ !is.na(result_tol) ],
         function(x) x[ x$rank == 'family', ]$name),
  length)) == 0))

todo5 = c(todo5_1, todo5_2)

# Only keep valid results
result_tol = result_tol[ !is.na(result_tol) & !names(result_tol) %in% todo5 ]

message('The queries took: ', format(Sys.time() - time_tot, format = '%M'))
```

## not found & errata
```{r QUERY-not-found}
# sort(todo5)
# Taxa that couldn't be identified with the above methods
##DEPRECATED!!!!
# leftovers = list(`Agamermis unka` = data.table(name = 'Mermithidae',
#                                                         rank = 'family',
#                                                         id = NA,
#                                                         source = 'by_hand'),
#                  Neoniphargus = data.table(name = 'Neoniphargidae',
#                                                      rank = 'family',
#                                                      id = NA,
#                                                      source = 'by_hand'),
#                  Newnhamia = data.table(name = 'Notodromadidae',
#                                         rank = 'family',
#                                         id = NA,
#                                         source = 'by_hand'),
#                  `Parachlorella kessleri` = data.table(name = 'Chlorellaceae',
#                                            rank = 'family',
#                                            id = NA,
#                                            source = 'by_hand'),
#                  `Paramelita nigroculus` = data.table(name = 'Paramelitidae',
#                                                         rank = 'family',
#                                                         id = NA,
#                                                         source = 'by_hand'),
#                  Sphaerodema = data.table(name = 'Belostomatidae',
#                                           rank = 'family',
#                                           id = NA,
#                                           source = 'by_hand'),
leftovers = list(`Supputius cincticeps` = data.table(name = 'Pentatomoidea',
                                           rank = 'family',
                                           id = NA,
                                           source = 'by_hand'),
                 `Tetrahymena thermophila` = data.table(name = 'Tetrahymenidae',
                                                        rank = 'family',
                                                        id = NA,
                                                        source = 'by_hand'))
```

# Bind
```{r bind}
result_taxize = c(result_itis,
                  result_col,
                  result_nbn,
                  result_tol,
                  leftovers)

## Retreive name, family, source
# NOT very fast, NOT very R, but not as clumsy as lapply ;)
tx_list = list()
for (i in 1:length(result_taxize)) {
  dt = as.data.table(result_taxize[[i]], stringsAsFactors = FALSE)
  dt[ , source := as.character(source) ]
  latin_BIname = names(result_taxize[i])

  if (nrow(dt[ rank == 'family' ]) == 0) {
    tx = data.table(latin_BIname = latin_BIname,
                    family = 'NOT',
                    source_tax = unlist(dt[ nrow(dt), 'source' ]),
                    stringsAsFactors = FALSE)
  } else {
    tx = data.table(latin_BIname = latin_BIname,
                    family = unlist(dt[ rank == 'family', 'name' ]),
                    source_tax = unlist(dt[ nrow(dt), 'source' ]),
                    stringsAsFactors = FALSE)
  }
  tx_list[[i]] = tx
}

tx_dt = rbindlist(tx_list)
setnames(tx_dt, c("latin_BIname", "family_tax", "source_tax"))

# checks
family_check =
  tx_dt[ is.na(family_tax) | family_tax == '' ]

if (nrow(family_check) != 0) {
  warning(nrow(family_check), ' missing family names.')
}
```

# Errata

Final errata mostly due to change in taxonomic hierachies and names.
```{r errata}
# COL: 'not assigned'
tx_dt[ latin_BIname == 'Ceriodaphnia silvestrii', `:=`
       (family_tax = 'Daphniidae',
        source_tax = 'by_hand')]
tx_dt[ latin_BIname == 'Daphnia spinulata', `:=`
       (family_tax = 'Daphniidae',
        source_tax = 'by_hand')]
tx_dt[ latin_BIname == 'Oculimacula yallundae', `:=`
       (family_tax = 'Dermateaceae',
        source_tax = 'by_hand')]
tx_dt[ latin_BIname == 'Pseudosida ramosa', `:=`
       (family_tax = 'Sididae',
        source_tax = 'by_hand')] 	
tx_dt[ latin_BIname == 'Simocephalus elizabethae', `:=`
       (family_tax = 'Daphniidae',
        source_tax = 'by_hand')]

# Wrongly classified by taxize (from PPDB or Malaj)
tx_dt[ latin_BIname == 'Navicula seminulum', `:=`
       (family_tax = 'Naviculaceae',
        source_tax = 'by_hand')]

# Other:
tx_dt[ family_tax == 'Melolonthidae', `:=`
       (family_tax = 'Scarabaeidae',
        source_tax = 'by_hand') ]
tx_dt[ family_tax == 'Megaperidae', `:=`
       (family_tax = 'Apocreadiidae',
        source_tax = 'by_hand') ]
tx_dt[ family_tax == 'Pentatomoidea', `:=` # superfamily_tax
       (family_tax = 'Pentatomidae',
        source_tax = 'by_hand') ]
tx_dt[ family_tax == 'Lymantriidae', `:=` # subfamily_tax
       (family_tax = 'Erebidae',
        source_tax = 'by_hand')]
tx_dt[ family_tax == 'Filiniidae', `:=` # unaccepted synonym
       (family_tax = 'Trochosphaeridae',
        source_tax = 'by_hand')]
```

# Cleaning
```{r cleaning}
oldw = getOption("warn")
options(warn = -1)

rm(dt, family_check, tx_list, tx,
   leftovers, result_itis, result_col, result_nbn, result_tol,
   list = grep('todo', ls(), value = TRUE),
   latin_BIname, online, time_tot, i)

options(warn = oldw)
```

### 2018-08-02 TeX/organigram2.tex
added .tex for organigram of data preparation

\documentclass{article}
%\usepackage{graphicx} % includegraphics
\usepackage{tikz}
\usetikzlibrary{trees, arrows, shapes.geometric} % TikZ libraries
\usepackage{lscape}

\begin{document}

\thispagestyle{empty} % removes the 1

% tree styles (defined by template)
\tikzstyle{every node}=[draw=black,thick,anchor=west]
\tikzstyle{selected}=[draw=red,fill=red!30]
\tikzstyle{optional}=[dashed,fill=gray!50]
% self defined
\tikzstyle{io}=[rectangle, rounded corners, minimum width = 2cm, minimum height = 1cm, text centered, draw = black, fill = blue!10]
\tikzstyle{main}=[rectangle, rounded corners, minimum width = 3cm, minimum height = 1cm, text centered, draw = black, fill = red!30]
\tikzstyle{ids}=[diamond, minimum width = 3cm, minimum height = 1cm, text centered, draw = black, fill = green!30]
\tikzstyle{res}=[rectangle, rounded corners, minimum width = 2cm, minimum height = 1cm, text centered, draw = black, fill = blue!30]

%arrows
\tikzstyle{arrow} = [thick, ->, > = stealth]

\begin{landscape}
%\parindent=0pt % extra spaces towards the edges
%\eject\null
%\vfill
%\hspace*{-4cm}

\begin{tikzpicture}[
  grow via three points={one child at (0.5,-1) and
  two children at (0.5,-1) and (0.5,-1.7)},
  edge from parent path={(\tikzparentnode.south) |- (\tikzchildnode.west)},
  remember picture, overlay, % r.p. save picture size in the picture, o. puts picture out of text segm.
  inner sep = 0pt, outer sep = 0pt
]

%%% nodes %%%
%% input
% TODO coordinates (-3,4) is strange. Also rectangle around current page seems wrong
\node[text=red] (io) [io, align = left] at (-3,4) {Table of Chemicals \\ CAS, Compound name}; % coordinates: (current page.east)

\draw[thick] (current page.south west) rectangle (current page.north east); % debuging

%% chemicals
% id nodes
\node (cas) [ids, below of=io, yshift = -4cm] {CAS};
% main nodes
\node (epa) [main, right of=cas, yshift=1cm, xshift=3cm] {EPA ectox tests}
	child { node { qu\_epa.R} };
\node (chemscr) [main, right of=cas, yshift=-1cm, xshift=3cm] {Chemical scripts ch}
	child { node { Alan Wood Pesticide Compendium - \textbf{qu\_aw.R } }}
	child { node { Pesticide Action Network - \textbf{ch\_pp.R} }}
	child { node { PhysProp - \textbf{ch\_pp.R} }}
	child { node { PubChem - \textbf{qu\_pubchem.R} }
		child { node { Variable - InchiKey }} };
% result nodes
\node (toxres) [res, right of=cas, yshift=0cm, xshift=7cm] {ETOX Test Results}; % result
%\includegraphics{/home/andreas/Documents/Projects/etox-base/data/pubchemlogo2017.png};

%\animategraphics{https://cfpub.epa.gov/ecotox/epafiles/images/logo_epaseal.gif}

%% organisms
% id nodes
\node (taxon) [ids, below of=cas, yshift=-9cm, align = center] {Taxon}; % \\ {(Genus or Species)}}; % escape brackets usin curly brackets
% main nodes
\node (taxonquery) [main, right of=taxon, xshift=3cm, yshift=0cm] {Taxonomic data};
\node (habitat) [main, right of=taxon, xshift=3cm, yshift=1.5cm] {Habitat data}
	child { node { WORMS data base - \textbf{tx\_worms.R} }
		child { node { Family }}
		child { node { Species }} }
	child [dashed] { node { Freshwaterecology.info data }} ;
\node (region) [main, right of = taxon, xshift=3cm, yshift=-1.5cm] {Regional data};
% result nodes
\node (taxres) [res, right of=taxon, xshift=7cm] {Taxon and Habitat Result}; % result

%% final result
\node (testres) [res] at (15,-5) {Final result};

%%% arrows %%%
\draw [arrow] (io) -- (cas);
\draw [->] (cas) -- node[anchor = east] {receive} (chemscr);
\draw [->, dashed] (cas) -- node[anchor = south] {receive} (epa);
\draw [->] (epa) -- (toxres);
\draw [->] (chemscr) -- (toxres);
\draw [->] (toxres) .. controls (5,-6) and (1,-4) .. (taxon);
\draw [->] (taxon) -- (habitat);
\draw [->] (taxon) -- (region);
\draw [->] (taxon) -- (taxonquery);
\draw [->] (habitat) -- (taxres);
\draw [->] (taxonquery) -- (taxres);
\draw [->] (region) -- (taxres);
\draw [->] (toxres) -- (testres);
\draw [->] (taxres) -- (testres);
\draw [->, dashed] (testres) -- (20,-5);

\end{tikzpicture}
\hfill
\vfill
\end{landscape}

\clearpage

\begin{landscape}
\begin{tikzpicture}

%% entities
\node [res] (clean) at (-5,5) {Cleaning};
\node (filter) [res, right of = clean, xshift = 3cm] {Filter};
\node (agg1) [res, right of = filter, xshift = 3cm] {Aggregation 1};
\node (agg2) [res, right of = agg1, xshift = 3cm] {Aggregation 2};

%% arrows
\draw [->, dashed] (-5,8) -- (clean);

%% boxes
\draw [dashed] (1,6) rectangle (3,-11);
\draw [dashed] (4,5) rectangle (7,-11);
\draw [dashed] (9,5) rectangle (11,-11);
\draw [dashed] (14,5) rectangle (15,-11);

\end{tikzpicture}
\end{landscape}

\end{document}

### 2018-08-02 R/EC50aggregation.Rmd
added french psm

--- 2018-07-09 src/EC50aggregation.Rmd
@@
 srcdir = '/home/andreas/Documents/Projects/species-sensitivity/src'
+datadir = '/home/andreas/Documents/Projects/etox-base/data'
 
@@
 ```{r subst1}
-## load data
-psm = readRDS(file.path(cachedir, 'psm1.rds')) # created in EC50sensitivity_data.Rmd
-psm = psm[ !is.na(casnr) ] # only metabolite and sumed pesticides
-psm = psm[ !subst_name %in% c('Desethylsimazin') ] # casnr: 1007289; duplicated with Desethylsimazin
+## (1) bfg_monitoring pesticides
+psm_bfg = readRDS(file.path(cachedir, 'psm1.rds')) # created in EC50sensitivity_data.Rmd
+psm_bfg = psm_bfg[ !is.na(casnr) ] # only metabolite and sumed pesticides
+psm_bfg = psm_bfg[ !subst_name %in% c('Desethylsimazin') ] # casnr: 1007289; duplicated with Desethylsimazin
+psm_bfg[205, cas := '72490-01-8'] # bug (instead of 72490-0-1--8)
 
-## for Nadin and verena
+## (2) French pesticides
+# obtained from Mira
+psm_mi = fread(file.path(datadir, 'missingVal.FR.csv'))
+psm_mi[ , cas := casconv(casnr)] # convert to CAS
+psm_mi = psm_mi[!cas %in% psm_bfg$cas]
+
+## (3) for Nadin and verena
 # psm = fread('/home/andreas/Documents/data/romania/det_compounds_with_CAS.csv')
@@
 #       psm_type = NA)]
+
+# combine data.frame
+psm = rbindlist(list(psm_bfg, psm_mi), fill = TRUE)
 

### 2018-08-09 TeX/organigram2.tex
changed a lot

@@
 %\usepackage{graphicx} % includegraphics
+%% packages
 \usepackage{tikz}
 \usetikzlibrary{trees, arrows, shapes.geometric} % TikZ libraries
+%\usetikzlibrary{backgrounds}
 \usepackage{lscape}
+
+%% layers
+\pgfdeclarelayer{bg1}    % declare background layer
+\pgfdeclarelayer{bg2}
+\pgfsetlayers{bg2,bg1,main}  % set the order of the layers (main is the standard layer)
 
@@
 \tikzstyle{every node}=[draw=black,thick,anchor=west]
-\tikzstyle{selected}=[draw=red,fill=red!30]
-\tikzstyle{optional}=[dashed,fill=gray!50]
+%\tikzstyle{selected}=[draw=red,fill=red!30]
+%\tikzstyle{optional}=[dashed,fill=gray!50]
 % self defined
-\tikzstyle{io}=[rectangle, rounded corners, minimum width = 2cm, minimum height = 1cm, text centered, draw = black, fill = blue!10]
-\tikzstyle{main}=[rectangle, rounded corners, minimum width = 3cm, minimum height = 1cm, text centered, draw = black, fill = red!30]
+\tikzstyle{io}=[rectangle, minimum width = 4cm, minimum height = 1cm, text centered, draw = black, fill = blue!10]
+\tikzstyle{main}=[rectangle, rounded corners, minimum width = 3cm, minimum height = 1cm, text centered, draw = black, fill = purple!30]
 \tikzstyle{ids}=[diamond, minimum width = 3cm, minimum height = 1cm, text centered, draw = black, fill = green!30]
 \tikzstyle{res}=[rectangle, rounded corners, minimum width = 2cm, minimum height = 1cm, text centered, draw = black, fill = blue!30]
+\tikzstyle{resvariables}=[rectangle, minimum width = 2cm, minimum height = 1cm, text centered, draw = black]
+\tikzstyle{steps} = [rectangle, minimum width = 1cm, text centered, fill = yellow]
 
 %arrows
-\tikzstyle{arrow} = [thick, ->, > = stealth]
+\tikzstyle{arrow} = [->, > = stealth, draw = black, color = yellow!80, line width = 6pt]
+% TODO double colored arrows - don't know how it works
+\iffalse
+\tikzset{
+	darrow/.style args={#1 colored by #2 and #3}{
+		-stealth,line width=#1,#2, % first arrow
+		postaction={draw,-stealth,#3,line width=(#1)/3,
+			shorten <=(#1)/3,shorten >=2*(#1)/3}, % second arrow
+	}
+}
+\fi
 
@@
 
-\draw[thick] (current page.south west) rectangle (current page.north east); % debuging
+% \draw[thick] (current page.south west) rectangle (current page.north east); % debuging
 
@@
 % main nodes
-\node (epa) [main, right of=cas, yshift=1cm, xshift=3cm] {EPA ectox tests}
+\node (chemprop) [main, dashed, right of=cas, xshift=4.5cm, yshift=3cm] {CHEMPROP};
+\node (epa) [main, right of=cas, yshift=1cm, xshift=4.5cm] {EPA ectox tests}
 	child { node { qu\_epa.R} };
-\node (chemscr) [main, right of=cas, yshift=-1cm, xshift=3cm] {Chemical scripts ch}
+\node (chemscr) [main, right of=cas, yshift=-1cm, xshift=4.5cm] {Chemical scripts ch}
 	child { node { Alan Wood Pesticide Compendium - \textbf{qu\_aw.R } }}
@@
 	child { node { PhysProp - \textbf{ch\_pp.R} }}
-	child { node { PubChem - \textbf{qu\_pubchem.R} }
-		child { node { Variable - InchiKey }} };
+	child { node { PubChem - \textbf{qu\_pubchem.R} }};
 % result nodes
-\node (toxres) [res, right of=cas, yshift=0cm, xshift=7cm] {ETOX Test Results}; % result
+\node (toxres) [resvariables, right of=cas, yshift=0cm, xshift=16cm, align = left] {Results\\Test results\\Substance type:\\pesticides, metals, etc.}; % result
 %\includegraphics{/home/andreas/Documents/Projects/etox-base/data/pubchemlogo2017.png};
@@
 % main nodes
-\node (taxonquery) [main, right of=taxon, xshift=3cm, yshift=0cm] {Taxonomic data};
-\node (habitat) [main, right of=taxon, xshift=3cm, yshift=1.5cm] {Habitat data}
-	child { node { WORMS data base - \textbf{tx\_worms.R} }
-		child { node { Family }}
-		child { node { Species }} }
-	child [dashed] { node { Freshwaterecology.info data }} ;
-\node (region) [main, right of = taxon, xshift=3cm, yshift=-1.5cm] {Regional data};
+\node (taxonquery) [main, right of=taxon, xshift=4.5cm, yshift=3cm] {Taxonomic data};
+\node (habitat) [main, right of=taxon, xshift=4.5cm, yshift=1cm] {Habitat data}
+	child { node { WORMS data base - \textbf{tx\_worms.R} }}
+	child { node { Manual Search - \textbf{ha\_manual.R} }};
+\node (region) [main, right of = taxon, xshift=4.5cm, yshift=-2cm] {Regional data}
+	child { node { GBIF - \textbf{qu\_gbif.R} }};
 % result nodes
-\node (taxres) [res, right of=taxon, xshift=7cm] {Taxon and Habitat Result}; % result
+\node (taxres) [resvariables, right of=taxon, xshift=17cm, align = left] {Results:\\Validation of taxonomic ranks\\Habitat information\\Marine, Bracksih, Freshwater, Terrestrial, Parasites\\Occurrence}; % result
 
 %% final result
-\node (testres) [res] at (15,-5) {Final result};
+\node (testres) [res] at (16,-5) {Final result}
+	child { node { re\_merge.R} };
+\iffalse
+node[anchor = south]
+node[anchor = east]
+\fi
 
 %%% arrows %%%
-\draw [arrow] (io) -- (cas);
-\draw [->] (cas) -- node[anchor = east] {receive} (chemscr);
-\draw [->, dashed] (cas) -- node[anchor = south] {receive} (epa);
-\draw [->] (epa) -- (toxres);
-\draw [->] (chemscr) -- (toxres);
-\draw [->] (toxres) .. controls (5,-6) and (1,-4) .. (taxon);
-\draw [->] (taxon) -- (habitat);
-\draw [->] (taxon) -- (region);
-\draw [->] (taxon) -- (taxonquery);
-\draw [->] (habitat) -- (taxres);
-\draw [->] (taxonquery) -- (taxres);
-\draw [->] (region) -- (taxres);
-\draw [->] (toxres) -- (testres);
-\draw [->] (taxres) -- (testres);
-\draw [->, dashed] (testres) -- (20,-5);
+\begin{pgfonlayer}{bg1}
+\draw [arrow, black] (io) -- (cas);
+\draw [arrow, dashed] (cas) -- (chemprop);
+\draw [arrow] (cas) -- (chemscr);
+\draw [arrow] (cas) -- (epa);
+\draw [arrow] (epa) -- (toxres);
+\draw [arrow] (chemscr) -- (toxres);
+
+%\begin{scope}[on background layer]
+\begin{pgfonlayer}{bg2}
+	\draw [arrow, black] (epa) .. controls (-2,-4) and (1,-4) .. (taxon);
+\end{pgfonlayer}
+%\end{scope}
+
+\draw [arrow] (taxon) -- (habitat);
+\draw [arrow] (taxon) -- (region);
+\draw [arrow] (taxon) -- (taxonquery);
+\draw [arrow] (habitat) -- (taxres);
+\draw [arrow] (taxonquery) -- (taxres);
+\draw [arrow] (region) -- (taxres);
+\draw [arrow] (toxres) -- (testres);
+\draw [arrow] (taxres) -- (testres);
+
+\draw [arrow, dashed] (testres) -- (20,-5);
+\end{pgfonlayer}
 
@@
 
-\clearpage
+%\clearpage
 
@@
 %% entities
-\node [res] (clean) at (-5,5) {Cleaning};
-\node (filter) [res, right of = clean, xshift = 3cm] {Filter};
-\node (agg1) [res, right of = filter, xshift = 3cm] {Aggregation 1};
-\node (agg2) [res, right of = agg1, xshift = 3cm] {Aggregation 2};
+\node (clean) [res] at (-5,5) {Cleaning}
+	child { node { re\_clean.R} };
+\node (filter) [res, right of = clean, xshift = 3cm] {Filter}
+	child { node { re\_filter.R} };
+\node (plot) [res, right of = filter, xshift = 3cm] {Plots}
+	child { node { re\_plot.R} };
+\node (agg1) [res, right of = plot, xshift = 3cm] {Aggregation};
+
+\node (write) [io, right of = agg1, xshift = 3cm, yshift = -5cm] {Write tables to data base};
+
+%% cleaning and filter steps
+\node (solub) [steps, below of = clean, yshift = -2cm] {Solubility check};
+\node (substtype) [steps, below of = solub, yshift = -2cm] {Substance-type: F, AI, U, T};
+
+\node (habitat) [steps, right of = solub, xshift = 7cm] {Habitat: Marine, Brackish, Freshwater, Terrestrial};
+\node (region) [steps, right of = solub, xshift = 7cm, yshift = -2cm] {Regions: Americas, Africa, Freshwater, Terrestrial};
+\node (duration) [steps, right of = solub, xshift = 7cm, yshift = -4cm] {Tests between 22 and 338h};
+\node (taxagroups) [steps, right of = solub, xshift = 7cm, yshift = -6cm] {};
+
+\node (agg1) [steps, right of = solub, xshift = 7cm, yshift = -6cm] {by casnr, taxon and duration};
 

### 2018-08-09 TeX/organigram2.tex
additional changes

@@
 \tikzstyle{arrow} = [->, > = stealth, draw = black, color = yellow!80, line width = 6pt]
-% TODO double colored arrows - don't know how it works
-\iffalse
-\tikzset{
-	darrow/.style args={#1 colored by #2 and #3}{
-		-stealth,line width=#1,#2, % first arrow
-		postaction={draw,-stealth,#3,line width=(#1)/3,
-			shorten <=(#1)/3,shorten >=2*(#1)/3}, % second arrow
-	}
-}
-\fi
 
@@
 % TODO coordinates (-3,4) is strange. Also rectangle around current page seems wrong
-\node[text=red] (io) [io, align = left] at (-3,4) {Table of Chemicals \\ CAS, Compound name}; % coordinates: (current page.east)
+\node[text=red] (io) [io, align = left] at (-3cm,4cm) {Table of Chemicals \\ CAS, Compound name}; % coordinates: (current page.east)
 
@@
 	child { node { Alan Wood Pesticide Compendium - \textbf{qu\_aw.R } }}
-	child { node { Pesticide Action Network - \textbf{ch\_pp.R} }}
+	child { node { Pesticide Action Network - \textbf{ch\_pan.R} }}
 	child { node { PhysProp - \textbf{ch\_pp.R} }}
@@
 %% final result
-\node (testres) [res] at (16,-5) {Final result}
+\node (testres) [res] at (18,-5) {Final result}
 	child { node { re\_merge.R} };
@@
 
-\draw [arrow, dashed] (testres) -- (20,-5);
+\draw [arrow] (testres) -- (22,-5);
 \end{pgfonlayer}
@@
 \end{tikzpicture}
-\hfill
-\vfill
 \end{landscape}
@@
 %\clearpage
+%\thispagestyle{empty} % removes the 1
 
@@
 
+%\draw[thick] (current page.south west) rectangle (current page.north east); % debuging
+
 %% entities
+\node (test) [res] at (-10,-10) {Test}; %debuging
+
 \node (clean) [res] at (-5,5) {Cleaning}

### 2018-08-09 TeX/organigram1.tex
added 2 new organigram .tex files

\documentclass[tikz, border={1000 750}]{standalone}
%%%% packages
\usepackage{tikz}
\usetikzlibrary{trees, arrows, shapes.geometric} % TikZ libraries
%%%% layers
\pgfdeclarelayer{bg1}    % declare background layer
\pgfdeclarelayer{bg2}
\pgfsetlayers{bg2,bg1,main}  % set the order of the layers (main is the standard layer)

%%%% styles
\tikzset{
  every rectangle node/.style = {draw=black,thick,anchor=west},
  io/.style = {rectangle, minimum width = 4cm, minimum height = 3cm, text centered, draw = black, fill = blue!10},
  main/.style = {rectangle, rounded corners, minimum width = 3cm, text centered, fill = purple!30},
  ids/.style = {diamond, minimum width = 3cm, minimum height = 2cm, text centered, draw = black, fill = green!30},
  res/.style = {rectangle, rounded corners, minimum width = 2cm, minimum height = 1cm, text centered, draw = black, fill = blue!30},
  resvariables/.style = {rectangle, minimum width = 2cm, minimum height = 1cm, text centered, draw = black},
  steps/.style = {rectangle, minimum width = 1cm, text centered, fill = yellow},
  arrow/.style = {->, > = stealth, draw = black, color = yellow!80, line width = 6pt}
}

\begin{document}
	\begin{tikzpicture}[auto,
	grow via three points={one child at (0.5,-1) and
		two children at (0.5,-1) and (0.5,-1.7)},
	edge from parent path={(\tikzparentnode.south) |- (\tikzchildnode.west)},
	remember picture, overlay, % r.p. save picture size in the picture, o. puts picture out of text segm.
	inner sep = 0pt, outer sep = 0pt
	]

		\node[text=red] (io) [io, align = left] at (0cm,0cm) {Table of Chemicals \\ CAS, Compound name};
		%% chemicals
		% id nodes

		\node (cas) [ids, below of=io, yshift = -4cm] {CAS};

		% main nodes
		\node (chemprop) [main, dashed, right of=cas, xshift=4.5cm, yshift=3cm] {CHEMPROP};
		\node (epa) [main, right of=cas, yshift=1cm, xshift=4.5cm] {EPA ectox tests}
			child { node { qu\_epa.R} };
		\node (chemscr) [main, right of=cas, yshift=-1cm, xshift=4.5cm] {Chemical scripts ch}
			child { node { Alan Wood Pesticide Compendium - \textbf{qu\_aw.R } }}
			child { node { Pesticide Action Network - \textbf{ch\_pan.R} }}
			child { node { PhysProp - \textbf{ch\_pp.R} }}
			child { node { PubChem - \textbf{qu\_pubchem.R} }};
		% result nodes
		\node (toxres) [resvariables, right of=cas, yshift=0cm, xshift=18cm, align = left] {Results\\Test results\\Substance type:\\pesticides, metals, etc.}; % result
		%\includegraphics{/home/andreas/Documents/Projects/etox-base/data/pubchemlogo2017.png};

		%\animategraphics{https://cfpub.epa.gov/ecotox/epafiles/images/logo_epaseal.gif}

		%% organisms
		% id nodes
		\node (taxon) [ids, below of=cas, yshift=-9cm, align = center] {Taxon}; % \\ {(Genus or Species)}}; % escape brackets usin curly brackets
		% main nodes
		\node (taxonquery) [main, right of=taxon, xshift = 4.5cm, yshift = 3cm] {Taxonomic data};
		\node (habitat) [main, right of=taxon, xshift = 4.5cm, yshift = 1cm] {Habitat data}
			child { node { WORMS data base - \textbf{tx\_worms.R} }}
			child { node { Manual Search - \textbf{ha\_manual.R} }};
		\node (region) [main, right of = taxon, xshift = 4.5cm, yshift = -2cm] {Regional data}
			child { node { GBIF - \textbf{qu\_gbif.R} }};
		% result nodes
		\node (taxres) [resvariables, right of=taxon, xshift = 17cm, align = left] {Results:\\Validation of taxonomic ranks\\Habitat information\\Marine, Bracksih, Freshwater, Terrestrial, Parasites\\Occurrence}; % result

		%% final result
		\node (testres) [res, right of = cas, xshift = 22cm, yshift = -5cm] {Final result}
			child { node { re\_merge.R} };

		%%% arrows %%%
		\begin{pgfonlayer}{bg1}
		\draw [arrow, black] (io) -- (cas);
		\draw [arrow, dashed] (cas.east) -- (chemprop.west);
		\draw [arrow] (cas.east) -- (chemscr.west);
		\draw [arrow] (cas.east) -- (epa.west);
		\draw [arrow] (epa.east) -- (toxres.west);
		\draw [arrow] (chemscr.east) -- (toxres.west);

		%\begin{scope}[on background layer]
		\begin{pgfonlayer}{bg2}
			\draw [arrow, black] (epa) .. controls (3cm,-8cm) .. (taxon);
		\end{pgfonlayer}
		%\end{scope}

		\draw [arrow] (taxon.east) -- (habitat.west);
		\draw [arrow] (taxon.east) -- (region.west);
		\draw [arrow] (taxon.east) -- (taxonquery.west);
		\draw [arrow] (habitat.east) -- (taxres.west);
		\draw [arrow] (taxonquery.east) -- (taxres.west);
		\draw [arrow] (region.east) -- (taxres.west);
		\draw [arrow] (toxres) -- (testres.north);
		\draw [arrow] (taxres) -- (testres.south);

		\draw [arrow, right of = testres] (testres) -- (20cm,-2cm);

		\end{pgfonlayer}
	\end{tikzpicture}
\end{document}

### 2018-08-22 TeX/organigram1.tex
a lot of changes

@@
-\documentclass[tikz, border={1000 750}]{standalone}
+\documentclass[border = 1.5cm]{standalone}
 %%%% packages
 \usepackage{tikz}
-\usetikzlibrary{trees, arrows, shapes.geometric} % TikZ libraries
+\usetikzlibrary{trees, arrows, shapes.geometric, positioning} % TikZ libraries
 %%%% layers
@@
 \tikzset{
-  every rectangle node/.style = {draw=black,thick,anchor=west},
+	every rectangle node/.style = {thick,anchor=west},
 	io/.style = {rectangle, minimum width = 4cm, minimum height = 3cm, text centered, draw = black, fill = blue!10},
-  main/.style = {rectangle, rounded corners, minimum width = 3cm, text centered, fill = purple!30},
+	main/.style = {rectangle, draw = black, rounded corners, minimum width = 3cm, minimum height = 1.25cm, text centered},
 	ids/.style = {diamond, minimum width = 3cm, minimum height = 2cm, text centered, draw = black, fill = green!30},
 	res/.style = {rectangle, rounded corners, minimum width = 2cm, minimum height = 1cm, text centered, draw = black, fill = blue!30},
-  resvariables/.style = {rectangle, minimum width = 2cm, minimum height = 1cm, text centered, draw = black},
+	resvariables/.style = {rectangle, minimum width = 2cm, minimum height = 1cm, text centered},
 	steps/.style = {rectangle, minimum width = 1cm, text centered, fill = yellow},
-  arrow/.style = {->, > = stealth, draw = black, color = yellow!80, line width = 6pt}
+	arrow/.style = {->, > = stealth, draw = black, color = gray!80, line width = 3pt}
 }
@@
 \begin{document}
-	\begin{tikzpicture}[auto,
+	\begin{tikzpicture}[
 	grow via three points={one child at (0.5,-1) and
@@
 	edge from parent path={(\tikzparentnode.south) |- (\tikzchildnode.west)},
-	remember picture, overlay, % r.p. save picture size in the picture, o. puts picture out of text segm.
 	inner sep = 0pt, outer sep = 0pt
@@
 
-		\node[text=red] (io) [io, align = left] at (0cm,0cm) {Table of Chemicals \\ CAS, Compound name};
+		\node (io) [io, align = left] at (0cm,0cm) {Table of Chemicals \\ CAS, Compound name};
 		%% chemicals
@@
 			child { node { qu\_epa.R} };
-		\node (chemscr) [main, right of=cas, yshift=-1cm, xshift=4.5cm] {Chemical scripts ch}
+		\node (chemscr) [main, right of=cas, yshift=-1cm, xshift=4.5cm] {Chemical scripts}
 			child { node { Alan Wood Pesticide Compendium - \textbf{qu\_aw.R } }}
@@
 		% result nodes
-		\node (toxres) [resvariables, right of=cas, yshift=0cm, xshift=18cm, align = left] {Results\\Test results\\Substance type:\\pesticides, metals, etc.}; % result
+		\node (toxres) [resvariables, right of=cas, yshift=0cm, xshift=16cm, align = left] {Results\\Test results\\Substance type:\\pesticides, metals, etc.}; % result
 		%\includegraphics{/home/andreas/Documents/Projects/etox-base/data/pubchemlogo2017.png};
@@
 		% result nodes
-		\node (taxres) [resvariables, right of=taxon, xshift = 17cm, align = left] {Results:\\Validation of taxonomic ranks\\Habitat information\\Marine, Bracksih, Freshwater, Terrestrial, Parasites\\Occurrence}; % result
+		\node (taxres) [resvariables, right of=taxon, xshift = 14cm, align = left] {Results:\\Validation of taxonomic ranks\\Habitat information\\Marine, Bracksih, Freshwater, Terrestrial, Parasites\\Occurrence}; % result
 
 		%% final result
-		\node (testres) [res, right of = cas, xshift = 22cm, yshift = -5cm] {Final result}
+		\node (testres) [res, right of = cas, xshift = 18cm, yshift = -5cm] {Result}
 			child { node { re\_merge.R} };
+		\node (filtres) [res, right of = testres, xshift = 2cm] {Filter};
+		\node (aggres) [res, right of = filtres, xshift = 2cm] {Aggregate};
 
@@
 		\draw [arrow] (chemscr.east) -- (toxres.west);
+
+%		\draw [arrow] (epa.east) -| (testres.north);
+%		\draw [arrow] (chemscr.east) -| (testres.north);
 
@@
 
-		\draw [arrow, right of = testres] (testres) -- (20cm,-2cm);
+		\draw [arrow, right of = testres] (testres) -- ++(2cm,0cm);
 

### 2018-08-22 TeX/organigram2.tex
a lot of changes

@@
-\documentclass{article}
-%\usepackage{graphicx} % includegraphics
-%% packages
+\documentclass[border = 1.5cm]{standalone}
+%%%% packages
 \usepackage{tikz}
 \usetikzlibrary{trees, arrows, shapes.geometric} % TikZ libraries
-%\usetikzlibrary{backgrounds}
-\usepackage{lscape}
-
-%% layers
+%%%% layers
 \pgfdeclarelayer{bg1}    % declare background layer
@@
 
+%%%% styles
+\tikzset{
+	every rectangle node/.style = {draw=black,thick},
+	io/.style = {rectangle, minimum width = 4cm, minimum height = 3cm, text centered, draw = black, fill = blue!10},
+	main/.style = {rectangle, rounded corners, minimum width = 3cm, text centered, fill = purple!30},
+	ids/.style = {diamond, minimum width = 3cm, minimum height = 2cm, text centered, draw = black, fill = green!30},
+	res/.style = {rectangle, rounded corners, minimum width = 2cm, minimum height = 1cm, text centered, draw = black, fill = blue!30},
+	resvariables/.style = {rectangle, minimum width = 2cm, minimum height = 1cm, text centered, draw = black},
+	steps/.style = {rectangle, minimum width = 1cm, text centered, fill = yellow},
+	txtfield/.style = {rectangle, rounded corners, text width = 3cm, minimum height = 6cm},
+	arrow/.style = {->, > = stealth, draw = black, color = gray!80, shorten > = 2pt, line width = 3pt}
+}
+
 \begin{document}
-
-\thispagestyle{empty} % removes the 1
-
-% tree styles (defined by template)
-\tikzstyle{every node}=[draw=black,thick,anchor=west]
-%\tikzstyle{selected}=[draw=red,fill=red!30]
-%\tikzstyle{optional}=[dashed,fill=gray!50]
-% self defined
-\tikzstyle{io}=[rectangle, minimum width = 4cm, minimum height = 1cm, text centered, draw = black, fill = blue!10]
-\tikzstyle{main}=[rectangle, rounded corners, minimum width = 3cm, minimum height = 1cm, text centered, draw = black, fill = purple!30]
-\tikzstyle{ids}=[diamond, minimum width = 3cm, minimum height = 1cm, text centered, draw = black, fill = green!30]
-\tikzstyle{res}=[rectangle, rounded corners, minimum width = 2cm, minimum height = 1cm, text centered, draw = black, fill = blue!30]
-\tikzstyle{resvariables}=[rectangle, minimum width = 2cm, minimum height = 1cm, text centered, draw = black]
-\tikzstyle{steps} = [rectangle, minimum width = 1cm, text centered, fill = yellow]
-
-%arrows
-\tikzstyle{arrow} = [->, > = stealth, draw = black, color = yellow!80, line width = 6pt]
-
-\begin{landscape}
-%\parindent=0pt % extra spaces towards the edges
-%\eject\null
-%\vfill
-%\hspace*{-4cm}
 
@@
 	edge from parent path={(\tikzparentnode.south) |- (\tikzchildnode.west)},
-  remember picture, overlay, % r.p. save picture size in the picture, o. puts picture out of text segm.
-  inner sep = 0pt, outer sep = 0pt
+	anchor = west, inner sep = 0pt, outer sep = 0pt
 	]
-
-%%% nodes %%%
-%% input
-% TODO coordinates (-3,4) is strange. Also rectangle around current page seems wrong
-\node[text=red] (io) [io, align = left] at (-3cm,4cm) {Table of Chemicals \\ CAS, Compound name}; % coordinates: (current page.east)
-
-% \draw[thick] (current page.south west) rectangle (current page.north east); % debuging
-
-%% chemicals
-% id nodes
-\node (cas) [ids, below of=io, yshift = -4cm] {CAS};
-% main nodes
-\node (chemprop) [main, dashed, right of=cas, xshift=4.5cm, yshift=3cm] {CHEMPROP};
-\node (epa) [main, right of=cas, yshift=1cm, xshift=4.5cm] {EPA ectox tests}
-	child { node { qu\_epa.R} };
-\node (chemscr) [main, right of=cas, yshift=-1cm, xshift=4.5cm] {Chemical scripts ch}
-	child { node { Alan Wood Pesticide Compendium - \textbf{qu\_aw.R } }}
-	child { node { Pesticide Action Network - \textbf{ch\_pan.R} }}
-	child { node { PhysProp - \textbf{ch\_pp.R} }}
-	child { node { PubChem - \textbf{qu\_pubchem.R} }};
-% result nodes
-\node (toxres) [resvariables, right of=cas, yshift=0cm, xshift=16cm, align = left] {Results\\Test results\\Substance type:\\pesticides, metals, etc.}; % result
-%\includegraphics{/home/andreas/Documents/Projects/etox-base/data/pubchemlogo2017.png};
-
-%\animategraphics{https://cfpub.epa.gov/ecotox/epafiles/images/logo_epaseal.gif}
-
-%% organisms
-% id nodes
-\node (taxon) [ids, below of=cas, yshift=-9cm, align = center] {Taxon}; % \\ {(Genus or Species)}}; % escape brackets usin curly brackets
-% main nodes
-\node (taxonquery) [main, right of=taxon, xshift=4.5cm, yshift=3cm] {Taxonomic data};
-\node (habitat) [main, right of=taxon, xshift=4.5cm, yshift=1cm] {Habitat data}
-	child { node { WORMS data base - \textbf{tx\_worms.R} }}
-	child { node { Manual Search - \textbf{ha\_manual.R} }};
-\node (region) [main, right of = taxon, xshift=4.5cm, yshift=-2cm] {Regional data}
-	child { node { GBIF - \textbf{qu\_gbif.R} }};
-% result nodes
-\node (taxres) [resvariables, right of=taxon, xshift=17cm, align = left] {Results:\\Validation of taxonomic ranks\\Habitat information\\Marine, Bracksih, Freshwater, Terrestrial, Parasites\\Occurrence}; % result
-
-%% final result
-\node (testres) [res] at (18,-5) {Final result}
-	child { node { re\_merge.R} };
-\iffalse
-node[anchor = south]
-node[anchor = east]
-\fi
-
-%%% arrows %%%
-\begin{pgfonlayer}{bg1}
-\draw [arrow, black] (io) -- (cas);
-\draw [arrow, dashed] (cas) -- (chemprop);
-\draw [arrow] (cas) -- (chemscr);
-\draw [arrow] (cas) -- (epa);
-\draw [arrow] (epa) -- (toxres);
-\draw [arrow] (chemscr) -- (toxres);
-
-%\begin{scope}[on background layer]
-\begin{pgfonlayer}{bg2}
-	\draw [arrow, black] (epa) .. controls (-2,-4) and (1,-4) .. (taxon);
-\end{pgfonlayer}
-%\end{scope}
-
-\draw [arrow] (taxon) -- (habitat);
-\draw [arrow] (taxon) -- (region);
-\draw [arrow] (taxon) -- (taxonquery);
-\draw [arrow] (habitat) -- (taxres);
-\draw [arrow] (taxonquery) -- (taxres);
-\draw [arrow] (region) -- (taxres);
-\draw [arrow] (toxres) -- (testres);
-\draw [arrow] (taxres) -- (testres);
-
-\draw [arrow] (testres) -- (22,-5);
-\end{pgfonlayer}
-
-\end{tikzpicture}
-\end{landscape}
-
-%\clearpage
-%\thispagestyle{empty} % removes the 1
-
-\begin{landscape}
-\begin{tikzpicture}
 
@@
 
-%% entities
-\node (test) [res] at (-10,-10) {Test}; %debuging
-
-\node (clean) [res] at (-5,5) {Cleaning}
+	%% nodes	
+	\node (clean) [res] at (0cm,0cm) {Cleaning}
 	child { node { re\_clean.R} };
@@
 	child { node { re\_plot.R} };
-\node (agg1) [res, right of = plot, xshift = 3cm] {Aggregation};
+	\node (agg) [res, right of = plot, xshift = 3cm] {Aggregation}
+	child { node { re\_aggregation.R} };
 
-\node (write) [io, right of = agg1, xshift = 3cm, yshift = -5cm] {Write tables to data base};
+	\node (write) [io, right of = agg, xshift = 4cm, yshift = -3cm] {Write tables to data base};
+
+	%% text fields
+	%\node (txtclean2) [txtfield, below of = clean, yshift = -4cm] {Test2};
+	\node (txtclean2) [txtfield, below of = clean, yshift = -4cm] {Try};
+	\node (txtclean2) [txtfield, below of = filter, yshift = -4cm] {Tryasdf};
+	\node (txtclean2) [txtfield, below of = plot, yshift = -4cm] {Tryasdf};
+	\node (txtclean2) [txtfield, below of = agg, yshift = -4cm] {Tryasdf};
+
+	%% arrows
+	\draw [arrow] (clean) -- (filter);
+	\draw [arrow] (filter) -- (plot);
+	\draw [arrow] (plot) -- (agg);
+	\draw [arrow, to path={-| (\tikztotarget)}] (agg) edge (write.north);
 
 	%% cleaning and filter steps
+
+	\iffalse
 	\node (solub) [steps, below of = clean, yshift = -2cm] {Solubility check};
-\node (substtype) [steps, below of = solub, yshift = -2cm] {Substance-type: F, AI, U, T};
+	\node (substtype) [steps, below of = clean, yshift = -4cm] {Substance-type: F, AI, U, T};
 
@@
 	\draw [dashed] (14,5) rectangle (15,-11);
-
+	\fi
 	\end{tikzpicture}
-\end{landscape}
-
 \end{document}

### 2018-08-22 TeX/organigram1.tex
renewes TikX picture

@@
 \documentclass[border = 1.5cm]{standalone}
+%\documentclass[border = 1.5cm, convert={density=300,size=1080x800,outext=.png}]{standalone} % doesn't work
 %%%% packages
@@
 \pgfdeclarelayer{bg2}
-\pgfsetlayers{bg2,bg1,main}  % set the order of the layers (main is the standard layer)
+\pgfdeclarelayer{bg3}
+\pgfsetlayers{bg3,bg2,bg1,main}  % set the order of the layers (main is the standard layer)
 
@@
 \tikzset{
-	every rectangle node/.style = {thick,anchor=west},
+	every rectangle node/.style = {thick, anchor=west},
+	every picture/.style = {/utils/exec={\sffamily}},
+	font = {\fontsize{13pt}{12}\selectfont},
 	io/.style = {rectangle, minimum width = 4cm, minimum height = 3cm, text centered, draw = black, fill = blue!10},
@@
 	steps/.style = {rectangle, minimum width = 1cm, text centered, fill = yellow},
-	arrow/.style = {->, > = stealth, draw = black, color = gray!80, line width = 3pt}
+	arrow/.style = {->, > = stealth, draw = black, color = gray!80, line width = 2.5pt}
 }
@@
 
-		\node (io) [io, align = left] at (0cm,0cm) {Table of Chemicals \\ CAS, Compound name};
+		\node (io) [io, align = center] at (0cm,0cm) {.csv or .xlsx \\ CAS \\ Compound name};
 		%% chemicals
@@
 		\node (chemprop) [main, dashed, right of=cas, xshift=4.5cm, yshift=3cm] {CHEMPROP};
-		\node (epa) [main, right of=cas, yshift=1cm, xshift=4.5cm] {EPA ectox tests}
+		\node (epa) [main, ultra thick, right of=cas, yshift=1cm, xshift=4.5cm] {EPA ECOTOX}
 			child { node { qu\_epa.R} };
-		\node (chemscr) [main, right of=cas, yshift=-1cm, xshift=4.5cm] {Chemical scripts}
+		\node (chemscr) [main, right of=cas, yshift=-1cm, xshift=4.5cm] {PHYS-CHEM}
 			child { node { Alan Wood Pesticide Compendium - \textbf{qu\_aw.R } }}
@@
 			child { node { PubChem - \textbf{qu\_pubchem.R} }};
-		% result nodes
-		\node (toxres) [resvariables, right of=cas, yshift=0cm, xshift=16cm, align = left] {Results\\Test results\\Substance type:\\pesticides, metals, etc.}; % result
-		%\includegraphics{/home/andreas/Documents/Projects/etox-base/data/pubchemlogo2017.png};
-
-		%\animategraphics{https://cfpub.epa.gov/ecotox/epafiles/images/logo_epaseal.gif}
 
@@
 		% main nodes
-		\node (taxonquery) [main, right of=taxon, xshift = 4.5cm, yshift = 3cm] {Taxonomic data};
-		\node (habitat) [main, right of=taxon, xshift = 4.5cm, yshift = 1cm] {Habitat data}
+		\node (taxonquery) [main, right of=taxon, xshift = 4.5cm, yshift = 3cm] {Taxonomy};
+		\node (habitat) [main, right of=taxon, xshift = 4.5cm, yshift = 1cm] {Habitats}
 			child { node { WORMS data base - \textbf{tx\_worms.R} }}
 			child { node { Manual Search - \textbf{ha\_manual.R} }};
-		\node (region) [main, right of = taxon, xshift = 4.5cm, yshift = -2cm] {Regional data}
+		\node (region) [main, right of = taxon, xshift = 4.5cm, yshift = -2cm] {Continents}
 			child { node { GBIF - \textbf{qu\_gbif.R} }};
-		% result nodes
-		\node (taxres) [resvariables, right of=taxon, xshift = 14cm, align = left] {Results:\\Validation of taxonomic ranks\\Habitat information\\Marine, Bracksih, Freshwater, Terrestrial, Parasites\\Occurrence}; % result
 
 		%% final result
-		\node (testres) [res, right of = cas, xshift = 18cm, yshift = -5cm] {Result}
-			child { node { re\_merge.R} };
+		\begin{pgfonlayer}{bg3}
+			\node[rectangle, draw, rounded corners, minimum height = 2cm, minimum width = 7cm, fill = gray!10] (appbox) at (20.5cm,-10cm) {};
+		\end{pgfonlayer}
+		\node (apptext) [above of = appbox, yshift = 0.25cm, xshift = -3cm] {Etox-Base App}; 
+		\node (testres) [res, right of = cas, xshift = 14cm, yshift = -5cm] {Result};
 		\node (filtres) [res, right of = testres, xshift = 2cm] {Filter};
@@
 		\draw [arrow] (cas.east) -- (epa.west);
-		\draw [arrow] (epa.east) -- (toxres.west);
-		\draw [arrow] (chemscr.east) -- (toxres.west);
 
-%		\draw [arrow] (epa.east) -| (testres.north);
-%		\draw [arrow] (chemscr.east) -| (testres.north);
+		\draw [arrow] (epa.east) -| (testres.north);
+		\draw [arrow] (chemscr.east) -| (testres.north);
 
@@
 		\begin{pgfonlayer}{bg2}
-			\draw [arrow, black] (epa) .. controls (3cm,-8cm) .. (taxon);
+			\draw [arrow, black] (epa) .. controls (3cm,-8cm) .. (taxon.north);
 		\end{pgfonlayer}
@@
 		\draw [arrow] (taxon.east) -- (taxonquery.west);
-		\draw [arrow] (habitat.east) -- (taxres.west);
-		\draw [arrow] (taxonquery.east) -- (taxres.west);
-		\draw [arrow] (region.east) -- (taxres.west);
-		\draw [arrow] (toxres) -- (testres.north);
-		\draw [arrow] (taxres) -- (testres.south);
+		\draw [arrow] (taxonquery.east) -| (testres.south);
+		\draw [arrow] (habitat.east) -| (testres.south);
+		\draw [arrow] (region.east) -| (testres.south);
 
-		\draw [arrow, right of = testres] (testres) -- ++(2cm,0cm);
+		\draw [arrow] (testres) -- (filtres); %++(2cm,0cm);
+		\draw [arrow] (filtres) -- (aggres);
 

### 2018-08-22 presentation/talk_internal.Rmd
added internal talk

---
title: "[Decluttering ecotoxicological test data](https://cfpub.epa.gov/ecotox/)"
subtitle: "how to obtain valuable information from the EPA ECOTOX data base"
author: "Andreas Scharmüller"
date: "2018/08/23"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---

```{r setup, echo=FALSE, include=FALSE}
source('/home/andreas/Documents/Projects/etox-base/R/setup.R')
require(knitr)
```

# The EPA ECOTOX data base

<https://cfpub.epa.gov/ecotox/>

<img src="/home/andreas/Documents/Projects/etox-base/presentation/data/epa_screenshot.png" width="600px">

---
# The EPA ECOTOX data base

<img src="/home/andreas/Documents/Projects/etox-base/presentation/data/epa_screenshot.png" width="300px" style="position:absolute; left:85px; top:230px">

<img src="https://cdn.pixabay.com/photo/2012/04/24/11/42/arrow-39526_960_720.png" width="200px" style="position:absolute; left:410px; top:280px">

<img src="https://zdnet4.cbsistatic.com/hub/i/r/2018/04/19/092cbf81-acac-4f3a-91a1-5a26abc1721f/resize/370xauto/ce84e38cb1c1a7c5a2c9e4c337e108ba/postgresql-logo.png" width="250px" style="position:absolute; left:610px; top:255px">

---
# The EPA ECOTOX data base

<img src="/home/andreas/Documents/Projects/etox-base/presentation/data/epa_screenshot.png" width="300px" style="position:absolute; right:40px">

information on ecotoxicological tests

__filters in online version:__

* endpoints (e.g. NOEC, LOEC, __EC50__, ECXX )

* chemicals (CAS, compound name)

* organism (upper-level groups)

__not possible:__

* duration

* distinguish between pesticide groups (e.g. fungicides, herbicides, ...)

* filter organism groups

* refine to specific regions where tested organisms occur

* attribute organism to habitats (e.g. marine, brackish, freshwater, ...)

---
# The EPA ECOTOX data base

<img src="/home/andreas/Documents/Projects/etox-base/presentation/data/epa_screenshot.png" width="300px" style="position:absolute; right:40px">

information on ecotoxicological tests

__filters in online version:__

* endpoints (e.g. NOEC, LOEC, __EC50__, ECXX )

* chemicals (CAS, compound name)

* organism (upper-level groups)

<img src="/home/andreas/Documents/Projects/etox-base/data/pubchemlogo2017.png" width="150px" style="position:absolute; top:430px;">

<img src="http://www.panna.org/sites/all/themes/velocity/new-panlogo.png" width="150px" style="position:absolute; top:470px; left:200px;">

<img src="https://www.itis.gov/Static/images/maintoplogo.gif" width="75px" style="position:absolute; top:510px; left:75px;">

<img src="http://3.bp.blogspot.com/-Wpp0vCJQDHo/U_ZMorXEynI/AAAAAAAAA7Y/Y-f-3brn99k/s1600/GBIF_logo_short_form.gif" width="150px" style="position:absolute; top:520px; right:20px;">

<img src="https://www.easterncapescubadiving.co.za/downloads/Worms.jpg" width="250px" style="position:absolute; top:470px; right:200px;">

---
# The data

```{r test-plot, echo=FALSE, fig.width=9}
gg_test = readRDS(file.path(plotdir, 'gg_test.rds'))
gg_test
```

---
# Aims

* to separate the wheat from the chaff

* EC50 values

* improve filters

* to aggreagte many test results

* usable and accessible

* establish effect relationships between compounds and organisms (not only: _D. magna_, _P. subcapitata_, _P. promelas_)

---
class: right

# Workflow

```{r organigram-conversion, echo=FALSE, include=TRUE, out.width="100%"}
pdf_file = '/home/andreas/Documents/Projects/etox-base/TeX/organigram1.pdf'
png_file = '/home/andreas/Documents/Projects/etox-base/TeX/organigram1.png'

system(sprintf('convert -density 300 %s %s', pdf_file, png_file))
```

<img src="/home/andreas/Documents/Projects/etox-base/TeX/organigram1.png" width="1000px" style="position:absolute; top:10px; left:10px">

--

<img src="https://zdnet4.cbsistatic.com/hub/i/r/2018/04/19/092cbf81-acac-4f3a-91a1-5a26abc1721f/resize/370xauto/ce84e38cb1c1a7c5a2c9e4c337e108ba/postgresql-logo.png" width="150px" style="position:absolute; right:75px; top:210px">

<img src="https://www.rstudio.com/wp-content/uploads/2014/04/shiny.png" width="100px" style="position:absolute; right:105px; bottom:95px">

---
# Filter

* EC50

* CAS

* different organism groups (e.g. Algae, Macro-Invertebrates, Chironomidae, ... )

* test duration

* habitat

* regions

* concentration type (Active ingredient, Formulation)

<img src="/home/andreas/Documents/Projects/etox-base/TeX/organigram1_app.png" width="380px" style="position:absolute; right:0px; top:10px;">

---
# Aggregation

* (1) by CAS, test organism and test duration

  * __if:__ N<sub>tests</sub> > 2 __then:__ median
  * __else:__ minimum

* (2) by CAS

  * mean, median, min, max (can be chosen)

<img src="/home/andreas/Documents/Projects/etox-base/TeX/organigram1_app.png" width="380px" style="position:absolute; right:0px; top:10px">

---
class: middle

[the app](future/link/to/host)

<img src="/home/andreas/Documents/Projects/etox-base/presentation/data/etox_base_screenshot.png" width="600px" style="position:absolute; left:200px; top:150px">

---
class: right, middle

# [Thank you for your attention!](https://www.researchgate.net/profile/Andreas_Scharmueller)
Andreas Scharmüller<br/>

Quantitative Landscape Ecology<br/>
University Koblenz-Landau

<img src="https://upload.wikimedia.org/wikipedia/de/thumb/9/9f/Twitter_bird_logo_2012.svg/1259px-Twitter_bird_logo_2012.svg.png" width="20"> @andschar <br/>
<img src="https://seeklogo.com/images/M/mail-icon-logo-28FE0635D0-seeklogo.com.png" width="20"> scharmueller@uni-landau.de

<img src="https://www.uni-koblenz-landau.de/de/landau/fb7/umweltwissenschaften/logos/land-ecol-neu.jpg" width="100">
<img src="https://www.uni-koblenz-landau.de/de/uni/organisation/verwaltung/abteilungen/abt-1/dokumente-ab1/uni-logo-farbig.jpg" width="200">

### 2018-08-22 TeX/organigram1.tex
changed names

@@
 			child { node { Manual Search - \textbf{ha\_manual.R} }};
-		\node (region) [main, right of = taxon, xshift = 4.5cm, yshift = -2cm] {Continents}
+		\node (region) [main, right of = taxon, xshift = 4.5cm, yshift = -2cm] {Regions}
 			child { node { GBIF - \textbf{qu\_gbif.R} }};
@@
 		\end{pgfonlayer}
-		\node (apptext) [above of = appbox, yshift = 0.25cm, xshift = -3cm] {Etox-Base App}; 
+		\node (apptext) [above of = appbox, yshift = 0.25cm, xshift = -3cm] {Etox-Base Function}; 
 		\node (testres) [res, right of = cas, xshift = 14cm, yshift = -5cm] {Result};

### 2018-08-30 TeX/organigram1.tex
changed TiX picture

@@
 
-		\node (io) [io, align = center] at (0cm,0cm) {.csv or .xlsx \\ CAS \\ Compound name};
+	\node (io) [io, align = center] at (0cm,0cm) {EPA ECOTOX\\data base};
 	%% chemicals
@@
 
-		\node (cas) [ids, below of=io, yshift = -4cm] {CAS};
+	\node (cas) [ids, below of=io, yshift = -4cm, xshift = 3cm] {CAS};
 
 	% main nodes
-		\node (chemprop) [main, dashed, right of=cas, xshift=4.5cm, yshift=3cm] {CHEMPROP};
-		\node (epa) [main, ultra thick, right of=cas, yshift=1cm, xshift=4.5cm] {EPA ECOTOX}
-			child { node { qu\_epa.R} };
+	\node (chemprop) [main, dashed, right of=cas, xshift=4.5cm, yshift=1.5cm] {CHEMPROP};
 	\node (chemscr) [main, right of=cas, yshift=-1cm, xshift=4.5cm] {PHYS-CHEM}
@@
 	%% final result
-		\begin{pgfonlayer}{bg3}
-			\node[rectangle, draw, rounded corners, minimum height = 2cm, minimum width = 7cm, fill = gray!10] (appbox) at (20.5cm,-10cm) {};
-		\end{pgfonlayer}
-		\node (apptext) [above of = appbox, yshift = 0.25cm, xshift = -3cm] {Etox-Base Function}; 
 	\node (testres) [res, right of = cas, xshift = 14cm, yshift = -5cm] {Result};
@@
 
+	\begin{pgfonlayer}{bg3}
+	\node[rectangle, draw, rounded corners, minimum height = 2cm, minimum width = 7cm, fill = gray!10, right of = testres, xshift = 1.5cm] (appbox) {};
+	\end{pgfonlayer}
+
+	\node (apptext) [above of = appbox, yshift = 0.25cm, xshift = -3cm] {Etox-Base Function};
+
 	%%% arrows %%%
 	\begin{pgfonlayer}{bg1}
-		\draw [arrow, black] (io) -- (cas);
+	\draw [arrow, black] (io) |- (cas.west);
 	\draw [arrow, dashed] (cas.east) -- (chemprop.west);
 	\draw [arrow] (cas.east) -- (chemscr.west);
-		\draw [arrow] (cas.east) -- (epa.west);
-
-		\draw [arrow] (epa.east) -| (testres.north);
 	\draw [arrow] (chemscr.east) -| (testres.north);
 
-		%\begin{scope}[on background layer]
-		\begin{pgfonlayer}{bg2}
-			\draw [arrow, black] (epa) .. controls (3cm,-8cm) .. (taxon.north);
-		\end{pgfonlayer}
-		%\end{scope}
+	\draw [arrow, black] (io) |- (taxon.west);
 

### 2018-09-04 article/article.Rmd
added article

---
title: "Etox-Base - A tool and data set in ecotoxicology"
author: "Andreas Scharmüller"
date: "August 29, 2018"
output: html_document
bibliography: references-etox-base.bib
---

# Abstract

# Introduction

Chemical pollutants in the environment, such as pesticides, heavy metals, (?? main pollutant groups ??) belong after @rockstrom_safe_2009 to one of nine planetary boundary groups that should not exceed certain thresholds. Yet, the influence of chemical pollutants is not sufficiently studied to circumvent this boundary.

"Currently, pesticide use is wide-spread in agriculture all over the world, but still only very few countries have established wildlife poisoning surveillance programs (2). As a result, many data on pesticides remain scattered and/or
not publicly available" [@kohler_wildlife_2013]

what is really new and an improvement of the ECOTOX data base is the querying of regional and habitat data and automated checking!!

The EPA ECOTOX data base !citation! is a valuable resource for ecotoxicological test data. When it was last updated (June 7, 2018) it contained 910,465 test results of 11,583 different chemicals that were tested on 12,599 organisms.

# Methods

The EPA ECOTOX data base was downloaded (version: June 7, 2018) and built into a local PostgreSQL data base [@szoecs_build_2016]. The data base is subsequently queried for EC50 values

The whole process is automated and scheduled to be re-run four times a year.

Building the the data base is updated quaterly whenever a new version from the EPA can be downloaded. Based on the EPA ECOTOX data base CAS and taxa information are extracted and queried against other online available data bases to retrieve information which is lackign in the ECOTOX data base.

## Habitat data

## Regionl data

## Additional chemical parameters

water solubility

maybe also include degredation (water and soil) ?
information for the classification of chemicals

Data is stored at a [github repository](https://github.com/andreasLD/etox-base).

```{r organigram-conversion, echo=FALSE, include=TRUE, eval=FALSE}
# TODO can this be done at the end of the TiX file?
pdf_file = '/home/andreas/Documents/Projects/etox-base/TeX/organigram1.pdf'
png_file = '/home/andreas/Documents/Projects/etox-base/TeX/organigram1.png'

system(sprintf('convert -density 300 %s %s', pdf_file, png_file))
```

![Figure](/home/andreas/Documents/Projects/etox-base/TeX/organigram1.png)

# Discussion

quality of data in GBIF for instance in Africa is low. Continental refinement should be used with caution.

some ecotox test results might be flawed

Generally information on habitats is not easy to find. Only worms, algaebase and some others provide it, therefore own list.

it would be valuable to include other resources such as PPDB and algaebase. however they are proprietary.

# References

### 2018-10-10 article/article.Rmd
added text

@@
 ---
-title: "Etox-Base - A tool and data set in ecotoxicology"
+title: "Etox-Base - A tool and data set in ecotoxicology [max. 110 words]"
 author: "Andreas Scharmüller"
 date: "August 29, 2018"
-output: html_document
+output:
+  # bookdown::html_book:
+  # bookdown::pdf_document2:
+  bookdown::html_document2:
+  # html_document:
 bibliography: references-etox-base.bib
+editor_options: 
+  chunk_output_type: console
 ---
 
-# Abstract
+```{r setup, echo=FALSE}
+prj = '/home/andreas/Documents/Projects/etox-base'
+datadir = file.path(prj, 'article/data') # in sub folder article
+cachedir = file.path(prj, 'cache')
+```
 
-# Introduction
+# Abstract [max. 170 words]
+
+# Background & Summary (i.e. Introduction)
 
@@
 main pollutant groups ??) belong after @rockstrom_safe_2009 to one of nine planetary boundary groups that should not exceed certain thresholds.
-Yet, the influence of chemical pollutants is not sufficiently studied to circumvent this boundary.
+
+A vast amount of chemical substances such as pharmaceuticals, pesticides, synthetic hormones and many more are used all over the world with insufficient knowledge about possible adverse effects on the environment.
+In Europe alone some 100,000 substances are estimated to be in current use, whereof 30,000 are produced in quantities larger than one ton per year.
+Thereof less than one percent are thoroughly tested [@breithaupt_costs_2006].
+However, for some substance groups adverse effects could be detected.
+Pesticides for exmaple reduce macro-invertebrate biodiversity in streams [@beketov_pesticides_2013], estrogenic-associated effects in fish could be proven [@vethaak_integrated_2005] + HEAVY METALS.
+However for many substance groups, especially newly emerging ones authorative test results remain scarce.
+To overcome such knowledge paucities it is paramount to accumulate and distribute existing knowledge on possible adverse effects of substances widely.
+The ECOTOXicology knowledgebase (ECOTOX) created by U.S.
+Environmental Protection Agency (EPA) represents one such initiative [@elonen_ecotoxicology_2018].
+It collects ecotoxicological test data on on aquatic and terrestrial wildlife as well as plants and publishes it on a quarterly basis freely accessible on the web.
+By the time of writing it contained 919,123 test results, comprising 11,655 chemicla substances tested on 12,630 different species.
+Tests results that can exhibit over ten different endpoints such as Effective Concentrations - EC50 values, No-observed effect concentratons NOEC or lowest observed effect concentraitons - LOEC, where different effects such as Mortality, Intoxication, Growth etc.
+and test durations ranging from days to weeks can be measured.
+
+All this information can be queried through a web interface or the whole data base can be fully downloaded.
+Endpoints from ecotoxicological test data can be used to evaluate ocurrences of substances in the envrionment.
+Widely used endpoints are EC50s which determine the concentration where 50% of e.g.
+plant growth is inhibited, 50% of individuals have died etc.
+These can be realted to measured concnetrations by calculating Toxic Units (TUs) \@ref(eq:tu) [@schafer_effects_2011] [+other citations!!].
+
+\begin{equation*}
+TU = \frac{conc_i}{EC50}
+(\#eq:tu)
+\end{equation*}
+
+Test results vary naturally
+
+Effects of chemical substances vary with time [@sanchez-bayo_simple_2009].
+Therefore we propose to aggregate in a first step tests according to the tested compoud, the tested species and the test duration.
+Hereby if the number of tests is smaller or equal to two the minimum otherwise the median of the reported test concentrations is taken.
+Secondly this data can then be filtered \@ref(tab:table-filters) and aggregated by the user through the tool/app.
+Filters steps are applied before the aggregation.
+
+```{r table-filters, echo=FALSE}
+epa_stats_l = readRDS(file.path(cachedir, 'epa_stats_l.rds'))
+
+conc_type = epa_stats_l[['coty']]$conc_type
+auth = c('Africa', 'Asia', 'Europe', 'North America', 'South America', 'Australia')
+chem_type = c('metal', 'pesticide')
+
+filter_chem = data.frame(filters = c('Concentration type', 'Chemical class', 'Authorization'),
+                         options = c(paste0(conc_type, collapse = ', '),
+                                     paste0(chem_type, collapse = ', '),
+                                     paste0(auth, collapse = ', ')),
+                         explanations = 'link')
+
+habi = c('all', 'marine', 'brackish', 'freshwater', 'terrestrial')
+conti = 'bla'
+
+filter_taxa = data.frame(filters = c('Organism habitat', 'Continent'),
+                         options = paste0(habi, collapse = ', '))
+
+knitr::kable(filter_chem, caption = 'Chemical compound filter option')
+knitr::kable(filter_taxa, caption = 'Taxa filter options')
+```
+
+Filters according to concentration type (AI - Active ingredient, F-Formulation)
+
+However, due to the large number of tests it is not always trivial to determine single EC50 values for chemical substances.
+Therefore we propose a toolkit, data set that allows to choose between multiple aggregation methods.
+
+Similar data bases exist and one prominent example is the Pesticide Property Data Base (PPDB) [citation!] which is operated by the university of Head
+
+# Introduction IDEAS
+
+checkout citation in [@koehler_wildlife_2013]: Rohr 2008  
+"A field survey of the northern leopard
+frog, Rana pipiens, revealed that atrazine pol-
+lution and inorganic phosphate accounted for
+74% of the variation in the abundance of trem-
+atodes (62)."
 
@@
 not publicly available" [@kohler_wildlife_2013]
+
+"bias towards domestic and lab models" [@kohler_wildlife_2013]
+
+"Even though the banning of highly persistent or-
+ganochlorines in developed countries has shifted
+pesticide use toward a vast diversity of readily
+biodegradable ingredients, the explosiveness of
+organochlorines on a global scale cannot be ig-
+nored." [@koehler_wildlife_2013]
+
+-> half a milion tons in developing countries
+
+"to date 120 endocrine disruptive chemcials are known" [@koehler_wildlife_2013] 
+R. McKinlay, J. A. Plant, J. N. B. Bell, N. Voulvoulis,
+Environ. Int. 34, 168–183 (2008). covering numerous classes
+
+important effects: thyroid disruption
+
+"A recent meta-analysis revealed over-
+all environmental pollution to have large effects
+on abnormality frequencies but only medium
+effects on survival and no effects on time of de-
+velopment (29)." [@koehler_wildlife_2013]
+
+"Generally, single-
+chemical risk assessment will probably under-
+estimate the actual risks of pesticide mixtures to
+fish, as combinations of organophosphates and
+carbamates were shown to exert synergistic neu-
+rotoxicity and unpredicted mortality in Pacific
+salmon (35)." [@koehler_wildlife_2013]
+
+" So, despite all efforts to increase the
+specificity of insecticides, there is as yet no com-
+pound that both targets insect pests and leaves
+nontarget insects unaffected" [@koehler_wildlife_2013]
+
+köhler and triebskorn pesticide classification: where do they have it from?
 
@@
 
+there ain't no habitat data base on the web
+except gbif fuzzy coded data$habitat column
+
 # Methods
 
-The EPA ECOTOX data base was downloaded (version: June 7, 2018) and built into a local PostgreSQL data base [@szoecs_build_2016]. The data base is subsequently queried for EC50 values
+### EPA Ecotox data base
+The EPA ECOTOX data base was downloaded (version: June 7, 2018) and built into a local PostgreSQL data base [@szocs_build_2016].
+The data base is subsequently queried for __endpoints__ such as _half maximal effective concentrations - LC50, EC50, LD50, LT50, IC50, ED50 and ET50 (further denoted as EC50)_, _No-observed-adverse-effect level - NOEC, NOEL_ and _lowest observed concentrations - LOEC, LOEL_ [@elonen_ecotoxicology_2018].
+Data on other enpoints such as EC25 are ignored.
+Only the following __effect groups__ are considered: MOR|POP|ITX|GRO.
+Reported _compound concentrations_ are converted to µg/L and µg/kg respecitvely.
+
+### Ohter chemical data bases
+CAS numbers (e.g.
+210880-92-5) that are retrieved from the EPA ECOTOX data base are then used to query other publically available data bases with the help of the webchme R-package [@szocs_webchem_2015] to fetch additional chemical parameters \@ref(chem-refs).
+
+```{r chem-refs, echo=FALSE}
+chem_refs = read.table(file.path(datadir, 'chem_refs.txt'), sep = ',', header = TRUE)
+
+knitr::kable(
+  chem_refs,
+  caption = 'Chemical parameters'
+)
+```
+
+### Habitat data
+
+```{r habi-refs, echo=FALSE}
+habi_refs = NA
+
+knitr::kable(
+  habi_refs,
+  caption = 'Habitat parameters'
+)
+
+# From EPA:
+# media_type: SW, FW
+# habitat: Soil
+
+```
+
+The habitat parameters _isFre_, _isBra_, _isMar_, _isTer_ are assorted from data from different data bases such as the World Register of Marine Species (WoRMS) [@worms_editorial_board_world_2018] and the Global Biodiversity Information Facility (GBIF) [@gbif.org_gbif_2018].
+Moreover habitat types are also derived from EPA data entries on test media type [EPA SOURCE]
+Data from WoRMS are prefered as they are checked.
+
+* WORMS
+* algaebase?
+
+### Regional data
+
+Taxonomic data retrieved from the EPA ECOTOX data base is cleaned and only taxa that are determined until Genus level are kept.
+Subsequently taxa are queried using the rgbif-package [@chamberlain_rgbif_2018].
+
+__what about different chemical class columns?__
+
+__show overview plots or in suplimentary materials__
 
@@
 
-## Habitat data
+## Additional chemical parameters
 
-## Regionl data
+### Parameter table
 
-## Additional chemical parameters
+```{r parameter-table, echo=FALSE}
+# tests_data = readRDS(file.path(cachedir, 'tests_fl.rds'))
+# TODO read in tests data and list the parameters
+
+```
 
@@
 
+discuss problem of lacking data in GBIF on Africa. And Asia is way to big.
+
+## Classification of chemicals
+classification of chemical substances or compounds is not trivial.
+* Approaches
+  * Use-oriented:
+    * Pesticide classification according to degree of harzard based on the acute oral and dermal toxicity to rats [@international_program_on_chemical_safety_who_2010]
+  * Environment
+    * MOA
+  * Chemical classes
+    * ohp document - private company for selling pesticides
+
 quality of data in GBIF for instance in Africa is low. Continental refinement should be used with caution.
@@
 
+There is a need for a __habitat__ and __chemical classification__ data base!
+
+## Taxonomy:
+
+Dinoflagelates are considered autotroph, although mixotrophy in this taxon has been proven [@]stoecker_mixotrophy_1999
+
 # References

### 2018-10-23 TeX/organigram1.tex
added database and stats field to TikX picture

@@
 	\node (testres) [res, right of = cas, xshift = 14cm, yshift = -5cm] {Result};
-	\node (filtres) [res, right of = testres, xshift = 2cm] {Filter};
-	\node (aggres) [res, right of = filtres, xshift = 2cm] {Aggregate};
 
+	% shiny
 	\begin{pgfonlayer}{bg3}
-	\node[rectangle, draw, rounded corners, minimum height = 2cm, minimum width = 7cm, fill = gray!10, right of = testres, xshift = 1.5cm] (appbox) {};
+	\node[rectangle, draw, rounded corners, minimum height = 2cm, minimum width = 7cm, fill = gray!10, right of = testres, xshift = 1.5cm, yshift = 4cm] (appbox_shiny) {};
 	\end{pgfonlayer}
+	\node (apptext) [above of = appbox_shiny, yshift = 0.5cm, xshift = -3cm] {Etox-Base Function};
+	\node (filtres) [res, right of = appbox_shiny, xshift = -3.5cm] {Filter};
+	\node (aggres) [res, right of = appbox_shiny, xshift = 0cm] {Aggregate};
 
-	\node (apptext) [above of = appbox, yshift = 0.25cm, xshift = -3cm] {Etox-Base Function};
+	% database output
+	\begin{pgfonlayer}{bg3}
+	\node[rectangle, draw, rounded corners, minimum height = 2cm, minimum width = 7cm, fill = gray!10, right of = testres, xshift = 1.5cm, yshift = 0cm] (appbox_db) {};
+	\end{pgfonlayer}
+	\node (apptext) [above of = appbox_db, yshift = 0.5cm, xshift = -3cm] {Database output};
+	\node (tableres) [res, right of = appbox_db, xshift = -2cm] {Postgres Table};
+	% stats
+	\node (stats) [res, below of = tableres, yshift = -2cm] {re\_stats.R};
 
@@
 
-	\draw [arrow] (testres) -- (filtres); %++(2cm,0cm);
-	\draw [arrow] (filtres) -- (aggres);
+	% Result arrows
+	\draw [arrow] (testres) -- (appbox_shiny.west); %++(2cm,0cm);
+	\draw [arrow] (testres) -- (appbox_db.west);
 

### 2018-10-30 TeX/organigram1.tex
Updates from ShareLaTeX

@@
 
+	%%%%%%%% 1st part: downloading data %%%%%%%%
 	\node (io) [io, align = center] at (0cm,0cm) {EPA ECOTOX\\data base};
@@
 
+	%%%%%%%% 2nd part: processing chain %%%%%%%%
+	% nodes
+	\node (testres2) [res] at (0cm,-24cm) {Result re\_merge.R}; % maybe connect it with testres ?
+	\node (filter) [res, right of = testres2, xshift = 2cm] {re\_filter.R};
+	\node (analyses) [res, right of = filter, xshift = 2cm] {re\_analyses.R};
+	\node (checks) [res, right of = analyses, xshift = 2cm] {re\_checks.R};
+	\node (final) [res, right of = checks, xshift = 2cm] {re\_final.R};
+	    \node (stats) [res, right of = final, xshift = 2cm, yshift = 2cm] {re\_stats.R};
+	    \node (write) [res, right of = final, xshift = 2cm, yshift = 0cm] {re\_write.R};
+	    \node (shinyvar) [res, right of = final, xshift = 2cm, yshift = -2cm] {re\_shiny\_variables.R};
+
+	% arrows
+	\draw [arrow] (testres2.east) -- (filter.west);
+	\draw [arrow] (filter.east) -- (analyses.west);
+	\draw [arrow] (analyses.east) -- (checks.west);
+	\draw [arrow] (checks.east) -- (final.west);
+	\draw [arrow] (final.east) -- (stats.west);
+	\draw [arrow] (final.east) -- (write.west);
+	\draw [arrow] (final.east) -- (shinyvar.west);
+
 	\end{pgfonlayer}

### 2018-10-31 README.Rmd
new README

---
title: "Etox-Base"
author: "Andras Scharmüller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}

```

## Etox Base

Etox Base is a tool that aggregates ecotoxicological test data (EC50 values, NOEC etc.) from the [US EPA ECOTOTX data base](https://cfpub.epa.gov/ecotox/). Each new version of the EPA ECOTOX database which is released on a quarterly basis is downloaded and rebuilt locally. Besides the ecotoxicological test data therein, other databases on chemicals and organism parameters are then queried to extend and fill data gaps of the EPA ECOTOX database in order to allow for more adequat filter steps according to ecotoxicological needs. Chemical parameters on for example the classification of chemicals, water solubility of compounds amongst others are obtained through databases such as [Alan Wood's Compendium of Pesticide Common Names](http://www.alanwood.net/pesticides/index.html), the [PubChem](https://pubchem.ncbi.nlm.nih.gov/) database, the [PHYSPROP Database](https://www.srcinc.com/what-we-do/environmental/scientific-databases.html) and the [chemspider Database](http://www.chemspider.com/), all of which have publically available programming interfaces (APIs). Likewise databases on organisms such as the [World Refister of Marine Species (WORMS)](http://marinespecies.org/) and the [Global Biodiversity Information Facility (GBIF)](https://www.gbif.org/) are queried to obtain information on habitat preferences and organisms occurrence patterns. CAS-Numbers are used to retrieve the additional chemical inforamtion whereas taxa names are used to query organism parameters. These queries largely rely on R-packages such as [webchem](https://github.com/ropensci/webchem), [taxize](https://github.com/ropensci/taxize/) and [rgbif](https://github.com/ropensci/rgbif) besides some self-written functions that can be accesed in the two project repositories: [etox-base](https://github.com/andreasLD/etox-base) and [etox-base-shiny](https://github.com/andreasLD/etox-base).

The compiled data is than accessed through the Etox-Base application which filters it according to the users inputs. In the end the tool aggregates the many filtered test data to obtain a single value for a given set of test parameters.

An article on the project can be accessed [here]().

### 2018-10-31 R/run_build.R

@@ -24,6 +24,9 @@ if (nodename == 'scharmueller') {
 # (0) setup ---------------------------------------------------------------
 source(file.path(prj, 'R/setup.R'), max.deparse.length = 1e6)
 
+# copy files to shiny repo:
+file.copy('README.Rmd', file.path(shinydir, 'README.Rmd'),
+          overwrite = TRUE)
 
 # (0a) console log ---------------------------------------------------------
 if (nodename == 'uwigis') {

### 2018-11-30 article/article.Rmd
added scripts overview table

@@
   # bookdown::html_book:
-  # bookdown::pdf_document2:
-  bookdown::html_document2:
+  bookdown::pdf_document2:
+  # bookdown::html_document2:
   # html_document:
@@
 ```{r setup, echo=FALSE}
+# projectdir
 prj = '/home/andreas/Documents/Projects/etox-base'
+
+# packages
+require(data.table)
+
+# variables
 datadir = file.path(prj, 'article/data') # in sub folder article
@@
 
+\newcommand{\etoxbase}{Etox-Base}
+\newcommand{\epa}{EPA ECOTOX data base}
+\newcommand{\app}{http://139.14.20.252:3838/etox-base-shiny/}
+
 # Abstract [max. 170 words]
 
-# Background & Summary (i.e. Introduction)
+# Background & Summary (i.e. Introduction) [max 700 words]
 
-Chemical pollutants in the environment, such as pesticides, heavy metals, (??
-main pollutant groups ??) belong after @rockstrom_safe_2009 to one of nine planetary boundary groups that should not exceed certain thresholds.
-
-A vast amount of chemical substances such as pharmaceuticals, pesticides, synthetic hormones and many more are used all over the world with insufficient knowledge about possible adverse effects on the environment.
-In Europe alone some 100,000 substances are estimated to be in current use, whereof 30,000 are produced in quantities larger than one ton per year.
+A vast amount of chemical compounds such as pharmaceuticals, pesticides, synthetic hormones and many more are used all over the world with insufficient knowledge about possible adverse effects on the environment.
+In Europe alone some 100,000 compounds are estimated to be in current use, whereof 30,000 are produced in quantities larger than one ton per year.
 Thereof less than one percent are thoroughly tested [@breithaupt_costs_2006].
-However, for some substance groups adverse effects could be detected.
-Pesticides for exmaple reduce macro-invertebrate biodiversity in streams [@beketov_pesticides_2013], estrogenic-associated effects in fish could be proven [@vethaak_integrated_2005] + HEAVY METALS.
-However for many substance groups, especially newly emerging ones authorative test results remain scarce.
+For substances that have been tested adverse effects could be detected.
+Pesticides for exmaple reduce macro-invertebrate biodiversity in streams [@beketov_pesticides_2013] and estrogenic-associated effects were found in fish [@vethaak_integrated_2005].
+However for many substance groups, especially newly emerging ones authorative test results remain scarce or not hard to find.
 To overcome such knowledge paucities it is paramount to accumulate and distribute existing knowledge on possible adverse effects of substances widely.
 The ECOTOXicology knowledgebase (ECOTOX) created by U.S.
-Environmental Protection Agency (EPA) represents one such initiative [@elonen_ecotoxicology_2018].
+Environmental Protection Agency (EPA) represents one such initiative.
 It collects ecotoxicological test data on on aquatic and terrestrial wildlife as well as plants and publishes it on a quarterly basis freely accessible on the web.
-By the time of writing it contained 919,123 test results, comprising 11,655 chemicla substances tested on 12,630 different species.
-Tests results that can exhibit over ten different endpoints such as Effective Concentrations - EC50 values, No-observed effect concentratons NOEC or lowest observed effect concentraitons - LOEC, where different effects such as Mortality, Intoxication, Growth etc.
-and test durations ranging from days to weeks can be measured.
+By the time of writing it contained 919,123 test results, comprising 11,655 chemical compounds tested on 12,630 different species [@elonen_ecotoxicology_2018].
+The ecotoxicological tests are collected without limitation and comprise all sorts of measured endpoints such as Effective Concentrations - EC50 values, No-observed effect concentratons NOEC or lowest observed effect concentraitons - LOEC.
+Likewise the data set contains many different effect measures such as mortality, intoxication, growth etc.
+and test durations ranging from seconds to weeks or even years.
+These test results can be used to derive toxicity risk thresholds.
+As ecotoxicological risk assessment often relies on national and international quality thresholds such as EU-Environmtal quality standards (EQS) or Regulatory Aceptable concentrations (RAC) in Germany.
+In order to make this data more widely accessible we created a tool that downloads every new version of the \epa{} and performs several quality checks, e.g.
+unit harmonization, and cleaning steps on it.
+In addition to the test parameters that are retrieved from the \epa{} we query compound and organism parameters from other open data bases to adress missing information such as compound solubility or organism habitat or regional ocurrence patterns.
+Subsequently the data is provided online via a web app where users can filter the data according to needed test parameters.
+Furthermore the user can choose between several aggregation methods to retrieve individual values per compound.
+Compounds are identified by their CAS-number as this is also the key identifier in the \epa{}.
+Optionally the data base can also be downloaded as a whole for local useage.
+This is the first approach that tries to handle large amounts of toxicity test data to derive meaningful information about the susceptability of organisms towards individual compounds by aggregating test results.
+Researchers relying on ecotoxicological test data will benefit from the \etoxbase{} due to concise representation ecotoxicological test data.
 
-All this information can be queried through a web interface or the whole data base can be fully downloaded.
-Endpoints from ecotoxicological test data can be used to evaluate ocurrences of substances in the envrionment.
+# Methods [no word limit]
+
+The EPA releases \epa{} on a quarterly basis through their website.
+The software downloads each new version and rebuilds it locally in a PostgreSQL data base.
+A processing pipline of several R \cite{r_core_team_r_2017} scripts prepares the data for the tool.
+The tool itself is a R function that is executed by a user acessing the R shiny \citep{chang_shiny_2018} application.
+In the following the steps for processing the data are explained in more detail (PATH_TO_OVERVIEW) as well as the scripts that have been used for it (REFER_TO_SCRIPTS_TABLE).
+
+## Downloading and preparing the data
+
+The data is downloaded (bd_epa_download.R) and subsequently built into a PostgreSQL data base (bd_epa_postgres.R).
+Thereafter the data is imported into R for preparing it (da_epa1.R, da_epa2.R, da_epa3.R), by removing special characters type conversions, unit harmonizations and by reducing the columns.
+At the end the data is stored locally (epa3.rds, epa2_taxa.rds, epa2_chem.rds).
+Obtained CAS numbers and taxon names are then used to query other data bases for additional information on compounds and biota respectively.
+Compound information is retrieved from the Pubchem data base \citep{CITE_PUBCHEM} (qu_pc.R), from the Compendium of Pesticide Common Names \citep{CITE_AW} (qu_aw.R), from the Physprop data base \citep{CITE_PHYSPROP} (qu_pp.R), from EUROSTAT and the Chemspider data base \citep(CITE_CHEMSPIDER).
+These queries mostly rely on the webchem R-package \citep{szocs_webchem_2015-1}.
+Additional habitat and occurrence information for organisms is queried from the World Register of Marine Species (WORMS) \citep{WORMS} and from the Global Biodiversity Information Facility (GBIF) \citep{CITE_RGBIF} using the rgbif R-package.
+The acquired information is then merged (re_merge.R), variables are combined and created (re_combine.R), checked (re_checks_internal.R) and finally compiled in one final table (re_final.R) This table is then used in the application to perform search queries on.
+For a detailed overview of what information is collected see table (REFERNCE_TO_VARIABLE_TABLE)
+
+## The Application
+The application is accesible through \app{}
+
+### functionalities
+
+###
+
+# Supporting information
+
+Tables:
+
+(1) script-table
+(2) varaiable-table
+
+# OLD
+
+Due to its size allowing for a more broad compound assessment.
+
 Widely used endpoints are EC50s which determine the concentration where 50% of e.g.
@@
 
-### EPA Ecotox data base
-The EPA ECOTOX data base was downloaded (version: June 7, 2018) and built into a local PostgreSQL data base [@szocs_build_2016].
-The data base is subsequently queried for __endpoints__ such as _half maximal effective concentrations - LC50, EC50, LD50, LT50, IC50, ED50 and ET50 (further denoted as EC50)_, _No-observed-adverse-effect level - NOEC, NOEL_ and _lowest observed concentrations - LOEC, LOEL_ [@elonen_ecotoxicology_2018].
+For all concentration columns patterns like + and * are deleted
+
+### EPA ECOTOX data
+In a first step the latest version of the EPA ECOTOX data base was downloaded (version: June 7, 2018) and built into a local PostgreSQL data base [@szocs_build_2016].
+The data base was subsequently queried for __endpoints__ such as _half maximal effective concentrations - LC50, EC50, LD50, LT50, IC50, ED50 and ET50 (further denoted as EC50)_, _No-observed-adverse-effect level - NOEC, NOEL_ and _lowest observed concentrations - LOEC, LOEL_ [@elonen_ecotoxicology_2018].
 Data on other enpoints such as EC25 are ignored.
-Only the following __effect groups__ are considered: MOR|POP|ITX|GRO.
-Reported _compound concentrations_ are converted to µg/L and µg/kg respecitvely.
+Reported _compound concentrations_ are converted to µg/L and µg/kg respecitvely whenever possible.
 
-### Ohter chemical data bases
-CAS numbers (e.g.
-210880-92-5) that are retrieved from the EPA ECOTOX data base are then used to query other publically available data bases with the help of the webchme R-package [@szocs_webchem_2015] to fetch additional chemical parameters \@ref(chem-refs).
+### Other data sources
+In order to retrieve additional __chemical parameters__, CAS numbers (e.g.
+210880-92-5) that are retrieved from the EPA ECOTOX data base are then used to query other publically available data bases with the help of the webchme R-package [@szocs_webchem_2015] or directly through respective APIs.
+Likewise __habitat parameters__ are retrieved from online database such as the World Register of Marine Species (WoRMS) [@worms_editorial_board_world_2018] and the Global Biodiversity Information Facility (GBIF) [@gbif.org_gbif_2018].
+To some extent the EPA ECOTOX data base itself also contains information on organism hatbitats [@elonen_ecotoxicology_2018].
+__Regional parameters__, that consitute natural occurrences of tested organisms are retrieved from GBIF.
+Original names and newly assigned names of the parameters are listed in \@ref(meta-refs).
 
-```{r chem-refs, echo=FALSE}
-chem_refs = read.table(file.path(datadir, 'chem_refs.txt'), sep = ',', header = TRUE)
-
-knitr::kable(
-  chem_refs,
-  caption = 'Chemical parameters'
-)
+```{r meta-stats, echo=FALSE}
+meta = fread(file.path(cachedir, 'meta.csv'))
+knitr::kable(meta)
 ```
-
-### Habitat data
-
-```{r habi-refs, echo=FALSE}
-habi_refs = NA
-
-knitr::kable(
-  habi_refs,
-  caption = 'Habitat parameters'
-)
-
-# From EPA:
-# media_type: SW, FW
-# habitat: Soil
-
-```
-
-The habitat parameters _isFre_, _isBra_, _isMar_, _isTer_ are assorted from data from different data bases such as the World Register of Marine Species (WoRMS) [@worms_editorial_board_world_2018] and the Global Biodiversity Information Facility (GBIF) [@gbif.org_gbif_2018].
-Moreover habitat types are also derived from EPA data entries on test media type [EPA SOURCE]
-Data from WoRMS are prefered as they are checked.
-
-* WORMS
-* algaebase?
 
@@
 
-## Additional chemical parameters
-
-### Parameter table
-
-```{r parameter-table, echo=FALSE}
-# tests_data = readRDS(file.path(cachedir, 'tests_fl.rds'))
-# TODO read in tests data and list the parameters
-
-```
-
 water solubility
@@
 information for the classification of chemicals
-
-Data is stored at a [github repository](https://github.com/andreasLD/etox-base).
 

### 2018-12-04 TeX/organigram1.tex

@@
 	main/.style = {rectangle, draw = black, rounded corners, minimum width = 3cm, minimum height = 1.25cm, text centered},
+	etox-base/.style = {rectangle, rounded corners},
+	etox-base-shiny/.style = {circle},
 	ids/.style = {diamond, minimum width = 3cm, minimum height = 2cm, text centered, draw = black, fill = green!30},
@@
 	\node (io) [io, align = center] at (0cm,0cm) {EPA ECOTOX\\data base};
+	\node (chemprop) [io, dashed, right of=io, xshift=2cm] {CHEMPROP};
+	\node (etox) [io, dashed, right of=chemprop, xshift=2cm] {UBA etox};
 	%% chemicals
 	% id nodes
-
 	\node (cas) [ids, below of=io, yshift = -4cm, xshift = 3cm] {CAS};
-
 	% main nodes
-	\node (chemprop) [main, dashed, right of=cas, xshift=4.5cm, yshift=1.5cm] {CHEMPROP};
-	\node (chemscr) [main, right of=cas, yshift=-1cm, xshift=4.5cm] {PHYS-CHEM}
+	\node (chemscr) [main, right of=cas, xshift=4.5cm] {PHYS-CHEM}
 	child { node { Alan Wood Pesticide Compendium - \textbf{qu\_aw.R } }}
@@
 	\node (apptext) [above of = appbox_shiny, yshift = 0.5cm, xshift = -3cm] {Etox-Base Function};
-	\node (filtres) [res, right of = appbox_shiny, xshift = -3.5cm] {Filter};
+	\node (filtres) [res, right of = appbox_shiny, xshift = -3.5cm] {Combine};
 	\node (aggres) [res, right of = appbox_shiny, xshift = 0cm] {Aggregate};
@@
 	\draw [arrow, black] (io) |- (cas.west);
-	\draw [arrow, dashed] (cas.east) -- (chemprop.west);
 	\draw [arrow] (cas.east) -- (chemscr.west);
@@
 	\node (testres2) [res] at (0cm,-24cm) {Result re\_merge.R}; % maybe connect it with testres ?
-	\node (filter) [res, right of = testres2, xshift = 2cm] {re\_filter.R};
-	\node (analyses) [res, right of = filter, xshift = 2cm] {re\_analyses.R};
+	\node (combine) [res, right of = testres2, xshift = 2cm] {re\_combine.R};
+	\node (analyses) [res, right of = combine, xshift = 2cm] {re\_analyses.R};
 	\node (checks) [res, right of = analyses, xshift = 2cm] {re\_checks.R};
@@
 	    \node (stats) [res, right of = final, xshift = 2cm, yshift = 2cm] {re\_stats.R};
+	    \node (meta) [res, right of = final, xshift = 2cm, yshift = 4cm] {re\_meta.R};
 	    \node (write) [res, right of = final, xshift = 2cm, yshift = 0cm] {re\_write.R};
@@
 	% arrows
-	\draw [arrow] (testres2.east) -- (filter.west);
-	\draw [arrow] (filter.east) -- (analyses.west);
+	\draw [arrow] (testres2.east) -- (combine.west);
+	\draw [arrow] (combine.east) -- (analyses.west);
 	\draw [arrow] (analyses.east) -- (checks.west);
@@
 	\draw [arrow] (final.east) -- (stats.west);
+	\draw [arrow] (final.east) -- (meta.west);
 	\draw [arrow] (final.east) -- (write.west);

### 2018-12-04 article/article.tex
added .tex article

\documentclass[english]{article}
\usepackage{natbib}
%\usepackage{pgfplotstable, filecontents}
%\usepackage{csvsimple}

\newcommand{\etoxbase}{Etox-Base}
\newcommand{\epa}{EPA ECOTOX data base}
\newcommand{\app}{http://139.14.20.252:3838/etox-base-shiny/}
\newcommand{\git}{https://github.com/andreasLD/etox-base}

\listfiles

\begin{document}

\title{Etox-Base - a tool for ecotoxicologists}

\begin{abstract}

\end{abstract}

\section*{Background \& Summary [max 700 words]}

A vast amount of chemical compounds such as pharmaceuticals, pesticides, synthetic hormones and many more are used all over the world with insufficient knowledge about possible adverse effects on the environment. In Europe alone some 100,000 compounds are estimated to be in current use, whereof 30,000 are produced in quantities larger than one ton per year. Thereof less than one percent are thoroughly tested \citep{breithaupt_costs_2006}. For substances that have been tested adverse effects could be detected. Pesticides for exmaple reduce macro-invertebrate biodiversity in streams \citep{beketov_pesticides_2013} and estrogenic-associated effects were found in fish \citep{vethaak_integrated_2005}. However for many substance groups, especially newly emerging ones authorative test results remain scarce or not hard to find. To overcome such knowledge paucities it is paramount to accumulate and distribute existing knowledge on possible adverse effects of substances widely. The ECOTOXicology knowledgebase (ECOTOX) created by U.S. Environmental Protection Agency (EPA) represents one such initiative. It collects ecotoxicological test data on on aquatic and terrestrial wildlife as well as plants and publishes it on a quarterly basis freely accessible on the web. By the time of writing it contained 919,123 test results, comprising 11,655 chemical compounds tested on 12,630 different species \citep{elonen_ecotoxicology_2018}. The ecotoxicological tests are collected without limitation and comprise all sorts of measured endpoints such as Effective Concentrations - EC50 values, No-observed effect concentratons NOEC or lowest observed effect concentraitons - LOEC. Likewise the data set contains many different effect measures such as mortality, intoxication, growth etc. and test durations ranging from seconds to weeks or even years. These test results can be used to derive toxicity risk thresholds. As ecotoxicological risk assessment often relies on national and international quality thresholds such as EU-Environmtal quality standards (EQS) or Regulatory Aceptable concentrations (RAC) in Germany. In order to make this data more widely accessible we created a tool that downloads every new version of the \epa{} and performs several quality checks, e.g. unit harmonization, and cleaning steps on it. In addition to the test parameters that are retrieved from the \epa{} we query compound and organism parameters from other open data bases to adress missing information such as compound solubility or organism habitat or regional ocurrence patterns. Subsequently the data is provided online via a web app where users can filter the data according to needed test parameters. Furthermore the user can choose between several aggregation methods to retrieve individual values per compound. Compounds are identified by their CAS-number as this is also the key identifier in the \epa{}. Optionally the data base can also be downloaded as a whole for local useage. This is the first approach that tries to handle large amounts of toxicity test data to derive meaningful information about the susceptability of organisms towards individual compounds by aggregating test results. Researchers relying on ecotoxicological test data will benefit from the \etoxbase{} due to concise representation ecotoxicological test data.

\section*{Methods}
%%The Methods should include detailed text describing any steps or procedures 
%%used in producing the data, including full descriptions of the experimental 
%%design, data acquisition assays, and any computational processing (e.g. 
%%normalization, image feature extraction). Related methods should be grouped 
%%under corresponding subheadings where possible, and methods should be described 
%%in enough detail to allow other researchers to interpret and repeat, if required, 
%%the full study. Specific data outputs should be explicitly referenced via data 
%%citation (see Data Records and Data Citations, below). Authors should cite 
%%previous descriptions of the methods under use, but ideally the method 
%%descriptions should be complete enough for others to understand and reproduce 
%%the methods and processing steps without referring to associated publications. 
%%There is no limit to the length of the Methods section.

The EPA releases \epa{} on a quarterly basis through their website. The software downloads each new version and rebuilds it locally in a PostgreSQL data base. A processing pipline of several R \cite{r_core_team_r_2017} scripts prepares the data for the tool. The tool itself is a R function that is executed by a user acessing the R shiny \citep{chang_shiny_2018} application. In the following the steps for processing the data are explained in more detail (PATH-TO-OVERVIEW) as well as the scripts that have been used for it (REFER-TO-SCRIPTS-TABLE).

\subsection*{Data acquisition \& preparation}

The data is downloaded (bd\_epa\_download.R) and subsequently built into a PostgreSQL data base (bd\_epa\_postgres.R). Thereafter the data is imported into R for preparing it (da\_epa1.R, da\_epa2.R, da\_epa3.R), by removing special characters type conversions, unit harmonizations and by reducing the columns. At the end the data is stored locally (epa3.rds, epa2\_taxa.rds, epa2\_chem.rds). Obtained CAS numbers and taxon names are then used to query other data bases for additional information on compounds and biota respectively. Compound information is retrieved from the Pubchem data base \citep{CITE_PUBCHEM} (qu\_pc.R), from the Compendium of Pesticide Common Names \citep{CITE_AW} (qu\_aw.R), from the Physprop data base \citep{CITE_PHYSPROP} (qu\_pp.R), from EUROSTAT and the Chemspider data base \citep(CITE-CHEMSPIDER). These queries mostly rely on the webchem R-package \citep{szocs_webchem_2015-1}. Additional habitat and occurrence information for organisms is queried from the World Register of Marine Species (WORMS) \citep{WORMS} and from the Global Biodiversity Information Facility (GBIF) \citep{CITE_RGBIF} using the rgbif R-package. The acquired information is then merged (re\_merge.R), variables are combined and created (re\_combine.R), checked (re\_checks\_internal.R) and finally compiled in one final table (re\_final.R) This table is then used in the application to perform search queries on. For a detailed overview of what information is collected see table (REFERNCE-TO-VARIABLE-TABLE)

\subsection*{The application}
The application is accessible through \app{} and consitutes

%
%\begin{filecontents*}{data/scripts.csv}
%	
%\end{filecontents*}

%\input{data/scripts.csv}

%
%\begin{tabular}{l|c}
%	
%	\csvreader[head to column names]{data/scripts.csv}{}
%	\end{tabular}

\subsection*{Code availability}
%%For all studies using custom code in the generation or processing of datasets, 
%%a statement must be included here, indicating whether and how the code can be 
%%accessed, including any restrictions to access. This section should also include 
%%information on the versions of any software used, if relevant, and any specific 
%%variables or parameters used to generate, test, or process the current dataset.

The code for processing the data is stored in a Github reopsitory (\git{})

\section*{Data Records}
%%Please explain each data record associated with this work, including
%%the repository where this information is stored, and an overview of
%%the data files and their formats. Each external data record should
%%be listed in Data Citation section at the end of this template, and 
%%records should be cited throughout the manuscript as, for example 
%%(Data Citation 1). 
%%
%%Tables should be used to support the data records, and should clearly indicate 
%%the samples and subjects, their provenance, and the experimental manipulations 
%%performed on each. They should also specify the data output resulting from each 
%%data-collection or analytical step, should these form part of the archived record. 
%%Please see the submission guidelines at the \emph{Scientific Data} website, and 
%%our Word templates for more information on preparing such tables.

\section*{Technical Validation}
%%This section presents any experiments or analyses that are needed
%%to support the technical quality of the dataset. This section may
%%be supported by up figures and tables, as needed. This is a required
%%section; authors must present information justifying the reliability
%%of their data.

\section*{Usage Notes}
%%Brief instructions that may help other researchers reuse these dataset.
%%This is an optional section, but strongly encouraged when helpful
%%to readers. This may include discussion of software packages that
%%are suitable for analyzing the assay data files, suggested downstream
%%processing steps (e.g. normalization, etc.), or tips for integrating
%%or comparing this with other datasets. If needed, authors are encouraged
%%to provide code, programs, or data processing workflows when they may help 
%%others analyse the data. We encourage authors to archive related code in 
%%a DOI-issuing archive when possible, but code may also be supplied as 
%%supplementary information files. 
%%
%%For studies involving privacy or safety controls on public access
%%to the data, this section should describe in detail these controls,
%%including how authors can apply to access the data, and what criteria
%%will be used to determine who may access the data, and any limitations
%%on data use.

\section*{Acknowledgements}
%%Text acknowledging non-author contributors. Acknowledgements should
%%be brief, and should not include thanks to anonymous referees and
%%editors, or effusive comments. Grant or contribution numbers may be
%%acknowledged. Author contributions Please describe briefly the contributions
%%of each author to this work on a separate line. 
%%
%%AK did this and that. 
%%
%%BG did this and that and the other.

\section*{Competing financial interests}
%%A competing financial interests statement is required for all accepted
%%papers published in \emph{Scientific Data}. If none exist simply write,
%%``The author(s) declare no competing financial interests''.

\section*{Figures Legends}
%%Figure should be referred to using a consistent numbering scheme through
%%the entire Data Descriptor. For initial submissions, authors may choose
%%to supply this document as a single PDF with embedded figures, but
%%separate figure image files must be provided for revisions and accepted
%%manuscripts. In most cases, a Data Descriptor should not contain more
%%than three figures, but more may be allowed when needed. We discourage
%%the inclusion of figures in the Supplementary Information \textendash{}
%%all key figures should be included here in the main Figure section. 
%%
%%Figure legends begin with a brief title sentence for the whole figure
%%and continue with a short description of what is shown in each panel,
%%as well as explaining any symbols used. Legend must total no more
%%than 350 words, and may contain literature references.

\section*{Tables}
%%Tables supporting the Data Descriptor. These can provide summary information
%%(sample numbers, demographics, etc.), but they should generally not
%%be used to present primary data (i.e. measurements). Tables containing
%%primary data should be submitted to an appropriate data repository. 
%%
%%Tables may be provided within the \LaTeX{} document or as separate
%%files (tab-delimited text or Excel files). Legends, where needed,
%%should be included here. Generally, a Data Descriptor should have
%%fewer than ten Tables, but more may be allowed when needed. Tables
%%may be of any size, but only Tables which fit onto a single printed
%%page will be included in the PDF version of the article (up to a maximum
%%of three).

\bibliography{references-etox-base}

\end{document}

### 2018-12-11 article/article.tex

@@
-\documentclass[english]{article}
+d\documentclass[english]{article}
 \usepackage{natbib}
-%\usepackage{pgfplotstable, filecontents}
-%\usepackage{csvsimple}
+\usepackage{csvsimple}
+\usepackage{tikz}
+\usetikzlibrary{trees, arrows, shapes.geometric, positioning}
+\usepackage{standalone}
+%% text packages
+\usepackage{siunitx} % fake ° symbol \ang{}
+\usepackage[super]{nth} % for 1st, 2nd etc. \nth{}
 
@@
 \newcommand{\git}{https://github.com/andreasLD/etox-base}
+\newcommand{\gitapp}{https://github.com/andreasLD/etox-base-shiny}
 
@@
 
-A vast amount of chemical compounds such as pharmaceuticals, pesticides, synthetic hormones and many more are used all over the world with insufficient knowledge about possible adverse effects on the environment.
-In Europe alone some 100,000 compounds are estimated to be in current use, whereof 30,000 are produced in quantities larger than one ton per year.
+A large number of chemical compounds such as pharmaceuticals, pesticides and synthetic hormones are in daily use all over the world with insufficient knowledge about possible adverse effects on the environment.
+In Europe alone some 100,000 compounds are estimated to be in current use, whereof 30,000 are produced in quantities larger than one ton per year \citep{breithaupt_costs_2006}.
+Chemical compounds are brought to the environment deliberately - as it is the case for pesticides or arrive in the environment as a byproduct from other processes (e.g.
+as atmospheric emissions or as wastewater) \citep{schwarzenbach_challenge_2006}.
+Ecosystems provide essential services to human societies such as drinking and irrigation water, food and climate regulation.
+These ecosystem services are products of different ecosystem functions that crucially depend on the integrity of the populations and communities that drive these ecosystem functions.
+
+Pollution with man-made chemical toxicants was identified as one of three major environmental problems for which research gaps hamper the derivation of planetary boundaries, i.e.
+thresholds beyond which irreversible state shifts may occur { | Steffen, et al., 2015 | | |zu:0:4565WXWV}.
+
+# Überleitung Biodiversitätsverlust – Rolle von Chemikalien
+Hier ein Text von mir von einem internen Antrag, vielleicht kannst Du davon was recyclen:
+Ecosystems provide essential services to human societies such as drinking and irrigation water, food and climate regulation.
+These ecosystem services are products of different ecosystem functions that crucially depend on the integrity of the populations and communities that drive these ecosystem functions.
+Humans are altering ecosystems from regional (e.g.
+through land use change) to global (e.g.
+through ocean acidification) scales, which has been captured in defining the current era the Anthropocene { | Crutzen, 2002 | | |zu:0:QMD3ZN8K}.
+The anthropogenic alteration of ecosystems is associated with an accelerating rate of species extinctions { | Butchart, et al., 2010 | | |zu:0:PAKGEKAH}{ | Barnosky, et al., 2011 | | |zu:0:5WZP7PB6} that threaten the provisioning of the abovementioned services { | MEA, 2005 | | |zu:0:7MXCXGXS}{ | Barnosky, et al., 2014 | | |zu:0:DBW4NPGX}.
+The major anthropogenic stressors that have caused the deterioration of ecosystems include habitat degradation, climate change, nutrient enrichment, and chemical pollution (e.g.
+pesticides).
+Several of these stressors, such as climate change and nutrient enrichment, are close to or are already exceeding their planetary boundaries { | Steffen, et al., 2015 | | |zu:0:4565WXWV}, which may result in irreversible shifts in the biosphere { | Barnosky, et al., 2012 | | |zu:0:SXKWUBMG}.
+Pollution with man-made chemical toxicants was identified as one of three major environmental problems for which research gaps hamper the derivation of planetary boundaries, i.e.
+thresholds beyond which irreversible state shifts may occur { | Steffen, et al., 2015 | | |zu:0:4565WXWV}.
+Bernhardt et al.
+{ | Bernhardt, et al., 2017 | | |zu:0:MCKMS8AZ} argue that the knowledge gap how chemicals affect biodiversity and ecosystem functions, and in turn ecosystem services, would also impede society’s ability to accomplish the Sustainable Development Goals of the United Nations.
+Hence, research on the effects of chemicals on ecosystems is required for establishing thresholds ensuring ecosystem integrity and enabling societies to effectively manage chemical risks.
+From the perspective of basic science, the scope of such a research program also involves basic scientific questions such as to which extent functions (e.g.
+nutrients and carbon cycling) and services are interdependent in their response to chemical stress, and to what extent non-linear feedbacks, as a potential consequence of transgenerational effects (such as induced mutation, population bottlenecks and selection ), occur.
+
+Biodiversitätsverlust - Rolle von chemikalien?
+
 Thereof less than one percent are thoroughly tested \citep{breithaupt_costs_2006}.
@@
 \section*{Methods}
+
 %%The Methods should include detailed text describing any steps or procedures 
@@
 The acquired information is then merged (re\_merge.R), variables are combined and created (re\_combine.R), checked (re\_checks\_internal.R) and finally compiled in one final table (re\_final.R) This table is then used in the application to perform search queries on.
-For a detailed overview of what information is collected see table (REFERNCE-TO-VARIABLE-TABLE)
+For a detailed overview of what information is collected see table \ref{table:processing-scripts}.
 
 \subsection*{The application}
-The application is accessible through \app{} and consitutes
+The application is accessible through \etox-base{} and was built by using the shiny web application framework \citep{chang_shiny_2018} in R.
+Therein the user can choose to filter the data by several parameters such as chemical class, organism habitat, test duration and test endpoints amongst others.
+A detailed list with example entries of the possible filter parameters can be found in \ref{table:meta}.
+In order to derive single endpoints for an organism groups and a compound of the partly extensive test data, the \etox-base{} aggregates the test results according to the chosen filters in two steps.
+Firstly the filtered test results are aggregated by the CAS number, the chosen taxon plus the duration and secondly only by the CAS number.
+The former can't be influenced by the user and calculates either the minimum or the median depending on the amount of results to aggregate (n <= 2: minimum, if n > 2: median).
+Thereof the second step calculates the minimum, the maximum, the median, the mean or the geometric mean as aggregate.
+Besides the aggregation some cleaning can also be performed.
+The \etox-base{} allows to exclude test results exhibiting concentrations that are higher than the actual water solubility of the respective compound at \ang{20} C.
+In any case the user can choose to exclude outliers.
+The outliers are selected by excluding values that exceed lower (0.25) and the upper (0.75) quartile by 1.5 times the interquartile range.
+All the steps described in this section are executed at each click in the app which runs the function fun\_ec50filter\_aggregation.R (cf.
+table \ref{table:processing-scripts}).
+Fast, adequate and live data manipulation is possible thanks to the data.table package \citep{dowle_data.table_2018}.
 
-%
-%\begin{filecontents*}{data/scripts.csv}
-%	
-%\end{filecontents*}
+\subsection*{Future of the project}
+As new versions of the \epa{} are published on a quarterly basis the data acquisition process is automated and new versions of the data and the \app{} are planed to be released regularly.
 
-%\input{data/scripts.csv}
-
-%
-%\begin{tabular}{l|c}
-%	
-%	\csvreader[head to column names]{data/scripts.csv}{}
-%	\end{tabular}
+\begin{figure}
+    %\includestandalone[width=\textwidth]{article/tikz/organigram}
+    \input{article/tikz/organigram}
+    \caption{Organigram}
+    \label{fig:organigram}
+\end{figure}
 
@@
 
-The code for processing the data is stored in a Github reopsitory (\git{})
+The code for processing the data is stored in a Github reopsitory (\git{}).
+All processing scripts were written in R version 3.4.4.
+PostgreSQL 9.5 has been used to build the data base.
+Hence R together with its packages and PostgreSQL suffice to rebuild the process.
+The \app{} itself is stored in a second Github repostitory (\gitapp{}).
 
@@
 %%on data use.
+This data set is designed to support ecotoxicologists in assessing the risk of chemical compounds on the environment.
+As there is more and more ecotoxicological test data available, it becomes fundamental to provide such information in reasonable formats.
+This means, that it should be accessible for humans as well as computers.
+Up to now ecotoxicoloists rely on effect values of single publications or data bases such as the Pesticide Property Data Base (PPDB), having certain limitations such as encompassing only a few substance groups.
+The \epa{} tries to fill these gaps by collecting in peer-review articles published ecotoxicological test results.
+However due to lacking test information and un-harmonized data sets it is not always easily accessible.
+The \app{} tries to mitigate this gap.
+An ecotoxicologist can now access the \app{} and retrieve cleaned, harmonized and aggregated test data according to his or her needs.
+
+\subsection*{Limitations}
+Due to missing entries,...
 
@@
 
+\begin{table}
+    \csvautotabular[respect all]{data/scripts.csv}
+    \caption{Processing scripts}
+    \label{table:processing-scripts}
+\end{table}
+
+\section*{Figure Legends}
+
+\begin{thebibliography}{}
 \bibliography{references-etox-base}
+\end{thebibliography}
+
+\section*{Data Citations}
 

### 2018-12-17 article/article.tex
changes to tikz picture and article

@@
-d\documentclass[english]{article}
-\usepackage{natbib}
+\documentclass[english]{article}
+\usepackage{natbib} % citations
 \usepackage{csvsimple}
+\usepackage{rotating} % to rotate tables
+%% tikz packages
 \usepackage{tikz}
@@
 These ecosystem services are products of different ecosystem functions that crucially depend on the integrity of the populations and communities that drive these ecosystem functions.
+Pollution with man-made chemical toxicants threatens ecosystems in various ways which are currently not fully understood \citep{steffen_anthropocene_2007}.
+Besides habitat degradation, climate change and nutrient enrichment, chemical pollution poses a significant threat to earth's biodiversity
 
 Pollution with man-made chemical toxicants was identified as one of three major environmental problems for which research gaps hamper the derivation of planetary boundaries, i.e.
-thresholds beyond which irreversible state shifts may occur { | Steffen, et al., 2015 | | |zu:0:4565WXWV}.
+thresholds beyond which irreversible state shifts may occur \citep{steffen_anthropocene_2007}.
+Bernhardt et al.
+\citet{bernhardt_synthetic_2017} argue that the knowledge gap how chemicals affect biodiversity and ecosystem functions, and in turn ecosystem services, would also impede society’s ability to accomplish the Sustainable Development Goals of the United Nations.
+Hence, research on the effects of chemicals on ecosystems is required for establishing thresholds ensuring ecosystem integrity and enabling societies to effectively manage chemical risks.
 
-# Überleitung Biodiversitätsverlust – Rolle von Chemikalien
-Hier ein Text von mir von einem internen Antrag, vielleicht kannst Du davon was recyclen:
-Ecosystems provide essential services to human societies such as drinking and irrigation water, food and climate regulation.
-These ecosystem services are products of different ecosystem functions that crucially depend on the integrity of the populations and communities that drive these ecosystem functions.
-Humans are altering ecosystems from regional (e.g.
-through land use change) to global (e.g.
-through ocean acidification) scales, which has been captured in defining the current era the Anthropocene { | Crutzen, 2002 | | |zu:0:QMD3ZN8K}.
-The anthropogenic alteration of ecosystems is associated with an accelerating rate of species extinctions { | Butchart, et al., 2010 | | |zu:0:PAKGEKAH}{ | Barnosky, et al., 2011 | | |zu:0:5WZP7PB6} that threaten the provisioning of the abovementioned services { | MEA, 2005 | | |zu:0:7MXCXGXS}{ | Barnosky, et al., 2014 | | |zu:0:DBW4NPGX}.
-The major anthropogenic stressors that have caused the deterioration of ecosystems include habitat degradation, climate change, nutrient enrichment, and chemical pollution (e.g.
-pesticides).
-Several of these stressors, such as climate change and nutrient enrichment, are close to or are already exceeding their planetary boundaries { | Steffen, et al., 2015 | | |zu:0:4565WXWV}, which may result in irreversible shifts in the biosphere { | Barnosky, et al., 2012 | | |zu:0:SXKWUBMG}.
-Pollution with man-made chemical toxicants was identified as one of three major environmental problems for which research gaps hamper the derivation of planetary boundaries, i.e.
-thresholds beyond which irreversible state shifts may occur { | Steffen, et al., 2015 | | |zu:0:4565WXWV}.
-Bernhardt et al.
-{ | Bernhardt, et al., 2017 | | |zu:0:MCKMS8AZ} argue that the knowledge gap how chemicals affect biodiversity and ecosystem functions, and in turn ecosystem services, would also impede society’s ability to accomplish the Sustainable Development Goals of the United Nations.
-Hence, research on the effects of chemicals on ecosystems is required for establishing thresholds ensuring ecosystem integrity and enabling societies to effectively manage chemical risks.
-From the perspective of basic science, the scope of such a research program also involves basic scientific questions such as to which extent functions (e.g.
-nutrients and carbon cycling) and services are interdependent in their response to chemical stress, and to what extent non-linear feedbacks, as a potential consequence of transgenerational effects (such as induced mutation, population bottlenecks and selection ), occur.
+plan:
+- occurring in vast numbers
+- pathways
+- ecosystem services 
+- biodiversity loss
+- Übergang abschätzung von chemikalien
 
-Biodiversitätsverlust - Rolle von chemikalien?
-
-Thereof less than one percent are thoroughly tested \citep{breithaupt_costs_2006}.
+However, less than one percent are thoroughly tested \citep{breithaupt_costs_2006}.
 For substances that have been tested adverse effects could be detected.
@@
 Researchers relying on ecotoxicological test data will benefit from the \etoxbase{} due to concise representation ecotoxicological test data.
+
+\citep{kefford_definition_2011} % paper that shows SSD species richness approximates effect of salinity on macroinvertebrates
+
+\citep{posthuma_species_2002} % describes concept of SSDs first
+
+two prominent concepts to assess effects on organisms or biodiversity in ecotoxicology are Species Sensitivity Distributions \citep{posthuma_species_2002} (SSDs) and Toxic Units (TUs) <CITATION>.
+Kefford et al.
+\citet{kefford_definition_2011} showed how SSDs to effectively approximate salinity effects on species richness (i.e.
+macroinvertebrates) by comparing SSDs with measurements of species richness determined in field studies.
 
@@
 \subsection*{The application}
-The application is accessible through \etox-base{} and was built by using the shiny web application framework \citep{chang_shiny_2018} in R.
+The application is accessible through \etoxbase{} and was built by using the shiny web application framework \citep{chang_shiny_2018} in R.
 Therein the user can choose to filter the data by several parameters such as chemical class, organism habitat, test duration and test endpoints amongst others.
 A detailed list with example entries of the possible filter parameters can be found in \ref{table:meta}.
-In order to derive single endpoints for an organism groups and a compound of the partly extensive test data, the \etox-base{} aggregates the test results according to the chosen filters in two steps.
+In order to derive single endpoints for an organism groups and a compound of the partly extensive test data, the \etoxbase{} aggregates the test results according to the chosen filters in two steps.
 Firstly the filtered test results are aggregated by the CAS number, the chosen taxon plus the duration and secondly only by the CAS number.
@@
 Besides the aggregation some cleaning can also be performed.
-The \etox-base{} allows to exclude test results exhibiting concentrations that are higher than the actual water solubility of the respective compound at \ang{20} C.
+The \etoxbase{} allows to exclude test results exhibiting concentrations that are higher than the actual water solubility of the respective compound at \ang{20} C.
 In any case the user can choose to exclude outliers.
@@
 \begin{figure}
-    %\includestandalone[width=\textwidth]{article/tikz/organigram}
-    \input{article/tikz/organigram}
+    %\includestandalone[]{tikz/test}
+    \includestandalone[width=\textwidth]{tikz/organigram}
+    %\includestandalone[width=\textwidth]{tikz/organigram}
     \caption{Organigram}
@@
 
+pH data is not really available for, temperature data either \ref{table:epa-variables}.
+therefore it is hard to evaluate single toxicities.
+We call for more concise reporting of ecotoxicological test data.
+It's all in articles but hardyl in a concise machine-readable format!
+
 \section*{Acknowledgements}
@@
 
-\begin{table}
+\begin{sidewaystable}
     \csvautotabular[respect all]{data/scripts.csv}
@@
     \label{table:processing-scripts}
-\end{table}
+\end{sidewaystable}
 
@@
 
-\begin{thebibliography}{}
-\bibliography{references-etox-base}
-\end{thebibliography}
+\bibliographystyle{apalike}
+\bibliography{refs/references-etox-base}
 

### 2018-12-17 article/tikz/organigram.tex

\documentclass[border = 1.5cm]{standalone}

%%%% packages
\usepackage{tikz}
\usetikzlibrary{trees,
                arrows,
                shapes.geometric,
                positioning,
                calc,
                backgrounds,
                fit}

%%%% document
\begin{document}

%%%% layers
\pgfdeclarelayer{bg1}    % declare background layer
\pgfdeclarelayer{bg2}
\pgfdeclarelayer{bg3}
\pgfsetlayers{bg3,bg2,bg1,main}  % set the order of the layers (main is the standard layer)

%%%% styles
\tikzset{
	font = {\fontsize{13pt}{12}\sffamily},
	children/.style = {anchor = west},
	box/.style = {rectangle, text centered, draw = black, anchor = north west},
	iomain/.style = {rectangle, minimum width = 4cm, minimum height = 3cm, text centered, draw = black, fill = blue!10},
	main/.style = {rectangle, draw = black, rounded corners, minimum width = 3cm, minimum height = 1.25cm, text centered, anchor = west},
	etox-base/.style = {rectangle, rounded corners},
	etox-base-shiny/.style = {circle},
	ids/.style = {diamond, minimum width = 3cm, minimum height = 2cm, text centered, draw = black, fill = green!30},
	res/.style = {rectangle, rounded corners, minimum width = 2cm, minimum height = 1cm, text centered, draw = black, fill = blue!30},
	resvariables/.style = {rectangle, minimum width = 2cm, minimum height = 1cm, text centered, anchor = west},
	arrow/.style = {->, > = stealth, draw = black, color = gray!80, line width = 2.5pt},
	background/.style = {rectangle, rounded corners, minimum width=23cm, fill=gray!10},
	arrow2/.style = {->, > = latex, gray!10, line width=10pt}
}

    %%%% picture
    \begin{tikzpicture}[
	grow via three points={one child at (0.5,-1) and
		two children at (0.5,-1) and (0.5,-1.7)},
	edge from parent path={(\tikzparentnode.south) |- (\tikzchildnode.west)},
	inner sep = 0pt, outer sep = 0pt
	]

	%% EPA ECOTOX node
	\node (io) [iomain, align = center] at (0cm,0cm) {EPA ECOTOX\\data base\\(cleaned)};
	%\node (chemprop) [io, dashed, right of=io, xshift=2cm] {CHEMPROP};
	%\node (etox) [io, dashed, right of=chemprop, xshift=2cm] {UBA etox};

	%% run + setup scripts
	\begin{scope}[local bounding box = run]
	\node (run) [main, above of = io, yshift=9cm, xshift=3cm, align=center] {Run Algorithm}
	    child{ node [children] { \textbf{run\_build.R} }};
	\node (setup) [main, right of = run, xshift = 5cm, align = center] {Setup\\Environment}
	    child{ node [children] { \textbf{setup.R} }};
	\end{scope}

	%% build & data scripts
	\begin{scope}[local bounding box = build]
    % main nodes
	\node (download) [main, above of=io, yshift=4.5cm, xshift=16cm, align=center,
	                  minimum width=4cm, minimum height=2cm] {Download\\EPA ECOTOX\\data base}
	    child { node [children, yshift = -0.5cm] { \textbf{bd\_epa\_download.R} }};
    \node (build) [main, left of=download, xshift=-5cm, align=center,
                   minimum width=4cm, minimum height=2cm] {Build\\EPA ECOTOX\\data base}
	    child { node [children, yshift = -0.5cm] { \textbf{bd\_epa\_postgres.R} }};
	\node (data) [main, left of=build, xshift=-5cm, align = center,
	              minimum width=4cm, minimum height = 2cm] {Prepare\\EPA ECOTOX\\data base}
	    child { node [children, yshift = -0.5cm] { \textbf{da\_epa1.R} }} 
	    child { node [children, yshift = -0.5cm] { \textbf{da\_epa2.R} }}
	    child { node [children, yshift = -0.5cm] { \textbf{da\_epa3.R} }};
	\end{scope}
	% bbox
	\begin{pgfonlayer}{bg3}
	\node (back_data) [background, minimum height=5cm, above of=download, xshift=-6cm, yshift=-2cm];
	\node (back_data_label) [right of = back_data, xshift = 11.5cm, rotate = 90, align=center]
	    {\huge{Data \& Build}\\\huge{scripts}};
	\end{pgfonlayer}

	%% chemicals and organisms
	\begin{scope}[local bounding box = query]
	\node (cas) [ids, below of=io, yshift = -2.5cm, xshift = 3cm] {CAS};
	% main nodes
	\node (chemscr) [main, right of=cas, xshift=4.5cm] {PHYS-CHEM}
    	child { node [children] { Alan Wood Pesticide Compendium - \textbf{qu\_aw.R } }}
    	child { node [children] { Pesticide Action Network - \textbf{qu\_pan.R} }}
    	child { node [children] { PhysProp - \textbf{qu\_pp.R} }}
    	child { node [children] { PubChem - \textbf{qu\_pubchem.R} }};
	\node (taxon) [ids, below of=cas, yshift=-6cm, align = center] {Taxon};
	% main nodes
	\node (habitat) [main, right of = taxon, xshift = 4.5cm, yshift = 1.5cm] {Habitats}
    	child { node [children] { WORMS data base - \textbf{qu\_worms.R} }}
    	child { node [children] { Manual Search - \textbf{qu\_manual.R} }};
	\node (region) [main, right of = taxon, xshift = 4.5cm, yshift = -1.5cm] {Regions}
	    child { node [children] { GBIF - \textbf{qu\_gbif.R} }};
	\end{scope}
	% bbox
	\begin{pgfonlayer}{bg3}
	\node (back_query) [background, minimum height=12cm, above of=cas, xshift=7cm, yshift=-5.5cm];
	\node (back_query_label) [right of = back_query, xshift = 11.5cm, rotate = 90]
	    {\huge{Query scripts}};
	\end{pgfonlayer}

	%% result
	\node (testres) [main, right of = cas, xshift = 14cm, yshift = -13cm] {Merge}
	    child { node [children]  { \textbf{re\_merge.R} }}
	    child { node [children]  { \textbf{re\_combine.R} }};
	\node (analysesres) [main, left of = testres, xshift = -4.5cm] {Analyses}
	    child { node [children]  { \textbf{re\_analyses.R} }};
	\node (checksres) [main, left of = analysesres, xshift = -4.5cm] {checks}
	    child { node [children]  { \textbf{re\_checks.R} }};
	\node (finalres) [main, left of = checksres, xshift = -4.5cm] {Final Table}
	    child { node [children]  { \textbf{re\_final.R} }};
	% bbox
	\begin{pgfonlayer}{bg3}
	\node (back_result) [background, minimum height=4cm, above of=finalres, xshift=8.5cm, yshift=-1cm];
	\node (back_result_label) [right of = back_result, xshift = 11.5cm, rotate = 90, align = center]
	    {\huge{Result}\\\huge{scripts}};
	\end{pgfonlayer}

	%% writing
	\node (writedb) [main, below of = finalres, yshift=-3cm, align = center] {Write to\\Data Base}
	    child { node [children] { \textbf{wr\_dbase.R} }};
	\node (writeapp) [main, right of = writedb, xshift=3cm] {Write to app}
	    child { node [children] { \textbf{wr\_app.R}  }};
	\node (writemeta) [main, right of = writeapp, xshift=3cm] {Write Meta data}
	    child { node [children] { \textbf{wr\_meta.R} }};
	\node (writestats) [main, right of = writemeta, xshift=3cm] {Write statistics}
	    child { node [children] { \textbf{wr\_stats.R} }};

	\begin{pgfonlayer}{bg3}
	\node (back_writing) [background, minimum height=3cm, above of = writedb, xshift=8.5cm, yshift=-1cm];
	\node (back_writing_label) [right of = back_writing, xshift = 11.5cm, rotate = 90, align = center]
	    {\huge{Writing}\\\huge{scripts}};
	\end{pgfonlayer}

    %%% arrows %%%
	\begin{pgfonlayer}{bg1}
	% io
	\draw [arrow] (io.east) -| (testres);
	% build & data
	\draw [arrow] (download.west) -- (build.east);
	\draw [arrow] (build.west) -- (data.east);
	\draw [arrow] (data.west) -| (io.north);
	% query
	\draw [arrow] (io) |- (cas.west);
	\draw [arrow] (cas.east) -- (chemscr.west);
	\draw [arrow] (chemscr.east) -| (testres.north);
	\draw [arrow] (io) |- (taxon.west);
	\draw [arrow] (taxon.east) -- (habitat.west);
	\draw [arrow] (taxon.east) -- (region.west);
	\draw [arrow] (habitat.east) -| (testres.north);
	\draw [arrow] (region.east) -| (testres.north);
	% result
	\draw [arrow] (testres) -- (analysesres);
	\draw [arrow] (analysesres) -- (checksres);
	\draw [arrow] (checksres) -- (finalres);
    % writing
    \foreach \x in {writedb, writeapp, writemeta, writestats}
	    { \draw [arrow] (finalres) -- +(0,-2.75) -| (\x.north); }
	% run + setup
	\draw [arrow2] (run.east) -- (setup.west);
	\foreach \x in {back_data, back_query, back_result, back_writing}
	    { \draw [arrow2] (run.west) -- +(-4,0) |- (\x.west); }

	%% picture
	% \node (epalogo) [] { \includegraphics[]{https://worldvectorlogo.com/fr/logo/epa-1}};

	\end{pgfonlayer}
	\end{tikzpicture}
\end{document}

### 2019-02-05 article/tikz/tree_postgres.tex
Updates from Overleaf

\documentclass[border=5pt]{standalone}

\usepackage{forest}

\begin{document}
\begin{forest}
for tree={font=\ttfamily, grow'=0, child anchor=west, parent anchor=south, anchor=west, calign=first,
    inner xsep=7pt,
    edge path={
      \noexpand\path [draw, \forestoption{edge}]
      (!u.south west) +(7.5pt,0) |- (.child anchor) pic {folder} \forestoption{edge label};
    },
    before typesetting nodes={
      if n=1
        {insert before={[,phantom]}}
        {}
    },
    fit=band,
    before computing xy={l=15pt}
}

[Postgres data base
    [etoxYYYYMMDD
        [ecotox]
        [public]
    ]
]

\end{forest}
\end{document}

### 2019-02-05 article/article.tex

@@
 \documentclass[english]{article}
+\usepackage[T1]{fontenc}
 \usepackage{natbib} % citations
@@
 \usepackage{tikz}
+\usepackage{forest}
 \usetikzlibrary{trees, arrows, shapes.geometric, positioning}
@@
 \usepackage[super]{nth} % for 1st, 2nd etc. \nth{}
+\usepackage{hyperref}
 
@@
 \newcommand{\gitapp}{https://github.com/andreasLD/etox-base-shiny}
+\newcommand{\ecfifty}{EC\textsubscript{50}}
 
@@
 
-\title{Etox-Base - a tool for ecotoxicologists}
+\title{Etox-Base: A tool to derive exposure endpoints from chemical test results}
+
+\author{Andreas Scharm{\"u}ller\textsuperscript{1{*}},
+        Verena Schreiner\textsuperscript{1},
+        Ralf B. Sch{\"a}fer\textsuperscript{1}}
+
+\maketitle
+\thispagestyle{fancy}
+
+1. Institute of Environmental Sciences, University of Koblenz-Landau Fortstraße 7, 76829 Landau, Germany {*}corresponding author(s):
+Andreas Scharm{\"u}ller (scharmueller@uni-landau.de)
 
@@
 
+this data will be useful for
+
 \end{abstract}
-
-\section*{Background \& Summary [max 700 words]}
-
-A large number of chemical compounds such as pharmaceuticals, pesticides and synthetic hormones are in daily use all over the world with insufficient knowledge about possible adverse effects on the environment.
-In Europe alone some 100,000 compounds are estimated to be in current use, whereof 30,000 are produced in quantities larger than one ton per year \citep{breithaupt_costs_2006}.
-Chemical compounds are brought to the environment deliberately - as it is the case for pesticides or arrive in the environment as a byproduct from other processes (e.g.
-as atmospheric emissions or as wastewater) \citep{schwarzenbach_challenge_2006}.
-Ecosystems provide essential services to human societies such as drinking and irrigation water, food and climate regulation.
-These ecosystem services are products of different ecosystem functions that crucially depend on the integrity of the populations and communities that drive these ecosystem functions.
-Pollution with man-made chemical toxicants threatens ecosystems in various ways which are currently not fully understood \citep{steffen_anthropocene_2007}.
-Besides habitat degradation, climate change and nutrient enrichment, chemical pollution poses a significant threat to earth's biodiversity
-
-Pollution with man-made chemical toxicants was identified as one of three major environmental problems for which research gaps hamper the derivation of planetary boundaries, i.e.
-thresholds beyond which irreversible state shifts may occur \citep{steffen_anthropocene_2007}.
-Bernhardt et al.
-\citet{bernhardt_synthetic_2017} argue that the knowledge gap how chemicals affect biodiversity and ecosystem functions, and in turn ecosystem services, would also impede society’s ability to accomplish the Sustainable Development Goals of the United Nations.
-Hence, research on the effects of chemicals on ecosystems is required for establishing thresholds ensuring ecosystem integrity and enabling societies to effectively manage chemical risks.
 
@@
 
-However, less than one percent are thoroughly tested \citep{breithaupt_costs_2006}.
-For substances that have been tested adverse effects could be detected.
-Pesticides for exmaple reduce macro-invertebrate biodiversity in streams \citep{beketov_pesticides_2013} and estrogenic-associated effects were found in fish \citep{vethaak_integrated_2005}.
-However for many substance groups, especially newly emerging ones authorative test results remain scarce or not hard to find.
-To overcome such knowledge paucities it is paramount to accumulate and distribute existing knowledge on possible adverse effects of substances widely.
-The ECOTOXicology knowledgebase (ECOTOX) created by U.S.
-Environmental Protection Agency (EPA) represents one such initiative.
-It collects ecotoxicological test data on on aquatic and terrestrial wildlife as well as plants and publishes it on a quarterly basis freely accessible on the web.
-By the time of writing it contained 919,123 test results, comprising 11,655 chemical compounds tested on 12,630 different species \citep{elonen_ecotoxicology_2018}.
-The ecotoxicological tests are collected without limitation and comprise all sorts of measured endpoints such as Effective Concentrations - EC50 values, No-observed effect concentratons NOEC or lowest observed effect concentraitons - LOEC.
-Likewise the data set contains many different effect measures such as mortality, intoxication, growth etc.
-and test durations ranging from seconds to weeks or even years.
-These test results can be used to derive toxicity risk thresholds.
-As ecotoxicological risk assessment often relies on national and international quality thresholds such as EU-Environmtal quality standards (EQS) or Regulatory Aceptable concentrations (RAC) in Germany.
-In order to make this data more widely accessible we created a tool that downloads every new version of the \epa{} and performs several quality checks, e.g.
+\section*{Background \& Summary [max 700 words]}
+
+A large number of chemicals such as pharmaceuticals, pesticides and synthetic hormones are in daily use all over the world with insufficient knowledge about possible adverse effects on the environment.
+In Europe alone some 100,000 chemicals are estimated to be in current use, whereof 30,000 are produced in quantities larger than one ton per year \citep{breithaupt_costs_2006}.
+Chemicals are brought to the environment deliberately - as it is the case of pesticides, or arrive therein as a byproduct from other processes (e.g.
+atmospheric emissions or wastewater) \citep{schwarzenbach_challenge_2006}.
+Ecosystems provide essential services to human societies such as drinking and irrigation water, food and climate regulation.
+These services are products of different ecosystem functions that crucially depend on the integrity of the populations and communities that drive these ecosystem functions.
+However, besides habitat degradation, climate change and nutrient enrichment, pollution with man-made chemical toxicants threatens these populations and communities in various ways which are currently not fully understood \citep{steffen_anthropocene_2007}.
+Pollution with man-made chemical toxicants was indeed identified as one of three major environmental problems for which research gaps hamper the derivation of planetary boundaries, i.e.
+thresholds beyond which irreversible state shifts may occur \citep{steffen_anthropocene_2007}.
+Bernhardt et al.
+\citet{bernhardt_synthetic_2017} argue further that the knowledge gap how chemicals effect populations and communities and hence ecosystem functions and ecosystem services, would also impede society’s ability to accomplish the Sustainable Development Goals of the United Nations.
+According to Breithaupt \citet{breithaupt_costs_2006} less than one percent of chemicals released to the environment are thoroughly tested.
+Although for some chemicals advanced and standardized \citep{oecd_oecd_2018} test methods have been developed, ecotoxicological tests for a lot of chemicals, especially newly emerging ones remain scarce and if existent, the information are represented sparsely and in inconsistent formats.
+\citep{gessner_fostering_2016}.
+Therefore we developed a tool that collects, processes and aggregates ecotoxicological tests results to then represent them in an harmonized form in a web tool - the \etoxbase tool: \href{http://139.14.20.252:3838/etox-base-shiny/}.
+
+As a resource for ecotoxicological tests we used the ECOTOXicology knowledgebase (ECOTOX) created by the U.S.
+Environmental Protection Agency (EPA), which collects raw, non harmonized ecotoxicological test data on on aquatic and terrestrial wildlife as well as plants and publishes it on a quarterly basis.
+By the time of writing it contained 926,108 test results, comprising 11,685 chemicals tested on 12,668 different species \citep{elonen_ecotoxicology_2018}.
+The collected ecotoxicological tests contain all sorts of measured endpoints such as Effective Concentrations (\ecfifty{}) values, No-observed effect concentrations (NOEC) or lowest observed effect concentrations (LOEC).
+Likewise the data set contains different effect measures such as mortality, intoxication and growth as well as test durations ranging from seconds to weeks or even years.
+
+The \etoxbase aggregates this information according to the user's inputs in order to provide reliable concentrations for the derivation of toxicity effect measures such as Species Sensitivity Distributions (SSDs) \citep{posthuma_species_2002} and Toxic Units (TUs), which represent two prominent concepts in ecotoxicology to assess effects on organisms.
+The former combines toxicity data of several organism groups towards a chemicals to estimate effects on biotic communities, the latter refers to effects of a specific organism (group) towards a chemical.
+The two concepts are widely used in ecotoxicology \citep{kefford_definition_2011, schafer_effects_2011} since the allow for a comparison of toxicities across multiple chemicals and biological communities, respectively.
+Although performed on one and the same organism and chemical, outcomes of multiple ecotoxicological tests can vary greatly due to different test parameters, different populations or other individual factors which in turn leads to ambiguities in choosing test results for proper risk assessment.
+
+The \etoxbase tool therefore aggregates test results to obtain single toxicity values for organism, chemical and test duration combinations.
+The tool downloads every new version of the \epa{} and performs several quality checks, e.g.
 unit harmonization, and cleaning steps on it.
-In addition to the test parameters that are retrieved from the \epa{} we query compound and organism parameters from other open data bases to adress missing information such as compound solubility or organism habitat or regional ocurrence patterns.
-Subsequently the data is provided online via a web app where users can filter the data according to needed test parameters.
-Furthermore the user can choose between several aggregation methods to retrieve individual values per compound.
-Compounds are identified by their CAS-number as this is also the key identifier in the \epa{}.
-Optionally the data base can also be downloaded as a whole for local useage.
-This is the first approach that tries to handle large amounts of toxicity test data to derive meaningful information about the susceptability of organisms towards individual compounds by aggregating test results.
-Researchers relying on ecotoxicological test data will benefit from the \etoxbase{} due to concise representation ecotoxicological test data.
-
-\citep{kefford_definition_2011} % paper that shows SSD species richness approximates effect of salinity on macroinvertebrates
-
-\citep{posthuma_species_2002} % describes concept of SSDs first
-
-two prominent concepts to assess effects on organisms or biodiversity in ecotoxicology are Species Sensitivity Distributions \citep{posthuma_species_2002} (SSDs) and Toxic Units (TUs) <CITATION>.
-Kefford et al.
-\citet{kefford_definition_2011} showed how SSDs to effectively approximate salinity effects on species richness (i.e.
-macroinvertebrates) by comparing SSDs with measurements of species richness determined in field studies.
+In addition to the test parameters that are retrieved from the \epa{} the tool includes chemical and organism parameters from other open data bases to address missing information such as water solubility, organism habitat or regional occurrence patterns.
 
@@
 The software downloads each new version and rebuilds it locally in a PostgreSQL data base.
-A processing pipline of several R \cite{r_core_team_r_2017} scripts prepares the data for the tool.
-The tool itself is a R function that is executed by a user acessing the R shiny \citep{chang_shiny_2018} application.
+A processing pipeline of several R \citep{r_core_team_r_2017} scripts prepares the data for the tool.
+The tool itself is a R function that is executed by a user accessing the R shiny \citep{chang_shiny_2018} application.
 In the following the steps for processing the data are explained in more detail (PATH-TO-OVERVIEW) as well as the scripts that have been used for it (REFER-TO-SCRIPTS-TABLE).
@@
 The data is downloaded (bd\_epa\_download.R) and subsequently built into a PostgreSQL data base (bd\_epa\_postgres.R).
-Thereafter the data is imported into R for preparing it (da\_epa1.R, da\_epa2.R, da\_epa3.R), by removing special characters type conversions, unit harmonizations and by reducing the columns.
+Thereafter the data is imported into R for pre-processing (da\_epa1.R, da\_epa2.R, da\_epa3.R), which includes validity checks, removal of special characters, unit harmonizations and selection of relevant variables.
 At the end the data is stored locally (epa3.rds, epa2\_taxa.rds, epa2\_chem.rds).
-Obtained CAS numbers and taxon names are then used to query other data bases for additional information on compounds and biota respectively.
-Compound information is retrieved from the Pubchem data base \citep{CITE_PUBCHEM} (qu\_pc.R), from the Compendium of Pesticide Common Names \citep{CITE_AW} (qu\_aw.R), from the Physprop data base \citep{CITE_PHYSPROP} (qu\_pp.R), from EUROSTAT and the Chemspider data base \citep(CITE-CHEMSPIDER).
+CAS numbers, unique identifiers and taxon names are then used to query other data bases for additional information on chemicals and biota respectively.
+Chemical information is retrieved from the Pubchem data base \citep{CITE_PUBCHEM} (qu\_pc.R), from the Compendium of Pesticide Common Names \citep{CITE_AW} (qu\_aw.R), from the Physprop data base \citep{CITE_PHYSPROP} (qu\_pp.R), from EUROSTAT and the Chemspider data base \citep{CITE-CHEMSPIDER}.
 These queries mostly rely on the webchem R-package \citep{szocs_webchem_2015-1}.
-Additional habitat and occurrence information for organisms is queried from the World Register of Marine Species (WORMS) \citep{WORMS} and from the Global Biodiversity Information Facility (GBIF) \citep{CITE_RGBIF} using the rgbif R-package.
+For biota, habitat and occurrence information is queried from the World Register of Marine Species (WORMS) \citep{WORMS} and from the Global Biodiversity Information Facility (GBIF) \citep{CITE_RGBIF} using the rgbif R-package.
 The acquired information is then merged (re\_merge.R), variables are combined and created (re\_combine.R), checked (re\_checks\_internal.R) and finally compiled in one final table (re\_final.R) This table is then used in the application to perform search queries on.
@@
 Therein the user can choose to filter the data by several parameters such as chemical class, organism habitat, test duration and test endpoints amongst others.
-A detailed list with example entries of the possible filter parameters can be found in \ref{table:meta}.
-In order to derive single endpoints for an organism groups and a compound of the partly extensive test data, the \etoxbase{} aggregates the test results according to the chosen filters in two steps.
+A detailed list for possible filter parameters can be found in \ref{table:meta}.
+
+In order to be able to calculate TUs and SSDs, derivation of single endpoint values for specific organism groups and a chemicals of the partly extensive test data, the \etoxbase{} aggregates the test results according to the chosen filters in two steps.
 Firstly the filtered test results are aggregated by the CAS number, the chosen taxon plus the duration and secondly only by the CAS number.
@@
 Besides the aggregation some cleaning can also be performed.
-The \etoxbase{} allows to exclude test results exhibiting concentrations that are higher than the actual water solubility of the respective compound at \ang{20} C.
+The \etoxbase{} allows to exclude test results exhibiting concentrations that are higher than the actual water solubility of the respective chemical at \ang{20} C.
 In any case the user can choose to exclude outliers.
-The outliers are selected by excluding values that exceed lower (0.25) and the upper (0.75) quartile by 1.5 times the interquartile range.
+The outliers are selected as values that exceed lower (0.25) and the upper (0.75) quartile by 1.5 times the inter-quartile range.
 All the steps described in this section are executed at each click in the app which runs the function fun\_ec50filter\_aggregation.R (cf.
 table \ref{table:processing-scripts}).
-Fast, adequate and live data manipulation is possible thanks to the data.table package \citep{dowle_data.table_2018}.
+\citep{dowle_data.table_2018}.
 
 \subsection*{Future of the project}
-As new versions of the \epa{} are published on a quarterly basis the data acquisition process is automated and new versions of the data and the \app{} are planed to be released regularly.
+As new versions of the \epa{} is published on a quarterly basis the data acquisition process is automated and new versions of the data and the \app{} are planed to be released regularly.
 
 \begin{figure}
-    %\includestandalone[]{tikz/test}
-    \includestandalone[width=\textwidth]{tikz/organigram}
-    %\includestandalone[width=\textwidth]{tikz/organigram}
+    \includestandalone[width=\textwidth, scale=0.1]{tikz/pipeline_organigram}
     \caption{Organigram}
@@
 
+\begin{figure}
+    \includestandalone[scale=0.5]{tikz/tree_source}
+    \caption{Processing pipeline}
+    \label{fig:organigram}
+\end{figure}
+
 \section*{Technical Validation}
@@
 %%on data use.
-This data set is designed to support ecotoxicologists in assessing the risk of chemical compounds on the environment.
+This data set is designed to support ecotoxicologists in assessing the risk of chemicals on the environment.
 As there is more and more ecotoxicological test data available, it becomes fundamental to provide such information in reasonable formats.
@@
 \subsection*{Limitations}
-Due to missing entries,...
 
-pH data is not really available for, temperature data either \ref{table:epa-variables}.
-therefore it is hard to evaluate single toxicities.
-We call for more concise reporting of ecotoxicological test data.
-It's all in articles but hardyl in a concise machine-readable format!
+SSDs aim to extrapolate from species to the community level while not ignoring environmental variables (e.g.
+pH, temperature, etc.) which can effect the toxicity of a specific chemical greatly \citep{posthuma_species_2002}.
+For a large number of the \epa{} tests, environmental variables are not recorded (See \ref{table:meta-variables} for proportions of missing entries per variable).
+Hence to some extent the reported test results inherit uncertainty.
+In this context we also call for more emphasis on machine-readability when publishing ecotoxicological test results <CITATION ON AS OF HOW MANY DATA POINTS MACHINES ARE NEEDED>.
+A lot of important information is still confined in continuous text of articles and therefore not ascertainable for large scale analyses.
+
+When using the data from the tool it is essential to think about which endpoint groups (e.g.
+EC50, LC50, NOEC etc.) and which effect measurements to use (e.g.
+MOR - Mortality, ITX - Intoxication, GRO - Growth etc.).
+
+Likewise the data of the \epa{} doesn't allow for the in ecotoxicology so frequently used distinction of acute and chronic tests. Even the OECD guidlines are presented only as pdfs inproperly.
+
+discuss flaws in column types (e.g some concentration entries contain '+' or '~' and are therefore not convertable to numeric) - This hampers large scale analysis greatly!
+
+write here that it is meant as an addition to PPDB
+
+in the futrue there will probably more of such resources. Als NORMAN which even performs encompasses expert judgment on the validity of tests
+
+other approaches: https://envirotoxdatabase.org/ - performs no aggregation, not open source
 
@@
 
+\begin{table}
+%\begin{sidewaystable}
+    \csvautotabular[respect underscore=true]{data/scripts.csv}
+    \caption{R Scripts to query and process data obtained from various resources. The script tree, i.e. the order which script is processed can be found in}
+    \label{table:processing-scripts};
+%\end{sidewaystable}
+\end{table}
+
+\csvreader[tabular=|l|l|,
+    table head=\hline & category & description\\\hline,
+    late after line=\\\hline]
+{data/scripts.csv}{name=\category,surname=\description}
+%{\thecsvrow & \surname~\name & \age}
+
 \begin{sidewaystable}
-    \csvautotabular[respect all]{data/scripts.csv}
-    \caption{Processing scripts}
-    \label{table:processing-scripts}
+    \csvautotabular[respect all]{data/meta.csv};
+    \caption{Meta variables table};
+    \label{table:meta-variables};
 \end{sidewaystable}
@@
 
+\section*{not included citations}
+Put in citation \citep{hartung_chemical_2009} ?
+which claims that REACH won't meet their assumptions and say 54 million vertebrate animals are needed for tests that would cost €9.5 billion over the next ten years.
+I.e 20 times more animals, 6 times the costs in comparison to the official estimates
+
 \section*{Data Citations}

### 2019-02-05 article/test.tex

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{csvsimple}
\usepackage{pgfplotstable}

\begin{document}

\begin{table}
\pgfplotstabletypeset[
	col sep=comma,
	%string type,
	columns={description,script}
	columns/description/.style={column name=Description, column type={|l}},
	columns/script/.style={column name=Script, column type={|l}},
	%columns/name/.style={column name=Name, column type={|l}},
	%columns/surname/.style={column name=Surname, column type={|l}},
	%columns/age/.style={column name=Age, column type={|c|}},
	%every head row/.style={before row=\hline,after row=\hline},
	%every last row/.style={after row=\hline},
	]{data/scripts.csv}

\end{table}

\pgfplotstabletypeset[
    col sep=comma,
    string type,
    columns={script,[index]2},
    ]{data/scripts.csv}

\end{document}

### 2019-02-05 article/test.tex
Updates from Overleaf

@@
     string type,
-    columns={script,[index]2},
+    columns={Script,Description},
+    columns/Description/.style={column type={|l}},
+    columns/Script/.style={column type={|l}, % needed for replacing underscore
+                           postproc cell content/.code={%
+                                \pgfplotsutilstrreplace{_}{\_}{##1}%
+                                \pgfkeyslet{/pgfplots/table/@cell content}\pgfplotsretval
+                                },
+                            },
+    every head row/.style={before row=\hline,after row=\hline},
+	every last row/.style={after row=\hline},
     ]{data/scripts.csv}

### 2019-02-05 article/article.tex

@@
 \usepackage{csvsimple}
+\usepackage{pgfplotstable}
 \usepackage{rotating} % to rotate tables
@@
 
-\begin{table}
-%\begin{sidewaystable}
-    \csvautotabular[respect underscore=true]{data/scripts.csv}
+%\begin{table}
+\begin{sidewaystable}
+    \pgfplotstabletypeset[
+    col sep=comma,
+    string type,
+    columns={Script,Description},
+    columns/Description/.style={column type={|l}},
+    columns/Script/.style={column type={|l}, % needed for replacing underscore
+                           postproc cell content/.code={%
+                                \pgfplotsutilstrreplace{_}{\_}{##1}%
+                                \pgfkeyslet{/pgfplots/table/@cell content}\pgfplotsretval
+                                },
+                            },
+    every head row/.style={before row=\hline,after row=\hline},
+	every last row/.style={after row=\hline},
+    ]{data/scripts.csv}
     \caption{R Scripts to query and process data obtained from various resources. The script tree, i.e. the order which script is processed can be found in}
     \label{table:processing-scripts};
-%\end{sidewaystable}
-\end{table}
-
-\csvreader[tabular=|l|l|,
-    table head=\hline & category & description\\\hline,
-    late after line=\\\hline]
-{data/scripts.csv}{name=\category,surname=\description}
-%{\thecsvrow & \surname~\name & \age}
+\end{sidewaystable}
+%\end{table}
 

### 2019-02-08 article/article.tex

--- 2019-02-05 article/article.tex
@@
 
-this data will be useful for
+<WRITE ABSTRACT>
 
 \end{abstract}
-
-plan:
-- occurring in vast numbers
-- pathways
-- ecosystem services 
-- biodiversity loss
-- Übergang abschätzung von chemikalien
 
@@
 
-The EPA releases \epa{} on a quarterly basis through their website.
-The software downloads each new version and rebuilds it locally in a PostgreSQL data base.
-A processing pipeline of several R \citep{r_core_team_r_2017} scripts prepares the data for the tool.
-The tool itself is a R function that is executed by a user accessing the R shiny \citep{chang_shiny_2018} application.
-In the following the steps for processing the data are explained in more detail (PATH-TO-OVERVIEW) as well as the scripts that have been used for it (REFER-TO-SCRIPTS-TABLE).
+The \etoxbase{} consists of two parts of software: Firstly a processing pipeline \ref{fig:pipeline} downloads quarterly released new versions of the \epa{} and performs several preparation steps in R \citep{r_core_team_r_2017}, to then store the outcome in a PostgreSQL data base.
+This constitutes the basis upon which the R shiny web application is accessing the data \citep{r_core_team_r_2017}.
+The application itself is set of R functions that filter and aggregate the data according to a user's inputs \ref{fig:app}.
+Finally a data set, which can be downloaded and an interactive plot, for data exploration are returned.
+In the following the two parts are explained in detail.
 
@@
 
-The data is downloaded (bd\_epa\_download.R) and subsequently built into a PostgreSQL data base (bd\_epa\_postgres.R).
-Thereafter the data is imported into R for pre-processing (da\_epa1.R, da\_epa2.R, da\_epa3.R), which includes validity checks, removal of special characters, unit harmonizations and selection of relevant variables.
-At the end the data is stored locally (epa3.rds, epa2\_taxa.rds, epa2\_chem.rds).
-CAS numbers, unique identifiers and taxon names are then used to query other data bases for additional information on chemicals and biota respectively.
+The data is downloaded (bd\_epa\_download.R) and then built into a PostgreSQL data base (bd\_epa\_postgres.R).
+Subsequently the data is imported into R for pre-processing (da\_epa1.R, da\_epa2.R, da\_epa3.R), which includes validity checks, removal of special characters, unit harmonizations and selection of relevant variables.
+Chemical abstract service (CAS) numbers, unique chemical identifiers together we taxon names are then used to query other data bases for additional information on chemicals and biota respectively.
 Chemical information is retrieved from the Pubchem data base \citep{CITE_PUBCHEM} (qu\_pc.R), from the Compendium of Pesticide Common Names \citep{CITE_AW} (qu\_aw.R), from the Physprop data base \citep{CITE_PHYSPROP} (qu\_pp.R), from EUROSTAT and the Chemspider data base \citep{CITE-CHEMSPIDER}.
 These queries mostly rely on the webchem R-package \citep{szocs_webchem_2015-1}.
-For biota, habitat and occurrence information is queried from the World Register of Marine Species (WORMS) \citep{WORMS} and from the Global Biodiversity Information Facility (GBIF) \citep{CITE_RGBIF} using the rgbif R-package.
-The acquired information is then merged (re\_merge.R), variables are combined and created (re\_combine.R), checked (re\_checks\_internal.R) and finally compiled in one final table (re\_final.R) This table is then used in the application to perform search queries on.
-For a detailed overview of what information is collected see table \ref{table:processing-scripts}.
+For biota, habitat and occurrence information is queried from the World Register of Marine Species (WORMS) \citep{WORMS} and from the Global Biodiversity Information Facility (GBIF) \citep{CITE_RGBIF} using the rgbif R-package \citep{chamberlain_rgbif_2018}.
+The acquired information is then merged (re\_merge.R), variables are combined and created (re\_combine.R), checked (re\_checks\_internal.R) and finally compiled in one table (re\_final.R) This table is then used in the application to perform search queries on.
+For a detailed overview of what information is collected see table \ref{tab:scripts-pipeline}.
+
+\begin{figure}
+    \includestandalone[width=\textwidth, scale=0.1]{tikz/pipeline_organigram}
+    \caption{Processing Pipeline}
+    \label{fig:pipeline}
+\end{figure}
 
 \subsection*{The application}
-The application is accessible through \etoxbase{} and was built by using the shiny web application framework \citep{chang_shiny_2018} in R.
-Therein the user can choose to filter the data by several parameters such as chemical class, organism habitat, test duration and test endpoints amongst others.
-A detailed list for possible filter parameters can be found in \ref{table:meta}.
+The application is accessible through \app{} and was built in R by using the shiny web application framework \citep{chang_shiny_2018}.
+Users can define several input parameters \ref{tab:inputs} in the application's graphical user interface (GUI), handeled in the ui.R script in order to adjust the aggregation according specific requirements.
+In doing so users execute R functions (fun\_filter.R, fun\_aggregation.R, fun\_filagg\_plot\_ly.R) in the server.R script, which produce a 'filtered' and an 'aggregated' data set as well as an interactive overview plot.
+The 'aggregated' data set is presented in the GUI.
+Both data sets can be downloaded via a download button.
 
+\begin{figure}
+    \includestandalone[scale=0.75]{tikz/application}
+    \caption{Application}
+    \label{fig:app}
+\end{figure}
+
+\subsection*{Aggregating the data}
 In order to be able to calculate TUs and SSDs, derivation of single endpoint values for specific organism groups and a chemicals of the partly extensive test data, the \etoxbase{} aggregates the test results according to the chosen filters in two steps.
@@
 The former can't be influenced by the user and calculates either the minimum or the median depending on the amount of results to aggregate (n <= 2: minimum, if n > 2: median).
-Thereof the second step calculates the minimum, the maximum, the median, the mean or the geometric mean as aggregate.
-Besides the aggregation some cleaning can also be performed.
-The \etoxbase{} allows to exclude test results exhibiting concentrations that are higher than the actual water solubility of the respective chemical at \ang{20} C.
-In any case the user can choose to exclude outliers.
+Thereof the second step calculates the minimum, the maximum, the median, the geometric mean, or the arithmetic mean as an aggregate.
+Besides the aggregation two cleaning can also be performed.
+Firstly the application allows to exclude test results exhibiting concentrations that are higher than the actual water solubility of the respective chemical at \ang{20} C.
+Secondly the user can choose to exclude outliers.
 The outliers are selected as values that exceed lower (0.25) and the upper (0.75) quartile by 1.5 times the inter-quartile range.
-All the steps described in this section are executed at each click in the app which runs the function fun\_ec50filter\_aggregation.R (cf.
-table \ref{table:processing-scripts}).
+All the steps described in this section are selectable in the GUI and are executed at each click by the server.
+(cf.
+table \ref{tab:scripts-app}).
+
 \citep{dowle_data.table_2018}.
@@
 \subsection*{Future of the project}
-As new versions of the \epa{} is published on a quarterly basis the data acquisition process is automated and new versions of the data and the \app{} are planed to be released regularly.
-
-\begin{figure}
-    \includestandalone[width=\textwidth, scale=0.1]{tikz/pipeline_organigram}
-    \caption{Organigram}
-    \label{fig:organigram}
-\end{figure}
 
@@
 PostgreSQL 9.5 has been used to build the data base.
-Hence R together with its packages and PostgreSQL suffice to rebuild the process.
+Hence, R together with its packages and PostgreSQL suffice to rebuild the process.
 The \app{} itself is stored in a second Github repostitory (\gitapp{}).
+Since new versions of the \epa{} are published on a quarterly basis the data acquisition process is automated and new versions of the data and the \app{} are scheduled to be released regularly.
+
+\begin{table}[]
+    \centering
+    \begin{tabular}{c|c}
+         Category       & Package       & Version   & Reference \\
+         processing     & RCurl         & \\
+         & 
+    \end{tabular}
+    \caption{List of R packages used in compiling the data}
+    \label{tab:rpackages}
+\end{table}
 
@@
 
-\begin{figure}
-    \includestandalone[scale=0.5]{tikz/tree_source}
-    \caption{Processing pipeline}
-    \label{fig:organigram}
-\end{figure}
-
 \section*{Technical Validation}
@@
 %%of their data.
+
+<PUT COMPARISON WITH PPDB DATA HERE?>
 
@@
 %%on data use.
-This data set is designed to support ecotoxicologists in assessing the risk of chemicals on the environment.
-As there is more and more ecotoxicological test data available, it becomes fundamental to provide such information in reasonable formats.
-This means, that it should be accessible for humans as well as computers.
-Up to now ecotoxicoloists rely on effect values of single publications or data bases such as the Pesticide Property Data Base (PPDB), having certain limitations such as encompassing only a few substance groups.
-The \epa{} tries to fill these gaps by collecting in peer-review articles published ecotoxicological test results.
-However due to lacking test information and un-harmonized data sets it is not always easily accessible.
-The \app{} tries to mitigate this gap.
-An ecotoxicologist can now access the \app{} and retrieve cleaned, harmonized and aggregated test data according to his or her needs.
+
+The \etoxbase{} can be accessed via an application here: \app{}. The user can filter among different test parameters \ref{tab:inputs}.
+
+The \etoxbase{} is designed to support ecotoxicologists in assessing the risk of chemicals on the environment.
+With an increase of ecotoxicological test data available, it becomes fundamental to provide such information in reasonable formats, meaning accessible for humans as well as machines.
+Currently chemical risk assessment relies on test results of a few, well tested organism groups, such as \textit{Daphnia magna}, \textit{Pimephales promelas}, \textit{Selenastrum capricornutum} reported in publications on test procedures <CITE>, compiled published data sets \citep{malaj_organic_2014} or data bases such as the Pesticide Property Data Base (PPDB) \citep{lewis_international_2016}, whereby each source having its limitations: Collecting toxicity information from individual publications is laborious and using already compiled data from publications or data bases is often limited to specific chemical groups, such as pesticides, biozides etc.
+The \etoxbase{}, making use of the \epa{} data aims to provide holistic aggregates (\ecfifty{}, NOEC, LOEC) of ecotoxicological tests results in adequate formats in order to support researchers to in the compilation of SSDs, TUs and other risk indicators.
+
+Generally it has to be stated here that the returned values are subject to change, as new tests will be constantly incorporated into the data set.
+However, we expect values of well tested chemical-organism combinations (i.e.
+high N <MAYBE PUT NUMBER HERE - CITATION>) to be relatively stable.
+Yet, returned values from the \etoxbase{} are certainly influenced by individual tests procedures and characteristics, such as pH, temperature or conductivity to just name a few.
+However, using the geometric mean as an aggregate allows for an adequate estimation of the central tendency of the toxicity of a chemical towards an organism group.
+The geometric mean is in comparison to the arithmetic mean considerably less influenced by outliers and is suitable for skewed data.
+Also, the geometric mean is preferable in contrast to the median, since the median completely ignores the influence of large or small values, making it unreliable for small data sets <CITATION GEOMETRIC MEAN>.
+
+Other initiatives, with different, though partly overlapping aims accumulating existing ecotoxicological test data are published online.
+The PPDB provides validated ecotoxicological test results for pesticides, narrowing chemicals to this group \citep{lewis_international_2016}.
+The Network of reference laboratories, research centres and related organisations for monitoring of emerging environmental substances (NORMAN) focuses on assembling river basin specific pollutants\citep{von_der_ohe_new_2011}.
+The EnviroTox data base \href{https://envirotoxdatabase.org/} makes also use of the \epa{} data in order to derive Predicted no-effect concentrations (PNEC) \citep{health_and_environmental_sciences_institute_hesi_envirotox_2019}.
+In comparison to the \etoxbase, these examples partly rely on the same data, yet not aiming to derive \ecfifty, LOEC and NOEC from it automatically.
 
@@
 
+PUT SSD HERE
+
 When using the data from the tool it is essential to think about which endpoint groups (e.g.
@@
 
-write here that it is meant as an addition to PPDB
-
 in the futrue there will probably more of such resources. Als NORMAN which even performs encompasses expert judgment on the validity of tests
-
-other approaches: https://envirotoxdatabase.org/ - performs no aggregation, not open source
 
@@
 
-\begin{table}
 %\begin{sidewaystable}
-    \csvautotabular[respect underscore=true]{data/scripts.csv}
-    \caption{R Scripts to query and process data obtained from various resources. The script tree, i.e. the order which script is processed can be found in}
-    \label{table:processing-scripts};
+%    \csvautotabular[respect all]{data/meta.csv};
+%    \caption{Meta variables table};
+%    \label{table:meta-variables};
 %\end{sidewaystable}
+
+\begin{table}[]
+\begin{tabular}{|l|l|l|l|}
+\hline
+Procedure   & Group       & Subgroups              & Options \\ \hline
+Filter      & Compound    & CAS                    & input CAS \\ \hline
+Filter      & Compound    & Concentration type     & Active Ingredient; Formulation; Total \\ \hline
+Filter      & Compound    & Chemical class         & Insecticides; Herbicides; Metals \\ \hline
+Filter      & Taxon       & Taxon                  & Choose a taxon \\ \hline
+Filter      & Taxon       & Organism habitat       & freshwater; terrestrial; marine; brackish \\ \hline
+Filter      & Taxon       & Continent              & North America; Asia; Europe \\ \hline
+Filter      & Test        & Duration               & enter duration in hours \\ \hline
+Filter      & Test        & Effect groups          & MOR - Mortality; POP - Population \\ \hline
+Filter      & Test        & Endpoins               & EC50, LOEC, LOEL \\ \hline
+Filter      & Checks      & Water solubility check & yes/no \\ \hline
+Filter      & Checks      & Remove outliers        & yes/no \\ \hline
+Aggregation & Aggregation & Aggregate              & Minimum; Maximum; Median \\ \hline
+\end{tabular}
+\caption{\etoxbase input possibilities}
+\label{tab:inputs}
 \end{table}
 
-\csvreader[tabular=|l|l|,
-    table head=\hline & category & description\\\hline,
-    late after line=\\\hline]
-{data/scripts.csv}{name=\category,surname=\description}
-%{\thecsvrow & \surname~\name & \age}
+\begin{table}[]
+\begin{tabular}{|l|l|}
+\hline
+Script                      & Description                                      \\ \hline
+run_build.R                 & runs all sub scripts                             \\ \hline
+setup.R                     & sets up the environment                          \\ \hline
+bd\_epa\_download.R         & download EPA Ecotox data base                    \\ \hline
+bd\_epa\_postgres.R         & build EPA Ecotox data base locally               \\ \hline
+da\_epa1.R                  & raw EPA Ecotox data export                       \\ \hline
+da\_epa2.R                  & process EPA Ecotox data                          \\ \hline
+da\_epa3.R                  & reduce EPA Ecotox data                           \\ \hline
+qu\_pc.R                    & query PubChem data base                          \\ \hline
+qu\_aw.R                    & query Compendium of Pesticide Common Names       \\ \hline
+qu\_pp.R                    & query Physprop data base                         \\ \hline
+qu\_eurostat\_chem\_class.R & query Eurostat data set                          \\ \hline
+qu\_chemspider\_scrape.R    & query Chemspider data base                       \\ \hline
+qu\_worms2.R                & query WORMS data base                            \\ \hline
+qu\_gbif.R                  & query GBIF data base                             \\ \hline
+re\_merge.R                 & merge all data                                   \\ \hline
+re\_combine.R               & combine information                              \\ \hline
+re\_analyses.R              & ???                                              \\ \hline
+re\_checks\_internal.R      & ???                                              \\ \hline
+re\_final.R                 & create and write final table                     \\ \hline
+wr\_dbase.R                 & write final table to data base                   \\ \hline
+wr\_app.R                   & write final table and variables to app directory \\ \hline
+wr\_meta.R                  & ???                                              \\ \hline
+wr\_stats.R                 & ???                                              \\ \hline
+\end{tabular}
+\caption{Scripts that are used to process the data}
+\label{tab:scripts-pipeline}
+\end{table}
 
-\begin{sidewaystable}
-    \csvautotabular[respect all]{data/meta.csv};
-    \caption{Meta variables table};
-    \label{table:meta-variables};
-\end{sidewaystable}
+\begin{table}[]
+\begin{tabular}{|l|l|}
+\hline
+Script                     & Description                                               \\ \hline
+setup.R                    & sets up the environment                                   \\ \hline
+ui.R                       & creates the graphical user interface of the application   \\ \hline
+server.R                   & handles functions and data input and output               \\ \hline
+fun\_filter.R              & function that filters the data                            \\ \hline
+fun\_aggregation.R         & function that performs the aggregation                    \\ \hline
+fun\_filagg\_plot\_ly.R    & creates interactive plot                                  \\ \hline
+\end{tabular}
+\caption{Scripts that used for the web application}
+\label{tab:scripts-app}
+\end{table}
 
@@
 \bibliographystyle{apalike}
-\bibliography{refs/references-etox-base}
+\bibliography{refs/references-etox-base,
+              refs/references-etox-base-rpackages,
+              refs/references-etox-base-rshinypackages}
+
+\section*{Data Citations}
+
+\pagebreak
+\section*{TODO}
+
+\begin{itemize}
+
+\item compare Etox-BAse output with PPDB
+
+\item plot example SSD and TU
+
+\pagebreak
+\section*{MISC}
+
+\begin{figure}
+    \includestandalone[scale=0.5]{tikz/tree_source}
+    \caption{Alternative processing pipeline}
+    \label{fig:pipeline-tree}
+\end{figure}
 
@@
 
-\section*{Data Citations}
+\end{itemize}
 

### 2019-02-08 article/tikz/application.tex

\documentclass[border = 1.5cm]{standalone}

%%%% packages
\usepackage{tikz}
\usetikzlibrary{trees,
                arrows,
                shapes.geometric,
                positioning,
                calc,
                backgrounds,
                fit}

%%%% document
\begin{document}

%%%% layers
\pgfdeclarelayer{bg1}    % declare background layer
\pgfdeclarelayer{bg2}
\pgfdeclarelayer{bg3}
\pgfsetlayers{bg3,bg2,bg1,main}  % set the order of the layers (main is the standard layer)

%%%% styles
\tikzset{
	font = {\fontsize{13pt}{12}\sffamily},
	children/.style = {anchor = west},
	main/.style = {rectangle, draw = black, rounded corners, minimum width = 3cm, minimum height = 1.25cm, text centered, anchor = west},
	arrow/.style = {<->, > = stealth, draw = black, color = gray!80, line width = 2.5pt},
	background/.style = {rectangle, rounded corners, minimum width=23cm, fill=gray!10},
}

    %%%% picture
    \begin{tikzpicture}[
	grow via three points={one child at (0.5,-1) and
		two children at (0.5,-1) and (0.5,-1.7)},
	edge from parent path={(\tikzparentnode.south) |- (\tikzchildnode.west)},
	inner sep = 0pt, outer sep = 0pt
	]

    \begin{scope}
	\node (server) [main, at(0cm,0cm), xshift = 4cm, align = center] {Server \\ \textbf{server.R}}
	    child{ node [children, yshift = -0.5cm] { \textbf{setup.R} }}
	    child{ node [children, yshift = -0.5cm] { \textbf{fun\_filter.R} }}
	    child{ node [children, yshift = -0.5cm] { \textbf{fun\_aggregation.R} }};

	\node (ui) [main, right of = server, xshift = 5cm, align = center] {User Interface \\ \textbf{ui.R}};
	\end{scope}

	\begin{pgfonlayer}{bg3}
	\node (back_query) [background, minimum height=5cm, minimum width=14cm, above of=server, xshift=4cm, yshift=-2.25cm];
	\node (back_query_label) [below of = back_query, xshift = 5cm, yshift = -1cm, rotate = 0]
	    {\huge{Application}};
	\end{pgfonlayer}

	\draw [arrow] (server.east) -- (ui.west);

	\end{tikzpicture}
\end{document}

### 2019-02-08 article/article.tex
Updates from Overleaf

@@
 
-\citep{dowle_data.table_2018}.
-
-\subsection*{Future of the project}
-
 \subsection*{Code availability}
@@
 PostgreSQL 9.5 has been used to build the data base.
-Hence, R together with its packages and PostgreSQL suffice to rebuild the process.
+Hence, R together with its packages \ref{tab:rpackages} and PostgreSQL suffice to rebuild the process.
 The \app{} itself is stored in a second Github repostitory (\gitapp{}).
@@
     \centering
-    \begin{tabular}{c|c}
-         Category       & Package       & Version   & Reference \\
-         processing     & RCurl         & \\
-         & 
+    \begin{tabular}{|l|l|l|l|}
+        \hline
+        Category        & Package       & Version   & Reference                     \\ \hline
+        processing      & RCurl         & 1.95-4.11 & \citep{lang_rcurl_2018}       \\ \hline
+        processing      & data.table    & 1.12.0    & \citep{dowle_data.table_2018} \\ \hline
+        application     & shiny         & 1.1.0     & \citep{chang_shiny_2018}      \\ \hline
     \end{tabular}
-    \caption{List of R packages used in compiling the data}
+    \caption{List of R packages used in compiling the data.\newline{}
+    TODO: Automate version input and name generation - whole table\newline{}
+    TODO2: debug chang\_shiny\_2018 reference}
     \label{tab:rpackages}
@@
 
-The \etoxbase{} can be accessed via an application here: \app{}. The user can filter among different test parameters \ref{tab:inputs}.
-
+The \etoxbase{} can be accessed via an application here: \app{}.
+The user can filter among different test parameters \ref{tab:inputs}.
 The \etoxbase{} is designed to support ecotoxicologists in assessing the risk of chemicals on the environment.
 With an increase of ecotoxicological test data available, it becomes fundamental to provide such information in reasonable formats, meaning accessible for humans as well as machines.
-Currently chemical risk assessment relies on test results of a few, well tested organism groups, such as \textit{Daphnia magna}, \textit{Pimephales promelas}, \textit{Selenastrum capricornutum} reported in publications on test procedures <CITE>, compiled published data sets \citep{malaj_organic_2014} or data bases such as the Pesticide Property Data Base (PPDB) \citep{lewis_international_2016}, whereby each source having its limitations: Collecting toxicity information from individual publications is laborious and using already compiled data from publications or data bases is often limited to specific chemical groups, such as pesticides, biozides etc.
-The \etoxbase{}, making use of the \epa{} data aims to provide holistic aggregates (\ecfifty{}, NOEC, LOEC) of ecotoxicological tests results in adequate formats in order to support researchers to in the compilation of SSDs, TUs and other risk indicators.
+Currently chemical risk assessment relies on test results of a few, well tested organism groups, such as \textit{Daphnia magna}, \textit{Pimephales promelas}, \textit{Selenastrum capricornutum} reported in publications on test procedures <CITE>, compiled and published data sets \citep{malaj_organic_2014} or in data bases such as the Pesticide Property Data Base (PPDB) \citep{lewis_international_2016}.
+Each data source having its limitations: Collecting toxicity information from individual publications is laborious and using already compiled data from publications or data bases is often limited to specific chemical groups, such as pesticides, biozides etc.
+The \etoxbase{}, making use of the \epa{}, the largest publicly available collection of ecotoxicological test results, aims to provide holistic aggregates (\ecfifty{}, NOEC, LOEC) of ecotoxicological tests results in adequate formats in order to support researchers in fetching ecotoxicological data for the calculation of SSDs, TUs and other risk indicators.
+SSDs try to extrapolate effects of chemicals towards organisms from species to a community level by fitting a statistical distribution to toxicity test results on organisms of different trophic levels \citep{posthuma_species_2002}.
+As an example
+
+\begin{figure}
+    \dentering
+    \includegraphics[width=1\linewidth]{figures/ssd2}
+
+\end{figure}
 
@@
 \bibliographystyle{apalike}
-\bibliography{refs/references-etox-base,
-              refs/references-etox-base-rpackages,
-              refs/references-etox-base-rshinypackages}
+\bibliography{refs/references-etox-base.bib,refs/references-etox-base-rpackages.bib,refs/references-etox-base-rshinypackages.bib}
+% multiple refs have to be written w/o space
 

### 2019-02-09 article/article.tex

@@
 According to Breithaupt \citet{breithaupt_costs_2006} less than one percent of chemicals released to the environment are thoroughly tested.
-Although for some chemicals advanced and standardized \citep{oecd_oecd_2018} test methods have been developed, ecotoxicological tests for a lot of chemicals, especially newly emerging ones remain scarce and if existent, the information are represented sparsely and in inconsistent formats.
+Although for some chemicals advanced and standardized \citep{oecd_oecd_2018} test methods have been developed, ecotoxicological tests for a lot of chemicals, especially newly emerging ones remain scarce and if existent, the information is represented sparsely and in inconsistent formats.
 \citep{gessner_fostering_2016}.
-Therefore we developed a tool that collects, processes and aggregates ecotoxicological tests results to then represent them in an harmonized form in a web tool - the \etoxbase tool: \href{http://139.14.20.252:3838/etox-base-shiny/}.
-
-As a resource for ecotoxicological tests we used the ECOTOXicology knowledgebase (ECOTOX) created by the U.S.
-Environmental Protection Agency (EPA), which collects raw, non harmonized ecotoxicological test data on on aquatic and terrestrial wildlife as well as plants and publishes it on a quarterly basis.
+Therefore we developed a data base that collects, processes and aggregates ecotoxicological tests results in order to subsequently publish it in an harmonized form as a web application - the \etoxbase tool: \href{http://139.14.20.252:3838/etox-base-shiny/}.
+This should facilitate researchers the access to ecotoxicological test results and support modern chemical risk assessment (CRA) workflows.
+The \etoxbase makes use of ecotoxicological test data, obtained from the ECOTOXicology knowledgebase (ECOTOX) created by the U.S.
+Environmental Protection Agency (EPA).
+ECOTOX collects raw, non harmonized ecotoxicological test data on on aquatic and terrestrial wildlife as well as plants and publishes it on a quarterly basis.
 By the time of writing it contained 926,108 test results, comprising 11,685 chemicals tested on 12,668 different species \citep{elonen_ecotoxicology_2018}.
-The collected ecotoxicological tests contain all sorts of measured endpoints such as Effective Concentrations (\ecfifty{}) values, No-observed effect concentrations (NOEC) or lowest observed effect concentrations (LOEC).
-Likewise the data set contains different effect measures such as mortality, intoxication and growth as well as test durations ranging from seconds to weeks or even years.
-
-The \etoxbase aggregates this information according to the user's inputs in order to provide reliable concentrations for the derivation of toxicity effect measures such as Species Sensitivity Distributions (SSDs) \citep{posthuma_species_2002} and Toxic Units (TUs), which represent two prominent concepts in ecotoxicology to assess effects on organisms.
+The collected ecotoxicological test data contain a variety of measured endpoints such as Effective Concentrations (\ecfifty{}) values, No-observed effect concentrations (NOEC) or lowest observed effect concentrations (LOEC).
+Likewise the data set contains different effect measures such as mortality, intoxication and growth as well as a multiplicity of test durations ranging from seconds to weeks or even years.
+The \etoxbase aggregates this information according to the user's inputs.
+These can then be used for the derivation of risk indicators such as Species Sensitivity Distributions (SSDs) \citep{posthuma_species_2002} and Toxic Units (TUs), which represent two prominent concepts to assess effects on organisms in ecotoxicology .
 The former combines toxicity data of several organism groups towards a chemicals to estimate effects on biotic communities, the latter refers to effects of a specific organism (group) towards a chemical.
 The two concepts are widely used in ecotoxicology \citep{kefford_definition_2011, schafer_effects_2011} since the allow for a comparison of toxicities across multiple chemicals and biological communities, respectively.
-Although performed on one and the same organism and chemical, outcomes of multiple ecotoxicological tests can vary greatly due to different test parameters, different populations or other individual factors which in turn leads to ambiguities in choosing test results for proper risk assessment.
-
-The \etoxbase tool therefore aggregates test results to obtain single toxicity values for organism, chemical and test duration combinations.
-The tool downloads every new version of the \epa{} and performs several quality checks, e.g.
+Although performed on one and the same organism and chemical, outcomes of different ecotoxicological tests vary greatly due to differences in test parameters, genetic differences in individual populations or other irrepressible factors making a selection of appropriate test results for CRA laborious.
+The \etoxbase{} aims to relief this process and provide single toxicity values for organism, chemical and test duration combinations.
+The \etoxbase{} downloads every new version of the ECOTOX data base and performs several quality checks, e.g.
 unit harmonization, and cleaning steps on it.
-In addition to the test parameters that are retrieved from the \epa{} the tool includes chemical and organism parameters from other open data bases to address missing information such as water solubility, organism habitat or regional occurrence patterns.
+Hence, new scientific test results are constantly incorporated.
+Additionally the \etoxbase{} queries chemical- and organism-specific variables from other open data bases to fill information gaps such as water solubility, organism habitat or regional occurrence patterns in the ECOTOX data base.
 
@@
 
-The \etoxbase{} consists of two parts of software: Firstly a processing pipeline \ref{fig:pipeline} downloads quarterly released new versions of the \epa{} and performs several preparation steps in R \citep{r_core_team_r_2017}, to then store the outcome in a PostgreSQL data base.
+The \etoxbase{} consists of two parts of software: Firstly a processing pipeline (Fig.
+\ref{fig:pipeline}) downloads quarterly released new versions of the \epa{} and performs several preparation steps in R \citep{r_core_team_r_2017}, to then store the outcome in a PostgreSQL data base.
 This constitutes the basis upon which the R shiny web application is accessing the data \citep{r_core_team_r_2017}.
 The application itself is set of R functions that filter and aggregate the data according to a user's inputs \ref{fig:app}.
-Finally a data set, which can be downloaded and an interactive plot, for data exploration are returned.
-In the following the two parts are explained in detail.
+Finally in the web application, a `filtered` and an `aggregated` data set as well as an interactive plot, for data exploration are returned.
+These two data sets can be downloaded.
+In the following these two parts are explained in detail, naming the respective R scripts in brackets.
 
@@
 
-The data is downloaded (bd\_epa\_download.R) and then built into a PostgreSQL data base (bd\_epa\_postgres.R).
-Subsequently the data is imported into R for pre-processing (da\_epa1.R, da\_epa2.R, da\_epa3.R), which includes validity checks, removal of special characters, unit harmonizations and selection of relevant variables.
+The EPA ECOTOX data is downloaded (bd\_epa\_download.R) and subsequently built into a PostgreSQL data base (bd\_epa\_postgres.R).
+Afterwards the data is imported into R for pre-processing (da\_epa1.R, da\_epa2.R, da\_epa3.R), which includes validity checks, removal of special characters, unit harmonizations and selection of relevant variables.
 Chemical abstract service (CAS) numbers, unique chemical identifiers together we taxon names are then used to query other data bases for additional information on chemicals and biota respectively.
-Chemical information is retrieved from the Pubchem data base \citep{CITE_PUBCHEM} (qu\_pc.R), from the Compendium of Pesticide Common Names \citep{CITE_AW} (qu\_aw.R), from the Physprop data base \citep{CITE_PHYSPROP} (qu\_pp.R), from EUROSTAT and the Chemspider data base \citep{CITE-CHEMSPIDER}.
+Chemical information is retrieved from the Compendium of Pesticide Common Names (qu\_aw.R) \citep{CITE_AW}, the Chemspider data base (qu\_chemspieder\_scrape.R) \citep{CITE-CHEMSPIDER}, from Eurostat <CITATION>, the Pubchem data base (qu\_pc.R) \citep{CITE_PUBCHEM}, and from the Physprop data base (qu\_pp.R) \citep{CITE_PHYSPROP}.
 These queries mostly rely on the webchem R-package \citep{szocs_webchem_2015-1}.
-For biota, habitat and occurrence information is queried from the World Register of Marine Species (WORMS) \citep{WORMS} and from the Global Biodiversity Information Facility (GBIF) \citep{CITE_RGBIF} using the rgbif R-package \citep{chamberlain_rgbif_2018}.
-The acquired information is then merged (re\_merge.R), variables are combined and created (re\_combine.R), checked (re\_checks\_internal.R) and finally compiled in one table (re\_final.R) This table is then used in the application to perform search queries on.
+For biota, habitat and occurrence information is queried from the World Register of Marine Species (WORMS) (qu\_worms2.R) \citep{WORMS} and from the Global Biodiversity Information Facility (GBIF) (qu\_rgbif.R) \citep{CITE_RGBIF} using the rgbif R-package \citep{chamberlain_rgbif_2018}.
+The acquired information is then merged (re\_merge.R) and information is combined (re\_combine.R).
+From there on the data set is analysed (re\_analyses.R) and checked (re\_checks\_internal.R) and finally compiled in one table (re\_final.R).
+This table is then used as the input for the filter and aggregation functions \ref{fig:pipeline}.
 For a detailed overview of what information is collected see table \ref{tab:scripts-pipeline}.
@@
 \subsection*{The application}
-The application is accessible through \app{} and was built in R by using the shiny web application framework \citep{chang_shiny_2018}.
-Users can define several input parameters \ref{tab:inputs} in the application's graphical user interface (GUI), handeled in the ui.R script in order to adjust the aggregation according specific requirements.
-In doing so users execute R functions (fun\_filter.R, fun\_aggregation.R, fun\_filagg\_plot\_ly.R) in the server.R script, which produce a 'filtered' and an 'aggregated' data set as well as an interactive overview plot.
-The 'aggregated' data set is presented in the GUI.
-Both data sets can be downloaded via a download button.
+The application is accessible through \app{} and was built in R, using the shiny web application framework \citep{chang_shiny_2018}.
+Users can define several input parameters \ref{tab:inputs} in the application's graphical user interface (GUI), handled in the ui.R script in order to adjust the aggregation according specific requirements.
+In doing so users execute R functions (fun\_filter.R, fun\_aggregation.R, fun\_filagg\_plot\_ly.R) loaded in the server.R script, which produce a 'filtered' and an 'aggregated' data set as well as an interactive overview plot.
+Only the 'aggregated' data set is presented in the GUI.
+However, both data sets can be downloaded via a download button.
 
@@
 
-\subsection*{Aggregating the data}
-In order to be able to calculate TUs and SSDs, derivation of single endpoint values for specific organism groups and a chemicals of the partly extensive test data, the \etoxbase{} aggregates the test results according to the chosen filters in two steps.
-Firstly the filtered test results are aggregated by the CAS number, the chosen taxon plus the duration and secondly only by the CAS number.
+\subsection*{Aggregation}
+The \etoxbase{} application aggregates the test results according to chosen filters in a two step process.
+Firstly the filtered test results are aggregated by the CAS number, the chosen taxon and the selected test duration.
+Secondly, the returned data is then aggregated by the CAS number.
 The former can't be influenced by the user and calculates either the minimum or the median depending on the amount of results to aggregate (n <= 2: minimum, if n > 2: median).
 Thereof the second step calculates the minimum, the maximum, the median, the geometric mean, or the arithmetic mean as an aggregate.
-Besides the aggregation two cleaning can also be performed.
-Firstly the application allows to exclude test results exhibiting concentrations that are higher than the actual water solubility of the respective chemical at \ang{20} C.
-Secondly the user can choose to exclude outliers.
-The outliers are selected as values that exceed lower (0.25) and the upper (0.75) quartile by 1.5 times the inter-quartile range.
-All the steps described in this section are selectable in the GUI and are executed at each click by the server.
-(cf.
-table \ref{tab:scripts-app}).
 
@@
 
-The code for processing the data is stored in a Github reopsitory (\git{}).
+The processing code is stored in a Github reopsitory (\git{}).
 All processing scripts were written in R version 3.4.4.
@@
 Hence, R together with its packages \ref{tab:rpackages} and PostgreSQL suffice to rebuild the process.
-The \app{} itself is stored in a second Github repostitory (\gitapp{}).
-Since new versions of the \epa{} are published on a quarterly basis the data acquisition process is automated and new versions of the data and the \app{} are scheduled to be released regularly.
+The code of the application itself is stored in a second Github repostitory (\gitapp{}).
+Since new versions of the EPA ECOTOX data base are published on a quarterly basis the data acquisition process is automated and new versions of the application are scheduled to be released regularly.
 
-\begin{table}[]
+\begin{table}[h!]
     \centering
@@
 
-The \etoxbase{} can be accessed via an application here: \app{}.
-The user can filter among different test parameters \ref{tab:inputs}.
-The \etoxbase{} is designed to support ecotoxicologists in assessing the risk of chemicals on the environment.
-With an increase of ecotoxicological test data available, it becomes fundamental to provide such information in reasonable formats, meaning accessible for humans as well as machines.
-Currently chemical risk assessment relies on test results of a few, well tested organism groups, such as \textit{Daphnia magna}, \textit{Pimephales promelas}, \textit{Selenastrum capricornutum} reported in publications on test procedures <CITE>, compiled and published data sets \citep{malaj_organic_2014} or in data bases such as the Pesticide Property Data Base (PPDB) \citep{lewis_international_2016}.
-Each data source having its limitations: Collecting toxicity information from individual publications is laborious and using already compiled data from publications or data bases is often limited to specific chemical groups, such as pesticides, biozides etc.
-The \etoxbase{}, making use of the \epa{}, the largest publicly available collection of ecotoxicological test results, aims to provide holistic aggregates (\ecfifty{}, NOEC, LOEC) of ecotoxicological tests results in adequate formats in order to support researchers in fetching ecotoxicological data for the calculation of SSDs, TUs and other risk indicators.
-SSDs try to extrapolate effects of chemicals towards organisms from species to a community level by fitting a statistical distribution to toxicity test results on organisms of different trophic levels \citep{posthuma_species_2002}.
-As an example
+The \etoxbase{} is designed to support CRA of ecotoxicologists.
+With an increased amount of available ecotoxicological test data, it becomes fundamental to provide and distribute such information in reasonable formats, meaning accessible for humans as well as machines.
+Currently CRA often relies on test results of a few, well tested standard organism, such as \textit{Daphnia magna}, \textit{Pimephales promelas}, \textit{Selenastrum capricornutum}, although researchers have conducted experiments on a much greater variety of organisms (e.g.
+the 926,108 test results collected in the EPA ECOTOX data base).
+In order to locate ecotoxicological test data, researchers can rely on individual test results reported in publications <CITE>, on compiled and published data sets \citep{malaj_organic_2014} or on data bases such as the Pesticide Property Data Base (PPDB) \citep{lewis_international_2016}.
+Each data source has its limitations though: Collecting toxicity information from individual publications is laborious and using already compiled data from publications or data bases is often limited to a specific groups of chemicals, such as pesticides, biozides etc.
+The \etoxbase{}, making use of the EPA ECOTOX data base, the largest publicly available collection of ecotoxicological test results, aims to facilitate the data retrieval process by providing adequate aggregates (\ecfifty{}, NOEC, LOEC) of the tests results.
+These are then used for the calculation of SSDs, TUs or other risk indicators.
+SSDs try to extrapolate effects of chemicals towards organisms from species to a community level by fitting a statistical distribution to toxicity test results on organisms of different trophic levels in order to derive concentrations at which a specific fraction of organisms are effected (often at 5\% - HC\textsubscript{5} \citep{posthuma_species_2002}.
+Figure \ref{fig:ssd-isoproturon} shows an example of a SSD and the derived HC\textsubscript{5} value (red dassed line) created for the herbicide Isoproturon.
+TUs are another indicator of effects of chemicals on organisms.
+They are calculated by dividing a measured concentration (c) by an \ecfifty{} value (cf.
+Equation \ref{eq:tu}).
 
-\begin{figure}
+\begin{equation}
+    TU_i = \frac{c_i}{\ecfifty}
+    \label{eq:tu}
+\end{equation}
+
+\begin{figure}[h!]
     \dentering
-    \includegraphics[width=1\linewidth]{figures/ssd2}
-
+    \includegraphics[width=1\linewidth]{article/figures/ssd2_boot.png}
+    \caption{SSD plot showing the susceptibility of organisms towards the herbicide Isoproturon.
+The red line represents the fit and the red dashed line marks the HC\textsubscript{5} value, which is a common measure in CRA.
+The legend denotes common organism groups used in ecotoxicology.}
+    \label{fig:ssd-isoproturon}
 \end{figure}
 
-Generally it has to be stated here that the returned values are subject to change, as new tests will be constantly incorporated into the data set.
+\subsection*{Filter and Aggregate}
+
+In the application's GUI, the user can choose among different parameters to filter and aggregate the test data accordingly \ref{tab:inputs}.
+Percentage values next to inputs show the proportion of the respective option (e.g.
+\textit{Active Ingredient - 43\%} indicates that 43\% of the data are Active Ingredients).
+The user can filter for compound-specific filters, including the CAS number, the concentration type (e.g.
+Active Ingredient, Formulation), the chemical class (e.g.
+Insecticides, Metals) and taxon-specific filters, including common taxonomic groups (e.g.
+Daphniidae, Algae), the organism habitat (e.g.
+freshwater) and the continent (e.g.
+Europe, Asia).
+Regarding the test parameters the user can choose test durations (in hours), Effect groups (e.g.
+Mortaliy, Population, Growth) and Endpoints (\ecfifty, EC\textsubscript{10}, LOEC or NOEC).
+Besides that, two cleaning steps can also be chosen.
+Firstly the application allows to exclude test results exhibiting concentrations that are higher than the actual water solubility of the respective chemical at \ang{20} C.
+Secondly the user can choose to exclude outliers (cf.
+Table \ref{tab:scripts-app}).
+The outliers are selected as values that exceed lower (0.25) and the upper (0.75) quartile by 1.5 times the inter-quartile range.
+As aggregates the user can choose the minimum, the maximum, the median, the geometric mean or the arithmetic mean.
+All the steps described in this section are selectable in the GUI and are executed at each click by the application.
+The drop down menu `Download data` allows to download a filtered data set as well as an aggregated data set.
+The \etoxbase{} application can be accessed here: \app{}.
+
+\begin{table}[h!]
+\begin{tabular}{|l|l|l|l|}
+\hline
+Menu   & Group       & Subgroups              & Options \\ \hline
+Filter      & Compound    & CAS                    & input CAS \\ \hline
+Filter      & Compound    & Concentration type     & Active Ingredient; Formulation; Total \\ \hline
+Filter      & Compound    & Chemical class         & Insecticides; Herbicides; Metals \\ \hline
+Filter      & Taxon       & Taxon                  & Choose a taxon \\ \hline
+Filter      & Taxon       & Organism habitat       & freshwater; terrestrial; marine; brackish \\ \hline
+Filter      & Taxon       & Continent              & North America; Asia; Europe \\ \hline
+Filter      & Test        & Duration               & enter duration in hours \\ \hline
+Filter      & Test        & Effect groups          & MOR - Mortality; POP - Population \\ \hline
+Filter      & Test        & Endpoins               & EC50, LOEC, LOEL \\ \hline
+Filter      & Checks      & Water solubility check & yes/no \\ \hline
+Filter      & Checks      & Remove outliers        & yes/no \\ \hline
+Aggregation & Aggregation & Aggregate              & Minimum; Maximum; Median \\ \hline
+\end{tabular}
+\caption{\etoxbase GUI paramters to choose.}
+\label{tab:inputs}
+\end{table}
+
+Generally it has to be stated here that the returned values are subject to change, as new tests will be constantly incorporated into the \etoxbase{}.
 However, we expect values of well tested chemical-organism combinations (i.e.
@@
 Yet, returned values from the \etoxbase{} are certainly influenced by individual tests procedures and characteristics, such as pH, temperature or conductivity to just name a few.
-However, using the geometric mean as an aggregate allows for an adequate estimation of the central tendency of the toxicity of a chemical towards an organism group.
-The geometric mean is in comparison to the arithmetic mean considerably less influenced by outliers and is suitable for skewed data.
+However, using the median or the geometric mean as an aggregate allows for an adequate estimation of the central tendency of the toxicity of a chemical towards an organism group.
+The geometric mean is recommended, as it is in comparison to the arithmetic mean considerably less influenced by outliers and is suitable for skewed data.
 Also, the geometric mean is preferable in contrast to the median, since the median completely ignores the influence of large or small values, making it unreliable for small data sets <CITATION GEOMETRIC MEAN>.
@@
 A lot of important information is still confined in continuous text of articles and therefore not ascertainable for large scale analyses.
-
-PUT SSD HERE
-
-When using the data from the tool it is essential to think about which endpoint groups (e.g.
-EC50, LC50, NOEC etc.) and which effect measurements to use (e.g.
-MOR - Mortality, ITX - Intoxication, GRO - Growth etc.).
-
-Likewise the data of the \epa{} doesn't allow for the in ecotoxicology so frequently used distinction of acute and chronic tests. Even the OECD guidlines are presented only as pdfs inproperly.
-
-discuss flaws in column types (e.g some concentration entries contain '+' or '~' and are therefore not convertable to numeric) - This hampers large scale analysis greatly!
-
-in the futrue there will probably more of such resources. Als NORMAN which even performs encompasses expert judgment on the validity of tests
 
@@
 
-\begin{table}[]
-\begin{tabular}{|l|l|l|l|}
-\hline
-Procedure   & Group       & Subgroups              & Options \\ \hline
-Filter      & Compound    & CAS                    & input CAS \\ \hline
-Filter      & Compound    & Concentration type     & Active Ingredient; Formulation; Total \\ \hline
-Filter      & Compound    & Chemical class         & Insecticides; Herbicides; Metals \\ \hline
-Filter      & Taxon       & Taxon                  & Choose a taxon \\ \hline
-Filter      & Taxon       & Organism habitat       & freshwater; terrestrial; marine; brackish \\ \hline
-Filter      & Taxon       & Continent              & North America; Asia; Europe \\ \hline
-Filter      & Test        & Duration               & enter duration in hours \\ \hline
-Filter      & Test        & Effect groups          & MOR - Mortality; POP - Population \\ \hline
-Filter      & Test        & Endpoins               & EC50, LOEC, LOEL \\ \hline
-Filter      & Checks      & Water solubility check & yes/no \\ \hline
-Filter      & Checks      & Remove outliers        & yes/no \\ \hline
-Aggregation & Aggregation & Aggregate              & Minimum; Maximum; Median \\ \hline
-\end{tabular}
-\caption{\etoxbase input possibilities}
-\label{tab:inputs}
-\end{table}
-
-\begin{table}[]
+\begin{table}[h!]
 \begin{tabular}{|l|l|}
@@
 Script                      & Description                                      \\ \hline
-run_build.R                 & runs all sub scripts                             \\ \hline
+run\_build.R                 & runs all sub scripts                             \\ \hline
 setup.R                     & sets up the environment                          \\ \hline
@@
 da\_epa3.R                  & reduce EPA Ecotox data                           \\ \hline
+qu\_aw.R                    & query Compendium of Pesticide Common Names       \\ \hline
+qu\_chemspider\_scrape.R    & query Chemspider data base                       \\ \hline
+qu\_eurostat\_chem\_class.R & query Eurostat data set                          \\ \hline
 qu\_pc.R                    & query PubChem data base                          \\ \hline
-qu\_aw.R                    & query Compendium of Pesticide Common Names       \\ \hline
 qu\_pp.R                    & query Physprop data base                         \\ \hline
-qu\_eurostat\_chem\_class.R & query Eurostat data set                          \\ \hline
-qu\_chemspider\_scrape.R    & query Chemspider data base                       \\ \hline
 qu\_worms2.R                & query WORMS data base                            \\ \hline
@@
 
-\begin{table}[]
+\begin{table}[h!]
 \begin{tabular}{|l|l|}
@@
 
+\item acute vs chronic
+
+\item OECD guidlines are not machine readable
+
+\item discuss flaws in column types (e.g some concentration entries contain '+' or '~' and are therefore not convertable to numeric) - This hampers large scale analysis greatly!
+
 \pagebreak

### 2019-02-09 article/tikz/pipeline_organigram.tex

--- 2018-12-17 article/tikz/organigram.tex
@@
     	child { node [children] { Alan Wood Pesticide Compendium - \textbf{qu\_aw.R } }}
-    	child { node [children] { Pesticide Action Network - \textbf{qu\_pan.R} }}
-    	child { node [children] { PhysProp - \textbf{qu\_pp.R} }}
-    	child { node [children] { PubChem - \textbf{qu\_pubchem.R} }};
+    	child { node [children] { Chemspider - \textbf{qu\_chemspider\_scrape.R } }}
+    	child { node [children] { Eurostat - \textbf{qu\_eurostat\_chem\_class.R} }}
+    	child { node [children] { PubChem - \textbf{qu\_pc.R} }}
+    	child { node [children] { PhysProp - \textbf{qu\_pp.R} }};
 	\node (taxon) [ids, below of=cas, yshift=-6cm, align = center] {Taxon};
@@
 	\node (habitat) [main, right of = taxon, xshift = 4.5cm, yshift = 1.5cm] {Habitats}
-    	child { node [children] { WORMS data base - \textbf{qu\_worms.R} }}
-    	child { node [children] { Manual Search - \textbf{qu\_manual.R} }};
+    	child { node [children] { WORMS data base - \textbf{qu\_worms2.R} }};
 	\node (region) [main, right of = taxon, xshift = 4.5cm, yshift = -1.5cm] {Regions}
@@
 	\node (checksres) [main, left of = analysesres, xshift = -4.5cm] {checks}
-	    child { node [children]  { \textbf{re\_checks.R} }};
+	    child { node [children]  { \textbf{re\_checks\_internal.R} }};
 	\node (finalres) [main, left of = checksres, xshift = -4.5cm] {Final Table}

### 2019-02-09 article/article.tex
Updates from Overleaf

@@
 
+\pagebreak
+
 \section*{Background \& Summary [max 700 words]}
@@
 Additionally the \etoxbase{} queries chemical- and organism-specific variables from other open data bases to fill information gaps such as water solubility, organism habitat or regional occurrence patterns in the ECOTOX data base.
+
+\pagebreak
 
@@
 
+\pagebreak
+
 \section*{Data Records}
@@
 
+\pagebreak
+
 \section*{Technical Validation}
@@
 <PUT COMPARISON WITH PPDB DATA HERE?>
+
+\pagebreak
 
@@
 
-Generally it has to be stated here that the returned values are subject to change, as new tests will be constantly incorporated into the \etoxbase{}.
-However, we expect values of well tested chemical-organism combinations (i.e.
-high N <MAYBE PUT NUMBER HERE - CITATION>) to be relatively stable.
-Yet, returned values from the \etoxbase{} are certainly influenced by individual tests procedures and characteristics, such as pH, temperature or conductivity to just name a few.
+\subsection*{Limitations}
+
+Generally it has to be stated that the returned values are subject to change, as new tests will be constantly incorporated into the \etoxbase{}.
+However, we expect values of well tested chemical-organism combinations to be relatively stable.
+The returned values are certainly influenced by individual tests characteristics, such as pH, temperature or conductivity to just name a few, which influences the effect relationship between chemicals and organisms \citep{rosenkrantz_influence_2013,li_temperature_2011}.
+Unfortunately the EPA ECOTOX data base has only rare records on test parameters, which prevents incorporating such parameters in an aggregation <proportions of available data: Temperature:77\%, pH:56\%, hardness:27\%, dissolved oxygen:18\%, Alkalinity:15\%, salinity:9\%, others:<=5\%>.
 However, using the median or the geometric mean as an aggregate allows for an adequate estimation of the central tendency of the toxicity of a chemical towards an organism group.
@@
 
-\subsection*{Limitations}
-
-SSDs aim to extrapolate from species to the community level while not ignoring environmental variables (e.g.
-pH, temperature, etc.) which can effect the toxicity of a specific chemical greatly \citep{posthuma_species_2002}.
-For a large number of the \epa{} tests, environmental variables are not recorded (See \ref{table:meta-variables} for proportions of missing entries per variable).
-Hence to some extent the reported test results inherit uncertainty.
-In this context we also call for more emphasis on machine-readability when publishing ecotoxicological test results <CITATION ON AS OF HOW MANY DATA POINTS MACHINES ARE NEEDED>.
-A lot of important information is still confined in continuous text of articles and therefore not ascertainable for large scale analyses.
+\pagebreak
 
@@
 
-\section*{Figure Legends}
+\pagebreak
 

### 2019-06-11 article/parts/04_technical_validation.tex
split up article text

--- 2019-02-08 article/article.tex
@@
-\documentclass[english]{article}
-\usepackage[T1]{fontenc}
-\usepackage{natbib} % citations
-\usepackage{csvsimple}
-\usepackage{pgfplotstable}
-\usepackage{rotating} % to rotate tables
-%% tikz packages
-\usepackage{tikz}
-\usepackage{forest}
-\usetikzlibrary{trees, arrows, shapes.geometric, positioning}
-\usepackage{standalone}
-%% text packages
-\usepackage{siunitx} % fake ° symbol \ang{}
-\usepackage[super]{nth} % for 1st, 2nd etc. \nth{}
-\usepackage{hyperref}
-
-\newcommand{\etoxbase}{Etox-Base}
-\newcommand{\epa}{EPA ECOTOX data base}
-\newcommand{\app}{http://139.14.20.252:3838/etox-base-shiny/}
-\newcommand{\git}{https://github.com/andreasLD/etox-base}
-\newcommand{\gitapp}{https://github.com/andreasLD/etox-base-shiny}
-\newcommand{\ecfifty}{EC\textsubscript{50}}
-
-\listfiles
-
-\begin{document}
-
-\title{Etox-Base: A tool to derive exposure endpoints from chemical test results}
-
-\author{Andreas Scharm{\"u}ller\textsuperscript{1{*}},
-        Verena Schreiner\textsuperscript{1},
-        Ralf B. Sch{\"a}fer\textsuperscript{1}}
-
-\maketitle
-\thispagestyle{fancy}
-
-1. Institute of Environmental Sciences, University of Koblenz-Landau Fortstraße 7, 76829 Landau, Germany {*}corresponding author(s):
-Andreas Scharm{\"u}ller (scharmueller@uni-landau.de)
-
-\begin{abstract}
-
-<WRITE ABSTRACT>
-
-\end{abstract}
-
-\section*{Background \& Summary [max 700 words]}
-
-A large number of chemicals such as pharmaceuticals, pesticides and synthetic hormones are in daily use all over the world with insufficient knowledge about possible adverse effects on the environment.
-In Europe alone some 100,000 chemicals are estimated to be in current use, whereof 30,000 are produced in quantities larger than one ton per year \citep{breithaupt_costs_2006}.
-Chemicals are brought to the environment deliberately - as it is the case of pesticides, or arrive therein as a byproduct from other processes (e.g.
-atmospheric emissions or wastewater) \citep{schwarzenbach_challenge_2006}.
-Ecosystems provide essential services to human societies such as drinking and irrigation water, food and climate regulation.
-These services are products of different ecosystem functions that crucially depend on the integrity of the populations and communities that drive these ecosystem functions.
-However, besides habitat degradation, climate change and nutrient enrichment, pollution with man-made chemical toxicants threatens these populations and communities in various ways which are currently not fully understood \citep{steffen_anthropocene_2007}.
-Pollution with man-made chemical toxicants was indeed identified as one of three major environmental problems for which research gaps hamper the derivation of planetary boundaries, i.e.
-thresholds beyond which irreversible state shifts may occur \citep{steffen_anthropocene_2007}.
-Bernhardt et al.
-\citet{bernhardt_synthetic_2017} argue further that the knowledge gap how chemicals effect populations and communities and hence ecosystem functions and ecosystem services, would also impede society’s ability to accomplish the Sustainable Development Goals of the United Nations.
-According to Breithaupt \citet{breithaupt_costs_2006} less than one percent of chemicals released to the environment are thoroughly tested.
-Although for some chemicals advanced and standardized \citep{oecd_oecd_2018} test methods have been developed, ecotoxicological tests for a lot of chemicals, especially newly emerging ones remain scarce and if existent, the information are represented sparsely and in inconsistent formats.
-\citep{gessner_fostering_2016}.
-Therefore we developed a tool that collects, processes and aggregates ecotoxicological tests results to then represent them in an harmonized form in a web tool - the \etoxbase tool: \href{http://139.14.20.252:3838/etox-base-shiny/}.
-
-As a resource for ecotoxicological tests we used the ECOTOXicology knowledgebase (ECOTOX) created by the U.S.
-Environmental Protection Agency (EPA), which collects raw, non harmonized ecotoxicological test data on on aquatic and terrestrial wildlife as well as plants and publishes it on a quarterly basis.
-By the time of writing it contained 926,108 test results, comprising 11,685 chemicals tested on 12,668 different species \citep{elonen_ecotoxicology_2018}.
-The collected ecotoxicological tests contain all sorts of measured endpoints such as Effective Concentrations (\ecfifty{}) values, No-observed effect concentrations (NOEC) or lowest observed effect concentrations (LOEC).
-Likewise the data set contains different effect measures such as mortality, intoxication and growth as well as test durations ranging from seconds to weeks or even years.
-
-The \etoxbase aggregates this information according to the user's inputs in order to provide reliable concentrations for the derivation of toxicity effect measures such as Species Sensitivity Distributions (SSDs) \citep{posthuma_species_2002} and Toxic Units (TUs), which represent two prominent concepts in ecotoxicology to assess effects on organisms.
-The former combines toxicity data of several organism groups towards a chemicals to estimate effects on biotic communities, the latter refers to effects of a specific organism (group) towards a chemical.
-The two concepts are widely used in ecotoxicology \citep{kefford_definition_2011, schafer_effects_2011} since the allow for a comparison of toxicities across multiple chemicals and biological communities, respectively.
-Although performed on one and the same organism and chemical, outcomes of multiple ecotoxicological tests can vary greatly due to different test parameters, different populations or other individual factors which in turn leads to ambiguities in choosing test results for proper risk assessment.
-
-The \etoxbase tool therefore aggregates test results to obtain single toxicity values for organism, chemical and test duration combinations.
-The tool downloads every new version of the \epa{} and performs several quality checks, e.g.
-unit harmonization, and cleaning steps on it.
-In addition to the test parameters that are retrieved from the \epa{} the tool includes chemical and organism parameters from other open data bases to address missing information such as water solubility, organism habitat or regional occurrence patterns.
-
-\section*{Methods}
-
-%%The Methods should include detailed text describing any steps or procedures 
-%%used in producing the data, including full descriptions of the experimental 
-%%design, data acquisition assays, and any computational processing (e.g. 
-%%normalization, image feature extraction). Related methods should be grouped 
-%%under corresponding subheadings where possible, and methods should be described 
-%%in enough detail to allow other researchers to interpret and repeat, if required, 
-%%the full study. Specific data outputs should be explicitly referenced via data 
-%%citation (see Data Records and Data Citations, below). Authors should cite 
-%%previous descriptions of the methods under use, but ideally the method 
-%%descriptions should be complete enough for others to understand and reproduce 
-%%the methods and processing steps without referring to associated publications. 
-%%There is no limit to the length of the Methods section.
-
-The \etoxbase{} consists of two parts of software: Firstly a processing pipeline \ref{fig:pipeline} downloads quarterly released new versions of the \epa{} and performs several preparation steps in R \citep{r_core_team_r_2017}, to then store the outcome in a PostgreSQL data base.
-This constitutes the basis upon which the R shiny web application is accessing the data \citep{r_core_team_r_2017}.
-The application itself is set of R functions that filter and aggregate the data according to a user's inputs \ref{fig:app}.
-Finally a data set, which can be downloaded and an interactive plot, for data exploration are returned.
-In the following the two parts are explained in detail.
-
-\subsection*{Data acquisition \& preparation}
-
-The data is downloaded (bd\_epa\_download.R) and then built into a PostgreSQL data base (bd\_epa\_postgres.R).
-Subsequently the data is imported into R for pre-processing (da\_epa1.R, da\_epa2.R, da\_epa3.R), which includes validity checks, removal of special characters, unit harmonizations and selection of relevant variables.
-Chemical abstract service (CAS) numbers, unique chemical identifiers together we taxon names are then used to query other data bases for additional information on chemicals and biota respectively.
-Chemical information is retrieved from the Pubchem data base \citep{CITE_PUBCHEM} (qu\_pc.R), from the Compendium of Pesticide Common Names \citep{CITE_AW} (qu\_aw.R), from the Physprop data base \citep{CITE_PHYSPROP} (qu\_pp.R), from EUROSTAT and the Chemspider data base \citep{CITE-CHEMSPIDER}.
-These queries mostly rely on the webchem R-package \citep{szocs_webchem_2015-1}.
-For biota, habitat and occurrence information is queried from the World Register of Marine Species (WORMS) \citep{WORMS} and from the Global Biodiversity Information Facility (GBIF) \citep{CITE_RGBIF} using the rgbif R-package \citep{chamberlain_rgbif_2018}.
-The acquired information is then merged (re\_merge.R), variables are combined and created (re\_combine.R), checked (re\_checks\_internal.R) and finally compiled in one table (re\_final.R) This table is then used in the application to perform search queries on.
-For a detailed overview of what information is collected see table \ref{tab:scripts-pipeline}.
-
-\begin{figure}
-    \includestandalone[width=\textwidth, scale=0.1]{tikz/pipeline_organigram}
-    \caption{Processing Pipeline}
-    \label{fig:pipeline}
-\end{figure}
-
-\subsection*{The application}
-The application is accessible through \app{} and was built in R by using the shiny web application framework \citep{chang_shiny_2018}.
-Users can define several input parameters \ref{tab:inputs} in the application's graphical user interface (GUI), handeled in the ui.R script in order to adjust the aggregation according specific requirements.
-In doing so users execute R functions (fun\_filter.R, fun\_aggregation.R, fun\_filagg\_plot\_ly.R) in the server.R script, which produce a 'filtered' and an 'aggregated' data set as well as an interactive overview plot.
-The 'aggregated' data set is presented in the GUI.
-Both data sets can be downloaded via a download button.
-
-\begin{figure}
-    \includestandalone[scale=0.75]{tikz/application}
-    \caption{Application}
-    \label{fig:app}
-\end{figure}
-
-\subsection*{Aggregating the data}
-In order to be able to calculate TUs and SSDs, derivation of single endpoint values for specific organism groups and a chemicals of the partly extensive test data, the \etoxbase{} aggregates the test results according to the chosen filters in two steps.
-Firstly the filtered test results are aggregated by the CAS number, the chosen taxon plus the duration and secondly only by the CAS number.
-The former can't be influenced by the user and calculates either the minimum or the median depending on the amount of results to aggregate (n <= 2: minimum, if n > 2: median).
-Thereof the second step calculates the minimum, the maximum, the median, the geometric mean, or the arithmetic mean as an aggregate.
-Besides the aggregation two cleaning can also be performed.
-Firstly the application allows to exclude test results exhibiting concentrations that are higher than the actual water solubility of the respective chemical at \ang{20} C.
-Secondly the user can choose to exclude outliers.
-The outliers are selected as values that exceed lower (0.25) and the upper (0.75) quartile by 1.5 times the inter-quartile range.
-All the steps described in this section are selectable in the GUI and are executed at each click by the server.
-(cf.
-table \ref{tab:scripts-app}).
-
-\subsection*{Code availability}
-%%For all studies using custom code in the generation or processing of datasets, 
-%%a statement must be included here, indicating whether and how the code can be 
-%%accessed, including any restrictions to access. This section should also include 
-%%information on the versions of any software used, if relevant, and any specific 
-%%variables or parameters used to generate, test, or process the current dataset.
-
-The code for processing the data is stored in a Github reopsitory (\git{}).
-All processing scripts were written in R version 3.4.4.
-PostgreSQL 9.5 has been used to build the data base.
-Hence, R together with its packages \ref{tab:rpackages} and PostgreSQL suffice to rebuild the process.
-The \app{} itself is stored in a second Github repostitory (\gitapp{}).
-Since new versions of the \epa{} are published on a quarterly basis the data acquisition process is automated and new versions of the data and the \app{} are scheduled to be released regularly.
-
-\begin{table}[]
-    \centering
-    \begin{tabular}{|l|l|l|l|}
-        \hline
-        Category        & Package       & Version   & Reference                     \\ \hline
-        processing      & RCurl         & 1.95-4.11 & \citep{lang_rcurl_2018}       \\ \hline
-        processing      & data.table    & 1.12.0    & \citep{dowle_data.table_2018} \\ \hline
-        application     & shiny         & 1.1.0     & \citep{chang_shiny_2018}      \\ \hline
-    \end{tabular}
-    \caption{List of R packages used in compiling the data.\newline{}
-    TODO: Automate version input and name generation - whole table\newline{}
-    TODO2: debug chang\_shiny\_2018 reference}
-    \label{tab:rpackages}
-\end{table}
-
-\section*{Data Records}
-%%Please explain each data record associated with this work, including
-%%the repository where this information is stored, and an overview of
-%%the data files and their formats. Each external data record should
-%%be listed in Data Citation section at the end of this template, and 
-%%records should be cited throughout the manuscript as, for example 
-%%(Data Citation 1). 
-%%
-%%Tables should be used to support the data records, and should clearly indicate 
-%%the samples and subjects, their provenance, and the experimental manipulations 
-%%performed on each. They should also specify the data output resulting from each 
-%%data-collection or analytical step, should these form part of the archived record. 
-%%Please see the submission guidelines at the \emph{Scientific Data} website, and 
-%%our Word templates for more information on preparing such tables.
-
-\section*{Technical Validation}
 %%This section presents any experiments or analyses that are needed
@@
 
+\section{Technical validation}
+
 <PUT COMPARISON WITH PPDB DATA HERE?>
-
-\section*{Usage Notes}
-%%Brief instructions that may help other researchers reuse these dataset.
-%%This is an optional section, but strongly encouraged when helpful
-%%to readers. This may include discussion of software packages that
-%%are suitable for analyzing the assay data files, suggested downstream
-%%processing steps (e.g. normalization, etc.), or tips for integrating
-%%or comparing this with other datasets. If needed, authors are encouraged
-%%to provide code, programs, or data processing workflows when they may help 
-%%others analyse the data. We encourage authors to archive related code in 
-%%a DOI-issuing archive when possible, but code may also be supplied as 
-%%supplementary information files. 
-%%
-%%For studies involving privacy or safety controls on public access
-%%to the data, this section should describe in detail these controls,
-%%including how authors can apply to access the data, and what criteria
-%%will be used to determine who may access the data, and any limitations
-%%on data use.
-
-The \etoxbase{} can be accessed via an application here: \app{}.
-The user can filter among different test parameters \ref{tab:inputs}.
-The \etoxbase{} is designed to support ecotoxicologists in assessing the risk of chemicals on the environment.
-With an increase of ecotoxicological test data available, it becomes fundamental to provide such information in reasonable formats, meaning accessible for humans as well as machines.
-Currently chemical risk assessment relies on test results of a few, well tested organism groups, such as \textit{Daphnia magna}, \textit{Pimephales promelas}, \textit{Selenastrum capricornutum} reported in publications on test procedures <CITE>, compiled and published data sets \citep{malaj_organic_2014} or in data bases such as the Pesticide Property Data Base (PPDB) \citep{lewis_international_2016}.
-Each data source having its limitations: Collecting toxicity information from individual publications is laborious and using already compiled data from publications or data bases is often limited to specific chemical groups, such as pesticides, biozides etc.
-The \etoxbase{}, making use of the \epa{}, the largest publicly available collection of ecotoxicological test results, aims to provide holistic aggregates (\ecfifty{}, NOEC, LOEC) of ecotoxicological tests results in adequate formats in order to support researchers in fetching ecotoxicological data for the calculation of SSDs, TUs and other risk indicators.
-SSDs try to extrapolate effects of chemicals towards organisms from species to a community level by fitting a statistical distribution to toxicity test results on organisms of different trophic levels \citep{posthuma_species_2002}.
-As an example
-
-\begin{figure}
-    \dentering
-    \includegraphics[width=1\linewidth]{figures/ssd2}
-
-\end{figure}
-
-Generally it has to be stated here that the returned values are subject to change, as new tests will be constantly incorporated into the data set.
-However, we expect values of well tested chemical-organism combinations (i.e.
-high N <MAYBE PUT NUMBER HERE - CITATION>) to be relatively stable.
-Yet, returned values from the \etoxbase{} are certainly influenced by individual tests procedures and characteristics, such as pH, temperature or conductivity to just name a few.
-However, using the geometric mean as an aggregate allows for an adequate estimation of the central tendency of the toxicity of a chemical towards an organism group.
-The geometric mean is in comparison to the arithmetic mean considerably less influenced by outliers and is suitable for skewed data.
-Also, the geometric mean is preferable in contrast to the median, since the median completely ignores the influence of large or small values, making it unreliable for small data sets <CITATION GEOMETRIC MEAN>.
-
-Other initiatives, with different, though partly overlapping aims accumulating existing ecotoxicological test data are published online.
-The PPDB provides validated ecotoxicological test results for pesticides, narrowing chemicals to this group \citep{lewis_international_2016}.
-The Network of reference laboratories, research centres and related organisations for monitoring of emerging environmental substances (NORMAN) focuses on assembling river basin specific pollutants\citep{von_der_ohe_new_2011}.
-The EnviroTox data base \href{https://envirotoxdatabase.org/} makes also use of the \epa{} data in order to derive Predicted no-effect concentrations (PNEC) \citep{health_and_environmental_sciences_institute_hesi_envirotox_2019}.
-In comparison to the \etoxbase, these examples partly rely on the same data, yet not aiming to derive \ecfifty, LOEC and NOEC from it automatically.
-
-\subsection*{Limitations}
-
-SSDs aim to extrapolate from species to the community level while not ignoring environmental variables (e.g.
-pH, temperature, etc.) which can effect the toxicity of a specific chemical greatly \citep{posthuma_species_2002}.
-For a large number of the \epa{} tests, environmental variables are not recorded (See \ref{table:meta-variables} for proportions of missing entries per variable).
-Hence to some extent the reported test results inherit uncertainty.
-In this context we also call for more emphasis on machine-readability when publishing ecotoxicological test results <CITATION ON AS OF HOW MANY DATA POINTS MACHINES ARE NEEDED>.
-A lot of important information is still confined in continuous text of articles and therefore not ascertainable for large scale analyses.
-
-PUT SSD HERE
-
-When using the data from the tool it is essential to think about which endpoint groups (e.g.
-EC50, LC50, NOEC etc.) and which effect measurements to use (e.g.
-MOR - Mortality, ITX - Intoxication, GRO - Growth etc.).
-
-Likewise the data of the \epa{} doesn't allow for the in ecotoxicology so frequently used distinction of acute and chronic tests. Even the OECD guidlines are presented only as pdfs inproperly.
-
-discuss flaws in column types (e.g some concentration entries contain '+' or '~' and are therefore not convertable to numeric) - This hampers large scale analysis greatly!
-
-in the futrue there will probably more of such resources. Als NORMAN which even performs encompasses expert judgment on the validity of tests
-
-\section*{Acknowledgements}
-%%Text acknowledging non-author contributors. Acknowledgements should
-%%be brief, and should not include thanks to anonymous referees and
-%%editors, or effusive comments. Grant or contribution numbers may be
-%%acknowledged. Author contributions Please describe briefly the contributions
-%%of each author to this work on a separate line. 
-%%
-%%AK did this and that. 
-%%
-%%BG did this and that and the other.
-
-\section*{Competing financial interests}
-%%A competing financial interests statement is required for all accepted
-%%papers published in \emph{Scientific Data}. If none exist simply write,
-%%``The author(s) declare no competing financial interests''.
-
-\section*{Figures Legends}
-%%Figure should be referred to using a consistent numbering scheme through
-%%the entire Data Descriptor. For initial submissions, authors may choose
-%%to supply this document as a single PDF with embedded figures, but
-%%separate figure image files must be provided for revisions and accepted
-%%manuscripts. In most cases, a Data Descriptor should not contain more
-%%than three figures, but more may be allowed when needed. We discourage
-%%the inclusion of figures in the Supplementary Information \textendash{}
-%%all key figures should be included here in the main Figure section. 
-%%
-%%Figure legends begin with a brief title sentence for the whole figure
-%%and continue with a short description of what is shown in each panel,
-%%as well as explaining any symbols used. Legend must total no more
-%%than 350 words, and may contain literature references.
-
-\section*{Tables}
-%%Tables supporting the Data Descriptor. These can provide summary information
-%%(sample numbers, demographics, etc.), but they should generally not
-%%be used to present primary data (i.e. measurements). Tables containing
-%%primary data should be submitted to an appropriate data repository. 
-%%
-%%Tables may be provided within the \LaTeX{} document or as separate
-%%files (tab-delimited text or Excel files). Legends, where needed,
-%%should be included here. Generally, a Data Descriptor should have
-%%fewer than ten Tables, but more may be allowed when needed. Tables
-%%may be of any size, but only Tables which fit onto a single printed
-%%page will be included in the PDF version of the article (up to a maximum
-%%of three).
-
-%\begin{sidewaystable}
-%    \csvautotabular[respect all]{data/meta.csv};
-%    \caption{Meta variables table};
-%    \label{table:meta-variables};
-%\end{sidewaystable}
-
-\begin{table}[]
-\begin{tabular}{|l|l|l|l|}
-\hline
-Procedure   & Group       & Subgroups              & Options \\ \hline
-Filter      & Compound    & CAS                    & input CAS \\ \hline
-Filter      & Compound    & Concentration type     & Active Ingredient; Formulation; Total \\ \hline
-Filter      & Compound    & Chemical class         & Insecticides; Herbicides; Metals \\ \hline
-Filter      & Taxon       & Taxon                  & Choose a taxon \\ \hline
-Filter      & Taxon       & Organism habitat       & freshwater; terrestrial; marine; brackish \\ \hline
-Filter      & Taxon       & Continent              & North America; Asia; Europe \\ \hline
-Filter      & Test        & Duration               & enter duration in hours \\ \hline
-Filter      & Test        & Effect groups          & MOR - Mortality; POP - Population \\ \hline
-Filter      & Test        & Endpoins               & EC50, LOEC, LOEL \\ \hline
-Filter      & Checks      & Water solubility check & yes/no \\ \hline
-Filter      & Checks      & Remove outliers        & yes/no \\ \hline
-Aggregation & Aggregation & Aggregate              & Minimum; Maximum; Median \\ \hline
-\end{tabular}
-\caption{\etoxbase input possibilities}
-\label{tab:inputs}
-\end{table}
-
-\begin{table}[]
-\begin{tabular}{|l|l|}
-\hline
-Script                      & Description                                      \\ \hline
-run_build.R                 & runs all sub scripts                             \\ \hline
-setup.R                     & sets up the environment                          \\ \hline
-bd\_epa\_download.R         & download EPA Ecotox data base                    \\ \hline
-bd\_epa\_postgres.R         & build EPA Ecotox data base locally               \\ \hline
-da\_epa1.R                  & raw EPA Ecotox data export                       \\ \hline
-da\_epa2.R                  & process EPA Ecotox data                          \\ \hline
-da\_epa3.R                  & reduce EPA Ecotox data                           \\ \hline
-qu\_pc.R                    & query PubChem data base                          \\ \hline
-qu\_aw.R                    & query Compendium of Pesticide Common Names       \\ \hline
-qu\_pp.R                    & query Physprop data base                         \\ \hline
-qu\_eurostat\_chem\_class.R & query Eurostat data set                          \\ \hline
-qu\_chemspider\_scrape.R    & query Chemspider data base                       \\ \hline
-qu\_worms2.R                & query WORMS data base                            \\ \hline
-qu\_gbif.R                  & query GBIF data base                             \\ \hline
-re\_merge.R                 & merge all data                                   \\ \hline
-re\_combine.R               & combine information                              \\ \hline
-re\_analyses.R              & ???                                              \\ \hline
-re\_checks\_internal.R      & ???                                              \\ \hline
-re\_final.R                 & create and write final table                     \\ \hline
-wr\_dbase.R                 & write final table to data base                   \\ \hline
-wr\_app.R                   & write final table and variables to app directory \\ \hline
-wr\_meta.R                  & ???                                              \\ \hline
-wr\_stats.R                 & ???                                              \\ \hline
-\end{tabular}
-\caption{Scripts that are used to process the data}
-\label{tab:scripts-pipeline}
-\end{table}
-
-\begin{table}[]
-\begin{tabular}{|l|l|}
-\hline
-Script                     & Description                                               \\ \hline
-setup.R                    & sets up the environment                                   \\ \hline
-ui.R                       & creates the graphical user interface of the application   \\ \hline
-server.R                   & handles functions and data input and output               \\ \hline
-fun\_filter.R              & function that filters the data                            \\ \hline
-fun\_aggregation.R         & function that performs the aggregation                    \\ \hline
-fun\_filagg\_plot\_ly.R    & creates interactive plot                                  \\ \hline
-\end{tabular}
-\caption{Scripts that used for the web application}
-\label{tab:scripts-app}
-\end{table}
-
-\section*{Figure Legends}
-
-\bibliographystyle{apalike}
-\bibliography{refs/references-etox-base.bib,refs/references-etox-base-rpackages.bib,refs/references-etox-base-rshinypackages.bib}
-% multiple refs have to be written w/o space
-
-\section*{Data Citations}
-
-\pagebreak
-\section*{TODO}
-
-\begin{itemize}
-
-\item compare Etox-BAse output with PPDB
-
-\item plot example SSD and TU
-
-\pagebreak
-\section*{MISC}
-
-\begin{figure}
-    \includestandalone[scale=0.5]{tikz/tree_source}
-    \caption{Alternative processing pipeline}
-    \label{fig:pipeline-tree}
-\end{figure}
-
-\section*{not included citations}
-Put in citation \citep{hartung_chemical_2009} ?
-which claims that REACH won't meet their assumptions and say 54 million vertebrate animals are needed for tests that would cost €9.5 billion over the next ten years.
-I.e 20 times more animals, 6 times the costs in comparison to the official estimates
-
-\end{itemize}
-
-\end{document}

### 2019-06-11 article/parts/05_usage_notes.tex
split up article text

--- 2019-02-09 article/article.tex
@@
-\documentclass[english]{article}
-\usepackage[T1]{fontenc}
-\usepackage{natbib} % citations
-\usepackage{csvsimple}
-\usepackage{pgfplotstable}
-\usepackage{rotating} % to rotate tables
-%% tikz packages
-\usepackage{tikz}
-\usepackage{forest}
-\usetikzlibrary{trees, arrows, shapes.geometric, positioning}
-\usepackage{standalone}
-%% text packages
-\usepackage{siunitx} % fake ° symbol \ang{}
-\usepackage[super]{nth} % for 1st, 2nd etc. \nth{}
-\usepackage{hyperref}
-
-\newcommand{\etoxbase}{Etox-Base}
-\newcommand{\epa}{EPA ECOTOX data base}
-\newcommand{\app}{http://139.14.20.252:3838/etox-base-shiny/}
-\newcommand{\git}{https://github.com/andreasLD/etox-base}
-\newcommand{\gitapp}{https://github.com/andreasLD/etox-base-shiny}
-\newcommand{\ecfifty}{EC\textsubscript{50}}
-
-\listfiles
-
-\begin{document}
-
-\title{Etox-Base: A tool to derive exposure endpoints from chemical test results}
-
-\author{Andreas Scharm{\"u}ller\textsuperscript{1{*}},
-        Verena Schreiner\textsuperscript{1},
-        Ralf B. Sch{\"a}fer\textsuperscript{1}}
-
-\maketitle
-\thispagestyle{fancy}
-
-1. Institute of Environmental Sciences, University of Koblenz-Landau Fortstraße 7, 76829 Landau, Germany {*}corresponding author(s):
-Andreas Scharm{\"u}ller (scharmueller@uni-landau.de)
-
-\begin{abstract}
-
-<WRITE ABSTRACT>
-
-\end{abstract}
-
-\pagebreak
-
-\section*{Background \& Summary [max 700 words]}
-
-A large number of chemicals such as pharmaceuticals, pesticides and synthetic hormones are in daily use all over the world with insufficient knowledge about possible adverse effects on the environment.
-In Europe alone some 100,000 chemicals are estimated to be in current use, whereof 30,000 are produced in quantities larger than one ton per year \citep{breithaupt_costs_2006}.
-Chemicals are brought to the environment deliberately - as it is the case of pesticides, or arrive therein as a byproduct from other processes (e.g.
-atmospheric emissions or wastewater) \citep{schwarzenbach_challenge_2006}.
-Ecosystems provide essential services to human societies such as drinking and irrigation water, food and climate regulation.
-These services are products of different ecosystem functions that crucially depend on the integrity of the populations and communities that drive these ecosystem functions.
-However, besides habitat degradation, climate change and nutrient enrichment, pollution with man-made chemical toxicants threatens these populations and communities in various ways which are currently not fully understood \citep{steffen_anthropocene_2007}.
-Pollution with man-made chemical toxicants was indeed identified as one of three major environmental problems for which research gaps hamper the derivation of planetary boundaries, i.e.
-thresholds beyond which irreversible state shifts may occur \citep{steffen_anthropocene_2007}.
-Bernhardt et al.
-\citet{bernhardt_synthetic_2017} argue further that the knowledge gap how chemicals effect populations and communities and hence ecosystem functions and ecosystem services, would also impede society’s ability to accomplish the Sustainable Development Goals of the United Nations.
-According to Breithaupt \citet{breithaupt_costs_2006} less than one percent of chemicals released to the environment are thoroughly tested.
-Although for some chemicals advanced and standardized \citep{oecd_oecd_2018} test methods have been developed, ecotoxicological tests for a lot of chemicals, especially newly emerging ones remain scarce and if existent, the information is represented sparsely and in inconsistent formats.
-\citep{gessner_fostering_2016}.
-Therefore we developed a data base that collects, processes and aggregates ecotoxicological tests results in order to subsequently publish it in an harmonized form as a web application - the \etoxbase tool: \href{http://139.14.20.252:3838/etox-base-shiny/}.
-This should facilitate researchers the access to ecotoxicological test results and support modern chemical risk assessment (CRA) workflows.
-The \etoxbase makes use of ecotoxicological test data, obtained from the ECOTOXicology knowledgebase (ECOTOX) created by the U.S.
-Environmental Protection Agency (EPA).
-ECOTOX collects raw, non harmonized ecotoxicological test data on on aquatic and terrestrial wildlife as well as plants and publishes it on a quarterly basis.
-By the time of writing it contained 926,108 test results, comprising 11,685 chemicals tested on 12,668 different species \citep{elonen_ecotoxicology_2018}.
-The collected ecotoxicological test data contain a variety of measured endpoints such as Effective Concentrations (\ecfifty{}) values, No-observed effect concentrations (NOEC) or lowest observed effect concentrations (LOEC).
-Likewise the data set contains different effect measures such as mortality, intoxication and growth as well as a multiplicity of test durations ranging from seconds to weeks or even years.
-The \etoxbase aggregates this information according to the user's inputs.
-These can then be used for the derivation of risk indicators such as Species Sensitivity Distributions (SSDs) \citep{posthuma_species_2002} and Toxic Units (TUs), which represent two prominent concepts to assess effects on organisms in ecotoxicology .
-The former combines toxicity data of several organism groups towards a chemicals to estimate effects on biotic communities, the latter refers to effects of a specific organism (group) towards a chemical.
-The two concepts are widely used in ecotoxicology \citep{kefford_definition_2011, schafer_effects_2011} since the allow for a comparison of toxicities across multiple chemicals and biological communities, respectively.
-Although performed on one and the same organism and chemical, outcomes of different ecotoxicological tests vary greatly due to differences in test parameters, genetic differences in individual populations or other irrepressible factors making a selection of appropriate test results for CRA laborious.
-The \etoxbase{} aims to relief this process and provide single toxicity values for organism, chemical and test duration combinations.
-The \etoxbase{} downloads every new version of the ECOTOX data base and performs several quality checks, e.g.
-unit harmonization, and cleaning steps on it.
-Hence, new scientific test results are constantly incorporated.
-Additionally the \etoxbase{} queries chemical- and organism-specific variables from other open data bases to fill information gaps such as water solubility, organism habitat or regional occurrence patterns in the ECOTOX data base.
-
-\pagebreak
-
-\section*{Methods}
-
-%%The Methods should include detailed text describing any steps or procedures 
-%%used in producing the data, including full descriptions of the experimental 
-%%design, data acquisition assays, and any computational processing (e.g. 
-%%normalization, image feature extraction). Related methods should be grouped 
-%%under corresponding subheadings where possible, and methods should be described 
-%%in enough detail to allow other researchers to interpret and repeat, if required, 
-%%the full study. Specific data outputs should be explicitly referenced via data 
-%%citation (see Data Records and Data Citations, below). Authors should cite 
-%%previous descriptions of the methods under use, but ideally the method 
-%%descriptions should be complete enough for others to understand and reproduce 
-%%the methods and processing steps without referring to associated publications. 
-%%There is no limit to the length of the Methods section.
-
-The \etoxbase{} consists of two parts of software: Firstly a processing pipeline (Fig.
-\ref{fig:pipeline}) downloads quarterly released new versions of the \epa{} and performs several preparation steps in R \citep{r_core_team_r_2017}, to then store the outcome in a PostgreSQL data base.
-This constitutes the basis upon which the R shiny web application is accessing the data \citep{r_core_team_r_2017}.
-The application itself is set of R functions that filter and aggregate the data according to a user's inputs \ref{fig:app}.
-Finally in the web application, a `filtered` and an `aggregated` data set as well as an interactive plot, for data exploration are returned.
-These two data sets can be downloaded.
-In the following these two parts are explained in detail, naming the respective R scripts in brackets.
-
-\subsection*{Data acquisition \& preparation}
-
-The EPA ECOTOX data is downloaded (bd\_epa\_download.R) and subsequently built into a PostgreSQL data base (bd\_epa\_postgres.R).
-Afterwards the data is imported into R for pre-processing (da\_epa1.R, da\_epa2.R, da\_epa3.R), which includes validity checks, removal of special characters, unit harmonizations and selection of relevant variables.
-Chemical abstract service (CAS) numbers, unique chemical identifiers together we taxon names are then used to query other data bases for additional information on chemicals and biota respectively.
-Chemical information is retrieved from the Compendium of Pesticide Common Names (qu\_aw.R) \citep{CITE_AW}, the Chemspider data base (qu\_chemspieder\_scrape.R) \citep{CITE-CHEMSPIDER}, from Eurostat <CITATION>, the Pubchem data base (qu\_pc.R) \citep{CITE_PUBCHEM}, and from the Physprop data base (qu\_pp.R) \citep{CITE_PHYSPROP}.
-These queries mostly rely on the webchem R-package \citep{szocs_webchem_2015-1}.
-For biota, habitat and occurrence information is queried from the World Register of Marine Species (WORMS) (qu\_worms2.R) \citep{WORMS} and from the Global Biodiversity Information Facility (GBIF) (qu\_rgbif.R) \citep{CITE_RGBIF} using the rgbif R-package \citep{chamberlain_rgbif_2018}.
-The acquired information is then merged (re\_merge.R) and information is combined (re\_combine.R).
-From there on the data set is analysed (re\_analyses.R) and checked (re\_checks\_internal.R) and finally compiled in one table (re\_final.R).
-This table is then used as the input for the filter and aggregation functions \ref{fig:pipeline}.
-For a detailed overview of what information is collected see table \ref{tab:scripts-pipeline}.
-
-\begin{figure}
-    \includestandalone[width=\textwidth, scale=0.1]{tikz/pipeline_organigram}
-    \caption{Processing Pipeline}
-    \label{fig:pipeline}
-\end{figure}
-
-\subsection*{The application}
-The application is accessible through \app{} and was built in R, using the shiny web application framework \citep{chang_shiny_2018}.
-Users can define several input parameters \ref{tab:inputs} in the application's graphical user interface (GUI), handled in the ui.R script in order to adjust the aggregation according specific requirements.
-In doing so users execute R functions (fun\_filter.R, fun\_aggregation.R, fun\_filagg\_plot\_ly.R) loaded in the server.R script, which produce a 'filtered' and an 'aggregated' data set as well as an interactive overview plot.
-Only the 'aggregated' data set is presented in the GUI.
-However, both data sets can be downloaded via a download button.
-
-\begin{figure}
-    \includestandalone[scale=0.75]{tikz/application}
-    \caption{Application}
-    \label{fig:app}
-\end{figure}
-
-\subsection*{Aggregation}
-The \etoxbase{} application aggregates the test results according to chosen filters in a two step process.
-Firstly the filtered test results are aggregated by the CAS number, the chosen taxon and the selected test duration.
-Secondly, the returned data is then aggregated by the CAS number.
-The former can't be influenced by the user and calculates either the minimum or the median depending on the amount of results to aggregate (n <= 2: minimum, if n > 2: median).
-Thereof the second step calculates the minimum, the maximum, the median, the geometric mean, or the arithmetic mean as an aggregate.
-
-\subsection*{Code availability}
-%%For all studies using custom code in the generation or processing of datasets, 
-%%a statement must be included here, indicating whether and how the code can be 
-%%accessed, including any restrictions to access. This section should also include 
-%%information on the versions of any software used, if relevant, and any specific 
-%%variables or parameters used to generate, test, or process the current dataset.
-
-The processing code is stored in a Github reopsitory (\git{}).
-All processing scripts were written in R version 3.4.4.
-PostgreSQL 9.5 has been used to build the data base.
-Hence, R together with its packages \ref{tab:rpackages} and PostgreSQL suffice to rebuild the process.
-The code of the application itself is stored in a second Github repostitory (\gitapp{}).
-Since new versions of the EPA ECOTOX data base are published on a quarterly basis the data acquisition process is automated and new versions of the application are scheduled to be released regularly.
-
-\begin{table}[h!]
-    \centering
-    \begin{tabular}{|l|l|l|l|}
-        \hline
-        Category        & Package       & Version   & Reference                     \\ \hline
-        processing      & RCurl         & 1.95-4.11 & \citep{lang_rcurl_2018}       \\ \hline
-        processing      & data.table    & 1.12.0    & \citep{dowle_data.table_2018} \\ \hline
-        application     & shiny         & 1.1.0     & \citep{chang_shiny_2018}      \\ \hline
-    \end{tabular}
-    \caption{List of R packages used in compiling the data.\newline{}
-    TODO: Automate version input and name generation - whole table\newline{}
-    TODO2: debug chang\_shiny\_2018 reference}
-    \label{tab:rpackages}
-\end{table}
-
-\pagebreak
-
-\section*{Data Records}
-%%Please explain each data record associated with this work, including
-%%the repository where this information is stored, and an overview of
-%%the data files and their formats. Each external data record should
-%%be listed in Data Citation section at the end of this template, and 
-%%records should be cited throughout the manuscript as, for example 
-%%(Data Citation 1). 
-%%
-%%Tables should be used to support the data records, and should clearly indicate 
-%%the samples and subjects, their provenance, and the experimental manipulations 
-%%performed on each. They should also specify the data output resulting from each 
-%%data-collection or analytical step, should these form part of the archived record. 
-%%Please see the submission guidelines at the \emph{Scientific Data} website, and 
-%%our Word templates for more information on preparing such tables.
-
-\pagebreak
-
-\section*{Technical Validation}
-%%This section presents any experiments or analyses that are needed
-%%to support the technical quality of the dataset. This section may
-%%be supported by up figures and tables, as needed. This is a required
-%%section; authors must present information justifying the reliability
-%%of their data.
-
-<PUT COMPARISON WITH PPDB DATA HERE?>
-
-\pagebreak
-
 \section*{Usage Notes}
@@
 \pagebreak
-
-\section*{Acknowledgements}
-%%Text acknowledging non-author contributors. Acknowledgements should
-%%be brief, and should not include thanks to anonymous referees and
-%%editors, or effusive comments. Grant or contribution numbers may be
-%%acknowledged. Author contributions Please describe briefly the contributions
-%%of each author to this work on a separate line. 
-%%
-%%AK did this and that. 
-%%
-%%BG did this and that and the other.
-
-\section*{Competing financial interests}
-%%A competing financial interests statement is required for all accepted
-%%papers published in \emph{Scientific Data}. If none exist simply write,
-%%``The author(s) declare no competing financial interests''.
-
-\section*{Figures Legends}
-%%Figure should be referred to using a consistent numbering scheme through
-%%the entire Data Descriptor. For initial submissions, authors may choose
-%%to supply this document as a single PDF with embedded figures, but
-%%separate figure image files must be provided for revisions and accepted
-%%manuscripts. In most cases, a Data Descriptor should not contain more
-%%than three figures, but more may be allowed when needed. We discourage
-%%the inclusion of figures in the Supplementary Information \textendash{}
-%%all key figures should be included here in the main Figure section. 
-%%
-%%Figure legends begin with a brief title sentence for the whole figure
-%%and continue with a short description of what is shown in each panel,
-%%as well as explaining any symbols used. Legend must total no more
-%%than 350 words, and may contain literature references.
-
-\section*{Tables}
-%%Tables supporting the Data Descriptor. These can provide summary information
-%%(sample numbers, demographics, etc.), but they should generally not
-%%be used to present primary data (i.e. measurements). Tables containing
-%%primary data should be submitted to an appropriate data repository. 
-%%
-%%Tables may be provided within the \LaTeX{} document or as separate
-%%files (tab-delimited text or Excel files). Legends, where needed,
-%%should be included here. Generally, a Data Descriptor should have
-%%fewer than ten Tables, but more may be allowed when needed. Tables
-%%may be of any size, but only Tables which fit onto a single printed
-%%page will be included in the PDF version of the article (up to a maximum
-%%of three).
-
-%\begin{sidewaystable}
-%    \csvautotabular[respect all]{data/meta.csv};
-%    \caption{Meta variables table};
-%    \label{table:meta-variables};
-%\end{sidewaystable}
-
-\begin{table}[h!]
-\begin{tabular}{|l|l|}
-\hline
-Script                      & Description                                      \\ \hline
-run\_build.R                 & runs all sub scripts                             \\ \hline
-setup.R                     & sets up the environment                          \\ \hline
-bd\_epa\_download.R         & download EPA Ecotox data base                    \\ \hline
-bd\_epa\_postgres.R         & build EPA Ecotox data base locally               \\ \hline
-da\_epa1.R                  & raw EPA Ecotox data export                       \\ \hline
-da\_epa2.R                  & process EPA Ecotox data                          \\ \hline
-da\_epa3.R                  & reduce EPA Ecotox data                           \\ \hline
-qu\_aw.R                    & query Compendium of Pesticide Common Names       \\ \hline
-qu\_chemspider\_scrape.R    & query Chemspider data base                       \\ \hline
-qu\_eurostat\_chem\_class.R & query Eurostat data set                          \\ \hline
-qu\_pc.R                    & query PubChem data base                          \\ \hline
-qu\_pp.R                    & query Physprop data base                         \\ \hline
-qu\_worms2.R                & query WORMS data base                            \\ \hline
-qu\_gbif.R                  & query GBIF data base                             \\ \hline
-re\_merge.R                 & merge all data                                   \\ \hline
-re\_combine.R               & combine information                              \\ \hline
-re\_analyses.R              & ???                                              \\ \hline
-re\_checks\_internal.R      & ???                                              \\ \hline
-re\_final.R                 & create and write final table                     \\ \hline
-wr\_dbase.R                 & write final table to data base                   \\ \hline
-wr\_app.R                   & write final table and variables to app directory \\ \hline
-wr\_meta.R                  & ???                                              \\ \hline
-wr\_stats.R                 & ???                                              \\ \hline
-\end{tabular}
-\caption{Scripts that are used to process the data}
-\label{tab:scripts-pipeline}
-\end{table}
-
-\begin{table}[h!]
-\begin{tabular}{|l|l|}
-\hline
-Script                     & Description                                               \\ \hline
-setup.R                    & sets up the environment                                   \\ \hline
-ui.R                       & creates the graphical user interface of the application   \\ \hline
-server.R                   & handles functions and data input and output               \\ \hline
-fun\_filter.R              & function that filters the data                            \\ \hline
-fun\_aggregation.R         & function that performs the aggregation                    \\ \hline
-fun\_filagg\_plot\_ly.R    & creates interactive plot                                  \\ \hline
-\end{tabular}
-\caption{Scripts that used for the web application}
-\label{tab:scripts-app}
-\end{table}
-
-\pagebreak
-
-\bibliographystyle{apalike}
-\bibliography{refs/references-etox-base.bib,refs/references-etox-base-rpackages.bib,refs/references-etox-base-rshinypackages.bib}
-% multiple refs have to be written w/o space
-
-\section*{Data Citations}
-
-\pagebreak
-\section*{TODO}
-
-\begin{itemize}
-
-\item compare Etox-BAse output with PPDB
-
-\item plot example SSD and TU
-
-\item acute vs chronic
-
-\item OECD guidlines are not machine readable
-
-\item discuss flaws in column types (e.g some concentration entries contain '+' or '~' and are therefore not convertable to numeric) - This hampers large scale analysis greatly!
-
-\pagebreak
-\section*{MISC}
-
-\begin{figure}
-    \includestandalone[scale=0.5]{tikz/tree_source}
-    \caption{Alternative processing pipeline}
-    \label{fig:pipeline-tree}
-\end{figure}
-
-\section*{not included citations}
-Put in citation \citep{hartung_chemical_2009} ?
-which claims that REACH won't meet their assumptions and say 54 million vertebrate animals are needed for tests that would cost €9.5 billion over the next ten years.
-I.e 20 times more animals, 6 times the costs in comparison to the official estimates
-
-\end{itemize}
-
-\end{document}

### 2019-06-12 article/article.tex

@@
 
-\newcommand{\etoxbase}{Etox-Base}
+\newcommand{\standartox}{STANDARTOX}
 \newcommand{\epa}{EPA ECOTOX data base}
@@
 \begin{abstract}
+%%5This is a manuscript template for Data Descriptor submissions to \emph{Scientific
+%%Data} (http://www.nature.com/scientificdata). The abstract must be
+%%no longer than 170 words, and should succinctly describe the study,
+%%the assay(s) performed, the resulting data, and the reuse potential,
+%%but should not make any claims regarding new scientific findings.
+%%No references are allowed in this section.
 
@@
 \end{abstract}
-
 \pagebreak
 
-\section*{Background \& Summary [max 700 words]}
-
-A large number of chemicals such as pharmaceuticals, pesticides and synthetic hormones are in daily use all over the world with insufficient knowledge about possible adverse effects on the environment.
-In Europe alone some 100,000 chemicals are estimated to be in current use, whereof 30,000 are produced in quantities larger than one ton per year \citep{breithaupt_costs_2006}.
-Chemicals are brought to the environment deliberately - as it is the case of pesticides, or arrive therein as a byproduct from other processes (e.g.
-atmospheric emissions or wastewater) \citep{schwarzenbach_challenge_2006}.
-Ecosystems provide essential services to human societies such as drinking and irrigation water, food and climate regulation.
-These services are products of different ecosystem functions that crucially depend on the integrity of the populations and communities that drive these ecosystem functions.
-However, besides habitat degradation, climate change and nutrient enrichment, pollution with man-made chemical toxicants threatens these populations and communities in various ways which are currently not fully understood \citep{steffen_anthropocene_2007}.
-Pollution with man-made chemical toxicants was indeed identified as one of three major environmental problems for which research gaps hamper the derivation of planetary boundaries, i.e.
-thresholds beyond which irreversible state shifts may occur \citep{steffen_anthropocene_2007}.
-Bernhardt et al.
-\citet{bernhardt_synthetic_2017} argue further that the knowledge gap how chemicals effect populations and communities and hence ecosystem functions and ecosystem services, would also impede society’s ability to accomplish the Sustainable Development Goals of the United Nations.
-According to Breithaupt \citet{breithaupt_costs_2006} less than one percent of chemicals released to the environment are thoroughly tested.
-Although for some chemicals advanced and standardized \citep{oecd_oecd_2018} test methods have been developed, ecotoxicological tests for a lot of chemicals, especially newly emerging ones remain scarce and if existent, the information is represented sparsely and in inconsistent formats.
-\citep{gessner_fostering_2016}.
-Therefore we developed a data base that collects, processes and aggregates ecotoxicological tests results in order to subsequently publish it in an harmonized form as a web application - the \etoxbase tool: \href{http://139.14.20.252:3838/etox-base-shiny/}.
-This should facilitate researchers the access to ecotoxicological test results and support modern chemical risk assessment (CRA) workflows.
-The \etoxbase makes use of ecotoxicological test data, obtained from the ECOTOXicology knowledgebase (ECOTOX) created by the U.S.
-Environmental Protection Agency (EPA).
-ECOTOX collects raw, non harmonized ecotoxicological test data on on aquatic and terrestrial wildlife as well as plants and publishes it on a quarterly basis.
-By the time of writing it contained 926,108 test results, comprising 11,685 chemicals tested on 12,668 different species \citep{elonen_ecotoxicology_2018}.
-The collected ecotoxicological test data contain a variety of measured endpoints such as Effective Concentrations (\ecfifty{}) values, No-observed effect concentrations (NOEC) or lowest observed effect concentrations (LOEC).
-Likewise the data set contains different effect measures such as mortality, intoxication and growth as well as a multiplicity of test durations ranging from seconds to weeks or even years.
-The \etoxbase aggregates this information according to the user's inputs.
-These can then be used for the derivation of risk indicators such as Species Sensitivity Distributions (SSDs) \citep{posthuma_species_2002} and Toxic Units (TUs), which represent two prominent concepts to assess effects on organisms in ecotoxicology .
-The former combines toxicity data of several organism groups towards a chemicals to estimate effects on biotic communities, the latter refers to effects of a specific organism (group) towards a chemical.
-The two concepts are widely used in ecotoxicology \citep{kefford_definition_2011, schafer_effects_2011} since the allow for a comparison of toxicities across multiple chemicals and biological communities, respectively.
-Although performed on one and the same organism and chemical, outcomes of different ecotoxicological tests vary greatly due to differences in test parameters, genetic differences in individual populations or other irrepressible factors making a selection of appropriate test results for CRA laborious.
-The \etoxbase{} aims to relief this process and provide single toxicity values for organism, chemical and test duration combinations.
-The \etoxbase{} downloads every new version of the ECOTOX data base and performs several quality checks, e.g.
-unit harmonization, and cleaning steps on it.
-Hence, new scientific test results are constantly incorporated.
-Additionally the \etoxbase{} queries chemical- and organism-specific variables from other open data bases to fill information gaps such as water solubility, organism habitat or regional occurrence patterns in the ECOTOX data base.
-
+%%%%%%%%%%%%%%%%%%%%%% 01 Background
+\input{article/parts/01_background.tex}
 \pagebreak
 
-\section*{Methods}
-
-%%The Methods should include detailed text describing any steps or procedures 
-%%used in producing the data, including full descriptions of the experimental 
-%%design, data acquisition assays, and any computational processing (e.g. 
-%%normalization, image feature extraction). Related methods should be grouped 
-%%under corresponding subheadings where possible, and methods should be described 
-%%in enough detail to allow other researchers to interpret and repeat, if required, 
-%%the full study. Specific data outputs should be explicitly referenced via data 
-%%citation (see Data Records and Data Citations, below). Authors should cite 
-%%previous descriptions of the methods under use, but ideally the method 
-%%descriptions should be complete enough for others to understand and reproduce 
-%%the methods and processing steps without referring to associated publications. 
-%%There is no limit to the length of the Methods section.
-
-The \etoxbase{} consists of two parts of software: Firstly a processing pipeline (Fig.
-\ref{fig:pipeline}) downloads quarterly released new versions of the \epa{} and performs several preparation steps in R \citep{r_core_team_r_2017}, to then store the outcome in a PostgreSQL data base.
-This constitutes the basis upon which the R shiny web application is accessing the data \citep{r_core_team_r_2017}.
-The application itself is set of R functions that filter and aggregate the data according to a user's inputs \ref{fig:app}.
-Finally in the web application, a `filtered` and an `aggregated` data set as well as an interactive plot, for data exploration are returned.
-These two data sets can be downloaded.
-In the following these two parts are explained in detail, naming the respective R scripts in brackets.
-
-\subsection*{Data acquisition \& preparation}
-
-The EPA ECOTOX data is downloaded (bd\_epa\_download.R) and subsequently built into a PostgreSQL data base (bd\_epa\_postgres.R).
-Afterwards the data is imported into R for pre-processing (da\_epa1.R, da\_epa2.R, da\_epa3.R), which includes validity checks, removal of special characters, unit harmonizations and selection of relevant variables.
-Chemical abstract service (CAS) numbers, unique chemical identifiers together we taxon names are then used to query other data bases for additional information on chemicals and biota respectively.
-Chemical information is retrieved from the Compendium of Pesticide Common Names (qu\_aw.R) \citep{CITE_AW}, the Chemspider data base (qu\_chemspieder\_scrape.R) \citep{CITE-CHEMSPIDER}, from Eurostat <CITATION>, the Pubchem data base (qu\_pc.R) \citep{CITE_PUBCHEM}, and from the Physprop data base (qu\_pp.R) \citep{CITE_PHYSPROP}.
-These queries mostly rely on the webchem R-package \citep{szocs_webchem_2015-1}.
-For biota, habitat and occurrence information is queried from the World Register of Marine Species (WORMS) (qu\_worms2.R) \citep{WORMS} and from the Global Biodiversity Information Facility (GBIF) (qu\_rgbif.R) \citep{CITE_RGBIF} using the rgbif R-package \citep{chamberlain_rgbif_2018}.
-The acquired information is then merged (re\_merge.R) and information is combined (re\_combine.R).
-From there on the data set is analysed (re\_analyses.R) and checked (re\_checks\_internal.R) and finally compiled in one table (re\_final.R).
-This table is then used as the input for the filter and aggregation functions \ref{fig:pipeline}.
-For a detailed overview of what information is collected see table \ref{tab:scripts-pipeline}.
-
-\begin{figure}
-    \includestandalone[width=\textwidth, scale=0.1]{tikz/pipeline_organigram}
-    \caption{Processing Pipeline}
-    \label{fig:pipeline}
-\end{figure}
-
-\subsection*{The application}
-The application is accessible through \app{} and was built in R, using the shiny web application framework \citep{chang_shiny_2018}.
-Users can define several input parameters \ref{tab:inputs} in the application's graphical user interface (GUI), handled in the ui.R script in order to adjust the aggregation according specific requirements.
-In doing so users execute R functions (fun\_filter.R, fun\_aggregation.R, fun\_filagg\_plot\_ly.R) loaded in the server.R script, which produce a 'filtered' and an 'aggregated' data set as well as an interactive overview plot.
-Only the 'aggregated' data set is presented in the GUI.
-However, both data sets can be downloaded via a download button.
-
-\begin{figure}
-    \includestandalone[scale=0.75]{tikz/application}
-    \caption{Application}
-    \label{fig:app}
-\end{figure}
-
-\subsection*{Aggregation}
-The \etoxbase{} application aggregates the test results according to chosen filters in a two step process.
-Firstly the filtered test results are aggregated by the CAS number, the chosen taxon and the selected test duration.
-Secondly, the returned data is then aggregated by the CAS number.
-The former can't be influenced by the user and calculates either the minimum or the median depending on the amount of results to aggregate (n <= 2: minimum, if n > 2: median).
-Thereof the second step calculates the minimum, the maximum, the median, the geometric mean, or the arithmetic mean as an aggregate.
-
-\subsection*{Code availability}
-%%For all studies using custom code in the generation or processing of datasets, 
-%%a statement must be included here, indicating whether and how the code can be 
-%%accessed, including any restrictions to access. This section should also include 
-%%information on the versions of any software used, if relevant, and any specific 
-%%variables or parameters used to generate, test, or process the current dataset.
-
-The processing code is stored in a Github reopsitory (\git{}).
-All processing scripts were written in R version 3.4.4.
-PostgreSQL 9.5 has been used to build the data base.
-Hence, R together with its packages \ref{tab:rpackages} and PostgreSQL suffice to rebuild the process.
-The code of the application itself is stored in a second Github repostitory (\gitapp{}).
-Since new versions of the EPA ECOTOX data base are published on a quarterly basis the data acquisition process is automated and new versions of the application are scheduled to be released regularly.
-
-\begin{table}[h!]
-    \centering
-    \begin{tabular}{|l|l|l|l|}
-        \hline
-        Category        & Package       & Version   & Reference                     \\ \hline
-        processing      & RCurl         & 1.95-4.11 & \citep{lang_rcurl_2018}       \\ \hline
-        processing      & data.table    & 1.12.0    & \citep{dowle_data.table_2018} \\ \hline
-        application     & shiny         & 1.1.0     & \citep{chang_shiny_2018}      \\ \hline
-    \end{tabular}
-    \caption{List of R packages used in compiling the data.\newline{}
-    TODO: Automate version input and name generation - whole table\newline{}
-    TODO2: debug chang\_shiny\_2018 reference}
-    \label{tab:rpackages}
-\end{table}
-
+%%%%%%%%%%%%%%%%%%%%%% 02 Methods
+\input{article/parts/02_methods.tex}
 \pagebreak
 
-\section*{Data Records}
-%%Please explain each data record associated with this work, including
-%%the repository where this information is stored, and an overview of
-%%the data files and their formats. Each external data record should
-%%be listed in Data Citation section at the end of this template, and 
-%%records should be cited throughout the manuscript as, for example 
-%%(Data Citation 1). 
-%%
-%%Tables should be used to support the data records, and should clearly indicate 
-%%the samples and subjects, their provenance, and the experimental manipulations 
-%%performed on each. They should also specify the data output resulting from each 
-%%data-collection or analytical step, should these form part of the archived record. 
-%%Please see the submission guidelines at the \emph{Scientific Data} website, and 
-%%our Word templates for more information on preparing such tables.
-
+%%%%%%%%%%%%%%%%%%%%%% 03 Data records
+\input{article/parts/03_data_records.tex}
 \pagebreak
 
-\section*{Technical Validation}
-%%This section presents any experiments or analyses that are needed
-%%to support the technical quality of the dataset. This section may
-%%be supported by up figures and tables, as needed. This is a required
-%%section; authors must present information justifying the reliability
-%%of their data.
-
-<PUT COMPARISON WITH PPDB DATA HERE?>
-
+%%%%%%%%%%%%%%%%%%%%%% 04 Technical validation
+\input{article/parts/04_technical_validation.tex}
 \pagebreak
 
-\section*{Usage Notes}
-%%Brief instructions that may help other researchers reuse these dataset.
-%%This is an optional section, but strongly encouraged when helpful
-%%to readers. This may include discussion of software packages that
-%%are suitable for analyzing the assay data files, suggested downstream
-%%processing steps (e.g. normalization, etc.), or tips for integrating
-%%or comparing this with other datasets. If needed, authors are encouraged
-%%to provide code, programs, or data processing workflows when they may help 
-%%others analyse the data. We encourage authors to archive related code in 
-%%a DOI-issuing archive when possible, but code may also be supplied as 
-%%supplementary information files. 
-%%
-%%For studies involving privacy or safety controls on public access
-%%to the data, this section should describe in detail these controls,
-%%including how authors can apply to access the data, and what criteria
-%%will be used to determine who may access the data, and any limitations
-%%on data use.
-
-The \etoxbase{} is designed to support CRA of ecotoxicologists.
-With an increased amount of available ecotoxicological test data, it becomes fundamental to provide and distribute such information in reasonable formats, meaning accessible for humans as well as machines.
-Currently CRA often relies on test results of a few, well tested standard organism, such as \textit{Daphnia magna}, \textit{Pimephales promelas}, \textit{Selenastrum capricornutum}, although researchers have conducted experiments on a much greater variety of organisms (e.g.
-the 926,108 test results collected in the EPA ECOTOX data base).
-In order to locate ecotoxicological test data, researchers can rely on individual test results reported in publications <CITE>, on compiled and published data sets \citep{malaj_organic_2014} or on data bases such as the Pesticide Property Data Base (PPDB) \citep{lewis_international_2016}.
-Each data source has its limitations though: Collecting toxicity information from individual publications is laborious and using already compiled data from publications or data bases is often limited to a specific groups of chemicals, such as pesticides, biozides etc.
-The \etoxbase{}, making use of the EPA ECOTOX data base, the largest publicly available collection of ecotoxicological test results, aims to facilitate the data retrieval process by providing adequate aggregates (\ecfifty{}, NOEC, LOEC) of the tests results.
-These are then used for the calculation of SSDs, TUs or other risk indicators.
-SSDs try to extrapolate effects of chemicals towards organisms from species to a community level by fitting a statistical distribution to toxicity test results on organisms of different trophic levels in order to derive concentrations at which a specific fraction of organisms are effected (often at 5\% - HC\textsubscript{5} \citep{posthuma_species_2002}.
-Figure \ref{fig:ssd-isoproturon} shows an example of a SSD and the derived HC\textsubscript{5} value (red dassed line) created for the herbicide Isoproturon.
-TUs are another indicator of effects of chemicals on organisms.
-They are calculated by dividing a measured concentration (c) by an \ecfifty{} value (cf.
-Equation \ref{eq:tu}).
-
-\begin{equation}
-    TU_i = \frac{c_i}{\ecfifty}
-    \label{eq:tu}
-\end{equation}
-
-\begin{figure}[h!]
-    \dentering
-    \includegraphics[width=1\linewidth]{article/figures/ssd2_boot.png}
-    \caption{SSD plot showing the susceptibility of organisms towards the herbicide Isoproturon.
-The red line represents the fit and the red dashed line marks the HC\textsubscript{5} value, which is a common measure in CRA.
-The legend denotes common organism groups used in ecotoxicology.}
-    \label{fig:ssd-isoproturon}
-\end{figure}
-
-\subsection*{Filter and Aggregate}
-
-In the application's GUI, the user can choose among different parameters to filter and aggregate the test data accordingly \ref{tab:inputs}.
-Percentage values next to inputs show the proportion of the respective option (e.g.
-\textit{Active Ingredient - 43\%} indicates that 43\% of the data are Active Ingredients).
-The user can filter for compound-specific filters, including the CAS number, the concentration type (e.g.
-Active Ingredient, Formulation), the chemical class (e.g.
-Insecticides, Metals) and taxon-specific filters, including common taxonomic groups (e.g.
-Daphniidae, Algae), the organism habitat (e.g.
-freshwater) and the continent (e.g.
-Europe, Asia).
-Regarding the test parameters the user can choose test durations (in hours), Effect groups (e.g.
-Mortaliy, Population, Growth) and Endpoints (\ecfifty, EC\textsubscript{10}, LOEC or NOEC).
-Besides that, two cleaning steps can also be chosen.
-Firstly the application allows to exclude test results exhibiting concentrations that are higher than the actual water solubility of the respective chemical at \ang{20} C.
-Secondly the user can choose to exclude outliers (cf.
-Table \ref{tab:scripts-app}).
-The outliers are selected as values that exceed lower (0.25) and the upper (0.75) quartile by 1.5 times the inter-quartile range.
-As aggregates the user can choose the minimum, the maximum, the median, the geometric mean or the arithmetic mean.
-All the steps described in this section are selectable in the GUI and are executed at each click by the application.
-The drop down menu `Download data` allows to download a filtered data set as well as an aggregated data set.
-The \etoxbase{} application can be accessed here: \app{}.
-
-\begin{table}[h!]
-\begin{tabular}{|l|l|l|l|}
-\hline
-Menu   & Group       & Subgroups              & Options \\ \hline
-Filter      & Compound    & CAS                    & input CAS \\ \hline
-Filter      & Compound    & Concentration type     & Active Ingredient; Formulation; Total \\ \hline
-Filter      & Compound    & Chemical class         & Insecticides; Herbicides; Metals \\ \hline
-Filter      & Taxon       & Taxon                  & Choose a taxon \\ \hline
-Filter      & Taxon       & Organism habitat       & freshwater; terrestrial; marine; brackish \\ \hline
-Filter      & Taxon       & Continent              & North America; Asia; Europe \\ \hline
-Filter      & Test        & Duration               & enter duration in hours \\ \hline
-Filter      & Test        & Effect groups          & MOR - Mortality; POP - Population \\ \hline
-Filter      & Test        & Endpoins               & EC50, LOEC, LOEL \\ \hline
-Filter      & Checks      & Water solubility check & yes/no \\ \hline
-Filter      & Checks      & Remove outliers        & yes/no \\ \hline
-Aggregation & Aggregation & Aggregate              & Minimum; Maximum; Median \\ \hline
-\end{tabular}
-\caption{\etoxbase GUI paramters to choose.}
-\label{tab:inputs}
-\end{table}
-
-\subsection*{Limitations}
-
-Generally it has to be stated that the returned values are subject to change, as new tests will be constantly incorporated into the \etoxbase{}.
-However, we expect values of well tested chemical-organism combinations to be relatively stable.
-The returned values are certainly influenced by individual tests characteristics, such as pH, temperature or conductivity to just name a few, which influences the effect relationship between chemicals and organisms \citep{rosenkrantz_influence_2013,li_temperature_2011}.
-Unfortunately the EPA ECOTOX data base has only rare records on test parameters, which prevents incorporating such parameters in an aggregation <proportions of available data: Temperature:77\%, pH:56\%, hardness:27\%, dissolved oxygen:18\%, Alkalinity:15\%, salinity:9\%, others:<=5\%>.
-However, using the median or the geometric mean as an aggregate allows for an adequate estimation of the central tendency of the toxicity of a chemical towards an organism group.
-The geometric mean is recommended, as it is in comparison to the arithmetic mean considerably less influenced by outliers and is suitable for skewed data.
-Also, the geometric mean is preferable in contrast to the median, since the median completely ignores the influence of large or small values, making it unreliable for small data sets <CITATION GEOMETRIC MEAN>.
-
-Other initiatives, with different, though partly overlapping aims accumulating existing ecotoxicological test data are published online.
-The PPDB provides validated ecotoxicological test results for pesticides, narrowing chemicals to this group \citep{lewis_international_2016}.
-The Network of reference laboratories, research centres and related organisations for monitoring of emerging environmental substances (NORMAN) focuses on assembling river basin specific pollutants\citep{von_der_ohe_new_2011}.
-The EnviroTox data base \href{https://envirotoxdatabase.org/} makes also use of the \epa{} data in order to derive Predicted no-effect concentrations (PNEC) \citep{health_and_environmental_sciences_institute_hesi_envirotox_2019}.
-In comparison to the \etoxbase, these examples partly rely on the same data, yet not aiming to derive \ecfifty, LOEC and NOEC from it automatically.
-
-\pagebreak
+%%%%%%%%%%%%%%%%%%%%%% 05 Usage notes
+\input{article/parts/05_usage_notes.tex}
 
@@
 \section*{not included citations}
-Put in citation \citep{hartung_chemical_2009} ?
+
+\item Put in citation \citep{hartung_chemical_2009} ?
 which claims that REACH won't meet their assumptions and say 54 million vertebrate animals are needed for tests that would cost €9.5 billion over the next ten years.
 I.e 20 times more animals, 6 times the costs in comparison to the official estimates
+
+\item "comparison of rate of change in the production and variety of pesticides, pharmaceuticals and other chemicals over the past four decades to CO2 nutrient pollution and other global change factors" \citep{bernhardt_synthetic_2017}
+
+\item worldwide decline in entomofauna: \citep{sanchez-bayo_worldwide_2019}
 

### 2019-06-12 article/parts/01_background.tex

\section*{Background \& Summary [max 700 words]}

A large number of chemicals such as pharmaceuticals, pesticides and synthetic hormones are in daily use all over the world with insufficient knowledge about possible adverse effects on the environment. In Europe alone some 100,000 chemicals are estimated to be in current use, whereof 30,000 are produced in quantities larger than one ton per year \citep{breithaupt_costs_2006}. Although for some chemicals advanced and standardized test methods \citep{oecd_oecd_2018} have been developed, data on toxicity tests for a lot of chemicals, especially newly emerging ones remains scarce and if existent, the information is represented sparsely and in inconsistent formats \citep{gessner_fostering_2016}. Therefore we developed \standardtox{}, a tool to aggregate and standardize toxicity test data from various sources [EPA and ETOX] which provides the information in a reproducible manner.

To our knowledge this ist the first approach

reporducible

time saving

data base that collects, processes and aggregates ecotoxicological tests results in order to subsequently publish it in an harmonized form as a web application - the \etoxbase{} tool: \href{http://139.14.20.252:3838/etox-base-shiny/}.

Chemicals are brought to the environment deliberately - as it is the case of pesticides, or arrive therein as a byproduct from other processes (e.g. atmospheric emissions or wastewater) \citep{schwarzenbach_challenge_2006}. Ecosystems provide essential services to human societies such as drinking and irrigation water, food and climate regulation. These services are products of different ecosystem functions that crucially depend on the integrity of the populations and communities that drive these ecosystem functions. However, besides habitat degradation, climate change and nutrient enrichment, pollution with man-made chemical toxicants threatens these populations and communities in various ways which are currently not fully understood \citep{steffen_anthropocene_2007}. Pollution with man-made chemical toxicants was indeed identified as one of three major environmental problems for which research gaps hamper the derivation of planetary boundaries, i.e. thresholds beyond which irreversible state shifts may occur \citep{steffen_anthropocene_2007}. Bernhardt et al. \citet{bernhardt_synthetic_2017} argue further that the knowledge gap how chemicals effect populations and communities and hence ecosystem functions and ecosystem services, would also impede society's ability to accomplish the Sustainable Development Goals of the United Nations. According to Breithaupt \citet{breithaupt_costs_2006} less than one percent of chemicals released to the environment are thoroughly tested.

This should facilitate researchers the access to ecotoxicological test results and support modern chemical risk assessment (CRA) workflows. The \etoxbase{} makes use of ecotoxicological test data, obtained from the ECOTOXicology knowledgebase (ECOTOX) created by the U.S. Environmental Protection Agency (EPA). ECOTOX collects raw, non harmonized ecotoxicological test data on aquatic and terrestrial wildlife as well as plants and publishes it on a quarterly basis. By the time of writing it contained 926,108 test results, comprising 11,685 chemicals tested on 12,668 different species \citep{elonen_ecotoxicology_2018}. The collected ecotoxicological test data contain a variety of measured endpoints such as Effective Concentrations (\ecfifty{}) values, No-observed effect concentrations (NOEC) or lowest observed effect concentrations (LOEC). Likewise the data set contains different effect measures such as mortality, intoxication and growth as well as a multiplicity of test durations ranging from seconds to weeks or even years. The \etoxbase{} aggregates this information according to the user's inputs. These can then be used for the derivation of risk indicators such as Species Sensitivity Distributions (SSDs) \citep{posthuma_species_2002} and Toxic Units (TUs), which represent two prominent concepts to assess effects on organisms in ecotoxicology. The former combines toxicity data of several organism groups towards a chemicals to estimate effects on biotic communities, the latter refers to effects of a specific organism (group) towards a chemical. The two concepts are widely used in ecotoxicology \citep{kefford_definition_2011, schafer_effects_2011} since the allow for a comparison of toxicities across multiple chemicals and biological communities, respectively. Although performed on one and the same organism and chemical, outcomes of different ecotoxicological tests vary greatly due to differences in test parameters, genetic differences in individual populations or other irrepressible factors making a selection of appropriate test results for CRA laborious. The \etoxbase{} aims to relief this process and provide single toxicity values for organism, chemical and test duration combinations. The \etoxbase{} downloads every new version of the ECOTOX data base and performs several quality checks, e.g. unit harmonization, and cleaning steps on it. Hence, new scientific test results are constantly incorporated. Additionally the \etoxbase{} queries chemical- and organism-specific variables from other open data bases to fill information gaps such as water solubility, organism habitat or regional occurrence patterns in the ECOTOX data base.

    \item standardized toxicity values

    \item reproducible

    \item time saving

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% OLD %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\iffalse

A large number of chemicals such as pharmaceuticals, pesticides and synthetic hormones are in daily use all over the world with insufficient knowledge about possible adverse effects on the environment. In Europe alone some 100,000 chemicals are estimated to be in current use, whereof 30,000 are produced in quantities larger than one ton per year \citep{breithaupt_costs_2006}. Chemicals are brought to the environment deliberately - as it is the case of pesticides, or arrive therein as a byproduct from other processes (e.g. atmospheric emissions or wastewater) \citep{schwarzenbach_challenge_2006}. Ecosystems provide essential services to human societies such as drinking and irrigation water, food and climate regulation. These services are products of different ecosystem functions that crucially depend on the integrity of the populations and communities that drive these ecosystem functions. However, besides habitat degradation, climate change and nutrient enrichment, pollution with man-made chemical toxicants threatens these populations and communities in various ways which are currently not fully understood \citep{steffen_anthropocene_2007}. Pollution with man-made chemical toxicants was indeed identified as one of three major environmental problems for which research gaps hamper the derivation of planetary boundaries, i.e. thresholds beyond which irreversible state shifts may occur \citep{steffen_anthropocene_2007}. Bernhardt et al. \citet{bernhardt_synthetic_2017} argue further that the knowledge gap how chemicals effect populations and communities and hence ecosystem functions and ecosystem services, would also impede society's ability to accomplish the Sustainable Development Goals of the United Nations. According to Breithaupt \citet{breithaupt_costs_2006} less than one percent of chemicals released to the environment are thoroughly tested. Although for some chemicals advanced and standardized \citep{oecd_oecd_2018} test methods have been developed, ecotoxicological tests for a lot of chemicals, especially newly emerging ones remain scarce and if existent, the information is represented sparsely and in inconsistent formats. \citep{gessner_fostering_2016}. Therefore we developed a data base that collects, processes and aggregates ecotoxicological tests results in order to subsequently publish it in an harmonized form as a web application - the \etoxbase{} tool: \href{http://139.14.20.252:3838/etox-base-shiny/}. This should facilitate researchers the access to ecotoxicological test results and support modern chemical risk assessment (CRA) workflows. The \etoxbase{} makes use of ecotoxicological test data, obtained from the ECOTOXicology knowledgebase (ECOTOX) created by the U.S. Environmental Protection Agency (EPA). ECOTOX collects raw, non harmonized ecotoxicological test data on aquatic and terrestrial wildlife as well as plants and publishes it on a quarterly basis. By the time of writing it contained 926,108 test results, comprising 11,685 chemicals tested on 12,668 different species \citep{elonen_ecotoxicology_2018}. The collected ecotoxicological test data contain a variety of measured endpoints such as Effective Concentrations (\ecfifty{}) values, No-observed effect concentrations (NOEC) or lowest observed effect concentrations (LOEC). Likewise the data set contains different effect measures such as mortality, intoxication and growth as well as a multiplicity of test durations ranging from seconds to weeks or even years. The \etoxbase{} aggregates this information according to the user's inputs. These can then be used for the derivation of risk indicators such as Species Sensitivity Distributions (SSDs) \citep{posthuma_species_2002} and Toxic Units (TUs), which represent two prominent concepts to assess effects on organisms in ecotoxicology. The former combines toxicity data of several organism groups towards a chemicals to estimate effects on biotic communities, the latter refers to effects of a specific organism (group) towards a chemical. The two concepts are widely used in ecotoxicology \citep{kefford_definition_2011, schafer_effects_2011} since the allow for a comparison of toxicities across multiple chemicals and biological communities, respectively. Although performed on one and the same organism and chemical, outcomes of different ecotoxicological tests vary greatly due to differences in test parameters, genetic differences in individual populations or other irrepressible factors making a selection of appropriate test results for CRA laborious. The \etoxbase{} aims to relief this process and provide single toxicity values for organism, chemical and test duration combinations. The \etoxbase{} downloads every new version of the ECOTOX data base and performs several quality checks, e.g. unit harmonization, and cleaning steps on it. Hence, new scientific test results are constantly incorporated. Additionally the \etoxbase{} queries chemical- and organism-specific variables from other open data bases to fill information gaps such as water solubility, organism habitat or regional occurrence patterns in the ECOTOX data base.

\fi

### 2019-06-12 article/tikz/pipeline_organigram.tex

@@
 	    child { node [children] { \textbf{wr\_dbase.R} }};
-	\node (writeapp) [main, right of = writedb, xshift=3cm] {Write to app}
+	\node (writeapp) [main, right of = writedb, xshift=3cm, align = center] {Write to\\application}
 	    child { node [children] { \textbf{wr\_app.R}  }};
-	\node (writemeta) [main, right of = writeapp, xshift=3cm] {Write Meta data}
+	\node (writemeta) [main, right of = writeapp, xshift=3cm, align = center] {Write\\Meta data}
 	    child { node [children] { \textbf{wr\_meta.R} }};
-	\node (writestats) [main, right of = writemeta, xshift=3cm] {Write statistics}
+	\node (writestats) [main, right of = writemeta, xshift=3cm, align = center] {Write\\statistics}
 	    child { node [children] { \textbf{wr\_stats.R} }};

### 2019-06-12 article/parts/02_methods.tex

\section*{Methods}

%%The Methods should include detailed text describing any steps or procedures 
%%used in producing the data, including full descriptions of the experimental 
%%design, data acquisition assays, and any computational processing (e.g. 
%%normalization, image feature extraction). Related methods should be grouped 
%%under corresponding subheadings where possible, and methods should be described 
%%in enough detail to allow other researchers to interpret and repeat, if required, 
%%the full study. Specific data outputs should be explicitly referenced via data 
%%citation (see Data Records and Data Citations, below). Authors should cite 
%%previous descriptions of the methods under use, but ideally the method 
%%descriptions should be complete enough for others to understand and reproduce 
%%the methods and processing steps without referring to associated publications. 
%%There is no limit to the length of the Methods section.

The \etoxbase{} consists of two parts of software: Firstly a processing pipeline (Fig. \ref{fig:pipeline}) downloads quarterly released new versions of the \epa{} and performs several preparation steps in R \citep{r_core_team_r_2017}, to then store the outcome in a PostgreSQL data base. This constitutes the basis upon which the R shiny web application is accessing the data \citep{r_core_team_r_2017}. The application itself is set of R functions that filter and aggregate the data according to a user's inputs \ref{fig:app}. Finally in the web application, a `filtered` and an `aggregated` data set as well as an interactive plot, for data exploration are returned. These two data sets can be downloaded. In the following these two parts are explained in detail, naming the respective R scripts in brackets.

\subsection*{Data acquisition \& preparation}

(bd\_epa\_download.R) 
(bd\_epa\_postgres.R)
(da\_epa1.R, da\_epa2.R, da\_epa3.R)
(qu\_aw.R)
(qu\_chemspieder\_scrape.R) 
(qu\_pc.R) 
(qu\_pp.R)
(qu\_worms2.R) 
(qu\_rgbif.R) 
(re\_merge.R) 
(re\_combine.R)
(re\_checks\_internal.R) 
(re\_analyses.R)
(re\_final.R)

The EPA ECOTOX data is downloaded and subsequently built into a PostgreSQL data base. Afterwards the data is imported into R for pre-processing, which includes validity checks, removal of special characters, unit harmonizations and selection of relevant variables. Chemical abstract service (CAS) numbers, unique chemical identifiers together we taxon names are then used to query other data bases for additional information on chemicals and biota respectively. Chemical information is retrieved from the Compendium of Pesticide Common Names \citep{CITE_AW}, the Chemspider data base \citep{CITE-CHEMSPIDER}, from Eurostat <CITATION>, the Pubchem data base \citep{CITE_PUBCHEM}, and from the Physprop data base \citep{CITE_PHYSPROP}. These queries mostly rely on the webchem R-package \citep{szocs_webchem_2015-1}. For biota, habitat and occurrence information is queried from the World Register of Marine Species (WORMS) \citep{WORMS} and from the Global Biodiversity Information Facility (GBIF) \citep{CITE_RGBIF} using the rgbif R-package \citep{chamberlain_rgbif_2018}. The acquired information is then merged and information is combined. From there on the data set is analyzed and checked and finally compiled in one table. This table is then used as the input for the filter and aggregation functions \ref{fig:pipeline}. For a detailed overview of what information is collected see table \ref{tab:scripts-pipeline}.

\begin{figure}
    \includestandalone[width=\textwidth, scale=0.1]{tikz/pipeline_organigram}
    \caption{Processing Pipeline}
    \label{fig:pipeline}
\end{figure}

\subsection*{The application}
The application is accessible through \app{} and was built in R, using the shiny web application framework \citep{chang_shiny_2018}. Users can define several input parameters \ref{tab:inputs} in the application's graphical user interface (GUI), handled in the ui.R script in order to adjust the aggregation according specific requirements. In doing so users execute R functions (fun\_filter.R, fun\_aggregation.R, fun\_filagg\_plot\_ly.R) loaded in the server.R script, which produce a 'filtered' and an 'aggregated' data set as well as an interactive overview plot. Only the 'aggregated' data set is presented in the GUI. However, both data sets can be downloaded via a download button.

\begin{figure}
    \includestandalone[scale=0.75]{tikz/application}
    \caption{Application}
    \label{fig:app}
\end{figure}

\subsection*{Aggregation}
The \etoxbase{} application aggregates the test results according to chosen filters in a two step process. Firstly the filtered test results are aggregated by the CAS number, the chosen taxon and the selected test duration. Secondly, the returned data is then aggregated by the CAS number. The former can't be influenced by the user and calculates either the minimum or the median depending on the amount of results to aggregate (n <= 2: minimum, if n > 2: median). Thereof the second step calculates the minimum, the maximum, the median, the geometric mean, or the arithmetic mean as an aggregate.

\subsection*{Code availability}
%%For all studies using custom code in the generation or processing of datasets, 
%%a statement must be included here, indicating whether and how the code can be 
%%accessed, including any restrictions to access. This section should also include 
%%information on the versions of any software used, if relevant, and any specific 
%%variables or parameters used to generate, test, or process the current dataset.

The processing code is stored in a Github reopsitory (\git{}). All processing scripts were written in R version 3.4.4. PostgreSQL 9.5 has been used to build the data base. Hence, R together with its packages \ref{tab:rpackages} and PostgreSQL suffice to rebuild the process. The code of the application itself is stored in a second Github repostitory (\gitapp{}). Since new versions of the EPA ECOTOX data base are published on a quarterly basis the data acquisition process is automated and new versions of the application are scheduled to be released regularly.

\begin{table}[h!]
    \centering
    \begin{tabular}{|l|l|l|l|}
        \hline
        Category        & Package       & Version   & Reference                     \\ \hline
        processing      & RCurl         & 1.95-4.11 & \citep{lang_rcurl_2018}       \\ \hline
        processing      & data.table    & 1.12.0    & \citep{dowle_data.table_2018} \\ \hline
        application     & shiny         & 1.1.0     & \citep{chang_shiny_2018}      \\ \hline
    \end{tabular}
    \caption{List of R packages used in compiling the data.\newline{}
    TODO: Automate version input and name generation - whole table\newline{}
    TODO2: debug chang\_shiny\_2018 reference}
    \label{tab:rpackages}
\end{table}

\section{TODO}

    \item it is planed to extend this tool to other data bases if they provide new test data  (e.g. UBA ETOC)

### 2019-06-29 talk/talk169.Rmd
added talk

---
title: "[Standartox](http://standartox.uni-landau.de)"
subtitle: "A tool for assessing the risk of chemicals"
author: "Andreas Scharmüller<br/>Verena C. Schreiner<br/>Ralf B. Schäfer<br/>"
institute: "Quantitative Landscape Ecology<br/>iES Landau, Institute for Environmental Sciences<br/>"
date: "1.7.2019 (updated: `r format(Sys.Date(), '%d.%m.%Y')`)"
output:
  xaringan::moon_reader:
    css: [metropolis, metropolis-fonts, default]
    #css: ['css/tweeks.css', metropolis, metropolis-fonts, default]
    #css: ['css/metropolis_as.css', 'css/metropolis-fonts_as.css', 'css/default.css']
    lib_dir: libs
    #classoption: "aspectratio=169"
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
    seal: false # omits title slide
editor_options: 
  chunk_output_type: console
---

```{r setup, echo=FALSE, include=FALSE, warning=FALSE, message=FALSE}
src = '/home/scharmueller/Projects/etox-base/R'
source(file.path(src, 'gn_setup.R'))
```

class: title-slide

# [Standartox](http://standartox.uni-landau.de)
## A tool for assessing the risk of chemicals

### Andreas Scharmüller<br>
### Verena C. Schreiner<br>
### Ralf B. Schäfer
<br><br>
### Quantitative Landscape Ecology<br/>
### iES Landau, Institute for Environmental Sciences<br>
### University Koblenz-Landau
<br>
<img src="https://www.uni-koblenz-landau.de/de/landau/fb7/umweltwissenschaften/logos/land-ecol-neu.jpg" width="100">
<img src="https://www.uni-koblenz-landau.de/de/uni/organisation/verwaltung/abteilungen/abt-1/dokumente-ab1/uni-logo-farbig.jpg" width="200">

---
# Chemicals in the environment

.pull-left[
- Increase in world-wide production of chemicals

- Increase in chemical diversity

- Arrive in freshwater systems
  - Pharmaceuticals
  - Pesticides
  - Personal care products
  - ...
]

.pull-right[

Bernhardt et al. 2017
<img src="/home/scharmueller/Projects/etox-base/talk/data/bernhardt_et_al2.png" width="600px">

]

???
light-red: diversity of US pharmaceuticals<br>
red: application of pesticides (globally and in the US)

---
background-image: url(/home/scharmueller/Projects/etox-base/talk/data/stream.jpg)
background-size: 450px
background-position: 90% 50%

# Assessing the risk of chemicals

- Regulatory thresholds

- SPEAR

- Species Sensitivity Distributions

- Toxic Units (TUs)

  - $TU = \log10(\frac{concentration}{EC_{50}})$

--
<br><br><br><br>
## Toxicity Data?

???
Presenter notes can be put here!

---
# US EPA ECOTOX data base

<https://cfpub.epa.gov/ecotox/>
<br/>

--
<img src="/home/scharmueller/Projects/etox-base/talk/data/epa_ecotox_screenshot.png" width="1000px">

---
# US EPA ECOTOX data base

<https://cfpub.epa.gov/ecotox/>
<br/>

.pull-left[

<img src="/home/scharmueller/Projects/etox-base/talk/data/epa_ecotox_screenshot.png" width="500px">

]

.pull-right[
- Toxicity tests data base

  - ~12,000 chemicals

  - ~12,000 taxa

  - ~1,000,000 results

  - Released quarterly

  - Open for download
]

---
# US EPA ECOTOX ➡️ Standartox

<img src="/home/scharmueller/Projects/etox-base/talk/data/organigram2.png" width="1100px" style="position:absolute; left:10px; top:80px;">

---
# Standartox

<http://standartox.uni-landau.de/>
<br/>

--
<img src="/home/scharmueller/Projects/etox-base/talk/data/standartox_screenshot.png" width="1000px">

---
# Standartox

<http://standartox.uni-landau.de/>
<br/>

.pull-left[
<img src="/home/scharmueller/Projects/etox-base/talk/data/standartox_screenshot.png" width="500px">
### Aim: Prepare ecotoxicological test data in a __reproducible__ and __standardized__ way
]

---
# Standartox

<http://standartox.uni-landau.de/>
<br/>

.pull-left[
<img src="/home/scharmueller/Projects/etox-base/talk/data/standartox_screenshot.png" width="500px">
### Aim: Prepare ecotoxicological test data in a __reproducible__ and __standardized__ way
]

.pull-right[
- Web tool / data base

  - Prepares ecotoxicological test data + additional data

  - Cleaning & Harmonisation (e.g. units)

  - __Filters__ test results

  - __Aggregates__ test results

  - Facilitates data retrieval

  - GUI & R-package (in preparation)

  - Manuscript (in preparation)
]

---
# Filter

.pull-left[
- Test parameters

  - Duration (e.g. 48h, 96h)
  - Endpoint (e.g. EC50, NOEC)
  - Effect groups (e.g. Mortality, Growth)
]

---
# Filter

.pull-left[
- Test parameters
  - Duration (e.g. 48h, 96h)
  - Endpoint (e.g. EC50, NOEC)
  - Effect groups (e.g. Mortality, Growth)

- Chemical parameters
  - Chemicals
  - Chemical classes (e.g. Herbicide, Fungicide)
]

.pull-right[
```{r echo=FALSE, message=FALSE, warning=FALSE}
gg_chem_class = readRDS(file.path(datadir_tk, 'gg_chem_class.rds'))
gg_chem_class
```
]

---
# Filter

.pull-left[
- Test parameters
  - Duration (e.g. 48h, 96h)
  - Endpoint (e.g. EC50, NOEC)
  - Effect groups (e.g. Mortality, Growth)

- Chemical parameters
  - Chemicals
  - Chemical classes (e.g. Herbicide, Fungicide)

- Organism parameters
  - Habitat
]

.pull-right[
```{r echo=FALSE, message=FALSE, warning=FALSE}
gg_habi_class = readRDS(file.path(datadir_tk, 'gg_habi_class.rds'))
gg_habi_class
```
]

---
# Filter

.pull-left[
- Test parameters
  - Duration (e.g. 48h, 96h)
  - Endpoint (e.g. EC50, NOEC)
  - Effect groups (e.g. Mortality, Growth)

- Chemical parameters
  - Chemicals
  - Chemical classes (e.g. Herbicide, Fungicide)

- Organism parameters
  - Habitat
  - Occurrence
]

.pull-right[
```{r echo=FALSE, message=FALSE, warning=FALSE}
gg_regi_class = readRDS(file.path(datadir_tk, 'gg_regi_class.rds'))
gg_regi_class
```
]

---
# Aggregate

- __Geometric mean__
- Median
- Minimum
- Maximum

---
# Example: Trichlorofon

.pull-left[
- Filters

  - Chemical: 52-68-6 (Trichlorofon)  
  - Taxon: _Oncorhynchus mykiss_
  - Duration: 48 to 96h
  - Endpoint: EC50

- Aggregate

  - Geometric mean
]

--

.pull-right[
```{r fig.width=9, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE}
gg_trichlorofon = readRDS(file.path(datadir_tk, 'gg_oncor_mykiss_trichlorofon_standartox_ppdb_comparison.rds'))
gg_trichlorofon
```
]

--
.pull-right[
__Result: Single EC50 value for a specific test setting__
]

---
# Example: All chemicals

```{r fig.width=12, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE}
gg_comparison = readRDS(file.path(datadir_tk, 'gg_oncor_mykiss_standartox_output.rds'))
gg_comparison
```

---
# Example: Comparison

```{r fig.width=12, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE}
gg_comparison = readRDS(file.path(datadir_tk, 'gg_oncor_mykiss_standartox_ppdb_comparison.rds'))
gg_comparison
```

<img src="https://sitem.herts.ac.uk/aeru/ppdb/en/new_logo_small.jpg" width="300px" style="position:absolute; top:120px; right:150px;">

--
__92% of Standartox values diverge by less than a factor of 10 from PPDB values__

---
# Conclusions

.pull-left[

`r emo::ji('heavy_minus_sign')`

- Poorly conducted studies included
  - Average them out
]

.pull-right[

`r emo::ji('heavy_plus_sign')`

- Cleaned & easily-accessible ecotoxicological data base
- Constantly updated
- Version control
- Standardized & reproducible aggregations
- Not restricted to specific chemical classes
  - PPDB: pesticides
]

---
# Outlook: R-package

- standartox::
  - Application Programming Interface (API)
  - scriptable
  - __reproducible__

```{r eval=FALSE}
require(standartox)

standartox(cas = '52-68-6',
           endpoint = 'EC50',
           effect = 'MOR',
           duration = c(48, 96),
           taxon = 'Oncorhynchus',
           filtered = TRUE,
           aggregated = TRUE)
```

---
# Appeal

__Ecotoxicology and Ecology need to embrace the computer age and data sources must become machine-readable!__

- Open Source --> Open Data

- Application Programming Interface (API)

- Big challenges...
  - 12,000 chemicals
  - 12,000 taxa 
  - 1,000,000 test results

- ...demand big, scalable solutions

<img src="http://3.bp.blogspot.com/-Wpp0vCJQDHo/U_ZMorXEynI/AAAAAAAAA7Y/Y-f-3brn99k/s1600/GBIF_logo_short_form.gif" width="120px" style="position:absolute; top:200px; right:70px;">

<img src="/home/scharmueller/Projects/etox-base/talk/data/logos/worms.png" width="250px" style="position:absolute; top:300px; right:260px;">

<img src="/home/scharmueller/Projects/etox-base/talk/data/logos/chebi.png" width="250px" style="position:absolute; top:450px; right:150px;">

<img src="/home/scharmueller/Projects/etox-base/talk/data/logos/pubchem.png" width="250px" style="position:absolute; top:570px; right:200px;">

---
layout: false
# [Thank you for your attention!](https://www.researchgate.net/profile/Andreas_Scharmueller)

## Standartox
### A tool for assessing the risk of chemicals

Andreas Scharmüller<br>
Verena C. Schreiner<br>
Ralf B. Schäfer

Quantitative Landscape Ecology<br/>
iES Landau, Institute for Environmental Sciences<br>
University Koblenz-Landau

<img src="https://upload.wikimedia.org/wikipedia/de/thumb/9/9f/Twitter_bird_logo_2012.svg/1259px-Twitter_bird_logo_2012.svg.png" width="20"> @andschar <br/>
<img src="https://seeklogo.com/images/M/mail-icon-logo-28FE0635D0-seeklogo.com.png" width="20"> scharmueller@uni-landau.de

<img src="https://www.uni-koblenz-landau.de/de/landau/fb7/umweltwissenschaften/logos/land-ecol-neu.jpg" width="100">
<img src="https://www.uni-koblenz-landau.de/de/uni/organisation/verwaltung/abteilungen/abt-1/dokumente-ab1/uni-logo-farbig.jpg" width="200">

### 2019-06-29 talk/talk43.Rmd
added talk

--- 2019-06-29 talk/talk169.Rmd
@@
     nature:
-      ratio: '16:9'
+      ratio: '4:3'
       highlightStyle: github
@@
 Bernhardt et al. 2017
-<img src="/home/scharmueller/Projects/etox-base/talk/data/bernhardt_et_al2.png" width="600px">
+<img src="/home/scharmueller/Projects/etox-base/talk/data/bernhardt_et_al_chem.png" width="260px">
 
@@
 background-image: url(/home/scharmueller/Projects/etox-base/talk/data/stream.jpg)
-background-size: 450px
+background-size: 375px
 background-position: 90% 50%
@@
 <br><br><br><br>
-## Toxicity Data?
+### Toxicity Data?
 
@@
 --
-<img src="/home/scharmueller/Projects/etox-base/talk/data/epa_ecotox_screenshot.png" width="1000px">
+<img src="/home/scharmueller/Projects/etox-base/talk/data/epa_ecotox_screenshot.png" width="800px">
 
@@
 
-<img src="/home/scharmueller/Projects/etox-base/talk/data/organigram2.png" width="1100px" style="position:absolute; left:10px; top:80px;">
+<img src="/home/scharmueller/Projects/etox-base/talk/data/organigram2.png" width="850px" style="position:absolute; left:10px; top:130px;">
 
@@
 --
-<img src="/home/scharmueller/Projects/etox-base/talk/data/standartox_screenshot.png" width="1000px">
+<img src="/home/scharmueller/Projects/etox-base/talk/data/standartox_screenshot.png" width="800px">
 
@@
 
-.pull-left[
 - Filters
@@
   - Geometric mean
-]
+
+---
+# Example: Trichlorofon
+
+```{r fig.width=8, fig.height=6, echo=FALSE, warning=FALSE, message=FALSE}
+gg_trichlorofon = readRDS(file.path(datadir_tk, 'gg_oncor_mykiss_trichlorofon_standartox_ppdb_comparison.rds'))
+gg_trichlorofon
+```
 
@@
 
-.pull-right[
-```{r fig.width=9, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE}
-gg_trichlorofon = readRDS(file.path(datadir_tk, 'gg_oncor_mykiss_trichlorofon_standartox_ppdb_comparison.rds'))
-gg_trichlorofon
-```
-]
-
---
-.pull-right[
-__Result: Single EC50 value for a specific test setting__
-]
+### Result: Single EC50 value for a specific test setting
 
@@
 
-```{r fig.width=12, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE}
+```{r fig.width=10, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE}
 gg_comparison = readRDS(file.path(datadir_tk, 'gg_oncor_mykiss_standartox_output.rds'))
@@
 ---
-# Example: Comparison
+# Comparison: PPDB
 
-```{r fig.width=12, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE}
+```{r fig.width=10, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE}
 gg_comparison = readRDS(file.path(datadir_tk, 'gg_oncor_mykiss_standartox_ppdb_comparison.rds'))
@@
 
-<img src="https://sitem.herts.ac.uk/aeru/ppdb/en/new_logo_small.jpg" width="300px" style="position:absolute; top:120px; right:150px;">
+<img src="https://sitem.herts.ac.uk/aeru/ppdb/en/new_logo_small.jpg" width="250px" style="position:absolute; top:140px; right:50px;">
 
 --
-__92% of Standartox values diverge by less than a factor of 10 from PPDB values__
+92% of Standartox values diverge by less than a factor of 10 from PPDB values
 
@@
 
-<img src="http://3.bp.blogspot.com/-Wpp0vCJQDHo/U_ZMorXEynI/AAAAAAAAA7Y/Y-f-3brn99k/s1600/GBIF_logo_short_form.gif" width="120px" style="position:absolute; top:200px; right:70px;">
+<img src="http://3.bp.blogspot.com/-Wpp0vCJQDHo/U_ZMorXEynI/AAAAAAAAA7Y/Y-f-3brn99k/s1600/GBIF_logo_short_form.gif" width="120px" style="position:absolute; bottom:20px; left:70px;">
 
-<img src="/home/scharmueller/Projects/etox-base/talk/data/logos/worms.png" width="250px" style="position:absolute; top:300px; right:260px;">
+<img src="/home/scharmueller/Projects/etox-base/talk/data/logos/worms.png" width="250px" style="position:absolute; bottom:30px; left:260px;">
 
-<img src="/home/scharmueller/Projects/etox-base/talk/data/logos/chebi.png" width="250px" style="position:absolute; top:450px; right:150px;">
+<img src="/home/scharmueller/Projects/etox-base/talk/data/logos/chebi.png" width="250px" style="position:absolute; bottom:140px; right:110px;">
 
-<img src="/home/scharmueller/Projects/etox-base/talk/data/logos/pubchem.png" width="250px" style="position:absolute; top:570px; right:200px;">
+<img src="/home/scharmueller/Projects/etox-base/talk/data/logos/pubchem.png" width="250px" style="position:absolute; bottom:25px; right:90px;">
 

### 2019-07-23 Rmd/summary_db_cols_norman.Rmd
changes

---
date: "`r format(Sys.Date(), '%d.%m.%Y')`"
output:
  html_vignette
  #html_document
editor_options: 
  chunk_output_type: console
params:
  title: "My Title!"
  schema: "information_schema"
  table: "tables"
  output_dir: NULL # https://community.rstudio.com/t/is-it-possible-to-save-the-html-output-in-a-directory-which-is-not-the-one-where-the-rmd-file-resides/3588/6
  cols: NULL
title: "`r params$title`"
---

```{r params, echo=FALSE, include=FALSE}
schema = params$schema
table = params$table
cols = params$cols
dir = file.path(params$output_dir, paste0(schema, '_', table))
# debuging
# schema = 'norman'
# table = 'data1'
# cols = c('nor3', 'nor4')
```

```{r setup, echo=FALSE, include=FALSE}
prj = '/home/scharmueller/Projects/etox-base'
source(file.path(prj, 'R', 'gn_setup.R'))
```

Summary output from data base table __`r paste0(schema, '.', table)`__

```{r echo=FALSE, message=FALSE}
# schema = 'norman'
# table = 'data2_newest'
# cols = c('nor88test', 'nor88')

## table
tab = summary_db_all(schema = schema, table = table)
## cols
l = summary_db_cols(schema = schema, table = table, cols = cols)
## names lookup table
q = "SELECT *
     FROM norman.variables"
var = read_tbl(user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
               query = q)
```

```{r preparation, echo=FALSE, message=FALSE}
## table
tab[var, header := i.norman_header, on = c(column = 'id') ]
setcolorder(tab, c('column', 'header', 'type'))
## cols
# TODO names(l) = var[ var$id %in% names(l), norman_header ]
# doesn't work b/c of get(col) in ggplot call
```

```{r function, echo=FALSE}
## plot function
gg_summary = function(l, col, limit, var) {

  dat = l[[col]]
  dat[ , var20 := substr(get(col), 1, 20) ]
  header = var[ match(names(dat)[1], var$id), norman_header ]

  gg = ggplot(dat[ 1:limit ], aes(x = reorder(var20, -n), y = n)) +
    geom_bar(stat = 'identity') +
    geom_text(aes(label = n), position=position_dodge(width=0.9), vjust=-0.25) +
    scale_y_continuous(expand = expand_scale(mult = c(0, 0.05))) +
    labs(title = header, subtitle = col, x = NULL) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

  return(gg)
}
```

## Table
```{r table, echo=FALSE, warning=FALSE}
knitr::kable(tab)
```

## Plots
```{r plots, echo=FALSE, fig.height=7, fig.width=15, dpi=72, warning=FALSE}
# options , dpi=36
mapply(gg_summary,
       col = names(l),
       MoreArgs = list(l = l,
                       limit = 50,
                       var = var),
       SIMPLIFY = FALSE)
```

```{r write, echo=FALSE, message=FALSE}
unlink(dir, recursive = TRUE)
mkdirs(dir)
for (i in seq_along(l)) {
  dat = l[[i]]
  nam = names(l)[i]
  fl = file.path(dir, paste0(nam, '.csv'))

  message('Writing: ', fl)
  fwrite(dat, fl)
}
```

### 2019-07-23 Rmd/summary_db_cols_standartox.Rmd
changes

--- 2019-07-23 Rmd/summary_db_cols_norman.Rmd
@@
 # debuging
-# schema = 'norman'
-# table = 'data1'
-# cols = c('nor3', 'nor4')
+# schema = 'standartox'
+# table = 'refs'
+# cols = NULL
 ```
@@
 # schema = 'norman'
-# table = 'data2_newest'
+# table = 'data2'
 # cols = c('nor88test', 'nor88')
@@
 tab = summary_db_all(schema = schema, table = table)
-## cols
+
+## column
 l = summary_db_cols(schema = schema, table = table, cols = cols)
-## names lookup table
-q = "SELECT *
-     FROM norman.variables"
-var = read_tbl(user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
-               query = q)
-```
-
-```{r preparation, echo=FALSE, message=FALSE}
-## table
-tab[var, header := i.norman_header, on = c(column = 'id') ]
-setcolorder(tab, c('column', 'header', 'type'))
-## cols
-# TODO names(l) = var[ var$id %in% names(l), norman_header ]
-# doesn't work b/c of get(col) in ggplot call
 ```
@@
 ## plot function
-gg_summary = function(l, col, limit, var) {
+gg_summary = function(l, col, limit) {
 
@@
   dat[ , var20 := substr(get(col), 1, 20) ]
-  header = var[ match(names(dat)[1], var$id), norman_header ]
 
@@
     scale_y_continuous(expand = expand_scale(mult = c(0, 0.05))) +
-    labs(title = header, subtitle = col, x = NULL) +
+    labs(title = col, x = NULL) +
     theme(axis.text.x = element_text(angle = 45, hjust = 1))
@@
        MoreArgs = list(l = l,
-                       limit = 50,
-                       var = var),
+                       limit = 50),
        SIMPLIFY = FALSE)

### 2019-07-23 R/bd_epa_changes.R

@@ -12,13 +12,13 @@ drv = dbDriver("PostgreSQL")
 con = dbConnect(drv, user = DBuser, dbname = DBetox, host = DBhost, port = DBport, password = DBpassword)
 
 dbSendQuery(con, q1)
-dbSendQuery(con, q2) # TODO takes long!
+dbSendQuery(con, q2)
 
 dbDisconnect(con)
 dbUnloadDriver(drv)
 
 # log ---------------------------------------------------------------------
-log_msg("ECOTOX added functions")
+log_msg("ECOTOX changes")
 
 # clean -------------------------------------------------------------------
 clean_workspace()

### 2019-07-23 R/bd_epa_download.R

@@ -53,8 +53,7 @@ saveRDS(etoxdir, file.path(cachedir, 'etox_data_path.rds'))
 
 # check -------------------------------------------------------------------
 if (length(releases) == 0) {
-  msg = 'Newest etox file (DBetox) file can not be found.'
-  log_msg(msg); rm(msg)
+  log_msg('Newest etox file (DBetox) file can not be found.')
 }
 
 # cleaning ----------------------------------------------------------------

### 2019-07-23 R/bd_norman.R

@@ -4,10 +4,17 @@
 source(file.path(src, 'gn_setup.R'))
 
 # data --------------------------------------------------------------------
+nor_lookup = fread(file.path(normandir, 'lookup', 'lookup_variables.csv'))
 fl = file.path(sql, 'norman_data1.sql')
 q1 = readChar(fl, file.info(fl)$size)
 fl = file.path(sql, 'norman_data2.sql')
 q2 = readChar(fl, file.info(fl)$size)
+fl = file.path(sql, 'norman_data1_newest.sql')
+q1_newest = readChar(fl, file.info(fl)$size)
+fl = file.path(sql, 'norman_data2_newest.sql')
+q2_newest = readChar(fl, file.info(fl)$size)
+fl = file.path(sql, 'norman_data2_removed.sql')
+q2_removed = readChar(fl, file.info(fl)$size)
 
 # query -------------------------------------------------------------------
 drv = dbDriver("PostgreSQL")
@@ -16,11 +23,49 @@ con = dbConnect(drv, user = DBuser, dbname = DBetox, host = DBhost, port = DBpor
 dbSendQuery(con, "DROP SCHEMA IF EXISTS norman CASCADE;")
 dbSendQuery(con, "CREATE SCHEMA norman;")
 
+dbWriteTable(con, nor_lookup, name = c('norman', 'variables'),
+             overwrite = TRUE, row.names = FALSE)
+
 dbSendQuery(con, q1) # cleaned data
 dbSendQuery(con, q2) # converted data
+dbSendQuery(con, q1_newest) # cleaned newest data (acc to published date)
+dbSendQuery(con, q2_newest) # cleaned newest data (acc to published date)
+dbSendQuery(con, q2_removed) # removed tests
 
 dbDisconnect(con)
 dbUnloadDriver(drv)
+
+# summary -----------------------------------------------------------------
+output_dir = file.path(normandir, 'summary')
+rmarkdown::render(file.path(srcrmd, 'summary_db_cols_norman.Rmd'),
+                  output_dir = output_dir,
+                  output_file = 'summary_NORMAN_raw',
+                  params = list(
+                    title = 'NORMAN raw data summary',
+                    schema = 'norman',
+                    table = 'data1',
+                    # cols = c('nor88test', 'nor88')
+                    output_dir = output_dir #! strange that output_dir has to be specified 2 times
+                  ))
+rmarkdown::render(file.path(srcrmd, 'summary_db_cols_norman.Rmd'),
[…]

### 2019-07-23 R/bd_postgres_permissions.R

@@ -14,7 +14,7 @@ con = dbConnect(drv, user = DBuser, dbname = DBetox, host = DBhost, port = DBpor
 dbSendQuery(con, paste0("GRANT CONNECT ON DATABASE ", DBetox, " TO ", role, ";"))
 dbSendQuery(con, paste0("ALTER DEFAULT PRIVILEGES IN SCHEMA public
                           GRANT SELECT ON TABLES TO ", role, ";"))
-dbSendQuery(con, "GRANT USAGE ON SCHEMA application TO ecotox_read;")
+dbSendQuery(con, "GRANT USAGE ON SCHEMA standartox TO ecotox_read;")
 
 dbDisconnect(con)
 dbUnloadDriver(drv)

### 2019-07-23 R/bd_standartox.R

@@ -1,0 +1,84 @@
+# script to build application data
+
+# setup -------------------------------------------------------------------
+source(file.path(src, 'gn_setup.R'))
+
+# data --------------------------------------------------------------------
+fl = file.path(sql, 'standartox_data1.sql')
+q1 = readChar(fl, file.info(fl)$size)
+fl = file.path(sql, 'standartox_data2.sql')
+q2 = readChar(fl, file.info(fl)$size)
+fl = file.path(sql, 'standartox_data2_fin.sql')
+q2_fin = readChar(fl, file.info(fl)$size)
+
+# query -------------------------------------------------------------------
+drv = dbDriver("PostgreSQL")
+con = dbConnect(drv, user = DBuser, dbname = DBetox, host = DBhost, port = DBport, password = DBpassword)
+
+dbSendQuery(con, "DROP SCHEMA IF EXISTS standartox CASCADE;")
+dbSendQuery(con, "CREATE SCHEMA standartox;")
+
+# dbSendQuery(con, q1) # cleaned data
+dbSendQuery(con, q2) # converted data
+dbSendQuery(con, q2_fin) # combined data
+
+# dbSendQuery(con, "GRANT USAGE ON SCHEMA standartox TO jupke;") # TODO only temporarily! remove!
+# dbSendQuery(con, "GRANT SELECT ON TABLE standartox.data TO jupke;")
+# dbSendQuery(con, "GRANT SELECT ON TABLE standartox.data2 TO jupke;")
+
+dbDisconnect(con)
+dbUnloadDriver(drv)
+
+# summary -----------------------------------------------------------------
+rmarkdown::render(file.path(srcrmd, 'summary_db_cols_standartox.Rmd'),
+                  output_dir = summdir,
+                  output_file = 'summary_standartox_tests',
+                  params = list(
+                    title = 'Standartox raw data summary',
+                    schema = 'standartox',
+                    table = 'tests',
+                    output_dir = summdir #! strange that output_dir has to be specified 2 times
+                  ))
+rmarkdown::render(file.path(srcrmd, 'summary_db_cols_standartox.Rmd'),
+                  output_dir = summdir,
+                  output_file = 'summary_standartox_taxa',
+                  params = list(
+                    title = 'Standartox raw data summary',
+                    schema = 'standartox',
+                    table = 'taxa',
+                    output_dir = summdir #! strange that output_dir has to be specified 2 times
+                  ))
[…]

### 2019-07-23 R/cpy_shiny.R

@@ -1,0 +1,18 @@
+# script to copy files to shiny directory
+
+# setup -------------------------------------------------------------------
+source(file.path(src, 'gn_setup.R'))
+dir = gsub('etox', '', DBetox)
+
+# copy --------------------------------------------------------------------
+files = list.files(file.path(export, dir),
+                   full.names = TRUE)
+to = file.path(shinydata, dir)
+
+file_cpy(files, to)
+
+# log ---------------------------------------------------------------------
+log_msg(paste0('Export: application data copied to: ', to))
+
+# cleaning ----------------------------------------------------------------
+clean_workspace()

### 2019-07-23 R/da_epa1.R

@@ -1,157 +1,0 @@
-# script to prepare EPA ECOTOX data
-
-# setup -------------------------------------------------------------------
-source(file.path(src, 'gn_setup.R'))
-# query
-source(file.path(src, 'da_epa_query.R'))
-# # merge: addition
-# epa_taxa = readRDS(file.path(cachedir, 'epa_taxa.rds'))
-# source(file.path(src, 'da_epa_doses.R'))
-# source(file.path(src, 'da_epa_endpoints.R'))
-# source(file.path(src, 'da_epa_exposure_type.R'))
-# source(file.path(src, 'da_epa_media.R'))
-# source(file.path(src, 'da_epa_statistics.R'))
-
-# query -------------------------------------------------------------------
-online_db = FALSE
-if (online_db) {
-  drv = dbDriver("PostgreSQL")
-  con = dbConnect(drv, user = DBuser, dbname = DBetox, host = DBhost, port = DBport, password = DBpassword)
-  
-  res = dbGetQuery(con, "SELECT DISTINCT ON (test_cas) test_cas
-                         FROM ecotox.tests
-                         ORDER BY test_cas ASC")
-  todo_cas = sort(res$test_cas) # all the CAS in the EPA ECOTOX database
-  # todo_cas = todo_cas[1:10] # debug me!
-  
-  epa1_l <- list()
-  for (i in seq_along(todo_cas)) {
-    casnr <- todo_cas[i]
-    d <- dbGetQuery(con, sprintf(q, casnr))
-    
-    message('Returning ', '(', i, '/', length(todo_cas), '): ', casnr, ' (n = ', nrow(d), ')')
-    
-    epa1_l[[i]] <- d
-    names(epa1_l)[i] <- casnr
-    
-  }
-  
-  dbDisconnect(con)
-  dbUnloadDriver(drv)
-  
-  ## cleaning
-  rm(d, con, drv, i) # larger objects
-  rm(casnr, todo_cas) # vectors
-  
-  ## save
-  saveRDS(epa1_l, file.path(cachedir, 'source_epa1_list.rds'))
-  
-} else {
-  epa1_l = readRDS(file.path(cachedir, 'source_epa1_list.rds'))
-}
-
-epa1 = rbindlist(epa1_l)
-
-# preparation -------------------------------------------------------------
[…]

### 2019-07-23 R/da_epa2.R

@@ -1,111 +1,0 @@
-# script to prepare and clean EPA ECOTOX data
-
-# setup -------------------------------------------------------------------
-source(file.path(src, 'gn_setup.R'))
-
-# data --------------------------------------------------------------------
-epa2 = readRDS(file.path(cachedir, 'epa1.rds'))
-
-## conversions
-source(file.path(src, 'da_epa_result_unit.R'))
-source(file.path(src, 'da_epa_test_duration.R'))
-
-## additions
-# chemical classification
-ch_info_fin = readRDS(file.path(cachedir, 'ch_info_fin.rds'))
-# taxonomic information
-tx_info_fin = readRDS(file.path(cachedir, 'tx_info_fin.rds'))
-
-# preparation -------------------------------------------------------------
-## special characters
-for (col in names(epa2)) {
-  set(epa2, i = which(epa2[[col]] %in% c('NC', 'NR', '+ NR', '--', '')), j = col, value = NA)
-}
-## type conversion
-## concentration
-epa2[ grep('ca|x|~', conc1_mean), conc1_mean := NA ]
-epa2[ , conc1_mean := gsub(',', '', conc1_mean) ]
-# clean concentration column
-epa2[ , conc1_mean := trimws(gsub('^\\+|<|>|\\*|=', '', conc1_mean)) ] # remove '*' and '+' (TODO don't know '+')
-
-# type conversion ---------------------------------------------------------
-## concentration
-epa2[ , conc1_mean := as.numeric(conc1_mean) ]
-## duration
-epa2[ , obs_duration_mean := as.numeric(obs_duration_mean) ]
-
-# conversions -------------------------------------------------------------
-## units
-# remove unuseable units
-epa2 = merge(epa2, unit_fin, by = 'conc1_unit', all.x = TRUE)
-epa2[ unit_noscience %in% 1L, rm := 'unit' ] # remove nonsense units
-epa2[ ! unit_convert == 1L, rm := 'unit' ] # keep conversion units only
-epa2[ unit_convert == 1L, conc1_mean_conv := conc1_mean %*na% unit_multi ]
-epa2[ unit_convert == 1L, conc1_unit_conv := unit_conv_to ]
-# cleaning
-cols_rm = grep('^unit_', names(epa2), value = TRUE)
-epa2[ , (cols_rm) := NULL ]
-
-## durations
-epa2 = merge(epa2, look_dur, by.x = 'obs_duration_unit', by.y = 'dur_unit',
[…]

### 2019-07-23 R/da_epa_endpoints.R

@@ -1,95 +1,0 @@
-# script to retrieve all endpoints from the current EPA ECOTOX data base
-
-# setup -------------------------------------------------------------------
-source(file.path(src, 'gn_setup.R'))
-
-# data base
-DBetox = readRDS(file.path(cachedir, 'data_base_name_version.rds'))
-
-# query -------------------------------------------------------------------
-if (online_db) {
-  drv = dbDriver("PostgreSQL")
-  con = dbConnect(drv, user = DBuser, dbname = DBetox, host = DBhost, port = DBport, password = DBpassword)
-  
-  epts = dbGetQuery(con, "select distinct endpoint, count(endpoint) n
-                          from ecotox.results
-                          group by endpoint
-                          order by n desc")
-  setDT(epts)
-  
-  dbDisconnect(con)
-  dbUnloadDriver(drv)
-  
-  saveRDS(epts, file.path(cachedir, 'epts.rds'))
-  
-} else {
-  
-  epts = readRDS(file.path(cachedir, 'epts.rds'))
-}
-
-# preparation -------------------------------------------------------------
-# remove / and *
-epts[ , endpoint_cl := gsub('/|\\*', '', endpoint) ] # clean endpoint
-# combine endpoints
-epts[ grep('^[A-z]{0,2}0$', endpoint_cl), 
-      endpoint_grp := 'XX0' ]
-epts[ grep('^[A-z]{0,2}01$', endpoint_cl), 
-      endpoint_grp := 'XX01' ]
-epts[ grep('^[A-z]{0,2}10$', endpoint_cl), 
-      endpoint_grp := 'XX10' ]
-epts[ grep('^[A-z]{0,2}15$', endpoint_cl), 
-      endpoint_grp := 'XX15' ]
-epts[ grep('^[A-z]{0,2}16$', endpoint_cl), 
-      endpoint_grp := 'XX16' ]
-epts[ grep('^[A-z]{0,2}20', endpoint_cl),
-      endpoint_grp := 'XX20' ]
-epts[ grep('^[A-z]{0,2}50', endpoint_cl),
-      endpoint_grp := 'XX50' ]
-epts[ grep('^[A-z]{0,2}95', endpoint_cl),
-      endpoint_grp := 'XX95' ]
-epts[ grep('^[A-z]{0,2}99', endpoint_cl),
-      endpoint_grp := 'XX99' ]
-epts[ grep('^[A-z]{0,2}100', endpoint_cl),
-      endpoint_grp := 'XX100' ]
[…]

### 2019-07-23 R/da_epa_exposure_type.R

@@ -1,76 +1,0 @@
-# script to assign exposure type entries (e.g. S - Static (water), GV - Gavage, etc.)
-
-# setup -------------------------------------------------------------------
-source(file.path(src, 'gn_setup.R'))
-
-# data base
-DBetox = readRDS(file.path(cachedir, 'data_base_name_version.rds'))
-
-# (1) query ---------------------------------------------------------------
-if (online_db) {
-  drv = dbDriver("PostgreSQL")
-  con = dbConnect(drv, user = DBuser, dbname = DBetox, host = DBhost, port = DBport, password = DBpassword)
-  
-  exp_typ = dbGetQuery(con, "SELECT *
-                             FROM ecotox.exposure_type_codes")
-  setDT(exp_typ)
-    
-  dbDisconnect(con)
-  dbUnloadDriver(drv)
-  
-  saveRDS(exp_typ, file.path(cachedir, 'source_epa_exposure_type.rds'))
-  
-} else {
-  
-  exp_typ = readRDS(file.path(cachedir, 'source_epa_exposure_type.rds'))
-}
-
-
-# abbreviations -----------------------------------------------------------
-
-## Diet (D) exposure codes ----
-diet = c('CH', 'DR', 'DT', 'FD', 'GE', 'GV', 'IG', 'LC', 'OR')
-# FD - chemical incorporated in the Food, GV - Gavage (oral intubation)
-
-## Injection (I) codes ----
-inj = c('IA', 'IAC', 'IB', 'IC', 'ICL', 'ID', 'IE', 'IF', 'IH', 'II', 'IJ', 'IK', 'ILP', 'IM', 'IO', 'IP', 'IQ', 'IS', 'IU', 'IV', 'IY', 'IZ', 'OP', 'SC', 'SD', 'YK')
-
-## Multiple (M) Application codes ----
-multi = 'MU'
-# MU - Multiple routes between application codes (e.g. dermal and injection)
-
-## Aquatic Lab exposure types ----
-aqu_lab = c('S', 'R', 'F', 'P', 'L', 'AQUA - NR')
-# S - Static, R - Renewal, F - Flow through, AQUA-NR - Aquatic - not reported
-
-## Aquatic Field Exposure types ----
-aqu_field = c('B', 'E', 'O')
-# B - Tidal, E - Lentic, O - Lotic
-
-## Topical (T) Application codes ----
[…]

### 2019-07-23 R/da_epa_query.R

@@ -1,101 +1,0 @@
-q = "
-    SELECT
-      -- substances
-         tests.test_id,
-         results.result_id,
-         tests.test_cas::varchar AS casnr,
-         chemicals.chemical_name,
-         chemical_carriers.chem_name AS carr_name,
-         chemical_carriers.characteristics AS carr_characteristics,
-         chemicals.ecotox_group,
-      -- result variables
-         results.conc1_min,
-         results.conc1_mean,
-         results.conc1_max,
-         results.conc1_unit,
-         results.conc1_type,
-         results.conc1_comments,
-         results.bcf1_min,
-         results.bcf1_mean,
-         results.bcf1_max,
-         results.ion1,
-         results.measurement,
-         measurement_codes.description,
-         results.chem_analysis_method,
-         results.lipid_pct_min,
-         results.lipid_pct_mean,
-         results.lipid_pct_max,
-      -- test duration
-         results.obs_duration_mean,
-         results.obs_duration_unit,
-         tests.study_duration_mean,
-         tests.study_duration_unit,
-      -- test media
-         tests.substrate,
-         tests.substrate_description,
-      -- application
-         tests.application_freq_mean,
-         tests.application_freq_unit,
-         tests.application_type, -- aerial, 
-      -- result types
-         results.endpoint,
-         results.effect,
-      -- statistical analyses
-         tests.experimental_design,
-         results.significance_comments,
-      -- test dose numbers
-         tests.num_doses_min,
-         tests.num_doses_mean,
-         tests.num_doses_max,
-         tests.created_date,
-         tests.modified_date,
-         tests.published_date,
-         tests.exposure_type,
-         tests.exposure_duration_mean,
-         tests.exposure_duration_unit,
-         tests.test_method,
-         tests.control_type,
-         tests.test_radiolabel,
-         tests.test_purity_min,
-         tests.test_purity_mean_op,
-         tests.test_purity_mean,
-         tests.test_purity_max,
-         tests.media_type,
-         tests.test_type,
-         tests.test_characteristics,
-         tests.test_location,
-         tests.organism_habitat, -- ('soil')
-         tests.subhabitat, -- ('P', 'R', 'L', 'E', 'D', 'F', 'G', 'M') -- Palustrine, Riverine, Lacustrine, Estuarine
-      -- species parameters
-         species.species_number,
-         tests.organism_init_wt_mean,
-         tests.organism_init_wt_unit,
[…]

### 2019-07-23 R/da_epa_stats.R

@@ -1,117 +1,0 @@
-# script to query the EPA data base for descriptive statistics on reported concentrations
-
-# setup -------------------------------------------------------------------
-source(file.path(src, 'gn_setup.R'))
-
-# data base
-DBetox = readRDS(file.path(cachedir, 'data_base_name_version.rds'))
-
-# query -------------------------------------------------------------------
-if (online_db) {
-  drv = dbDriver("PostgreSQL")
-  con = dbConnect(drv, user = DBuser, dbname = DBetox, host = DBhost, port = DBport, password = DBpassword)
-  
-  ## filters ----
-  # concentration type
-  q = 'select conc1_type as conc_type, count(conc1_type) as n
-       from ecotox.results
-       group by conc1_type'
-  coty = as.data.table(dbGetQuery(con, q))
-  coty[ conc_type %in% c('NR', '--', 'NC', ''), conc_type := NA ]
-  coty = coty[ , .(n = sum(n)), conc_type ]
-  setorder(coty, -n)
-  
-  # endpoints
-  q = 'select endpoint, count(endpoint) as n
-       from ecotox.results
-       group by endpoint'
-  epts = as.data.table(dbGetQuery(con, q))
-  epts[ , endpoint := gsub('/|\\*', '', endpoint) ]
-  epts = epts[ , .(n = sum(n)), endpoint]
-  setorder(epts, -n)
-  epts[grep('50', endpoint),]; nrow(epts)
-  
-  # effect groups
-  q = 'select effect, count(effect) as n
-       from ecotox.results
-       group by effect'
-  effe = as.data.table(dbGetQuery(con, q))
-  effe[ , effect := gsub('~|/', '', effect) ]
-  effe = effe[ , .(n = sum(n)), effect]
-  setorder(effe, -n)
-  effe; nrow(effe)
-  
-  # exposure type
-  q = 'select exposure_type, count(exposure_type) n
-       from ecotox.tests
-       group by exposure_type'
-  expo = as.data.table(dbGetQuery(con, q))
-  expo[ , exposure_type := gsub('/|\\*', '', exposure_type) ]
-  expo = expo[ , .(n = sum(n)), exposure_type]
-  setorder(expo, -n)
-  expo; nrow(expo)
-  
-  # test dates
-  q = 'select published_date, created_date, application_date
-       from ecotox.tests'
[…]

### 2019-07-23 R/da_epa_test_duration.R

@@ -1,46 +1,0 @@
-# script to convert duration units
-
-# setup -------------------------------------------------------------------
-source(file.path(src, 'gn_setup.R'))
-
-# data --------------------------------------------------------------------
-## postgres
-if (online_db) {
-  drv = dbDriver("PostgreSQL")
-  con = dbConnect(drv,
-                  user = DBuser,
-                  dbname = DBetox,
-                  host = DBhost,
-                  port = DBport,
-                  password = DBpassword)
-  
-  dur = dbGetQuery(con, 'SELECT distinct on (obs_duration_unit) obs_duration_unit
-                         FROM ecotox.results')
-  setDT(dur)
-  dur[ obs_duration_unit == '', obs_duration_unit := NA ]
-  dur = dur[ !is.na(obs_duration_unit) ]
-  
-  dbDisconnect(con)
-  dbUnloadDriver(drv)
-  
-  saveRDS(dur, file.path(cachedir, 'epa_durations.rds'))
-  
-} else {
-  
-  dur = readRDS(file.path(cachedir, 'epa_durations.rds'))
-}
-## lookup csv
-look_dur = fread(file.path(lookupdir, 'lookup_duration.csv'), na.strings = '')
-
-# check -------------------------------------------------------------------
-chck_duration = nrow(dur[ !obs_duration_unit %in% look_dur$unit ])
-if (chck_duration != 0) {
-  log_warn('Not classified duration units.')
-}
-
-# preparation -------------------------------------------------------------
-setnames(look_dur, paste0('dur_', names(look_dur)))
-
-
-
-

### 2019-07-23 R/exp_application.R

@@ -1,16 +1,0 @@
-# script to export application data
-
-# setup -------------------------------------------------------------------
-source(file.path(src, 'gn_setup.R'))
-
-# save --------------------------------------------------------------------
-export_tbl(schema = 'application', table = 'data2', type = 'fst', debug = TRUE,
-           user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox)
-
-# log ---------------------------------------------------------------------
-log_msg('Export: application data exported')
-
-# cleaning ----------------------------------------------------------------
-clean_workspace()
-
-

### 2019-07-23 R/exp_norman.R

@@ -5,18 +5,30 @@ source(file.path(src, 'gn_setup.R'))
 
 # save --------------------------------------------------------------------
 ## raw data
-export_tbl(schema = 'norman', table = 'data', type = 'csv',
-           user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox)
-## cleaned data
-export_tbl(schema = 'norman', table = 'data2', type = 'csv',
-           user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox)
+# export_tbl(schema = 'norman', table = 'data1', type = 'csv',
+#            user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
+#            dir = file.path(normandir, 'export'))
+export_tbl(schema = 'norman', table = 'data1_newest', type = 'csv',
+           user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
+           dir = file.path(normandir, 'export'))
+# export_tbl(schema = 'norman', table = 'data2', type = 'csv',
+#            user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
+#            dir = file.path(normandir, 'export'))
+export_tbl(schema = 'norman', table = 'data2_newest', type = 'csv',
+           user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
+           dir = file.path(normandir, 'export'))
+export_tbl(schema = 'norman', table = 'data2_removed', type = 'csv',
+           user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
+           dir = file.path(normandir, 'export'))
+export_tbl(schema = 'norman', table = 'variables', type = 'csv',
+           user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
+           dir = file.path(normandir, 'export'))
 
-# summary -----------------------------------------------------------------
-summary_nor1 = summary_db_all(schema = 'norman', table = 'data1')
-summary_nor2 = summary_db_all(schema = 'norman', table = 'data2')
-
-fwrite(summary_nor1, file.path(normandir, 'summary_norman1.csv'))
-fwrite(summary_nor2, file.path(normandir, 'summary_norman2.csv'))
+# copy --------------------------------------------------------------------
+file.copy(from = normandir,
+          to = cloud,
+          recursive = TRUE,
+          overwrite = TRUE)
 
[…]

### 2019-07-23 R/exp_standartox.R

@@ -1,0 +1,44 @@
+# script to export application data
+
+# setup -------------------------------------------------------------------
+source(file.path(src, 'gn_setup.R'))
+
+# meta information --------------------------------------------------------
+q = paste0("SELECT *
+            FROM meta.info")
+
+meta = read_tbl(user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
+                query = q)
+v = meta$etox_version
+## output directory
+dir = file.path(export, v)
+unlink(dir)
+mkdirs(dir) # TODO put this into export_tbl()
+
+# export to etox-base repo ------------------------------------------------
+## data
+# as .fst object
+export_tbl(schema = 'standartox', table = 'data2', type = 'fst', debug = FALSE,
+           user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
+           dir = dir, file_name = paste0('standartox', v))
+# as .feather object
+export_tbl(schema = 'standartox', table = 'data2', type = 'feather', debug = FALSE,
+           user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
+           dir = dir, file_name = paste0('standartox', v))
+# as .rds object
+export_tbl(schema = 'standartox', table = 'data2', type = 'rds', debug = FALSE,
+           user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
+           dir = dir, file_name = paste0('standartox', v))
+# as .csv object
+export_tbl(schema = 'standartox', table = 'data2', type = 'csv', debug = FALSE,
+           user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
+           dir = dir, file_name = paste0('standartox', v))
+## meta data
+fwrite(meta, file.path(dir, paste0('standartox', v, '_meta.csv')))
+
+# log ---------------------------------------------------------------------
+log_msg('Export: application data exported')
+
+# cleaning ----------------------------------------------------------------
+clean_workspace()
+

### 2019-07-23 R/exp_standartox_shiny_stats.R

@@ -1,0 +1,61 @@
+# script to export summary statistics to be used in shiny application
+
+# setup -------------------------------------------------------------------
+source(file.path(src, 'gn_setup.R'))
+v = gsub('etox', '', DBetox)
+
+# data --------------------------------------------------------------------
+header = names(read_fst(file.path(export, v, paste0('standartox', v, '.fst')), to = 1))
+
+# query -------------------------------------------------------------------
+# cols
+cols = c('effect', 'endpoint', 'conc1_type', 'test_location', 'obs_duration_mean2',
+         grep('hab_', header, value = TRUE),
+         grep('reg_', header, value = TRUE),
+         grep('ccl_', header, value = TRUE))
+# loop
+l = list()
+for (i in seq_along(cols)) {
+  
+  col = cols[i]
+  message('Fetching: ', col)
+  con = DBI::dbConnect(RPostgreSQL::PostgreSQL(),
+                       dbname = DBetox,
+                       host = DBhost,
+                       port = DBport,
+                       user = DBuser,
+                       password = DBpassword)
+  dat = summary_db_perc(con, 'standartox', 'data2', col)
+  setDT(dat)
+  
+  l[[i]] = dat
+  names(l)[i] = col
+}
+
+# preparation -------------------------------------------------------------
+stat = rbindlist(l, idcol = TRUE)
+setnames(stat, c('variable', 'value', 'n', 'n_tot'))
+stat = stat[ !is.na(value) ]
+stat[ value == '1', value := variable ]
+stat[ , name := firstup(gsub('^([a-z]{3}_)', '', value)) ]
+stat[ name == 'America_north', name := 'North America' ]
+stat[ name == 'America_south', name := 'South America' ]
+stat[ grep('hab_', variable), variable := 'habitat' ]
+stat[ grep('reg_', variable), variable := 'region' ]
+stat[ grep('ccl_', variable), variable := 'chemical_class' ]
+stat[ , name_perc := paste0(name, ' (', round(n / n_tot * 100), '%)') ]
+stat_l = split(stat, stat$variable)
+
+stat_l$obs_duration_mean2 = stat_l$obs_duration_mean2[ , range(na.omit(as.numeric(value))) ]
+
+# write -------------------------------------------------------------------
+saveRDS(stat_l, file.path(export, v, paste0('standartox', v, '_shiny_stats.rds')))
+save(stat_l, file = file.path(export, v, paste0('standartox', v, '_shiny_stats.rda'))) # for standartox::
+
[…]

### 2019-07-23 R/fun_export_db.R

@@ -1,6 +1,6 @@
 # function to export postgres tables
 
-export_tbl = function(schema, table, dir = NULL, type = c('csv', 'rds', 'fst'), debug_mode = NULL,
+export_tbl = function(schema, table, dir = NULL, file_name = NULL, type = c('csv', 'rds', 'fst', 'feather'), debug_mode = FALSE,
                       user = NULL, host = NULL, port = NULL, password = NULL, dbname = NULL) {
   # file
   if (is.null(dir)) {
@@ -8,22 +8,26 @@ export_tbl = function(schema, table, dir = NULL, type = c('csv', 'rds', 'fst'), 
   }
   # query
   q = paste0("SELECT * FROM ", schema, ".", table)
-  if (!is.null(debug_mode)) {
-    paste0(q, ' LIMIT 100')
+  if (debug_mode) {
+    q = paste0(q, ' LIMIT 100')
   }
   # export
   drv = dbDriver("PostgreSQL")
   con = dbConnect(drv, user = DBuser, dbname = DBetox, host = DBhost, port = DBport, password = DBpassword)
   
-  message('Exporting table: ', schema, '.', table, ' to ', dir)
+  message('Exporting table: ', schema, '.', table, ' (.', type, ') to ', dir)
   dat = dbGetQuery(con, q)
   setDT(dat)
   
   dbDisconnect(con)
   dbUnloadDriver(drv)
   # save
-  fl = file.path(dir, table)
-  type = match.arg(type, c('csv', 'rds', 'fst'))
+  if (is.null(file_name)) {
+    fl = file.path(dir, table)
+  } else {
+    fl = file.path(dir, file_name)
+  }
+  type = match.arg(type, c('csv', 'rds', 'fst', 'feather'))
   if (type == 'csv') {
     fwrite(dat, paste0(fl, '.csv'))
   }
@@ -32,6 +36,9 @@ export_tbl = function(schema, table, dir = NULL, type = c('csv', 'rds', 'fst'), 
   }
   if (type == 'fst') {
     write_fst(dat, paste0(fl, '.fst'))
+  }
+  if (type == 'feather') {
+    write_feather(dat, paste0(fl, '.feather'))
   }
 }
 

### 2019-07-23 R/fun_file_cpy.R

@@ -1,0 +1,15 @@
+# function to copy files and create folders if needed
+# TODO add to asmisc
+
+file_cpy = function(files = NULL, to = NULL, overwrite_dir = TRUE, ...) {
+  # remove folder
+  unlink(to)
+  mkdirs(to)
+  # paths
+  from = files
+  to2 = file.path(to, basename(files))
+  # copy
+  file.copy(from = from,
+            to = to2,
+            overwrite = TRUE)
+}

### 2019-07-23 R/fun_sql_builder.R

@@ -1,19 +1,5 @@
 # functions to build sql queries
 
-
-# chemical names ----------------------------------------------------------
-# function(dat, ) {
-#   
-#   dat$column_name_full = with(dat, paste0(table_name, '.', column_name))
-#   
-#   select = 
-#   
-#   
-#   
-# }
-
-
-
 # chemical class flags ----------------------------------------------------
 # find the greatest (0,1) value of multiple columns in muktiple tables joined together by one key column
 

### 2019-07-23 R/fun_summary_db.R

@@ -1,76 +1,85 @@
 # function to summarise Postgres data base columns
 
 # categorical varaibles ---------------------------------------------------
-summary_db_cat = function(schema, table, table_description = NULL, col) {
-  
-  drv = dbDriver("PostgreSQL")
-  con = dbConnect(drv, user = DBuser, dbname = DBetox, host = DBhost, port = DBport, password = DBpassword)
-  
-  if (is.null(table_description)) {
-    q = paste0("SELECT t1.", col, ", count(*) AS n
-                FROM ", paste0(schema, '.', table), " AS t1
-                GROUP BY t1.", col, "
-                ORDER BY n DESC")
-  } else {
-    q = paste0("SELECT t1.", col, ", t2.description, count(*) AS n
-                FROM ", paste0(schema, '.', table), " AS t1
-                JOIN ", paste0('ecotox', '.', table_description), " AS t2 ON ", paste0('t1.', col), " = t2.code", "
-                GROUP BY t1.", col, ", t2.description
-                ORDER BY n DESC")
-  }
-  message('Fetching: ', paste(schema, table, col, sep = '.'))
-  dat = dbGetQuery(con, q)
-  setDT(dat)
-  
-  dbDisconnect(con)
-  dbUnloadDriver(drv)
-  
-  return(dat)
-}
+# TODO deprecate
+# summary_db_cat = function(schema, table, table_description = NULL, col) {
+#   
+#   drv = dbDriver("PostgreSQL")
+#   con = dbConnect(drv, user = DBuser, dbname = DBetox, host = DBhost, port = DBport, password = DBpassword)
+#   
+#   if (is.null(table_description)) {
+#     q = paste0("SELECT t1.", col, ", count(*) AS n
+#                 FROM ", paste0(schema, '.', table), " AS t1
+#                 GROUP BY t1.", col, "
+#                 ORDER BY n DESC")
+#   } else {
+#     q = paste0("SELECT t1.", col, ", t2.description, count(*) AS n
+#                 FROM ", paste0(schema, '.', table), " AS t1
+#                 JOIN ", paste0('ecotox', '.', table_description), " AS t2 ON ", paste0('t1.', col), " = t2.code", "
+#                 GROUP BY t1.", col, ", t2.description
+#                 ORDER BY n DESC")
+#   }
+#   message('Fetching: ', paste(schema, table, col, sep = '.'))
+#   dat = dbGetQuery(con, q)
+#   setDT(dat)
+#   
+#   dbDisconnect(con)
[…]

### 2019-07-23 R/fun_write_db.R

@@ -20,8 +20,14 @@ write_tbl = function(df = NULL, user = NULL, host = NULL, port = NULL, password 
   }
   
   ## data base
-  drv = dbDriver('PostgreSQL')
-  con = dbConnect(drv, dbname = dbname, user = user, host = host, port = port, password = password)
+  con = DBI::dbConnect(RPostgreSQL::PostgreSQL(), # RPostgres::Postgres(),
+                       dbname = dbname,
+                       host = host,
+                       port = port,
+                       user = user,
+                       password = password)
+                       # bigint = 'integer') # to not return integer64 (only for RPostgres::)
+  on.exit(DBI::dbDisconnect(con))
   
   dbSendQuery(con, paste0("CREATE SCHEMA IF NOT EXISTS ", schema, ";"))
   # write to DB
@@ -47,11 +53,7 @@ write_tbl = function(df = NULL, user = NULL, host = NULL, port = NULL, password 
     dbSendQuery(con, paste0("ALTER TABLE ", schema, ".", tbl,
                             " ADD PRIMARY KEY (", key, ");"))
   }
-  
-  
-  dbDisconnect(con)
-  dbUnloadDriver(drv)
-  
+
   message(Sys.time(), ' Table created in: ', path)
 }
 

### 2019-07-23 R/gn_setup.R

@@ -30,6 +30,7 @@ pkg_cran = c(
   'purrr',
   'data.table',
   'RPostgreSQL',
+  'DBI',
   'vegan',
   'plyr',
   'outliers',
@@ -77,28 +78,31 @@ sink_console = TRUE # sink console to file
 # variables ---------------------------------------------------------------
 cachedir = file.path(prj, 'cache')
 option = file.path(prj, 'options')
-missingdir = file.path(prj, 'missing')
 meta = file.path(prj, 'meta')
-fundir = file.path(prj, 'functions')
 plotdir = file.path(prj, 'plots')
 src = file.path(prj, 'R')
+srcrmd = file.path(prj, 'Rmd')
 data = file.path(prj, 'data')
 data_ecotox = file.path(data, 'ecotox')
 data_chebi = file.path(data, 'chebi')
 lookupdir = file.path(prj, 'lookup')
-shinydir = paste0(prj, '-shiny')
-shinydata = file.path(shinydir, 'data')
 cred = file.path(prj, 'cred')
-share = file.path(prj, 'share')
-normandir = file.path(prj, 'norman')
 sql = file.path(prj, 'sql')
 export = file.path(prj, 'export')
 summdir = file.path(prj, 'summary')
-# article subfolder
+## article subfolder
 article = file.path(prj, 'article')
 datadir_ar = file.path(article, 'data')
 ## talk subfolder
 datadir_tk = file.path(prj, 'talk', 'data')
+## shiny application
+shinydir = paste0(prj, '-shiny')
+shinydata = file.path(shinydir, 'data')
+## standartox R-package
+standartoxdir = gsub(basename(prj), 'standartox', prj, fixed = TRUE)
+## NORMAN
+normandir = file.path(prj, 'norman')
+cloud = file.path('/home/scharmueller/Nextcloud/norman')
 
 # data base to write to
 if (debug_mode) {
@@ -125,9 +129,11 @@ mdl = 1e6 # max deparse length for writing to sink console
 
 # source ------------------------------------------------------------------
 source(file.path(cred, 'credentials.R')) # data base credentials
-source(file.path(fundir, 'casconv.R')) # convert between CAS and CASNR
+source(file.path(src, 'fun_casconv.R')) # convert between CAS and CASNR
 source(file.path(src, 'fun_clean_workspace.R'))
[…]

### 2019-07-23 R/look_concentration_units.R

@@ -207,7 +207,7 @@ chck_dupl(units, 'conc1_unit')
 # write -------------------------------------------------------------------
 ## postgres
 write_tbl(units, user = DBuser, host = DBhost, port = DBport, password = DBpassword,
-          dbname = DBetox, schema = 'ecotox', tbl = 'conc_lookup',
+          dbname = DBetox, schema = 'ecotox', tbl = 'concentration_unit_lookup',
           key = 'conc1_unit',
           comment = 'Lookup table for concentration units')
 

### 2019-07-23 R/look_norman.R

@@ -6,12 +6,22 @@ source(file.path(src, 'gn_setup.R'))
 # retrieve values ---------------------------------------------------------
 drv = dbDriver("PostgreSQL")
 con = dbConnect(drv, user = DBuser, dbname = DBetox, host = DBhost, port = DBport, password = DBpassword)
-
 chemical_analysis = data.table(dbGetQuery(con,
-  "SELECT * FROM ecotox.chemical_analysis_codes"
+                                          "SELECT codes.*, COUNT(chem_analysis_method) n
+   FROM ecotox.results
+   RIGHT OUTER JOIN ecotox.chemical_analysis_codes codes ON results.chem_analysis_method = codes.code
+   GROUP BY code, description
+   ORDER BY n DESC"
+))
+chemical_analysis = data.table(dbGetQuery(con,
+  "SELECT chem_analysis_method AS code, COUNT(*) n
+   FROM ecotox.results
+   LEFT JOIN ecotox.chemical_analysis_codes codes ON results.chem_analysis_method = codes.code
+   GROUP BY chem_analysis_method, description
+   ORDER BY n DESC"
 ))
 
-effect_codes = data.table(dbGetQuery(con,
+effect = data.table(dbGetQuery(con,
   "SELECT effect_codes.*, COUNT(effect) n
    FROM ecotox.results
    RIGHT OUTER JOIN ecotox.effect_codes ON results.effect = effect_codes.code
@@ -19,31 +29,27 @@ effect_codes = data.table(dbGetQuery(con,
    ORDER BY n DESC"
 ))
 
-endpoint = data.table(dbGetQuery(con,
-  "SELECT endpoint_codes.*, COUNT(endpoint) n
-   FROM ecotox.results
-   RIGHT OUTER JOIN ecotox.endpoint_codes ON results.endpoint = endpoint_codes.code
-   GROUP BY code, description
-   ORDER BY n DESC"
-))
-
 media_type = data.table(dbGetQuery(con,
-  "SELECT media_type_codes.*, COUNT(media_type) n
-   FROM ecotox.tests
-   RIGHT OUTER JOIN ecotox.media_type_codes ON media_type = code
-   GROUP BY media_type, code
+  "SELECT media_type AS code, description, COUNT(*) n
+   FROM ecotox.tests tests
+   LEFT JOIN ecotox.media_type_codes codes ON tests.media_type = codes.code
+   GROUP BY media_type, description
[…]

### 2019-07-23 R/look_norman_acute_chronic_standard.R

@@ -1,0 +1,25 @@
+# useage, acute-chronic and standard-test classification for NORMAN
+# data provided by Peter v. d. Ohe
+
+# setup -------------------------------------------------------------------
+source(file.path(src, 'gn_setup.R'))
+
+# data --------------------------------------------------------------------
+dat = fread(file.path(normandir, 'data', 'lookup_acute_chronic_standard.csv'), na.strings = '')
+
+# chck --------------------------------------------------------------------
+chck_dupl(dat, 'result_id')
+
+# write -------------------------------------------------------------------
+write_tbl(dat, user = DBuser, host = DBhost, port = DBport, password = DBpassword,
+          dbname = DBetox, schema = 'ecotox', tbl = 'lookup_acute_chronic_standard',
+          key = 'result_id',
+          comment = 'NORMAN use, acute-chronic and standard-test lookup table')
+
+# log ---------------------------------------------------------------------
+log_msg('LOOK: NORMAN use, acute-chronic and standard-test tables script run.')
+
+# cleaning ----------------------------------------------------------------
+clean_workspace()
+
+

### 2019-07-23 R/qu_phch_fin.R

@@ -17,8 +17,9 @@ phch_cols = dbGetQuery(con, "SELECT table_schema, table_name, column_name
 setDT(phch_cols)
 
 # chemical classes --------------------------------------------------------
-# cols_exact = c('agrochemical', 'fungicide', 'herbicide', 'insecticide', 'pesticide', 'metal', 'drug')
-cols_exact = phch_cols[ table_name == 'chebi_envi', column_name ]
+cols = c('agrochemical', 'fungicide', 'herbicide', 'insecticide', 'pesticide', 'metal', 'drug')
+cols_exact = phch_cols[ table_name == 'chebi_envi' |
+                          column_name %in% cols, column_name ]
 cols_exact = cols_exact[ ! cols_exact %in% c('cas', 'chebiid') ]
 cols_exact = gsub('(.+)', '^\\1$', cols_exact)
 class_cols = phch_cols[ grep(paste0(cols_exact, collapse = '|'), column_name) ]
@@ -27,6 +28,8 @@ class_cols = phch_cols[ grep(paste0(cols_exact, collapse = '|'), column_name) ]
 q = q_join(class_cols, schema = 'phch', main_tbl = 'epa', col_join = 'cas',
            fun = 'GREATEST', debug = FALSE)
 q = paste0("CREATE TABLE phch_fin.chem_class AS ( ", q, ")")
+# TODO intermediate solution - change!
+q = gsub('SELECT epa.cas', 'SELECT epa.cas_number, epa.cas', q, fixed = TRUE)
 
 dbSendQuery(con, "DROP TABLE IF EXISTS phch_fin.chem_class;")
 dbSendQuery(con, q)
@@ -35,6 +38,7 @@ dbSendQuery(con, "ALTER TABLE phch_fin.chem_class ADD PRIMARY KEY (cas);")
 # chemical names ----------------------------------------------------------
 q = "
   SELECT
+    ep.cas_number,
     ep.cas,
     COALESCE(ci.cname, ch.cname, pp.cname, aw.cname, ep.cname) cname,
     COALESCE(ch.iupac_name, pc.iupac_name, aw.iupac_name) iupacname,

### 2019-07-23 R/qu_worms_prep.R

@@ -23,6 +23,11 @@ cols = c('marin', 'brack', 'fresh', 'terre', 'extinct')
 wo2[ , (cols) := lapply(.SD, as.numeric), .SDcols = cols ]
 wo2 = unique(wo2, by = 'aphiaid') # for safety
 wo2 = wo2[ rank %in% c('Family', 'Genus', 'Species') ] # only fm, gn, sp
+# 0 to NA
+wo2[ marin == 0L, marin := NA_integer_ ]
+wo2[ brack == 0L, brack := NA_integer_ ]
+wo2[ fresh == 0L, fresh := NA_integer_ ]
+wo2[ terre == 0L, terre := NA_integer_ ]
 # split
 wo2_l = split(wo2, wo2$rank)
 names(wo2_l) = c('fm', 'gn', 'sp')

### 2019-07-23 R/run_build.R

@@ -54,9 +54,7 @@ if (build_db) {
   # lookup tables
   source(file.path(src, 'bd_epa_lookup.R'), max.deparse.length = mdl)
   # changes
-  # TODO rwork EPA build process
   source(file.path(src, 'bd_epa_changes.R'), max.deparse.length = mdl)
-
 }
 
 # identifiers -------------------------------------------------------------
@@ -99,34 +97,21 @@ if (build) {
 
 # Application -------------------------------------------------------------
 if (build) {
-  source(file.path(src, 'bd_application.R'), max.deparse.length = mdl)
+  source(file.path(src, 'bd_standartox.R'), max.deparse.length = mdl)
   source(file.path(src, 'bd_norman.R'), max.deparse.length = mdl)
 }
 
 # Export ------------------------------------------------------------------
 if (export) {
-  source(file.path(src, 'exp_application.R'), max.deparse.length = mdl) #! also to shinydir!
+  source(file.path(src, 'exp_standartox.R'), max.deparse.length = mdl)
+  source(file.path(src, 'exp_standartox_shiny_stats.R'), max.deparse.length = mdl)
   source(file.path(src, 'exp_norman.R'), max.deparse.length = mdl)
 }
+
+# copy to shiny dir -------------------------------------------------------
+source(file.path(src, 'cpy_shiny.R'), max.deparse.length = mdl)
+
 # NORMAN ------------------------------------------------------------------
-
-
-
-
-# TODO
-# EPA data scripts
-# source(file.path(src, 'da_epa_run.R'), max.deparse.length = mdl)
-# NORMAN export scripts
-# source(file.path(src, 'da_norman_run.R'), max.deparse.length = mdl)
-
-# file copies -------------------------------------------------------------
-# TODO what's this?
-# copy README.md to shiny repo
-# file.copy('README.md', file.path(shinydir, 'README.md'),
-#           overwrite = TRUE)
-# if (nodename == 'scharmueller') {
-#   source(file.path(src, 'no_share.R'))
-# }
 
 
 # QSAR and PPDB comparison data -------------------------------------------

### 2019-07-23 options/ggplot_theme_etox_base.R

@@ -16,7 +16,7 @@ theme_minimal_etox_base_sans <- theme_minimal(base_size = 12, base_family = 'Ope
   theme(#panel.grid.major = element_blank(),
         #panel.grid.minor = element_blank(),
         text = element_text(size = 14),
-        axis.text = element_text(size = 13),
+        axis.text = element_text(size = 15),
         axis.title.x = element_text(size = 14, face = 'bold', vjust = 0),
         axis.title.y = element_text(size = 14, face = 'bold', vjust = 1),
         legend.position = 'right',

### 2019-08-30 talk/talk43.Rmd
changes

@@
   - ...
+
+### Effect?
 ]
@@
 Bernhardt et al. 2017
-<img src="/home/scharmueller/Projects/etox-base/talk/data/bernhardt_et_al_chem.png" width="260px">
+<img src="/home/scharmueller/Projects/etox-base/talk/data/bernhardt_et_al_chem.png" width="300px">
 
@@
 background-image: url(/home/scharmueller/Projects/etox-base/talk/data/stream.jpg)
-background-size: 375px
+background-size: 400px
 background-position: 90% 50%
@@
 
+.pull-left[
 - Regulatory thresholds
@@
 
---
-<br><br><br><br>
+<br>
 ### Toxicity Data?
+]
 
@@
 
-<img src="/home/scharmueller/Projects/etox-base/talk/data/organigram2.png" width="850px" style="position:absolute; left:10px; top:130px;">
+<img src="/home/scharmueller/Projects/etox-base/talk/data/organigram2.png" width="950px" style="position:absolute; left:10px; top:130px;">
 
@@
 - Web tool / data base
-
-  - Prepares ecotoxicological test data + additional data
 
@@
 
-```{r fig.width=10, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE}
+```{r fig.width=11, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE}
 gg_comparison = readRDS(file.path(datadir_tk, 'gg_oncor_mykiss_standartox_output.rds'))
@@
 
-```{r fig.width=10, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE}
+```{r fig.width=11, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE}
 gg_comparison = readRDS(file.path(datadir_tk, 'gg_oncor_mykiss_standartox_ppdb_comparison.rds'))
@@
   - Application Programming Interface (API)
-  - scriptable
-  - __reproducible__
+  - Scriptable
+  - __Reproducible__
 
@@
            duration = c(48, 96),
-           taxon = 'Oncorhynchus',
-           filtered = TRUE,
-           aggregated = TRUE)
+           taxon = 'Oncorhynchus mykiss')
 ```
@@
 
-<img src="http://3.bp.blogspot.com/-Wpp0vCJQDHo/U_ZMorXEynI/AAAAAAAAA7Y/Y-f-3brn99k/s1600/GBIF_logo_short_form.gif" width="120px" style="position:absolute; bottom:20px; left:70px;">
+<img src="http://3.bp.blogspot.com/-Wpp0vCJQDHo/U_ZMorXEynI/AAAAAAAAA7Y/Y-f-3brn99k/s1600/GBIF_logo_short_form.gif" width="100px" style="position:absolute; bottom:20px; left:70px;">
 

### 2019-08-30 R/exp_norman.R

@@ -5,21 +5,24 @@ source(file.path(src, 'gn_setup.R'))
 
 # save --------------------------------------------------------------------
 ## raw data
+# TODO uncomment section
 # export_tbl(schema = 'norman', table = 'data1', type = 'csv',
 #            user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
 #            dir = file.path(normandir, 'export'))
 export_tbl(schema = 'norman', table = 'data1_newest', type = 'csv',
            user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
            dir = file.path(normandir, 'export'))
+# TODO uncomment section
 # export_tbl(schema = 'norman', table = 'data2', type = 'csv',
 #            user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
 #            dir = file.path(normandir, 'export'))
 export_tbl(schema = 'norman', table = 'data2_newest', type = 'csv',
            user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
            dir = file.path(normandir, 'export'))
-export_tbl(schema = 'norman', table = 'data2_removed', type = 'csv',
-           user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
-           dir = file.path(normandir, 'export'))
+# TODO uncomment section
+# export_tbl(schema = 'norman', table = 'data2_removed', type = 'csv',
+#            user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
+#            dir = file.path(normandir, 'export'))
 export_tbl(schema = 'norman', table = 'variables', type = 'csv',
            user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
            dir = file.path(normandir, 'export'))

### 2019-08-30 R/exp_standartox.R

@@ -19,14 +19,26 @@ mkdirs(dir) # TODO put this into export_tbl()
 ## data
 # as .fst object
 export_tbl(schema = 'standartox', table = 'data2', type = 'fst', debug = FALSE,
+           compress = FALSE,
            user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
            dir = dir, file_name = paste0('standartox', v))
+# as .fst object (compressed)
+export_tbl(schema = 'standartox', table = 'data2', type = 'fst', debug = FALSE,
+           compress = TRUE,
+           user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
+           dir = dir, file_name = paste0('standartox_comp', v))
 # as .feather object
 export_tbl(schema = 'standartox', table = 'data2', type = 'feather', debug = FALSE,
            user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
            dir = dir, file_name = paste0('standartox', v))
 # as .rds object
 export_tbl(schema = 'standartox', table = 'data2', type = 'rds', debug = FALSE,
+           compress = TRUE,
+           user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
+           dir = dir, file_name = paste0('standartox', v))
+# as .rda object
+export_tbl(schema = 'standartox', table = 'data2', type = 'rda', debug = FALSE,
+           compress = TRUE,
            user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
            dir = dir, file_name = paste0('standartox', v))
 # as .csv object

### 2019-08-30 R/fun_export_db.R

@@ -1,6 +1,7 @@
 # function to export postgres tables
 
-export_tbl = function(schema, table, dir = NULL, file_name = NULL, type = c('csv', 'rds', 'fst', 'feather'), debug_mode = FALSE,
+export_tbl = function(schema, table, dir = NULL, file_name = NULL, type = c('csv', 'rds', 'rda', 'fst', 'feather'),
+                      compress = FALSE, debug_mode = FALSE,
                       user = NULL, host = NULL, port = NULL, password = NULL, dbname = NULL) {
   # file
   if (is.null(dir)) {
@@ -27,15 +28,18 @@ export_tbl = function(schema, table, dir = NULL, file_name = NULL, type = c('csv
   } else {
     fl = file.path(dir, file_name)
   }
-  type = match.arg(type, c('csv', 'rds', 'fst', 'feather'))
+  type = match.arg(type)
   if (type == 'csv') {
     fwrite(dat, paste0(fl, '.csv'))
   }
   if (type == 'rds') {
-    saveRDS(dat, paste0(fl, '.rds'))
+    saveRDS(dat, paste0(fl, '.rds'), compress = compress)
+  }
+  if (type == 'rda') {
+    save(dat, file = paste0(fl, '.rda'))
   }
   if (type == 'fst') {
-    write_fst(dat, paste0(fl, '.fst'))
+    write_fst(dat, paste0(fl, '.fst'), compress = compress)
   }
   if (type == 'feather') {
     write_feather(dat, paste0(fl, '.feather'))

### 2019-08-30 R/look_norman_acute_chronic_standard.R

@@ -12,7 +12,7 @@ chck_dupl(dat, 'result_id')
 
 # write -------------------------------------------------------------------
 write_tbl(dat, user = DBuser, host = DBhost, port = DBport, password = DBpassword,
-          dbname = DBetox, schema = 'ecotox', tbl = 'lookup_acute_chronic_standard',
+          dbname = DBetox, schema = 'lookup', tbl = 'lookup_acute_chronic_standard',
           key = 'result_id',
           comment = 'NORMAN use, acute-chronic and standard-test lookup table')
 

### 2019-08-30 R/look_norman_ecotox_group.R

@@ -1,0 +1,44 @@
+# lookup table for ecotox groups and habitat information
+# TODO not finished
+
+# setup -------------------------------------------------------------------
+source(file.path(src, 'gn_setup.R'))
+
+# query -------------------------------------------------------------------
+drv = dbDriver("PostgreSQL")
+con = dbConnect(drv, user = DBuser, dbname = DBetox, host = DBhost, port = DBport, password = DBpassword)
+
+ecotox_grp = data.table(dbGetQuery(
+  con,
+  "SELECT ecotox_group, COUNT(ecotox_group) n
+   FROM ecotox.species
+   GROUP BY ecotox_group
+   ORDER BY n DESC"
+))
+
+dbDisconnect(con)
+dbUnloadDriver(drv)
+
+# data --------------------------------------------------------------------
+ecotox_grp_look = fread(file.path(lookupdir, 'lookup_norman_ecotox_grp.csv'))
+ecotox_grp_look = ecotox_grp_look[ , .SD, .SDcols = c('ecotox_group', 'ecotox_group_conv') ]
+
+# merge -------------------------------------------------------------------
+ecotox_grp_fin = merge(ecotox_grp, ecotox_grp_look, by = 'ecotox_group')
+
+# write -------------------------------------------------------------------
+write_tbl(ecotox_grp_fin, user = DBuser, host = DBhost, port = DBport, password = DBpassword,
+          dbname = DBetox, schema = 'ecotox', tbl = 'ecotox_group_lookup',
+          key = 'ecotox_group',
+          comment = 'NORMAN ecotox_group lookup table')
+
+# log ---------------------------------------------------------------------
+log_msg('LOOK: NORMAN lookup tables script run.')
+
+# cleaning ----------------------------------------------------------------
+clean_workspace()
+
+
+
+
+

### 2019-08-30 R/qu_pc_prop_prep.R

@@ -7,6 +7,7 @@ source(file.path(src, 'gn_setup.R'))
 # Pubchem
 pc_prop_l = readRDS(file.path(cachedir, 'pc_prop_l.rds'))
 # CIR
+
 drv = dbDriver("PostgreSQL")
 con = dbConnect(drv, user = DBuser, dbname = DBetox, host = DBhost, port = DBport, password = DBpassword)
 

### 2019-08-30 R/run_build.R

@@ -42,9 +42,9 @@ if (build_db) {
   # build
   source(file.path(src, 'bd_epa_postgres.R'), max.deparse.length = mdl) # TODO rethink structure
   # DB roles
-  source(file.path(src, 'bd_postgres_roles.R'), max.deparse.length = mdl) # TODO  rethink structure
+  # TODO source(file.path(src, 'bd_postgres_roles.R'), max.deparse.length = mdl) # TODO  rethink structure
   # Permissions
-  source(file.path(src, 'bd_postgres_permissions.R', max_deparse.length = mdl)) # TODO rethink structure
+  # TODO source(file.path(src, 'bd_postgres_permissions.R', max_deparse.length = mdl)) # TODO rethink structure
   # functions
   source(file.path(src, 'bd_sql_functions.R'), max.deparse.length = mdl)
   # errata
@@ -113,11 +113,11 @@ if (export) {
 
 # backup ------------------------------------------------------------------
 if (general) {
-  source(file.path(src, 'gn_backup.R'))
+  source(file.path(src, 'gn_backup.R'), max.deparse.length = mdl)
 }
 
 # end ---------------------------------------------------------------------
-source(file.path(src, 'gn_end.R'))
+source(file.path(src, 'gn_end.R'), max.deparse.length = mdl)
 
 # # (0b) console log 2 --------------------------------------------------------
 if (sink_console) {

### 2019-09-02 Rmd/summary_db_cols_norman.Rmd
added cbind_fill()

@@
 # table = 'data2_newest'
-# cols = c('nor88test', 'nor88')
+# cols = c('nor85', 'nor88')
 
@@
 ```{r write, echo=FALSE, message=FALSE}
+# write summary to multiple csv
 unlink(dir, recursive = TRUE)
@@
 ```
+
+```{r write-single-file, message=FALSE}
+# write summary to single file
+dat = cbind_fill(dflist = l, limit = NULL)
+names(dat)[ seq(1, length(names(dat)), 2) + 1 ] = paste0(names(dat)[ seq(1, length(names(dat)), 2) ], '_n')
+
+fwrite(dat, file.path(params$output_dir, paste0(params$table, '_summary.csv')))
+```

### 2019-09-02 R/fun_cbind_fill.R

@@ -1,0 +1,17 @@
+# function to bind data.frame s in a list to a single data.frame
+# https://stackoverflow.com/questions/7962267/cbind-a-dataframe-with-an-empty-dataframe-cbind-fill
+
+cbind_fill <- function(..., dflist = NULL, limit = NULL){
+  nm <- c(list(...), dflist)
+  nm <- lapply(nm, as.matrix)
+  n <- max(sapply(nm, nrow)) 
+  out <- do.call(cbind, lapply(nm, function (x) 
+    rbind(x, matrix(, n-nrow(x), ncol(x)))))
+  out <- data.frame(out)
+  
+  if (!is.null(limit)) {
+    out <- out[1:limit, ]
+  }
+  
+  return(out)
+}

### 2019-09-02 R/gn_setup.R

@@ -131,6 +131,7 @@ source(file.path(src, 'fun_write_db.R'))
 source(file.path(src, 'fun_read_db.R'))
 source(file.path(src, 'fun_product_na.R'))
 source(file.path(src, 'fun_extr_vec.R'))
+source(file.path(src, 'fun_cbind_fill.R'))
 source(file.path(src, 'fun_log.R'))
 source(file.path(src, 'fun_scrape_phantomjs.R'))
 source(file.path(src, 'fun_worms_query.R'))

### 2019-09-13 Rmd/summary_db_cols_norman.Rmd
changes

@@
 # debuging
+# dir = '/tmp'
 # schema = 'norman'
-# table = 'data1'
-# cols = c('nor3', 'nor4')
+# table = 'data2_newest'
+# cols = c('nor1', 'nor3', 'nor4', 'nor5', 'nor6')
 ```
@@
 ```{r echo=FALSE, message=FALSE}
-# schema = 'norman'
-# table = 'data2_newest'
-# cols = c('nor85', 'nor88')
-
 ## table
@@
 
-  dat = l[[col]]
+  dat = copy(l[[col]])
   dat[ , var20 := substr(get(col), 1, 20) ]
@@
   message('Writing: ', fl)
-  fwrite(dat, fl)
+  fwrite(dat, fl, sep = ';')
 }
@@
 # write summary to single file
-dat = cbind_fill(dflist = l, limit = NULL)
-names(dat)[ seq(1, length(names(dat)), 2) + 1 ] = paste0(names(dat)[ seq(1, length(names(dat)), 2) ], '_n')
+dat = cbind_fill(dflist = l, limit = 5000)
+# names(dat)[ seq(1, length(names(dat)), 2) + 1 ] = paste0(names(dat)[ seq(1, length(names(dat)), 2) ], '_n')
 

### 2019-09-13 R/bd_norman.R

@@ -13,8 +13,8 @@ fl = file.path(sql, 'norman_data1_newest.sql')
 q1_newest = readChar(fl, file.info(fl)$size)
 fl = file.path(sql, 'norman_data2_newest.sql')
 q2_newest = readChar(fl, file.info(fl)$size)
-fl = file.path(sql, 'norman_data2_removed.sql')
-q2_removed = readChar(fl, file.info(fl)$size)
+# fl = file.path(sql, 'norman_data2_removed.sql')
+# q2_removed = readChar(fl, file.info(fl)$size)
 
 # query -------------------------------------------------------------------
 drv = dbDriver("PostgreSQL")
@@ -30,7 +30,7 @@ dbSendQuery(con, q1) # cleaned data
 dbSendQuery(con, q2) # converted data
 dbSendQuery(con, q1_newest) # cleaned newest data (acc to published date)
 dbSendQuery(con, q2_newest) # cleaned newest data (acc to published date)
-dbSendQuery(con, q2_removed) # removed tests
+# dbSendQuery(con, q2_removed) # removed tests
 
 dbDisconnect(con)
 dbUnloadDriver(drv)
@@ -39,32 +39,23 @@ dbUnloadDriver(drv)
 output_dir = file.path(normandir, 'summary')
 rmarkdown::render(file.path(srcrmd, 'summary_db_cols_norman.Rmd'),
                   output_dir = output_dir,
-                  output_file = 'summary_NORMAN_raw',
+                  output_file = 'data1',
                   params = list(
-                    title = 'NORMAN raw data summary',
+                    title = 'data1: NORMAN raw data summary',
                     schema = 'norman',
                     table = 'data1',
                     # cols = c('nor88test', 'nor88')
-                    output_dir = output_dir #! strange that output_dir has to be specified 2 times
+                    output_dir = output_dir
                   ))
 rmarkdown::render(file.path(srcrmd, 'summary_db_cols_norman.Rmd'),
                   output_dir = output_dir, #! strange that output_dir has to be specified 2 times
-                  output_file = 'summary_NORMAN_cleaned',
+                  output_file = 'data2',
                   params = list(
-                    title = 'NORMAN cleaned data summary',
+                    title = 'data2: NORMAN cleaned data summary',
                     schema = 'norman',
                     table = 'data2',
                     #cols = c('nor84', 'nor84b'),
[…]

### 2019-09-13 R/exp_norman.R

@@ -5,24 +5,18 @@ source(file.path(src, 'gn_setup.R'))
 
 # save --------------------------------------------------------------------
 ## raw data
-# TODO uncomment section
-# export_tbl(schema = 'norman', table = 'data1', type = 'csv',
+# export_tbl(schema = 'norman', table = 'data1', type = c('xlsx', 'rds'),
 #            user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
 #            dir = file.path(normandir, 'export'))
-export_tbl(schema = 'norman', table = 'data1_newest', type = 'csv',
+export_tbl(schema = 'norman', table = 'data1_newest', type = c('xlsx', 'rds'),
            user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
            dir = file.path(normandir, 'export'))
-# TODO uncomment section
-# export_tbl(schema = 'norman', table = 'data2', type = 'csv',
+# export_tbl(schema = 'norman', table = 'data2', type = c('xlsx', 'rds'),
 #            user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
 #            dir = file.path(normandir, 'export'))
-export_tbl(schema = 'norman', table = 'data2_newest', type = 'csv',
+export_tbl(schema = 'norman', table = 'data2_newest', type = c('xlsx', 'rds'),
            user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
            dir = file.path(normandir, 'export'))
-# TODO uncomment section
-# export_tbl(schema = 'norman', table = 'data2_removed', type = 'csv',
-#            user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
-#            dir = file.path(normandir, 'export'))
 export_tbl(schema = 'norman', table = 'variables', type = 'csv',
            user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
            dir = file.path(normandir, 'export'))

### 2019-09-13 R/exp_standartox_shiny_stats.R

@@ -14,23 +14,27 @@ cols = c('effect', 'endpoint', 'conc1_type', 'test_location', 'obs_duration_mean
          grep('reg_', header, value = TRUE),
          grep('ccl_', header, value = TRUE))
 # loop
+con = DBI::dbConnect(RPostgreSQL::PostgreSQL(),
+                     dbname = DBetox,
+                     host = DBhost,
+                     port = DBport,
+                     user = DBuser,
+                     password = DBpassword)
+
 l = list()
 for (i in seq_along(cols)) {
   
   col = cols[i]
   message('Fetching: ', col)
-  con = DBI::dbConnect(RPostgreSQL::PostgreSQL(),
-                       dbname = DBetox,
-                       host = DBhost,
-                       port = DBport,
-                       user = DBuser,
-                       password = DBpassword)
+  
   dat = summary_db_perc(con, 'standartox', 'data2', col)
   setDT(dat)
   
   l[[i]] = dat
   names(l)[i] = col
 }
+
+DBI::dbDisconnect(con)
 
 # preparation -------------------------------------------------------------
 stat = rbindlist(l, idcol = TRUE)
@@ -49,8 +53,10 @@ stat_l = split(stat, stat$variable)
 stat_l$obs_duration_mean2 = stat_l$obs_duration_mean2[ , range(na.omit(as.numeric(value))) ]
 
 # write -------------------------------------------------------------------
-saveRDS(stat_l, file.path(exportdir, v, paste0('standartox', v, '_shiny_stats.rds')))
-save(stat_l, file = file.path(exportdir, v, paste0('standartox', v, '_shiny_stats.rda'))) # for standartox::
+# TODO can't be read from R 3.4 (changed with 3.5)
+saveRDS(stat_l, file.path(exportdir, v, paste0('standartox', v, '_shiny_stats.rds')),
+        version = 2) # TODO change the server R version
+# TODO if API doesn t work save(stat_l, file = file.path(exportdir, v, paste0('standartox', v, '_shiny_stats.rda'))) # for standartox::
 
 # log ---------------------------------------------------------------------
 log_msg('Export: application summary stats exported')

### 2019-09-13 R/fun_cbind_fill.R

@@ -7,7 +7,7 @@ cbind_fill <- function(..., dflist = NULL, limit = NULL){
   n <- max(sapply(nm, nrow)) 
   out <- do.call(cbind, lapply(nm, function (x) 
     rbind(x, matrix(, n-nrow(x), ncol(x)))))
-  out <- data.frame(out)
+  out <- data.table(out)
   
   if (!is.null(limit)) {
     out <- out[1:limit, ]

### 2019-09-13 R/fun_export_db.R

@@ -1,6 +1,6 @@
 # function to export postgres tables
 
-export_tbl = function(schema, table, dir = NULL, file_name = NULL, type = c('csv', 'rds', 'rda', 'fst', 'feather'),
+export_tbl = function(schema, table, dir = NULL, file_name = NULL, type = c('csv', 'xlsx', 'rds', 'rda', 'fst', 'feather'),
                       compress = FALSE, debug_mode = FALSE,
                       user = NULL, host = NULL, port = NULL, password = NULL, dbname = NULL) {
   # file
@@ -16,7 +16,6 @@ export_tbl = function(schema, table, dir = NULL, file_name = NULL, type = c('csv
   drv = dbDriver("PostgreSQL")
   con = dbConnect(drv, user = DBuser, dbname = DBetox, host = DBhost, port = DBport, password = DBpassword)
   
-  message('Exporting table: ', schema, '.', table, ' (.', type, ') to ', dir)
   dat = dbGetQuery(con, q)
   setDT(dat)
   
@@ -28,20 +27,29 @@ export_tbl = function(schema, table, dir = NULL, file_name = NULL, type = c('csv
   } else {
     fl = file.path(dir, file_name)
   }
-  type = match.arg(type)
-  if (type == 'csv') {
+  type = match.arg(type, several.ok = TRUE)
+  if (any(type %in% 'csv')) {
+    message('Exporting table: ', schema, '.', table, ' (.csv) to ', dir)
     fwrite(dat, paste0(fl, '.csv'))
   }
-  if (type == 'rds') {
+  if (any(type %in% 'xlsx')) {
+    message('Exporting table: ', schema, '.', table, ' (.xlsx) to ', dir)
+    write.xlsx(dat, paste0(fl, '.xlsx'))
+  }
+  if (any(type %in% 'rds')) {
+    message('Exporting table: ', schema, '.', table, ' (.rds) to ', dir)
     saveRDS(dat, paste0(fl, '.rds'), compress = compress)
   }
-  if (type == 'rda') {
+  if (any(type %in% 'rda')) {
+    message('Exporting table: ', schema, '.', table, ' (.rda) to ', dir)
     save(dat, file = paste0(fl, '.rda'))
   }
-  if (type == 'fst') {
+  if (any(type %in% 'fst')) {
+    message('Exporting table: ', schema, '.', table, ' (.fst) to ', dir)
     write_fst(dat, paste0(fl, '.fst'), compress = compress)
   }
[…]

### 2019-09-13 R/look_norman_acute_chronic_standard.R

@@ -7,6 +7,53 @@ source(file.path(src, 'gn_setup.R'))
 # data --------------------------------------------------------------------
 dat = fread(file.path(normandir, 'data', 'lookup_acute_chronic_standard.csv'), na.strings = '')
 
+# additions ---------------------------------------------------------------
+# (1) make 'biomass', 'chlorophyll a concentration', 'abundance' to chronic
+time = Sys.time()
+drv = dbDriver("PostgreSQL")
+con = dbConnect(drv, user = DBuser, dbname = DBetox, host = DBhost, port = DBport, password = DBpassword)
+
+q = "SELECT results.result_id,
+            clean(results.obs_duration_mean) obs_duration_mean,
+            clean(results.obs_duration_mean)::numeric * duration_unit_lookup.multiplier obs_duration_mean2,
+            clean(results.obs_duration_unit) obs_duration_unit,
+            duration_unit_lookup.unit_conv obs_duration_unit2,
+            clean(results.endpoint) endpoint,
+            clean(results.effect) effect,
+            clean(results.measurement) measurement,
+            species.latin_name,
+            species.family,
+            ac_cr.*
+     FROM ecotox.tests
+     LEFT JOIN ecotox.results ON tests.test_id = results.test_id
+     LEFT JOIN ecotox.refs ON tests.reference_number = refs.reference_number
+     LEFT JOIN ecotox.chemicals ON tests.test_cas = chemicals.cas_number
+     LEFT JOIN ecotox.species ON tests.species_number = species.species_number
+     LEFT JOIN ecotox.duration_unit_lookup ON results.obs_duration_unit = duration_unit_lookup.obs_duration_unit
+     LEFT JOIN lookup.lookup_acute_chronic_standard ac_cr ON results.result_id = ac_cr.result_id"
+
+add = dbGetQuery(con, q)
+setDT(add)
+
+dbDisconnect(con)
+dbUnloadDriver(drv)
+Sys.time() - time
+
+fam_ok = add[ norman_use == 'yes',
+              sort(unique(family)) ]
+idx_chronic = add[ family %in% fam_ok &
+                     measurement %in% c('ABND', 'CHLA', 'BMAS'), result_id ]
+idx_acute = add[ family %in% fam_ok &
+                   measurement == 'PGRT', result_id ]
+dat[ result_id %in% idx_chronic, `:=`
+     (norman_use = 'yes',
+       acute_chronic = 'chronic',
[…]

### 2019-09-13 R/qu_chebi_prep.R

@@ -44,7 +44,7 @@ chebi_fin = chebi_fin[ !duplicated(cas) & !is.na(cas) ]
 # envi = c('biocide', 'fungicide', 'herbicide', 'insecticide', 'pesticide', 'environmental.contaminent', 'agrochemical')
 # cols = grep(paste0(envi, collapse = '|'), names(ont_par2), value = TRUE)
 # chebi_envi = ont_par2[ , .SD, .SDcols = c('chebiid', cols) ]
-chebi_envi = ont_par2
+chebi_envi = ont_par
 
 #! necessary 'cause it can be that a chemical is an azole fungicide but not classifed as a fungicide
 fung = grep('fungicide', names(chebi_envi), value = TRUE)
@@ -90,8 +90,8 @@ clean_names(chebi_envi)
 
 ## drugs
 drug = c('drug')
-cols = grep(paste0(drug, collapse = '|'), names(ont_par2), value = TRUE)
-chebi_drug = ont_par2[ , .SD, .SDcols = c('chebiid', cols) ]
+cols = grep(paste0(drug, collapse = '|'), names(ont_par), value = TRUE)
+chebi_drug = ont_par[ , .SD, .SDcols = c('chebiid', cols) ]
 chebi_drug[ , drug := do.call(pmin, c(.SD, na.rm = TRUE)), .SDcols = cols ]
 chebi_drug[chebi_fin, cas := i.cas, on = 'chebiid' ] # merge cas
 chebi_drug = chebi_drug[ , unique(.SD, by = 'cas') ] # take the first in case of duplicates

### 2019-09-13 R/run_build.R

@@ -116,6 +116,10 @@ if (general) {
   source(file.path(src, 'gn_backup.R'), max.deparse.length = mdl)
 }
 
+# sync --------------------------------------------------------------------
+# TODO
+# create rsync between server and my local machine
+
 # end ---------------------------------------------------------------------
 source(file.path(src, 'gn_end.R'), max.deparse.length = mdl)
 

### 2019-09-26 README.Rmd
changes

@@
 ---
-title: "Etox-Base"
-author: "Andras Scharmüller"
-date: "`r format(Sys.time(), '%d %B, %Y')`"
-output: html_document
+title: "Standartox"
+output:
+  github_document:
+  keep_md: true
 editor_options: 
@@
 
-```{r setup, include=FALSE}
+Standartox is a data base tool standartizing the retrieval and aggregation of ecotoxicological test data.
+It is based on the [EPA ECOTOX data base](https://cfpub.epa.gov/ecotox/) as well as on data from several other chemical data bases and allows users to filter and aggregate ecotoxicological test data in an easy way.
+It can either be accessed via a [web-app](http://standartox.uni-landau.de/) or via the R-package [standartox](https://github.com/andschar/standartox).
+Ecotoxicological test data are used in environmental risk assessment to calculate effect measures such as [TU - Toxic Units](https://en.wikipedia.org/wiki/Toxic_unit) or [SSD - Species Sensitivity Distributions](https://edild.github.io/ssd/) to asses toxicity of chemicals.
 
-```
+The project lives in these three repositories:
 
-## Etox Base
-
-Etox Base is a tool that aggregates ecotoxicological test data (EC50 values, NOEC etc.) from the [US EPA ECOTOTX data base](https://cfpub.epa.gov/ecotox/).
-Each new version of the EPA ECOTOX database which is released on a quarterly basis is downloaded and rebuilt locally.
-Besides the ecotoxicological test data therein, other databases on chemicals and organism parameters are then queried to extend and fill data gaps of the EPA ECOTOX database in order to allow for more adequat filter steps according to ecotoxicological needs.
-Chemical parameters on for example the classification of chemicals, water solubility of compounds amongst others are obtained through databases such as [Alan Wood's Compendium of Pesticide Common Names](http://www.alanwood.net/pesticides/index.html), the [PubChem](https://pubchem.ncbi.nlm.nih.gov/) database, the [PHYSPROP Database](https://www.srcinc.com/what-we-do/environmental/scientific-databases.html) and the [chemspider Database](http://www.chemspider.com/), all of which have publically available programming interfaces (APIs).
-Likewise databases on organisms such as the [World Refister of Marine Species (WORMS)](http://marinespecies.org/) and the [Global Biodiversity Information Facility (GBIF)](https://www.gbif.org/) are queried to obtain information on habitat preferences and organisms occurrence patterns.
-CAS-Numbers are used to retrieve the additional chemical inforamtion whereas taxa names are used to query organism parameters.
-These queries largely rely on R-packages such as [webchem](https://github.com/ropensci/webchem), [taxize](https://github.com/ropensci/taxize/) and [rgbif](https://github.com/ropensci/rgbif) besides some self-written functions that can be accesed in the two project repositories: [etox-base](https://github.com/andreasLD/etox-base) and [etox-base-shiny](https://github.com/andreasLD/etox-base).
-
-The compiled data is than accessed through the Etox-Base application which filters it according to the users inputs.
-In the end the tool aggregates the many filtered test data to obtain a single value for a given set of test parameters.
-
-An article on the project can be accessed [here]().
+- [standartox-build](https://github.com/andschar/standartox-build) - Compiles the tool
+- [standartox-app](https://github.com/andschar/standartox-build) - Serves the website and API
+- [standartox](https://github.com/andschar/standartox) - R-package

### 2019-09-26 R/fun_aggregate.R

@@ -1,0 +1,23 @@
+# function to aggregate input from fun_filter()
+
+fun_aggregate = function(dt = NULL, 
+                         comp = NULL,
+                         vl = 'concentration',
+                         agg = c('min', 'gmn', 'med', 'max'),
+                         info = NULL) {
+  agg = match.arg(agg, several.ok = TRUE)
+  out = dt[ ,
+            j = .(min = min(get(vl), na.rm = TRUE),
+                  med = median(get(vl), na.rm = TRUE),
+                  gmn = gm_mean(get(vl)),
+                  max = max(get(vl), na.rm = TRUE),
+                  n = .N,
+                  taxa = paste0(unique(tax_taxon), collapse = '-')),
+            by = .(casnr, cname) ]
+  sdcols = unique(c(comp, agg, info))
+  out = out[ , .SD, .SDcols = sdcols ]
+  
+  return(out)
+}
+
+

### 2019-09-26 R/fun_filter.R

@@ -2,40 +2,32 @@
 
 fun_filter = function(dt,
                       cas_ = NULL,
-                      conc1_type_ = NULL,
+                      concentration_type_ = NULL,
                       chemical_class_ = NULL,
                       taxa_ = NULL,
                       habitat_ = NULL,
                       region_ = NULL,
                       duration_ = NULL,
                       publ_year_ = NULL, # TODO
-                      acch_ = NULL, # TODO
-                      exposure_ = NULL, # TODO
+                      # acch_ = NULL, # TODO
+                      # exposure_ = NULL, # TODO
                       effect_ = NULL,
-                      endpoint_ = NULL,
-                      chck_solub_ = FALSE # TODO
+                      endpoint_ = NULL
+                      # chck_solub_ = FALSE # TODO
                       ) {
-  
-  # debug me!
-  # source('R/setup.R')
-  # dt = read_fst('/home/scharmueller/Projects/etox-base-shiny/data/20190314/standartox20190314.fst')
-  # conc1_type = NULL; chemical_class = 'ccl_herbicide'; tax = NULL; habitat = 'hab_freshwater'; region = 'reg_europe'; duration_ = c(24,48); publ_year = NULL; acch = NULL; exposure = NULL; effect = NULL; endpoint = NULL; chck_solub = NULL; cas = dt$casnr[1:10];
-  
   # checks -----------------------------------------------------------------
   if (!is.data.frame(dt)) {
     stop('Input object is not a data.frame!')
   }
   data.table::setDT(dt)
-  
   # CAS ---------------------------------------------------------------------
   if (!is.null(cas_)) {
-    casnr_todo = casconv(cas_, direction = 'tocasnr')
+    casnr_todo = gsub('-', '', cas_)
     dt = dt[ casnr %in% casnr_todo ]
   }
-  
   # filters -----------------------------------------------------------------
-  if (!is.null(conc1_type_)) {
-    dt = dt[ conc1_type %in% conc1_type_ ]
+  if (!is.null(concentration_type_)) {
+    dt = dt[ concentration_type %in% concentration_type_ ]
   }
   if (!is.null(effect_)) {
     dt = dt[ effect %in% effect_ ]
@@ -57,13 +49,13 @@ fun_filter = function(dt,
[…]

### 2019-09-26 R/functions/fun_aggregation.R

@@ -1,38 +1,0 @@
-# function to aggregate input from fun_filter()
-
-# setup -------------------------------------------------------------------
-fun_aggregate = function(dt, 
-                         comp = NULL,
-                         vl = 'conc1_mean2',
-                         agg = c('min', 'md', 'gm', 'max'),
-                         info = NULL,
-                         rm_outl = FALSE) {
-  
-  agg = match.arg(agg, several.ok = TRUE)
-  # outliers ---------------------------------------------------------------
-  dt[ , outl := rm_outliers(get(vl), lim = 1.5, na.rm = TRUE) ]
-  if (rm_outl) {
-    dt = dt[ !is.na(outl) ]
-  }
-  # aggregation -------------------------------------------------------------
-  out = dt[ ,
-            j = .(min = min(get(vl), na.rm = TRUE),
-                  md = median(get(vl), na.rm = TRUE),
-                  gm = gm_mean(get(vl)),
-                  max = max(get(vl), na.rm = TRUE),
-                  n = .N,
-                  info = paste0(unlist(lapply(strsplit(tax_taxon, '\\s'), paste0, collapse = '_')),
-                                '_', obs_duration_mean2, 'h_', '(', sort(get(vl)), 'ug/L)',
-                                collapse = ' - '),
-                  vls = paste0(as.character(sort(get(vl))), collapse = '-'),
-                  taxa = paste0(unique(tax_taxon), collapse = '-')
-                  ),
-            by = .(casnr, cname) ]
-  # return ------------------------------------------------------------------
-  sdcols = c('casnr', comp, agg, info)
-  out = out[ , .SD, .SDcols = sdcols ]
-  
-  return(out)
-}
-
-

### 2019-09-26 R/functions/fun_ec50filter_meta_plots.R

@@ -1,8 +1,0 @@
-# counter_l = readRDS(file.path(cache, 'dt_counter.rds'))
-# 
-# counter = rbindlist(counter_l)
-# #setorder(counter, N)
-# 
-# gg_counter = ggplot(counter, aes(y = N, x = reorder(Variable, N))) +
-#   geom_bar(stat = 'identity') +
-#   coord_flip()

### 2019-09-26 R/functions/fun_outliers.R

@@ -1,12 +1,0 @@
-# function to mark outliers as NA
-# taken from: https://stackoverflow.com/questions/4787332/how-to-remove-outliers-from-a-dataset/4788102#4788102
-
-rm_outliers <- function(x, lim = 1.5, na.rm = TRUE, ...) {
-  qnt <- quantile(x, probs=c(.25, .75), na.rm = na.rm, ...)
-  H <- lim * IQR(x, na.rm = na.rm)
-  y <- x
-  y[x < (qnt[1] - H)] <- NA
-  y[x > (qnt[2] + H)] <- NA
-  y
-}
-

### 2019-09-26 R/functions/fun_output_stats.R

@@ -1,23 +1,0 @@
-# create counts for individual factors in the output data.table
-
-# for long variables (i.e. conc_type = A,F,...)
-out_stats_lng = function(dt, var) {
-  dtmp = copy(dt)
-  var = as.character(substitute(var))
-  out = dtmp[ , .N, c(var) ][ order(-N) ]
-  out[ , nam := paste0(unlist(out[ , 1]), ' (', out$N, ')')]
-  
-  return(out)
-}
-
-# for wide variables (i.e. is_fresh = 1,NA; is_terre = 1,NA)
-out_stats_wid = function(dt, vars) {
-  dtmp = copy(dt)
-  dtmp = dtmp[ , .SD, .SDcols = vars ]
-  m_dtmp = melt(dtmp)
-  out = m_dtmp[ value == 1, .N, variable ][ order(-N) ]
-  
-  return(out)
-}
-
-

### 2019-09-26 R/functions/fun_plotly.R

@@ -1,138 +1,0 @@
-# function to create plot for application
-
-plotly_prep = function(agg,
-                       fil,
-                       cutoff = 25) {
-  ## prepartaion
-  setDT(agg)
-  agg[ , casnr := as.character(casnr) ]
-  setDT(fil)
-  fil[ , casnr := as.character(casnr) ]
-  ## aggregated data
-  cols = grep('gm|md|mn|min|max',
-              names(agg),
-              ignore.case = TRUE,
-              value = TRUE)
-  agg_m = melt(agg, id.vars = 'casnr', measure.vars = cols)
-  agg_m[fil, `:=` (
-    cname = i.cname
-    # TODO outl = i.outl,
-    # TODO ref_num = i.ref_num
-  ), on = 'casnr']
-  # split (in case of multiple variables)
-  l_agg = split(agg_m, agg_m$variable)
-  # order + cutoff
-  l_agg = lapply(l_agg, setorder, value)
-  l_agg = lapply(l_agg, head, cutoff)
-  ## filtered data
-  col_fil = c('casnr', 'cname', 'conc1_mean2')
-  fil2 = fil[ , .SD, .SDcols = col_fil ]
-  fil2 = fil2[ casnr %in% agg$casnr ]
-  fil2[ ,
-        outl := rm_outliers(conc1_mean2, lim = 1.5, na.rm = TRUE),
-        by = casnr ]
-  fil2[ , outl := ifelse(is.na(outl), TRUE, FALSE)]
-  
-  ## return
-  out_l = list(l_agg = l_agg,
-               fil2 = fil2)
-  
-  return(out_l)
-}
-  
-plotly_plot = function(agg,
-                       fil,
-                       yaxis = 'casnr',
-                       xaxis = 'limout') {
-  ## preparation
-  fil2 = fil[ casnr %in% agg$casnr ]
-  ## axis
-  yform = list(
-    categoryorder = "array",
-    categoryarray = agg[ order(value, decreasing = TRUE) ][ , get(yaxis) ],
-    type = 'category',
-    nticks = 1000
-  ) #! dummy to plot all y entries
-  if (xaxis == 'log') {
-    xform = list(title = 'Concentration (ug/L)',
-                 zeroline = FALSE,
-                 type = 'log')
-  }
-  if (xaxis == 'limout') {
-    xform = list(title = 'Concentration (ug/L)',
-                 zeroline = FALSE,
-                 range = range(agg$value))
-  }
-  ## plot
-  out = plot_ly() %>% 
[…]

### 2019-09-26 R/functions/fun_ssd.R

@@ -1,37 +1,0 @@
-# SSD function script 
-# https://github.com/bcgov/ssdtools
-
-### debuging
-# dt = data.table(ssdtools::boron_data) # debug me!
-# names(dt)
-# setnames(dt, paste0(names(dt), 1:5))
-### end
-# 
-# dt = d2
-# left = 'value_fin'
-
-fun_ssd = function(dt, left = NULL) {
-  
-  dt_dists = ssd_fit_dists(dt, left = left)
-  # dt_hc5 = ssd_hc(dt_dists, nboot = 10000)
-  dt_predict = predict(dt_dists)
-  
-  ssd_pl = ssd_plot(dt, dt_predict, shape = 'Group', label = 'Species',
-                    xlab = "Concentration (mg/L)", ribbon = TRUE) + 
-    expand_limits(x = 3000)
-  
-  # ssd_gof(dt_dists)
-  # autoplot(dt_dists)
-  # out_l = list()
-  # out_l[['plot']] == ssd_pl
-  # out_l[['hc5']] == dt_hc5
-  
-  return(ssd_pl)
-}
-
-
-
-
-
-
-

### 2019-09-26 api.R

@@ -1,69 +1,137 @@
 # API script
 
 # setup -------------------------------------------------------------------
-source('setup.R')
+source('setup.R', local = FALSE)
 
 # description -------------------------------------------------------------
 #* @apiTitle Standartox API
 #* @apiDescription Retieve filtered data
 
-# data --------------------------------------------------------------------
-# this path would have to be adapted if you would deploy this
-# dat = fst::read_fst('/home/scharmueller/Projects/etox-base-shiny/data/20180314/standartox20180314.fst')
-# stat = readRDS('/home/scharmueller/Projects/etox-base-shiny/data/20180314/standartox20180314_shiny_stats.rds')
-# data.table::setDT(dat)
-
-# logger ------------------------------------------------------------------
+# filter: logger ----------------------------------------------------------
 #* Log system time, request method and HTTP user agent of the incoming request
 #* @filter logger
 function(req) {
-  cat("System time:", as.character(Sys.time()), "\n",
-      "Request method:", req$REQUEST_METHOD, req$PATH_INFO, "\n",
-      "HTTP user agent:", req$HTTP_USER_AGENT, "@", req$REMOTE_ADDR, "\n")
+  log_df = data.frame(system_time = as.character(Sys.time()),
+                      request_method = req$REQUEST_METHOD,
+                      path_info = req$PATH_INFO,
+                      http_user_agent = paste0(req$HTTP_USER_AGENT, "@", req$REMOTE_ADDR))
+  fwrite(log_df, 'request.log', append = TRUE)
+  cat(unlist(log_df))
   
   plumber::forward()
 }
 
-# version path ------------------------------------------------------------
-#* version path
-#* @filter version_epa
+# filter: debuging --------------------------------------------------------
+# save req objects to /tmp for debugging
+#* @filter debugger
 function(req) {
-  dirs = list.dirs('data', recursive = FALSE, full.names = FALSE)
-  # req_debug <<- req # debuging
-  # version_epa1 <<- req$args$version_epa # debuging
-  # version_epa = req$args$version_epa
-  # version_epa <<- version_epa # debuging
-  v = req$args$version_epa
+  req_envir <<- req
   
-  if (is.null(v)) {
-    req$args$version_epa2 = max(dirs)
-  } else {
-    version_epa2 = as.integer(gsub('\\W', '', v)) # sanitizing
[…]

### 2019-09-26 app.R

@@ -48,17 +48,17 @@ sidebar = dashboardSidebar(
         )
       ),
       prettyCheckboxGroup(
-        inputId = 'conc1_type',
+        inputId = 'concentration_type',
         label = 'Concentration type',
-        choiceValues = stat_l$conc1_type$value,
-        choiceNames = stat_l$conc1_type$name_perc,
-        selected = grep('active.ingredien', stat_l$conc1_type$value, ignore.case = TRUE, value = TRUE)[1]
+        choiceValues = catalog_l$concentration_type$variable,
+        choiceNames = catalog_l$concentration_type$name_perc,
+        selected = grep('active.ingredien', catalog_l$concentration_type$variable, ignore.case = TRUE, value = TRUE)[1]
       ),
       prettyCheckboxGroup(
         inputId = 'chemical_class',
         label = 'Chemical class',
-        choiceValues = stat_l$chemical_class$value,
-        choiceNames = stat_l$chemical_class$name_perc,
+        choiceValues = catalog_l$chemical_class$variable,
+        choiceNames = catalog_l$chemical_class$name_perc,
         selected = NULL
       )
     ),
@@ -67,23 +67,23 @@ sidebar = dashboardSidebar(
       tabName = 'taxon',
       selectizeInput(inputId = 'tax',
                      label = 'Taxa',
-                     choices = taxa_all_list,
+                     choices = catalog_l$taxa$variable,
                      selected = NULL,
                      multiple = TRUE,
                      options = list(create = FALSE)),
       prettyCheckboxGroup(
         inputId = 'habitat',
         label = 'Organism hatbitat',
-        choiceValues = stat_l$habitat$value,
-        choiceNames = stat_l$habitat$name_perc,
-        selected = grep('fresh', stat_l$habitat$value, ignore.case = TRUE, value = TRUE)[1]
+        choiceValues = catalog_l$habitat$variable,
+        choiceNames = catalog_l$habitat$name_perc,
+        selected = grep('fresh', catalog_l$habitat$variable, ignore.case = TRUE, value = TRUE)[1]
       ),
       prettyCheckboxGroup(
         inputId = 'region',
-        label = 'Continent',
-        choiceValues = stat_l$region$value,
-        choiceNames = stat_l$region$name_perc,
-        selected = grep('europe', stat_l$region$value, ignore.case = TRUE, value = TRUE)[1]
+        label = 'Region',
+        choiceValues = catalog_l$region$variable,
[…]

### 2019-09-26 data.R

@@ -2,24 +2,41 @@
 
 # data --------------------------------------------------------------------
 time = Sys.time()
-dat = read_fst(file.path(datadir, epa_versions_newest, paste0('standartox', epa_versions_newest, '.fst')))
+dat = read_fst(file.path(datadir,
+                         epa_versions_newest,
+                         paste0('standartox', epa_versions_newest, '.fst')))
 setDT(dat)
 Sys.time() - time
-# list all taxa for auto completion
-taxa_all_list = sort(unique(unlist(dat[ , .SD, .SDcols = grep('tax_', names(dat)) ])))
-# meta data
-stat_l = readRDS(file.path(datadir, epa_versions_newest, paste0('standartox', epa_versions_newest, '_shiny_stats.rds')))
+# IDEA rewrite filter function for data.frame to be able to use disk fst file directly
+# dat_fst = fst(file.path(datadir,
+#                         epa_versions_newest,
+#                         paste0('standartox', epa_versions_newest, '.fst')))
+# test = dat_fst[1:10, ]
+# time = Sys.time()
+# test = dat_fst[ dat_fst$endpoint == 'XX50', ]
+# Sys.time() - time
+### END
 
-# time = Sys.time()
-# dat = read_feather(file.path(export, 'standartox.feather'))
-# Sys.time() - time
-# 
-# time = Sys.time()
-# dat = readRDS(file.path(export, 'standartox.rds'))
-# Sys.time() - time
-# 
-# time = Sys.time()
-# dat = fread(file.path(export, 'standartox.csv'))
-# Sys.time() - time
-# 
+# catalog -----------------------------------------------------------------
+catalog_l = readRDS(file.path(datadir,
+                              epa_versions_newest,
+                              paste0('standartox', epa_versions_newest, '_catalog.rds')))
+# add name percetnage column
+catalog_l = lapply(catalog_l,
+                   function(x) {
+                     if (is.data.table(x)) {
+                       x[ , name_perc := paste0(variable, '(', perc, '%)') ]
+                     } else {
+                       x
+                     }
+                   })
+# add ccl_, hab_, reg_ prefixes
+catalog_l$chemical_class$variable = paste0('ccl_', catalog_l$chemical_class$variable)
+catalog_l$habitat$variable = paste0('hab_', catalog_l$habitat$variable)
+catalog_l$region$variable = paste0('reg_', catalog_l$region$variable)
+
+# meta --------------------------------------------------------------------
[…]

### 2019-09-26 setup.R

@@ -3,11 +3,9 @@
 # project folder ----------------------------------------------------------
 nodename = Sys.info()[4]
 if (nodename == 'scharmueller-t460s') {
-  prj = '/home/scharmueller/Projects/etox-base-shiny'
-  prj_data = '/home/scharmueller/Projects/etox-base'
+  prj = '/home/scharmueller/Projects/standartox-app'
 } else if (nodename == 'uwigis') {
-  prj = '/home/scharmueller/Projects/etox-base-shiny'
-  prj_data = '/home/scharmueller/Projects/etox-base'
+  prj = '/home/scharmueller/Projects/standartox-app'
 } else {
   stop('New system. Define prj and shinydir variables.')
 }
@@ -39,7 +37,7 @@ pkg_cran = c(
 
 pacman::p_load(char = pkg_cran)
 
-# p_update() #! do this manually, as unexpected consequences might occur
+# p_update() # TODO do this manually, as unexpected consequences might occur
 
 # options -----------------------------------------------------------------
 options(stringsAsFactors = FALSE,
@@ -47,28 +45,21 @@ options(stringsAsFactors = FALSE,
         shiny.trace = TRUE)
 
 # variables ---------------------------------------------------------------
-# /etox-base-shiny
 src = file.path(prj, 'R')
-fundir = file.path(src,'functions')
 datadir = file.path(prj, 'data')
-cache = file.path(prj, 'cache')
-# /etox-base
-article = file.path(prj_data, 'article')
+# folder for article references
+article = file.path(gsub('-app', '-build', prj), 'article')
 
 # source ------------------------------------------------------------------
 # functions
-source(file.path(fundir, 'fun_filter.R'))
-source(file.path(fundir, 'fun_aggregation.R'))
-# TODO source(file.path(fundir, 'fun_ssd.R'))
-source(file.path(fundir, 'fun_plotly.R'))
-source(file.path(fundir, 'fun_outliers.R'))
-source(file.path(fundir, 'fun_geometric_mean.R'))
-
-# source(file.path(fundir, 'fun_ec50filter_aggregation_plots.R'))
-# source(file.path(fundir, 'fun_ec50filter_meta_plots.R'))
-# source(file.path(fundir, 'fun_output_stats.R'))
-source(file.path(fundir, 'fun_casconv.R'))
-# plot themes
+source(file.path(src, 'fun_filter.R'))
[…]

### 2019-09-27 README.Rmd
changes

@@
 - [standartox-build](https://github.com/andschar/standartox-build) - Compiles the tool
-- [standartox-app](https://github.com/andschar/standartox-build) - Serves the website and API
+- [standartox-app](https://github.com/andschar/standartox-app) - Serves the website and API
 - [standartox](https://github.com/andschar/standartox) - R-package

### 2019-09-27 R/fun_aggregate.R

@@ -1,15 +1,15 @@
 # function to aggregate input from fun_filter()
 
 fun_aggregate = function(dt = NULL, 
-                         comp = NULL,
+                         comp = c('cname', 'casnr'),
                          vl = 'concentration',
                          agg = c('min', 'gmn', 'med', 'max'),
-                         info = NULL) {
+                         info = c('taxa', 'n')) {
   agg = match.arg(agg, several.ok = TRUE)
   out = dt[ ,
             j = .(min = min(get(vl), na.rm = TRUE),
                   med = median(get(vl), na.rm = TRUE),
-                  gmn = gm_mean(get(vl)),
+                  gmn = exp(mean(log((get(vl))))),
                   max = max(get(vl), na.rm = TRUE),
                   n = .N,
                   taxa = paste0(unique(tax_taxon), collapse = '-')),

### 2019-09-27 api.R

@@ -16,7 +16,7 @@ function(req) {
                       path_info = req$PATH_INFO,
                       http_user_agent = paste0(req$HTTP_USER_AGENT, "@", req$REMOTE_ADDR))
   fwrite(log_df, 'request.log', append = TRUE)
-  cat(unlist(log_df))
+  cat(paste0(unlist(log_df), collapse = '\n'))
   
   plumber::forward()
 }
@@ -145,59 +145,15 @@ function(req) {
   plumber::forward()
 }
 
-# endpoint: data: json ----------------------------------------------------
-# define parameters with type and description
-# name endpoint
-# return output as html/text
-# specify 200 (okay) return
-#* @param:character cas
-#* @param:character concentration_type:character Concentration type (e.g. 'A', 'F')
-#* @param:character chemical_class Chemical class (e.g. 'herbicide')
-#* @param:character taxa taxonomic name (e.g. Algae)
-#* @param:character habitat Organism habitat (e.g. freshwater)
-#* @param:character region Organism geographical region (e.g. europe)
-#* @param:int duration Test duration (e.g. ????)
-#* @param:character effect
-#* @param:character endpoint
-#* @post /filter/json
-#* @json
-function(cas = NULL,
-         concentration_type = NULL,
-         chemical_class = NULL,
-         taxa = NULL,
-         habitat = NULL,
-         region = NULL,
-         duration = NULL,
-         effect = NULL,
-         endpoint = NULL
-         # publ_year = publ_year, # NOTE possible addition
-         # acch = acch, # NOTE possible addition
-         # exposure = exposure, # NOTE possible addition
-         ) {
-  # browser()
-  # data
-  dat = fst::read_fst(file.path('data',
-                                v,
-                                paste0('standartox', v, '.fst')))
-  setDT(dat)
-  # function
-  out = fun_filter(dt = dat,
-                   cas_ = cas,
-                   concentration_type_ = concentration_type,
-                   chemical_class_ = chemical_class,
-                   taxa_ = taxa,
-                   habitat_ = habitat,
-                   region_ = region,
-                   duration_ = duration,
-                   effect_ = effect,
-                   endpoint_ = endpoint)
-  
-  # setnames(out, gsub('hab_|ccl_|reg_', '', names(out)))
-  
-  # jsonlite::write_json(jsonlite::toJSON(out), '/tmp/out.json')
[…]

### 2019-09-30 README.Rmd
bug

@@
 
-Standartox is a data base tool standartizing the retrieval and aggregation of ecotoxicological test data.
+Standartox is a data base tool facilitating the retrieval of ecotoxicological test data.
 It is based on the [EPA ECOTOX data base](https://cfpub.epa.gov/ecotox/) as well as on data from several other chemical data bases and allows users to filter and aggregate ecotoxicological test data in an easy way.
 It can either be accessed via a [web-app](http://standartox.uni-landau.de/) or via the R-package [standartox](https://github.com/andschar/standartox).
-Ecotoxicological test data are used in environmental risk assessment to calculate effect measures such as [TU - Toxic Units](https://en.wikipedia.org/wiki/Toxic_unit) or [SSD - Species Sensitivity Distributions](https://edild.github.io/ssd/) to asses toxicity of chemicals.
+Ecotoxicological test data is used in environmental risk assessment to calculate effect measures such as [TU - Toxic Units](https://en.wikipedia.org/wiki/Toxic_unit) or [SSD - Species Sensitivity Distributions](https://edild.github.io/ssd/) in order to asses environmental toxicity of chemicals..
 

### 2019-09-30 R/fun_geometric_mean.R

@@ -1,7 +1,7 @@
 # function to calculate the geometric mean
 # https://stackoverflow.com/questions/2602583/geometric-mean-is-there-a-built-in
 
-gm_mean = function(x, na.rm=TRUE, zero.propagate = FALSE){
+gm_mean = function(x, na.rm = TRUE, zero.propagate = FALSE){
   if(any(x < 0, na.rm = TRUE)){
     return(NaN)
   }
@@ -15,4 +15,4 @@ gm_mean = function(x, na.rm=TRUE, zero.propagate = FALSE){
   }
 }
 
-gm_mean = function(x) { exp(mean(log(x))) }
+# gm_mean = function(x) { exp(mean(log(x))) }

### 2019-09-30 api.R

@@ -21,15 +21,6 @@ function(req) {
   plumber::forward()
 }
 
-# filter: debuging --------------------------------------------------------
-# save req objects to /tmp for debugging
-#* @filter debugger
-function(req) {
-  req_envir <<- req
-  
-  plumber::forward()
-}
-
 # filter vers --------------------------------------------------------------
 #* version path
 #* @filter vers
@@ -208,16 +199,6 @@ function(cas = NULL,
   cat('\n', tmp)
   
   readBin(tmp, "raw", n = file.size(tmp))
-}
-
-# debug: results ----------------------------------------------------------
-#* 
-#* @filter result
-function(res) {
-  
-  res_envir <<- res
-  
-  plumber::forward()
 }
 
 # endpoint: catalog -----------------------------------------------------

### 2019-10-14 article/parts/04_technical_validation.tex
Updates from Overleaf

@@
 
+\begin{figure}
+    \includegraphics[width=1\linewidth]{article/figures/Standartox_PPDB.png}
+    \caption{
+        Comparison between Standartox and PPDB values
+    }
+    \label{fig:standartox_ppdb_diff}
+\end{figure}
+
+In order to validate \standartox we compared resulting values with values from other data bases, such as the PPDB.
+Of 542 chemicals, only 16\% diverged from the PPDB values by a factor greater than 10, supporting plausibility of the \standartox results \ref{fig:standartox_ppdb_diff}.
+
 <PUT COMPARISON WITH PPDB DATA HERE?>

### 2019-10-14 talk/talk43.Rmd

@@
 ```{r setup, echo=FALSE, include=FALSE, warning=FALSE, message=FALSE}
-src = '/home/scharmueller/Projects/etox-base/R'
+src = '/home/scharmueller/Projects/standartox-build/R'
 source(file.path(src, 'gn_setup.R'))
@@
 <img src="https://www.uni-koblenz-landau.de/de/uni/organisation/verwaltung/abteilungen/abt-1/dokumente-ab1/uni-logo-farbig.jpg" width="200">
+<img src="https://www.uni-koblenz-landau.de/de/uni/organisation/verwaltung/abteilungen/abt-1/dokumente-ab1/uni-logo-farbig.jpg" width="200">
+<img src="https://de.wikipedia.org/wiki/Universit%C3%A4t_Stra%C3%9Fburg#/media/Datei:Universit%C3%A9_de_Strasbourg.svg" width="100">
 
@@
 Bernhardt et al. 2017
-<img src="/home/scharmueller/Projects/etox-base/talk/data/bernhardt_et_al_chem.png" width="300px">
+<img src="/home/scharmueller/Projects/standartox-build/talk/data/bernhardt_et_al_chem.png" width="300px">
 
@@
 ---
-background-image: url(/home/scharmueller/Projects/etox-base/talk/data/stream.jpg)
+background-image: url(/home/scharmueller/Projects/standartox-build/talk/data/stream.jpg)
 background-size: 400px
@@
 --
-<img src="/home/scharmueller/Projects/etox-base/talk/data/epa_ecotox_screenshot.png" width="800px">
+<img src="/home/scharmueller/Projects/standartox-build/talk/data/epa_ecotox_screenshot.png" width="800px">
 
@@
 
-<img src="/home/scharmueller/Projects/etox-base/talk/data/epa_ecotox_screenshot.png" width="500px">
+<img src="/home/scharmueller/Projects/standartox-build/talk/data/epa_ecotox_screenshot.png" width="500px">
 
@@
 
-<img src="/home/scharmueller/Projects/etox-base/talk/data/organigram2.png" width="950px" style="position:absolute; left:10px; top:130px;">
+<img src="/home/scharmueller/Projects/standartox-build/talk/data/organigram2.png" width="950px" style="position:absolute; left:10px; top:130px;">
 
@@
 --
-<img src="/home/scharmueller/Projects/etox-base/talk/data/standartox_screenshot.png" width="800px">
+<img src="/home/scharmueller/Projects/standartox-build/talk/data/standartox_screenshot.png" width="800px">
 
@@
 .pull-left[
-<img src="/home/scharmueller/Projects/etox-base/talk/data/standartox_screenshot.png" width="500px">
+<img src="/home/scharmueller/Projects/standartox-build/talk/data/standartox_screenshot.png" width="500px">
 ### Aim: Prepare ecotoxicological test data in a __reproducible__ and __standardized__ way
@@
 
-standartox(cas = '52-68-6',
+stx_query(cas = '52-68-6',
           endpoint = 'EC50',
@@
 
-<img src="/home/scharmueller/Projects/etox-base/talk/data/logos/worms.png" width="250px" style="position:absolute; bottom:30px; left:260px;">
+<img src="/home/scharmueller/Projects/standartox-build/talk/data/logos/worms.png" width="250px" style="position:absolute; bottom:30px; left:260px;">
 
-<img src="/home/scharmueller/Projects/etox-base/talk/data/logos/chebi.png" width="250px" style="position:absolute; bottom:140px; right:110px;">
+<img src="/home/scharmueller/Projects/standartox-build/talk/data/logos/chebi.png" width="250px" style="position:absolute; bottom:140px; right:110px;">
 
-<img src="/home/scharmueller/Projects/etox-base/talk/data/logos/pubchem.png" width="250px" style="position:absolute; bottom:25px; right:90px;">
+<img src="/home/scharmueller/Projects/standartox-build/talk/data/logos/pubchem.png" width="250px" style="position:absolute; bottom:25px; right:90px;">
 

### 2019-10-14 Rmd/summary_db_cols_norman.Rmd
added supplement dir

@@
 ```{r setup, echo=FALSE, include=FALSE}
-prj = '/home/scharmueller/Projects/etox-base'
+prj = '/home/scharmueller/Projects/standartox-build'
 source(file.path(prj, 'R', 'gn_setup.R'))
@@
      FROM norman.variables"
-var = read_tbl(user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
+var = read_query(user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
                  query = q)

### 2019-10-14 article/supplement/r-package-list.tex
added supplement dir

Rinker TW, Kurkiewicz D (2018).
\emph{pacman: Package Management for R}.
version 0.5.0, \url{http://github.com/trinker/pacman}.
R Core Team (2019).
\emph{R: A Language and Environment for Statistical Computing}.
R Foundation for Statistical Computing, Vienna, Austria.
\url{https://www.R-project.org/}.
Wickham H, Hester J, Chang W (2019).
\emph{devtools: Tools to Make Developing R Packages Easier}.
R package version 2.2.1, \url{https://CRAN.R-project.org/package=devtools}.
Lang DT, team tC (2019).
\emph{RCurl: General Network (HTTP/FTP/...) Client Interface for R}.
R package version 1.95-4.12, \url{https://CRAN.R-project.org/package=RCurl}.
Wickham H (2019).
\emph{stringr: Simple, Consistent Wrappers for Common String Operations}.
R package version 1.4.0, \url{https://CRAN.R-project.org/package=stringr}.
Bengtsson H (2019).
\emph{R.utils: Various Programming Utilities}.
R package version 2.9.0, \url{https://CRAN.R-project.org/package=R.utils}.
Hiebert J (2016).
\emph{udunits2: Udunits-2 Bindings for R}.
R package version 0.13, \url{https://CRAN.R-project.org/package=udunits2}.
Wickham H (2019).
\emph{rvest: Easily Harvest (Scrape) Web Pages}.
R package version 0.3.4, \url{https://CRAN.R-project.org/package=rvest}.
Wickham H (2019).
\emph{httr: Tools for Working with URLs and HTTP}.
R package version 1.4.1, \url{https://CRAN.R-project.org/package=httr}.
Ooms J (2014).
``The jsonlite Package: A Practical and Consistent Mapping Between JSON Data and R Objects.''
\emph{arXiv:1403.2805 [stat.CO]}.
\url{https://arxiv.org/abs/1403.2805}.
Wickham H, Bryan J (2019).
\emph{readxl: Read Excel Files}.
R package version 1.3.1, \url{https://CRAN.R-project.org/package=readxl}.
Walker A (2019).
\emph{openxlsx: Read, Write and Edit XLSX Files}.
R package version 4.1.0.1, \url{https://CRAN.R-project.org/package=openxlsx}.
Henry L, Wickham H (2019).
\emph{purrr: Functional Programming Tools}.
R package version 0.3.2, \url{https://CRAN.R-project.org/package=purrr}.
Dowle M, Srinivasan A (2019).
\emph{data.table: Extension of `data.frame`}.
R package version 1.12.4, \url{https://CRAN.R-project.org/package=data.table}.
Conway J, Eddelbuettel D, Nishiyama T, Prayaga SK, Tiffin N (2017).
\emph{RPostgreSQL: R Interface to the 'PostgreSQL' Database System}.
R package version 0.6-2, \url{https://CRAN.R-project.org/package=RPostgreSQL}.
R Special Interest Group on Databases (R-SIG-DB), Wickham H, Müller K (2018).
\emph{DBI: R Database Interface}.
R package version 1.0.0, \url{https://CRAN.R-project.org/package=DBI}.
Oksanen J, Blanchet FG, Friendly M, Kindt R, Legendre P, McGlinn D, Minchin PR, O'Hara RB, Simpson GL, Solymos P, Stevens MHH, Szoecs E, Wagner H (2019).
\emph{vegan: Community Ecology Package}.
R package version 2.5-5, \url{https://CRAN.R-project.org/package=vegan}.
Wickham H (2011).
``The Split-Apply-Combine Strategy for Data Analysis.''
\emph{Journal of Statistical Software}, \bold{40}(1), 1--29.
\url{http://www.jstatsoft.org/v40/i01/}.
Komsta L (2011).
\emph{outliers: Tests for outliers}.
R package version 0.14, \url{https://CRAN.R-project.org/package=outliers}.
Wickham H (2019).
\emph{feather: R Bindings to the Feather 'API'}.
R package version 0.3.3, \url{https://CRAN.R-project.org/package=feather}.
Wickham H (2016).
\emph{ggplot2: Elegant Graphics for Data Analysis}.
Springer-Verlag New York.
ISBN 978-3-319-24277-4, \url{https://ggplot2.tidyverse.org}.
Slowikowski K (2019).
\emph{ggrepel: Automatically Position Non-Overlapping Text Labels with
'ggplot2'}.
R package version 0.8.1, \url{https://CRAN.R-project.org/package=ggrepel}.
Wilke C (2019).
\emph{cowplot: Streamlined Plot Theme and Plot Annotations for 'ggplot2'}.
R package version 0.9.4, \url{https://CRAN.R-project.org/package=cowplot}.
Neuwirth E (2014).
\emph{RColorBrewer: ColorBrewer Palettes}.
R package version 1.1-2, \url{https://CRAN.R-project.org/package=RColorBrewer}.
Tennekes M (2017).
\emph{treemap: Treemap Visualization}.
R package version 2.4-2, \url{https://CRAN.R-project.org/package=treemap}.
Chamberlain S, Barve V, Mcglinn D, Oldoni D, Desmet P, Geffert L, Ram K (2019).
\emph{rgbif: Interface to the Global Biodiversity Information Facility API}.
R package version 1.3.0, \url{https://CRAN.R-project.org/package=rgbif}.

Chamberlain S, Boettiger C (2017).
``R Python, and Ruby clients for GBIF species occurrence data.''
\emph{PeerJ PrePrints}.
\url{https://doi.org/10.7287/peerj.preprints.3304v1}.
Scott Chamberlain, Eduard Szocs (2013).
``taxize - taxonomic search and retrieval in R.''
\emph{F1000Research}.
\url{http://f1000research.com/articles/2-191/v2}.

Chamberlain S, Szoecs E, Foster Z, Arendsee Z, Boettiger C, Ram K, Bartomeus I, Baumgartner J, O'Donnell J, Oksanen J, Tzovaras BG, Marchand P, Tran V, Salmon M, Li G, Grenié M (2019).
\emph{taxize: Taxonomic information from around the web}.
R package version 0.9.7, \url{https://github.com/ropensci/taxize}.
Arel-Bundock V, Enevoldsen N, Yetman C (2018).
``countrycode: An R package to convert country names and country codes.''
\emph{Journal of Open Source Software}, \bold{3}(28), 848.
\url{https://doi.org/10.21105/joss.00848}.
Microsoft, Weston S (2017).
\emph{foreach: Provides Foreach Looping Construct for R}.
R package version 1.4.4, \url{https://CRAN.R-project.org/package=foreach}.
Corporation M, Weston S (2018).
\emph{doParallel: Foreach Parallel Adaptor for the 'parallel' Package}.
R package version 1.0.14, \url{https://CRAN.R-project.org/package=doParallel}.
Klik M (2019).
\emph{fst: Lightning Fast Serialization of Data Frames for R}.
R package version 0.9.0, \url{https://CRAN.R-project.org/package=fst}.
Xie Y, Cheng J, Tan X (2019).
\emph{DT: A Wrapper of the JavaScript Library 'DataTables'}.
R package version 0.7, \url{https://CRAN.R-project.org/package=DT}.
Tennekes M (2017).
\emph{treemap: Treemap Visualization}.
R package version 2.4-2, \url{https://CRAN.R-project.org/package=treemap}.

### 2019-10-14 R/fun_read_db.R

@@ -1,7 +1,7 @@
 # function to read tables from a PostgreSQL data base
 
-read_tbl = function(user = NULL, host = NULL, port = NULL, password = NULL,
-                    dbname = NULL, query = NULL) {
+read_query = function(user = NULL, host = NULL, port = NULL, password = NULL,
+                      dbname = NULL, query = NULL) {
   
   if (is.null(query)) {
     stop('No query supplied.')

### 2019-10-14 R/gn_setup.R

@@ -174,8 +174,13 @@ pkg = c('pacman', pkg_cran)
 
 for (i in pkg) {
   capture.output(
-    utils:::print.bibentry(citation(i), style = "Bibtex"),
+    print(citation(i), style = "Bibtex"),
     file = file.path(article, 'refs', 'references-standartox-build.bib'),
+    append = TRUE
+  )
+  capture.output(
+    print(citation(i), style = 'latex'),
+    file = file.path(article, 'supplement', 'r-package-list.tex'),
     append = TRUE
   )
 }

### 2019-10-14 run_build.R

@@ -115,7 +115,7 @@ source(file.path(src, 'gn_end.R'), max.deparse.length = mdl)
 if (sink_console) {
   # Restore output to console
   sink()
-  sink(type="message")
+  sink(type = "message")
 }
 
 

### 2019-10-14 article/00_article.tex
Updates from Overleaf

--- 2019-06-12 article/article.tex
@@
 \newcommand{\epa}{EPA ECOTOX data base}
-\newcommand{\app}{http://139.14.20.252:3838/etox-base-shiny/}
+\newcommand{\rpackage}{http://139.14.20.252:3838/etox-base-shiny/}
+\newcommand{\app}{www.standartox.uni-landau.de}
 \newcommand{\git}{https://github.com/andreasLD/etox-base}
@@
 
-\title{Etox-Base: A tool to derive exposure endpoints from chemical test results}
+\title{Standartox: A tool to derive exposure endpoints from chemical test results}
 
@@
 1. Institute of Environmental Sciences, University of Koblenz-Landau Fortstraße 7, 76829 Landau, Germany {*}corresponding author(s):
-Andreas Scharm{\"u}ller (scharmueller@uni-landau.de)
+Andreas Scharm{\"u}ller (andschar@protonmail.com)
 
@@
 \bibliographystyle{apalike}
-\bibliography{refs/references-etox-base.bib,refs/references-etox-base-rpackages.bib,refs/references-etox-base-rshinypackages.bib}
+\bibliography{refs/references-standartox-build.bib,refs/references-standartox-app.bib}
 % multiple refs have to be written w/o space

### 2019-10-14 article/parts/02_methods.tex
Updates from Overleaf

@@
 
-The \etoxbase{} consists of two parts of software: Firstly a processing pipeline (Fig.
-\ref{fig:pipeline}) downloads quarterly released new versions of the \epa{} and performs several preparation steps in R \citep{r_core_team_r_2017}, to then store the outcome in a PostgreSQL data base.
-This constitutes the basis upon which the R shiny web application is accessing the data \citep{r_core_team_r_2017}.
+Standartox consists of three software parts which are explained in detail here.
+Firstly a processing pipeline (Fig.
+\ref{fig:pipeline}) downloads quarterly released new versions of the EPA ECOTOX data base and performs several preparation steps in R \citep{r_core_team_r_2017}, to then store the outcome in a PostgreSQL <CITE> data base.
+This constitutes the basis upon which the web application and and application programming interface (API) together with a R package are accessing the data \citep{r_core_team_r_2017}.
+
+\subsection{}
+
 The application itself is set of R functions that filter and aggregate the data according to a user's inputs \ref{fig:app}.
@@
 
-(bd\_epa\_download.R) 
-(bd\_epa\_postgres.R)
-(da\_epa1.R, da\_epa2.R, da\_epa3.R)
-(qu\_aw.R)
-(qu\_chemspieder\_scrape.R) 
-(qu\_pc.R) 
-(qu\_pp.R)
-(qu\_worms2.R) 
-(qu\_rgbif.R) 
-(re\_merge.R) 
-(re\_combine.R)
-(re\_checks\_internal.R) 
-(re\_analyses.R)
-(re\_final.R)
-
 The EPA ECOTOX data is downloaded and subsequently built into a PostgreSQL data base.
 Afterwards the data is imported into R for pre-processing, which includes validity checks, removal of special characters, unit harmonizations and selection of relevant variables.
-Chemical abstract service (CAS) numbers, unique chemical identifiers together we taxon names are then used to query other data bases for additional information on chemicals and biota respectively.
-Chemical information is retrieved from the Compendium of Pesticide Common Names \citep{CITE_AW}, the Chemspider data base \citep{CITE-CHEMSPIDER}, from Eurostat <CITATION>, the Pubchem data base \citep{CITE_PUBCHEM}, and from the Physprop data base \citep{CITE_PHYSPROP}.
+Chemical abstract service (CAS) numbers, unique chemical identifiers together with taxon names are then used to query other data bases for additional information on chemicals and organisms respectively.
+Chemical information is retrieved inter alia from the Compendium of Pesticide Common Names \citep{CITE_AW}, the Chemspider data base \citep{CITE-CHEMSPIDER}, from Eurostat <CITATION>, the Pubchem data base \citep{CITE_PUBCHEM}, and from the Physprop data base \citep{CITE_PHYSPROP}.
 These queries mostly rely on the webchem R-package \citep{szocs_webchem_2015-1}.
@@
 
+\subsubsection*{Aggregation}
+The \etoxbase{} application aggregates the test results according to chosen filters in a two step process.
+Firstly the filtered test results are aggregated by the CAS number, the chosen taxon and the selected test duration.
+Secondly, the returned data is then aggregated by the CAS number.
+The former can't be influenced by the user and calculates either the minimum or the median depending on the amount of results to aggregate (n <= 2: minimum, if n > 2: median).
+Thereof the second step calculates the minimum, the maximum, the median, the geometric mean, or the arithmetic mean as an aggregate.
+
+\subsubsection*{Future of the project}
+Since new versions of the EPA ECOTOX data base are published on a quarterly basis the data acquisition process is automated and new versions of the Standartox are scheduled to be released regularly.
+
+The code for the web-application and the application programming interface are placed here: https://github.com/andschar/standartox-app
+
 \begin{figure}
@@
 
-\subsection*{Aggregation}
-The \etoxbase{} application aggregates the test results according to chosen filters in a two step process.
-Firstly the filtered test results are aggregated by the CAS number, the chosen taxon and the selected test duration.
-Secondly, the returned data is then aggregated by the CAS number.
-The former can't be influenced by the user and calculates either the minimum or the median depending on the amount of results to aggregate (n <= 2: minimum, if n > 2: median).
-Thereof the second step calculates the minimum, the maximum, the median, the geometric mean, or the arithmetic mean as an aggregate.
+\subsection*{The R package}
+
+The code for the R-package standartox is placed here: https://github.com/andschar/standartox
 
@@
 
-The processing code is stored in a Github reopsitory (\git{}).
-All processing scripts were written in R version 3.4.4.
-PostgreSQL 9.5 has been used to build the data base.
-Hence, R together with its packages \ref{tab:rpackages} and PostgreSQL suffice to rebuild the process.
-The code of the application itself is stored in a second Github repostitory (\gitapp{}).
-Since new versions of the EPA ECOTOX data base are published on a quarterly basis the data acquisition process is automated and new versions of the application are scheduled to be released regularly.
+The processing pipeline code is stored here https://github.com/andschar/standartox-build, the code for the web application is stored under https://github.com/andschar/standartox-app and the R-pckage code is stored in https://github.com/andschar/standartox.
+The code can freely be accessed under the MIT License <CITATION>.
+Most of the code is written in R 3.6.1 and some parts in PostgreSQL 9.6.1.
+An exhaustive list of additional R-pakage used can be found in the supplementary material <REFERENCE-TO-SUPPL>.
 

### 2019-10-14 article/parts/01_background.tex

@@
 
-A large number of chemicals such as pharmaceuticals, pesticides and synthetic hormones are in daily use all over the world with insufficient knowledge about possible adverse effects on the environment.
+The risk of adverse effects of chemicals on the environment is globally assessed with the help of ecotoxicological tests that are conducted on different organism groups, ideally representing a specific biocoenosis and an associated biotope.
+These tests are conducted by different research groups all over the world who publish the results independently.
+Only few initiatives, such as the United States Environmental Protection Agency (EPA) ECOTOXicology knowledgebase (EPA ECOTOX), the Pesticide Property Data Base (PPDB) or the ETOX (UBA ETOX) <cite UBA> exist which aim to consolidate test results.
+However, the EPA ECOTOX data base and the UBA ETOX only collect test results only partly checking for quality.
+In contrast the PPDB data base aims for quality, yet is confined to only one specific class of chemicals, pesticides.
+An overall holistic approach in processing ecotoxicological test results is lacking.
+Standartox, the tool suggested here aims to overcome these paucities by incorporating the ever-growing number of test results in an automated process workflow that ultimately leads to single aggregated test results for specific chemical-organism test combinations.
+
+This is important because an ever-growing number of chemicals such as pharmaceuticals, pesticides and synthetic hormones amongst others are in daily use all over the world with insufficient knowledge about possible adverse effects on the environment.
 In Europe alone some 100,000 chemicals are estimated to be in current use, whereof 30,000 are produced in quantities larger than one ton per year \citep{breithaupt_costs_2006}.
-Although for some chemicals advanced and standardized test methods \citep{oecd_oecd_2018} have been developed, data on toxicity tests for a lot of chemicals, especially newly emerging ones remains scarce and if existent, the information is represented sparsely and in inconsistent formats \citep{gessner_fostering_2016}.
-Therefore we developed \standardtox{}, a tool to aggregate and standardize toxicity test data from various sources [EPA and ETOX] which provides the information in a reproducible manner.
-
-To our knowledge this ist the first approach
-
-reporducible
-
-time saving
-
-data base that collects, processes and aggregates ecotoxicological tests results in order to subsequently publish it in an harmonized form as a web application - the \etoxbase{} tool: \href{http://139.14.20.252:3838/etox-base-shiny/}.
+However, data on toxicity tests for a lot of chemicals, especially newly emerging ones remains scarce and if existent, the information is represented sparsely and in inconsistent formats \citep{gessner_fostering_2016}.
 
@@
 By the time of writing it contained 926,108 test results, comprising 11,685 chemicals tested on 12,668 different species \citep{elonen_ecotoxicology_2018}.
+
 The collected ecotoxicological test data contain a variety of measured endpoints such as Effective Concentrations (\ecfifty{}) values, No-observed effect concentrations (NOEC) or lowest observed effect concentrations (LOEC).

### 2019-10-15 article/supplement/r-package-list.tex
article

@@
 version 0.5.0, \url{http://github.com/trinker/pacman}.
-R Core Team (2019).
+\newline R Core Team (2019).
 \emph{R: A Language and Environment for Statistical Computing}.
@@
 \url{https://www.R-project.org/}.
-Wickham H, Hester J, Chang W (2019).
+\newline Wickham H, Hester J, Chang W (2019).
 \emph{devtools: Tools to Make Developing R Packages Easier}.
 R package version 2.2.1, \url{https://CRAN.R-project.org/package=devtools}.
-Lang DT, team tC (2019).
+\newline Lang DT, team tC (2019).
 \emph{RCurl: General Network (HTTP/FTP/...) Client Interface for R}.
 R package version 1.95-4.12, \url{https://CRAN.R-project.org/package=RCurl}.
-Wickham H (2019).
+\newline Wickham H (2019).
 \emph{stringr: Simple, Consistent Wrappers for Common String Operations}.
 R package version 1.4.0, \url{https://CRAN.R-project.org/package=stringr}.
-Bengtsson H (2019).
+\newline Bengtsson H (2019).
 \emph{R.utils: Various Programming Utilities}.
 R package version 2.9.0, \url{https://CRAN.R-project.org/package=R.utils}.
-Hiebert J (2016).
+\newline Hiebert J (2016).
 \emph{udunits2: Udunits-2 Bindings for R}.
 R package version 0.13, \url{https://CRAN.R-project.org/package=udunits2}.
-Wickham H (2019).
+\newline Wickham H (2019).
 \emph{rvest: Easily Harvest (Scrape) Web Pages}.
 R package version 0.3.4, \url{https://CRAN.R-project.org/package=rvest}.
-Wickham H (2019).
+\newline Wickham H (2019).
 \emph{httr: Tools for Working with URLs and HTTP}.
 R package version 1.4.1, \url{https://CRAN.R-project.org/package=httr}.
-Ooms J (2014).
+\newline Ooms J (2014).
 ``The jsonlite Package: A Practical and Consistent Mapping Between JSON Data and R Objects.''
@@
 \url{https://arxiv.org/abs/1403.2805}.
-Wickham H, Bryan J (2019).
+\newline Wickham H, Bryan J (2019).
 \emph{readxl: Read Excel Files}.
 R package version 1.3.1, \url{https://CRAN.R-project.org/package=readxl}.
-Walker A (2019).
+\newline Walker A (2019).
 \emph{openxlsx: Read, Write and Edit XLSX Files}.
 R package version 4.1.0.1, \url{https://CRAN.R-project.org/package=openxlsx}.
-Henry L, Wickham H (2019).
+\newline Henry L, Wickham H (2019).
 \emph{purrr: Functional Programming Tools}.
 R package version 0.3.2, \url{https://CRAN.R-project.org/package=purrr}.
-Dowle M, Srinivasan A (2019).
+\newline Dowle M, Srinivasan A (2019).
 \emph{data.table: Extension of `data.frame`}.
 R package version 1.12.4, \url{https://CRAN.R-project.org/package=data.table}.
-Conway J, Eddelbuettel D, Nishiyama T, Prayaga SK, Tiffin N (2017).
+\newline Conway J, Eddelbuettel D, Nishiyama T, Prayaga SK, Tiffin N (2017).
 \emph{RPostgreSQL: R Interface to the 'PostgreSQL' Database System}.
 R package version 0.6-2, \url{https://CRAN.R-project.org/package=RPostgreSQL}.
-R Special Interest Group on Databases (R-SIG-DB), Wickham H, Müller K (2018).
+\newline R Special Interest Group on Databases (R-SIG-DB), Wickham H, Müller K (2018).
 \emph{DBI: R Database Interface}.
 R package version 1.0.0, \url{https://CRAN.R-project.org/package=DBI}.
-Oksanen J, Blanchet FG, Friendly M, Kindt R, Legendre P, McGlinn D, Minchin PR, O'Hara RB, Simpson GL, Solymos P, Stevens MHH, Szoecs E, Wagner H (2019).
+\newline Oksanen J, Blanchet FG, Friendly M, Kindt R, Legendre P, McGlinn D, Minchin PR, O'Hara RB, Simpson GL, Solymos P, Stevens MHH, Szoecs E, Wagner H (2019).
 \emph{vegan: Community Ecology Package}.
 R package version 2.5-5, \url{https://CRAN.R-project.org/package=vegan}.
-Wickham H (2011).
+\newline Wickham H (2011).
 ``The Split-Apply-Combine Strategy for Data Analysis.''
@@
 \url{http://www.jstatsoft.org/v40/i01/}.
-Komsta L (2011).
+\newline Komsta L (2011).
 \emph{outliers: Tests for outliers}.
 R package version 0.14, \url{https://CRAN.R-project.org/package=outliers}.
-Wickham H (2019).
+\newline Wickham H (2019).
 \emph{feather: R Bindings to the Feather 'API'}.
 R package version 0.3.3, \url{https://CRAN.R-project.org/package=feather}.
-Wickham H (2016).
+\newline Wickham H (2016).
 \emph{ggplot2: Elegant Graphics for Data Analysis}.
@@
 ISBN 978-3-319-24277-4, \url{https://ggplot2.tidyverse.org}.
-Slowikowski K (2019).
+\newline Slowikowski K (2019).
 \emph{ggrepel: Automatically Position Non-Overlapping Text Labels with
@@
 R package version 0.8.1, \url{https://CRAN.R-project.org/package=ggrepel}.
-Wilke C (2019).
+\newline Wilke C (2019).
 \emph{cowplot: Streamlined Plot Theme and Plot Annotations for 'ggplot2'}.
 R package version 0.9.4, \url{https://CRAN.R-project.org/package=cowplot}.
-Neuwirth E (2014).
+\newline Neuwirth E (2014).
 \emph{RColorBrewer: ColorBrewer Palettes}.
 R package version 1.1-2, \url{https://CRAN.R-project.org/package=RColorBrewer}.
-Tennekes M (2017).
+\newline Tennekes M (2017).
 \emph{treemap: Treemap Visualization}.
 R package version 2.4-2, \url{https://CRAN.R-project.org/package=treemap}.
-Chamberlain S, Barve V, Mcglinn D, Oldoni D, Desmet P, Geffert L, Ram K (2019).
+\newline Chamberlain S, Barve V, Mcglinn D, Oldoni D, Desmet P, Geffert L, Ram K (2019).
 \emph{rgbif: Interface to the Global Biodiversity Information Facility API}.
@@
 \url{https://doi.org/10.7287/peerj.preprints.3304v1}.
-Scott Chamberlain, Eduard Szocs (2013).
+\newline Scott Chamberlain, Eduard Szocs (2013).
 ``taxize - taxonomic search and retrieval in R.''
@@
 R package version 0.9.7, \url{https://github.com/ropensci/taxize}.
-Arel-Bundock V, Enevoldsen N, Yetman C (2018).
+\newline Arel-Bundock V, Enevoldsen N, Yetman C (2018).
 ``countrycode: An R package to convert country names and country codes.''
@@
 \url{https://doi.org/10.21105/joss.00848}.
-Microsoft, Weston S (2017).
+\newline Microsoft, Weston S (2017).
 \emph{foreach: Provides Foreach Looping Construct for R}.
 R package version 1.4.4, \url{https://CRAN.R-project.org/package=foreach}.
-Corporation M, Weston S (2018).
+\newline Corporation M, Weston S (2018).
 \emph{doParallel: Foreach Parallel Adaptor for the 'parallel' Package}.
 R package version 1.0.14, \url{https://CRAN.R-project.org/package=doParallel}.
-Klik M (2019).
+\newline Klik M (2019).
 \emph{fst: Lightning Fast Serialization of Data Frames for R}.
 R package version 0.9.0, \url{https://CRAN.R-project.org/package=fst}.
-Xie Y, Cheng J, Tan X (2019).
+\newline Xie Y, Cheng J, Tan X (2019).
 \emph{DT: A Wrapper of the JavaScript Library 'DataTables'}.
 R package version 0.7, \url{https://CRAN.R-project.org/package=DT}.
-Tennekes M (2017).
+\newline Tennekes M (2017).
 \emph{treemap: Treemap Visualization}.
 R package version 2.4-2, \url{https://CRAN.R-project.org/package=treemap}.
+\newline

### 2019-10-15 R/gn_setup.R

@@ -171,16 +171,20 @@ theme_set(theme_minimal_etox_base_sans)
 
 # cite packages -----------------------------------------------------------
 pkg = c('pacman', pkg_cran)
-
+fl_bib = file.path(article, 'refs', 'references-standartox-build.bib')
+fl_tex = file.path(article, 'supplement', 'r-package-list.tex')
+file.remove(fl_bib)
+file.remove(fl_tex)
 for (i in pkg) {
   capture.output(
     print(citation(i), style = "Bibtex"),
-    file = file.path(article, 'refs', 'references-standartox-build.bib'),
+    file = fl_bib,
     append = TRUE
   )
   capture.output(
     print(citation(i), style = 'latex'),
-    file = file.path(article, 'supplement', 'r-package-list.tex'),
+    cat('\\newline '), # HACK
+    file = fl_tex,
     append = TRUE
   )
 }

### 2019-10-15 article/parts/02_methods.tex

@@
 
-Standartox consists of three software parts which are explained in detail here.
-Firstly a processing pipeline (Fig.
+Standartox consists of three software parts which are explained in detail below.
+Firstly a data acquisition and preparation processing pipeline (Fig.
 \ref{fig:pipeline}) downloads quarterly released new versions of the EPA ECOTOX data base and performs several preparation steps in R \citep{r_core_team_r_2017}, to then store the outcome in a PostgreSQL <CITE> data base.
-This constitutes the basis upon which the web application and and application programming interface (API) together with a R package are accessing the data \citep{r_core_team_r_2017}.
+This constitutes the basis upon which the web application and and application programming interface (API) together with a R package are accessing the data.
 
-\subsection{}
+\subsection*{Data acquisition \& preparation}
+
+In order to build Standartox the EPA ECOTOX data is downloaded and subsequently built into a PostgreSQL data base.
+From there, the data is processed, including validity checks, removal of special characters, unit harmonizations and selection of relevant variables.
+Chemical abstract service (CAS) numbers, unique chemical identifiers together with taxon names are then used to query additional information from other data bases (Table \ref{tab:data-base}) on chemicals and organisms respectively.
+Standartox itself is finally built and exported as a binary object to be used by the web application and the API.
+
+\subsection*{Web application and application programming interface}
+
+Once called, both the web application and the API load the binary Standartox file into memory and allow the user to interact with it.
+
+\subsubsection*{Web application}
+The interactive web application was built in R, using the shiny web application framework \citep{chang_shiny_2018} which runs with the help of a shiny server \citep{HOW-TO-CITE-SHINY-SERVER} on www.standartox.uni-landau.de.
+The app itstel lies in the file ~/standartox-app/app.R.
+
+\subsubsection*{Application programming interface (API)}
+The API is built in R with the plumber-package \citep{}
 
@@
 
-\subsection*{Data acquisition \& preparation}
+\subsubsection*{R-package}
 
-The EPA ECOTOX data is downloaded and subsequently built into a PostgreSQL data base.
-Afterwards the data is imported into R for pre-processing, which includes validity checks, removal of special characters, unit harmonizations and selection of relevant variables.
-Chemical abstract service (CAS) numbers, unique chemical identifiers together with taxon names are then used to query other data bases for additional information on chemicals and organisms respectively.
-Chemical information is retrieved inter alia from the Compendium of Pesticide Common Names \citep{CITE_AW}, the Chemspider data base \citep{CITE-CHEMSPIDER}, from Eurostat <CITATION>, the Pubchem data base \citep{CITE_PUBCHEM}, and from the Physprop data base \citep{CITE_PHYSPROP}.
-These queries mostly rely on the webchem R-package \citep{szocs_webchem_2015-1}.
-For biota, habitat and occurrence information is queried from the World Register of Marine Species (WORMS) \citep{WORMS} and from the Global Biodiversity Information Facility (GBIF) \citep{CITE_RGBIF} using the rgbif R-package \citep{chamberlain_rgbif_2018}.
-The acquired information is then merged and information is combined.
-From there on the data set is analyzed and checked and finally compiled in one table.
-This table is then used as the input for the filter and aggregation functions \ref{fig:pipeline}.
-For a detailed overview of what information is collected see table \ref{tab:scripts-pipeline}.
-
-\begin{figure}
-    \includestandalone[width=\textwidth, scale=0.1]{tikz/pipeline_organigram}
-    \caption{Processing Pipeline}
-    \label{fig:pipeline}
-\end{figure}
-
-\subsection*{The application}
-The application is accessible through \app{} and was built in R, using the shiny web application framework \citep{chang_shiny_2018}.
-Users can define several input parameters \ref{tab:inputs} in the application's graphical user interface (GUI), handled in the ui.R script in order to adjust the aggregation according specific requirements.
-In doing so users execute R functions (fun\_filter.R, fun\_aggregation.R, fun\_filagg\_plot\_ly.R) loaded in the server.R script, which produce a 'filtered' and an 'aggregated' data set as well as an interactive overview plot.
-Only the 'aggregated' data set is presented in the GUI.
-However, both data sets can be downloaded via a download button.
+httr,
+jsonlite,
+fst,
+data.table
 

### 2019-10-15 article/parts/05_usage_notes.tex

@@
 %%on data use.
+
+\subsection*{Web application}
+
+Users can define several input parameters \ref{tab:inputs} in the application's graphical user interface (GUI), handled in the ui.R script in order to adjust the aggregation according specific requirements.
+In doing so users execute R functions (fun\_filter.R, fun\_aggregation.R, fun\_filagg\_plot\_ly.R) loaded in the server.R script, which produce a 'filtered' and an 'aggregated' data set as well as an interactive overview plot.
+Only the 'aggregated' data set is presented in the GUI.
+However, both data sets can be downloaded via a download button.
 

### 2019-10-15 article/00_article.tex
Updates from Overleaf

@@
 \newcommand{\rpackage}{http://139.14.20.252:3838/etox-base-shiny/}
-\newcommand{\app}{www.standartox.uni-landau.de}
 \newcommand{\git}{https://github.com/andreasLD/etox-base}
@@
 
-\begin{table}[h!]
-\begin{tabular}{|l|l|}
-\hline
-Script                      & Description                                      \\ \hline
-run\_build.R                 & runs all sub scripts                             \\ \hline
-setup.R                     & sets up the environment                          \\ \hline
-bd\_epa\_download.R         & download EPA Ecotox data base                    \\ \hline
-bd\_epa\_postgres.R         & build EPA Ecotox data base locally               \\ \hline
-da\_epa1.R                  & raw EPA Ecotox data export                       \\ \hline
-da\_epa2.R                  & process EPA Ecotox data                          \\ \hline
-da\_epa3.R                  & reduce EPA Ecotox data                           \\ \hline
-qu\_aw.R                    & query Compendium of Pesticide Common Names       \\ \hline
-qu\_chemspider\_scrape.R    & query Chemspider data base                       \\ \hline
-qu\_eurostat\_chem\_class.R & query Eurostat data set                          \\ \hline
-qu\_pc.R                    & query PubChem data base                          \\ \hline
-qu\_pp.R                    & query Physprop data base                         \\ \hline
-qu\_worms2.R                & query WORMS data base                            \\ \hline
-qu\_gbif.R                  & query GBIF data base                             \\ \hline
-re\_merge.R                 & merge all data                                   \\ \hline
-re\_combine.R               & combine information                              \\ \hline
-re\_analyses.R              & ???                                              \\ \hline
-re\_checks\_internal.R      & ???                                              \\ \hline
-re\_final.R                 & create and write final table                     \\ \hline
-wr\_dbase.R                 & write final table to data base                   \\ \hline
-wr\_app.R                   & write final table and variables to app directory \\ \hline
-wr\_meta.R                  & ???                                              \\ \hline
-wr\_stats.R                 & ???                                              \\ \hline
-\end{tabular}
-\caption{Scripts that are used to process the data}
-\label{tab:scripts-pipeline}
-\end{table}
-
-\begin{table}[h!]
-\begin{tabular}{|l|l|}
-\hline
-Script                     & Description                                               \\ \hline
-setup.R                    & sets up the environment                                   \\ \hline
-ui.R                       & creates the graphical user interface of the application   \\ \hline
-server.R                   & handles functions and data input and output               \\ \hline
-fun\_filter.R              & function that filters the data                            \\ \hline
-fun\_aggregation.R         & function that performs the aggregation                    \\ \hline
-fun\_filagg\_plot\_ly.R    & creates interactive plot                                  \\ \hline
-\end{tabular}
-\caption{Scripts that used for the web application}
-\label{tab:scripts-app}
-\end{table}
-
 \pagebreak
@@
 \bibliographystyle{apalike}
-\bibliography{refs/references-standartox-build.bib,refs/references-standartox-app.bib}
-% multiple refs have to be written w/o space
+\bibliography{refs/references-standartox.bib} % only reference from Zotero to avoid key conflicts
+% NOTE multiple refs would have to be written w/o space and separated by <,>
 
@@
 
+\pagebreak
+\section*{Supplement}
+\subsection*{R-package list} %% TODO put in appropriate list
+\input{article/supplement/r_package_list.tex}
+\pagebreak
+
+\subsection*{Data base tables}
+\input{article/supplement/data_base_tables.tex}
+
 \end{document}

### 2019-10-18 article/00_article.tex
Updates from Overleaf

@@
 \author{Andreas Scharm{\"u}ller\textsuperscript{1{*}},
-        Verena Schreiner\textsuperscript{1},
+        Verena C. Schreiner\textsuperscript{1},
         Ralf B. Sch{\"a}fer\textsuperscript{1}}
@@
 
+\subsubsection*{Future of the project}
+Since new versions of the EPA ECOTOX data base are published on a quarterly basis the data acquisition process is automated and new versions of the Standartox are scheduled to be released regularly.
+
 \pagebreak
@@
 \subsection*{R-package list} %% TODO put in appropriate list
-\input{article/supplement/r_package_list.tex}
+\label{list:r-packages}
+%% TODO how to refernce this?
+\input{article/tables/r_package_list.tex}
 \pagebreak
@@
 \subsection*{Data base tables}
-\input{article/supplement/data_base_tables.tex}
+\input{article/tables/data_base_tables.tex}
+
+\subsection*{Application parameters}
+\input{article/tables/app_parameters.tex}
 

### 2019-10-18 article/tables/DEPRECATED_old_tables.tex
Updates from Overleaf

--- 2019-06-12 article/article.tex
@@
-\documentclass[english]{article}
-\usepackage[T1]{fontenc}
-\usepackage{natbib} % citations
-\usepackage{csvsimple}
-\usepackage{pgfplotstable}
-\usepackage{rotating} % to rotate tables
-%% tikz packages
-\usepackage{tikz}
-\usepackage{forest}
-\usetikzlibrary{trees, arrows, shapes.geometric, positioning}
-\usepackage{standalone}
-%% text packages
-\usepackage{siunitx} % fake ° symbol \ang{}
-\usepackage[super]{nth} % for 1st, 2nd etc. \nth{}
-\usepackage{hyperref}
-
-\newcommand{\standartox}{STANDARTOX}
-\newcommand{\epa}{EPA ECOTOX data base}
-\newcommand{\app}{http://139.14.20.252:3838/etox-base-shiny/}
-\newcommand{\git}{https://github.com/andreasLD/etox-base}
-\newcommand{\gitapp}{https://github.com/andreasLD/etox-base-shiny}
-\newcommand{\ecfifty}{EC\textsubscript{50}}
-
-\listfiles
-
-\begin{document}
-
-\title{Etox-Base: A tool to derive exposure endpoints from chemical test results}
-
-\author{Andreas Scharm{\"u}ller\textsuperscript{1{*}},
-        Verena Schreiner\textsuperscript{1},
-        Ralf B. Sch{\"a}fer\textsuperscript{1}}
-
-\maketitle
-\thispagestyle{fancy}
-
-1. Institute of Environmental Sciences, University of Koblenz-Landau Fortstraße 7, 76829 Landau, Germany {*}corresponding author(s):
-Andreas Scharm{\"u}ller (scharmueller@uni-landau.de)
-
-\begin{abstract}
-%%5This is a manuscript template for Data Descriptor submissions to \emph{Scientific
-%%Data} (http://www.nature.com/scientificdata). The abstract must be
-%%no longer than 170 words, and should succinctly describe the study,
-%%the assay(s) performed, the resulting data, and the reuse potential,
-%%but should not make any claims regarding new scientific findings.
-%%No references are allowed in this section.
-
-<WRITE ABSTRACT>
-
-\end{abstract}
-\pagebreak
-
-%%%%%%%%%%%%%%%%%%%%%% 01 Background
-\input{article/parts/01_background.tex}
-\pagebreak
-
-%%%%%%%%%%%%%%%%%%%%%% 02 Methods
-\input{article/parts/02_methods.tex}
-\pagebreak
-
-%%%%%%%%%%%%%%%%%%%%%% 03 Data records
-\input{article/parts/03_data_records.tex}
-\pagebreak
-
-%%%%%%%%%%%%%%%%%%%%%% 04 Technical validation
-\input{article/parts/04_technical_validation.tex}
-\pagebreak
-
-%%%%%%%%%%%%%%%%%%%%%% 05 Usage notes
-\input{article/parts/05_usage_notes.tex}
-
-\section*{Acknowledgements}
-%%Text acknowledging non-author contributors. Acknowledgements should
-%%be brief, and should not include thanks to anonymous referees and
-%%editors, or effusive comments. Grant or contribution numbers may be
-%%acknowledged. Author contributions Please describe briefly the contributions
-%%of each author to this work on a separate line. 
-%%
-%%AK did this and that. 
-%%
-%%BG did this and that and the other.
-
-\section*{Competing financial interests}
-%%A competing financial interests statement is required for all accepted
-%%papers published in \emph{Scientific Data}. If none exist simply write,
-%%``The author(s) declare no competing financial interests''.
-
-\section*{Figures Legends}
-%%Figure should be referred to using a consistent numbering scheme through
-%%the entire Data Descriptor. For initial submissions, authors may choose
-%%to supply this document as a single PDF with embedded figures, but
-%%separate figure image files must be provided for revisions and accepted
-%%manuscripts. In most cases, a Data Descriptor should not contain more
-%%than three figures, but more may be allowed when needed. We discourage
-%%the inclusion of figures in the Supplementary Information \textendash{}
-%%all key figures should be included here in the main Figure section. 
-%%
-%%Figure legends begin with a brief title sentence for the whole figure
-%%and continue with a short description of what is shown in each panel,
-%%as well as explaining any symbols used. Legend must total no more
-%%than 350 words, and may contain literature references.
-
-\section*{Tables}
-%%Tables supporting the Data Descriptor. These can provide summary information
-%%(sample numbers, demographics, etc.), but they should generally not
-%%be used to present primary data (i.e. measurements). Tables containing
-%%primary data should be submitted to an appropriate data repository. 
-%%
-%%Tables may be provided within the \LaTeX{} document or as separate
-%%files (tab-delimited text or Excel files). Legends, where needed,
-%%should be included here. Generally, a Data Descriptor should have
-%%fewer than ten Tables, but more may be allowed when needed. Tables
-%%may be of any size, but only Tables which fit onto a single printed
-%%page will be included in the PDF version of the article (up to a maximum
-%%of three).
-
-%\begin{sidewaystable}
-%    \csvautotabular[respect all]{data/meta.csv};
-%    \caption{Meta variables table};
-%    \label{table:meta-variables};
-%\end{sidewaystable}
-
 \begin{table}[h!]
@@
 \end{table}
-
-\pagebreak
-
-\bibliographystyle{apalike}
-\bibliography{refs/references-etox-base.bib,refs/references-etox-base-rpackages.bib,refs/references-etox-base-rshinypackages.bib}
-% multiple refs have to be written w/o space
-
-\section*{Data Citations}
-
-\pagebreak
-\section*{TODO}
-
-\begin{itemize}
-
-\item compare Etox-BAse output with PPDB
-
-\item plot example SSD and TU
-
-\item acute vs chronic
-
-\item OECD guidlines are not machine readable
-
-\item discuss flaws in column types (e.g some concentration entries contain '+' or '~' and are therefore not convertable to numeric) - This hampers large scale analysis greatly!
-
-\pagebreak
-\section*{MISC}
-
-\begin{figure}
-    \includestandalone[scale=0.5]{tikz/tree_source}
-    \caption{Alternative processing pipeline}
-    \label{fig:pipeline-tree}
-\end{figure}
-
-\section*{not included citations}
-
-\item Put in citation \citep{hartung_chemical_2009} ?
-which claims that REACH won't meet their assumptions and say 54 million vertebrate animals are needed for tests that would cost €9.5 billion over the next ten years.
-I.e 20 times more animals, 6 times the costs in comparison to the official estimates
-
-\item "comparison of rate of change in the production and variety of pesticides, pharmaceuticals and other chemicals over the past four decades to CO2 nutrient pollution and other global change factors" \citep{bernhardt_synthetic_2017}
-
-\item worldwide decline in entomofauna: \citep{sanchez-bayo_worldwide_2019}
-
-\end{itemize}
-
-\end{document}

### 2019-10-18 article/tables/data_base_tables.tex
Updates from Overleaf

\begin{table}[ht]
  \caption{Table of queried data bases}
  \label{tab:data-base}
  \centering
\begin{tabular}{ll}
  \hline
  Data Base & URL \\ 
  \hline
  Chemical Entities of Biological Interest (ChEBI) & https://www.ebi.ac.uk/chebi \\ 
  Chemical Identifier Resolver & https://cactus.nci.nih.gov/chemical/structure \\
  ChemSpider & http://www.chemspider.com \\
  Eurostat & https://ec.europa.eu/eurostat/home \\
  PHYSPROP Database & https://www.srcinc.com/what-we-do/environmental/scientific-databases.html \\
  PubChem & https://pubchem.ncbi.nlm.nih.gov \\ 
  Wikidata & https://www.wikidata.org/wiki/Wikidata:Main_Page \\
  Global Biodiversity Information Facility (GBIF) & https://www.gbif.org \\
  World Register of Marine Species (WoRMS) & http://marinespecies.org \\
  \hline
\end{tabular}
\end{table}

### 2019-10-18 article/parts/02_methods.tex
Updates from Overleaf

@@
 
-Standartox consists of three software parts which are explained in detail below.
+Standartox consists of three parts of software explained in detail below.
 Firstly a data acquisition and preparation processing pipeline (Fig.
-\ref{fig:pipeline}) downloads quarterly released new versions of the EPA ECOTOX data base and performs several preparation steps in R \citep{r_core_team_r_2017}, to then store the outcome in a PostgreSQL <CITE> data base.
+\ref{fig:pipeline}) downloads quarterly released new versions of the EPA ECOTOX data base and performs several preparation steps in R \citep{rcoreteam_language_2017}, to then store the outcome in a PostgreSQL9.6.1 data base.
 This constitutes the basis upon which the web application and and application programming interface (API) together with a R package are accessing the data.
@@
 From there, the data is processed, including validity checks, removal of special characters, unit harmonizations and selection of relevant variables.
-Chemical abstract service (CAS) numbers, unique chemical identifiers together with taxon names are then used to query additional information from other data bases (Table \ref{tab:data-base}) on chemicals and organisms respectively.
-Standartox itself is finally built and exported as a binary object to be used by the web application and the API.
+Chemical abstract service (CAS) numbers, unique chemical identifiers and taxa names are then used to query additional information from other data bases (Table \ref{tab:data-base}) on chemicals and organisms respectively.
+Standartox itself is finally built and exported as a compressed serialized object to be then used by the web application and the API.
 
-\subsection*{Web application and application programming interface}
+\subsection*{Web application \& Application Programming Interface (API)}
 
-Once called, both the web application and the API load the binary Standartox file into memory and allow the user to interact with it.
+Once called, both the web application and the API load the compressed serialized Standartox file into memory and allow a client to interact with it.
+The client can then call the functions stx\_filter() and stx\_aggregate() to filter and aggregate the data according to specific parameters (Table \ref{tab:app-parameters}).
 
@@
 The interactive web application was built in R, using the shiny web application framework \citep{chang_shiny_2018} which runs with the help of a shiny server \citep{HOW-TO-CITE-SHINY-SERVER} on www.standartox.uni-landau.de.
-The app itstel lies in the file ~/standartox-app/app.R.
-
-\subsubsection*{Application programming interface (API)}
-The API is built in R with the plumber-package \citep{}
+The app itself lies in the file ~/standartox-app/app.R.
 
@@
 
+\subsubsection*{Application programming interface (API)}
+The API is built by using the plumber-package \citep{trestletechnologyllc_plumber_2018} which allows for the creation of REST APIs from R.
+Four API-endpoints ('/catalog', '/filter/rds', 'aggregate' and 'meta') that can be queried from the client side were created.
+The \textit{catalog} API-endpoint returns a JavaScript Object Notation (JSON) file containing a catalog of possible filter parameters to choose from.
+The \textit{filter-rds} returns the filtered test results as a compressed serialized file \citep{klik_fst_2019}, to reduce size and allow for fast user queries.
+The \textit{aggregate} API-endpoint returns a R function to the client allowing the client the aggregation of the filtered data.
+Lastly, the \textit{meta} API-endpoint returns a JSON file with a timestamp of the request and the used Standartox version.
+The API is designed to be used with the R-package standartox and therefore uses serialization methods specific to R (rds() from the base-package and fst() from the fst-package).
+The code for the the API lies in the file ~/standartox-app/api.R.
+
 \subsubsection*{R-package}
+The R-package standartox contains two functions \textit{stx\_catalog()} and \textit{stx\_query()}.
+The former accesses the \textit{catalog} API-endpoint and returns a R-list of possible filter parameters that can be used in \textit{stx\_query()}.
+The latter returns a R list of three tables (R data.frames) containing a filtered data set, an aggregated data set and a table with the meta information retrieved from the \textit{meta} API-endpoint.
 
-httr,
-jsonlite,
-fst,
-data.table
-
-\subsubsection*{Aggregation}
-The \etoxbase{} application aggregates the test results according to chosen filters in a two step process.
+\subsubsection*{Standartox Aggregation}
+Standartox aggregates the test results according to chosen filters in a two step process.
 Firstly the filtered test results are aggregated by the CAS number, the chosen taxon and the selected test duration.
@@
 Thereof the second step calculates the minimum, the maximum, the median, the geometric mean, or the arithmetic mean as an aggregate.
-
-\subsubsection*{Future of the project}
-Since new versions of the EPA ECOTOX data base are published on a quarterly basis the data acquisition process is automated and new versions of the Standartox are scheduled to be released regularly.
-
-The code for the web-application and the application programming interface are placed here: https://github.com/andschar/standartox-app
-
-\begin{figure}
-    \includestandalone[scale=0.75]{tikz/application}
-    \caption{Application}
-    \label{fig:app}
-\end{figure}
-
-\subsection*{The R package}
-
-The code for the R-package standartox is placed here: https://github.com/andschar/standartox
 
@@
 
-The processing pipeline code is stored here https://github.com/andschar/standartox-build, the code for the web application is stored under https://github.com/andschar/standartox-app and the R-pckage code is stored in https://github.com/andschar/standartox.
-The code can freely be accessed under the MIT License <CITATION>.
-Most of the code is written in R 3.6.1 and some parts in PostgreSQL 9.6.1.
-An exhaustive list of additional R-pakage used can be found in the supplementary material <REFERENCE-TO-SUPPL>.
+Standartox's code is distributed on three Github repositories below.
+All code can freely be accessed under the MIT License <CITATION>.
+Most of the code is written in R 3.6.1 and associated packages (compare Supplement: \ref{list:r-packages}).
+Some parts in PostgreSQL 9.6.1.
 
-\begin{table}[h!]
-    \centering
-    \begin{tabular}{|l|l|l|l|}
-        \hline
-        Category        & Package       & Version   & Reference                     \\ \hline
-        processing      & RCurl         & 1.95-4.11 & \citep{lang_rcurl_2018}       \\ \hline
-        processing      & data.table    & 1.12.0    & \citep{dowle_data.table_2018} \\ \hline
-        application     & shiny         & 1.1.0     & \citep{chang_shiny_2018}      \\ \hline
-    \end{tabular}
-    \caption{List of R packages used in compiling the data.\newline{}
-    TODO: Automate version input and name generation - whole table\newline{}
-    TODO2: debug chang\_shiny\_2018 reference}
-    \label{tab:rpackages}
-\end{table}
+\begin{itemize}
 
-\section{TODO}
+\item \url{https://github.com/andschar/standartox-build} \newline
+Contains the code for the data acquisition and processing pipeline.
 
-    \item it is planed to extend this tool to other data bases if they provide new test data  (e.g. UBA ETOC)
+\item \url{https://github.com/andschar/standartox-app} \newline
+Contains the code for the web application and the API.
+
+\item \url{https://github.com/andschar/standartox} \newline
+Contains the code for the R-package.
+
+\end{itemize}

### 2019-10-18 article/parts/04_technical_validation.tex

@@
 \begin{figure}
-    \includegraphics[width=1\linewidth]{article/figures/Standartox_PPDB.png}
+    \includegraphics[width=1\linewidth]{article/figures/.png}
     \caption{
@@
 
-In order to validate \standartox we compared resulting values with values from other data bases, such as the PPDB.
+\subsection{PPDB Comparsion}
+The PPDB also offers effect values.
+
+\subsection{Unit testing}
+
+\subsection{Overall distribution of tests}
+Heatmap: don't Include for now. Too large
+
+In order to validate Standartox we compared resulting values with values from other data bases, such as the PPDB.
 Of 542 chemicals, only 16\% diverged from the PPDB values by a factor greater than 10, supporting plausibility of the \standartox results \ref{fig:standartox_ppdb_diff}.

### 2019-10-22 article/parts/04_technical_validation.tex

@@
 
+\subsection{Comparison to other data bases}
+
+In order to asses the validity of the Standartox approach to aggregate test results we compared Standartox values with effect values for the same chemicals and the same test duration from the PPDB (n=3601) \ref{fig:standartox_ppdb_diff}.
+This showed that 91.3\% of the Standartox values diverge by less than a factor of 10 from the PPDB values.
+Considering only Standartox values for which at least five of ten tests were aggregated the value increases to 91.7\% and 93.4\% respectively.
+Admittedly a few Standartox values diverge by a large margin.
+
 \begin{figure}
-    \includegraphics[width=1\linewidth]{article/figures/.png}
-    \caption{
-        Comparison between Standartox and PPDB values
-    }
+    \includegraphics[width=1\linewidth]{article/figures/gg_ppdb_stan_compare_continous.png}
+    \caption{Comparison between Standartox and PPDB values}
     \label{fig:standartox_ppdb_diff}
@@
 
-\subsection{PPDB Comparsion}
-The PPDB also offers effect values.
+\subsection{Unit Harmonisation}
 
-\subsection{Unit testing}
+In order to assure appropriate conversion and harmonisation of units we took a random sample of the 50 most occurring units and converted them manually and subsequently compared them to the conversions done by Standartox.
+
+We did this for concentration units and duration units
+
+in R/chck\_unit\_conversion.R
 
 \subsection{Overall distribution of tests}
-Heatmap: don't Include for now. Too large
+Heatmap: don't Include for now. Too large.
+Put maybe a plotly on wep app
+Mybe search for the 50 most sold chemicals and asses the number of tests
 
@@
 Of 542 chemicals, only 16\% diverged from the PPDB values by a factor greater than 10, supporting plausibility of the \standartox results \ref{fig:standartox_ppdb_diff}.
-
-<PUT COMPARISON WITH PPDB DATA HERE?>

### 2019-10-22 article/parts/05_usage_notes.tex
Updates from Overleaf

@@
 
+\subsubsection*{Standartox Filters}
+
 \subsection*{Web application}
@@
 However, both data sets can be downloaded via a download button.
+
+\subsection*{R package}
+
+The R-package standartox
+
+\input{article/scripts/standartox_example.tex}
 
@@
 The drop down menu `Download data` allows to download a filtered data set as well as an aggregated data set.
-The \etoxbase{} application can be accessed here: \app{}.
-
-\begin{table}[h!]
-\begin{tabular}{|l|l|l|l|}
-\hline
-Menu   & Group       & Subgroups              & Options \\ \hline
-Filter      & Compound    & CAS                    & input CAS \\ \hline
-Filter      & Compound    & Concentration type     & Active Ingredient; Formulation; Total \\ \hline
-Filter      & Compound    & Chemical class         & Insecticides; Herbicides; Metals \\ \hline
-Filter      & Taxon       & Taxon                  & Choose a taxon \\ \hline
-Filter      & Taxon       & Organism habitat       & freshwater; terrestrial; marine; brackish \\ \hline
-Filter      & Taxon       & Continent              & North America; Asia; Europe \\ \hline
-Filter      & Test        & Duration               & enter duration in hours \\ \hline
-Filter      & Test        & Effect groups          & MOR - Mortality; POP - Population \\ \hline
-Filter      & Test        & Endpoins               & EC50, LOEC, LOEL \\ \hline
-Filter      & Checks      & Water solubility check & yes/no \\ \hline
-Filter      & Checks      & Remove outliers        & yes/no \\ \hline
-Aggregation & Aggregation & Aggregate              & Minimum; Maximum; Median \\ \hline
-\end{tabular}
-\caption{\etoxbase GUI paramters to choose.}
-\label{tab:inputs}
-\end{table}
 
@@
 In comparison to the \etoxbase, these examples partly rely on the same data, yet not aiming to derive \ecfifty, LOEC and NOEC from it automatically.
+\subsection*{Other data bases}
+
+Ecotox tool for PENEC
+Lara's work for regulatory thresholds
+
+greater and greater demand to numerically  analyse large amount of ecotoxicological data.
+
+Other data bases are also thought of. Compare Sascha Bub's work.
+
+\subsection*{Identify chemicals to be tested}
+
+\begin{figure}
+    \includegraphics[width=1\linewidth]{article/figures/heatmap_tests_n.png}
+    \caption{30 most tested chemicals in EPA ECOTOX.}
+    \label{fig:standartox_ppdb_diff}
+\end{figure}
+
+There is a clear need in ecotoxicology for harmonised approaches to collect ecotoxicological data as other data base compilation attempts recently published show.
+
+Hitchock et al.
+\citet{hitchcock_improving_2018} argue for an increased awareness concerning newly published ecotoxicological test results to facilitate meta analyses across studies.
+Important statistical parameters, such as mean estimates, variances ans sample sizes.
+Likewise, state variables, including medium temperature, pH, mineral contents or organism age and sex should be reported (if not in the text, in an online repository).
+Standartox doesn't include such information mainly because the EPA ECOTOX data base reports only a few of these variables.
+A future research aim could be to use machine learning techniques to retrieve this information from the referenced articles.
+
+Agrochemicals contribute to a decline in biodiversity \citep{schafer_future_2019}
+contrary to EU pesticide regulation
+
+Current risk assessment or pesticide regulation consists of Stadard Toxicity Tests and Safety factors. They are not sufficient
+
+tiered framework critizized, especiall the two main assumptions:
+(1) 1st tier provides and overly protective measure
+(2) higher tiers provide more ecologically relevant predictions
+
+"Inacurate predictions of exposure and effect" \citep{schafer_future_2019}
+
+In their policy brief for a future risk assessment Schäfer et al.
+\citet{schafer_future_2019} propose to include a greater ecological context, meaning that not only single but communities need to evaluated and a landscape-oriented contemplation should be envisaged inter alia.
+Standartox partly follows such a route in making ecotoxicological data more easily available by standardizing ?????
+(1st tier) and integrating habitat and occurrence data, allowing for a more geographically detailed refinement.
+
+Hallmann et al. \citet{hallmann_declines_2014} showed a decline of insectivorous birds associated with neonicotinoid usage.
+
+Fungicides paper zitieren?
+
+Standartox follows such a route in trying to synthesize information from a plethora of studies.
 

### 2019-10-29 R/bd_standartox.R
added unit checks

@@ -11,6 +11,8 @@ q2 = readChar(fl, file.info(fl)$size)
 fl = file.path(sql, 'standartox_data2_fin.sql')
 q2_fin = readChar(fl, file.info(fl)$size)
 explanation = fread(file.path(lookupdir, 'lookup_explanation.csv'), na.strings = '')
+fl = file.path(sql, 'fun_molconv.sql') # TODO put somewhere better
+molconv = readChar(fl , file.info(fl)$size)
 
 # query -------------------------------------------------------------------
 drv = dbDriver("PostgreSQL")
@@ -19,6 +21,7 @@ con = dbConnect(drv, user = DBuser, dbname = DBetox, host = DBhost, port = DBpor
 dbSendQuery(con, "DROP SCHEMA IF EXISTS standartox CASCADE;")
 dbSendQuery(con, "CREATE SCHEMA standartox;")
 
+dbSendQuery(con, molconv)
 # dbSendQuery(con, q1) # cleaned data
 dbSendQuery(con, q2) # converted data
 dbSendQuery(con, q2_fin) # combined data

### 2019-10-29 R/chck_unit_conversions_conc.R

@@ -1,0 +1,112 @@
+# script to check concentration and duration units
+
+# setup -------------------------------------------------------------------
+source(file.path(src, 'gn_setup.R'))
+
+# query -------------------------------------------------------------------
+# NOTE query to retrieve a sample of the top50 units
+# q = '
+# WITH t1 AS (
+# SELECT conc1_unit, count(*) n
+# FROM ecotox.results
+# GROUP BY conc1_unit
+# ORDER BY n DESC
+# LIMIT 50
+# )
+# SELECT DISTINCT ON (conc1_unit) result_id, conc1_mean, conc1_unit
+# FROM ecotox.results
+# RIGHT JOIN t1 USING (conc1_unit)
+# ORDER BY conc1_unit;
+# '
+
+# sample ------------------------------------------------------------------
+# 50 most occurring result_id s
+ids = c(2227707L, 2144511L, 1177875L, 1012510L, 2267424L, 1059003L, 
+        2109679L, 2459421L, 1145750L, 1165988L, 2262868L, 2179588L, 2117263L, 
+        1172177L, 798810L, 2453697L, 2345310L, 1128312L, 1039743L, 775822L, 
+        206500L, 1042482L, 2295969L, 2371433L, 1117335L, 1228663L, 1065969L, 
+        168785L, 1186575L, 1138853L, 1153817L, 244062L, 2074585L, 2226852L, 
+        239960L, 2126050L, 2287429L, 1110015L, 2060384L, 1190262L, 2333451L, 
+        1189176L, 1050168L, 88612L, 2248365L, 1219977L, 55416L, 1219748L, 
+        2229405L, 2348737L, 756609L)
+
+q = paste0("SELECT stx.result_id, stx.casnr, che.molecularweight, stx.conc1_mean, stx.conc1_unit, stx.conc1_mean2, stx.conc1_unit2
+            FROM standartox.tests stx
+            LEFT JOIN standartox.chemicals che ON stx.casnr = che.casnr
+            WHERE stx.result_id IN (", paste0(ids, collapse = ','), ")
+              AND conc1_mean != 'NR';")
+
[…]

### 2019-10-29 R/chck_unit_conversions_duration.R

@@ -1,0 +1,67 @@
+# script to check concentration and duration units
+
+# setup -------------------------------------------------------------------
+source(file.path(src, 'gn_setup.R'))
+
+# query -------------------------------------------------------------------
+# NOTE query to retrieve a sample of the top50 units
+# q = '
+# WITH t1 AS (
+#   SELECT obs_duration_unit, count(*) n
+#   FROM ecotox.results
+#   GROUP BY obs_duration_unit
+#   HAVING obs_duration_unit IN ('s', 'mi', 'h', 'd', 'wk', 'mo')
+#   ORDER BY n DESC
+# )
+# SELECT DISTINCT ON (obs_duration_unit) result_id, obs_duration_mean, obs_duration_unit
+# FROM ecotox.results
+# RIGHT JOIN t1 USING (obs_duration_unit)
+# WHERE obs_duration_unit IN ('s', 'mi', 'h', 'd', 'wk', 'mo') AND obs_duration_mean NOT IN ('NR')
+# ORDER BY obs_duration_unit;
+# '
+
+# sample ------------------------------------------------------------------
+# 50 most occurring result_id s
+ids = c(2231908L, 2038088L, 717310L, 2369905L, 756251L, 2318654L)
+
+q = paste0("SELECT stx.result_id, stx.casnr, stx.obs_duration_mean, stx.obs_duration_unit, stx.obs_duration_mean2, stx.obs_duration_unit2
+            FROM standartox.tests stx
+            LEFT JOIN standartox.chemicals che ON stx.casnr = che.casnr
+            WHERE stx.result_id IN (", paste0(ids, collapse = ','), ")
+              AND conc1_mean != 'NR';")
+
+dur = read_query(user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
+                 query = q)
+
+# manual comparison -------------------------------------------------------
+dur[ result_id == 717310L, `:=` (obs_duration_mean3 = 3.500007, obs_duration_unit3 = 'h') ] # * 0,0166667
+dur[ result_id == 756251L, `:=` (obs_duration_mean3 = 0.0027778, obs_duration_unit3 = 'h') ] # * 0.00027778
+dur[ result_id == 2038088L, `:=` (obs_duration_mean3 = 48, obs_duration_unit3 = 'h') ]
+dur[ result_id == 2231908L, `:=` (obs_duration_mean3 = 24, obs_duration_unit3 = 'h') ]
[…]

### 2019-10-29 R/look_concentration_units.R

@@ -4,30 +4,12 @@
 source(file.path(src, 'gn_setup.R'))
 
 # data --------------------------------------------------------------------
-drv = dbDriver("PostgreSQL")
-con = dbConnect(
-  drv,
-  user = DBuser,
-  dbname = DBetox,
-  host = DBhost,
-  port = DBport,
-  password = DBpassword
-)
-
-unit = dbGetQuery(
-  con,
-  "SELECT conc1_unit, count(conc1_unit) AS n
-   FROM ecotox.results
-   GROUP BY conc1_unit
-   ORDER BY n DESC"
-)
-setDT(unit)
-unit[conc1_unit == '', conc1_unit := NA]
-unit = unit[!is.na(conc1_unit)]
-
-dbDisconnect(con)
-dbUnloadDriver(drv)
-
+q = "SELECT conc1_unit, count(conc1_unit) AS n
+     FROM ecotox.results
+     GROUP BY conc1_unit
+     ORDER BY n DESC"
+unit = read_query(user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
+                  query = q)
 ## lookup
 look_unit = fread(file.path(lookupdir, 'lookup_concentration_unit.csv'), na.strings = '')
 look_str = paste0(look_unit$unit, collapse = '|')
@@ -53,6 +35,7 @@ info = c(
   'wetbdwt',
   'soil',
   'ai\\s',
+  'AI\\s',
   acid_equivalent = 'ae',
   water_soluble_fraction = 'wsf',
   'media',
@@ -146,16 +129,16 @@ units[, conc1_unit_clean := gsub('bq', 'Bq', conc1_unit_clean)]
 # classification ----------------------------------------------------------
 u1 = c(
   fraction = 'ppb',
-  mass = 'g',
+  mass = 'ug',
   percent = '%',
   volume = 'ml',
   length = 'cm'
 )
 u2 = c(
-  `mass/mass` = 'g/g',
+  `mass/mass` = 'mg/kg',
   `mass/volume` = 'ug/l',
   `mass/area` = 'g/m2',
-  `volume/volume` = 'ml/l',
+  `volume/volume` = 'ul/l',
   `volume/area` = 'ml/m2',
   `volume/mass` = 'ml/g',
   `mass/length` = 'g/cm',
@@ -202,9 +185,7 @@ units[ conv == 'TRUE', conv := 'yes' ]
 units[ conv == 'FALSE', conv := 'no' ]
 
 # mol ---------------------------------------------------------------------
[…]

### 2019-10-29 R/look_duration_units.R

@@ -4,39 +4,24 @@
 source(file.path(src, 'gn_setup.R'))
 
 # data --------------------------------------------------------------------
-drv = dbDriver("PostgreSQL")
-con = dbConnect(
-  drv,
-  user = DBuser,
-  dbname = DBetox,
-  host = DBhost,
-  port = DBport,
-  password = DBpassword
-)
-
-unit_test = data.table(dbGetQuery(
-  con,
-  "SELECT obs_duration_unit, count(obs_duration_unit) AS n
-   FROM ecotox.results
-   GROUP BY obs_duration_unit
-   ORDER BY n DESC"
-))
-unit_study = data.table(dbGetQuery(
-  con,
-  "SELECT study_duration_unit, count(study_duration_unit) AS n
-   FROM ecotox.tests
-   GROUP BY study_duration_unit
-   ORDER BY n DESC"
-))
-
-dbDisconnect(con)
-dbUnloadDriver(drv)
+q1 = "SELECT obs_duration_unit, count(obs_duration_unit) AS n
+      FROM ecotox.results
+      GROUP BY obs_duration_unit
+      ORDER BY n DESC"
+unit_test = read_query(user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
+                       query = q1)
+q2 = "SELECT study_duration_unit, count(study_duration_unit) AS n
+      FROM ecotox.tests
+      GROUP BY study_duration_unit
+      ORDER BY n DESC"
+unit_study = read_query(user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
+                        query = q2)
+## lookup csv
+look_dur = fread(file.path(lookupdir, 'lookup_duration.csv'), na.strings = '')
 
 # prepare -----------------------------------------------------------------
 unit = merge(unit_test, unit_study, by.x = 'obs_duration_unit', by.y = 'study_duration_unit',
              all = TRUE, suffixes = c('_test', '_study'))
-## lookup csv
-look_dur = fread(file.path(lookupdir, 'lookup_duration.csv'), na.strings = '')
 units = merge(unit, look_dur, by.x = 'obs_duration_unit', by.y = 'unit',
               all = TRUE)
 units = units[ !is.na(obs_duration_unit) ]

### 2019-10-29 run_build.R

@@ -103,6 +103,9 @@ if (build_norman) {
   source(file.path(src, 'cpy_norman.R'), max.deparse.length = mdl)
 }
 
+# check scripts -----------------------------------------------------------
+source(file.path(src, 'chck_unit_conversions.R'))
+
 # backup ------------------------------------------------------------------
 if (general) {
   source(file.path(src, 'gn_backup.R'), max.deparse.length = mdl)

### 2019-10-29 article/parts/04_technical_validation.tex
Updates from Overleaf

@@
 \subsection{Comparison to other data bases}
+
+We sampled an entry for each of the 50 most occurring concentration units and compared the Standartox conversions to manually calculated conversions for these, to guarantee the correctness of unit conversions.
+The same was done for six samples of duration units (See \ref{tab:chck-unit}
+
+To guarantee correctness of unit conversions we sampled an entry for each of the 50 most frequent concentration units. and the six most frequent duration units, respectively.
+
+Out of the 1229 concentration units in the EPA ECOTOX data base, Standartox retains only those (n = X) whih are convertible to one of the following units: ug/l, ul/l, ppb, g/m2, mg/kg, \%, resulting in 873473 out of 952625 test results.
+Likewise, out of the 125 test duration units in the EPA ECOTOX, Standartox retains only those (n = X) which are accurately convertible to hours, leading to 905110 out of 952625 test results (see table \ref{tab:units-duration}).
+% TODO
+
+The here described Standartox version uses the EPA ECOTOX data base released on the 12.09.2019.
 

### 2019-10-29 article/parts/05_usage_notes.tex
Updates from Overleaf

@@
 
-\subsubsection*{Standartox Filters}
-
 \subsection*{Web application}
 
-Users can define several input parameters \ref{tab:inputs} in the application's graphical user interface (GUI), handled in the ui.R script in order to adjust the aggregation according specific requirements.
-In doing so users execute R functions (fun\_filter.R, fun\_aggregation.R, fun\_filagg\_plot\_ly.R) loaded in the server.R script, which produce a 'filtered' and an 'aggregated' data set as well as an interactive overview plot.
-Only the 'aggregated' data set is presented in the GUI.
-However, both data sets can be downloaded via a download button.
+In Standartox web application (app) users can define several input parameters \ref{tab:inputs} through the app's graphical user interface (GUI) in order to filter the data specifically and perform the aggregation.
+The aggregated result is then shown as a table as well as an overview plot and can then be downloaded as a 'filtered' and an 'aggregated' data set \ref{fig:standartox_app}.
+
+\begin{figure}
+    \includegraphics[width=1\linewidth]{article/figures/screenshot_standartox_app.png}
+    \caption{Standartox web application}
+    \label{fig:standartox_app}
+\end{figure}
 
@@
 
-The R-package standartox
+The R-package standartox consists of the two functions stx\_catalog() and stx\_query().
+The former allows the retrieval of a catalog of possible parameters that can be used as an input for the latter.
+stx\_query() fetches the toxicity values from the data base.
+It returns a R list object following the app containing a 'filtered' and 'aggregated' R data.frame object.
+Additionally ...................
 
@@
 
+The units are denoted as ppm or ppb which comprises ug/l and mg/l
+
+List the units which are kept.
+
+Thereby 'food', soil, ai and other information are lost but these are coded in other variables anyway. And they don't belong in a unit variable. Characteristic for ecotoxicology
+
+units such as per day (mg/kg/day) are generally excluded.
+
 Standartox follows such a route in trying to synthesize information from a plethora of studies.

### 2019-10-31 article/parts/02_methods.tex

@@
 Standartox consists of three parts of software explained in detail below.
-Firstly a data acquisition and preparation processing pipeline (Fig.
-\ref{fig:pipeline}) downloads quarterly released new versions of the EPA ECOTOX data base and performs several preparation steps in R \citep{rcoreteam_language_2017}, to then store the outcome in a PostgreSQL9.6.1 data base.
+Firstly a data acquisition and preparation processing pipeline downloads quarterly released new versions of the EPA ECOTOX data base and performs several preparation steps in R \citep{rcoreteam_language_2017}.
+The output is then stored in a PostgreSQL9.6.1 data base.
 This constitutes the basis upon which the web application and and application programming interface (API) together with a R package are accessing the data.
 
-\subsection*{Data acquisition \& preparation}
+\subsection*{Data acquisition pipe line}
+\subsubsection*{Building the data base}
+In order to build Standartox the EPA ECOTOX data base is downloaded and subsequently built into a PostgreSQL data base.
+This is handled by the script run\_build.R.
+Therein, the EPA ECOTOX data base is downloaded (bd\_epa\_download.R), built into a PostgreSQL data base (bd\_epa\_postgres.R) and SQL function for further processing are added (bd\_sql\_functions.R) to it.
+Located errors are fixed and updated in the data base (bd\_epa\_errata.R).
+Subsequently lookup tables for filters and conversions are uploaded (bd\_epa\_lookup.R).
+Lastly, a meta data table is created in the data base, comprising the version number of the EPA ECOTOX data base (bd\_epa\_meta.R).
+\subsubsection*{Querying additional identifiers}
+In a next step Chemical Abstracts Service (CAS) identifiers are used to query identifier data bases, including the Chemical Identifier Resolver (CIR) service <CITATION> (id\_pc\_cid\_dwld.R) and the Pubchem data base <CITATION>.
+Chemical identifiers (CAS, InchI) and taxonomic names are then used to query chemical and taxonomic specific data bases (Table \ref{tab:data-base-additional}) to add information to Standartox.
+The data is aggregated and unique values are derived.
+PUT LOOKUPTABLES HERE.
+From here on the Standartox data set is compiled and exported.
+Additionally a catalog object listing possible categorical values and continous value ranges, respectively is created.
+To guarantee the accuracy of the data check scripts are run at the end.
+Single processing scripts are listed in figure \ref{fig:pipeline-tree}.
 
-In order to build Standartox the EPA ECOTOX data is downloaded and subsequently built into a PostgreSQL data base.
-From there, the data is processed, including validity checks, removal of special characters, unit harmonizations and selection of relevant variables.
+CONTINUE HERE!!!!!!!!!!
+
+which first downloads the data, builds it into a PostgresSQL data base SQL functions are then
+
+From there, the data is processed including, removal of special characters, unit harmonizations and selection of relevant variables.
+
 Chemical abstract service (CAS) numbers, unique chemical identifiers and taxa names are then used to query additional information from other data bases (Table \ref{tab:data-base}) on chemicals and organisms respectively.
-Standartox itself is finally built and exported as a compressed serialized object to be then used by the web application and the API.
+The data for Standartox is finally compiled and exported as a compressed serialized object to then be used by the web application and the API.
+
+\subsubsection*{Data aggregation}
+Standartox uses the geometric mean to aggregate the ecotoxicological test results.
+
+\textbf{OLD: Standartox aggregates the test results according to chosen filters in a two step process.
+Firstly the filtered test results are aggregated by the CAS number, the chosen taxon and the selected test duration.
+Secondly, the returned data is then aggregated by the CAS number.
+The former can't be influenced by the user and calculates either the minimum or the median depending on the amount of results to aggregate (n <= 2: minimum, if n > 2: median).
+Thereof the second step calculates the minimum, the maximum, the median, the geometric mean, or the arithmetic mean as an aggregate.}
 
@@
 
-Once called, both the web application and the API load the compressed serialized Standartox file into memory and allow a client to interact with it.
+The web application and the API load the compressed serialized Standartox data into memory and allow a client to interact with it.
 The client can then call the functions stx\_filter() and stx\_aggregate() to filter and aggregate the data according to specific parameters (Table \ref{tab:app-parameters}).
@@
 
+\begin{figure}
+    \includestandalone[scale=0.5]{article/tikz/pipeline-organigram}
+    %\input[scale = 0.5]{article/tikz/pipeline-organigram.tex}
+    \caption{Organigram of Standartox.}
+    \label{fig:stx-organigram}
+\end{figure}
+
 \subsubsection*{R-package}
@@
 The latter returns a R list of three tables (R data.frames) containing a filtered data set, an aggregated data set and a table with the meta information retrieved from the \textit{meta} API-endpoint.
-
-\subsubsection*{Standartox Aggregation}
-Standartox aggregates the test results according to chosen filters in a two step process.
-Firstly the filtered test results are aggregated by the CAS number, the chosen taxon and the selected test duration.
-Secondly, the returned data is then aggregated by the CAS number.
-The former can't be influenced by the user and calculates either the minimum or the median depending on the amount of results to aggregate (n <= 2: minimum, if n > 2: median).
-Thereof the second step calculates the minimum, the maximum, the median, the geometric mean, or the arithmetic mean as an aggregate.
 

### 2019-10-31 article/tikz/pipeline-organigram.tex

\documentclass[border = 1.5cm]{standalone}

%%%% packages
\usepackage{tikz}
\usetikzlibrary{trees,
                arrows,
                shapes.geometric,
                positioning,
                calc,
                backgrounds,
                fit}
%% document
\begin{document}
%% layers
\pgfdeclarelayer{bg1}    % declare background layer
\pgfdeclarelayer{bg2}
\pgfdeclarelayer{bg3}
\pgfsetlayers{bg3, bg2, bg1, main}  % set the order of the layers (main is the standard layer)
%% styles
\tikzset{
	font = {\fontsize{13pt}{12}\sffamily},
	children/.style = {anchor = west},
	ecotox/.style = {rectangle, minimum width = 4cm, minimum height = 3cm, text centered, draw = black, fill = blue!10},
	data/.style = {rectangle, minimum height = 3cm, minimum width = 7cm, rounded corners, draw = black},
	app/.style = {circle, minimum height = 3cm, fill = red!40},
	api/.style = {circle, minimum height = 3cm, fill = red!40},
	rpackage/.style = {circle, minimum height = 3cm, fill = red!40},
	ids/.style = {diamond, minimum width = 3cm, minimum height = 2cm, text centered, draw = black, fill = green!30},
	db/.style = {rectangle, text centered, draw = black, rounded corners, minimum width = 3cm, minimum height = 1.25cm},
	arrow1/.style = {->, > = stealth, draw = black, color = gray!80, line width = 2.5pt},
	arrow2/.style = {->, > = latex, draw = black, line width=10pt},
	background/.style = {rectangle, rounded corners, draw = black, fill=gray!0.01}
}

\begin{tikzpicture}
%% nodes
% pipeline
\node (epa-ecotox) [ecotox, align = center] at (0cm,0cm) {EPA ECOTOX\\data base\\(cleaned)};
\node (cas) [ids, below of = epa-ecotox, xshift = 6cm, yshift = 3cm] {CAS};
\node (taxa) [ids, below of = cas, xshift = 0cm, yshift = -2.5cm] {Taxa};
\node (db-chemical) [db, right of = cas, xshift = 4cm] {Chemical\\data bases};
\node (db-habitat) [db, right of = taxa, xshift = 4cm, yshift = 1cm] {Habitat\\data bases};
\node (db-region) [db, right of = taxa, xshift = 4cm, yshift = -1cm] {Region\\data bases};
% standartox data
\node (stx-data) [data, below of = bg1-pipeline, xshift = 5.5cm, yshift = -6cm] {Standatox\\data set};
% user interfaces
\node (stx-app) [app, below of = stx-data, yshift = -3.5cm, xshift = -2cm] {Web\\application};
\node (stx-api) [api, below of = stx-data, yshift = -3.5cm, xshift = 2cm] {API};
\node (stx-rpackage) [rpackage, below of = stx-api, yshift = -2cm, xshift = 2cm] {R-package};
%% backgrounds
\begin{pgfonlayer}{bg1}
    \node (bg1-pipeline) [background, minimum height = 9cm, minimum width = 16cm, above of = epa-ecotox, xshift = 5.5cm, yshift = -1cm];
    \node (bg1-user) [background, below of = bg1-pipeline, minimum height = 7cm, minimum width = 12cm, yshift = -12cm];
\end{pgfonlayer}
%% arrows
% small
\draw [arrow1] (epa-ecotox.east) -- (cas.west);
\draw [arrow1] (epa-ecotox.east) -- (taxa.west);
\draw [arrow1] (cas.east)-- (db-chemical.west);
\draw [arrow1] (taxa.east) -- (db-habitat.west);
\draw [arrow1] (taxa.east) -- (db-region.west);
\begin{pgfonlayer}{bg1}
    \draw [arrow2] (bg1-pipeline.south) -- (stx-data.north);
    \draw [arrow2] (stx-data.south) -- (stx-app.north);
    \draw [arrow2] (stx-data.south) -- (stx-api.north);
\end{pgfonlayer}
\draw [arrow1] (stx-api) -- (stx-rpackage);
% large

\end{tikzpicture}
\end{document}

### 2019-11-05 article/parts/02_methods.tex
Updates from Overleaf

--- 2019-10-18 article/parts/02_methods.tex
@@
-\section*{Methods}
-
 %%The Methods should include detailed text describing any steps or procedures 
@@
 %%There is no limit to the length of the Methods section.
+\section*{Methods}
 
 Standartox consists of three parts of software explained in detail below.
-Firstly a data acquisition and preparation processing pipeline (Fig.
-\ref{fig:pipeline}) downloads quarterly released new versions of the EPA ECOTOX data base and performs several preparation steps in R \citep{rcoreteam_language_2017}, to then store the outcome in a PostgreSQL9.6.1 data base.
-This constitutes the basis upon which the web application and and application programming interface (API) together with a R package are accessing the data.
+Firstly a data acquisition and preparation processing pipeline which downloads quarterly released new versions of the EPA ECOTOX, upon which Standartox is built.
+Secondly, Standartox is accessible via a web application (APP) and an application programming interface (API).
+Lastly, the R package standartox constitutes an easily usable way of accessing the tool via the API.
+All the software is written in R \citep{rcoreteam_language_2017} and SQL scripts.
 
-\subsection*{Data acquisition \& preparation}
+\subsection*{Data acquisition and preparation pipe line}
+The EPA ECOTOX, which is released on a quarterly basis (current release: 12.09.2019) is checked for new releases in monthly intervals and downloaded as soon as a new release is published.
+Subsequently the EPA ECOTOX is built into a PostgreSQL data base and SQL functions for further processing are added and known errors are corrected.
+Lookup tables for filtering and converting the data are uploaded.
+Thereafter, a meta data table is created in the data base, comprising the version number of the EPA ECOTOX.
+In a next step Chemical Abstracts Service (CAS) identifiers are used to query identifier data bases, including the Chemical Identifier Resolver (CIR) service <CITATION> and the Pubchem data base <CITATION>.
+Chemical identifiers (CAS, InchI) and taxonomic names are then used to query chemical and taxonomic specific data bases (Table \ref{tab:data-base-additional}) to add information to the local data base.
+The data is aggregated and unique values are derived.
+<PUT LOOKUPTABLES HERE>.
+From here on the Standartox data set is compiled, whereupon concentration and duration units are converted (see tables \ref{tab:conv-concentration}, \ref{tab:conv-duration}) in order to harmonize data.
+A catalog object, listing unique values for categorical variables and ranges for continuous variables, respectively is created.
+To guarantee the accuracy of the data check scripts are run at the end.
+Single processing scripts are listed in figure \ref{fig:pipeline-tree}.
+The final Standartox table together with the catalog are exported.
+Both the APP and the API access these objects.
 
-In order to build Standartox the EPA ECOTOX data is downloaded and subsequently built into a PostgreSQL data base.
-From there, the data is processed, including validity checks, removal of special characters, unit harmonizations and selection of relevant variables.
-Chemical abstract service (CAS) numbers, unique chemical identifiers and taxa names are then used to query additional information from other data bases (Table \ref{tab:data-base}) on chemicals and organisms respectively.
-Standartox itself is finally built and exported as a compressed serialized object to be then used by the web application and the API.
+\subsection*{Web application (APP) and Application Programming Interface (API)}
 
-\subsection*{Web application \& Application Programming Interface (API)}
-
-Once called, both the web application and the API load the compressed serialized Standartox file into memory and allow a client to interact with it.
+The web application and the API load the compressed serialized Standartox data into memory and allow a client to interact with it.
 The client can then call the functions stx\_filter() and stx\_aggregate() to filter and aggregate the data according to specific parameters (Table \ref{tab:app-parameters}).
 
-\subsubsection*{Web application}
-The interactive web application was built in R, using the shiny web application framework \citep{chang_shiny_2018} which runs with the help of a shiny server \citep{HOW-TO-CITE-SHINY-SERVER} on www.standartox.uni-landau.de.
-The app itself lies in the file ~/standartox-app/app.R.
+\subsubsection*{Web application (APP)}
+The interactive APP was built in R, using the shiny web application framework \citep{chang_shiny_2018} which runs with the help of a shiny server \citep{HOW-TO-CITE-SHINY-SERVER} on www.standartox.uni-landau.de.
+It is set of R functions that filter and aggregate the data according to a user's inputs \ref{fig:app}.
 
-The application itself is set of R functions that filter and aggregate the data according to a user's inputs \ref{fig:app}.
 Finally in the web application, a `filtered` and an `aggregated` data set as well as an interactive plot, for data exploration are returned.
@@
 \end{itemize}
+
+\textbf{OLD: Standartox aggregates the test results according to chosen filters in a two step process.
+Firstly the filtered test results are aggregated by the CAS number, the chosen taxon and the selected test duration.
+Secondly, the returned data is then aggregated by the CAS number.
+The former can't be influenced by the user and calculates either the minimum or the median depending on the amount of results to aggregate (n <= 2: minimum, if n > 2: median).
+Thereof the second step calculates the minimum, the maximum, the median, the geometric mean, or the arithmetic mean as an aggregate.}

### 2019-11-05 article/parts/06_supplement.tex

\section*{Supplement}
\subsection*{R-package list} %% TODO put in appropriate list
\label{list:r-packages}
%% TODO how to refernce this?
\input{article/tables/r_package_list.tex}
\pagebreak

\subsection*{Data base tables}
\input{article/tables/data_base_tables.tex}

\subsection*{Unit conversion}

\begin{table}
    \csvautotabular[respect all]{lookup/lookup_concentration_unit.csv}
    \caption{Concentration unit conversion}
    \label{tab:conv-concentration}
\end{table}

\begin{table}
    \csvautotabular[respect all]{lookup/lookup_duration.csv}
    \caption{Duration unit conversion}
    \label{tab:conv-duration}
\end{table}

\subsection*{Application parameters}
\input{article/tables/app_parameters.tex}

\input{article/tables/chck-units-concentration.tex}

\input{article/tables/chck-units-duration.tex}

### 2019-11-06 article/tables/look-concentration.units.tex
added latex table

\begin{tabular}{l|l|l|l|r|l|l|l|l|l}
\hline
unit & prefix & word & si & multiplier & conv & unit\_conv & type & remove & notes\\
\hline
fg & femto & gram & yes & 0.000000e+00 & yes & g & mass & no & \\
\hline
pg & pico & gram & yes & 0.000000e+00 & yes & g & mass & no & \\
\hline
ng & nano & gram & yes & 0.000000e+00 & yes & g & mass & no & \\
\hline
ug & micro & gram & yes & 1.000000e-06 & yes & g & mass & no & \\
\hline
mg & milli & gram & yes & 1.000000e-03 & yes & g & mass & no & \\
\hline
g &  & gram & yes & 1.000000e+00 & yes & g & mass & no & \\
\hline
kg & kilo & gram & yes & 1.000000e+03 & yes & g & mass & no & \\
\hline
t & mega & gram & no & 1.000000e+06 & yes & g & mass & no & \\
\hline
tons & mega & gram & no & 1.000000e+06 & yes & g & mass & no & \\
\hline
pl & pico & liter & yes & 0.000000e+00 & yes & l & volume & no & \\
\hline
nl & nano & liter & yes & 0.000000e+00 & yes & l & volume & no & \\
\hline
ul & micro & liter & yes & 1.000000e-06 & yes & l & volume & no & \\
\hline
ml & milli & liter & yes & 1.000000e-03 & yes & l & volume & no & \\
\hline
l &  & liter & yes & 1.000000e+00 & yes & l & volume & no & \\
\hline
hl & hecto & liter & yes & 1.000000e+02 & yes & l & volume & no & \\
\hline
bu &  & bushel & no & 3.523907e+01 & yes & l & volume & no & \\
\hline
ppq & quadrillion & partsper & no & 1.000000e-06 & yes & ppb & fraction & no & \\
\hline
ppt & trillion & partsper & no & 1.000000e-03 & yes & ppb & fraction & no & \\
\hline
ppb & billion & partsper & no & 1.000000e+00 & yes & ppb & fraction & no & \\
\hline
ppm & million & partsper & no & 1.000000e+03 & yes & ppb & fraction & no & \\
\hline
no &  & number & no & NA & no &  & amount & no & \\
\hline
amol & atto & mol & yes & 0.000000e+00 & yes & mol & mol & no & \\
\hline
fmol & femto & mol & yes & 0.000000e+00 & yes & mol & mol & no & \\
\hline
pmol & pico & mol & yes & 0.000000e+00 & yes & mol & mol & no & \\
\hline
nmol & nano & mol & yes & 0.000000e+00 & yes & mol & mol & no & \\
\hline
umol & micro & mol & yes & 1.000000e-06 & yes & mol & mol & no & \\
\hline
mmol & milli & mol & yes & 1.000000e-03 & yes & mol & mol & no & \\
\hline
mol & one & mol & yes & 1.000000e+00 & yes & mol & mol & no & \\
\hline
kmol & kilo & mol & yes & 1.000000e+03 & yes & mol & mol & no & \\
\hline
nM & nano & mol & yes & 0.000000e+00 & yes & mol & mol & no & \\
\hline
uM & micro & mol & yes & 1.000000e-06 & yes & mol & mol & no & \\
\hline
mM & milli & mol & yes & 1.000000e-03 & yes & mol & mol & no & \\
\hline
M & one & mol & yes & 1.000000e+00 & yes & mol & mol & no & \\
\hline
yd &  & yard & no & 9.144000e-01 & yes & m & length & no & \\
\hline
ft &  & foot & no & 3.048000e-01 & yes & m & length & no & \\
\hline
cm & centi & meter & yes & 1.000000e-02 & yes & m & length & no & \\
\hline
mm & milli & meter & yes & NA & yes &  & length &  & \\
\hline
m &  & meter & yes & NA &  &  & length &  & \\
\hline
neq & nano & equivalent & no & NA & no &  & noscience & yes & \\
\hline
meq & milli & equivalent & no & NA & no &  & noscience & yes & \\
\hline
ueq & micro & equivalent & no & NA & no &  & noscience & yes & \\
\hline
mm2 & milli &  & yes & 1.000000e-03 & yes & m2 & area & no & \\
\hline
cm2 & centi &  & yes & 1.000000e-02 & yes & m2 & area & no & \\
\hline
m2 & one &  & yes & 1.000000e+00 & yes & m2 & area & no & \\
\hline
ha &  & hectare & yes & 1.000000e+04 & yes & m2 & area & no & \\
\hline
acre &  &  & no & 4.050000e+03 & yes & m2 & area & no & \\
\hline
ac &  &  & no & 4.050000e+03 & yes & m2 & area & no & \\
\hline
mi2 &  & squaremile & no & NA & yes & m2 & area & no & \\
\hline
km2 & kilo & squaremeter & yes & 1.000000e+06 & yes & m2 & area & no & \\
\hline
ft2 &  & squarefeet & no & 9.290000e-02 & yes & m2 & area & no & \\
\hline
feddan &  & feddan & no & 4.200000e+03 & yes & m2 & area & no & \\
\hline
dn(Cyprus) &  & donum & no & 3.300000e+02 & yes & m2 & area & no & \\
\hline
\% &  & percent & no & 1.000000e+07 & yes & ppb & percentage & no & \\
\hline
‰ &  & permille & no & 1.000000e+06 & yes & ppb & percentage & no & \\
\hline
d &  & day & no & 2.400000e+01 & yes & h & time & no & \\
\hline
h &  & hour & no & 1.000000e+00 & yes & h & time &  & \\
\hline
mi &  & minute & no & NA &  &  & time &  & \\
\hline
wk &  & week & no & NA &  &  & time &  & \\
\hline
yr &  & year & no & NA & no &  & time &  & \\
\hline
ft3 &  & squarefeet & no & NA &  &  & volume & no & \\
\hline
mm3 & milli & cubic & no & 0.000000e+00 & yes & m3 & volume & no & \\
\hline
cm3 & centi & cubic & no & 1.000000e-06 & yes & m3 & volume & no & \\
\hline
dm3 & deci & cubic & no & 1.000000e-03 & yes & m3 & volume & no & \\
\hline
m3 &  & cubic & no & 1.000000e+00 & yes & m3 & volume & no & \\
\hline
fl oz &  &  & no & 2.960000e-02 & yes & l & volume & no & 28.4ml in the imperial system – 29.6ml in the US\\
\hline
pt &  & pint & no & NA & no &  & volume & yes & three different pints\\
\hline
oz &  & ounce & no & 2.834000e+01 & yes & g & volume & no & \\
\hline
gal &  &  & no & 4.546090e+00 & yes & l & volume & no & \\
\hline
qt &  & quarts & no & NA &  &  & volume &  & \\
\hline
lb &  & pound & no & 4.535920e+02 & yes & g & mass & no & \\
\hline
psu &  & practical salinity unit & no & NA & no &  & salinity & no & \\
\hline
-- &  & no data & no & NA & no &  & nodata & yes & \\
\hline
nr &  & no data & no & NA & no &  & nodata & yes & \\
\hline
ai &  & active ingredient & no & NA & no &  & noscience & yes & \\
\hline
eu &  & experimental unit & no & 1.000000e+00 & no & eu & experimentalunit & yes & \\
\hline
kbq & kilo & becquerel & no & NA &  &  & radioactivity &  & \\
\hline
bq &  & becquerel & no & NA & no &  & radioactivity & no & \\
\hline
mbq & milli & becquerel &  & NA &  &  & radioactivity &  & \\
\hline
ubq & micro & becquerel & no & NA & no &  & radioactivity & yes & \\
\hline
ci &  & curie & no & 0.000000e+00 & yes & Bq & radioactivity & no & \\
\hline
mci & milli & curie &  & NA &  &  & radioactivity &  & \\
\hline
uci & micro & curie & no & NA &  &  & radioactivity &  & \\
\hline
nci & nano & curie &  & NA &  &  & radioactivity &  & \\
\hline
icu &  & international chick units & no & NA & no &  & noscience & yes & \\
\hline
iu &  & international units & no & NA & no &  & noscience & yes & \\
\hline
fibers &  & fibers & no & NA & no &  & noscience & yes & \\
\hline
kj & kilo & joule & no & NA & no &  & energy & yes & \\
\hline
n &  & normal & no & NA & no &  & noscience & yes & \\
\hline
ra &  & ratio & no & NA & no &  & ratio & yes & \\
\hline
unit &  & unit & no & NA & no &  & noscience & yes & \\
\hline
org &  & organism & no & 1.000000e+00 & no & org & organism & yes & \\
\hline
v &  & volume & no & NA & no &  & volume & yes & \\
\hline
\% v &  & percentage volume &  & NA &  &  & percentage &  & \\
\hline
cwt &  & hundred weight & no & NA & no &  & noscience & yes & \\
\hline
w &  & weight & no & NA & no &  & noscience & yes & \\
\hline
\% w &  & percentage weight &  & NA &  &  & percentage &  & \\
\hline
in dia &  & inch diameter & no & NA & no &  & noscience & yes & \\
\hline
egg &  & egg & no & NA & no &  & noscience & yes & \\
\hline
pellets &  & pellets & no & NA & no &  & noscience & yes & \\
\hline
bee &  & bee & no & NA & no &  & noscience & yes & \\
\hline
fish &  & fish & no & NA & no &  & noscience & yes & \\
\hline
u &  & units & no & NA & no &  & noscience & yes & \\
\hline
dpm &  & disintegrations per minute & no & NA & no &  & radioactivity & yes & \\
\hline
cwt &  & hundredweight & no & NA & no &  & noscience & yes & \\
\hline
sd &  & seed & no & NA & no &  & noscience & yes & \\
\hline
cntr &  & container &  & NA &  &  & noscience &  & \\
\hline
plot &  & plot &  & NA &  &  & noscience &  & \\
\hline
cpm &  & counts per minute &  & NA &  &  & counts &  & \\
\hline
\end{tabular}

### 2019-11-06 article/parts/01_background.tex

--- 2019-06-12 article/parts/01_background.tex
@@
 
-A large number of chemicals such as pharmaceuticals, pesticides and synthetic hormones are in daily use all over the world with insufficient knowledge about possible adverse effects on the environment.
-In Europe alone some 100,000 chemicals are estimated to be in current use, whereof 30,000 are produced in quantities larger than one ton per year \citep{breithaupt_costs_2006}.
-Although for some chemicals advanced and standardized test methods \citep{oecd_oecd_2018} have been developed, data on toxicity tests for a lot of chemicals, especially newly emerging ones remains scarce and if existent, the information is represented sparsely and in inconsistent formats \citep{gessner_fostering_2016}.
-Therefore we developed \standardtox{}, a tool to aggregate and standardize toxicity test data from various sources [EPA and ETOX] which provides the information in a reproducible manner.
-
-To our knowledge this ist the first approach
-
-reporducible
-
-time saving
-
-data base that collects, processes and aggregates ecotoxicological tests results in order to subsequently publish it in an harmonized form as a web application - the \etoxbase{} tool: \href{http://139.14.20.252:3838/etox-base-shiny/}.
-
-Chemicals are brought to the environment deliberately - as it is the case of pesticides, or arrive therein as a byproduct from other processes (e.g.
+The risk of adverse effects of chemicals on the environment is globally assessed with the help of ecotoxicological tests that are conducted on different organism groups, ideally representing a specific biocoenosis and associated biotopes.
+These tests are conducted by different research groups all over the world who publish the results independently.
+Up to now, only few initiatives, such as the United States Environmental Protection Agency ECOTOXicology knowledgebase (EPA ECOTOX) \citep{elonen_ecotoxicology_2018}, the ETOX (UBA ETOX) <cite UBA> or the Pesticide Property Data Base (PPDB) which aim to consolidate ecotoxicological test results, exist.
+However, the former two only collect test results and hence partly check for quality whereas the latter aims for quality, yet being confined to specific classes of chemicals, in this case pesticides.
+An overall holistic approach in processing the large number of ecotoxicological test results is lacking.
+Therefore, we suggest Standartox, a tool and data base that aims to overcome these paucities by steadily incorporating the ever-growing number of test results in an automated process workflow that ultimately leads to single aggregated test results for specific chemical-organism test combinations, representing the toxicity of a chemical.
+Automated approaches are increasingly important because a growing number of chemicals such as pharmaceuticals, pesticides and synthetic hormones, amongst others are in daily use all over the world with insufficient knowledge about possible adverse effects on the environment.
+Generally, chemicals are brought to the environment deliberately - as it is the case of pesticides, or arrive therein as a byproduct from other processes (e.g.
 atmospheric emissions or wastewater) \citep{schwarzenbach_challenge_2006}.
-Ecosystems provide essential services to human societies such as drinking and irrigation water, food and climate regulation.
+In turn, ecosystems provide essential services to human societies such as drinking and irrigation water, food and climate regulation.
 These services are products of different ecosystem functions that crucially depend on the integrity of the populations and communities that drive these ecosystem functions.
-However, besides habitat degradation, climate change and nutrient enrichment, pollution with man-made chemical toxicants threatens these populations and communities in various ways which are currently not fully understood \citep{steffen_anthropocene_2007}.
-Pollution with man-made chemical toxicants was indeed identified as one of three major environmental problems for which research gaps hamper the derivation of planetary boundaries, i.e.
+However, besides habitat degradation, climate change and nutrient enrichment, pollution with man-made chemicals threatens these populations and communities in various ways which are currently not fully understood \citep{steffen_anthropocene_2007}.
+Pollution with man-made chemicals was indeed identified as one of three major environmental problems for which research gaps hamper the derivation of planetary boundaries, i.e.
 thresholds beyond which irreversible state shifts may occur \citep{steffen_anthropocene_2007}.
@@
 \citet{bernhardt_synthetic_2017} argue further that the knowledge gap how chemicals effect populations and communities and hence ecosystem functions and ecosystem services, would also impede society's ability to accomplish the Sustainable Development Goals of the United Nations.
-According to Breithaupt \citet{breithaupt_costs_2006} less than one percent of chemicals released to the environment are thoroughly tested.
+In Europe alone some 100,000 chemicals are estimated to be in current use, whereof 30,000 are produced in quantities larger than one ton per year \citep{breithaupt_costs_2006}.
+Data on toxicity tests for a lot of chemicals, especially newly emerging ones remains scarce and if existent, the information is represented sparsely and in inconsistent formats \citep{gessner_fostering_2016}.
+Therefore, tools such as Standartox are a needed step adding a layer of abstraction satisfying modern environmental risk assessment demands.
+Standartox aggregates ecotoxicological test results in a standardized and reproducible way, providing the basis for the derivation of risk indicators such as Species Sensitivity Distributions (SSDs) \citep{posthuma_species_2002} and Toxic Units (TUs), which represent two prominent concepts to assess effects on organisms in ecotoxicology \citep{kefford_definition_2011, schafer_effects_2011}.
+Besides aggregating ecotoxicological test results, Standartox provides a concise overview of the status of tested chemicals, allowing the identification of potential research gaps.
+Furthermore Standartox comes as an R package largely reducing search time and adding convenience structures for researchers \citep{rcoreteam_language_2017}.
 
-This should facilitate researchers the access to ecotoxicological test results and support modern chemical risk assessment (CRA) workflows.
-The \etoxbase{} makes use of ecotoxicological test data, obtained from the ECOTOXicology knowledgebase (ECOTOX) created by the U.S.
-Environmental Protection Agency (EPA).
-ECOTOX collects raw, non harmonized ecotoxicological test data on aquatic and terrestrial wildlife as well as plants and publishes it on a quarterly basis.
-By the time of writing it contained 926,108 test results, comprising 11,685 chemicals tested on 12,668 different species \citep{elonen_ecotoxicology_2018}.
-The collected ecotoxicological test data contain a variety of measured endpoints such as Effective Concentrations (\ecfifty{}) values, No-observed effect concentrations (NOEC) or lowest observed effect concentrations (LOEC).
-Likewise the data set contains different effect measures such as mortality, intoxication and growth as well as a multiplicity of test durations ranging from seconds to weeks or even years.
-The \etoxbase{} aggregates this information according to the user's inputs.
-These can then be used for the derivation of risk indicators such as Species Sensitivity Distributions (SSDs) \citep{posthuma_species_2002} and Toxic Units (TUs), which represent two prominent concepts to assess effects on organisms in ecotoxicology.
-The former combines toxicity data of several organism groups towards a chemicals to estimate effects on biotic communities, the latter refers to effects of a specific organism (group) towards a chemical.
-The two concepts are widely used in ecotoxicology \citep{kefford_definition_2011, schafer_effects_2011} since the allow for a comparison of toxicities across multiple chemicals and biological communities, respectively.
-Although performed on one and the same organism and chemical, outcomes of different ecotoxicological tests vary greatly due to differences in test parameters, genetic differences in individual populations or other irrepressible factors making a selection of appropriate test results for CRA laborious.
-The \etoxbase{} aims to relief this process and provide single toxicity values for organism, chemical and test duration combinations.
-The \etoxbase{} downloads every new version of the ECOTOX data base and performs several quality checks, e.g.
-unit harmonization, and cleaning steps on it.
-Hence, new scientific test results are constantly incorporated.
-Additionally the \etoxbase{} queries chemical- and organism-specific variables from other open data bases to fill information gaps such as water solubility, organism habitat or regional occurrence patterns in the ECOTOX data base.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% IDEAS %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\iffalse
+
+Curently \citep{schafer_future_2019}
+
+include \citep{yamamuro_neonicotinoids_2019} as a new prominent examples for effects on the ecosystem
 
@@
     \item time saving
+\fi
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% OLD %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \iffalse

### 2019-11-06 article/tikz/pipeline-organigram.tex
Updates from Overleaf

@@
 	ecotox/.style = {rectangle, minimum width = 4cm, minimum height = 3cm, text centered, draw = black, fill = blue!10},
-	data/.style = {rectangle, minimum height = 3cm, minimum width = 7cm, rounded corners, draw = black},
-	app/.style = {circle, minimum height = 3cm, fill = red!40},
-	api/.style = {circle, minimum height = 3cm, fill = red!40},
-	rpackage/.style = {circle, minimum height = 3cm, fill = red!40},
+	data/.style = {rectangle, minimum width = 7cm, minimum height = 3cm, text centered, rounded corners, draw = black},
+	app/.style = {circle, minimum height = 3cm, text centered, fill = red!40},
+	api/.style = {circle, minimum height = 3cm, text centered, fill = red!40},
+	rpackage/.style = {circle, minimum height = 3cm, text centered, fill = red!40},
 	ids/.style = {diamond, minimum width = 3cm, minimum height = 2cm, text centered, draw = black, fill = green!30},
@@
 \node (taxa) [ids, below of = cas, xshift = 0cm, yshift = -2.5cm] {Taxa};
-\node (db-chemical) [db, right of = cas, xshift = 4cm] {Chemical\\data bases};
-\node (db-habitat) [db, right of = taxa, xshift = 4cm, yshift = 1cm] {Habitat\\data bases};
-\node (db-region) [db, right of = taxa, xshift = 4cm, yshift = -1cm] {Region\\data bases};
+\node (db-chemical) [db, right of = cas, xshift = 4cm, align = center] {Chemical\\data bases};
+\node (db-habitat) [db, right of = taxa, xshift = 4cm, yshift = 1cm, align = center] {Habitat\\data bases};
+\node (db-region) [db, right of = taxa, xshift = 4cm, yshift = -1cm, align = center] {Region\\data bases};
 % standartox data
-\node (stx-data) [data, below of = bg1-pipeline, xshift = 5.5cm, yshift = -6cm] {Standatox\\data set};
+\node (stx-data) [data, below of = bg1-pipeline, xshift = 5.5cm, yshift = -6cm, align = center] {Standatox\\data set};
 % user interfaces
-\node (stx-app) [app, below of = stx-data, yshift = -3.5cm, xshift = -2cm] {Web\\application};
+\node (stx-app) [app, below of = stx-data, yshift = -3.5cm, xshift = -2cm, align = center] {Web\\application};
+\begin{pgfonlayer}{bg1}
+    \node (stx-app-screenshot) [below of = stx-app, xshift = -2cm] {\includegraphics[width=.5\textwidth]{article/figures/screenshot_standartox_app.png}};
+\end{pgfonlayer}
 \node (stx-api) [api, below of = stx-data, yshift = -3.5cm, xshift = 2cm] {API};
 \node (stx-rpackage) [rpackage, below of = stx-api, yshift = -2cm, xshift = 2cm] {R-package};
+\begin{pgfonlayer}{bg1}
+    \node (stx-rpackage-screenshot) [below of = stx-rpackage] {\includegraphics[width=.7\textwidth]{article/figures/screenshot_standartox_rpackage.png}};
+\end{pgfonlayer}
 %% backgrounds
-\begin{pgfonlayer}{bg1}
-    \node (bg1-pipeline) [background, minimum height = 9cm, minimum width = 16cm, above of = epa-ecotox, xshift = 5.5cm, yshift = -1cm];
-    \node (bg1-user) [background, below of = bg1-pipeline, minimum height = 7cm, minimum width = 12cm, yshift = -12cm];
+\begin{pgfonlayer}{bg2}
+    \node (bg1-pipeline) [background, minimum height = 9cm, minimum width = 17cm, above of = epa-ecotox, xshift = 5.5cm, yshift = -1cm];
+    \node (bg1-user) [background, below of = bg1-pipeline, minimum height = 7cm, minimum width = 17cm, yshift = -12cm];
 \end{pgfonlayer}
@@
 \draw [arrow1] (taxa.east) -- (db-region.west);
-\begin{pgfonlayer}{bg1}
+\begin{pgfonlayer}{bg2}
     \draw [arrow2] (bg1-pipeline.south) -- (stx-data.north);
-    \draw [arrow2] (stx-data.south) -- (stx-app.north);
-    \draw [arrow2] (stx-data.south) -- (stx-api.north);
+    \draw [arrow2] (stx-data.south) -- (stx-app.north east);
+    \draw [arrow2] (stx-data.south) -- (stx-api.north west);
 \end{pgfonlayer}

### 2019-11-06 article/parts/02_methods.tex

@@
 The interactive APP was built in R, using the shiny web application framework \citep{chang_shiny_2018} which runs with the help of a shiny server \citep{HOW-TO-CITE-SHINY-SERVER} on www.standartox.uni-landau.de.
-It is set of R functions that filter and aggregate the data according to a user's inputs \ref{fig:app}.
-
-Finally in the web application, a `filtered` and an `aggregated` data set as well as an interactive plot, for data exploration are returned.
-These two data sets can be downloaded.
-In the following these two parts are explained in detail, naming the respective R scripts in brackets.
+It is set of R functions that allow a user to filter and aggregate the data.
+The data can then be downloaded.
 
 \subsubsection*{Application programming interface (API)}
-The API is built by using the plumber-package \citep{trestletechnologyllc_plumber_2018} which allows for the creation of REST APIs from R.
+The API is built by using the R package plumber \citep{trestletechnologyllc_plumber_2018} which allows for the creation of RESTful APIs from R.
 Four API-endpoints ('/catalog', '/filter/rds', 'aggregate' and 'meta') that can be queried from the client side were created.
 The \textit{catalog} API-endpoint returns a JavaScript Object Notation (JSON) file containing a catalog of possible filter parameters to choose from.
-The \textit{filter-rds} returns the filtered test results as a compressed serialized file \citep{klik_fst_2019}, to reduce size and allow for fast user queries.
+The \textit{filter} returns the filtered Standartox table as a compressed serialized binary file \citep{klik_fst_2019}, to reduce size and allow for fast user queries.
 The \textit{aggregate} API-endpoint returns a R function to the client allowing the client the aggregation of the filtered data.
-Lastly, the \textit{meta} API-endpoint returns a JSON file with a timestamp of the request and the used Standartox version.
-The API is designed to be used with the R-package standartox and therefore uses serialization methods specific to R (rds() from the base-package and fst() from the fst-package).
-The code for the the API lies in the file ~/standartox-app/api.R.
+Lastly, the \textit{meta} API-endpoint returns a JSON file with meta information, such as the timestamp of the request and the used Standartox version.
+The API is designed to be used with the R-package standartox and therefore uses serialization methods specific to R (rds() from the R package base and fst() from the package fst).
 
@@
 The former accesses the \textit{catalog} API-endpoint and returns a R-list of possible filter parameters that can be used in \textit{stx\_query()}.
-The latter returns a R list of three tables (R data.frames) containing a filtered data set, an aggregated data set and a table with the meta information retrieved from the \textit{meta} API-endpoint.
+The latter returns a R list of three tables (R data.frames) containing the filtered data set, the aggregated data set and a table with the meta information retrieved from the \textit{meta} API-endpoint.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% OLD %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\iffalse
 \textbf{OLD: Standartox aggregates the test results according to chosen filters in a two step process.
@@
 Thereof the second step calculates the minimum, the maximum, the median, the geometric mean, or the arithmetic mean as an aggregate.}
+\fi

### 2019-11-11 article/parts/02_methods.tex
Updates from Overleaf

@@
 Standartox consists of three parts of software explained in detail below.
-Firstly a data acquisition and preparation processing pipeline which downloads quarterly released new versions of the EPA ECOTOX, upon which Standartox is built.
+Firstly a data acquisition and preparation processing pipeline which downloads quarterly released new versions of the EPA ECOTOX data base, upon which Standartox is built.
 Secondly, Standartox is accessible via a web application (APP) and an application programming interface (API).
@@
 \subsection*{Data acquisition and preparation pipe line}
-The EPA ECOTOX, which is released on a quarterly basis (current release: 12.09.2019) is checked for new releases in monthly intervals and downloaded as soon as a new release is published.
-Subsequently the EPA ECOTOX is built into a PostgreSQL data base and SQL functions for further processing are added and known errors are corrected.
-Lookup tables for filtering and converting the data are uploaded.
-Thereafter, a meta data table is created in the data base, comprising the version number of the EPA ECOTOX.
-In a next step Chemical Abstracts Service (CAS) identifiers are used to query identifier data bases, including the Chemical Identifier Resolver (CIR) service <CITATION> and the Pubchem data base <CITATION>.
-Chemical identifiers (CAS, InchI) and taxonomic names are then used to query chemical and taxonomic specific data bases (Table \ref{tab:data-base-additional}) to add information to the local data base.
-The data is aggregated and unique values are derived.
-<PUT LOOKUPTABLES HERE>.
-From here on the Standartox data set is compiled, whereupon concentration and duration units are converted (see tables \ref{tab:conv-concentration}, \ref{tab:conv-duration}) in order to harmonize data.
-A catalog object, listing unique values for categorical variables and ranges for continuous variables, respectively is created.
-To guarantee the accuracy of the data check scripts are run at the end.
+The US Environmental Protection Agency (EPA) releases the EPA ECOTOX data base on a quarterly basis.
+The latest version, released on 12.09.2019 contains data on 11,756 chemicals, 12,906 taxa, 49,153 references and 952,634 results \citep{usepa_ecotox_2019} constituting a vast collection of ecotoxicological tests, which can be downloaded as pipe delimited ASCII files.
+Standartox downloads the EPA ECOTOX data base whenever a new version is published and builds into a local PostgreSQL data base \citep{szocs_build_2019}.
+Subsequently SQL functions for further processing are added and known errors are corrected (and also reported).
+Lookup tables <PUT LOOKUPTABLES REFERNCE HERE> for filtering and converting the data together with a meta table, containing inter alia the release version of the EPA ECOTOX data base are created as well.
+In a next step Chemical Abstracts Service (CAS) identifiers are used to query identifier data bases, including the Chemical Identifier Resolver (CIR) service \citep{nationalinstitutesofhealthnih_chemical_2019} and the Pubchem data base \citep{kim_pubchem_2016}.
+Chemical identifiers (CAS, InchI) and taxonomic names are then used to query chemical and taxonomic specific data bases (Table \ref{tab:data-base-additional}) to add information, such as chemical grouping or organism habitat preferences and global occurrence patterns to the local build.
+From here on the final Standartox data set is compiled, whereupon concentration and duration units are converted (see tables \ref{tab:conv-concentration}, \ref{tab:conv-duration}) in the process of harmonizing the data.
+Out of the unique 1229 concentration units in the EPA ECOTOX data base, Standartox retains only those (n = X) which are convertible to one of the following units: ug/l, ul/l, ppb, g/m2, mg/kg, \% and flags the remaining as 'other', resulting in 873473 out of 952625 test results.
+Likewise, out of the 125 test duration units in the EPA ECOTOX, Standartox retains only those (n = X) which are accurately convertible to hours, leading to 905110 out of 952625 test results).
+Thereby concentration unit additions such as 'food', soil, ai etc.
+are neglected since they are also coded in other variables and only hinder the processing of units.
+Units such as per day (mg/kg/day) are generally excluded.
+Along with the created Standartox data set, a catalog of it, listing all possible entries and value ranges, for continuous variables is created.
+As a last step, to guarantee the accuracy of the data, check scripts are run at the end.
 Single processing scripts are listed in figure \ref{fig:pipeline-tree}.
 The final Standartox table together with the catalog are exported.
-Both the APP and the API access these objects.
+Both the APP and the API access it.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% CONTINUE HERE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \subsection*{Web application (APP) and Application Programming Interface (API)}
-
 The web application and the API load the compressed serialized Standartox data into memory and allow a client to interact with it.
@@
 The API is built by using the R package plumber \citep{trestletechnologyllc_plumber_2018} which allows for the creation of RESTful APIs from R.
-Four API-endpoints ('/catalog', '/filter/rds', 'aggregate' and 'meta') that can be queried from the client side were created.
-The \textit{catalog} API-endpoint returns a JavaScript Object Notation (JSON) file containing a catalog of possible filter parameters to choose from.
-The \textit{filter} returns the filtered Standartox table as a compressed serialized binary file \citep{klik_fst_2019}, to reduce size and allow for fast user queries.
-The \textit{aggregate} API-endpoint returns a R function to the client allowing the client the aggregation of the filtered data.
-Lastly, the \textit{meta} API-endpoint returns a JSON file with meta information, such as the timestamp of the request and the used Standartox version.
+Four API-endpoints (\textit{/catalog}, \textit{/filter}, \textit{/aggregate} and \textit{/meta}) that can be queried from the client side were created.
+The \textit{/catalog} API-endpoint returns a JavaScript Object Notation (JSON) file containing a catalog of possible filter parameters to choose from.
+The \textit{/filter} returns the filtered Standartox table as a compressed serialized binary file created by the R fst package \citep{klik_fst_2019}, to reduce size and allow for fast user queries.
+The \textit{/aggregate} API-endpoint returns a R function to the client allowing the client the aggregation of the filtered data.
+Lastly, the \textit{/meta} API-endpoint returns a JSON file with meta information, such as the timestamp of the request and the used Standartox version.
 The API is designed to be used with the R-package standartox and therefore uses serialization methods specific to R (rds() from the R package base and fst() from the package fst).
@@
 The R-package standartox contains two functions \textit{stx\_catalog()} and \textit{stx\_query()}.
-The former accesses the \textit{catalog} API-endpoint and returns a R-list of possible filter parameters that can be used in \textit{stx\_query()}.
-The latter returns a R list of three tables (R data.frames) containing the filtered data set, the aggregated data set and a table with the meta information retrieved from the \textit{meta} API-endpoint.
+The former accesses the \textit{/catalog} API-endpoint and returns a R list object of possible filter parameters that can be used in \textit{stx\_query()}.
+The latter returns a R list of three tables (R data.frames) containing the filtered data set, the aggregated data set and a table with the meta information retrieved from the \textit{/meta} API-endpoint.
 

### 2019-11-11 article/tikz/pipeline-organigram.tex
Updates from Overleaf

@@
 	arrow1/.style = {->, > = stealth, draw = black, color = gray!80, line width = 2.5pt},
-	arrow2/.style = {->, > = latex, draw = black, line width=10pt},
+	arrow2/.style = {->, > = latex, draw = black, line width = 10pt},
+	arrow3/.style = {draw = gray!10, line width = 2cm},
 	background/.style = {rectangle, rounded corners, draw = black, fill=gray!0.01}
@@
 % pipeline
-\node (epa-ecotox) [ecotox, align = center] at (0cm,0cm) {EPA ECOTOX\\data base\\(cleaned)};
-\node (cas) [ids, below of = epa-ecotox, xshift = 6cm, yshift = 3cm] {CAS};
+\node (epa-ecotox) [ecotox, above of = epa-ecotox-clean, yshift = 5cm] {EPA};
+\node (epa-ecotox-clean) [ecotox, align = center] at (-3cm,-0.5cm) {EPA ECOTOX\\data base\\(cleaned)};
+\node (cas) [ids, below of = epa-ecotox-clean, xshift = 6cm, yshift = 3cm] {CAS};
 \node (taxa) [ids, below of = cas, xshift = 0cm, yshift = -2.5cm] {Taxa};
@@
 % standartox data
-\node (stx-data) [data, below of = bg1-pipeline, xshift = 5.5cm, yshift = -6cm, align = center] {Standatox\\data set};
+\node (stx-data) [data, below of = epa-ecotox-clean, xshift = 7.5cm, yshift = -5.9cm, align = center] {Standartox\\data set};
 % user interfaces
@@
 \begin{pgfonlayer}{bg2}
-    \node (bg1-pipeline) [background, minimum height = 9cm, minimum width = 17cm, above of = epa-ecotox, xshift = 5.5cm, yshift = -1cm];
-    \node (bg1-user) [background, below of = bg1-pipeline, minimum height = 7cm, minimum width = 17cm, yshift = -12cm];
+    %\node (bg1-pipeline) [background, minimum height = 9cm, minimum width = 17cm, above of = epa-ecotox-clean, xshift = 5.5cm, yshift = -1cm];
+    \node (bg1-user) [background, below of = bg1-pipeline, minimum height = 8cm, minimum width = 17cm, yshift = -13cm, xshift = 5cm];
+
+    \draw [arrow3] (12cm,0cm) arc[radius = 7.5cm, start angle = 0, delta angle = 360];
+
 \end{pgfonlayer}
@@
 % small
-\draw [arrow1] (epa-ecotox.east) -- (cas.west);
-\draw [arrow1] (epa-ecotox.east) -- (taxa.west);
+\draw [arrow1] (epa-ecotox-clean.east) -- (cas.west);
+\draw [arrow1] (epa-ecotox-clean.east) -- (taxa.west);
 \draw [arrow1] (cas.east)-- (db-chemical.west);
@@
 \begin{pgfonlayer}{bg2}
-    \draw [arrow2] (bg1-pipeline.south) -- (stx-data.north);
-    \draw [arrow2] (stx-data.south) -- (stx-app.north east);
-    \draw [arrow2] (stx-data.south) -- (stx-api.north west);
+    \draw [->, > = latex, draw = gray!10, line width = 1cm] (stx-data.south) -- node[auto] ++ (0cm,-8cm);
 \end{pgfonlayer}

### 2019-11-11 article/00_article.tex
Updates from Overleaf

@@
 \documentclass[english]{article}
+
+%% ---------------------------- packages ------------------------------------ %%
 \usepackage[T1]{fontenc}
 \usepackage{natbib} % citations
-\usepackage{csvsimple}
+\usepackage{csvsimple, longtable, booktabs}
 \usepackage{pgfplotstable}
@@
 \usepackage{hyperref}
+\usepackage{listings} % code highlighting
+\usepackage{color}
 
+%% ---------------------------- commands ------------------------------------ %%
 \newcommand{\standartox}{STANDARTOX}
@@
 \newcommand{\ecfifty}{EC\textsubscript{50}}
+
+%% ---------------------------- code ---------------------------------------- %%
+\definecolor{dkgreen}{rgb}{0,0.6,0}
+\definecolor{gray}{rgb}{0.5,0.5,0.5}
+\definecolor{mauve}{rgb}{0.58,0,0.82}
+
+\lstset{frame=tb,
+  language=R,
+  aboveskip=3mm,
+  belowskip=3mm,
+  showstringspaces=false,
+  columns=flexible,
+  basicstyle={\small\ttfamily},
+  numbers=none,
+  numberstyle=\tiny\color{gray},
+  keywordstyle=\color{blue},
+  commentstyle=\color{dkgreen},
+  stringstyle=\color{mauve},
+  breaklines=true,
+  breakatwhitespace=true,
+  tabsize=3
+}
 
@@
 \begin{figure}
-    \includestandalone[scale=0.5]{tikz/tree_source}
-    \caption{Alternative processing pipeline}
+    % TODO make better graphic
+    \includegraphics[width=1\linewidth]{article/figures/scripts_dendrogram.png}
+    \caption{Processing pipeline}
     \label{fig:pipeline-tree}
@@
 
-\pagebreak
-\section*{Supplement}
-\subsection*{R-package list} %% TODO put in appropriate list
-\label{list:r-packages}
-%% TODO how to refernce this?
-\input{article/tables/r_package_list.tex}
+\input{article/parts/06_supplement.tex}
 \pagebreak
 
-\subsection*{Data base tables}
-\input{article/tables/data_base_tables.tex}
-
-\subsection*{Application parameters}
-\input{article/tables/app_parameters.tex}
-
 \end{document}

### 2019-11-11 article/parts/01_background.tex

@@
 These tests are conducted by different research groups all over the world who publish the results independently.
-Up to now, only few initiatives, such as the United States Environmental Protection Agency ECOTOXicology knowledgebase (EPA ECOTOX) \citep{elonen_ecotoxicology_2018}, the ETOX (UBA ETOX) <cite UBA> or the Pesticide Property Data Base (PPDB) which aim to consolidate ecotoxicological test results, exist.
-However, the former two only collect test results and hence partly check for quality whereas the latter aims for quality, yet being confined to specific classes of chemicals, in this case pesticides.
+Up to now, only few initiatives, such as the United States Environmental Protection Agency ECOTOXicology knowledgebase (EPA ECOTOX) \citep{elonen_ecotoxicology_2018}, the German Umweltbundesamt's Information System Ecotoxicology and Environmental Quality Targets (ETOX) \citep{umweltbundesamt_etox_2019} or the Pesticide Property Data Base (PPDB) aiming to consolidate ecotoxicological test results, exist.
+The former two collect test results and hence only partly check for quality whereas the latter aims for quality, yet being confined to specific classes of chemicals, in this case pesticides.
 An overall holistic approach in processing the large number of ecotoxicological test results is lacking.
@@
 Data on toxicity tests for a lot of chemicals, especially newly emerging ones remains scarce and if existent, the information is represented sparsely and in inconsistent formats \citep{gessner_fostering_2016}.
-Therefore, tools such as Standartox are a needed step adding a layer of abstraction satisfying modern environmental risk assessment demands.
-Standartox aggregates ecotoxicological test results in a standardized and reproducible way, providing the basis for the derivation of risk indicators such as Species Sensitivity Distributions (SSDs) \citep{posthuma_species_2002} and Toxic Units (TUs), which represent two prominent concepts to assess effects on organisms in ecotoxicology \citep{kefford_definition_2011, schafer_effects_2011}.
+Therefore, tools such as Standartox are a required step adding a layer of abstraction satisfying the needs of modern environmental risk assessment demands.
+Standartox aggregates ecotoxicological test results in a standardized and reproducible way, providing the basis for the derivation of risk indicators such as Species Sensitivity Distributions (SSD) \citep{posthuma_species_2002} and Toxic Units (TU), which represent two prominent concepts to assess effects on organisms in ecotoxicology \citep{kefford_definition_2011, schafer_effects_2011}.
 Besides aggregating ecotoxicological test results, Standartox provides a concise overview of the status of tested chemicals, allowing the identification of potential research gaps.
-Furthermore Standartox comes as an R package largely reducing search time and adding convenience structures for researchers \citep{rcoreteam_language_2017}.
+Standartox comes with two front-ends, an R \citep{rcoreteam_language_2017} package and a web tool, adding convenience structures and largely reducing work time for users.
 
@@
 
-include \citep{yamamuro_neonicotinoids_2019} as a new prominent examples for effects on the ecosystem
+include \citep{yamamuro_neonicotinoids_2019} and \citep{seibold_arthropod_2019} as a new prominent examples for effects on the ecosystem
 

### 2019-11-11 article/parts/04_technical_validation.tex

--- 2019-10-22 article/parts/04_technical_validation.tex
@@
 %%of their data.
-
 \section{Technical validation}
 
+\subsection{Unit Harmonisation}
+In order to guarantee appropriate conversion and harmonisation of units we take a random sample of the 50 most occurring concentration units (Table \ref{tab:chck-concentration}) as well as six of the most occurring duration units (Table \ref{tab:chck-duration}) and convert them manually to subsequently compare them to the conversions done by Standartox.
+This is check is done whenever Standartox is rebuilt.
+
 \subsection{Comparison to other data bases}
-
 In order to asses the validity of the Standartox approach to aggregate test results we compared Standartox values with effect values for the same chemicals and the same test duration from the PPDB (n=3601) \ref{fig:standartox_ppdb_diff}.
@@
 
-\subsection{Unit Harmonisation}
-
-In order to assure appropriate conversion and harmonisation of units we took a random sample of the 50 most occurring units and converted them manually and subsequently compared them to the conversions done by Standartox.
-
-We did this for concentration units and duration units
-
-in R/chck\_unit\_conversion.R
-
 \subsection{Overall distribution of tests}

### 2019-11-11 article/parts/05_usage_notes.tex

--- 2019-10-15 article/parts/05_usage_notes.tex
@@
 %%on data use.
+Standartox can either be accessed through the APP or the API together with the R-package standartox.
+Both access pathways allow for the same filters and aggregation methods (Table \ref{tab:app-parameters} to be applied on the data.
+In the former the user can download the resulting data sets as a comma-separated values file whereas in the latter the user can directly load them in R.
+The user can choose among different parameters to filter and aggregate the test data accordingly \ref{tab:app-parameters}.
+The user can filter for compound-specific filters, including the CAS number, the concentration type (e.g.
+Active Ingredient, Formulation), the chemical class (e.g.
+Insecticides, Metals) and taxon-specific filters, including common taxonomic groups (e.g.
+Daphniidae, Algae), the habitat (e.g.
+freshwater) and the regional occurrence (e.g.
+Europe, Asia) of the organisms.
+Furthermore, the user can refine the results to specific test durations (in hours), effect groups (e.g.
+Mortaliy, Population, Growth) and endpoints (EC\textsubscript{50}, EC\textsubscript{10}, LOEC or NOEC).
+As aggregates the user can choose the minimum, the maximum, the median, the geometric mean or the arithmetic mean.
+All the steps described in this section are selectable in the GUI and are executed at each click by the application.
+The drop down menu `Download data` allows to download a filtered data set as well as an aggregated data set.
 
-\subsection*{Web application}
+Percentage values next to inputs show the proportion of the respective option (e.g. \textit{Active Ingredient - 43\%} indicates that 43\% of the data are Active Ingredients).
 
-Users can define several input parameters \ref{tab:inputs} in the application's graphical user interface (GUI), handled in the ui.R script in order to adjust the aggregation according specific requirements.
-In doing so users execute R functions (fun\_filter.R, fun\_aggregation.R, fun\_filagg\_plot\_ly.R) loaded in the server.R script, which produce a 'filtered' and an 'aggregated' data set as well as an interactive overview plot.
-Only the 'aggregated' data set is presented in the GUI.
-However, both data sets can be downloaded via a download button.
+\subsection*{R package}
+The R-package standartox consists of the two functions stx\_catalog() and stx\_query().
+The former allows the retrieval of a catalog of possible parameters that can be used as an input for the latter.
+stx\_query() fetches the toxicity values from the data base and returns them as a R list object (compare Code \ref{listing:standartox-example}).
 
-The \etoxbase{} is designed to support CRA of ecotoxicologists.
+\begin{lstlisting}[
+    caption = {standartox R-package code} example,
+    label = {listing:standartox-example}]
+# install
+install.packages('remotes')
+remotes::install_github('andschar/standartox') # package not yet on CRAN
+# retrieve catalog    
+require(standartox)
+catal = stx_catalog()
+# retrieve data
+l = stx_query(cas = '1071-83-6',
+                endpoint = 'XX50',
+                taxa = 'Oncorhynchus',
+                duration = c(24, 120))
+\end{lstlisting}
+
+Standartox is designed to support CRA.
 With an increased amount of available ecotoxicological test data, it becomes fundamental to provide and distribute such information in reasonable formats, meaning accessible for humans as well as machines.
@@
 Each data source has its limitations though: Collecting toxicity information from individual publications is laborious and using already compiled data from publications or data bases is often limited to a specific groups of chemicals, such as pesticides, biozides etc.
-The \etoxbase{}, making use of the EPA ECOTOX data base, the largest publicly available collection of ecotoxicological test results, aims to facilitate the data retrieval process by providing adequate aggregates (\ecfifty{}, NOEC, LOEC) of the tests results.
+Standartox, making use of the EPA ECOTOX, the largest publicly available collection of ecotoxicological test results, aims to facilitate the data retrieval process by providing adequate aggregates (EC\textsubscript{50}, NOEC, LOEC) of the tests results.
 These are then used for the calculation of SSDs, TUs or other risk indicators.
@@
 TUs are another indicator of effects of chemicals on organisms.
-They are calculated by dividing a measured concentration (c) by an \ecfifty{} value (cf.
+They are calculated by dividing a measured concentration (c) by an EC\textsubscript{50} value (cf.
 Equation \ref{eq:tu}).
@@
 \begin{figure}[h!]
-    \dentering
     \includegraphics[width=1\linewidth]{article/figures/ssd2_boot.png}
@@
 
-\subsection*{Filter and Aggregate}
+\subsection*{Limitations}
+Since the EPA ECOTOX data at hand merely represents a sample of the reality of the toxicity of chemicals on organisms, it has to be stated that the returned Standartox values are subject to change.
+New studies will be published and constantly incorporated into Standartox, improving the toxicity estimates.
+However, we expect values of well tested chemical-organism combinations to be relatively stable.
+A future analysis on the smallest number of tests needed to return trustworthy results, could shed more light on this aspect.
+Furthermore, toxicity test results are certainly influenced by individual tests parameters, such as pH, temperature or conductivity amongst others \citep{rosenkrantz_influence_2013, li_temperature_2011}.
+Such information is not included in the aggregation performed by Standartox, since the EPA ECOTOX only provides sparse records on those.
+Most frequently provided test parameters (and their percentage) are medium temperature (77\%), pH:56\%, hardness:27\%, dissolved oxygen:18\%, Alkalinity:15\% and salinity:9\%.
+Others are provided for less than 5\% of the tests.
+A test-mining approach, iterating through the individual test studies could potentially increase this number.
+However, we argue that using the geometric mean as an aggregate allows for an adequate estimation of the toxicity of a chemical towards an organism group.
+We prefer the geometric mean in comparison to the arithmetic mean, since it is considerably less influenced by outliers and is suitable for skewed data.
+Also, the geometric mean is preferable over the median, since the median completely ignores the influence of large or small values, making it unreliable for small data sets <CITATION GEOMETRIC MEAN>.
 
-In the application's GUI, the user can choose among different parameters to filter and aggregate the test data accordingly \ref{tab:inputs}.
-Percentage values next to inputs show the proportion of the respective option (e.g.
-\textit{Active Ingredient - 43\%} indicates that 43\% of the data are Active Ingredients).
-The user can filter for compound-specific filters, including the CAS number, the concentration type (e.g.
-Active Ingredient, Formulation), the chemical class (e.g.
-Insecticides, Metals) and taxon-specific filters, including common taxonomic groups (e.g.
-Daphniidae, Algae), the organism habitat (e.g.
-freshwater) and the continent (e.g.
-Europe, Asia).
-Regarding the test parameters the user can choose test durations (in hours), Effect groups (e.g.
-Mortaliy, Population, Growth) and Endpoints (\ecfifty, EC\textsubscript{10}, LOEC or NOEC).
+\subsection*{Other data bases}
+Other initiatives tidying the large number of ecotoxicological test results were also developed recently.
+They partly aim for overlapping goals, yet have limitations or objectives that differentiate them from Standartox.
+Since 2006, the PPDB provides validated ecotoxicological information on pesticides \citep{lewis_international_2016}.
+% TODO lookup NORMAN citation
+The Network of reference laboratories, research centres and related organisations for monitoring of emerging environmental substances (NORMAN) focuses on assembling river basin specific pollutants \citep{von_der_ohe_new_2011}.
+Connors et al.
+\citet{healthandenvironmentalsciencesinstitutehesi_envirotox_2019, connors_creation_2019} published the EnviroTox data base \href{https://envirotoxdatabase.org/} which also uses, amongst others the EPA ECOTOX data as an input.
+In comparison to Standartox, ENviroTox focuses only on aquatic organisms and uses an qualitative algorithm to exclude unreliable test results.
+They filter for example the data to fish, amphibian invertebrate and algae taxa, to test durations of above 24 hours and aquatic taxa only and add additional information on toxicity endpoints, such as chronic or acute tests as well as mode of action assignments.
+In contrast, Standartox doesn't refine to specific test durations, but leaves it up to the user to decide on such parameters.
+The EnvirTox data base also allows for an aggregation of test results to derive single toxicity values for individual taxa whereas Standartox performs this aggregation for individual chemicals.
+This has to be done in a cumbersome way by down- and uploading the data manually.
+Comptox, is a web tool published by the EPA which, similar to Standartox allows for filtering test results and the retrieval of additional chemical information.
+However, it lacks the possibility to aggregate toxicity test results \citep{usenvironmentalprotectionagency_comptox_2019}.
+Petschick et al.
+(in submission) modeled risk threshold level equivalents for aquatic organisms by using ecotoxicological effect data from the EPA ECOTOX.
+However, none of the above mentioned approaches aim for a standardized aggregation method of toxicity endpoints for individual chemicals.
+Besides, they also lack the possibility for automated and scriptable user requests for high level programming languages, such as R or Python.
+Along with newly created ecotoxicological data bases here, methods of how to efficiently store ecotoxicological data are also proposed as can be seen in the MAGIC Knowledge Base \citep{bub_graphing_2019}.
+This recent increase of efforts to compile ecotoxicological test data bases clearly emphasizes the need for holistic and automated analyses of large-scale ecotoxicological data.
+Standartox is a tool that puts its focus on the aggregation of toxicity data and by this adds a piece to the puzzle of modern ecotoxicological analyses.
+
+\subsection*{Identify chemicals to be tested}
+There is a clear need in ecotoxicology for harmonised approaches to collect ecotoxicological data as other data base compilation attempts recently published show.
+Hitchock et al.
+\citet{hitchcock_improving_2018} argue for an increased awareness concerning newly published ecotoxicological test results to facilitate meta analyses across studies.
+Important statistical parameters, such as mean estimates, variances and sample sizes.
+Likewise, state variables, including medium temperature, pH, mineral contents or organism age and sex should be reported (if not in the text, in an online repository).
+Standartox doesn't include such information mainly because the EPA ECOTOX reports only a few of these variables.
+A future research aim could be to use machine learning techniques to retrieve this information from the referenced articles.
+Above all, recent studies showed that, in defiance of many regulation efforts \citep{schafer_future_2019} chemicals can still cause strong adverse effects on organisms and in turn the environment.
+Neonicotinoids, a group of insecticides was for example related to declines in insectivorous birds \citep{hallmann_declines_2014} and fish yields due to cascading food web effects \citep{yamamuro_neonicotinoids_2019}.
+Standartox provides means to synthesise ecotoxicological test data, thereby identify research gaps.
+Furthermore, standardized approaches could potentially reduce unnecessary animal testing by synthesizing existing ecotoxicological test data.
+
+\begin{figure}
+    \includegraphics[width=1\linewidth]{article/figures/heatmap_tests_n.png}
+    \caption{30 most tested chemicals in EPA ECOTOX.}
+    \label{fig:standartox_ppdb_diff}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TODO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+1) \citep{staveley_challenge_2016} describe what is necessary for future ecotoxicology.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% OLD %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\iffalse
 Besides that, two cleaning steps can also be chosen.
@@
 The outliers are selected as values that exceed lower (0.25) and the upper (0.75) quartile by 1.5 times the inter-quartile range.
-As aggregates the user can choose the minimum, the maximum, the median, the geometric mean or the arithmetic mean.
-All the steps described in this section are selectable in the GUI and are executed at each click by the application.
-The drop down menu `Download data` allows to download a filtered data set as well as an aggregated data set.
-The \etoxbase{} application can be accessed here: \app{}.
-
-\begin{table}[h!]
-\begin{tabular}{|l|l|l|l|}
-\hline
-Menu   & Group       & Subgroups              & Options \\ \hline
-Filter      & Compound    & CAS                    & input CAS \\ \hline
-Filter      & Compound    & Concentration type     & Active Ingredient; Formulation; Total \\ \hline
-Filter      & Compound    & Chemical class         & Insecticides; Herbicides; Metals \\ \hline
-Filter      & Taxon       & Taxon                  & Choose a taxon \\ \hline
-Filter      & Taxon       & Organism habitat       & freshwater; terrestrial; marine; brackish \\ \hline
-Filter      & Taxon       & Continent              & North America; Asia; Europe \\ \hline
-Filter      & Test        & Duration               & enter duration in hours \\ \hline
-Filter      & Test        & Effect groups          & MOR - Mortality; POP - Population \\ \hline
-Filter      & Test        & Endpoins               & EC50, LOEC, LOEL \\ \hline
-Filter      & Checks      & Water solubility check & yes/no \\ \hline
-Filter      & Checks      & Remove outliers        & yes/no \\ \hline
-Aggregation & Aggregation & Aggregate              & Minimum; Maximum; Median \\ \hline
-\end{tabular}
-\caption{\etoxbase GUI paramters to choose.}
-\label{tab:inputs}
-\end{table}
-
-\subsection*{Limitations}
-
-Generally it has to be stated that the returned values are subject to change, as new tests will be constantly incorporated into the \etoxbase{}.
-However, we expect values of well tested chemical-organism combinations to be relatively stable.
-The returned values are certainly influenced by individual tests characteristics, such as pH, temperature or conductivity to just name a few, which influences the effect relationship between chemicals and organisms \citep{rosenkrantz_influence_2013,li_temperature_2011}.
-Unfortunately the EPA ECOTOX data base has only rare records on test parameters, which prevents incorporating such parameters in an aggregation <proportions of available data: Temperature:77\%, pH:56\%, hardness:27\%, dissolved oxygen:18\%, Alkalinity:15\%, salinity:9\%, others:<=5\%>.
-However, using the median or the geometric mean as an aggregate allows for an adequate estimation of the central tendency of the toxicity of a chemical towards an organism group.
-The geometric mean is recommended, as it is in comparison to the arithmetic mean considerably less influenced by outliers and is suitable for skewed data.
-Also, the geometric mean is preferable in contrast to the median, since the median completely ignores the influence of large or small values, making it unreliable for small data sets <CITATION GEOMETRIC MEAN>.
-
-Other initiatives, with different, though partly overlapping aims accumulating existing ecotoxicological test data are published online.
-The PPDB provides validated ecotoxicological test results for pesticides, narrowing chemicals to this group \citep{lewis_international_2016}.
-The Network of reference laboratories, research centres and related organisations for monitoring of emerging environmental substances (NORMAN) focuses on assembling river basin specific pollutants\citep{von_der_ohe_new_2011}.
-The EnviroTox data base \href{https://envirotoxdatabase.org/} makes also use of the \epa{} data in order to derive Predicted no-effect concentrations (PNEC) \citep{health_and_environmental_sciences_institute_hesi_envirotox_2019}.
-In comparison to the \etoxbase, these examples partly rely on the same data, yet not aiming to derive \ecfifty, LOEC and NOEC from it automatically.
-
-\pagebreak
+\fi

### 2019-11-13 article/tikz/pipeline-organigram.tex
Updates from Overleaf

@@
 \node (epa-ecotox) [ecotox, above of = epa-ecotox-clean, yshift = 5cm] {EPA};
-\node (epa-ecotox-clean) [ecotox, align = center] at (-3cm,-0.5cm) {EPA ECOTOX\\data base\\(cleaned)};
+\node (epa-ecotox-clean) [ecotox, align = center] at (-3cm,0cm) {EPA ECOTOX\\data base\\(cleaned)};
 \node (cas) [ids, below of = epa-ecotox-clean, xshift = 6cm, yshift = 3cm] {CAS};
@@
 % standartox data
-\node (stx-data) [data, below of = epa-ecotox-clean, xshift = 7.5cm, yshift = -5.9cm, align = center] {Standartox\\data set};
+\node (stx-data) [ecotox, below of = epa-ecotox-clean, xshift = 7.5cm, yshift = -6.4cm, align = center] {Standartox\\data set};
 % user interfaces
-\node (stx-app) [app, below of = stx-data, yshift = -3.5cm, xshift = -2cm, align = center] {Web\\application};
+\node (stx-app) [app, below of = stx-data, yshift = -5cm, xshift = -2cm, align = center] {Web\\application};
 \begin{pgfonlayer}{bg1}
@@
 \end{pgfonlayer}
-\node (stx-api) [api, below of = stx-data, yshift = -3.5cm, xshift = 2cm] {API};
+\node (stx-api) [api, below of = stx-data, yshift = -5cm, xshift = 2cm] {API};
 \node (stx-rpackage) [rpackage, below of = stx-api, yshift = -2cm, xshift = 2cm] {R-package};
@@
     %\node (bg1-pipeline) [background, minimum height = 9cm, minimum width = 17cm, above of = epa-ecotox-clean, xshift = 5.5cm, yshift = -1cm];
-    \node (bg1-user) [background, below of = bg1-pipeline, minimum height = 8cm, minimum width = 17cm, yshift = -13cm, xshift = 5cm];
-
+    % \node (bg1-user) [background, below of = bg1-pipeline, minimum height = 8cm, minimum width = 17cm, yshift = -13cm, xshift = 5cm];
     \draw [arrow3] (12cm,0cm) arc[radius = 7.5cm, start angle = 0, delta angle = 360];
-
 \end{pgfonlayer}
@@
 \begin{pgfonlayer}{bg2}
-    \draw [->, > = latex, draw = gray!10, line width = 1cm] (stx-data.south) -- node[auto] ++ (0cm,-8cm);
+    \draw [->, > = latex, draw = gray!10, line width = 1cm] (stx-data.south) -- node[auto] ++ (0cm,-4cm);
 \end{pgfonlayer}
 \draw [arrow1] (stx-api) -- (stx-rpackage);
-% large
 

### 2019-11-13 article/parts/02_methods.tex

@@
 From here on the final Standartox data set is compiled, whereupon concentration and duration units are converted (see tables \ref{tab:conv-concentration}, \ref{tab:conv-duration}) in the process of harmonizing the data.
-Out of the unique 1229 concentration units in the EPA ECOTOX data base, Standartox retains only those (n = X) which are convertible to one of the following units: ug/l, ul/l, ppb, g/m2, mg/kg, \% and flags the remaining as 'other', resulting in 873473 out of 952625 test results.
+Out of the unique 1229 concentration units in the EPA ECOTOX data base, Standartox retains only those (n = X) which are convertible to one of the following units: ug/l, g/m2, ppb, mg/kg, \% and flags the remaining as \textit{other}, resulting in 873473 out of 952625 test results.
 Likewise, out of the 125 test duration units in the EPA ECOTOX, Standartox retains only those (n = X) which are accurately convertible to hours, leading to 905110 out of 952625 test results).
-Thereby concentration unit additions such as 'food', soil, ai etc.
+Thereby concentration unit additions such as \textit{food}, \textit{soil}, \textit{ai} etc.
 are neglected since they are also coded in other variables and only hinder the processing of units.
 Units such as per day (mg/kg/day) are generally excluded.
+Since in the current version the EPA ECOTOX data base lists 298 distinct entries, endpoints were refined to three endpoint groups NOEX, LOEX and XX50.
+The former two summarise no-observed-adverse-effect level and lowest-observed-adverse-effect levels, respectively.
+The latter includes various half maximal effective concentration (e.g.
+EC50, LC50, LD50 etc.) commonly used in ecotoxicology (Table \ref{tab:???????}).
+Other endpoints are not included in Standartox.
+Beyond that, users can filter the data according to the CAS number, the concentration type (e.g.
+active ingredient, formulation etc.), the chemical class (e.g.
+fungicides, metals etc.), taxa, organism habitat and region, test duration as well as effect (e.g.
+Mortality, Growth etc.) groups (See Table \ref{tab:app-parameters}) in one of the two front-ends.
 Along with the created Standartox data set, a catalog of it, listing all possible entries and value ranges, for continuous variables is created.
@@
 
+TODO: How does this fit?
+\begin{table}
+    \input{article/tables/data_refinement_counts.tex}
+    \caption{Reduction of data in the compilation process for Standartox}
+    \label{tab:data-refinement}
+\end{table}
+
+\begin{figure}
+    \includestandalone[scale=0.5]{article/tikz/pipeline-organigram}
+    %\input[scale = 0.5]{article/tikz/pipeline-organigram.tex}
+    \caption{Organigram of Standartox.}
+    \label{fig:stx-organigram}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% CONTINUE HERE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \subsection*{Web application (APP) and Application Programming Interface (API)}
-The web application and the API load the compressed serialized Standartox data into memory and allow a client to interact with it.
-The client can then call the functions stx\_filter() and stx\_aggregate() to filter and aggregate the data according to specific parameters (Table \ref{tab:app-parameters}).
+The web application (APP) and the Application Programming Interface (API) load the compressed serialized Standartox data into memory and allow a client to interact with it.
+The client can then call the functions stx\_filter() and stx\_aggregate() which filter and aggregate the data according to specific parameters (Table \ref{tab:app-parameters}).
 
 \subsubsection*{Web application (APP)}
-The interactive APP was built in R, using the shiny web application framework \citep{chang_shiny_2018} which runs with the help of a shiny server \citep{HOW-TO-CITE-SHINY-SERVER} on www.standartox.uni-landau.de.
+The interactive APP was built in R, using the shiny web application framework which runs with the help of a shiny server \citep{chang_shiny_2018} on \url{www.standartox.uni-landau.de}.
 It is set of R functions that allow a user to filter and aggregate the data.
@@
 The API is designed to be used with the R-package standartox and therefore uses serialization methods specific to R (rds() from the R package base and fst() from the package fst).
-
-\begin{figure}
-    \includestandalone[scale=0.5]{article/tikz/pipeline-organigram}
-    %\input[scale = 0.5]{article/tikz/pipeline-organigram.tex}
-    \caption{Organigram of Standartox.}
-    \label{fig:stx-organigram}
-\end{figure}
 

### 2019-11-18 article/parts/02_results_discussion.tex
Updates from Overleaf

--- 2019-11-11 article/parts/05_usage_notes.tex
@@
-\section*{Usage Notes}
+\section*{Results and Discussion}
 %%Brief instructions that may help other researchers reuse these dataset.
@@
 %%on data use.
-Standartox can either be accessed through the APP or the API together with the R-package standartox.
-Both access pathways allow for the same filters and aggregation methods (Table \ref{tab:app-parameters} to be applied on the data.
-In the former the user can download the resulting data sets as a comma-separated values file whereas in the latter the user can directly load them in R.
-The user can choose among different parameters to filter and aggregate the test data accordingly \ref{tab:app-parameters}.
-The user can filter for compound-specific filters, including the CAS number, the concentration type (e.g.
-Active Ingredient, Formulation), the chemical class (e.g.
-Insecticides, Metals) and taxon-specific filters, including common taxonomic groups (e.g.
+Standartox is a tool that allows to filter and aggregate ecotoxicological test results to derive exposure endpoints for individual chemicals.
+It can either be accessed through the APP (Figure \ref{fig:app}) or the API together with the R-package standartox (Listing \ref{listing:standartox-example}).
+Both access pathways allow for the same filters and aggregation methods (Table \ref{tab:app-parameters}) to be applied on the same data set.
+Using the APP, the user can download the resulting data sets as a comma-separated values (csv) file whereas in the API the data can directly be loaded in R.
+Different parameters in order to filter the test data, such as compound-specific parameters, including the CAS number, the concentration type (e.g.
+Active Ingredient, Formulation) or the chemical class (e.g.
+Insecticides, Metals) as well as taxon-specific parameters, including common taxonomic groups (e.g.
 Daphniidae, Algae), the habitat (e.g.
-freshwater) and the regional occurrence (e.g.
-Europe, Asia) of the organisms.
-Furthermore, the user can refine the results to specific test durations (in hours), effect groups (e.g.
-Mortaliy, Population, Growth) and endpoints (EC\textsubscript{50}, EC\textsubscript{10}, LOEC or NOEC).
-As aggregates the user can choose the minimum, the maximum, the median, the geometric mean or the arithmetic mean.
-All the steps described in this section are selectable in the GUI and are executed at each click by the application.
-The drop down menu `Download data` allows to download a filtered data set as well as an aggregated data set.
+freshwater) or the regional occurrence (e.g.
+Europe, Asia) of the organisms can be chosen.
+Furthermore, users can refine the results to specific test durations (in hours), effect groups (e.g.
+Mortality, Population, Growth) and endpoints (EC\textsubscript{50}, LOEC or NOEC).
+As the method of aggregation users can choose the minimum, the maximum, or the geometric mean.
+To provide an easy overview of the data set, percentage values for each parameter entry are shown in the APP in the selection menu as well as by using the R package through querying the \textit{/catalog} API-endpoint.
 
-Percentage values next to inputs show the proportion of the respective option (e.g. \textit{Active Ingredient - 43\%} indicates that 43\% of the data are Active Ingredients).
+\pagebreak
+\begin{figure}
+    \caption{Web application (APP) Standartox allowing to set filter parameters and aggregation methods.
+Data can be downloaded as comma separated value (csv) files.
+An interactive overview plot and table allow for browsing of the data.}
+    \label{fig:app}
+    \includegraphics[width=0.75\linewidth]{article/figures/screenshot_standartox_app.png}
+\end{figure}
 
-\subsection*{R package}
-The R-package standartox consists of the two functions stx\_catalog() and stx\_query().
-The former allows the retrieval of a catalog of possible parameters that can be used as an input for the latter.
-stx\_query() fetches the toxicity values from the data base and returns them as a R list object (compare Code \ref{listing:standartox-example}).
-
+\pagebreak
 \begin{lstlisting}[
-    caption = {standartox R-package code} example,
+    caption = {Sample code to access the Standartox data base through the API with the help of the R-package.
+\textit{stx\_catalog()} returns a catalog of possible filter and aggregation parameters.
+\textit{stx\_query()} returns the Standartox object, a list of the filtered and the aggregated data as well as a meta data entry.},
     label = {listing:standartox-example}]
@@
 
-Standartox is designed to support CRA.
-With an increased amount of available ecotoxicological test data, it becomes fundamental to provide and distribute such information in reasonable formats, meaning accessible for humans as well as machines.
-Currently CRA often relies on test results of a few, well tested standard organism, such as \textit{Daphnia magna}, \textit{Pimephales promelas}, \textit{Selenastrum capricornutum}, although researchers have conducted experiments on a much greater variety of organisms (e.g.
-the 926,108 test results collected in the EPA ECOTOX data base).
-In order to locate ecotoxicological test data, researchers can rely on individual test results reported in publications <CITE>, on compiled and published data sets \citep{malaj_organic_2014} or on data bases such as the Pesticide Property Data Base (PPDB) \citep{lewis_international_2016}.
-Each data source has its limitations though: Collecting toxicity information from individual publications is laborious and using already compiled data from publications or data bases is often limited to a specific groups of chemicals, such as pesticides, biozides etc.
-Standartox, making use of the EPA ECOTOX, the largest publicly available collection of ecotoxicological test results, aims to facilitate the data retrieval process by providing adequate aggregates (EC\textsubscript{50}, NOEC, LOEC) of the tests results.
-These are then used for the calculation of SSDs, TUs or other risk indicators.
-SSDs try to extrapolate effects of chemicals towards organisms from species to a community level by fitting a statistical distribution to toxicity test results on organisms of different trophic levels in order to derive concentrations at which a specific fraction of organisms are effected (often at 5\% - HC\textsubscript{5} \citep{posthuma_species_2002}.
+\pagebreak
+
+\input{article/tables/app-parameters.tex}
+
+\pagebreak
+
+Standartox is designed to foster chemical risk assessment (CRA).
+Due to an increased amount of available ecotoxicological test data, it becomes fundamental to provide and distribute such information in adequate formats, both easily accessible for humans and easily processable for machines.
+Currently, CRA often relies on test results of a few, well tested standard organisms, such as \textit{Daphnia magna}, \textit{Pimephales promelas}, \textit{Selenastrum capricornutum} to name a few, although researchers have conducted experiments on a much greater variety of organisms, observable in the 12,906 taxa from 952,634 test results collected in the EPA ECOTOX data base.
+In order to locate ecotoxicological test data, researchers can rely on individual test results reported in publications, on compiled and published data sets \citep{malaj_organic_2014} or on data bases such as the Pesticide Property Data Base (PPDB) \citep{lewis_international_2016}.
+Each approach has its limitations though.
+Collecting toxicity information from individual publications is laborious and using already compiled data from publications or published data bases is often limited to specific groups of chemicals, such as pesticides, biozides etc.
+Just as well, these resources run at risk of providing out of date data sooner or later or rarely provide sophisticated means to directly process the data.
+Standartox overcomes these paucities by using the EPA ECOTOX data base, the largest publicly available collection of ecotoxicological test information and builds Standartox and its aggregation methods on top of it to provide adequate and readily available estimates of the toxicity of chemicals.
+Besides that, Standartox constitutes not only a one-time compiled data set but rather a scalable method that can steadily incorporate new ecotoxicological test data in an automated manner, allowing for constant improvement of toxicity estimations over time.
+Standartox allows users to access older versions of the data base as well and is created as open source software and therefore a fully reproducible method.
+
+\subsection*{Other data bases}
+Other initiatives tidying the large number of ecotoxicological test results were also developed recently.
+They partly aim for overlapping goals, yet have limitations or objectives that differentiate them from Standartox.
+Since 2006, the PPDB provides validated ecotoxicological information on pesticides \citep{lewis_international_2016}.
+In order to asses the validity of the Standartox aggregation approach, we compared Standartox results with effect values from the PPDB (n=3601) for the same chemicals (Figure \ref{fig:standartox_ppdb_diff}).
+This showed that 91.3\% of the Standartox values diverge by less than a factor of ten from the PPDB values.
+Considering only Standartox values for which at least five or ten tests were aggregated the value increases to 91.7\% and 93.4\%, respectively.
+Admittedly a few Standartox values diverge by a large margin.
+
+\begin{figure}
+    \includegraphics[width=1\linewidth]{article/figures/gg_ppdb_stan_compare_continous.png}
+    \caption{Comparison between Standartox and PPDB values. The black line marks exact coherence and red lines mark a divergence of a factor of 10. Compared organism groups are color coded.}
+    \label{fig:standartox_ppdb_diff}
+\end{figure}
+
+The Network of reference laboratories, research centres and related organisations for monitoring of emerging environmental substances (NORMAN) focuses on assembling river basin specific pollutants \citep{von_der_ohe_new_2011}.
+The EnviroTox data base \href{https://envirotoxdatabase.org/} which also uses, amongst others the EPA ECOTOX data base as an input was recently published \citep{healthandenvironmentalsciencesinstitutehesi_envirotox_2019, connors_creation_2019}.
+In comparison to Standartox, EnviroTox focuses only on aquatic organisms and uses an qualitative algorithm to exclude unreliable test results, such as pre-filtering, for example the data to fish, amphibian invertebrate and algae taxa, to test durations of above 24 hours and aquatic taxa only.
+Besides, they add additional information on toxicity endpoints, such as acute or chronic classifications and mode of action assignments.
+In contrast, Standartox doesn't refine to specific test durations, but leaves it up to the user to decide on such parameters.
+The EnvirTox data base also allows for an aggregation of test results to derive single toxicity values for individual taxa whereas Standartox performs this aggregation for individual chemicals.
+Comptox, is a web tool published by the EPA which, similar to Standartox allows for filtering test results, the retrieval of additional chemical information and predicting toxicity properties, such as 48h \textit{Daphnia magna} LC\textsubscript{50} values.
+However, toxicity predictions are limited to some standard organisms (e.g.
+\textit{Daphnia magna}), and the tool lacks the possibility to retrieve predicted values in an automated way \citep{williams_comptox_2017}.
+Petschick et al.
+(in submission) modeled risk threshold level equivalents for aquatic organisms by using ecotoxicological effect data from the EPA ECOTOX data base to support regulation.
+In summary, none of the above mentioned approaches aim for a holistic standardized aggregation method of exposure endpoints for individual chemicals and also lack the possibility for automated and scriptable user requests from common high level programming languages, such as R or Python.
+Along with newly created ecotoxicological data bases here, methods of how to efficiently store ecotoxicological data are also proposed as can be seen in the MAGIC Knowledge Base \citep{bub_graphing_2019}.
+This recent increase of efforts to compile ecotoxicological test data bases clearly emphasizes the need for holistic and automated analyses of large-scale ecotoxicological data.
+
+\subsection*{Limitations}
+Since the used data in Standartox merely represents a sample of the actual toxicity of chemicals on organisms, it has to be stated that the returned Standartox values are subject to change.
+New studies will be published and constantly incorporated into Standartox, improving the toxicity estimates.
+However, we expect values of well tested chemical-organism combinations to be relatively stable.
+A future analysis on the smallest number of tests needed to return trustworthy results, could shed more light on this aspect.
+Furthermore, toxicity test results are certainly influenced by individual tests parameters, such as pH, temperature or conductivity amongst others \citep{rosenkrantz_influence_2013, li_temperature_2011}.
+Such information is not included in the aggregation performed by Standartox, since the EPA ECOTOX data base only provides sparse records on those.
+Most frequently provided test parameters (and their percentage) are medium temperature (77\%), pH:56\%, hardness:27\%, dissolved oxygen:18\%, Alkalinity:15\% and salinity:9\%.
+Others are provided for less than 5\% of the tests.
+A text-mining approach, iterating through the individual publications could potentially increase this number.
+However, we argue that using the geometric mean as an aggregate allows for an adequate estimation of the toxicity of a chemical towards an organism group.
+We chose the geometric mean in comparison to the arithmetic mean, since it is considerably less influenced by outliers and is suitable for skewed data.
+Also, the geometric mean is preferable over the median, since the median completely ignores the influence of large or small values, making it unreliable for small data sets <CITATION GEOMETRIC MEAN>.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TODO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\iffalse
+- can also be incorporated in EPA Dashboard
+- Identify also underrepresented regions concerning organisms.
+1) \citep{staveley_challenge_2016} describe what is necessary for future ecotoxicology.
+2) Compare Standartox with predictions here: https://comptox.epa.gov/dashboard/predictions/index
+\fi
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% OLD %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\iffalse
+SSD try to extrapolate effects of chemicals towards organisms from species to a community level by fitting a statistical distribution to toxicity test results on organisms of different trophic levels in order to derive concentrations at which a specific fraction of organisms are effected (often at 5\% - HC\textsubscript{5} \citep{posthuma_species_2002}.
 Figure \ref{fig:ssd-isoproturon} shows an example of a SSD and the derived HC\textsubscript{5} value (red dassed line) created for the herbicide Isoproturon.
@@
 
-\subsection*{Limitations}
-Since the EPA ECOTOX data at hand merely represents a sample of the reality of the toxicity of chemicals on organisms, it has to be stated that the returned Standartox values are subject to change.
-New studies will be published and constantly incorporated into Standartox, improving the toxicity estimates.
-However, we expect values of well tested chemical-organism combinations to be relatively stable.
-A future analysis on the smallest number of tests needed to return trustworthy results, could shed more light on this aspect.
-Furthermore, toxicity test results are certainly influenced by individual tests parameters, such as pH, temperature or conductivity amongst others \citep{rosenkrantz_influence_2013, li_temperature_2011}.
-Such information is not included in the aggregation performed by Standartox, since the EPA ECOTOX only provides sparse records on those.
-Most frequently provided test parameters (and their percentage) are medium temperature (77\%), pH:56\%, hardness:27\%, dissolved oxygen:18\%, Alkalinity:15\% and salinity:9\%.
-Others are provided for less than 5\% of the tests.
-A test-mining approach, iterating through the individual test studies could potentially increase this number.
-However, we argue that using the geometric mean as an aggregate allows for an adequate estimation of the toxicity of a chemical towards an organism group.
-We prefer the geometric mean in comparison to the arithmetic mean, since it is considerably less influenced by outliers and is suitable for skewed data.
-Also, the geometric mean is preferable over the median, since the median completely ignores the influence of large or small values, making it unreliable for small data sets <CITATION GEOMETRIC MEAN>.
-
-\subsection*{Other data bases}
-Other initiatives tidying the large number of ecotoxicological test results were also developed recently.
-They partly aim for overlapping goals, yet have limitations or objectives that differentiate them from Standartox.
-Since 2006, the PPDB provides validated ecotoxicological information on pesticides \citep{lewis_international_2016}.
-% TODO lookup NORMAN citation
-The Network of reference laboratories, research centres and related organisations for monitoring of emerging environmental substances (NORMAN) focuses on assembling river basin specific pollutants \citep{von_der_ohe_new_2011}.
-Connors et al.
-\citet{healthandenvironmentalsciencesinstitutehesi_envirotox_2019, connors_creation_2019} published the EnviroTox data base \href{https://envirotoxdatabase.org/} which also uses, amongst others the EPA ECOTOX data as an input.
-In comparison to Standartox, ENviroTox focuses only on aquatic organisms and uses an qualitative algorithm to exclude unreliable test results.
-They filter for example the data to fish, amphibian invertebrate and algae taxa, to test durations of above 24 hours and aquatic taxa only and add additional information on toxicity endpoints, such as chronic or acute tests as well as mode of action assignments.
-In contrast, Standartox doesn't refine to specific test durations, but leaves it up to the user to decide on such parameters.
-The EnvirTox data base also allows for an aggregation of test results to derive single toxicity values for individual taxa whereas Standartox performs this aggregation for individual chemicals.
-This has to be done in a cumbersome way by down- and uploading the data manually.
-Comptox, is a web tool published by the EPA which, similar to Standartox allows for filtering test results and the retrieval of additional chemical information.
-However, it lacks the possibility to aggregate toxicity test results \citep{usenvironmentalprotectionagency_comptox_2019}.
-Petschick et al.
-(in submission) modeled risk threshold level equivalents for aquatic organisms by using ecotoxicological effect data from the EPA ECOTOX.
-However, none of the above mentioned approaches aim for a standardized aggregation method of toxicity endpoints for individual chemicals.
-Besides, they also lack the possibility for automated and scriptable user requests for high level programming languages, such as R or Python.
-Along with newly created ecotoxicological data bases here, methods of how to efficiently store ecotoxicological data are also proposed as can be seen in the MAGIC Knowledge Base \citep{bub_graphing_2019}.
-This recent increase of efforts to compile ecotoxicological test data bases clearly emphasizes the need for holistic and automated analyses of large-scale ecotoxicological data.
-Standartox is a tool that puts its focus on the aggregation of toxicity data and by this adds a piece to the puzzle of modern ecotoxicological analyses.
+Besides that, two cleaning steps can also be chosen.
+Firstly the application allows to exclude test results exhibiting concentrations that are higher than the actual water solubility of the respective chemical at \ang{20} C.
+Secondly the user can choose to exclude outliers (cf.
+Table \ref{tab:scripts-app}).
+The outliers are selected as values that exceed lower (0.25) and the upper (0.75) quartile by 1.5 times the inter-quartile range.
 
@@
 A future research aim could be to use machine learning techniques to retrieve this information from the referenced articles.
-Above all, recent studies showed that, in defiance of many regulation efforts \citep{schafer_future_2019} chemicals can still cause strong adverse effects on organisms and in turn the environment.
+Above all, recent studies showed that, in defiance of many regulation efforts \citep{schafer_future_2019} chemicals can still cause strong adverse effects on organisms and in turn on the environment.
 Neonicotinoids, a group of insecticides was for example related to declines in insectivorous birds \citep{hallmann_declines_2014} and fish yields due to cascading food web effects \citep{yamamuro_neonicotinoids_2019}.
 Standartox provides means to synthesise ecotoxicological test data, thereby identify research gaps.
-Furthermore, standardized approaches could potentially reduce unnecessary animal testing by synthesizing existing ecotoxicological test data.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TODO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-1) \citep{staveley_challenge_2016} describe what is necessary for future ecotoxicology.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% OLD %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\iffalse
-Besides that, two cleaning steps can also be chosen.
-Firstly the application allows to exclude test results exhibiting concentrations that are higher than the actual water solubility of the respective chemical at \ang{20} C.
-Secondly the user can choose to exclude outliers (cf.
-Table \ref{tab:scripts-app}).
-The outliers are selected as values that exceed lower (0.25) and the upper (0.75) quartile by 1.5 times the inter-quartile range.
 \fi

### 2019-11-18 article/parts/03_methods.tex
Updates from Overleaf

--- 2019-11-13 article/parts/02_methods.tex
@@
 Standartox consists of three parts of software explained in detail below.
-Firstly a data acquisition and preparation processing pipeline which downloads quarterly released new versions of the EPA ECOTOX data base, upon which Standartox is built.
-Secondly, Standartox is accessible via a web application (APP) and an application programming interface (API).
-Lastly, the R package standartox constitutes an easily usable way of accessing the tool via the API.
-All the software is written in R \citep{rcoreteam_language_2017} and SQL scripts.
+Firstly a data acquisition and preparation processing pipeline which downloads quarterly released new versions of the EPA ECOTOX data base, upon which Standartox is built and performs several processing steps on the data.
+Secondly, the final Standartox table is accessible via a web application (APP) and an application programming interface (API).
+Lastly, the R \citep{rcoreteam_language_2017} package standartox constitutes an easily usable way of accessing the tool via the API.
 
-\subsection*{Data acquisition and preparation pipe line}
+\subsection*{Processing pipe line}
 The US Environmental Protection Agency (EPA) releases the EPA ECOTOX data base on a quarterly basis.
@@
 Subsequently SQL functions for further processing are added and known errors are corrected (and also reported).
-Lookup tables <PUT LOOKUPTABLES REFERNCE HERE> for filtering and converting the data together with a meta table, containing inter alia the release version of the EPA ECOTOX data base are created as well.
+Lookup tables (Supplement \ref{sup:conv-concentration}, \ref{sup:conv-duration}) for filtering and converting the data together with a meta table, containing inter alia the release version of the EPA ECOTOX data base are created as well.
 In a next step Chemical Abstracts Service (CAS) identifiers are used to query identifier data bases, including the Chemical Identifier Resolver (CIR) service \citep{nationalinstitutesofhealthnih_chemical_2019} and the Pubchem data base \citep{kim_pubchem_2016}.
 Chemical identifiers (CAS, InchI) and taxonomic names are then used to query chemical and taxonomic specific data bases (Table \ref{tab:data-base-additional}) to add information, such as chemical grouping or organism habitat preferences and global occurrence patterns to the local build.
-From here on the final Standartox data set is compiled, whereupon concentration and duration units are converted (see tables \ref{tab:conv-concentration}, \ref{tab:conv-duration}) in the process of harmonizing the data.
+
+\input{article/tables/data_base_tables.tex}
+
+From here on the final Standartox data set is compiled, whereupon concentration and duration units are converted in the process of harmonizing the data.
 Out of the unique 1229 concentration units in the EPA ECOTOX data base, Standartox retains only those (n = X) which are convertible to one of the following units: ug/l, g/m2, ppb, mg/kg, \% and flags the remaining as \textit{other}, resulting in 873473 out of 952625 test results.
@@
 Units such as per day (mg/kg/day) are generally excluded.
-Since in the current version the EPA ECOTOX data base lists 298 distinct entries, endpoints were refined to three endpoint groups NOEX, LOEX and XX50.
+Endpoints were refined to three endpoint groups NOEX, LOEX and XX50.
 The former two summarise no-observed-adverse-effect level and lowest-observed-adverse-effect levels, respectively.
 The latter includes various half maximal effective concentration (e.g.
-EC50, LC50, LD50 etc.) commonly used in ecotoxicology (Table \ref{tab:???????}).
+EC50, LC50, LD50 etc.) commonly used in ecotoxicology.
 Other endpoints are not included in Standartox.
-Beyond that, users can filter the data according to the CAS number, the concentration type (e.g.
+Beyond that, filters for the CAS number, the concentration type (e.g.
 active ingredient, formulation etc.), the chemical class (e.g.
 fungicides, metals etc.), taxa, organism habitat and region, test duration as well as effect (e.g.
-Mortality, Growth etc.) groups (See Table \ref{tab:app-parameters}) in one of the two front-ends.
-Along with the created Standartox data set, a catalog of it, listing all possible entries and value ranges, for continuous variables is created.
-As a last step, to guarantee the accuracy of the data, check scripts are run at the end.
+Mortality, Growth etc.) groups (See Table \ref{tab:app-parameters}) are created.
+Along with the compiled Standartox data set, a catalog, listing all distinct entries and value ranges, for categorical and continuous variables, respectively is created.
+As the end, to guarantee the accuracy of the data, check scripts are run.
 Single processing scripts are listed in figure \ref{fig:pipeline-tree}.
-The final Standartox table together with the catalog are exported.
+The final Standartox table together with the catalog is exported.
 Both the APP and the API access it.
-
-TODO: How does this fit?
-\begin{table}
-    \input{article/tables/data_refinement_counts.tex}
-    \caption{Reduction of data in the compilation process for Standartox}
-    \label{tab:data-refinement}
-\end{table}
 
@@
 \end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% CONTINUE HERE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 The API is built by using the R package plumber \citep{trestletechnologyllc_plumber_2018} which allows for the creation of RESTful APIs from R.
-Four API-endpoints (\textit{/catalog}, \textit{/filter}, \textit{/aggregate} and \textit{/meta}) that can be queried from the client side were created.
+The API is reachable via the URL \url{http://139.14.20.252} and port 8000.
+Four API-endpoints (\textit{/catalog}, \textit{/filter}, \textit{/aggregate} and \textit{/meta}) can be queried from the client side.
 The \textit{/catalog} API-endpoint returns a JavaScript Object Notation (JSON) file containing a catalog of possible filter parameters to choose from.
@@
 \subsubsection*{R-package}
-The R-package standartox contains two functions \textit{stx\_catalog()} and \textit{stx\_query()}.
-The former accesses the \textit{/catalog} API-endpoint and returns a R list object of possible filter parameters that can be used in \textit{stx\_query()}.
-The latter returns a R list of three tables (R data.frames) containing the filtered data set, the aggregated data set and a table with the meta information retrieved from the \textit{/meta} API-endpoint.
+The R-package standartox accesses the API through two functions \textit{stx\_catalog()} and \textit{stx\_query()}.
+The former accesses the \textit{/catalog} API-endpoint and returns a R list object of possible filter parameters (Table \ref{tab:app-parameters}) that can be used in \textit{stx\_query()}.
+The latter returns a R list of three tables (i.e.
+R data.frames) containing the filtered data set, the aggregated data set and a table with the meta information retrieved from the API endpoints \textit{/filter} and \textit{/meta}.
 
-\subsection*{Code availability}
-%%For all studies using custom code in the generation or processing of datasets, 
-%%a statement must be included here, indicating whether and how the code can be 
-%%accessed, including any restrictions to access. This section should also include 
-%%information on the versions of any software used, if relevant, and any specific 
-%%variables or parameters used to generate, test, or process the current dataset.
+\subsection*{Technical Validity}
+In order to guarantee appropriate conversion and harmonisation of units we take a random sample of the 50 most occurring concentration and six of the most occurring duration units and compare units created by Standartox to manually calculated unit conversions.
+This is done at the end of the processing pipeline and stops the it if not all units match.
 
-Standartox's code is distributed on three Github repositories below.
-All code can freely be accessed under the MIT License <CITATION>.
-Most of the code is written in R 3.6.1 and associated packages (compare Supplement: \ref{list:r-packages}).
-Some parts in PostgreSQL 9.6.1.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TODO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\iffalse
+- Endpoint: For the last parameter here, only be exclusive.
 
-\begin{itemize}
+TODO: How does this fit?
+\begin{table}
+    \input{article/tables/data_refinement_counts.tex}
+    \caption{Reduction of data in the compilation process for Standartox}
+    \label{tab:data-refinement}
+\end{table}
 
-\item \url{https://github.com/andschar/standartox-build} \newline
-Contains the code for the data acquisition and processing pipeline.
-
-\item \url{https://github.com/andschar/standartox-app} \newline
-Contains the code for the web application and the API.
-
-\item \url{https://github.com/andschar/standartox} \newline
-Contains the code for the R-package.
-
-\end{itemize}
+\fi
 

### 2019-11-18 article/tables/app-parameters.tex
Updates from Overleaf

\begin{table}[ht]
  \caption{Input parameters for the Standartox Web application and the R-package standartox.}
  \label{tab:app-parameters}
  \centering
\begin{tabular}{ll}
  \hline
  parameter & example \\ 
  \hline
  cas & 7758987, 2921882, 1912249 \\
  concentration\_type & active ingredient, formulation, total \\
  chemical\_class & fungicide, herbicide, insecticide \\
  taxa & Oncorhynchus mykiss, Rattus norvegicus, Daphnia magna \\
  habitat & marine, brackish, freshwater \\
  region & africa, america_north, america_south \\
  duration & 24, 96 \\
  effect & Mortality, Population, Growth \\
  endpoint & NOEX, XX50, LOEX \\
  vers & 20190912 \\
  \hline
\end{tabular}
\end{table}

### 2019-11-18 article/00_article.tex

@@
 
-\title{Standartox: A tool to derive exposure endpoints from chemical test results}
+\title{Standartox: A tool to derive exposure endpoints from ecotoxicological test results}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%% 01 Background
-\input{article/parts/01_background.tex}
-\pagebreak
+%%%%%%%%%%%%%%%%%%%%%% 01 Introduction
+\input{article/parts/01_introduction.tex}
 
-%%%%%%%%%%%%%%%%%%%%%% 02 Methods
-\input{article/parts/02_methods.tex}
-\pagebreak
+%%%%%%%%%%%%%%%%%%%%%% 02 Results and Discussion
+\input{article/parts/02_results_discussion.tex}
 
-%%%%%%%%%%%%%%%%%%%%%% 03 Data records
-\input{article/parts/03_data_records.tex}
-\pagebreak
+%%%%%%%%%%%%%%%%%%%%%% 03 Methods
+\input{article/parts/03_methods.tex}
 
-%%%%%%%%%%%%%%%%%%%%%% 04 Technical validation
-\input{article/parts/04_technical_validation.tex}
-\pagebreak
+%%%%%%%%%%%%%%%%%%%%%% 04 Conclusion
+\input{article/parts/04_conclusion.tex}
 
-%%%%%%%%%%%%%%%%%%%%%% 05 Usage notes
-\input{article/parts/05_usage_notes.tex}
+%%%%%%%%%%%%%%%%%%%%%% 05 Code Availability
+\input{article/parts/05_code_availability.tex}
+
+%%%%%%%%%%%%%%%%%%%%%% References
+% only reference from Zotero to avoid key conflicts
+% NOTE multiple refs would have to be written w/o space and separated by <,>
+\bibliographystyle{apalike}
+\bibliography{refs/references-standartox.bib}
 
@@
 
-\section*{Figures Legends}
-%%Figure should be referred to using a consistent numbering scheme through
-%%the entire Data Descriptor. For initial submissions, authors may choose
-%%to supply this document as a single PDF with embedded figures, but
-%%separate figure image files must be provided for revisions and accepted
-%%manuscripts. In most cases, a Data Descriptor should not contain more
-%%than three figures, but more may be allowed when needed. We discourage
-%%the inclusion of figures in the Supplementary Information \textendash{}
-%%all key figures should be included here in the main Figure section. 
-%%
-%%Figure legends begin with a brief title sentence for the whole figure
-%%and continue with a short description of what is shown in each panel,
-%%as well as explaining any symbols used. Legend must total no more
-%%than 350 words, and may contain literature references.
+\section*{Author Information}
 
-\section*{Tables}
-%%Tables supporting the Data Descriptor. These can provide summary information
-%%(sample numbers, demographics, etc.), but they should generally not
-%%be used to present primary data (i.e. measurements). Tables containing
-%%primary data should be submitted to an appropriate data repository. 
-%%
-%%Tables may be provided within the \LaTeX{} document or as separate
-%%files (tab-delimited text or Excel files). Legends, where needed,
-%%should be included here. Generally, a Data Descriptor should have
-%%fewer than ten Tables, but more may be allowed when needed. Tables
-%%may be of any size, but only Tables which fit onto a single printed
-%%page will be included in the PDF version of the article (up to a maximum
-%%of three).
+\section*{Ethics Declaration}
 
-%\begin{sidewaystable}
-%    \csvautotabular[respect all]{data/meta.csv};
-%    \caption{Meta variables table};
-%    \label{table:meta-variables};
-%\end{sidewaystable}
+\section*{Additional Information}
 
-\pagebreak
+\section*{Rights and permissions}
 
-\bibliographystyle{apalike}
-\bibliography{refs/references-standartox.bib} % only reference from Zotero to avoid key conflicts
-% NOTE multiple refs would have to be written w/o space and separated by <,>
-
-\section*{Data Citations}
-
-\pagebreak
-\section*{TODO}
-
-\begin{itemize}
-
-\item compare Etox-BAse output with PPDB
-
-\item plot example SSD and TU
-
-\item acute vs chronic
-
-\item OECD guidlines are not machine readable
-
-\item discuss flaws in column types (e.g some concentration entries contain '+' or '~' and are therefore not convertable to numeric) - This hampers large scale analysis greatly!
-
-\pagebreak
-\section*{MISC}
-
-\begin{figure}
-    % TODO make better graphic
-    \includegraphics[width=1\linewidth]{article/figures/scripts_dendrogram.png}
-    \caption{Processing pipeline}
-    \label{fig:pipeline-tree}
-\end{figure}
-
-\section*{not included citations}
-
-\item Put in citation \citep{hartung_chemical_2009} ?
-which claims that REACH won't meet their assumptions and say 54 million vertebrate animals are needed for tests that would cost €9.5 billion over the next ten years.
-I.e 20 times more animals, 6 times the costs in comparison to the official estimates
-
-\item "comparison of rate of change in the production and variety of pesticides, pharmaceuticals and other chemicals over the past four decades to CO2 nutrient pollution and other global change factors" \citep{bernhardt_synthetic_2017}
-
-\item worldwide decline in entomofauna: \citep{sanchez-bayo_worldwide_2019}
-
-\end{itemize}
-
-\subsubsection*{Future of the project}
-Since new versions of the EPA ECOTOX data base are published on a quarterly basis the data acquisition process is automated and new versions of the Standartox are scheduled to be released regularly.
-
-\input{article/parts/06_supplement.tex}
-\pagebreak
+%%%%%%%%%%%%%%%%%%%%%% 10 Supplement
+\input{article/parts/10_supplement.tex}
 

### 2019-11-18 article/parts/01_introduction.tex

--- 2019-11-11 article/parts/01_background.tex
@@
-\section*{Background \& Summary [max 700 words]}
+\section*{Introduction}
 
@@
 These tests are conducted by different research groups all over the world who publish the results independently.
-Up to now, only few initiatives, such as the United States Environmental Protection Agency ECOTOXicology knowledgebase (EPA ECOTOX) \citep{elonen_ecotoxicology_2018}, the German Umweltbundesamt's Information System Ecotoxicology and Environmental Quality Targets (ETOX) \citep{umweltbundesamt_etox_2019} or the Pesticide Property Data Base (PPDB) aiming to consolidate ecotoxicological test results, exist.
+Up to now, only few initiatives, such as the United States Environmental Protection Agency ECOTOXicology knowledgebase (EPA ECOTOX) \citep{elonen_ecotoxicology_2018}, the German Umweltbundesamt's Information System Ecotoxicology and Environmental Quality Targets (ETOX) \citep{umweltbundesamt_etox_2019} or the Pesticide Property Data Base (PPDB) \citep{lewis_international_2016} aiming to consolidate ecotoxicological test results, exist.
 The former two collect test results and hence only partly check for quality whereas the latter aims for quality, yet being confined to specific classes of chemicals, in this case pesticides.
 An overall holistic approach in processing the large number of ecotoxicological test results is lacking.
-Therefore, we suggest Standartox, a tool and data base that aims to overcome these paucities by steadily incorporating the ever-growing number of test results in an automated process workflow that ultimately leads to single aggregated test results for specific chemical-organism test combinations, representing the toxicity of a chemical.
+Therefore, we suggest Standartox, a tool and data base that aims to overcome these paucities by steadily incorporating the ever-growing number of test results in an automated process workflow that ultimately leads to single aggregated exposure endpoints for specific chemical-organism test combinations, representing the toxicity of a chemical.
 Automated approaches are increasingly important because a growing number of chemicals such as pharmaceuticals, pesticides and synthetic hormones, amongst others are in daily use all over the world with insufficient knowledge about possible adverse effects on the environment.
@@
 thresholds beyond which irreversible state shifts may occur \citep{steffen_anthropocene_2007}.
-Bernhardt et al.
+Recent studies showed that, in defiance of many regulation efforts \citep{schafer_future_2019} chemicals can still cause strong adverse effects on organisms, such as birds \citep{hallmann_declines_2014} or fish \citep{yamamuro_neonicotinoids_2019} for example.
 \citet{bernhardt_synthetic_2017} argue further that the knowledge gap how chemicals effect populations and communities and hence ecosystem functions and ecosystem services, would also impede society's ability to accomplish the Sustainable Development Goals of the United Nations.
 In Europe alone some 100,000 chemicals are estimated to be in current use, whereof 30,000 are produced in quantities larger than one ton per year \citep{breithaupt_costs_2006}.
-Data on toxicity tests for a lot of chemicals, especially newly emerging ones remains scarce and if existent, the information is represented sparsely and in inconsistent formats \citep{gessner_fostering_2016}.
+Data on toxicity tests for a lot of chemicals, especially newly emerging ones remains scarce and if existent, the information is represented sparsely and in inconsistent formats \citep{gessner_fostering_2016, hitchcock_improving_2018}.
 Therefore, tools such as Standartox are a required step adding a layer of abstraction satisfying the needs of modern environmental risk assessment demands.
@@
 Besides aggregating ecotoxicological test results, Standartox provides a concise overview of the status of tested chemicals, allowing the identification of potential research gaps.
-Standartox comes with two front-ends, an R \citep{rcoreteam_language_2017} package and a web tool, adding convenience structures and largely reducing work time for users.
+Standartox comes with two front-ends,  a web application (APP) and a R \citep{rcoreteam_language_2017} package, which accesses a application programming interface (API) providing convenience structures and thereby largely reducing processing time for users.
 
@@
 \iffalse
+
+include \citep{sanchez-bayo_worldwide_2019} % worldwide decline in enthomofauna
+
+\item Put in citation \citep{hartung_chemical_2009} ?
+which claims that REACH won't meet their assumptions and say 54 million vertebrate animals are needed for tests that would cost €9.5 billion over the next ten years.
+I.e 20 times more animals, 6 times the costs in comparison to the official estimates
 
@@
 
-    \item reproducible
+    \item reproducible research % TODO mention much more often!
 

### 2019-11-18 article/parts/05_code_availability.tex

--- 2019-11-06 article/parts/02_methods.tex
@@
-%%The Methods should include detailed text describing any steps or procedures 
-%%used in producing the data, including full descriptions of the experimental 
-%%design, data acquisition assays, and any computational processing (e.g. 
-%%normalization, image feature extraction). Related methods should be grouped 
-%%under corresponding subheadings where possible, and methods should be described 
-%%in enough detail to allow other researchers to interpret and repeat, if required, 
-%%the full study. Specific data outputs should be explicitly referenced via data 
-%%citation (see Data Records and Data Citations, below). Authors should cite 
-%%previous descriptions of the methods under use, but ideally the method 
-%%descriptions should be complete enough for others to understand and reproduce 
-%%the methods and processing steps without referring to associated publications. 
-%%There is no limit to the length of the Methods section.
-\section*{Methods}
-
-Standartox consists of three parts of software explained in detail below.
-Firstly a data acquisition and preparation processing pipeline which downloads quarterly released new versions of the EPA ECOTOX, upon which Standartox is built.
-Secondly, Standartox is accessible via a web application (APP) and an application programming interface (API).
-Lastly, the R package standartox constitutes an easily usable way of accessing the tool via the API.
-All the software is written in R \citep{rcoreteam_language_2017} and SQL scripts.
-
-\subsection*{Data acquisition and preparation pipe line}
-The EPA ECOTOX, which is released on a quarterly basis (current release: 12.09.2019) is checked for new releases in monthly intervals and downloaded as soon as a new release is published.
-Subsequently the EPA ECOTOX is built into a PostgreSQL data base and SQL functions for further processing are added and known errors are corrected.
-Lookup tables for filtering and converting the data are uploaded.
-Thereafter, a meta data table is created in the data base, comprising the version number of the EPA ECOTOX.
-In a next step Chemical Abstracts Service (CAS) identifiers are used to query identifier data bases, including the Chemical Identifier Resolver (CIR) service <CITATION> and the Pubchem data base <CITATION>.
-Chemical identifiers (CAS, InchI) and taxonomic names are then used to query chemical and taxonomic specific data bases (Table \ref{tab:data-base-additional}) to add information to the local data base.
-The data is aggregated and unique values are derived.
-<PUT LOOKUPTABLES HERE>.
-From here on the Standartox data set is compiled, whereupon concentration and duration units are converted (see tables \ref{tab:conv-concentration}, \ref{tab:conv-duration}) in order to harmonize data.
-A catalog object, listing unique values for categorical variables and ranges for continuous variables, respectively is created.
-To guarantee the accuracy of the data check scripts are run at the end.
-Single processing scripts are listed in figure \ref{fig:pipeline-tree}.
-The final Standartox table together with the catalog are exported.
-Both the APP and the API access these objects.
-
-\subsection*{Web application (APP) and Application Programming Interface (API)}
-
-The web application and the API load the compressed serialized Standartox data into memory and allow a client to interact with it.
-The client can then call the functions stx\_filter() and stx\_aggregate() to filter and aggregate the data according to specific parameters (Table \ref{tab:app-parameters}).
-
-\subsubsection*{Web application (APP)}
-The interactive APP was built in R, using the shiny web application framework \citep{chang_shiny_2018} which runs with the help of a shiny server \citep{HOW-TO-CITE-SHINY-SERVER} on www.standartox.uni-landau.de.
-It is set of R functions that allow a user to filter and aggregate the data.
-The data can then be downloaded.
-
-\subsubsection*{Application programming interface (API)}
-The API is built by using the R package plumber \citep{trestletechnologyllc_plumber_2018} which allows for the creation of RESTful APIs from R.
-Four API-endpoints ('/catalog', '/filter/rds', 'aggregate' and 'meta') that can be queried from the client side were created.
-The \textit{catalog} API-endpoint returns a JavaScript Object Notation (JSON) file containing a catalog of possible filter parameters to choose from.
-The \textit{filter} returns the filtered Standartox table as a compressed serialized binary file \citep{klik_fst_2019}, to reduce size and allow for fast user queries.
-The \textit{aggregate} API-endpoint returns a R function to the client allowing the client the aggregation of the filtered data.
-Lastly, the \textit{meta} API-endpoint returns a JSON file with meta information, such as the timestamp of the request and the used Standartox version.
-The API is designed to be used with the R-package standartox and therefore uses serialization methods specific to R (rds() from the R package base and fst() from the package fst).
-
-\begin{figure}
-    \includestandalone[scale=0.5]{article/tikz/pipeline-organigram}
-    %\input[scale = 0.5]{article/tikz/pipeline-organigram.tex}
-    \caption{Organigram of Standartox.}
-    \label{fig:stx-organigram}
-\end{figure}
-
-\subsubsection*{R-package}
-The R-package standartox contains two functions \textit{stx\_catalog()} and \textit{stx\_query()}.
-The former accesses the \textit{catalog} API-endpoint and returns a R-list of possible filter parameters that can be used in \textit{stx\_query()}.
-The latter returns a R list of three tables (R data.frames) containing the filtered data set, the aggregated data set and a table with the meta information retrieved from the \textit{meta} API-endpoint.
-
-\subsection*{Code availability}
 %%For all studies using custom code in the generation or processing of datasets, 
@@
 %%variables or parameters used to generate, test, or process the current dataset. 
+\section*{Code Availability}
 
@@
 All code can freely be accessed under the MIT License <CITATION>.
-Most of the code is written in R 3.6.1 and associated packages (compare Supplement: \ref{list:r-packages}).
-Some parts in PostgreSQL 9.6.1.
+Most of the code is written in R 3.6.1 and associated packages (compare Supplement: \ref{list:r-packages}) as well as some parts in SQL for PostgreSQL 9.6.1.
 
@@
 \end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% OLD %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\iffalse
-\textbf{OLD: Standartox aggregates the test results according to chosen filters in a two step process.
-Firstly the filtered test results are aggregated by the CAS number, the chosen taxon and the selected test duration.
-Secondly, the returned data is then aggregated by the CAS number.
-The former can't be influenced by the user and calculates either the minimum or the median depending on the amount of results to aggregate (n <= 2: minimum, if n > 2: median).
-Thereof the second step calculates the minimum, the maximum, the median, the geometric mean, or the arithmetic mean as an aggregate.}
-\fi

### 2019-11-22 article/parts/01_introduction.tex
corrections co-authors

@@
 \section*{Introduction}
+An increasing number of chemicals such as pharmaceuticals, pesticides and synthetic hormones are in daily use all over the world.
+In Europe alone, some 100,000 chemicals are estimated to be in current use, whereof 30,000 are produced in quantities larger than one ton per year \citep{breithaupt_costs_2006}.
+Except for pesticides that are released into the environment deliberately, most chemicals enter the environment as a result of their use through different paths (e.g.
+atmospheric emission and deposition or discharge through wastewater) \citep{schwarzenbach_challenge_2006}.
+In the environment, chemicals can adversely affect populations and communities and in turn related ecosystem functions \citep{rhind_anthropogenic_2009, clements_community_2009, hallmann_declines_2014, barracaracciolo_pharmaceuticals_2015, johnston_review_2015, hallmann_more_2017, sanchez-bayo_worldwide_2019}.
+Ultimately, this may compromise natures contribution to human well-being, for example the ecosystem services clean drinking and irrigation water, food and climate regulation \citep{peters_review_2013, vandersluijs_neonicotinoids_2013, barracaracciolo_pharmaceuticals_2015, yamamuro_neonicotinoids_2019}.
+Pollution with man-made chemicals has been identified as one of three major environmental problems for which research gaps hamper the derivation of planetary boundaries, i.e.
+thresholds beyond which irreversible state shifts may occur \citep{steffen_anthropocene_2007, steffen_planetary_2015}.
+\citet{bernhardt_synthetic_2017} argue that the knowledge gap how chemicals affect populations, communities and in turn ecosystem functions and services, may impede the accomplishment of the Sustainable Development Goals of the United Nations.
+Even highly regulated chemical compounds such as pesticides have been shown to cause strong adverse effects on organisms, such as birds \citep{hallmann_declines_2014} or fish \citep{yamamuro_neonicotinoids_2019}, questioning the current regulation efforts \citep{schafer_future_2019}.
+To evaluate the risks from chemicals for ecosystems data on their toxicity, which is typically produced in standardised ecotoxicological laboratory tests is required.
+For example, \citet{morrissey_neonicotinoid_2015} used ecotoxicological test results from 49 insect and crustacean species to evaluate the effect of neonicotinoids, an insecticide class on the aquatic ecosystem.
+Also, \citet{malaj_organic_2014} compiled intera alia experimental toxicity test results for 223 various chemical compounds to assess the health of freshwater ecosystems in Europe.
+Similarly, permissible environmental concentrations are often derived from these test data, typically by a combination with safety factors to account for uncertainties.
+Given that data sets with large numbers of analysed chemicals are becoming increasingly available, ecotoxicity data are pivotal for regulatory risk assessment and science.
+To date, only few initiatives exist that aim to create a public resource of harmonized and quality controlled ecotoxicological data, such as the United States Environmental Protection Agency ECOTOXicology knowledgebase (EPA ECOTOX) \citep{elonen_ecotoxicology_2018}, the German Environmental Agency's Information System Ecotoxicology and Environmental Quality Targets (ETOX) \citep{umweltbundesamt_etox_2019} or the Pesticide Property Data Base (PPDB) \citep{lewis_international_2016}.
+The former two compile all available test results into a database.
+However, for many chemicals, multiple ecotoxicity values are available for the same test organism, which can vary strongly.
+Hence, these databases can contain multiple entries for the same test without quality information and with different units, hampering reproducible science if different values are selected for studies.
+The PPDB database is confined to pesticides and provides harmonised and quality controlled data for a few selected test organisms, thereby ignoring vast amounts of data for other species and chemicals.
+Moreover, data analyses often require links to additional data resources, for example to append additional chemical and species information (e.g.
+chemical properties, habitat of species), which calls for more automated procedures.\\
 
-The risk of adverse effects of chemicals on the environment is globally assessed with the help of ecotoxicological tests that are conducted on different organism groups, ideally representing a specific biocoenosis and associated biotopes.
-These tests are conducted by different research groups all over the world who publish the results independently.
-Up to now, only few initiatives, such as the United States Environmental Protection Agency ECOTOXicology knowledgebase (EPA ECOTOX) \citep{elonen_ecotoxicology_2018}, the German Umweltbundesamt's Information System Ecotoxicology and Environmental Quality Targets (ETOX) \citep{umweltbundesamt_etox_2019} or the Pesticide Property Data Base (PPDB) \citep{lewis_international_2016} aiming to consolidate ecotoxicological test results, exist.
-The former two collect test results and hence only partly check for quality whereas the latter aims for quality, yet being confined to specific classes of chemicals, in this case pesticides.
-An overall holistic approach in processing the large number of ecotoxicological test results is lacking.
-Therefore, we suggest Standartox, a tool and data base that aims to overcome these paucities by steadily incorporating the ever-growing number of test results in an automated process workflow that ultimately leads to single aggregated exposure endpoints for specific chemical-organism test combinations, representing the toxicity of a chemical.
-Automated approaches are increasingly important because a growing number of chemicals such as pharmaceuticals, pesticides and synthetic hormones, amongst others are in daily use all over the world with insufficient knowledge about possible adverse effects on the environment.
-Generally, chemicals are brought to the environment deliberately - as it is the case of pesticides, or arrive therein as a byproduct from other processes (e.g.
-atmospheric emissions or wastewater) \citep{schwarzenbach_challenge_2006}.
-In turn, ecosystems provide essential services to human societies such as drinking and irrigation water, food and climate regulation.
-These services are products of different ecosystem functions that crucially depend on the integrity of the populations and communities that drive these ecosystem functions.
-However, besides habitat degradation, climate change and nutrient enrichment, pollution with man-made chemicals threatens these populations and communities in various ways which are currently not fully understood \citep{steffen_anthropocene_2007}.
-Pollution with man-made chemicals was indeed identified as one of three major environmental problems for which research gaps hamper the derivation of planetary boundaries, i.e.
-thresholds beyond which irreversible state shifts may occur \citep{steffen_anthropocene_2007}.
-Recent studies showed that, in defiance of many regulation efforts \citep{schafer_future_2019} chemicals can still cause strong adverse effects on organisms, such as birds \citep{hallmann_declines_2014} or fish \citep{yamamuro_neonicotinoids_2019} for example.
-\citet{bernhardt_synthetic_2017} argue further that the knowledge gap how chemicals effect populations and communities and hence ecosystem functions and ecosystem services, would also impede society's ability to accomplish the Sustainable Development Goals of the United Nations.
-In Europe alone some 100,000 chemicals are estimated to be in current use, whereof 30,000 are produced in quantities larger than one ton per year \citep{breithaupt_costs_2006}.
-Data on toxicity tests for a lot of chemicals, especially newly emerging ones remains scarce and if existent, the information is represented sparsely and in inconsistent formats \citep{gessner_fostering_2016, hitchcock_improving_2018}.
-Therefore, tools such as Standartox are a required step adding a layer of abstraction satisfying the needs of modern environmental risk assessment demands.
-Standartox aggregates ecotoxicological test results in a standardized and reproducible way, providing the basis for the derivation of risk indicators such as Species Sensitivity Distributions (SSD) \citep{posthuma_species_2002} and Toxic Units (TU), which represent two prominent concepts to assess effects on organisms in ecotoxicology \citep{kefford_definition_2011, schafer_effects_2011}.
+We developed Standartox, a tool and database that aims to overcome the limitations of other databases by continuously incorporating the ever-growing number of test results in an automated process workflow that ultimately leads to single aggregated data point for a specific chemical-organism test combination, representing the toxicity of a chemical.
+Hence, Standartox provides the basis for reproducible science and combines information from different sources to simplify the derivation of more advanced risk indicators such as Species Sensitivity Distributions (SSD) and Toxic Units (TU), which represent two prominent concepts to assess effects on organisms in ecotoxicology \citep{posthuma_species_2002, kefford_definition_2011, schafer_effects_2011}.
 Besides aggregating ecotoxicological test results, Standartox provides a concise overview of the status of tested chemicals, allowing the identification of potential research gaps.
@@
 
-include \citep{sanchez-bayo_worldwide_2019} % worldwide decline in enthomofauna
-
 \item Put in citation \citep{hartung_chemical_2009} ?
@@
 
-Curently \citep{schafer_future_2019}
+For dicussion:
+Showed general decline in biodiversity and biomass, but no or not much causes: \citep{butchart_global_2010, hallmann_more_2017, seibold_arthropod_2019}
 
-include \citep{yamamuro_neonicotinoids_2019} and \citep{seibold_arthropod_2019} as a new prominent examples for effects on the ecosystem
+Direct effect of chemicals:
+include \citep{yamamuro_neonicotinoids_2019} % ecosystem services, food webs
+
+- gerneral stuff about small water bodies: \citep{riley_small_2018}
 

### 2019-11-22 article/tables/api-endpoints.tex
corrections co-authors

\begin{table}[ht]
  \caption{Application programming interface (API) endpoints, Http methods, Requests and Response objects. Object is abbreviated with obj.}
  \label{tab:api-endpoints}
  \centering
\begin{tabular}{cccc}
  \hline
  Endpoint & Http Method & Request & Response \\ 
  \hline
  /catalog & POST & \makecell{Standartox \\ version string} & \makecell{Catalog object of \\ possible filter parameters \\ JSON obj.}    \\[0.5cm]
  /filter & POST & \makecell{Standartox filter \\ parameters (Table \ref{tab:app-parameters})} & \makecell{Filtered Standartox data \\ serialized R data.frame \\ R fst:: \citep{klik_fst_2019} obj.}   \\[0.5cm]
  /aggregate & GET & / & \makecell{Aggregate function \\ serialized R function \\ R .rds \citep{rcoreteam_language_2017} obj.}    \\[0.5cm]
  /meta & POST & \makecell{Standartox \\ version string} & \makecell{Meta data on request \\ JSON obj.}    \\
  \hline
\end{tabular}
\end{table}

### 2019-11-22 article/tables/app-parameters.tex
corrections co-authors

@@
   \centering
-\begin{tabular}{ll}
+\begin{tabular}{cc}
   \hline
@@
   \hline
-  cas & 7758987, 2921882, 1912249 \\
-  concentration\_type & active ingredient, formulation, total \\
-  chemical\_class & fungicide, herbicide, insecticide \\
-  taxa & Oncorhynchus mykiss, Rattus norvegicus, Daphnia magna \\
-  habitat & marine, brackish, freshwater \\
-  region & africa, america_north, america_south \\
-  duration & 24, 96 \\
-  effect & Mortality, Population, Growth \\
-  endpoint & NOEX, XX50, LOEX \\
+  cas & 7758987, 2921882, 1912249   \\[0.5cm]
+  concentration\_type & active ingredient, formulation, total   \\[0.5cm]
+  chemical\_class & fungicide, herbicide, insecticide   \\[0.5cm]
+  taxa & Oncorhynchus mykiss, Rattus norvegicus, Daphnia magna   \\[0.5cm]
+  habitat & marine, brackish, freshwater   \\[0.5cm]
+  region & africa, america\_north, america\_south   \\[0.5cm]
+  duration & 24, 96   \\[0.5cm]
+  effect & Mortality, Population, Growth   \\[0.5cm]
+  endpoint & NOEX, XX50, LOEX   \\[0.5cm]
   vers & 20190912   \\

### 2019-11-28 article/code/standartox_example.tex

--- 2019-10-29 article/parts/05_usage_notes.tex
@@
-\section*{Usage Notes}
-%%Brief instructions that may help other researchers reuse these dataset.
-%%This is an optional section, but strongly encouraged when helpful
-%%to readers. This may include discussion of software packages that
-%%are suitable for analyzing the assay data files, suggested downstream
-%%processing steps (e.g. normalization, etc.), or tips for integrating
-%%or comparing this with other datasets. If needed, authors are encouraged
-%%to provide code, programs, or data processing workflows when they may help 
-%%others analyse the data. We encourage authors to archive related code in 
-%%a DOI-issuing archive when possible, but code may also be supplied as 
-%%supplementary information files. 
-%%
-%%For studies involving privacy or safety controls on public access
-%%to the data, this section should describe in detail these controls,
-%%including how authors can apply to access the data, and what criteria
-%%will be used to determine who may access the data, and any limitations
-%%on data use.
-
-\subsection*{Web application}
-
-In Standartox web application (app) users can define several input parameters \ref{tab:inputs} through the app's graphical user interface (GUI) in order to filter the data specifically and perform the aggregation.
-The aggregated result is then shown as a table as well as an overview plot and can then be downloaded as a 'filtered' and an 'aggregated' data set \ref{fig:standartox_app}.
-
-\begin{figure}
-    \includegraphics[width=1\linewidth]{article/figures/screenshot_standartox_app.png}
-    \caption{Standartox web application}
-    \label{fig:standartox_app}
-\end{figure}
-
-\subsection*{R package}
-
-The R-package standartox consists of the two functions stx\_catalog() and stx\_query().
-The former allows the retrieval of a catalog of possible parameters that can be used as an input for the latter.
-stx\_query() fetches the toxicity values from the data base.
-It returns a R list object following the app containing a 'filtered' and 'aggregated' R data.frame object.
-Additionally ...................
-
-\input{article/scripts/standartox_example.tex}
-
-The \etoxbase{} is designed to support CRA of ecotoxicologists.
-With an increased amount of available ecotoxicological test data, it becomes fundamental to provide and distribute such information in reasonable formats, meaning accessible for humans as well as machines.
-Currently CRA often relies on test results of a few, well tested standard organism, such as \textit{Daphnia magna}, \textit{Pimephales promelas}, \textit{Selenastrum capricornutum}, although researchers have conducted experiments on a much greater variety of organisms (e.g.
-the 926,108 test results collected in the EPA ECOTOX data base).
-In order to locate ecotoxicological test data, researchers can rely on individual test results reported in publications <CITE>, on compiled and published data sets \citep{malaj_organic_2014} or on data bases such as the Pesticide Property Data Base (PPDB) \citep{lewis_international_2016}.
-Each data source has its limitations though: Collecting toxicity information from individual publications is laborious and using already compiled data from publications or data bases is often limited to a specific groups of chemicals, such as pesticides, biozides etc.
-The \etoxbase{}, making use of the EPA ECOTOX data base, the largest publicly available collection of ecotoxicological test results, aims to facilitate the data retrieval process by providing adequate aggregates (\ecfifty{}, NOEC, LOEC) of the tests results.
-These are then used for the calculation of SSDs, TUs or other risk indicators.
-SSDs try to extrapolate effects of chemicals towards organisms from species to a community level by fitting a statistical distribution to toxicity test results on organisms of different trophic levels in order to derive concentrations at which a specific fraction of organisms are effected (often at 5\% - HC\textsubscript{5} \citep{posthuma_species_2002}.
-Figure \ref{fig:ssd-isoproturon} shows an example of a SSD and the derived HC\textsubscript{5} value (red dassed line) created for the herbicide Isoproturon.
-TUs are another indicator of effects of chemicals on organisms.
-They are calculated by dividing a measured concentration (c) by an \ecfifty{} value (cf.
-Equation \ref{eq:tu}).
-
-\begin{equation}
-    TU_i = \frac{c_i}{\ecfifty}
-    \label{eq:tu}
-\end{equation}
-
-\begin{figure}[h!]
-    \dentering
-    \includegraphics[width=1\linewidth]{article/figures/ssd2_boot.png}
-    \caption{SSD plot showing the susceptibility of organisms towards the herbicide Isoproturon.
-The red line represents the fit and the red dashed line marks the HC\textsubscript{5} value, which is a common measure in CRA.
-The legend denotes common organism groups used in ecotoxicology.}
-    \label{fig:ssd-isoproturon}
-\end{figure}
-
-\subsection*{Filter and Aggregate}
-
-In the application's GUI, the user can choose among different parameters to filter and aggregate the test data accordingly \ref{tab:inputs}.
-Percentage values next to inputs show the proportion of the respective option (e.g.
-\textit{Active Ingredient - 43\%} indicates that 43\% of the data are Active Ingredients).
-The user can filter for compound-specific filters, including the CAS number, the concentration type (e.g.
-Active Ingredient, Formulation), the chemical class (e.g.
-Insecticides, Metals) and taxon-specific filters, including common taxonomic groups (e.g.
-Daphniidae, Algae), the organism habitat (e.g.
-freshwater) and the continent (e.g.
-Europe, Asia).
-Regarding the test parameters the user can choose test durations (in hours), Effect groups (e.g.
-Mortaliy, Population, Growth) and Endpoints (\ecfifty, EC\textsubscript{10}, LOEC or NOEC).
-Besides that, two cleaning steps can also be chosen.
-Firstly the application allows to exclude test results exhibiting concentrations that are higher than the actual water solubility of the respective chemical at \ang{20} C.
-Secondly the user can choose to exclude outliers (cf.
-Table \ref{tab:scripts-app}).
-The outliers are selected as values that exceed lower (0.25) and the upper (0.75) quartile by 1.5 times the inter-quartile range.
-As aggregates the user can choose the minimum, the maximum, the median, the geometric mean or the arithmetic mean.
-All the steps described in this section are selectable in the GUI and are executed at each click by the application.
-The drop down menu `Download data` allows to download a filtered data set as well as an aggregated data set.
-
-\subsection*{Limitations}
-
-Generally it has to be stated that the returned values are subject to change, as new tests will be constantly incorporated into the \etoxbase{}.
-However, we expect values of well tested chemical-organism combinations to be relatively stable.
-The returned values are certainly influenced by individual tests characteristics, such as pH, temperature or conductivity to just name a few, which influences the effect relationship between chemicals and organisms \citep{rosenkrantz_influence_2013,li_temperature_2011}.
-Unfortunately the EPA ECOTOX data base has only rare records on test parameters, which prevents incorporating such parameters in an aggregation <proportions of available data: Temperature:77\%, pH:56\%, hardness:27\%, dissolved oxygen:18\%, Alkalinity:15\%, salinity:9\%, others:<=5\%>.
-However, using the median or the geometric mean as an aggregate allows for an adequate estimation of the central tendency of the toxicity of a chemical towards an organism group.
-The geometric mean is recommended, as it is in comparison to the arithmetic mean considerably less influenced by outliers and is suitable for skewed data.
-Also, the geometric mean is preferable in contrast to the median, since the median completely ignores the influence of large or small values, making it unreliable for small data sets <CITATION GEOMETRIC MEAN>.
-
-Other initiatives, with different, though partly overlapping aims accumulating existing ecotoxicological test data are published online.
-The PPDB provides validated ecotoxicological test results for pesticides, narrowing chemicals to this group \citep{lewis_international_2016}.
-The Network of reference laboratories, research centres and related organisations for monitoring of emerging environmental substances (NORMAN) focuses on assembling river basin specific pollutants\citep{von_der_ohe_new_2011}.
-The EnviroTox data base \href{https://envirotoxdatabase.org/} makes also use of the \epa{} data in order to derive Predicted no-effect concentrations (PNEC) \citep{health_and_environmental_sciences_institute_hesi_envirotox_2019}.
-In comparison to the \etoxbase, these examples partly rely on the same data, yet not aiming to derive \ecfifty, LOEC and NOEC from it automatically.
-\subsection*{Other data bases}
-
-Ecotox tool for PENEC
-Lara's work for regulatory thresholds
-
-greater and greater demand to numerically  analyse large amount of ecotoxicological data.
-
-Other data bases are also thought of. Compare Sascha Bub's work.
-
-\subsection*{Identify chemicals to be tested}
-
-\begin{figure}
-    \includegraphics[width=1\linewidth]{article/figures/heatmap_tests_n.png}
-    \caption{30 most tested chemicals in EPA ECOTOX.}
-    \label{fig:standartox_ppdb_diff}
-\end{figure}
-
-There is a clear need in ecotoxicology for harmonised approaches to collect ecotoxicological data as other data base compilation attempts recently published show.
-
-Hitchock et al.
-\citet{hitchcock_improving_2018} argue for an increased awareness concerning newly published ecotoxicological test results to facilitate meta analyses across studies.
-Important statistical parameters, such as mean estimates, variances ans sample sizes.
-Likewise, state variables, including medium temperature, pH, mineral contents or organism age and sex should be reported (if not in the text, in an online repository).
-Standartox doesn't include such information mainly because the EPA ECOTOX data base reports only a few of these variables.
-A future research aim could be to use machine learning techniques to retrieve this information from the referenced articles.
-
-Agrochemicals contribute to a decline in biodiversity \citep{schafer_future_2019}
-contrary to EU pesticide regulation
-
-Current risk assessment or pesticide regulation consists of Stadard Toxicity Tests and Safety factors. They are not sufficient
-
-tiered framework critizized, especiall the two main assumptions:
-(1) 1st tier provides and overly protective measure
-(2) higher tiers provide more ecologically relevant predictions
-
-"Inacurate predictions of exposure and effect" \citep{schafer_future_2019}
-
-In their policy brief for a future risk assessment Schäfer et al.
-\citet{schafer_future_2019} propose to include a greater ecological context, meaning that not only single but communities need to evaluated and a landscape-oriented contemplation should be envisaged inter alia.
-Standartox partly follows such a route in making ecotoxicological data more easily available by standardizing ?????
-(1st tier) and integrating habitat and occurrence data, allowing for a more geographically detailed refinement.
-
-Hallmann et al. \citet{hallmann_declines_2014} showed a decline of insectivorous birds associated with neonicotinoid usage.
-
-Fungicides paper zitieren?
-
-The units are denoted as ppm or ppb which comprises ug/l and mg/l
-
-List the units which are kept.
-
-Thereby 'food', soil, ai and other information are lost but these are coded in other variables anyway. And they don't belong in a unit variable. Characteristic for ecotoxicology
-
-units such as per day (mg/kg/day) are generally excluded.
-
-Standartox follows such a route in trying to synthesize information from a plethora of studies.
-
-\pagebreak
+\begin{lstlisting}[
+    caption = {Sample code to access the Standartox data base through the API with the help of the R-package.
+\textit{stx\_catalog()} returns a catalog of possible filter and aggregation parameters.
+\textit{stx\_query()} returns the Standartox object, a list of the filtered and the aggregated data as well as a meta data entry.},
+    label = {listing:standartox-example}]
+# install
+install.packages('remotes')
+remotes::install_github('andschar/standartox') # package not yet on CRAN
+# retrieve catalog    
+require(standartox)
+catal = stx_catalog()
+# retrieve data
+l = stx_query(cas = '1071-83-6',
+                endpoint = 'XX50',
+                taxa = 'Oncorhynchus',
+                duration = c(24, 120))
+\end{lstlisting}

### 2019-11-28 article/parts/02_results_discussion.tex

@@
 \section*{Results and Discussion}
-%%Brief instructions that may help other researchers reuse these dataset.
-%%This is an optional section, but strongly encouraged when helpful
-%%to readers. This may include discussion of software packages that
-%%are suitable for analyzing the assay data files, suggested downstream
-%%processing steps (e.g. normalization, etc.), or tips for integrating
-%%or comparing this with other datasets. If needed, authors are encouraged
-%%to provide code, programs, or data processing workflows when they may help 
-%%others analyse the data. We encourage authors to archive related code in 
-%%a DOI-issuing archive when possible, but code may also be supplied as 
-%%supplementary information files. 
-%%
-%%For studies involving privacy or safety controls on public access
-%%to the data, this section should describe in detail these controls,
-%%including how authors can apply to access the data, and what criteria
-%%will be used to determine who may access the data, and any limitations
-%%on data use.
-Standartox is a tool that allows to filter and aggregate ecotoxicological test results to derive exposure endpoints for individual chemicals.
-It can either be accessed through the APP (Figure \ref{fig:app}) or the API together with the R-package standartox (Listing \ref{listing:standartox-example}).
-Both access pathways allow for the same filters and aggregation methods (Table \ref{tab:app-parameters}) to be applied on the same data set.
-Using the APP, the user can download the resulting data sets as a comma-separated values (csv) file whereas in the API the data can directly be loaded in R.
+Standartox is a tool that allows to filter and aggregate ecotoxicological test results to derive single exposure endpoints for individual chemicals.
+It can either be accessed through the APP (Figure \ref{fig:app}) or the API \ref{tab:api-endpoints} together with the R-package standartox (Code example \ref{listing:standartox-example}).
+Both access pathways allow for the same filters and aggregation methods (Table \ref{tab:app-parameters}) to be applied and access the same data set.
+Using the APP, the user can download the resulting data sets as a comma-separated values (csv) file whereas by using the API, the data can directly be loaded in R.
 Different parameters in order to filter the test data, such as compound-specific parameters, including the CAS number, the concentration type (e.g.
 Active Ingredient, Formulation) or the chemical class (e.g.
-Insecticides, Metals) as well as taxon-specific parameters, including common taxonomic groups (e.g.
-Daphniidae, Algae), the habitat (e.g.
+insecticides, metals) as well as taxon-specific parameters, including common taxonomic groups (e.g.
+Daphniidae, algae), the habitat (e.g.
 freshwater) or the regional occurrence (e.g.
@@
 Furthermore, users can refine the results to specific test durations (in hours), effect groups (e.g.
-Mortality, Population, Growth) and endpoints (EC\textsubscript{50}, LOEC or NOEC).
+mortality, population, growth) and endpoints (EC\textsubscript{50}, LOEC or NOEC).
 As the method of aggregation users can choose the minimum, the maximum, or the geometric mean.
 To provide an easy overview of the data set, percentage values for each parameter entry are shown in the APP in the selection menu as well as by using the R package through querying the \textit{/catalog} API-endpoint.
+
+%% Add units part here
 
@@
 \pagebreak
-\begin{lstlisting}[
-    caption = {Sample code to access the Standartox data base through the API with the help of the R-package.
-\textit{stx\_catalog()} returns a catalog of possible filter and aggregation parameters.
-\textit{stx\_query()} returns the Standartox object, a list of the filtered and the aggregated data as well as a meta data entry.},
-    label = {listing:standartox-example}]
-# install
-install.packages('remotes')
-remotes::install_github('andschar/standartox') # package not yet on CRAN
-# retrieve catalog    
-require(standartox)
-catal = stx_catalog()
-# retrieve data
-l = stx_query(cas = '1071-83-6',
-                endpoint = 'XX50',
-                taxa = 'Oncorhynchus',
-                duration = c(24, 120))
-\end{lstlisting}
 
+\input{article/code/standartox_example.tex}
+\pagebreak
+
+\begin{figure}
+    \includestandalone[scale=0.4]{article/tikz/standartox_organigram}
+    \caption{Organigram of Standartox.}
+    \label{fig:stx-organigram}
+\end{figure}
 \pagebreak
@@
 \input{article/tables/app-parameters.tex}
+\pagebreak
 
+\input{article/tables/api-endpoints.tex}
 \pagebreak
@@
 Currently, CRA often relies on test results of a few, well tested standard organisms, such as \textit{Daphnia magna}, \textit{Pimephales promelas}, \textit{Selenastrum capricornutum} to name a few, although researchers have conducted experiments on a much greater variety of organisms, observable in the 12,906 taxa from 952,634 test results collected in the EPA ECOTOX data base.
-In order to locate ecotoxicological test data, researchers can rely on individual test results reported in publications, on compiled and published data sets \citep{malaj_organic_2014} or on data bases such as the Pesticide Property Data Base (PPDB) \citep{lewis_international_2016}.
+In order to locate ecotoxicological test data, researchers can rely on individual test results reported in publications, on compiled and published data sets \citep{malaj_organic_2014, morrissey_neonicotinoid_2015} or on data bases such as the Pesticide Property Data Base (PPDB) \citep{lewis_international_2016}.
 Each approach has its limitations though.
@@
 Just as well, these resources run at risk of providing out of date data sooner or later or rarely provide sophisticated means to directly process the data.
-Standartox overcomes these paucities by using the EPA ECOTOX data base, the largest publicly available collection of ecotoxicological test information and builds Standartox and its aggregation methods on top of it to provide adequate and readily available estimates of the toxicity of chemicals.
-Besides that, Standartox constitutes not only a one-time compiled data set but rather a scalable method that can steadily incorporate new ecotoxicological test data in an automated manner, allowing for constant improvement of toxicity estimations over time.
-Standartox allows users to access older versions of the data base as well and is created as open source software and therefore a fully reproducible method.
+Standartox overcomes these paucities by making use of the quarterly updated EPA ECOTOX data base, the largest publicly available collection of ecotoxicological test information.
+In doing so, Standartox allows for a quick access to ecotoxicological test data, not being confined to specific chemical classes.
+Furthermore, Standartox constitutes not only a one-time compiled data set but rather a scalable method that can steadily incorporate new ecotoxicological test data in an automated manner, allowing for constant improvement of toxicity estimations over time.
+Data retrieval is streamlined and can be easily done via the APP or the R package.
+On top of filtering the data, Standartox provides means to aggregate results from multiple ecotoxicological tests to retrieve single exposure endpoints for specific chemical, taxon and test parameter combinations.
+Test results for each chemical in the filtered data set are thereby aggregated by calculating the minimum, the geometric mean or the maximum.
+Standartox returns then a filtered and an aggregated data set.
+Besides, Standartox allows users to access older versions of the data base as well and is created as open source software and therefore a fully reproducible method.
 
@@
 The EnvirTox data base also allows for an aggregation of test results to derive single toxicity values for individual taxa whereas Standartox performs this aggregation for individual chemicals.
-Comptox, is a web tool published by the EPA which, similar to Standartox allows for filtering test results, the retrieval of additional chemical information and predicting toxicity properties, such as 48h \textit{Daphnia magna} LC\textsubscript{50} values.
+Comptox, is a web tool published by the EPA which, similar to Standartox allows for filtering test results, the retrieval of additional chemical information and predicting toxicity properties, such as 48 h \textit{Daphnia magna} LC\textsubscript{50} values.
 However, toxicity predictions are limited to some standard organisms (e.g.
@@
 Since the used data in Standartox merely represents a sample of the actual toxicity of chemicals on organisms, it has to be stated that the returned Standartox values are subject to change.
-New studies will be published and constantly incorporated into Standartox, improving the toxicity estimates.
+New studies will be published and constantly incorporated into EPA ECOTOX data base and therefore into Standartox, improving the toxicity estimates.
 However, we expect values of well tested chemical-organism combinations to be relatively stable.
@@
 2) Compare Standartox with predictions here: https://comptox.epa.gov/dashboard/predictions/index
+- \citep{gadaleta_new_2018} what does this software exactly do an how to access it?
 \fi

### 2019-11-28 article/parts/03_methods.tex

@@
 
-Standartox consists of three parts of software explained in detail below.
-Firstly a data acquisition and preparation processing pipeline which downloads quarterly released new versions of the EPA ECOTOX data base, upon which Standartox is built and performs several processing steps on the data.
+Standartox consists of three parts of software that are combined to ....
+Firstly a data acquisition and preparation part that downloads the quarterly released new versions of the EPA ECOTOX data base and automatically builds Standartox based on several processing steps on the data.
 Secondly, the final Standartox table is accessible via a web application (APP) and an application programming interface (API).
@@
 Subsequently SQL functions for further processing are added and known errors are corrected (and also reported).
-Lookup tables (Supplement \ref{sup:conv-concentration}, \ref{sup:conv-duration}) for filtering and converting the data together with a meta table, containing inter alia the release version of the EPA ECOTOX data base are created as well.
-In a next step Chemical Abstracts Service (CAS) identifiers are used to query identifier data bases, including the Chemical Identifier Resolver (CIR) service \citep{nationalinstitutesofhealthnih_chemical_2019} and the Pubchem data base \citep{kim_pubchem_2016}.
-Chemical identifiers (CAS, InchI) and taxonomic names are then used to query chemical and taxonomic specific data bases (Table \ref{tab:data-base-additional}) to add information, such as chemical grouping or organism habitat preferences and global occurrence patterns to the local build.
+In addition lookup tables (Supplement \ref{sup:conv-concentration}, \ref{sup:conv-duration}) are created that support filtering and converting the data as well as a meta table providing information such as the release version of the EPA ECOTOX data base.
+In the next step, Chemical Abstracts Service (CAS) identifiers are used to query identifier data bases, including the Chemical Identifier Resolver (CIR) service \citep{nationalinstitutesofhealthnih_chemical_2019} and the Pubchem data base \citep{kim_pubchem_2016}.
+Chemical identifiers (CAS, InchI) and taxonomic names are then used to query additional information such as chemical grouping or organism habitat preferences and global occurrence patterns from data bases (Table \ref{tab:data-base-additional}) and include them in the Standardtox database.
 
@@
 
-From here on the final Standartox data set is compiled, whereupon concentration and duration units are converted in the process of harmonizing the data.
-Out of the unique 1229 concentration units in the EPA ECOTOX data base, Standartox retains only those (n = X) which are convertible to one of the following units: ug/l, g/m2, ppb, mg/kg, \% and flags the remaining as \textit{other}, resulting in 873473 out of 952625 test results.
-Likewise, out of the 125 test duration units in the EPA ECOTOX, Standartox retains only those (n = X) which are accurately convertible to hours, leading to 905110 out of 952625 test results).
+Finally, the Standartox data set is compiled, which includes the harmonisation of data, e.g.
+through conversion of units related to the concentrations and test duration.
+Out of the unique 1229 concentration units in the EPA ECOTOX data base, Standartox retains only those (n = X) that are unambiguously convertible to one of the following units: ug/l, g/m2, ppb, mg/kg, \% and flags, the remaining as \textit{other}, resulting in 873473 out of 952625 test results.
+Furthermore, the units are cleaned, for example through removing additional information in the field such as \textit{food}, \textit{soil}, \textit{ai} that are also coded in other variables and hinder the processing of units.
+Concentrations that are given as rates such as per day (mg/kg/day) are generally excluded.
+Likewise, out of the 125 test duration units in the EPA ECOTOX, Standartox retains only those (n = X) that can be converted to hours, keeping 905110 out of 952625 test results).
 Thereby concentration unit additions such as \textit{food}, \textit{soil}, \textit{ai} etc.
@@
 Units such as per day (mg/kg/day) are generally excluded.
-Endpoints were refined to three endpoint groups NOEX, LOEX and XX50.
-The former two summarise no-observed-adverse-effect level and lowest-observed-adverse-effect levels, respectively.
-The latter includes various half maximal effective concentration (e.g.
-EC50, LC50, LD50 etc.) commonly used in ecotoxicology.
-Other endpoints are not included in Standartox.
+Test endpoints are refined to three endpoint groups NOEX, LOEX and XX50.
+The former two represent no-observed-adverse-effect levels and lowest-observed-adverse-effect levels, respectively.
+The latter includes various versions of halft of the maximal effective concentration (e.g.
+EC50, LC50, LD50 etc.) commonly used in ecotoxicology, e.g.
+the concentration where half of the test individuals show an effect.
+Other endpoints are removed.
 Beyond that, filters for the CAS number, the concentration type (e.g.
-active ingredient, formulation etc.), the chemical class (e.g.
-fungicides, metals etc.), taxa, organism habitat and region, test duration as well as effect (e.g.
-Mortality, Growth etc.) groups (See Table \ref{tab:app-parameters}) are created.
+active ingredient, formulation), the chemical class (e.g.
+fungicides, metals), taxon, organism habitat and region, test duration as well as effect type (e.g.
+mortality, growth) (See Table \ref{tab:app-parameters}) are created.
 Along with the compiled Standartox data set, a catalog, listing all distinct entries and value ranges, for categorical and continuous variables, respectively is created.
-As the end, to guarantee the accuracy of the data, check scripts are run.
-Single processing scripts are listed in figure \ref{fig:pipeline-tree}.
-The final Standartox table together with the catalog is exported.
-Both the APP and the API access it.
+Finally, we run quality control scripts that check the accuracy of the data.
+Single processing scripts are listed in Figure \ref{fig:pipeline-tree}.
+The final Standartox table together with the catalog is exported and accessible via web application and the API.
 
@@
 The web application (APP) and the Application Programming Interface (API) load the compressed serialized Standartox data into memory and allow a client to interact with it.
-The client can then call the functions stx\_filter() and stx\_aggregate() which filter and aggregate the data according to specific parameters (Table \ref{tab:app-parameters}).
+The client can then call the functions stx\_filter() and stx\_aggregate() that filter and aggregate the data according to specific parameters (Table \ref{tab:app-parameters}).
 
 \subsubsection*{Web application (APP)}
-The interactive APP was built in R, using the shiny web application framework which runs with the help of a shiny server \citep{chang_shiny_2018} on \url{www.standartox.uni-landau.de}.
-It is set of R functions that allow a user to filter and aggregate the data.
-The data can then be downloaded.
+The interactive APP was built in R using the shiny web application framework, which runs with the help of a shiny server \citep{chang_shiny_2018} on \url{www.standartox.uni-landau.de}.
+It is a set of R functions that allow a user to filter, aggregate and download the data.
 
 \subsubsection*{Application programming interface (API)}
-The API is built by using the R package plumber \citep{trestletechnologyllc_plumber_2018} which allows for the creation of RESTful APIs from R.
+The API is built by using the R package plumber \citep{trestletechnologyllc_plumber_2018}, which allows for the creation of RESTful APIs from R.
 The API is reachable via the URL \url{http://139.14.20.252} and port 8000.
@@
 The \textit{/filter} returns the filtered Standartox table as a compressed serialized binary file created by the R fst package \citep{klik_fst_2019}, to reduce size and allow for fast user queries.
-The \textit{/aggregate} API-endpoint returns a R function to the client allowing the client the aggregation of the filtered data.
+The \textit{/aggregate} API-endpoint returns an R function allowing the client to aggregate the filtered data.
 Lastly, the \textit{/meta} API-endpoint returns a JSON file with meta information, such as the timestamp of the request and the used Standartox version.
@@
 \subsection*{Technical Validity}
-In order to guarantee appropriate conversion and harmonisation of units we take a random sample of the 50 most occurring concentration and six of the most occurring duration units and compare units created by Standartox to manually calculated unit conversions.
+To guarantee appropriate conversion and harmonisation of units we take a random sample of the 50 most occurring concentration and six of the most occurring duration units and compare units created by Standartox to manually calculated unit conversions.
 This is done at the end of the processing pipeline and stops the it if not all units match.

### 2019-11-28 article/tables/data_base_tables.tex

@@
 \begin{table}[ht]
-  \caption{Table of queried data bases}
-  \label{tab:data-base}
+  \caption{Table of additionally queried publicly available data bases and their URLs.}
+  \label{tab:data-base-additional}
   \centering
-\begin{tabular}{ll}
+\begin{tabular}{cc}
   \hline
@@
   \hline
-  Chemical Entities of Biological Interest (ChEBI) & https://www.ebi.ac.uk/chebi \\ 
-  Chemical Identifier Resolver & https://cactus.nci.nih.gov/chemical/structure \\
-  ChemSpider & http://www.chemspider.com \\
-  Eurostat & https://ec.europa.eu/eurostat/home \\
-  PHYSPROP Database & https://www.srcinc.com/what-we-do/environmental/scientific-databases.html \\
-  PubChem & https://pubchem.ncbi.nlm.nih.gov \\ 
-  Wikidata & https://www.wikidata.org/wiki/Wikidata:Main_Page \\
-  Global Biodiversity Information Facility (GBIF) & https://www.gbif.org \\
-  World Register of Marine Species (WoRMS) & http://marinespecies.org \\
+  \makecell{Chemical Entities of \\ Biological Interest (ChEBI)} & https://www.ebi.ac.uk/chebi  \\
+  Chemical Identifier Resolver & https://cactus.nci.nih.gov/chemical/structure  \\[0.5cm]
+  ChemSpider & http://www.chemspider.com    \\[0.5cm]
+  Eurostat & https://ec.europa.eu/eurostat/home \\[0.5cm]
+  PHYSPROP & \makecell{https://www.srcinc.com/what-we-do/environmental/\\scientific-databases.html}  \\[0.5cm]
+  PubChem & https://pubchem.ncbi.nlm.nih.gov \\[0.5cm]
+  Wikidata & https://www.wikidata.org/wiki/Wikidata:Main_Page \\[0.5cm]
+  \makecell{Global Biodiversity \\ Information Facility (GBIF)} & https://www.gbif.org \\[0.5cm]
+  \makecell{World Register of \\ Marine Species (WoRMS)} & http://marinespecies.org \\
   \hline

### 2019-11-28 article/tikz/standartox_organigram.tex

\documentclass[border = 1.5cm]{standalone}

%%%% packages
\usepackage{tikz}
\usepackage{tikzpeople}
\usetikzlibrary{trees,
                arrows,
                shapes.geometric,
                positioning,
                fit}
%% document
\begin{document}
%% styles
\tikzset{
    font = {\fontsize{13pt}{12}\sffamily},
	element/.style = {rectangle, rounded corners, minimum width = 4cm, minimum height = 3cm, text centered, draw = black},
    filter/.style = {rectangle, rounded corners, align = center, anchor = west, yshift = -1cm},
    frame/.style = {draw, rectangle, rounded corners, dashed, blue, line width = 1pt, inner sep = 4mm}, %, dash pattern=on 1pt off 4pt on 6pt off 4pt},
    arrow1/.style = {->, thick},
    arrow2/.style = {->, line width = 3pt}
}
%% layers
\pgfdeclarelayer{bg1}
\pgfsetlayers{bg1, main}
%% tikzpicture
\begin{tikzpicture}[
	grow via three points = {one child at (0.5,-1) and
		                     two children at (0.5,-1) and
		                     (0.5,-1.7)},
	edge from parent path = {(\tikzparentnode.south) |- (\tikzchildnode.west)},
	inner sep = 0cm, outer sep = 0cm
	]
%% person
\node (person) [bob, minimum size = 2cm] at (0,0) {};
%% app
\node (stx-app-screenshot) [right of = person, xshift = 5cm, yshift = 3cm] {\includegraphics[width=.35\textwidth]{article/figures/screenshot_standartox_app.png}};
\node (stx-rpackage-screenshot) [right of = person, xshift = 5cm, yshift = -3cm] {\includegraphics[width=.35\textwidth]{article/figures/screenshot_standartox_rpackage.png}};
%% elements
\node (epa-ecotox) [element, right of = person, xshift = 13cm, yshift = 5.5cm, fill = gray] {EPA ECOTOX};
\node (filtered) [element, below of = epa-ecotox, yshift = -3cm] {Filtered Data}
    child {node [filter] {CAS}}
    child {node [filter] {Concentration type}}
    child {node [filter] {Chemical class}}
    child {node [filter] {Taxa}}
    child {node [filter] {Habitat}}
    child {node [filter] {Region}}
    child {node [filter] {Duration}}
    child {node [filter] {Effect}}
    child {node [filter] {Endpoint}}
    child {node (filter-last) [filter] {Version}}
    ;
\node (aggregated) [element, right of = filtered, xshift = 7cm] {Aggregated Data};
\node (meta) [element, below of = aggregated, yshift = -5cm] {Meta Information};
%% frames
\node (frame-stx) [frame, fit = (epa-ecotox) (filter-last) (meta) ] {};
\node (frame-stx-text) [above, blue, inner sep = 2mm] at (frame-stx.north) {Standartox data set};
\node (frame-app) [frame, fit = (stx-app-screenshot)] {};
\node (frame-app-text) [above, blue, inner sep = 2mm] at (frame-app.north) {APP};
\node (frame-api) [frame, fit = (stx-rpackage-screenshot)] {};
\node (frame-api-text) [above, blue, inner sep = 2mm] at (frame-api.north) {API};
\node (frame-rpackage) [frame, fit = (frame-api) (frame-api-text)] {};
\node (frame-rpackage-text) [above, blue, inner sep = 2mm] at (frame-rpackage.north) {R package};
%% arrows
\draw [arrow1] (filtered.east) -- (aggregated.west);
\draw [arrow1] (epa-ecotox.south) -- (filtered.north);
\begin{pgfonlayer}{bg1}
    \draw [arrow2] (person.north east) to[bend left = 10] (frame-stx.150);
    \draw [arrow2] (person.south east) to[bend right = 10] (frame-stx.210);
\end{pgfonlayer}
\end{tikzpicture}
\end{document}

### 2019-12-10 article/parts/01_introduction.tex
Updates from Overleaf

@@
 For example, \citet{morrissey_neonicotinoid_2015} used ecotoxicological test results from 49 insect and crustacean species to evaluate the effect of neonicotinoids, an insecticide class on the aquatic ecosystem.
-Also, \citet{malaj_organic_2014} compiled intera alia experimental toxicity test results for 223 various chemical compounds to assess the health of freshwater ecosystems in Europe.
+Also, \citet{malaj_organic_2014} compiled inter alia experimental toxicity test results for 223 various chemical compounds to assess the health of freshwater ecosystems in Europe.
 Similarly, permissible environmental concentrations are often derived from these test data, typically by a combination with safety factors to account for uncertainties.
@@
 
+Standartox makes use of the publicly available and quarterly released EPA ECOTOX database which comprises almost 1,000,000 ecotoxicological test results from almost 12,000 chemicals and 13,000 taxa \citep{usepa_ecotox_2019}.
+
 We developed Standartox, a tool and database that aims to overcome the limitations of other databases by continuously incorporating the ever-growing number of test results in an automated process workflow that ultimately leads to single aggregated data point for a specific chemical-organism test combination, representing the toxicity of a chemical.
@@
 Standartox comes with two front-ends,  a web application (APP) and a R \citep{rcoreteam_language_2017} package, which accesses a application programming interface (API) providing convenience structures and thereby largely reducing processing time for users.
+
+Standartox is needed due to variability of test results:
+
+125 distinct duration units
 

### 2019-12-10 article/setac2020/abstract_extended.tex
Updates from Overleaf

\documentclass[10pt, a4paper, english]{article}

%% packages
\usepackage[T1]{fontenc}
\usepackage{pgfplotstable} % include graphics
\usepackage{tikz}
\usetikzlibrary{trees, arrows, shapes.geometric, positioning}
\usepackage{tikzpeople}
\usepackage{forest}
\usepackage{standalone}
\usepackage{natbib}
\usepackage{geometry}
\usepackage{helvet} % Helvetica aka Arial
\usepackage[normalem]{ulem} % underline
\usepackage[affil-it]{authblk} 
\usepackage{mwe}

%% set geometry
\geometry{
 a4paper,
% total={170mm,257mm},
 top = 20mm,
 left = 20mm,
 right = 20mm,
 bottom = 20mm,
 bindingoffset = 0cm % gutter
}

%% renew commands
\renewcommand{\familydefault}{\sfdefault}
\renewcommand\Authfont{\fontsize{12}{14.4}\selectfont}
\renewcommand\Affilfont{\fontsize{10}{10.8}\selectfont}

%\renewcommand\email{\fontsize{10}{10}\selectfont}

\begin{document}

\title{\textbf{Standartox – a tool for assessing the risk of chemicals}}

\author{\uline{Scharmüller, Andreas}\textsuperscript{1}}
\author{Schreiner, Verena C.\textsuperscript{1}}
\author{Schäfer, Ralf B. \textsuperscript{1}}
\date{\vspace{-5ex}}

\affil{Institute for Environmental Sciences, University Koblenz-Landau, 76829 Landau in der Pfalz, Germany}

%\href{mailto:scharmueller@uni-landau.de}{scharmueller@uni-landau.de}
\maketitle

\section{Introduction}
An increasing number of chemicals such as pharmaceuticals, pesticides and synthetic hormones are in daily use all over the world. In Europe alone, some 100,000 chemicals are estimated to be in current use, whereof 30,000 are produced in quantities larger than one ton per year \citep{breithaupt_costs_2006}. Except for pesticides that are released into the environment deliberately, most chemicals enter the environment as a result of their use through different paths (e.g. atmospheric emission and deposition or discharge through wastewater) \citep{schwarzenbach_challenge_2006}. In the environment, chemicals can adversely affect populations and communities and in turn related ecosystem functions. Pollution with man-made chemicals has been identified as one of three major environmental problems for which research gaps hamper the derivation of planetary boundaries, i.e. thresholds beyond which irreversible state shifts may occur \citep{steffen_planetary_2015}. Even highly regulated chemical compounds such as pesticides have been shown to cause strong adverse effects on organisms, such as birds \citep{hallmann_declines_2014} or fish \citep{yamamuro_neonicotinoids_2019}, questioning the current regulation efforts \citep{schafer_future_2019}. To evaluate the risks from chemicals for ecosystems data on their toxicity, which is typically produced in standardised ecotoxicological laboratory tests is required \citep{malaj_organic_2014, morrissey_neonicotinoid_2015}. Similarly, permissible environmental concentrations are often derived from these test data, typically by a combination with safety factors to account for uncertainties. Given that data sets with large numbers of analysed chemicals are becoming increasingly available, ecotoxicity data are pivotal for regulatory risk assessment and science. To date, only few initiatives exist that aim to create a public resource of harmonized and quality controlled ecotoxicological data, such as the United States Environmental Protection Agency ECOTOXicology knowledgebase (EPA ECOTOX) \citep{elonen_ecotoxicology_2018}, the German Environmental Agency's Information System Ecotoxicology and Environmental Quality Targets (ETOX) \citep{umweltbundesamt_etox_2019} or the Pesticide Property Data Base (PPDB) \citep{lewis_international_2016}. The former two compile all available test results into a database. However, for many chemicals, multiple ecotoxicity values are available for the same test organism, which can vary strongly. Hence, these databases can contain multiple entries for the same test without quality information and with different units, hampering reproducible science if different values are selected for studies. The PPDB database is confined to pesticides and provides harmonised and quality controlled data for a few selected test organisms, thereby ignoring vast amounts of data for other species and chemicals. Moreover, data analyses often require links to additional data resources, for example to append additional chemical and species information (e.g. chemical properties, habitat of species), which calls for more automated procedures. We developed Standartox, a tool and database that aims to overcome the limitations of other databases by continuously incorporating the ever-growing number of test results in an automated process workflow that ultimately leads to single aggregated data point for a specific chemical-organism test combination, representing the toxicity of a chemical. Hence, Standartox provides the basis for reproducible science and combines information from different sources to simplify the derivation of more advanced risk indicators such as Species Sensitivity Distributions (SSD) and Toxic Units (TU), which represent two prominent concepts to assess effects on organisms in ecotoxicology \citep{posthuma_species_2002, kefford_definition_2011}. Besides aggregating ecotoxicological test results, Standartox provides a concise overview of the status of tested chemicals, allowing the identification of potential research gaps. Standartox comes with two front-ends, a web application (APP) (www.standartox.uni-landau.de) and a R package, which accesses a application programming interface (API) providing convenience structures and thereby largely reducing processing time for users.

\section{Methods}
Standartox downloads the EPA ECOTOX data base and builds it locally. It then extracts CAS numbers and taxonomic nomenclatures to query additional information from chemical and taxonomic specific data bases, respectively. Subsequently, measurement units are harmonized, filter parameters are created and a final data set is created and exported.

\section{Results and Discussion}
The exported object can then be accessed by a web application or an R-package to retrieve filtered and aggregated test results, as a comma-separated values (csv) file or directly within R. Users can filter the test data by CAS numbers, chemical classes (e.g. herbicides, metals), taxa, habitat-specific parameters (e.g. freshwater), test duration as well as effect (e.g. mortality, growth) and endpoint (e.g. EC\textsubscript{50}, NOEC, LOEC) groups. Furthermore test results can be aggregated by the geometric mean. On top of filtering the data, Standartox provides means to aggregate results from multiple ecotoxicological tests to retrieve single exposure endpoints for specific chemical, taxon and test parameter combinations. Test results for each chemical in the filtered data set are thereby aggregated by calculating the minimum, the geometric mean or the maximum. Standartox returns then a filtered and an aggregated data set (Figure \ref{fig:stx-organigram}). Besides, Standartox allows users to access older versions of the data base as well and is created as open source software and therefore a fully reproducible method. Standartox constitutes not only a one-time compiled data set but rather a scalable method that can steadily incorporate new ecotoxicological test data in an automated manner, allowing for constant improvement of toxicity estimations over time. To, date Standartox provides access to more than 605,000 test results and its aggregates. In order to asses the validity of the Standartox aggregation approach, we compared Standartox results with effect values from the PPDB (n=3601) for the same chemicals (Figure \ref{fig:standartox_ppdb_diff}). This showed that 91.3\% of the Standartox values diverge by less than a factor of ten from the PPDB values. Considering only Standartox values for which at least five or ten tests were aggregated the value increases to 91.7\% and 93.4\%, respectively.

\begin{figure}
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{article/figures/gg_ppdb_stan_compare_continous.png}
        \caption{Comparison between Standartox and PPDB values. The black line marks exact coherence and red lines mark a divergence of a factor of 10. Compared organism groups are color coded.}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includestandalone[width=0.9\textwidth]{article/tikz/standartox_organigram} % second figure itself
        \caption{Organigram of Standartox. Users can access Standartox via a web application or an R-package.}
    \end{minipage}
\end{figure}

\bibliographystyle{apalike}
\bibliography{refs/references-standartox.bib}

\end{document}

### 2019-12-10 article/tikz/standartox_organigram.tex

@@
 \node (frame-stx) [frame, fit = (epa-ecotox) (filter-last) (meta) ] {};
-\node (frame-stx-text) [above, blue, inner sep = 2mm] at (frame-stx.north) {Standartox data set};
+\node (frame-stx-text) [above, blue, inner sep = 2mm] at (frame-stx.north) {Standartox database};
 \node (frame-app) [frame, fit = (stx-app-screenshot)] {};
-\node (frame-app-text) [above, blue, inner sep = 2mm] at (frame-app.north) {APP};
+\node (frame-app-text) [above, blue, inner sep = 2mm, align=center] at (frame-app.north) {web application:\\ standartox.uni-landau.de};
 \node (frame-api) [frame, fit = (stx-rpackage-screenshot)] {};
-\node (frame-api-text) [above, blue, inner sep = 2mm] at (frame-api.north) {API};
+\node (frame-api-text) [above, blue, inner sep = 2mm] at (frame-api.north) {Application Programming Interface};
 \node (frame-rpackage) [frame, fit = (frame-api) (frame-api-text)] {};
-\node (frame-rpackage-text) [above, blue, inner sep = 2mm] at (frame-rpackage.north) {R package};
+\node (frame-rpackage-text) [above, blue, inner sep = 2mm] at (frame-rpackage.north) {R-package: standartox};
 %% arrows

### 2019-12-16 run_build.R
changes

@@ -6,8 +6,8 @@
 # 00 22 * * * /home/scharmueller/Projects/run_build.sh
 
 # setup -------------------------------------------------------------------
-saveRDS(Sys.time(), 'cache/start_time')
 source('~/Projects/standartox-build/R/gn_setup.R', max.deparse.length = mdl)
+saveRDS(Sys.time(), file.path(cachedir, 'start_time.rds'))
 
 # console log -------------------------------------------------------------
 if (sink_console) {
@@ -120,6 +120,7 @@ if (general) {
 
 # end ---------------------------------------------------------------------
 source(file.path(src, 'gn_end.R'), max.deparse.length = mdl)
+readRDS(file.path(cachedir, 'start_time.rds'))
 
 # console log 2 -----------------------------------------------------------
 if (sink_console) {

### 2019-12-16 article/parts/03_methods.tex

@@
 \section*{Methods}
-
-Standartox consists of three parts of software that are combined to ....
-Firstly a data acquisition and preparation part that downloads the quarterly released new versions of the EPA ECOTOX data base and automatically builds Standartox based on several processing steps on the data.
-Secondly, the final Standartox table is accessible via a web application (APP) and an application programming interface (API).
-Lastly, the R \citep{rcoreteam_language_2017} package standartox constitutes an easily usable way of accessing the tool via the API.
+An automated processing pipeline downloads the quarterly released EPA ECOTOX database, performs several preparation steps on it and exports a final Standartox data set.
+This data set is accessible via a web application (APP) and an application programming interface (API).
+To facilitate the API access, the R \citep{rcoreteam_language_2017} package \textit{standartox} is built.
 
 \subsection*{Processing pipe line}
-The US Environmental Protection Agency (EPA) releases the EPA ECOTOX data base on a quarterly basis.
-The latest version, released on 12.09.2019 contains data on 11,756 chemicals, 12,906 taxa, 49,153 references and 952,634 results \citep{usepa_ecotox_2019} constituting a vast collection of ecotoxicological tests, which can be downloaded as pipe delimited ASCII files.
-Standartox downloads the EPA ECOTOX data base whenever a new version is published and builds into a local PostgreSQL data base \citep{szocs_build_2019}.
-Subsequently SQL functions for further processing are added and known errors are corrected (and also reported).
-In addition lookup tables (Supplement \ref{sup:conv-concentration}, \ref{sup:conv-duration}) are created that support filtering and converting the data as well as a meta table providing information such as the release version of the EPA ECOTOX data base.
-In the next step, Chemical Abstracts Service (CAS) identifiers are used to query identifier data bases, including the Chemical Identifier Resolver (CIR) service \citep{nationalinstitutesofhealthnih_chemical_2019} and the Pubchem data base \citep{kim_pubchem_2016}.
-Chemical identifiers (CAS, InchI) and taxonomic names are then used to query additional information such as chemical grouping or organism habitat preferences and global occurrence patterns from data bases (Table \ref{tab:data-base-additional}) and include them in the Standardtox database.
+Standartox downloads the quarterly released EPA ECOTOX data base whenever a new version is published and builds it into a local PostgreSQL data base \citep{szocs_build_2019}.
+Subsequently Structured Query Language (SQL) functions for further processing the data are developed.
+In addition lookup tables (Supplement \ref{sup:conv-concentration}, \ref{sup:conv-duration}) that enable duration and concentration unit conversions of the data are created.
+A meta table providing information such as the release version of the EPA ECOTOX data base is then added.
+In the next step, Chemical Abstracts Service Numbers (CAS) and International Chemical Identifier (InChI) as well as taxonomic names are used to query additional information from publicly available databases on chemicals and organisms, respectively.
+This includes the Compendium of Pesticide Common Names \citep{wood_compendium_2019}, the Chemical Entities of Biological Interest (ChEBI) database \citep{hastings_chebi_2016}, the Chemical Identifier Resolver (CIR) service \citep{nationalinstitutesofhealthnih_chemical_2019}, the PHYSPROP database, the Pubchem data base \citep{kim_pubchem_2016}, Eurostat \citep{europeancommission_eurostat_2019} and Wikidata \citep{vrandecic_wikidata_2014} for chemicals and the World Register of Marine Species (WoRMS) \citep{wormseditorialboard_world_2018} and the Global Biodiversity Information Facility (GBIF) \citep{_gbif_2019} (Table \ref{tab:data-base-additional}) for habitat and spatial distribution of organisms.
+This information is added to Standartox to allow filtering for specific classes of chemicals as well as spatial distribution (i.e.
+continents) and habitat preferences (e.g.
+freshwater) of individual taxa.
 
@@
 through conversion of units related to the concentrations and test duration.
-Out of the unique 1229 concentration units in the EPA ECOTOX data base, Standartox retains only those (n = X) that are unambiguously convertible to one of the following units: ug/l, g/m2, ppb, mg/kg, \% and flags, the remaining as \textit{other}, resulting in 873473 out of 952625 test results.
+Out of the unique 1229 concentration units in the EPA ECOTOX data base, Standartox retains only those (n = X) that are unambiguously convertible to one of the following units: $\mu$g/l, g/m2, ppb, mg/kg, \% and flags, the remaining as \textit{other}, resulting in 873473 out of 952625 test results.
 Furthermore, the units are cleaned, for example through removing additional information in the field such as \textit{food}, \textit{soil}, \textit{ai} that are also coded in other variables and hinder the processing of units.
 Concentrations that are given as rates such as per day (mg/kg/day) are generally excluded.
-Likewise, out of the 125 test duration units in the EPA ECOTOX, Standartox retains only those (n = X) that can be converted to hours, keeping 905110 out of 952625 test results).
-Thereby concentration unit additions such as \textit{food}, \textit{soil}, \textit{ai} etc.
+Likewise, out of the 125 test duration units in the EPA ECOTOX, Standartox retains only those (n = X) that can be converted to hours, excluding ambiguous duration units such as harvest or lifetime and keeping 905110 out of 952625 test results.
+Thereby concentration unit additions such as \textit{food}, \textit{soil}, \textit{ai (active ingredient)} etc.
 are neglected since they are also coded in other variables and only hinder the processing of units.
 Units such as per day (mg/kg/day) are generally excluded.
-Test endpoints are refined to three endpoint groups NOEX, LOEX and XX50.
-The former two represent no-observed-adverse-effect levels and lowest-observed-adverse-effect levels, respectively.
-The latter includes various versions of halft of the maximal effective concentration (e.g.
-EC50, LC50, LD50 etc.) commonly used in ecotoxicology, e.g.
-the concentration where half of the test individuals show an effect.
-Other endpoints are removed.
+Test endpoints are restricted to three groups, named NOEX, LOEX and XX50.
+The former two represent no-observed-adverse-effect and lowest-observed-adverse-effect concentrations or levels, respectively.
+The latter includes various sub-groups of the half maximal effective concentration, where half of the tested individuals show an effect (e.g.
+EC50, LC50, LD50), which is a common measure in ecotoxicology \citep{malaj_organic_2014}.
+Other endpoints, such as Bioconcentration factors (BCF), non-half maximal effective concentrations (e.g.
+IC10, EC25, LD99) or maximum acceptable toxicant (MATC) concentrations are removed.
 Beyond that, filters for the CAS number, the concentration type (e.g.
 active ingredient, formulation), the chemical class (e.g.
-fungicides, metals), taxon, organism habitat and region, test duration as well as effect type (e.g.
-mortality, growth) (See Table \ref{tab:app-parameters}) are created.
+fungicides, metals), taxon, organism habitat (e.g freshwater, terrestrial) and region (e.g.
+Europe, Asia), test duration as well as effect type (e.g.
+mortality, growth) are created.
 Along with the compiled Standartox data set, a catalog, listing all distinct entries and value ranges, for categorical and continuous variables, respectively is created.
@@
 \subsubsection*{Web application (APP)}
-The interactive APP was built in R using the shiny web application framework, which runs with the help of a shiny server \citep{chang_shiny_2018} on \url{www.standartox.uni-landau.de}.
+The interactive APP was built in R using the shiny web application framework, which runs with the help of a shiny server \citep{chang_shiny_2018} on \url{standartox.uni-landau.de}.
 It is a set of R functions that allow a user to filter, aggregate and download the data.
@@
 \subsection*{Technical Validity}
-To guarantee appropriate conversion and harmonisation of units we take a random sample of the 50 most occurring concentration and six of the most occurring duration units and compare units created by Standartox to manually calculated unit conversions.
-This is done at the end of the processing pipeline and stops the it if not all units match.
+To guarantee appropriate unit conversion and harmonisation, we compare one automatically converted entry to a manual conversion for every distinct concentration (1229) and duration (129) unit.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TODO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%% Where to put?
+(See Table \ref{tab:app-parameters})
+
 \iffalse

### 2019-12-16 article/parts/04_user_notes.tex
Updates from Overleaf

--- 2019-11-28 article/parts/02_results_discussion.tex
@@
-\section*{Results and Discussion}
-Standartox is a tool that allows to filter and aggregate ecotoxicological test results to derive single exposure endpoints for individual chemicals.
-It can either be accessed through the APP (Figure \ref{fig:app}) or the API \ref{tab:api-endpoints} together with the R-package standartox (Code example \ref{listing:standartox-example}).
-Both access pathways allow for the same filters and aggregation methods (Table \ref{tab:app-parameters}) to be applied and access the same data set.
+\section{User notes}
+
 Using the APP, the user can download the resulting data sets as a comma-separated values (csv) file whereas by using the API, the data can directly be loaded in R.
-Different parameters in order to filter the test data, such as compound-specific parameters, including the CAS number, the concentration type (e.g.
-Active Ingredient, Formulation) or the chemical class (e.g.
-insecticides, metals) as well as taxon-specific parameters, including common taxonomic groups (e.g.
-Daphniidae, algae), the habitat (e.g.
-freshwater) or the regional occurrence (e.g.
-Europe, Asia) of the organisms can be chosen.
-Furthermore, users can refine the results to specific test durations (in hours), effect groups (e.g.
-mortality, population, growth) and endpoints (EC\textsubscript{50}, LOEC or NOEC).
-As the method of aggregation users can choose the minimum, the maximum, or the geometric mean.
 To provide an easy overview of the data set, percentage values for each parameter entry are shown in the APP in the selection menu as well as by using the R package through querying the \textit{/catalog} API-endpoint.
 
-%% Add units part here
-
-\pagebreak
-\begin{figure}
-    \caption{Web application (APP) Standartox allowing to set filter parameters and aggregation methods.
-Data can be downloaded as comma separated value (csv) files.
-An interactive overview plot and table allow for browsing of the data.}
-    \label{fig:app}
-    \includegraphics[width=0.75\linewidth]{article/figures/screenshot_standartox_app.png}
-\end{figure}
-
-\pagebreak
-
-\input{article/code/standartox_example.tex}
 \pagebreak
@@
 \end{figure}
-\pagebreak
-
-\input{article/tables/app-parameters.tex}
-\pagebreak
-
-\input{article/tables/api-endpoints.tex}
-\pagebreak
-
-Standartox is designed to foster chemical risk assessment (CRA).
-Due to an increased amount of available ecotoxicological test data, it becomes fundamental to provide and distribute such information in adequate formats, both easily accessible for humans and easily processable for machines.
-Currently, CRA often relies on test results of a few, well tested standard organisms, such as \textit{Daphnia magna}, \textit{Pimephales promelas}, \textit{Selenastrum capricornutum} to name a few, although researchers have conducted experiments on a much greater variety of organisms, observable in the 12,906 taxa from 952,634 test results collected in the EPA ECOTOX data base.
-In order to locate ecotoxicological test data, researchers can rely on individual test results reported in publications, on compiled and published data sets \citep{malaj_organic_2014, morrissey_neonicotinoid_2015} or on data bases such as the Pesticide Property Data Base (PPDB) \citep{lewis_international_2016}.
-Each approach has its limitations though.
-Collecting toxicity information from individual publications is laborious and using already compiled data from publications or published data bases is often limited to specific groups of chemicals, such as pesticides, biozides etc.
-Just as well, these resources run at risk of providing out of date data sooner or later or rarely provide sophisticated means to directly process the data.
-Standartox overcomes these paucities by making use of the quarterly updated EPA ECOTOX data base, the largest publicly available collection of ecotoxicological test information.
-In doing so, Standartox allows for a quick access to ecotoxicological test data, not being confined to specific chemical classes.
-Furthermore, Standartox constitutes not only a one-time compiled data set but rather a scalable method that can steadily incorporate new ecotoxicological test data in an automated manner, allowing for constant improvement of toxicity estimations over time.
-Data retrieval is streamlined and can be easily done via the APP or the R package.
-On top of filtering the data, Standartox provides means to aggregate results from multiple ecotoxicological tests to retrieve single exposure endpoints for specific chemical, taxon and test parameter combinations.
-Test results for each chemical in the filtered data set are thereby aggregated by calculating the minimum, the geometric mean or the maximum.
-Standartox returns then a filtered and an aggregated data set.
-Besides, Standartox allows users to access older versions of the data base as well and is created as open source software and therefore a fully reproducible method.
-
-\subsection*{Other data bases}
-Other initiatives tidying the large number of ecotoxicological test results were also developed recently.
-They partly aim for overlapping goals, yet have limitations or objectives that differentiate them from Standartox.
-Since 2006, the PPDB provides validated ecotoxicological information on pesticides \citep{lewis_international_2016}.
-In order to asses the validity of the Standartox aggregation approach, we compared Standartox results with effect values from the PPDB (n=3601) for the same chemicals (Figure \ref{fig:standartox_ppdb_diff}).
-This showed that 91.3\% of the Standartox values diverge by less than a factor of ten from the PPDB values.
-Considering only Standartox values for which at least five or ten tests were aggregated the value increases to 91.7\% and 93.4\%, respectively.
-Admittedly a few Standartox values diverge by a large margin.
-
-\begin{figure}
-    \includegraphics[width=1\linewidth]{article/figures/gg_ppdb_stan_compare_continous.png}
-    \caption{Comparison between Standartox and PPDB values. The black line marks exact coherence and red lines mark a divergence of a factor of 10. Compared organism groups are color coded.}
-    \label{fig:standartox_ppdb_diff}
-\end{figure}
-
-The Network of reference laboratories, research centres and related organisations for monitoring of emerging environmental substances (NORMAN) focuses on assembling river basin specific pollutants \citep{von_der_ohe_new_2011}.
-The EnviroTox data base \href{https://envirotoxdatabase.org/} which also uses, amongst others the EPA ECOTOX data base as an input was recently published \citep{healthandenvironmentalsciencesinstitutehesi_envirotox_2019, connors_creation_2019}.
-In comparison to Standartox, EnviroTox focuses only on aquatic organisms and uses an qualitative algorithm to exclude unreliable test results, such as pre-filtering, for example the data to fish, amphibian invertebrate and algae taxa, to test durations of above 24 hours and aquatic taxa only.
-Besides, they add additional information on toxicity endpoints, such as acute or chronic classifications and mode of action assignments.
-In contrast, Standartox doesn't refine to specific test durations, but leaves it up to the user to decide on such parameters.
-The EnvirTox data base also allows for an aggregation of test results to derive single toxicity values for individual taxa whereas Standartox performs this aggregation for individual chemicals.
-Comptox, is a web tool published by the EPA which, similar to Standartox allows for filtering test results, the retrieval of additional chemical information and predicting toxicity properties, such as 48 h \textit{Daphnia magna} LC\textsubscript{50} values.
-However, toxicity predictions are limited to some standard organisms (e.g.
-\textit{Daphnia magna}), and the tool lacks the possibility to retrieve predicted values in an automated way \citep{williams_comptox_2017}.
-Petschick et al.
-(in submission) modeled risk threshold level equivalents for aquatic organisms by using ecotoxicological effect data from the EPA ECOTOX data base to support regulation.
-In summary, none of the above mentioned approaches aim for a holistic standardized aggregation method of exposure endpoints for individual chemicals and also lack the possibility for automated and scriptable user requests from common high level programming languages, such as R or Python.
-Along with newly created ecotoxicological data bases here, methods of how to efficiently store ecotoxicological data are also proposed as can be seen in the MAGIC Knowledge Base \citep{bub_graphing_2019}.
-This recent increase of efforts to compile ecotoxicological test data bases clearly emphasizes the need for holistic and automated analyses of large-scale ecotoxicological data.
-
-\subsection*{Limitations}
-Since the used data in Standartox merely represents a sample of the actual toxicity of chemicals on organisms, it has to be stated that the returned Standartox values are subject to change.
-New studies will be published and constantly incorporated into EPA ECOTOX data base and therefore into Standartox, improving the toxicity estimates.
-However, we expect values of well tested chemical-organism combinations to be relatively stable.
-A future analysis on the smallest number of tests needed to return trustworthy results, could shed more light on this aspect.
-Furthermore, toxicity test results are certainly influenced by individual tests parameters, such as pH, temperature or conductivity amongst others \citep{rosenkrantz_influence_2013, li_temperature_2011}.
-Such information is not included in the aggregation performed by Standartox, since the EPA ECOTOX data base only provides sparse records on those.
-Most frequently provided test parameters (and their percentage) are medium temperature (77\%), pH:56\%, hardness:27\%, dissolved oxygen:18\%, Alkalinity:15\% and salinity:9\%.
-Others are provided for less than 5\% of the tests.
-A text-mining approach, iterating through the individual publications could potentially increase this number.
-However, we argue that using the geometric mean as an aggregate allows for an adequate estimation of the toxicity of a chemical towards an organism group.
-We chose the geometric mean in comparison to the arithmetic mean, since it is considerably less influenced by outliers and is suitable for skewed data.
-Also, the geometric mean is preferable over the median, since the median completely ignores the influence of large or small values, making it unreliable for small data sets <CITATION GEOMETRIC MEAN>.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TODO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\iffalse
-- can also be incorporated in EPA Dashboard
-- Identify also underrepresented regions concerning organisms.
-1) \citep{staveley_challenge_2016} describe what is necessary for future ecotoxicology.
-2) Compare Standartox with predictions here: https://comptox.epa.gov/dashboard/predictions/index
-- \citep{gadaleta_new_2018} what does this software exactly do an how to access it?
-\fi
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% OLD %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\iffalse
-SSD try to extrapolate effects of chemicals towards organisms from species to a community level by fitting a statistical distribution to toxicity test results on organisms of different trophic levels in order to derive concentrations at which a specific fraction of organisms are effected (often at 5\% - HC\textsubscript{5} \citep{posthuma_species_2002}.
-Figure \ref{fig:ssd-isoproturon} shows an example of a SSD and the derived HC\textsubscript{5} value (red dassed line) created for the herbicide Isoproturon.
-TUs are another indicator of effects of chemicals on organisms.
-They are calculated by dividing a measured concentration (c) by an EC\textsubscript{50} value (cf.
-Equation \ref{eq:tu}).
-
-\begin{equation}
-    TU_i = \frac{c_i}{\ecfifty}
-    \label{eq:tu}
-\end{equation}
-
-\begin{figure}[h!]
-    \includegraphics[width=1\linewidth]{article/figures/ssd2_boot.png}
-    \caption{SSD plot showing the susceptibility of organisms towards the herbicide Isoproturon.
-The red line represents the fit and the red dashed line marks the HC\textsubscript{5} value, which is a common measure in CRA.
-The legend denotes common organism groups used in ecotoxicology.}
-    \label{fig:ssd-isoproturon}
-\end{figure}
-
-Besides that, two cleaning steps can also be chosen.
-Firstly the application allows to exclude test results exhibiting concentrations that are higher than the actual water solubility of the respective chemical at \ang{20} C.
-Secondly the user can choose to exclude outliers (cf.
-Table \ref{tab:scripts-app}).
-The outliers are selected as values that exceed lower (0.25) and the upper (0.75) quartile by 1.5 times the inter-quartile range.
-
-\subsection*{Identify chemicals to be tested}
-There is a clear need in ecotoxicology for harmonised approaches to collect ecotoxicological data as other data base compilation attempts recently published show.
-Hitchock et al.
-\citet{hitchcock_improving_2018} argue for an increased awareness concerning newly published ecotoxicological test results to facilitate meta analyses across studies.
-Important statistical parameters, such as mean estimates, variances and sample sizes.
-Likewise, state variables, including medium temperature, pH, mineral contents or organism age and sex should be reported (if not in the text, in an online repository).
-Standartox doesn't include such information mainly because the EPA ECOTOX reports only a few of these variables.
-A future research aim could be to use machine learning techniques to retrieve this information from the referenced articles.
-Above all, recent studies showed that, in defiance of many regulation efforts \citep{schafer_future_2019} chemicals can still cause strong adverse effects on organisms and in turn on the environment.
-Neonicotinoids, a group of insecticides was for example related to declines in insectivorous birds \citep{hallmann_declines_2014} and fish yields due to cascading food web effects \citep{yamamuro_neonicotinoids_2019}.
-Standartox provides means to synthesise ecotoxicological test data, thereby identify research gaps.
-
-\begin{figure}
-    \includegraphics[width=1\linewidth]{article/figures/heatmap_tests_n.png}
-    \caption{30 most tested chemicals in EPA ECOTOX.}
-    \label{fig:standartox_ppdb_diff}
-\end{figure}
-
-\fi

### 2019-12-16 article/parts/05_conclusion.tex
Updates from Overleaf

\section{Conclusion}

Standartox is a tool that does this in a generalizing way and puts its focus on the aggregation of toxicity data, by this adding a piece to the puzzle of modern ecotoxicological analyses.

### 2019-12-16 article/parts/02_data_description.tex

--- 2019-11-28 article/parts/02_results_discussion.tex
@@
-\section*{Results and Discussion}
-Standartox is a tool that allows to filter and aggregate ecotoxicological test results to derive single exposure endpoints for individual chemicals.
-It can either be accessed through the APP (Figure \ref{fig:app}) or the API \ref{tab:api-endpoints} together with the R-package standartox (Code example \ref{listing:standartox-example}).
-Both access pathways allow for the same filters and aggregation methods (Table \ref{tab:app-parameters}) to be applied and access the same data set.
-Using the APP, the user can download the resulting data sets as a comma-separated values (csv) file whereas by using the API, the data can directly be loaded in R.
+\section{Data Description}
+
+\subsection{The tool}
+The EPA ECOTOX database contains almost 1,000,000 ecotoxicological test results from almost 12,000 chemicals and about 13,000 taxa \citep{usepa_ecotox_2019} and represents the largest publicly available collection of ecotoxicological test data.
+However, it constitutes only a collection of test results without aiming for summarising the information, which leads to the problem of multiple test results occurrences for individual chemicals.
+To tackle this Standartox introduces filter and aggregation methods that enable the derivation of single ecotoxicological endpoints for specific test parameter combinations.
+Hence, Standartox contains about 600,000 ecotoxicological test results comprising roughly 8000 distinct chemicals and taxa.
+The data can be restricted to three endpoints (XX50, LOEX and NOEX) and filtered to 23 distinct effect groups (e.g.
+mortality, population, growth).
+Furthermore, the data can be refined to five chemical classes (e.g.
+pesticides, metals, drugs) as well as to species-specific habitats (e.g.
+freshwater, marine, terrestrial) and regions (e.g.
+Europe, South America).
+
 Different parameters in order to filter the test data, such as compound-specific parameters, including the CAS number, the concentration type (e.g.
@@
 As the method of aggregation users can choose the minimum, the maximum, or the geometric mean.
-To provide an easy overview of the data set, percentage values for each parameter entry are shown in the APP in the selection menu as well as by using the R package through querying the \textit{/catalog} API-endpoint.
 
-%% Add units part here
+%%%%%%%%%%%%%%%%%%%%%%%%%%%% CONTINUE HERE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\pagebreak
-\begin{figure}
-    \caption{Web application (APP) Standartox allowing to set filter parameters and aggregation methods.
-Data can be downloaded as comma separated value (csv) files.
-An interactive overview plot and table allow for browsing of the data.}
-    \label{fig:app}
-    \includegraphics[width=0.75\linewidth]{article/figures/screenshot_standartox_app.png}
-\end{figure}
+Standartox is a tool that allows to filter and aggregate ecotoxicological test results to derive single exposure endpoints for individual chemicals.
+It can either be accessed through the APP (Figure \ref{fig:app}) or the API \ref{tab:api-endpoints} together with the R-package standartox (Code example \ref{listing:standartox-example}).
 
-\pagebreak
+%% TODO Add units part here
 
-\input{article/code/standartox_example.tex}
-\pagebreak
-
-\begin{figure}
-    \includestandalone[scale=0.4]{article/tikz/standartox_organigram}
-    \caption{Organigram of Standartox.}
-    \label{fig:stx-organigram}
-\end{figure}
 \pagebreak
@@
 Due to an increased amount of available ecotoxicological test data, it becomes fundamental to provide and distribute such information in adequate formats, both easily accessible for humans and easily processable for machines.
-Currently, CRA often relies on test results of a few, well tested standard organisms, such as \textit{Daphnia magna}, \textit{Pimephales promelas}, \textit{Selenastrum capricornutum} to name a few, although researchers have conducted experiments on a much greater variety of organisms, observable in the 12,906 taxa from 952,634 test results collected in the EPA ECOTOX data base.
+
 In order to locate ecotoxicological test data, researchers can rely on individual test results reported in publications, on compiled and published data sets \citep{malaj_organic_2014, morrissey_neonicotinoid_2015} or on data bases such as the Pesticide Property Data Base (PPDB) \citep{lewis_international_2016}.
 Each approach has its limitations though.
-Collecting toxicity information from individual publications is laborious and using already compiled data from publications or published data bases is often limited to specific groups of chemicals, such as pesticides, biozides etc.
-Just as well, these resources run at risk of providing out of date data sooner or later or rarely provide sophisticated means to directly process the data.
+Collecting toxicity information from individual publications is laborious and using already compiled data from publications or published data bases is often limited to specific groups of chemicals, such as pesticides, biocides etc.
+Additionally it is not always clear whether presented effect values are aggregated or single specific tests have been selected.
+These resources can provide out of date data and rarely provide sophisticated means to directly process the data.
+
 Standartox overcomes these paucities by making use of the quarterly updated EPA ECOTOX data base, the largest publicly available collection of ecotoxicological test information.
@@
 
-\subsection*{Other data bases}
-Other initiatives tidying the large number of ecotoxicological test results were also developed recently.
+\subsection{Accuracy Assessment}
+Other initiatives harmonizing the large number of ecotoxicological test results were also developed recently.
 They partly aim for overlapping goals, yet have limitations or objectives that differentiate them from Standartox.
@@
 
-The Network of reference laboratories, research centres and related organisations for monitoring of emerging environmental substances (NORMAN) focuses on assembling river basin specific pollutants \citep{von_der_ohe_new_2011}.
+The Network of reference laboratories, research centres and related organisations for monitoring of emerging environmental substances (NORMAN) focuses on assembling river basin specific pollutants from studies and databases \citep{von_der_ohe_new_2011}.
 The EnviroTox data base \href{https://envirotoxdatabase.org/} which also uses, amongst others the EPA ECOTOX data base as an input was recently published \citep{healthandenvironmentalsciencesinstitutehesi_envirotox_2019, connors_creation_2019}.
@@
 
-\subsection*{Limitations}
-Since the used data in Standartox merely represents a sample of the actual toxicity of chemicals on organisms, it has to be stated that the returned Standartox values are subject to change.
+\subsection{Limitations}
+Since the data in Standartox merely represents a sample of the actual toxicity of chemicals on organisms, it has to be stated that the returned Standartox values are subject to change.
 New studies will be published and constantly incorporated into EPA ECOTOX data base and therefore into Standartox, improving the toxicity estimates.
@@
 \iffalse
+- rename subsections as such:
+
+\subsection*{Accuracy Assessment}
+
+\subsection*{Perspective}
+
 - can also be incorporated in EPA Dashboard

### 2019-12-16 article/parts/03_methods.tex

@@
 %%There is no limit to the length of the Methods section.
-\section*{Methods}
+\section{Methods}
 An automated processing pipeline downloads the quarterly released EPA ECOTOX database, performs several preparation steps on it and exports a final Standartox data set.
@@
 
-\subsection*{Processing pipe line}
+\subsection{Processing pipe line}
 Standartox downloads the quarterly released EPA ECOTOX data base whenever a new version is published and builds it into a local PostgreSQL data base \citep{szocs_build_2019}.
@@
 
-\subsection*{Web application (APP) and Application Programming Interface (API)}
+\subsection{Web application (APP) and Application Programming Interface (API)}
 The web application (APP) and the Application Programming Interface (API) load the compressed serialized Standartox data into memory and allow a client to interact with it.
@@
 
-\subsubsection*{Web application (APP)}
+\subsubsection{Web application (APP)}
 The interactive APP was built in R using the shiny web application framework, which runs with the help of a shiny server \citep{chang_shiny_2018} on \url{standartox.uni-landau.de}.
@@
 
-\subsubsection*{Application programming interface (API)}
+\subsubsection{Application programming interface (API)}
 The API is built by using the R package plumber \citep{trestletechnologyllc_plumber_2018}, which allows for the creation of RESTful APIs from R.
@@
 
-\subsubsection*{R-package}
+\subsubsection{R-package}
 The R-package standartox accesses the API through two functions \textit{stx\_catalog()} and \textit{stx\_query()}.
@@
 
-\subsection*{Technical Validity}
-To guarantee appropriate unit conversion and harmonisation, we compare one automatically converted entry to a manual conversion for every distinct concentration (1229) and duration (129) unit.
+\subsection{Technical Validity}
+To guarantee appropriate unit conversion and harmonisation, we compare for each of the 1229 distinct concentration and for each of the 129 duration units the automatically converted units to one manually calculated one.
 

### 2019-12-16 article/tables/data_base_tables.tex

@@
   PHYSPROP & \makecell{https://www.srcinc.com/what-we-do/environmental/\\scientific-databases.html}  \\[0.5cm]
-  PubChem & https://pubchem.ncbi.nlm.nih.gov \\[0.5cm]
-  Wikidata & https://www.wikidata.org/wiki/Wikidata:Main_Page \\[0.5cm]
+  PubChem & \url{https://pubchem.ncbi.nlm.nih.gov} \\[0.5cm]
+  Wikidata & \url{https://www.wikidata.org/wiki/Wikidata:Main_Page} \\[0.5cm]
   \makecell{Global Biodiversity \\ Information Facility (GBIF)} & https://www.gbif.org \\[0.5cm]

### 2019-12-16 article/tikz/pipeline-organigram-detailed.tex

\documentclass[border = 1.5cm]{standalone}

\usepackage{tikz}
\usetikzlibrary{trees,
                arrows,
                shapes.geometric,
                positioning,
                calc,
                backgrounds,
                fit}
%% document
\begin{document}

\tikzset{
    font = {\fontsize{13pt}{12}\sffamily},
    epa/.style = {rectangle, text centered, draw=black, fill=blue!10, minimum width=3cm, minimum height=3cm},
    script-r/.style = {rectangle, text centered, draw=black}
}

\begin{tikzpicture}

\node (start) [epa, align=center] at (0cm,0cm) {EPA ECOTOX};
%\node (epa-ecotox) [ecotox, above of = epa-ecotox-clean, align=center, yshift = 5cm] {EPA ECOTOX\\raw data};
%\node (epa-ecotox-clean) [ecotox, align = center] at (-3cm,0cm) {EPA ECOTOX\\cleaned};

\node (epa-prepare) [script-r, align=center, right of=start, xshift=3cm] {EPA prepare};

\end{tikzpicture}
\end{document}

### 2019-12-19 article/parts/01_summary.tex

--- 2019-11-22 article/parts/01_introduction.tex
@@
-\section*{Introduction}
+\section{Summary}
 An increasing number of chemicals such as pharmaceuticals, pesticides and synthetic hormones are in daily use all over the world.
@@
 For example, \citet{morrissey_neonicotinoid_2015} used ecotoxicological test results from 49 insect and crustacean species to evaluate the effect of neonicotinoids, an insecticide class on the aquatic ecosystem.
-Also, \citet{malaj_organic_2014} compiled intera alia experimental toxicity test results for 223 various chemical compounds to assess the health of freshwater ecosystems in Europe.
-Similarly, permissible environmental concentrations are often derived from these test data, typically by a combination with safety factors to account for uncertainties.
-Given that data sets with large numbers of analysed chemicals are becoming increasingly available, ecotoxicity data are pivotal for regulatory risk assessment and science.
-To date, only few initiatives exist that aim to create a public resource of harmonized and quality controlled ecotoxicological data, such as the United States Environmental Protection Agency ECOTOXicology knowledgebase (EPA ECOTOX) \citep{elonen_ecotoxicology_2018}, the German Environmental Agency's Information System Ecotoxicology and Environmental Quality Targets (ETOX) \citep{umweltbundesamt_etox_2019} or the Pesticide Property Data Base (PPDB) \citep{lewis_international_2016}.
+Also, \citet{malaj_organic_2014} compiled experimental toxicity test results for 223 various chemical compounds to assess the health of freshwater ecosystems in Europe.
+Similarly, permissible environmental concentrations are often derived from these test data, typically by a combination with safety factors to account for uncertainties, thereby mostly relying on a few, well tested standard organisms, such as the water flea \textit{Daphnia magna}, the fathead minnow \textit{Pimephales promelas}, the microalga \textit{Raphidocelis subcapitata} to name a few, although researchers have conducted experiments on a much greater variety of organisms.
+Given that data sets with large numbers of analysed chemicals are becoming increasingly available, ecotoxicity data are pivotal for regulatory risk assessment (RA) and science.
+To date, only few initiatives exist that aim to create a public resource of harmonized and quality controlled ecotoxicological data, such as the United States Environmental Protection Agency ECOTOXicology knowledgebase (ECOTOX) (~1,000,000 test results, ~13,000 taxa, ~12,000 chemicals) \citep{elonen_ecotoxicology_2018}, the German Environmental Agency's Information System Ecotoxicology and Environmental Quality Targets (ETOX) \citep{umweltbundesamt_etox_2019} or the Pesticide Property Data Base (PPDB) (~ 2000 chemicals) \citep{lewis_international_2016}.
 The former two compile all available test results into a database.
@@
 Hence, these databases can contain multiple entries for the same test without quality information and with different units, hampering reproducible science if different values are selected for studies.
+
 The PPDB database is confined to pesticides and provides harmonised and quality controlled data for a few selected test organisms, thereby ignoring vast amounts of data for other species and chemicals.
 Moreover, data analyses often require links to additional data resources, for example to append additional chemical and species information (e.g.
-chemical properties, habitat of species), which calls for more automated procedures.\\
-
-We developed Standartox, a tool and database that aims to overcome the limitations of other databases by continuously incorporating the ever-growing number of test results in an automated process workflow that ultimately leads to single aggregated data point for a specific chemical-organism test combination, representing the toxicity of a chemical.
+chemical properties, habitat of species), which calls for more automated procedures.
+We therefore developed Standartox, a tool and database that aims to overcome the limitations of other databases by continuously incorporating the ever-growing number of test results in an automated process workflow that ultimately leads to single aggregated data points for a specific chemical-organism test combination.
+Standartox makes use of the publicly available and quarterly released EPA ECOTOX database \citep{usepa_ecotox_2019} and restricts the data to three commonly used endpoints in ecotoxicology (i.e.
+XX50, LOEX \& NOEX) and further harmonises and cleans the data, leading to roughly 600,000 ecotoxicological test results including 8000 chemicals, tested on about 7873 taxa.
+The tool allows user to filter test results according to several parameters, e.g.
+refining a search for ecotoxicity data on organisms occurring in specific habitats or regions of the world.
+Above all, Standartox aggregates ecotoxicological test results in a standardized way, by calculating the minimum, the geometric mean and the maximum of the result for each chemical for user-defined test parameters.
+This relaxes the commonly occurring problem of the variability of multiple ecotoxicological test results and which to choose for chemical risk assessment (CRA).
 Hence, Standartox provides the basis for reproducible science and combines information from different sources to simplify the derivation of more advanced risk indicators such as Species Sensitivity Distributions (SSD) and Toxic Units (TU), which represent two prominent concepts to assess effects on organisms in ecotoxicology \citep{posthuma_species_2002, kefford_definition_2011, schafer_effects_2011}.
 Besides aggregating ecotoxicological test results, Standartox provides a concise overview of the status of tested chemicals, allowing the identification of potential research gaps.
-Standartox comes with two front-ends,  a web application (APP) and a R \citep{rcoreteam_language_2017} package, which accesses a application programming interface (API) providing convenience structures and thereby largely reducing processing time for users.
+Standartox comes with two front-ends, a web application (APP): \url{standartox.uni-landau.de} and a R \citep{rcoreteam_language_2017} package \textit{standartox}, which accesses a application programming interface (API) providing convenience structures and thereby largely reducing processing time for users.
+
+Both access pathways allow for the same filters and aggregation methods (Table \ref{tab:app-parameters}) to be applied and access the same data set.
 
@@
 \iffalse
+
+\item \citep{fernandes_lethal_2016} example for ecotoxicological tests
 

### 2019-12-19 article/tikz/pipeline-organigram.tex
corrections co-authors

@@
 
+\iffalse
+ Könnten Aspekte von oben nicht noch in das Organigramm (e.g.
+Cleaning units and test durations)...
+Weitere Anmerkungen zur Abbildung: Die Daten EPA Ecotox und Standardtox sind irgendwie auf der gleichen Ebene - aber wozu ist die EPA da noch drin?
+Warum geht der flowchart nicht von so, dass irgendwas auf standartox data set mit Pfeil geht?
+Dann wird aus der API das R package?
+Insgesamt schön zur Visualisierung usw., aber die Logik ist unklar und kann einfacher gehalten werden
+\fi
+
 \begin{tikzpicture}
@@
 % pipeline
-\node (epa-ecotox) [ecotox, above of = epa-ecotox-clean, yshift = 5cm] {EPA};
-\node (epa-ecotox-clean) [ecotox, align = center] at (-3cm,0cm) {EPA ECOTOX\\data base\\(cleaned)};
+\node (epa-ecotox) [ecotox, above of = epa-ecotox-clean, align=center, yshift = 5cm] {EPA ECOTOX\\raw data};
+\node (epa-ecotox-clean) [ecotox, align = center] at (-3cm,0cm) {EPA ECOTOX\\cleaned};
 \node (cas) [ids, below of = epa-ecotox-clean, xshift = 6cm, yshift = 3cm] {CAS};

### 2019-12-19 article/parts/02_data_description.tex
Updates from Overleaf

@@
 Since 2006, the PPDB provides validated ecotoxicological information on pesticides \citep{lewis_international_2016}.
+The aggregation of Standartox matches PPDB (n=3601) values that have been manually quality checked in the 91 \% within 1 order of magnitude.
+Considering only Standartox values for which at least ten tests were aggregated the value increases to 93.4 \%.
+Similarly, we compared Standartox to values from the ChemProp \citep{ufzdepartmentofecologicalchemistry_chemprop_2019} software, which calculates LC\textsubscript{50} values by the Read Across method \citep{schuurmann_quantitative_2011} and found that (n=179) 95 \% lie within a factor 10 difference.
+However, the difference is not necessarily an indication of lower quality of these data in Standartox but may also be attributed to the inclusion of consideration of a wider range of test data and conditions as well as variance in the Read Across models, respectively.
+
 In order to asses the validity of the Standartox aggregation approach, we compared Standartox results with effect values from the PPDB (n=3601) for the same chemicals (Figure \ref{fig:standartox_ppdb_diff}).
-This showed that 91.3\% of the Standartox values diverge by less than a factor of ten from the PPDB values.
-Considering only Standartox values for which at least five or ten tests were aggregated the value increases to 91.7\% and 93.4\%, respectively.
+
   Admittedly a few Standartox values diverge by a large margin.
@@
 
-\subsection{Limitations}
+\citet{vandenberg_modeling_2019} used ECOTOX for trait-based sensitivity models. \include{article/values/publish_rate.tex}
+
+\subsection{Outlook}
+Standartox is not a one point in time effort, but an automated tool that steadily incorporates newly published ecotoxicity data.
+The ECOTOX database added on average \verbatiminput{article/values/test.txt} new test results in the five last years.
+
+Standartox is not a singular proposition of a tool that aggregates large ecotoxicity data, but constitutes an automated tool that
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%% CONTINUE HERE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%55
+
 Since the data in Standartox merely represents a sample of the actual toxicity of chemicals on organisms, it has to be stated that the returned Standartox values are subject to change.

### 2019-12-21 R/id_epa_tax_prep.R
changes

@@ -24,12 +24,16 @@ tax[ phylum_division == 'Cryptophycophyta', `:=` (algae = 1L, autotroph = 1L, ec
 tax[ phylum_division == 'Prasinophyta', `:=` (algae = 1L, autotroph = 1L, ecotox_group2 = 'Algae') ]
 tax[ phylum_division == 'Ochrophyta', `:=` (algae = 1L, autotroph = 1L, ecotox_group2 = 'Algae') ]
 tax[ class == 'Dinophyceae', `:=` (algae = 1L, autotroph = 1L, mixotroph = 1L, ecotox_group2 = 'Algae') ]
+tax[ class == 'Chloromonadophyceae', `:=` (algae = 1L, autotroph = 1L, mixotroph = 1L, ecotox_group2 = 'Algae') ]
 
 # plants ------------------------------------------------------------------
 tax[ kingdom == 'Plantae', `:=` (plants = 1L, autotroph = 1L, ecotox_group2 = 'Plantae') ]
 
 # mollusca ----------------------------------------------------------------
 tax[ phylum_division == 'Annelida', `:=` (annelida = 1L, heterotroph = 1L, invertebrates_makro = 1L, ecotox_group2 = 'Annelida') ]
+
+# Brachiopoda -------------------------------------------------------------
+tax[ phylum_division == 'Brachiopoda', `:=` (brachiopoda = 1L, heterotroph = 1L, invertebrates_makro = 1L, ecotox_group2 = 'Brachiopoda') ]
 
 # mollusca ----------------------------------------------------------------
 tax[ phylum_division == 'Mollusca', `:=` (mollusca = 1L, heterotroph = 1L, invertebrates_makro = 1L, ecotox_group2 = 'Mollusca') ]
@@ -41,10 +45,22 @@ tax[ phylum_division == 'Nemertea', `:=` (nemertea = 1L, heterotroph = 1L, inver
 tax[ phylum_division == 'Platyhelminthes', `:=` (nemertea = 1L, heterotroph = 1L, invertebrates_makro = 1L, ecotox_group2 = 'Platyhelminthes') ]
 
 # nematoda ----------------------------------------------------------------
-tax[ phylum_division == 'Nemata', `:=` (nematoda = 1L, heterotroph = 1L, invertebrates_mikro = 1L, ecotox_group2 = 'Nemata') ]
[…]

### 2019-12-21 R/qu_pp_prep.R

@@ -27,7 +27,8 @@ pp_l = lapply(pp_l, pp_resolve)
 pp = rbindlist(pp_l, fill = TRUE, idcol = 'cas')
 setnames(pp,
          c('Water Solubility', 'Log P (octanol-water)'),
-         c('solubility_water', 'p_log'))
+         c('solubility_water', 'p_log'),
+         skip_absent = TRUE)
 setnames(pp, tolower(names(pp)))
 setcolorder(pp, c('cas', 'cname'))
 pp[, cname := tolower(cname)]

### 2020-01-17 article/parts/02_data_description.tex
Updates from Overleaf

@@
 \subsection{The tool}
-The EPA ECOTOX database contains almost 1,000,000 ecotoxicological test results from almost 12,000 chemicals and about 13,000 taxa \citep{usepa_ecotox_2019} and represents the largest publicly available collection of ecotoxicological test data.
-However, it constitutes only a collection of test results without aiming for summarising the information, which leads to the problem of multiple test results occurrences for individual chemicals.
+%% Filter
+The ECOTOX database contains almost 1,000,000 ecotoxicological test results from almost 12,000 chemicals and about 13,000 taxa \citep{usepa_ecotox_2019} and represents the largest publicly available collection of ecotoxicological test data.
+However, it constitutes only a collection of test results without aiming for a complete quality check or for summarising the information, the latter leading to the problem of multiple test result occurrences for individual chemicals.
 To tackle this Standartox introduces filter and aggregation methods that enable the derivation of single ecotoxicological endpoints for specific test parameter combinations.
-Hence, Standartox contains about 600,000 ecotoxicological test results comprising roughly 8000 distinct chemicals and taxa.
+Hence, Standartox contains about 600,000 ecotoxicological test results comprising roughly 8000 distinct chemicals and 8000 taxa.
 The data can be restricted to three endpoints (XX50, LOEX and NOEX) and filtered to 23 distinct effect groups (e.g.
@@
 Europe, South America).
-
-Different parameters in order to filter the test data, such as compound-specific parameters, including the CAS number, the concentration type (e.g.
+Additionally, different parameters in order to filter the test data, such as compound-specific parameters, including the CAS number, the concentration type (e.g.
 Active Ingredient, Formulation) or the chemical class (e.g.
@@
 mortality, population, growth) and endpoints (EC\textsubscript{50}, LOEC or NOEC).
-As the method of aggregation users can choose the minimum, the maximum, or the geometric mean.
+
+%%%% TODO plot: ~/standartox-build/article/figures/standartox_paramteres.png %%%%
+
+list how many freshwater species, marine etc.
+how many regions
+how many chemical groups
+
+%% Aggregation
+
+In order to deal with multiple test result occurrences of individual chemicals, Standartox enables the aggregation of the data by calculating the minimum, the maximum, or the geometric mean.
+
+filtered data set + flags (for outliers)
+aggregated data set
+meta
+
+%%%% TODO: plot geometric mean above all data for a specific chemical (like on poster)
+
+%% Accuracy assessment
+
+%% General info and summary at the end
+Standartox provides quick access.
+
+Standartox is designed to foster chemical risk assessment (CRA).
+Due to an increased amount of available ecotoxicological test data, it becomes fundamental to provide and distribute such information in adequate formats, both easily accessible for humans and easily processable for machines.
 
@@
 \pagebreak
-
-Standartox is designed to foster chemical risk assessment (CRA).
-Due to an increased amount of available ecotoxicological test data, it becomes fundamental to provide and distribute such information in adequate formats, both easily accessible for humans and easily processable for machines.
 
@@
 The aggregation of Standartox matches PPDB (n=3601) values that have been manually quality checked in the 91 \% within 1 order of magnitude.
-Considering only Standartox values for which at least ten tests were aggregated the value increases to 93.4 \%.
+Considering only Standartox values for which at least five tests were aggregated the value increases to 92.7 \%.
 Similarly, we compared Standartox to values from the ChemProp \citep{ufzdepartmentofecologicalchemistry_chemprop_2019} software, which calculates LC\textsubscript{50} values by the Read Across method \citep{schuurmann_quantitative_2011} and found that (n=179) 95 \% lie within a factor 10 difference.
@@
 \textit{Daphnia magna}), and the tool lacks the possibility to retrieve predicted values in an automated way \citep{williams_comptox_2017}.
-Petschick et al.
-(in submission) modeled risk threshold level equivalents for aquatic organisms by using ecotoxicological effect data from the EPA ECOTOX data base to support regulation.
+\citet{petschick_modeling_2019} modeled risk threshold level equivalents for aquatic organisms by using ecotoxicological effect data from the EPA ECOTOX data base to support regulation.
 In summary, none of the above mentioned approaches aim for a holistic standardized aggregation method of exposure endpoints for individual chemicals and also lack the possibility for automated and scriptable user requests from common high level programming languages, such as R or Python.
@@
 
-\citet{vandenberg_modeling_2019} used ECOTOX for trait-based sensitivity models. \include{article/values/publish_rate.tex}
+\citet{vandenberg_modeling_2019} used ECOTOX for trait-based sensitivity models.
 
 \subsection{Outlook}
-Standartox is not a one point in time effort, but an automated tool that steadily incorporates newly published ecotoxicity data.
-The ECOTOX database added on average \verbatiminput{article/values/test.txt} new test results in the five last years.
-
-Standartox is not a singular proposition of a tool that aggregates large ecotoxicity data, but constitutes an automated tool that
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%% CONTINUE HERE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%55
-
-Since the data in Standartox merely represents a sample of the actual toxicity of chemicals on organisms, it has to be stated that the returned Standartox values are subject to change.
-New studies will be published and constantly incorporated into EPA ECOTOX data base and therefore into Standartox, improving the toxicity estimates.
-However, we expect values of well tested chemical-organism combinations to be relatively stable.
+Standartox is not singular effort, but an automated tool that steadily incorporates newly published ecotoxicity data.
+The ECOTOX database added on average 5103\rpm1317 new test results every four months in the five last years.
+This provides a constantly growing source of ecotoxicity data which is in turn directly incorporated into Standartox which steadily improves its toxicity estimates.
+Since Standartox toxicity estimates will change and to guarantee reproducibility we allow users to access older versions of Standartox.
+Yet, we are certain, that Standartox aggregates of well tested chemical-organism combinations to be stable when new test results are added, because the geometric mean is not strongly influenced by outliers.
+This can be seen from the, above mentioned comparison to the PPDB, which shows strong evidence for coherence already when aggregating only five values.
+At most, minima and maxima could potentially shift greatly.
+Therefore, we flag outliers in the filtered data set to point users to possible extreme values.
 A future analysis on the smallest number of tests needed to return trustworthy results, could shed more light on this aspect.
 Furthermore, toxicity test results are certainly influenced by individual tests parameters, such as pH, temperature or conductivity amongst others \citep{rosenkrantz_influence_2013, li_temperature_2011}.
-Such information is not included in the aggregation performed by Standartox, since the EPA ECOTOX data base only provides sparse records on those.
+Such information is not included in the aggregation performed by Standartox, since the ECOTOX database only provides sparse records on those.
 Most frequently provided test parameters (and their percentage) are medium temperature (77\%), pH:56\%, hardness:27\%, dissolved oxygen:18\%, Alkalinity:15\% and salinity:9\%.
@@
 A text-mining approach, iterating through the individual publications could potentially increase this number.
-However, we argue that using the geometric mean as an aggregate allows for an adequate estimation of the toxicity of a chemical towards an organism group.
+
 We chose the geometric mean in comparison to the arithmetic mean, since it is considerably less influenced by outliers and is suitable for skewed data.

### 2020-01-20 article/parts/02_data_description.tex

@@
 
-\subsection{The tool}
-%% Filter
-The ECOTOX database contains almost 1,000,000 ecotoxicological test results from almost 12,000 chemicals and about 13,000 taxa \citep{usepa_ecotox_2019} and represents the largest publicly available collection of ecotoxicological test data.
-However, it constitutes only a collection of test results without aiming for a complete quality check or for summarising the information, the latter leading to the problem of multiple test result occurrences for individual chemicals.
-To tackle this Standartox introduces filter and aggregation methods that enable the derivation of single ecotoxicological endpoints for specific test parameter combinations.
-Hence, Standartox contains about 600,000 ecotoxicological test results comprising roughly 8000 distinct chemicals and 8000 taxa.
-The data can be restricted to three endpoints (XX50, LOEX and NOEX) and filtered to 23 distinct effect groups (e.g.
+Standartox contains about 600,000 ecotoxicological test results comprising roughly 8000 distinct chemicals and 8000 taxa.
+The data can be restricted to three endpoint groups (XX50, LOEX and NOEX) and filtered to 23 distinct effect groups (e.g.
 mortality, population, growth).
@@
 mortality, population, growth) and endpoints (EC\textsubscript{50}, LOEC or NOEC).
+Figure (\ref{fig:stx-parameters}).
 
-%%%% TODO plot: ~/standartox-build/article/figures/standartox_paramteres.png %%%%
+\begin{figure}
+    \includegraphics[width=1.0\textwidth]{article/figures/standartox_parameters.png}
+    \caption{Parameters in Standartox.}
+    \label{fig:stx-parameters}
+\end{figure}
 
-list how many freshwater species, marine etc.
-how many regions
-how many chemical groups
-
-%% Aggregation
-
-In order to deal with multiple test result occurrences of individual chemicals, Standartox enables the aggregation of the data by calculating the minimum, the maximum, or the geometric mean.
+This constitutes a collection of quality checked test results without aggregating, which leads to the problem of multiple test result occurrences for individual chemicals.
+To tackle this, Standartox introduces aggregation methods that enable the derivation of single ecotoxicological endpoints for specific test parameter combinations.
+Standartox therefore calculates the minimum, the maximum and the geometric mean for chosen test combinations in order to derive a single toxicity endpoint from it.
+This aggregation step is important because multiple test results lead to highly variable results, which in turn can lead to differences in chemical risk assessment (CRA) by different researchers.
+To derive one single value out of many test results, people can perform qualitative assessments and pick the most adequate test and the associated toxicity equivalent or like Standartox use aggregation methods..
 
@@
 
+\begin{figure}
+    \includegraphics[width=1.0\textwidth]{article/figures/results_variability.png}
+    \caption{Variability of test results.}
+    \label{fig:stx-organigram}
+\end{figure}
 %%%% TODO: plot geometric mean above all data for a specific chemical (like on poster)
@@
 
+The ECOTOX database contains almost 1,000,000 ecotoxicological test results from almost 12,000 chemicals and about 13,000 taxa \citep{usepa_ecotox_2019} and represents the largest publicly available collection of ecotoxicological test data.
+However, it constitutes only a collection of test results without aiming for a complete quality check or for summarising the information, the latter leading to the problem of multiple test result occurrences for individual chemicals.
+
 \iffalse

### 2020-01-21 article/parts/02_data_description.tex

@@
 Standartox contains about 600,000 ecotoxicological test results comprising roughly 8000 distinct chemicals and 8000 taxa.
-The data can be restricted to three endpoint groups (XX50, LOEX and NOEX) and filtered to 23 distinct effect groups (e.g.
+The data can be restricted to three endpoint groups XX50, LOEX and NOEX, conflating commonly used endpoints such as half maximal effective concentration (EC\textsubscript{50}) values, no observed effect level/concentration (NOEC/L) values, as well as lowest observed effect level/concentration (LOEC/L) values respectively \ref{tab:endpoints-conflate}.
+Standartox allows users to filter to 23 distinct effect groups (e.g.
 mortality, population, growth).
@@
 
-This constitutes a collection of quality checked test results without aggregating, which leads to the problem of multiple test result occurrences for individual chemicals.
+Standartox constitutes a collection of quality checked test results without aggregating, which leads to the problem of multiple test result occurrences for individual chemicals.
 To tackle this, Standartox introduces aggregation methods that enable the derivation of single ecotoxicological endpoints for specific test parameter combinations.
 Standartox therefore calculates the minimum, the maximum and the geometric mean for chosen test combinations in order to derive a single toxicity endpoint from it.
-This aggregation step is important because multiple test results lead to highly variable results, which in turn can lead to differences in chemical risk assessment (CRA) by different researchers.
-To derive one single value out of many test results, people can perform qualitative assessments and pick the most adequate test and the associated toxicity equivalent or like Standartox use aggregation methods..
+This aggregation step is important because multiple test results exhibit highly variable results, which in turn lead to uncertainties in chemical risk assessment (CRA).
+Here, variability results from many factors, such as differences in the susceptibility of different taxa to a chemical (Fig.
+\ref{fig:stx-variability} A), test results variability of one and the same taxon and chemical due to different test durations (Fig.
+\ref{fig:stx-variability} B) as well as variability caused by unexplainable factors, due to different laboratories, physiological differences between organism populations or other un-accounted variation (Fig.
+\ref{fig:stx-variability} C).
+
+\begin{figure}
+    \includegraphics[width=1.0\textwidth]{article/figures/results_variability.png}
+    \caption{Variability of test results (EC\textsubscript{50}) in Standartox due to (A) the susceptibility of different taxa (i.e.
+\textit{Xenopus laevis}, \textit{Pseudokirchneriella subcapitata}, \textit{Oncorhynchus mykiss}, \textit{Lemna minor}) to the chemical atrazine tested for 96h, due to (B) different durations of tests of zinc sulfate on \textit{Daphnia magna} and due to (C) un-accountable variability of cupric sulfate tested on \textit{Pimephales promelas} for 96h.
+Yellow dots depict Standartox approximations and black and gray dots show un-aggregated values for each group.}
+    \label{fig:stx-variability}
+\end{figure}
+
+Another approach would be to derive single toxicity equivalents via rule-based approaches, demanding considerable amounts of time.
+
+Standartox will be subject to change
+
+In the course of the aggregation process, outliers that exceed 1.5 times the interquartile range are flagged to caution practitioners.
+However, since the geometric mean is not strongly influenced by such entries, outliers are not removed from the Standartox output.
+
+Researchers and practitioners spend a consdiderable amount of time on
+
+In order to derive one single value out of many test results, Standartox takes the
+
+people can perform qualitative assessments and pick the most adequate test and the associated toxicity equivalent or like Standartox use aggregation methods..
 
@@
 meta
-
-\begin{figure}
-    \includegraphics[width=1.0\textwidth]{article/figures/results_variability.png}
-    \caption{Variability of test results.}
-    \label{fig:stx-organigram}
-\end{figure}
-%%%% TODO: plot geometric mean above all data for a specific chemical (like on poster)
 

### 2020-01-21 article/tables/endpoint-conflate.tex

% Please add the following required packages to your document preamble:
% \usepackage[normalem]{ulem}
% \useunder{\uline}{\ul}{}
\begin{table}[]
\begin{tabular}{|l|l|}
\label{tab:endpoints-conflate}
\catption{Table of Standartox endpoints and their corresponding Ecotox endpoints and a description.}
\hline
Standartox endpoint & Ecotox endpoint & Ecotox endpoint description \\ \hline
XX50 & LC50 & Lethal concentration to 50\% of test organisms \\ \hline
XX50 & LD50 & Lethal dose to 50\% of test organisms \\ \hline
XX50 & EC50 & Effective concentration to 50\% of test organisms \\ \hline
XX50 & ED50 & Effective dose to 50\% of test organisms \\ \hline
XX50 & IC50 & Inhibition concentration to 50\% of test organisms \\ \hline
XX50 & ID50 & Inhibition dose to 50\% of test organisms \\ \hline
XX50 & ET50 & Effective response time to 50\% of test organisms \\ \hline
XX50 & LT50 & Time to 50\% mortality of organisms \\ \hline
NOEX & NOEC & No-observable-effect-concentration \\ \hline
NOEX & NOEL & No-observable-effect-level \\ \hline
LOEX & LOEC & Lowest observable effect concentration\\ \hline
LOEX & LOEL & Lowest-observable-effect-level \\ \hline
\end{tabular}
\end{table}

### 2020-01-27 article/parts/01_summary.tex
Updates from Overleaf

--- 2019-11-22 article/parts/01_introduction.tex
@@
 In the environment, chemicals can adversely affect populations and communities and in turn related ecosystem functions \citep{rhind_anthropogenic_2009, clements_community_2009, hallmann_declines_2014, barracaracciolo_pharmaceuticals_2015, johnston_review_2015, hallmann_more_2017, sanchez-bayo_worldwide_2019}.
-Ultimately, this may compromise natures contribution to human well-being, for example the ecosystem services clean drinking and irrigation water, food and climate regulation \citep{peters_review_2013, vandersluijs_neonicotinoids_2013, barracaracciolo_pharmaceuticals_2015, yamamuro_neonicotinoids_2019}.
+Ultimately, this may compromise natures contribution to human well-being, for example the ecosystem services clean drinking and irrigation water as well as food production \citep{peters_review_2013, vandersluijs_neonicotinoids_2013, barracaracciolo_pharmaceuticals_2015, yamamuro_neonicotinoids_2019}.
 Pollution with man-made chemicals has been identified as one of three major environmental problems for which research gaps hamper the derivation of planetary boundaries, i.e.
@@
 For example, \citet{morrissey_neonicotinoid_2015} used ecotoxicological test results from 49 insect and crustacean species to evaluate the effect of neonicotinoids, an insecticide class on the aquatic ecosystem.
-Also, \citet{malaj_organic_2014} compiled intera alia experimental toxicity test results for 223 various chemical compounds to assess the health of freshwater ecosystems in Europe.
-Similarly, permissible environmental concentrations are often derived from these test data, typically by a combination with safety factors to account for uncertainties.
-Given that data sets with large numbers of analysed chemicals are becoming increasingly available, ecotoxicity data are pivotal for regulatory risk assessment and science.
-To date, only few initiatives exist that aim to create a public resource of harmonized and quality controlled ecotoxicological data, such as the United States Environmental Protection Agency ECOTOXicology knowledgebase (EPA ECOTOX) \citep{elonen_ecotoxicology_2018}, the German Environmental Agency's Information System Ecotoxicology and Environmental Quality Targets (ETOX) \citep{umweltbundesamt_etox_2019} or the Pesticide Property Data Base (PPDB) \citep{lewis_international_2016}.
+Also, \citet{malaj_organic_2014} compiled experimental toxicity test results for 223 various chemical compounds to assess the health of freshwater ecosystems in Europe.
+Similarly, permissible environmental concentrations are often derived from these test data, typically by a combination with safety factors to account for uncertainties, thereby mostly relying on a few, well tested standard organisms, such as the water flea \textit{Daphnia magna}, the fathead minnow \textit{Pimephales promelas}, the microalga \textit{Raphidocelis subcapitata} to name a few.
+Researchers have conducted experiments on a much greater variety of organisms though.
+Given that data sets with large numbers of analysed chemicals are becoming increasingly available, ecotoxicity data are pivotal for regulatory risk assessment (RRA) and science.
+To date, only few initiatives exist that aim to create a public resource of harmonized and quality controlled ecotoxicological data, such as the United States Environmental Protection Agency ECOTOXicology knowledgebase (ECOTOX) (ca.
+1,000,000 test results, ca.
+13,000 taxa, ca.
+12,000 chemicals) \citep{elonen_ecotoxicology_2018}, the German Environmental Agency's Information System Ecotoxicology and Environmental Quality Targets (ETOX) \citep{umweltbundesamt_etox_2019}, the Pesticide Property Data Base (PPDB) (ca.
+2000 chemicals) \citep{lewis_international_2016} or the Envirotox database \citep{healthandenvironmentalsciencesinstitutehesi_envirotox_2019, connors_creation_2019}.
 The former two compile all available test results into a database.
-However, for many chemicals, multiple ecotoxicity values are available for the same test organism, which can vary strongly.
-Hence, these databases can contain multiple entries for the same test without quality information and with different units, hampering reproducible science if different values are selected for studies.
-The PPDB database is confined to pesticides and provides harmonised and quality controlled data for a few selected test organisms, thereby ignoring vast amounts of data for other species and chemicals.
+However, for many chemicals, multiple ecotoxicity values are available for the same test organism, which can vary strongly and the lack of associated quality information and different units hamper reproducible science.
+The PPDB database in contrast provides single ecotoxicity values only for pesticides and a few selected test organisms, thereby ignoring vast amounts of data for other species and chemicals.
+Envirotox puts its focus on aquatic organisms.
 Moreover, data analyses often require links to additional data resources, for example to append additional chemical and species information (e.g.
-chemical properties, habitat of species), which calls for more automated procedures.\\
-
-We developed Standartox, a tool and database that aims to overcome the limitations of other databases by continuously incorporating the ever-growing number of test results in an automated process workflow that ultimately leads to single aggregated data point for a specific chemical-organism test combination, representing the toxicity of a chemical.
+chemical properties, habitat of species), which calls for more automated procedures.
+We therefore developed Standartox, a tool and database that aims to overcome the limitations of other databases by continuously incorporating the ever-growing number of test results in an automated process workflow that ultimately leads to a cleaned and harmonised ecotoxicity collection and single aggregated ecotoxicity values for a specific chemical-organism test combination.
+Standartox makes use of the publicly available and quarterly released ECOTOX database \citep{usepa_ecotox_2019} and restricts the data to three commonly used endpoints in ecotoxicology (i.e.
+XX50, LOEX \& NOEX), leading to about 600,000 ecotoxicological test results including 8000 chemicals, tested on about 8000 taxa.
+Standartox allows users to filter test results according to several parameters, e.g.
+refining a search for ecotoxicity data on organisms occurring in specific habitats or regions of the world.
+Above that, Standartox aggregates ecotoxicological test results in a standardized way, by calculating the minimum, the geometric mean and the maximum of the result for each chemical and the associated, user-defined test parameters.
+This relaxes the commonly occurring problems of the variability of multiple ecotoxicological test results and which values to choose for chemical risk assessment (CRA).
 Hence, Standartox provides the basis for reproducible science and combines information from different sources to simplify the derivation of more advanced risk indicators such as Species Sensitivity Distributions (SSD) and Toxic Units (TU), which represent two prominent concepts to assess effects on organisms in ecotoxicology \citep{posthuma_species_2002, kefford_definition_2011, schafer_effects_2011}.
 Besides aggregating ecotoxicological test results, Standartox provides a concise overview of the status of tested chemicals, allowing the identification of potential research gaps.
-Standartox comes with two front-ends,  a web application (APP) and a R \citep{rcoreteam_language_2017} package, which accesses a application programming interface (API) providing convenience structures and thereby largely reducing processing time for users.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% IDEAS %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\iffalse
-
-\item Put in citation \citep{hartung_chemical_2009} ?
-which claims that REACH won't meet their assumptions and say 54 million vertebrate animals are needed for tests that would cost €9.5 billion over the next ten years.
-I.e 20 times more animals, 6 times the costs in comparison to the official estimates
-
-For dicussion:
-Showed general decline in biodiversity and biomass, but no or not much causes: \citep{butchart_global_2010, hallmann_more_2017, seibold_arthropod_2019}
-
-Direct effect of chemicals:
-include \citep{yamamuro_neonicotinoids_2019} % ecosystem services, food webs
-
-- gerneral stuff about small water bodies: \citep{riley_small_2018}
-
-\item standardized toxicity values
-
-\item reproducible research % TODO mention much more often!
-
-\item time saving
-\fi
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% OLD %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\iffalse
-
-A large number of chemicals such as pharmaceuticals, pesticides and synthetic hormones are in daily use all over the world with insufficient knowledge about possible adverse effects on the environment.
-In Europe alone some 100,000 chemicals are estimated to be in current use, whereof 30,000 are produced in quantities larger than one ton per year \citep{breithaupt_costs_2006}.
-Chemicals are brought to the environment deliberately - as it is the case of pesticides, or arrive therein as a byproduct from other processes (e.g.
-atmospheric emissions or wastewater) \citep{schwarzenbach_challenge_2006}.
-Ecosystems provide essential services to human societies such as drinking and irrigation water, food and climate regulation.
-These services are products of different ecosystem functions that crucially depend on the integrity of the populations and communities that drive these ecosystem functions.
-However, besides habitat degradation, climate change and nutrient enrichment, pollution with man-made chemical toxicants threatens these populations and communities in various ways which are currently not fully understood \citep{steffen_anthropocene_2007}.
-Pollution with man-made chemical toxicants was indeed identified as one of three major environmental problems for which research gaps hamper the derivation of planetary boundaries, i.e.
-thresholds beyond which irreversible state shifts may occur \citep{steffen_anthropocene_2007}.
-Bernhardt et al.
-\citet{bernhardt_synthetic_2017} argue further that the knowledge gap how chemicals effect populations and communities and hence ecosystem functions and ecosystem services, would also impede society's ability to accomplish the Sustainable Development Goals of the United Nations.
-According to Breithaupt \citet{breithaupt_costs_2006} less than one percent of chemicals released to the environment are thoroughly tested.
-Although for some chemicals advanced and standardized \citep{oecd_oecd_2018} test methods have been developed, ecotoxicological tests for a lot of chemicals, especially newly emerging ones remain scarce and if existent, the information is represented sparsely and in inconsistent formats.
-\citep{gessner_fostering_2016}.
-Therefore we developed a data base that collects, processes and aggregates ecotoxicological tests results in order to subsequently publish it in an harmonized form as a web application - the \etoxbase{} tool: \href{http://139.14.20.252:3838/etox-base-shiny/}.
-This should facilitate researchers the access to ecotoxicological test results and support modern chemical risk assessment (CRA) workflows.
-The \etoxbase{} makes use of ecotoxicological test data, obtained from the ECOTOXicology knowledgebase (ECOTOX) created by the U.S.
-Environmental Protection Agency (EPA).
-ECOTOX collects raw, non harmonized ecotoxicological test data on aquatic and terrestrial wildlife as well as plants and publishes it on a quarterly basis.
-By the time of writing it contained 926,108 test results, comprising 11,685 chemicals tested on 12,668 different species \citep{elonen_ecotoxicology_2018}.
-The collected ecotoxicological test data contain a variety of measured endpoints such as Effective Concentrations (\ecfifty{}) values, No-observed effect concentrations (NOEC) or lowest observed effect concentrations (LOEC).
-Likewise the data set contains different effect measures such as mortality, intoxication and growth as well as a multiplicity of test durations ranging from seconds to weeks or even years.
-The \etoxbase{} aggregates this information according to the user's inputs.
-These can then be used for the derivation of risk indicators such as Species Sensitivity Distributions (SSDs) \citep{posthuma_species_2002} and Toxic Units (TUs), which represent two prominent concepts to assess effects on organisms in ecotoxicology.
-The former combines toxicity data of several organism groups towards a chemicals to estimate effects on biotic communities, the latter refers to effects of a specific organism (group) towards a chemical.
-The two concepts are widely used in ecotoxicology \citep{kefford_definition_2011, schafer_effects_2011} since the allow for a comparison of toxicities across multiple chemicals and biological communities, respectively.
-Although performed on one and the same organism and chemical, outcomes of different ecotoxicological tests vary greatly due to differences in test parameters, genetic differences in individual populations or other irrepressible factors making a selection of appropriate test results for CRA laborious.
-The \etoxbase{} aims to relief this process and provide single toxicity values for organism, chemical and test duration combinations.
-The \etoxbase{} downloads every new version of the ECOTOX data base and performs several quality checks, e.g.
-unit harmonization, and cleaning steps on it.
-Hence, new scientific test results are constantly incorporated.
-Additionally the \etoxbase{} queries chemical- and organism-specific variables from other open data bases to fill information gaps such as water solubility, organism habitat or regional occurrence patterns in the ECOTOX data base.
-
-\fi
+Moreover, Standartox could help in reducing the millions of animals used for toxicity testing each year \citep{hartung_chemical_2009}.
+Standartox comes with two front-ends, a web application (APP): \url{standartox.uni-landau.de} and a R \citep{rcoreteam_language_2017} package \textit{standartox}, providing convenience structures and thereby largely reducing processing time for users.

### 2020-01-27 article/parts/02_data_description.tex
Updates from Overleaf

--- 2020-01-17 article/parts/02_data_description.tex
@@
 \section{Data Description}
+Standartox constitutes a collection of 600,000 quality checked ecotoxicological test results comprising roughly 8000 distinct chemicals and 8000 taxa that allows for the retrieval of single toxicity equivalents.
+It is build on the ECOTOX database \citep{usepa_ecotox_2019} and the data is processed in order to clean and harmonise entries to retrieve comparable toxicity endpoints.
+Subsequently, filter and aggregation methods are provided to allow for the retrieval of single toxicity equivalents for specific test parameter combinations.
+The underlying ECOTOX database is released quarterly and provides on average \input{article/values/publish_rate.txt} (2014 - 2019) new test results with each release.
+Standartox is build directly after each new ECOTOX release in an automated way and therefore incorporates new data steadily.
+To guarantee reproducibility, users can choose which Standartox version to access.
+Users can reach Standartox via an web app and an R-package (standartox) that accesses an application programming interface (API) providing the possibility of scriptable requests.
 
-\subsection{The tool}
-%% Filter
-The ECOTOX database contains almost 1,000,000 ecotoxicological test results from almost 12,000 chemicals and about 13,000 taxa \citep{usepa_ecotox_2019} and represents the largest publicly available collection of ecotoxicological test data.
-However, it constitutes only a collection of test results without aiming for a complete quality check or for summarising the information, the latter leading to the problem of multiple test result occurrences for individual chemicals.
-To tackle this Standartox introduces filter and aggregation methods that enable the derivation of single ecotoxicological endpoints for specific test parameter combinations.
-Hence, Standartox contains about 600,000 ecotoxicological test results comprising roughly 8000 distinct chemicals and 8000 taxa.
-The data can be restricted to three endpoints (XX50, LOEX and NOEX) and filtered to 23 distinct effect groups (e.g.
-mortality, population, growth).
-Furthermore, the data can be refined to five chemical classes (e.g.
-pesticides, metals, drugs) as well as to species-specific habitats (e.g.
-freshwater, marine, terrestrial) and regions (e.g.
-Europe, South America).
-Additionally, different parameters in order to filter the test data, such as compound-specific parameters, including the CAS number, the concentration type (e.g.
-Active Ingredient, Formulation) or the chemical class (e.g.
-insecticides, metals) as well as taxon-specific parameters, including common taxonomic groups (e.g.
-Daphniidae, algae), the habitat (e.g.
-freshwater) or the regional occurrence (e.g.
-Europe, Asia) of the organisms can be chosen.
-Furthermore, users can refine the results to specific test durations (in hours), effect groups (e.g.
-mortality, population, growth) and endpoints (EC\textsubscript{50}, LOEC or NOEC).
+\subsection{Filters}
+The data can be restricted to three endpoint groups XX50, LOEX and NOEX, conflating commonly used endpoints such as half maximal effective concentrations (EC\textsubscript{50}), no observed effect levels/concentrations (NOEC/L), as well as lowest observed effect levels/concentrations (LOEC/L) respectively (Table \ref{tab:endpoints-conflate}) (Fig.
+\ref{fig:stx-parameters}A).
+Standartox allows the ecotoxicity data to be filtered to 23 distinct effect groups (e.g.
+mortality, population, growth) (Fig.
+\ref{fig:stx-parameters}B) and concentration types (e.g.
+formulation, active ingredient).
+In addition to these test-specific parameters Standartox data entries can be filtered according to chemical-specific parameters such as, the CAS number and five chemical classes (e.g.
+pesticides, metals, drugs), commonly used in ecotoxicology (Fig.
+\ref{fig:stx-parameters}C).
+Furthermore, the Standartox data can be refined to organism-specific parameters, such as the organism habitat (e.g.
+freshwater, marine, terrestrial) (Fig.
+\ref{fig:stx-parameters}E), the occurrence regions (e.g.
+Europe, South America) (Fig.
+\ref{fig:stx-parameters}F) of the organisms as well as different taxonomic levels.
+Additionally users can refine the results to specific test durations (in hours).
 
-%%%% TODO plot: ~/standartox-build/article/figures/standartox_paramteres.png %%%%
+\begin{figure}[h!]
+    \includegraphics[width=1.0\textwidth]{article/figures/standartox_parameters.png}
+    \caption{Share of 10 most occurring entries for the parameters endpoint (A), effect (B), chemical class (C), taxonomic order (D), organism habitat (E) and organism occurrence region (F) in Standartox.}
+    \label{fig:stx-parameters}
+\end{figure}
 
-list how many freshwater species, marine etc.
-how many regions
-how many chemical groups
+\subsection{Aggregation}
+In order to reduce multiple ecotoxicity values for individual chemical-species combinations and to solve the associated problem of which value to choose for CRA, Standartox introduces aggregation methods that enable the derivation of single ecotoxicity values.
+This aggregation step is important because multiple test results exhibit highly variability in chemical risk assessment (CRA) due to several factors, such as differences in the susceptibility of different taxa to a chemical (Fig.
+\ref{fig:stx-variability}A), test results variability of one and the same taxon and chemical due to different test durations (Fig.
+\ref{fig:stx-variability}B) as well as variability caused by unexplainable factors, due to different laboratories, physiological differences between organism populations or other unaccounted variation (Fig.
+\ref{fig:stx-variability}C).
+Standartox calculates the minimum, the maximum and the geometric mean as aggregates from the filtered data set for each chemical.
+The geometric mean is preferred in comparison to the arithmetic mean, since it is considerably less influenced by outliers and is suitable for skewed data.
+Also, the geometric mean is preferable over the median, since the median completely ignores the influence of large or small values, making it unreliable for small data sets.
+In the course of the aggregation process, outliers that exceed 1.5 times the interquartile range are flagged to caution Standartox users.
+However, since the geometric mean is not strongly influenced by such entries, they are not removed.
 
-%% Aggregation
-
-In order to deal with multiple test result occurrences of individual chemicals, Standartox enables the aggregation of the data by calculating the minimum, the maximum, or the geometric mean.
-
-filtered data set + flags (for outliers)
-aggregated data set
-meta
-
-%%%% TODO: plot geometric mean above all data for a specific chemical (like on poster)
-
-%% Accuracy assessment
-
-%% General info and summary at the end
-Standartox provides quick access.
-
-Standartox is designed to foster chemical risk assessment (CRA).
-Due to an increased amount of available ecotoxicological test data, it becomes fundamental to provide and distribute such information in adequate formats, both easily accessible for humans and easily processable for machines.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%% CONTINUE HERE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Standartox is a tool that allows to filter and aggregate ecotoxicological test results to derive single exposure endpoints for individual chemicals.
-It can either be accessed through the APP (Figure \ref{fig:app}) or the API \ref{tab:api-endpoints} together with the R-package standartox (Code example \ref{listing:standartox-example}).
-
-%% TODO Add units part here
-
-\pagebreak
-
-\input{article/tables/app-parameters.tex}
-\pagebreak
-
-\input{article/tables/api-endpoints.tex}
-\pagebreak
-
-In order to locate ecotoxicological test data, researchers can rely on individual test results reported in publications, on compiled and published data sets \citep{malaj_organic_2014, morrissey_neonicotinoid_2015} or on data bases such as the Pesticide Property Data Base (PPDB) \citep{lewis_international_2016}.
-Each approach has its limitations though.
-Collecting toxicity information from individual publications is laborious and using already compiled data from publications or published data bases is often limited to specific groups of chemicals, such as pesticides, biocides etc.
-Additionally it is not always clear whether presented effect values are aggregated or single specific tests have been selected.
-These resources can provide out of date data and rarely provide sophisticated means to directly process the data.
-
-Standartox overcomes these paucities by making use of the quarterly updated EPA ECOTOX data base, the largest publicly available collection of ecotoxicological test information.
-In doing so, Standartox allows for a quick access to ecotoxicological test data, not being confined to specific chemical classes.
-Furthermore, Standartox constitutes not only a one-time compiled data set but rather a scalable method that can steadily incorporate new ecotoxicological test data in an automated manner, allowing for constant improvement of toxicity estimations over time.
-Data retrieval is streamlined and can be easily done via the APP or the R package.
-On top of filtering the data, Standartox provides means to aggregate results from multiple ecotoxicological tests to retrieve single exposure endpoints for specific chemical, taxon and test parameter combinations.
-Test results for each chemical in the filtered data set are thereby aggregated by calculating the minimum, the geometric mean or the maximum.
-Standartox returns then a filtered and an aggregated data set.
-Besides, Standartox allows users to access older versions of the data base as well and is created as open source software and therefore a fully reproducible method.
+\begin{figure}[h!]
+    \includegraphics[width=1.0\textwidth]{article/figures/results_variability.png}
+    \caption{Variability of test results (EC\textsubscript{50}) in Standartox due to (A) the susceptibility of different taxa (i.e.
+\textit{Xenopus laevis}, \textit{Pseudokirchneriella subcapitata}, \textit{Oncorhynchus mykiss}, \textit{Lemna minor}) to the chemical atrazine tested for 96h, due to (B) different durations of tests of zinc sulfate on \textit{Daphnia magna} and due to (C) un-accountable variability of cupric sulfate tested on \textit{Pimephales promelas} for 96h.
+Yellow dots depict Standartox approximations and black and gray dots show un-aggregated values for each group.}
+    \label{fig:stx-variability}
+\end{figure}
 
 \subsection{Accuracy Assessment}
-Other initiatives harmonizing the large number of ecotoxicological test results were also developed recently.
-They partly aim for overlapping goals, yet have limitations or objectives that differentiate them from Standartox.
+In order to validate Standartox results we compared the outputs to values from other databases.
 Since 2006, the PPDB provides validated ecotoxicological information on pesticides \citep{lewis_international_2016}.
-The aggregation of Standartox matches PPDB (n=3601) values that have been manually quality checked in the 91 \% within 1 order of magnitude.
-Considering only Standartox values for which at least five tests were aggregated the value increases to 92.7 \%.
-Similarly, we compared Standartox to values from the ChemProp \citep{ufzdepartmentofecologicalchemistry_chemprop_2019} software, which calculates LC\textsubscript{50} values by the Read Across method \citep{schuurmann_quantitative_2011} and found that (n=179) 95 \% lie within a factor 10 difference.
-However, the difference is not necessarily an indication of lower quality of these data in Standartox but may also be attributed to the inclusion of consideration of a wider range of test data and conditions as well as variance in the Read Across models, respectively.
+91 \% of the Standartox geometric mean aggregations lie within one order of magnitude to the manually quality checked PPDB (n=3601) values.
+Considering only Standartox geometric means for which at least five tests were aggregated, the value increases to 92.7 \%.
+Similarly, we compared Standartox to ecotoxicity values for \textit{D.
+magna} from the ChemProp \citep{ufzdepartmentofecologicalchemistry_chemprop_2019} software, which estimates LC\textsubscript{50} values via quantitative structure-activity relationship (QSAR) models \citep{schuurmann_quantitative_2011} and found that (n=179) 95 \% lie within a factor of below one order of magnitude in comparison to the Standartox aggregations.
+However, the difference is not necessarily an indication of lower quality of these data in Standartox but may also be attributed to the inclusion of consideration of a wider range of test data and conditions as well as variance in the QSAR models, respectively (Figure \ref{fig:standartox_ppdb_diff}).
 
-In order to asses the validity of the Standartox aggregation approach, we compared Standartox results with effect values from the PPDB (n=3601) for the same chemicals (Figure \ref{fig:standartox_ppdb_diff}).
-
-  Admittedly a few Standartox values diverge by a large margin.
-
-\begin{figure}
+\begin{figure}[h!]
     \includegraphics[width=1\linewidth]{article/figures/gg_ppdb_stan_compare_continous.png}
-    \caption{Comparison between Standartox and PPDB values. The black line marks exact coherence and red lines mark a divergence of a factor of 10. Compared organism groups are color coded.}
+    \caption{Comparison between Standartox and PPDB (A) and read across (B) values.
+The black lines mark exact coherence and red lines mark a divergence of a factor of 10.
+Compared organism groups are color coded.}
     \label{fig:standartox_ppdb_diff}
@@
 
+\subsection{Perspectives}
+Other initiatives harmonizing the large number of ecotoxicological test results were also developed recently.
+They partly aim for overlapping goals, yet have limitations or objectives that differentiate them from Standartox.
 The Network of reference laboratories, research centres and related organisations for monitoring of emerging environmental substances (NORMAN) focuses on assembling river basin specific pollutants from studies and databases \citep{von_der_ohe_new_2011}.
-The EnviroTox data base \href{https://envirotoxdatabase.org/} which also uses, amongst others the EPA ECOTOX data base as an input was recently published \citep{healthandenvironmentalsciencesinstitutehesi_envirotox_2019, connors_creation_2019}.
-In comparison to Standartox, EnviroTox focuses only on aquatic organisms and uses an qualitative algorithm to exclude unreliable test results, such as pre-filtering, for example the data to fish, amphibian invertebrate and algae taxa, to test durations of above 24 hours and aquatic taxa only.
+The EnviroTox database \href{https://envirotoxdatabase.org/} which also uses, amongst others the ECOTOX database as an input was recently published \citep{healthandenvironmentalsciencesinstitutehesi_envirotox_2019, connors_creation_2019}.
+In comparison to Standartox, EnviroTox focuses only on aquatic organisms and uses a qualitative algorithm to derive single ecotoxicity values.
+For example, they restrict the ecotoxicity data to fish, amphibian, invertebrate and algae taxa only and test durations to be at least above 24 hours.
+In contrast, Standartox doesn't refine to specific test durations, but leaves it up to the user to decide on such parameters.
 Besides, they add additional information on toxicity endpoints, such as acute or chronic classifications and mode of action assignments.
-In contrast, Standartox doesn't refine to specific test durations, but leaves it up to the user to decide on such parameters.
-The EnvirTox data base also allows for an aggregation of test results to derive single toxicity values for individual taxa whereas Standartox performs this aggregation for individual chemicals.
+An action that can also be allocated to the user, since there are several different ways to classify such parameters, especially those mentioned, guaranteeing user flexibility.
+The EnvirTox database also allows for an aggregation of test results to derive single toxicity values for individual taxa whereas Standartox performs this aggregation for individual chemicals taxa combinations.
 Comptox, is a web tool published by the EPA which, similar to Standartox allows for filtering test results, the retrieval of additional chemical information and predicting toxicity properties, such as 48 h \textit{Daphnia magna} LC\textsubscript{50} values.
@@
 \textit{Daphnia magna}), and the tool lacks the possibility to retrieve predicted values in an automated way \citep{williams_comptox_2017}.
-\citet{petschick_modeling_2019} modeled risk threshold level equivalents for aquatic organisms by using ecotoxicological effect data from the EPA ECOTOX data base to support regulation.
-In summary, none of the above mentioned approaches aim for a holistic standardized aggregation method of exposure endpoints for individual chemicals and also lack the possibility for automated and scriptable user requests from common high level programming languages, such as R or Python.
-Along with newly created ecotoxicological data bases here, methods of how to efficiently store ecotoxicological data are also proposed as can be seen in the MAGIC Knowledge Base \citep{bub_graphing_2019}.
-This recent increase of efforts to compile ecotoxicological test data bases clearly emphasizes the need for holistic and automated analyses of large-scale ecotoxicological data.
-
-\citet{vandenberg_modeling_2019} used ECOTOX for trait-based sensitivity models.
-
-\subsection{Outlook}
-Standartox is not singular effort, but an automated tool that steadily incorporates newly published ecotoxicity data.
-The ECOTOX database added on average 5103\rpm1317 new test results every four months in the five last years.
-This provides a constantly growing source of ecotoxicity data which is in turn directly incorporated into Standartox which steadily improves its toxicity estimates.
-Since Standartox toxicity estimates will change and to guarantee reproducibility we allow users to access older versions of Standartox.
-Yet, we are certain, that Standartox aggregates of well tested chemical-organism combinations to be stable when new test results are added, because the geometric mean is not strongly influenced by outliers.
-This can be seen from the, above mentioned comparison to the PPDB, which shows strong evidence for coherence already when aggregating only five values.
-At most, minima and maxima could potentially shift greatly.
-Therefore, we flag outliers in the filtered data set to point users to possible extreme values.
-A future analysis on the smallest number of tests needed to return trustworthy results, could shed more light on this aspect.
-Furthermore, toxicity test results are certainly influenced by individual tests parameters, such as pH, temperature or conductivity amongst others \citep{rosenkrantz_influence_2013, li_temperature_2011}.
-Such information is not included in the aggregation performed by Standartox, since the ECOTOX database only provides sparse records on those.
+\citet{petschick_modeling_2019} modeled risk threshold level equivalents for aquatic organisms by using ecotoxicological effect data from the ECOTOX database to support regulation.
+Along with newly created ecotoxicological databases, methods of how to efficiently store ecotoxicological data are also proposed as can be seen in the MAGIC Knowledge Base \citep{bub_graphing_2019}.
+Likewise, ideas to create new predictive frameworks in ecotoxicology, incorporating chemical mode of actions and species traits \citep{vandenberg_modeling_2019} together with the recent increase in efforts to compile ecotoxicological test databases clearly emphasizes the need for holistic and automated analyses of large-scale ecotoxicological data.
+In summary, none of the above mentioned approaches aim for a holistic standardized aggregation method of exposure endpoints for individual chemicals and also lack the possibility for automated and scriptable user requests from common high level programming languages, such as R or Python \citep{pythoncoreteam_python_2020} (Table \ref{tab:database-differences}).
+Another important feature to improve toxicity estimates is to include a large number of test parameters, such as pH, temperature or conductivity amongst others, since toxicity test results are highly influenced by those \citep{rosenkrantz_influence_2013, li_temperature_2011}.
+Such information is not included in the aggregation performed by Standartox, since the ECOTOX database only provides sparse records on test parameters.
 Most frequently provided test parameters (and their percentage) are medium temperature (77\%), pH:56\%, hardness:27\%, dissolved oxygen:18\%, Alkalinity:15\% and salinity:9\%.
@@
 
-We chose the geometric mean in comparison to the arithmetic mean, since it is considerably less influenced by outliers and is suitable for skewed data.
-Also, the geometric mean is preferable over the median, since the median completely ignores the influence of large or small values, making it unreliable for small data sets <CITATION GEOMETRIC MEAN>.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TODO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\iffalse
-- rename subsections as such:
-
-\subsection*{Accuracy Assessment}
-
-\subsection*{Perspective}
-
-- can also be incorporated in EPA Dashboard
-- Identify also underrepresented regions concerning organisms.
-1) \citep{staveley_challenge_2016} describe what is necessary for future ecotoxicology.
-2) Compare Standartox with predictions here: https://comptox.epa.gov/dashboard/predictions/index
-- \citep{gadaleta_new_2018} what does this software exactly do an how to access it?
-\fi
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% OLD %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\iffalse
-SSD try to extrapolate effects of chemicals towards organisms from species to a community level by fitting a statistical distribution to toxicity test results on organisms of different trophic levels in order to derive concentrations at which a specific fraction of organisms are effected (often at 5\% - HC\textsubscript{5} \citep{posthuma_species_2002}.
-Figure \ref{fig:ssd-isoproturon} shows an example of a SSD and the derived HC\textsubscript{5} value (red dassed line) created for the herbicide Isoproturon.
-TUs are another indicator of effects of chemicals on organisms.
-They are calculated by dividing a measured concentration (c) by an EC\textsubscript{50} value (cf.
-Equation \ref{eq:tu}).
-
-\begin{equation}
-    TU_i = \frac{c_i}{\ecfifty}
-    \label{eq:tu}
-\end{equation}
-
-\begin{figure}[h!]
-    \includegraphics[width=1\linewidth]{article/figures/ssd2_boot.png}
-    \caption{SSD plot showing the susceptibility of organisms towards the herbicide Isoproturon.
-The red line represents the fit and the red dashed line marks the HC\textsubscript{5} value, which is a common measure in CRA.
-The legend denotes common organism groups used in ecotoxicology.}
-    \label{fig:ssd-isoproturon}
-\end{figure}
-
-Besides that, two cleaning steps can also be chosen.
-Firstly the application allows to exclude test results exhibiting concentrations that are higher than the actual water solubility of the respective chemical at \ang{20} C.
-Secondly the user can choose to exclude outliers (cf.
-Table \ref{tab:scripts-app}).
-The outliers are selected as values that exceed lower (0.25) and the upper (0.75) quartile by 1.5 times the inter-quartile range.
-
-\subsection*{Identify chemicals to be tested}
-There is a clear need in ecotoxicology for harmonised approaches to collect ecotoxicological data as other data base compilation attempts recently published show.
-Hitchock et al.
-\citet{hitchcock_improving_2018} argue for an increased awareness concerning newly published ecotoxicological test results to facilitate meta analyses across studies.
-Important statistical parameters, such as mean estimates, variances and sample sizes.
-Likewise, state variables, including medium temperature, pH, mineral contents or organism age and sex should be reported (if not in the text, in an online repository).
-Standartox doesn't include such information mainly because the EPA ECOTOX reports only a few of these variables.
-A future research aim could be to use machine learning techniques to retrieve this information from the referenced articles.
-Above all, recent studies showed that, in defiance of many regulation efforts \citep{schafer_future_2019} chemicals can still cause strong adverse effects on organisms and in turn on the environment.
-Neonicotinoids, a group of insecticides was for example related to declines in insectivorous birds \citep{hallmann_declines_2014} and fish yields due to cascading food web effects \citep{yamamuro_neonicotinoids_2019}.
-Standartox provides means to synthesise ecotoxicological test data, thereby identify research gaps.
-
-\begin{figure}
-    \includegraphics[width=1\linewidth]{article/figures/heatmap_tests_n.png}
-    \caption{30 most tested chemicals in EPA ECOTOX.}
-    \label{fig:standartox_ppdb_diff}
-\end{figure}
-
-\fi
+\begin{sidewaystable}
+% \begin{table}
+    \input{article/tables/database-differences.tex}
+    \caption{Differences of databases that provide ecotoxicological data.}
+    \label{tab:database-differences}
+% \end{table}
+\end{sidewaystable}

### 2020-01-27 article/parts/03_methods.tex
Updates from Overleaf

@@
-%%The Methods should include detailed text describing any steps or procedures 
-%%used in producing the data, including full descriptions of the experimental 
-%%design, data acquisition assays, and any computational processing (e.g. 
-%%normalization, image feature extraction). Related methods should be grouped 
-%%under corresponding subheadings where possible, and methods should be described 
-%%in enough detail to allow other researchers to interpret and repeat, if required, 
-%%the full study. Specific data outputs should be explicitly referenced via data 
-%%citation (see Data Records and Data Citations, below). Authors should cite 
-%%previous descriptions of the methods under use, but ideally the method 
-%%descriptions should be complete enough for others to understand and reproduce 
-%%the methods and processing steps without referring to associated publications. 
-%%There is no limit to the length of the Methods section.
 \section{Methods}
-An automated processing pipeline downloads the quarterly released EPA ECOTOX database, performs several preparation steps on it and exports a final Standartox data set.
+An automated processing pipeline downloads the quarterly released ECOTOX database, performs several preparation steps on it and exports a final Standartox data set.
 This data set is accessible via a web application (APP) and an application programming interface (API).
+An API provides the means for machine communication between a host and a client and thus allowing scriptable data queries.
 To facilitate the API access, the R \citep{rcoreteam_language_2017} package \textit{standartox} is built.
 
-\subsection{Processing pipe line}
-Standartox downloads the quarterly released EPA ECOTOX data base whenever a new version is published and builds it into a local PostgreSQL data base \citep{szocs_build_2019}.
+\subsection{Processing}
+Standartox downloads the quarterly released EPA ECOTOX database whenever a new version is published and builds it into a local PostgreSQL database \citep{szocs_build_2019}.
 Subsequently Structured Query Language (SQL) functions for further processing the data are developed.
 In addition lookup tables (Supplement \ref{sup:conv-concentration}, \ref{sup:conv-duration}) that enable duration and concentration unit conversions of the data are created.
-A meta table providing information such as the release version of the EPA ECOTOX data base is then added.
+A meta table providing information such as the release version of the ECOTOX database is then added.
 In the next step, Chemical Abstracts Service Numbers (CAS) and International Chemical Identifier (InChI) as well as taxonomic names are used to query additional information from publicly available databases on chemicals and organisms, respectively.
-This includes the Compendium of Pesticide Common Names \citep{wood_compendium_2019}, the Chemical Entities of Biological Interest (ChEBI) database \citep{hastings_chebi_2016}, the Chemical Identifier Resolver (CIR) service \citep{nationalinstitutesofhealthnih_chemical_2019}, the PHYSPROP database, the Pubchem data base \citep{kim_pubchem_2016}, Eurostat \citep{europeancommission_eurostat_2019} and Wikidata \citep{vrandecic_wikidata_2014} for chemicals and the World Register of Marine Species (WoRMS) \citep{wormseditorialboard_world_2018} and the Global Biodiversity Information Facility (GBIF) \citep{_gbif_2019} (Table \ref{tab:data-base-additional}) for habitat and spatial distribution of organisms.
+This includes the Compendium of Pesticide Common Names \citep{wood_compendium_2019}, the Chemical Entities of Biological Interest (ChEBI) database \citep{hastings_chebi_2016}, the Chemical Identifier Resolver (CIR) service \citep{nationalinstitutesofhealthnih_chemical_2019}, the PHYSPROP database, the Pubchem database \citep{kim_pubchem_2016}, Eurostat \citep{europeancommission_eurostat_2019} and Wikidata \citep{vrandecic_wikidata_2014} for chemicals and the World Register of Marine Species (WoRMS) \citep{wormseditorialboard_world_2018} and the Global Biodiversity Information Facility (GBIF) \citep{_gbif_2019} (Table \ref{tab:data-base-additional}) for habitat and spatial distribution of organisms.
 This information is added to Standartox to allow filtering for specific classes of chemicals as well as spatial distribution (i.e.
@@
 freshwater) of individual taxa.
-
-\input{article/tables/data_base_tables.tex}
-
+Taxa that where not identified to at least genus level are excluded.
 Finally, the Standartox data set is compiled, which includes the harmonisation of data, e.g.
 through conversion of units related to the concentrations and test duration.
-Out of the unique 1229 concentration units in the EPA ECOTOX data base, Standartox retains only those (n = X) that are unambiguously convertible to one of the following units: $\mu$g/l, g/m2, ppb, mg/kg, \% and flags, the remaining as \textit{other}, resulting in 873473 out of 952625 test results.
+Out of the unique 1229 concentration units in the ECOTOX database, Standartox retains only those (n = X) that are unambiguously convertible to one of the following units: $\mu$g/l, g/m2, ppb, mg/kg, \% and flags, the remaining as \textit{other}, resulting in 873473 out of 952625 test results.
 Furthermore, the units are cleaned, for example through removing additional information in the field such as \textit{food}, \textit{soil}, \textit{ai} that are also coded in other variables and hinder the processing of units.
@@
 mortality, growth) are created.
+In order to detect possible outliers among the test results, we flag values that exceed 1.5 times the interquartile range (IQR) within groups with identical test parameters (e.g.
+chemical, taxon, duraiton).
 Along with the compiled Standartox data set, a catalog, listing all distinct entries and value ranges, for categorical and continuous variables, respectively is created.
@@
 
-\subsection{Web application (APP) and Application Programming Interface (API)}
-The web application (APP) and the Application Programming Interface (API) load the compressed serialized Standartox data into memory and allow a client to interact with it.
-The client can then call the functions stx\_filter() and stx\_aggregate() that filter and aggregate the data according to specific parameters (Table \ref{tab:app-parameters}).
-
-\subsubsection{Web application (APP)}
-The interactive APP was built in R using the shiny web application framework, which runs with the help of a shiny server \citep{chang_shiny_2018} on \url{standartox.uni-landau.de}.
-It is a set of R functions that allow a user to filter, aggregate and download the data.
-
-\subsubsection{Application programming interface (API)}
-The API is built by using the R package plumber \citep{trestletechnologyllc_plumber_2018}, which allows for the creation of RESTful APIs from R.
+\subsection{Access}
+The APP and the API load the compressed serialized Standartox data into memory and allow a client to interact with it.
+The client thereby calls the functions stx\_filter() and stx\_aggregate() that filter and aggregate the data according to specific parameters (Table \ref{tab:app-parameters}).
+The interactive APP is built in R using the shiny web application framework, which runs with the help of a shiny server \citep{chang_shiny_2018} on \url{standartox.uni-landau.de}.
+The API is built by using the R package \textit{plumber} \citep{trestletechnologyllc_plumber_2018}, which allows for the creation of Representational state transfer (REST) APIs from R.
+REST is a software standard that defines web service communication rules.
 The API is reachable via the URL \url{http://139.14.20.252} and port 8000.
@@
 The \textit{/catalog} API-endpoint returns a JavaScript Object Notation (JSON) file containing a catalog of possible filter parameters to choose from.
-The \textit{/filter} returns the filtered Standartox table as a compressed serialized binary file created by the R fst package \citep{klik_fst_2019}, to reduce size and allow for fast user queries.
+The \textit{/filter} returns the filtered Standartox table as a compressed serialized binary file created by the R \textit{fst} package \citep{klik_fst_2019}, to reduce size and allow for fast user queries.
 The \textit{/aggregate} API-endpoint returns an R function allowing the client to aggregate the filtered data.
 Lastly, the \textit{/meta} API-endpoint returns a JSON file with meta information, such as the timestamp of the request and the used Standartox version.
-The API is designed to be used with the R-package standartox and therefore uses serialization methods specific to R (rds() from the R package base and fst() from the package fst).
-
-\subsubsection{R-package}
-The R-package standartox accesses the API through two functions \textit{stx\_catalog()} and \textit{stx\_query()}.
-The former accesses the \textit{/catalog} API-endpoint and returns a R list object of possible filter parameters (Table \ref{tab:app-parameters}) that can be used in \textit{stx\_query()}.
-The latter returns a R list of three tables (i.e.
-R data.frames) containing the filtered data set, the aggregated data set and a table with the meta information retrieved from the API endpoints \textit{/filter} and \textit{/meta}.
+The API is designed to be used with the R-package \textit{standartox} and therefore uses serialization methods specific to R (rds() from the R package base and fst() from the package fst).
+In order to facilitate the API usage the R-package \textit{standartox} is created.
 
@@
 To guarantee appropriate unit conversion and harmonisation, we compare for each of the 1229 distinct concentration and for each of the 129 duration units the automatically converted units to one manually calculated one.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TODO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Where to put?
-(See Table \ref{tab:app-parameters})
-
-\iffalse
-- Endpoint: For the last parameter here, only be exclusive.
-
-TODO: How does this fit?
-\begin{table}
-    \input{article/tables/data_refinement_counts.tex}
-    \caption{Reduction of data in the compilation process for Standartox}
-    \label{tab:data-refinement}
-\end{table}
-
-\fi
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% OLD %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\iffalse
-\textbf{OLD: Standartox aggregates the test results according to chosen filters in a two step process.
-Firstly the filtered test results are aggregated by the CAS number, the chosen taxon and the selected test duration.
-Secondly, the returned data is then aggregated by the CAS number.
-The former can't be influenced by the user and calculates either the minimum or the median depending on the amount of results to aggregate (n <= 2: minimum, if n > 2: median).
-Thereof the second step calculates the minimum, the maximum, the median, the geometric mean, or the arithmetic mean as an aggregate.}
-\fi

### 2020-01-27 article/parts/04_user_notes.tex
Updates from Overleaf

@@
 \section{User notes}
+By accessing the APP users can filter and download the resulting data sets as a comma-separated values (csv) file.
+Users of the R-package can directly load the data in R.
+The R-package provides the two functions \textit{stx\_catalog()} and \textit{stx\_query()}.
+The first queries a catalog of possible Standartox parameters into a R list object.
+The latter allows users to set the Standartox filter parameters and to fetch the actual data.
+It returns a R list of three tables (i.e.
+R data.frames) containing the filtered data set, the aggregated data set and a table with the meta information retrieved from the API endpoints \textit{/filter} and \textit{/meta}.
+A short R-code example is given below \ref{lis:standartox-example} and a detailed description on the usage of the R package is provided on its Github page: \url{https://github.com/andschar/standartox}.
 
-Using the APP, the user can download the resulting data sets as a comma-separated values (csv) file whereas by using the API, the data can directly be loaded in R.
-To provide an easy overview of the data set, percentage values for each parameter entry are shown in the APP in the selection menu as well as by using the R package through querying the \textit{/catalog} API-endpoint.
+\input{article/code/standartox-example.tex}
 
-\pagebreak
-
-\begin{figure}
+\begin{figure}[h]
     \includestandalone[scale=0.4]{article/tikz/standartox_organigram}

### 2020-01-27 article/parts/05_conclusion.tex
Updates from Overleaf

@@
 
-Standartox is a tool that does this in a generalizing way and puts its focus on the aggregation of toxicity data, by this adding a piece to the puzzle of modern ecotoxicological analyses.
+Due to the steady incorporation of new ecotoxicity data, ecotoxicity values produced by Standartox will be subject to change and most certainly produce slightly different results in its aggregations in the future.
+However, we regard this as an advantage, since other published works that aim in similar directions often constitute a singular effort or require manual work for each update.
+Standartox, in contrast automates these processes and as already described would provide means to access older versions anyhow, to assure reproducibility.
+In comparison to rule-based approaches, such as the PPDB for the derivation of single ecotoxicity values, Standartox has the advantage to not be prone to the subjectivity, the exhaustiveness and the incompleteness of a human induced set of rules.
+Above all, Standartox provides quick access through its design to be queried via the R language.
+In general, Standartox is designed to foster chemical risk assessment (CRA).
+Due to an increased amount of available ecotoxicological test data, it becomes fundamental to provide and distribute ecotoxicity information in adequate formats, both easily accessible for humans and easily processable for machines.
+Standartox is a tool that does this in a generalizing way and puts its focus on the aggregation of toxicity data, thereby adding a piece to the puzzle of modern ecotoxicological analyses.

### 2020-01-27 article/parts/11_misc.tex
Updates from Overleaf

--- 2019-11-28 article/parts/02_results_discussion.tex
@@
-\section*{Results and Discussion}
-Standartox is a tool that allows to filter and aggregate ecotoxicological test results to derive single exposure endpoints for individual chemicals.
-It can either be accessed through the APP (Figure \ref{fig:app}) or the API \ref{tab:api-endpoints} together with the R-package standartox (Code example \ref{listing:standartox-example}).
-Both access pathways allow for the same filters and aggregation methods (Table \ref{tab:app-parameters}) to be applied and access the same data set.
-Using the APP, the user can download the resulting data sets as a comma-separated values (csv) file whereas by using the API, the data can directly be loaded in R.
-Different parameters in order to filter the test data, such as compound-specific parameters, including the CAS number, the concentration type (e.g.
-Active Ingredient, Formulation) or the chemical class (e.g.
-insecticides, metals) as well as taxon-specific parameters, including common taxonomic groups (e.g.
-Daphniidae, algae), the habitat (e.g.
-freshwater) or the regional occurrence (e.g.
-Europe, Asia) of the organisms can be chosen.
-Furthermore, users can refine the results to specific test durations (in hours), effect groups (e.g.
-mortality, population, growth) and endpoints (EC\textsubscript{50}, LOEC or NOEC).
-As the method of aggregation users can choose the minimum, the maximum, or the geometric mean.
-To provide an easy overview of the data set, percentage values for each parameter entry are shown in the APP in the selection menu as well as by using the R package through querying the \textit{/catalog} API-endpoint.
+%%%%%%%%%%%%% OLD SUMMARY TEXT %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+In order to locate ecotoxicological test data, researchers can rely on individual test results reported in publications, on compiled and published data sets \citep{malaj_organic_2014, morrissey_neonicotinoid_2015} or on data bases such as the Pesticide Property Data Base (PPDB) \citep{lewis_international_2016}.
+Each approach has its limitations though.
+Collecting toxicity information from individual publications is laborious and using already compiled data from publications or published data bases is often limited to specific groups of chemicals, such as pesticides, biocides etc.
+Additionally it is not always clear whether presented effect values are aggregated or single specific tests have been selected.
+These resources can provide out of date data and rarely provide sophisticated means to directly process the data.
+Standartox overcomes these paucities by making use of the quarterly updated EPA ECOTOX data base, the largest publicly available collection of ecotoxicological test information.
+In doing so, Standartox allows for a quick access to ecotoxicological test data, not being confined to specific chemical classes.
+Furthermore, Standartox constitutes not only a one-time compiled data set but rather a scalable method that can steadily incorporate new ecotoxicological test data in an automated manner, allowing for constant improvement of toxicity estimations over time.
+Data retrieval is streamlined and can be easily done via the APP or the R package.
+On top of filtering the data, Standartox provides means to aggregate results from multiple ecotoxicological tests to retrieve single exposure endpoints for specific chemical, taxon and test parameter combinations.
+Test results for each chemical in the filtered data set are thereby aggregated by calculating the minimum, the geometric mean or the maximum.
+Standartox returns then a filtered and an aggregated data set.
+Besides, Standartox allows users to access older versions of the data base as well and is created as open source software and therefore a fully reproducible method.
+%%% END
 
-%% Add units part here
-
-\pagebreak
-\begin{figure}
-    \caption{Web application (APP) Standartox allowing to set filter parameters and aggregation methods.
-Data can be downloaded as comma separated value (csv) files.
-An interactive overview plot and table allow for browsing of the data.}
-    \label{fig:app}
-    \includegraphics[width=0.75\linewidth]{article/figures/screenshot_standartox_app.png}
-\end{figure}
-
-\pagebreak
-
-\input{article/code/standartox_example.tex}
-\pagebreak
-
-\begin{figure}
-    \includestandalone[scale=0.4]{article/tikz/standartox_organigram}
-    \caption{Organigram of Standartox.}
-    \label{fig:stx-organigram}
-\end{figure}
+%%%%%%%%%%%%%%%% remove this? %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \pagebreak
@@
 
-Standartox is designed to foster chemical risk assessment (CRA).
-Due to an increased amount of available ecotoxicological test data, it becomes fundamental to provide and distribute such information in adequate formats, both easily accessible for humans and easily processable for machines.
-Currently, CRA often relies on test results of a few, well tested standard organisms, such as \textit{Daphnia magna}, \textit{Pimephales promelas}, \textit{Selenastrum capricornutum} to name a few, although researchers have conducted experiments on a much greater variety of organisms, observable in the 12,906 taxa from 952,634 test results collected in the EPA ECOTOX data base.
-In order to locate ecotoxicological test data, researchers can rely on individual test results reported in publications, on compiled and published data sets \citep{malaj_organic_2014, morrissey_neonicotinoid_2015} or on data bases such as the Pesticide Property Data Base (PPDB) \citep{lewis_international_2016}.
-Each approach has its limitations though.
-Collecting toxicity information from individual publications is laborious and using already compiled data from publications or published data bases is often limited to specific groups of chemicals, such as pesticides, biozides etc.
-Just as well, these resources run at risk of providing out of date data sooner or later or rarely provide sophisticated means to directly process the data.
-Standartox overcomes these paucities by making use of the quarterly updated EPA ECOTOX data base, the largest publicly available collection of ecotoxicological test information.
-In doing so, Standartox allows for a quick access to ecotoxicological test data, not being confined to specific chemical classes.
-Furthermore, Standartox constitutes not only a one-time compiled data set but rather a scalable method that can steadily incorporate new ecotoxicological test data in an automated manner, allowing for constant improvement of toxicity estimations over time.
-Data retrieval is streamlined and can be easily done via the APP or the R package.
-On top of filtering the data, Standartox provides means to aggregate results from multiple ecotoxicological tests to retrieve single exposure endpoints for specific chemical, taxon and test parameter combinations.
-Test results for each chemical in the filtered data set are thereby aggregated by calculating the minimum, the geometric mean or the maximum.
-Standartox returns then a filtered and an aggregated data set.
-Besides, Standartox allows users to access older versions of the data base as well and is created as open source software and therefore a fully reproducible method.
-
-\subsection*{Other data bases}
-Other initiatives tidying the large number of ecotoxicological test results were also developed recently.
-They partly aim for overlapping goals, yet have limitations or objectives that differentiate them from Standartox.
-Since 2006, the PPDB provides validated ecotoxicological information on pesticides \citep{lewis_international_2016}.
-In order to asses the validity of the Standartox aggregation approach, we compared Standartox results with effect values from the PPDB (n=3601) for the same chemicals (Figure \ref{fig:standartox_ppdb_diff}).
-This showed that 91.3\% of the Standartox values diverge by less than a factor of ten from the PPDB values.
-Considering only Standartox values for which at least five or ten tests were aggregated the value increases to 91.7\% and 93.4\%, respectively.
-Admittedly a few Standartox values diverge by a large margin.
-
-\begin{figure}
-    \includegraphics[width=1\linewidth]{article/figures/gg_ppdb_stan_compare_continous.png}
-    \caption{Comparison between Standartox and PPDB values. The black line marks exact coherence and red lines mark a divergence of a factor of 10. Compared organism groups are color coded.}
-    \label{fig:standartox_ppdb_diff}
-\end{figure}
-
-The Network of reference laboratories, research centres and related organisations for monitoring of emerging environmental substances (NORMAN) focuses on assembling river basin specific pollutants \citep{von_der_ohe_new_2011}.
-The EnviroTox data base \href{https://envirotoxdatabase.org/} which also uses, amongst others the EPA ECOTOX data base as an input was recently published \citep{healthandenvironmentalsciencesinstitutehesi_envirotox_2019, connors_creation_2019}.
-In comparison to Standartox, EnviroTox focuses only on aquatic organisms and uses an qualitative algorithm to exclude unreliable test results, such as pre-filtering, for example the data to fish, amphibian invertebrate and algae taxa, to test durations of above 24 hours and aquatic taxa only.
-Besides, they add additional information on toxicity endpoints, such as acute or chronic classifications and mode of action assignments.
-In contrast, Standartox doesn't refine to specific test durations, but leaves it up to the user to decide on such parameters.
-The EnvirTox data base also allows for an aggregation of test results to derive single toxicity values for individual taxa whereas Standartox performs this aggregation for individual chemicals.
-Comptox, is a web tool published by the EPA which, similar to Standartox allows for filtering test results, the retrieval of additional chemical information and predicting toxicity properties, such as 48 h \textit{Daphnia magna} LC\textsubscript{50} values.
-However, toxicity predictions are limited to some standard organisms (e.g.
-\textit{Daphnia magna}), and the tool lacks the possibility to retrieve predicted values in an automated way \citep{williams_comptox_2017}.
-Petschick et al.
-(in submission) modeled risk threshold level equivalents for aquatic organisms by using ecotoxicological effect data from the EPA ECOTOX data base to support regulation.
-In summary, none of the above mentioned approaches aim for a holistic standardized aggregation method of exposure endpoints for individual chemicals and also lack the possibility for automated and scriptable user requests from common high level programming languages, such as R or Python.
-Along with newly created ecotoxicological data bases here, methods of how to efficiently store ecotoxicological data are also proposed as can be seen in the MAGIC Knowledge Base \citep{bub_graphing_2019}.
-This recent increase of efforts to compile ecotoxicological test data bases clearly emphasizes the need for holistic and automated analyses of large-scale ecotoxicological data.
-
-\subsection*{Limitations}
-Since the used data in Standartox merely represents a sample of the actual toxicity of chemicals on organisms, it has to be stated that the returned Standartox values are subject to change.
-New studies will be published and constantly incorporated into EPA ECOTOX data base and therefore into Standartox, improving the toxicity estimates.
-However, we expect values of well tested chemical-organism combinations to be relatively stable.
-A future analysis on the smallest number of tests needed to return trustworthy results, could shed more light on this aspect.
-Furthermore, toxicity test results are certainly influenced by individual tests parameters, such as pH, temperature or conductivity amongst others \citep{rosenkrantz_influence_2013, li_temperature_2011}.
-Such information is not included in the aggregation performed by Standartox, since the EPA ECOTOX data base only provides sparse records on those.
-Most frequently provided test parameters (and their percentage) are medium temperature (77\%), pH:56\%, hardness:27\%, dissolved oxygen:18\%, Alkalinity:15\% and salinity:9\%.
-Others are provided for less than 5\% of the tests.
-A text-mining approach, iterating through the individual publications could potentially increase this number.
-However, we argue that using the geometric mean as an aggregate allows for an adequate estimation of the toxicity of a chemical towards an organism group.
-We chose the geometric mean in comparison to the arithmetic mean, since it is considerably less influenced by outliers and is suitable for skewed data.
-Also, the geometric mean is preferable over the median, since the median completely ignores the influence of large or small values, making it unreliable for small data sets <CITATION GEOMETRIC MEAN>.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TODO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%% OLD method %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \iffalse
-- can also be incorporated in EPA Dashboard
-- Identify also underrepresented regions concerning organisms.
-1) \citep{staveley_challenge_2016} describe what is necessary for future ecotoxicology.
-2) Compare Standartox with predictions here: https://comptox.epa.gov/dashboard/predictions/index
-- \citep{gadaleta_new_2018} what does this software exactly do an how to access it?
+\textbf{OLD: Standartox aggregates the test results according to chosen filters in a two step process.
+Firstly the filtered test results are aggregated by the CAS number, the chosen taxon and the selected test duration.
+Secondly, the returned data is then aggregated by the CAS number.
+The former can't be influenced by the user and calculates either the minimum or the median depending on the amount of results to aggregate (n <= 2: minimum, if n > 2: median).
+Thereof the second step calculates the minimum, the maximum, the median, the geometric mean, or the arithmetic mean as an aggregate.}
 \fi
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% OLD %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\iffalse
-SSD try to extrapolate effects of chemicals towards organisms from species to a community level by fitting a statistical distribution to toxicity test results on organisms of different trophic levels in order to derive concentrations at which a specific fraction of organisms are effected (often at 5\% - HC\textsubscript{5} \citep{posthuma_species_2002}.
-Figure \ref{fig:ssd-isoproturon} shows an example of a SSD and the derived HC\textsubscript{5} value (red dassed line) created for the herbicide Isoproturon.
-TUs are another indicator of effects of chemicals on organisms.
-They are calculated by dividing a measured concentration (c) by an EC\textsubscript{50} value (cf.
-Equation \ref{eq:tu}).
-
-\begin{equation}
-    TU_i = \frac{c_i}{\ecfifty}
-    \label{eq:tu}
-\end{equation}
-
-\begin{figure}[h!]
-    \includegraphics[width=1\linewidth]{article/figures/ssd2_boot.png}
-    \caption{SSD plot showing the susceptibility of organisms towards the herbicide Isoproturon.
-The red line represents the fit and the red dashed line marks the HC\textsubscript{5} value, which is a common measure in CRA.
-The legend denotes common organism groups used in ecotoxicology.}
-    \label{fig:ssd-isoproturon}
-\end{figure}
-
-Besides that, two cleaning steps can also be chosen.
-Firstly the application allows to exclude test results exhibiting concentrations that are higher than the actual water solubility of the respective chemical at \ang{20} C.
-Secondly the user can choose to exclude outliers (cf.
-Table \ref{tab:scripts-app}).
-The outliers are selected as values that exceed lower (0.25) and the upper (0.75) quartile by 1.5 times the inter-quartile range.
-
-\subsection*{Identify chemicals to be tested}
-There is a clear need in ecotoxicology for harmonised approaches to collect ecotoxicological data as other data base compilation attempts recently published show.
-Hitchock et al.
-\citet{hitchcock_improving_2018} argue for an increased awareness concerning newly published ecotoxicological test results to facilitate meta analyses across studies.
-Important statistical parameters, such as mean estimates, variances and sample sizes.
-Likewise, state variables, including medium temperature, pH, mineral contents or organism age and sex should be reported (if not in the text, in an online repository).
-Standartox doesn't include such information mainly because the EPA ECOTOX reports only a few of these variables.
-A future research aim could be to use machine learning techniques to retrieve this information from the referenced articles.
-Above all, recent studies showed that, in defiance of many regulation efforts \citep{schafer_future_2019} chemicals can still cause strong adverse effects on organisms and in turn on the environment.
-Neonicotinoids, a group of insecticides was for example related to declines in insectivorous birds \citep{hallmann_declines_2014} and fish yields due to cascading food web effects \citep{yamamuro_neonicotinoids_2019}.
-Standartox provides means to synthesise ecotoxicological test data, thereby identify research gaps.
-
-\begin{figure}
-    \includegraphics[width=1\linewidth]{article/figures/heatmap_tests_n.png}
-    \caption{30 most tested chemicals in EPA ECOTOX.}
-    \label{fig:standartox_ppdb_diff}
-\end{figure}
-
-\fi

### 2020-01-27 article/tables/database-differences.tex
Updates from Overleaf

%% table that compares different ecotoxicological databases

%%%% CONTINUE HERE

\begin{tabular}{|m{3cm}|m{3cm}|m{1cm}|m{2cm}|m{1cm}|m{2cm}|l|}
\hline
Database & Publisher & Filter & Aggregation, Selection & Scriptable & Access & website \\
\hline
Comptox & Environmental Protection Agency & yes & no & no & web & https://comptox.epa.gov/dashboard \\
\hline
Ecotox & Environmental Protection Agency & yes & no & no & web & https://webetox.uba.de/webETOX/index.do \\
\hline
EnviroTox & Health and Environmental Sciences Institute & yes & per species & no & web & https://envirotoxdatabase.org \\
\hline
Etox & Umweltbundesamt & yes & no & no & web & https://webetox.uba.de/webETOX/index.do \\
\hline
Pesticide Property Data Base (PPDB) & University of Hertfordshire & fixed values & manual selection & no & web, file & https://sitem.herts.ac.uk/aeru/ppdb/index.htm \\
\hline
\textbf{Standartox} & \textbf{this article} & \textbf{yes} & \textbf{per chemical and organism} & \textbf{yes} & \textbf{web, API} & \textbf{http://standartox.uni-landau.de} \\
\hline
\end{tabular}

### 2020-03-01 article/parts/01_summary.tex
Updates from Overleaf

@@
 atmospheric emission and deposition or discharge through wastewater) \citep{schwarzenbach_challenge_2006}.
-In the environment, chemicals can adversely affect populations and communities and in turn related ecosystem functions \citep{rhind_anthropogenic_2009, clements_community_2009, hallmann_declines_2014, barracaracciolo_pharmaceuticals_2015, johnston_review_2015, hallmann_more_2017, sanchez-bayo_worldwide_2019}.
-Ultimately, this may compromise natures contribution to human well-being, for example the ecosystem services clean drinking and irrigation water as well as food production \citep{peters_review_2013, vandersluijs_neonicotinoids_2013, barracaracciolo_pharmaceuticals_2015, yamamuro_neonicotinoids_2019}.
+In the environment, chemicals can adversely affect populations and communities and in turn related ecosystem functions \citep{schafer_thresholds_2012, malaj_organic_2014, hallmann_declines_2014, barracaracciolo_pharmaceuticals_2015, johnston_review_2015}.
+Ultimately, this may compromise natures contribution to human well-being, for example the ecosystem services clean drinking and irrigation water as well as food production \citep{peters_review_2013, vandersluijs_neonicotinoids_2013, yamamuro_neonicotinoids_2019}.
 Pollution with man-made chemicals has been identified as one of three major environmental problems for which research gaps hamper the derivation of planetary boundaries, i.e.
 thresholds beyond which irreversible state shifts may occur \citep{steffen_anthropocene_2007, steffen_planetary_2015}.
-\citet{bernhardt_synthetic_2017} argue that the knowledge gap how chemicals affect populations, communities and in turn ecosystem functions and services, may impede the accomplishment of the Sustainable Development Goals of the United Nations.
-Even highly regulated chemical compounds such as pesticides have been shown to cause strong adverse effects on organisms, such as birds \citep{hallmann_declines_2014} or fish \citep{yamamuro_neonicotinoids_2019}, questioning the current regulation efforts \citep{schafer_future_2019}.
-To evaluate the risks from chemicals for ecosystems data on their toxicity, which is typically produced in standardised ecotoxicological laboratory tests is required.
-For example, \citet{morrissey_neonicotinoid_2015} used ecotoxicological test results from 49 insect and crustacean species to evaluate the effect of neonicotinoids, an insecticide class on the aquatic ecosystem.
-Also, \citet{malaj_organic_2014} compiled experimental toxicity test results for 223 various chemical compounds to assess the health of freshwater ecosystems in Europe.
-Similarly, permissible environmental concentrations are often derived from these test data, typically by a combination with safety factors to account for uncertainties, thereby mostly relying on a few, well tested standard organisms, such as the water flea \textit{Daphnia magna}, the fathead minnow \textit{Pimephales promelas}, the microalga \textit{Raphidocelis subcapitata} to name a few.
-Researchers have conducted experiments on a much greater variety of organisms though.
-Given that data sets with large numbers of analysed chemicals are becoming increasingly available, ecotoxicity data are pivotal for regulatory risk assessment (RRA) and science.
-To date, only few initiatives exist that aim to create a public resource of harmonized and quality controlled ecotoxicological data, such as the United States Environmental Protection Agency ECOTOXicology knowledgebase (ECOTOX) (ca.
-1,000,000 test results, ca.
-13,000 taxa, ca.
-12,000 chemicals) \citep{elonen_ecotoxicology_2018}, the German Environmental Agency's Information System Ecotoxicology and Environmental Quality Targets (ETOX) \citep{umweltbundesamt_etox_2019}, the Pesticide Property Data Base (PPDB) (ca.
+\citet{bernhardt_synthetic_2017} argue that the knowledge gap how chemicals affect populations, communities and in turn ecosystem functions and services, may impede the accomplishment of the Sustainable Development Goals \citep{rosa_transforming_2017} of the United Nations.
+Even highly regulated chemical compounds such as pesticides have been shown to cause strong adverse effects on non-target organisms, such as birds \citep{hallmann_declines_2014} aquatic insects \citep{beketov_pesticides_2013} or fish \citep{yamamuro_neonicotinoids_2019}, questioning the current regulation efforts \citep{schafer_future_2019}.
+
+To evaluate the risks from chemicals for ecosystems, data on their toxicity is required, which is typically produced in standardised ecotoxicological laboratory tests.
+For example, \citet{morrissey_neonicotinoid_2015} used ecotoxicological test results from 49 insects and crustaceans to evaluate the effect of neonicotinoid insecticides in the aquatic ecosystem.
+Also, \citet{malaj_organic_2014} compiled experimental toxicity test results for 223 chemicals to assess the health of freshwater ecosystems in Europe.
+Similarly, permissible environmental concentrations are often derived from these test data, typically by a combination with safety factors to account for uncertainties.
+The test data mainly relate to a few, well tested standard organisms, such as the brown rat \textit{Rattus norvegicus}, the fathead minnow \textit{Pimephales promelas} and the microalga \textit{Raphidocelis subcapitata}.
+Nevertheless, a much greater variety of organisms has been used in ecotoxicological experiments.
+To date, only few initiatives exist that aim to create a public resource of ecotoxicological data, such as the United States Environmental Protection Agency ECOTOXicology knowledgebase (ECOTOX) (ca.
+1,000,000 test results, 13,000 taxa, 12,000 chemicals) \citep{elonen_ecotoxicology_2018}, the German Environmental Agency's Information System Ecotoxicology and Environmental Quality Targets (ETOX) \citep{umweltbundesamt_etox_2019}, the Pesticide Properties Data Base (PPDB) (ca.
 2000 chemicals) \citep{lewis_international_2016} or the Envirotox database \citep{healthandenvironmentalsciencesinstitutehesi_envirotox_2019, connors_creation_2019}.
-The former two compile all available test results into a database.
-However, for many chemicals, multiple ecotoxicity values are available for the same test organism, which can vary strongly and the lack of associated quality information and different units hamper reproducible science.
-The PPDB database in contrast provides single ecotoxicity values only for pesticides and a few selected test organisms, thereby ignoring vast amounts of data for other species and chemicals.
-Envirotox puts its focus on aquatic organisms.
+The former two compile all available results from experiments into a database.
+However, for many chemicals, multiple ecotoxicity values are available for the same test organism, which can vary strongly and the lack of associated quality information and heterogeneous units hamper reproducible science.
+The PPDB database, in contrast, provides single ecotoxicity values only for pesticides and a few selected test organisms, thereby covering only a minor fraction of the vast amount of ecotoxicological data.
+The Envirotox is limited to aquatic organisms.
 Moreover, data analyses often require links to additional data resources, for example to append additional chemical and species information (e.g.
@@
 refining a search for ecotoxicity data on organisms occurring in specific habitats or regions of the world.
-Above that, Standartox aggregates ecotoxicological test results in a standardized way, by calculating the minimum, the geometric mean and the maximum of the result for each chemical and the associated, user-defined test parameters.
+Above that, Standartox aggregates ecotoxicological test results in a standardized way, by calculating the minimum, the geometric mean and the maximum of the results for each chemical and the associated, user-defined test parameters.
 This relaxes the commonly occurring problems of the variability of multiple ecotoxicological test results and which values to choose for chemical risk assessment (CRA).
-Hence, Standartox provides the basis for reproducible science and combines information from different sources to simplify the derivation of more advanced risk indicators such as Species Sensitivity Distributions (SSD) and Toxic Units (TU), which represent two prominent concepts to assess effects on organisms in ecotoxicology \citep{posthuma_species_2002, kefford_definition_2011, schafer_effects_2011}.
-Besides aggregating ecotoxicological test results, Standartox provides a concise overview of the status of tested chemicals, allowing the identification of potential research gaps.
+Hence, Standartox provides the basis for reproducible science and combines information from different sources to simplify the derivation of risk indicators such as Species Sensitivity Distributions (SSD) and Toxic Units (TU), which represent two prominent concepts to assess effects on organisms in ecotoxicology \citep{posthuma_species_2002, kefford_definition_2011, schafer_effects_2011}.
+Besides aggregating ecotoxicological test results, Standartox provides a concise overview of the status of tested chemicals, allowing the identification of potential knowledge gaps.
 Moreover, Standartox could help in reducing the millions of animals used for toxicity testing each year \citep{hartung_chemical_2009}.

### 2020-03-01 article/parts/02_data_description.tex
Updates from Overleaf

@@
 Standartox constitutes a collection of 600,000 quality checked ecotoxicological test results comprising roughly 8000 distinct chemicals and 8000 taxa that allows for the retrieval of single toxicity equivalents.
-It is build on the ECOTOX database \citep{usepa_ecotox_2019} and the data is processed in order to clean and harmonise entries to retrieve comparable toxicity endpoints.
-Subsequently, filter and aggregation methods are provided to allow for the retrieval of single toxicity equivalents for specific test parameter combinations.
-The underlying ECOTOX database is released quarterly and provides on average \input{article/values/publish_rate.txt} (2014 - 2019) new test results with each release.
+It is build on the ECOTOX database \citep{usepa_ecotox_2019} and the data are processed to clean and harmonise entries to retrieve comparable toxicity endpoints.
+Subsequently, filter and aggregation methods are provided to allow for the retrieval of single toxicity equivalents for specific experimental conditions.
+The underlying ECOTOX database is released quarterly and provides on average \input{article/values/publish_rate.txt} (2014 - 2019) new toxicity entries with each release.
 Standartox is build directly after each new ECOTOX release in an automated way and therefore incorporates new data steadily.
-To guarantee reproducibility, users can choose which Standartox version to access.
-Users can reach Standartox via an web app and an R-package (standartox) that accesses an application programming interface (API) providing the possibility of scriptable requests.
+To guarantee reproducibility, Standartox comes with a version control and users can choose which Standartox version to access.
+Users can reach Standartox via a web app and an R-package (standartox) that accesses an application programming interface (API) providing the possibility of scriptable requests.
 
 \subsection{Filters}
-The data can be restricted to three endpoint groups XX50, LOEX and NOEX, conflating commonly used endpoints such as half maximal effective concentrations (EC\textsubscript{50}), no observed effect levels/concentrations (NOEC/L), as well as lowest observed effect levels/concentrations (LOEC/L) respectively (Table \ref{tab:endpoints-conflate}) (Fig.
+The data can be restricted to three endpoint groups XX50, LOEX and NOEX, conflating commonly used endpoints such as half maximal effective/lethal concentrations (E/LC\textsubscript{50}), no observed effect levels/concentrations (NOEC/L), as well as lowest observed effect levels/concentrations (LOEC/L), respectively (Table \ref{tab:endpoints-conflate}) (Fig.
 \ref{fig:stx-parameters}A).
-Standartox allows the ecotoxicity data to be filtered to 23 distinct effect groups (e.g.
+Standartox allows the ecotoxicity data to be filtered by 23 distinct effect groups (e.g.
 mortality, population, growth) (Fig.
@@
 formulation, active ingredient).
-In addition to these test-specific parameters Standartox data entries can be filtered according to chemical-specific parameters such as, the CAS number and five chemical classes (e.g.
-pesticides, metals, drugs), commonly used in ecotoxicology (Fig.
+In addition to these test-specific parameters, Standartox data entries can be filtered by chemical-specific parameters such as the CAS number and chemical classes (e.g.
+pesticides, metals, drugs) (Fig.
 \ref{fig:stx-parameters}C).
@@
 freshwater, marine, terrestrial) (Fig.
-\ref{fig:stx-parameters}E), the occurrence regions (e.g.
+\ref{fig:stx-parameters}E), their occurrence regions (e.g.
 Europe, South America) (Fig.
-\ref{fig:stx-parameters}F) of the organisms as well as different taxonomic levels.
+\ref{fig:stx-parameters}F) as well as different taxonomic levels.
 Additionally users can refine the results to specific test durations (in hours).
@@
     \includegraphics[width=1.0\textwidth]{article/figures/standartox_parameters.png}
-    \caption{Share of 10 most occurring entries for the parameters endpoint (A), effect (B), chemical class (C), taxonomic order (D), organism habitat (E) and organism occurrence region (F) in Standartox.}
+    \caption{Share of 10 most frequent entries for the parameters endpoint (A), effect group (B), chemical class (C), taxonomic order (D), organism habitat (E) and organism occurrence region (F) in Standartox.}
     \label{fig:stx-parameters}
@@
 \subsection{Aggregation}
-In order to reduce multiple ecotoxicity values for individual chemical-species combinations and to solve the associated problem of which value to choose for CRA, Standartox introduces aggregation methods that enable the derivation of single ecotoxicity values.
-This aggregation step is important because multiple test results exhibit highly variability in chemical risk assessment (CRA) due to several factors, such as differences in the susceptibility of different taxa to a chemical (Fig.
-\ref{fig:stx-variability}A), test results variability of one and the same taxon and chemical due to different test durations (Fig.
-\ref{fig:stx-variability}B) as well as variability caused by unexplainable factors, due to different laboratories, physiological differences between organism populations or other unaccounted variation (Fig.
-\ref{fig:stx-variability}C).
-Standartox calculates the minimum, the maximum and the geometric mean as aggregates from the filtered data set for each chemical.
-The geometric mean is preferred in comparison to the arithmetic mean, since it is considerably less influenced by outliers and is suitable for skewed data.
-Also, the geometric mean is preferable over the median, since the median completely ignores the influence of large or small values, making it unreliable for small data sets.
+Typically, species exhibit a differential sensitivity towards chemicals (Fig.
+\ref{fig:stx-variability}A).
+Moreover, multiple ecotoxicity values are available for individual species-chemical combinations and these can exhibit high variability due to several factors such as durations of ecotoxicity tests (Fig.
+\ref{fig:stx-variability}B), experimental conditions and physiology and genetic strain of the test individuals or populations.
+To aggregate multiple ecotoxicity values into a single value on the desired taxonomic level (e.g.
+for an individual species, across species of a genus or family), and chemical grouping (e.g.
+across all pesticides), Standartox provides several aggregation methods including the minimum, the maximum and the geometric mean as aggregates from the filtered data set for each chemical.
+The geometric mean is preferred in comparison to the arithmetic mean, because it is considerably less influenced by outliers and is suitable for skewed data.
+Also, the geometric mean is preferable over the median, because the median completely ignores the influence of large or small values, making it unreliable for small data sets.
 In the course of the aggregation process, outliers that exceed 1.5 times the interquartile range are flagged to caution Standartox users.
-However, since the geometric mean is not strongly influenced by such entries, they are not removed.
+However, since the geometric mean is relatively robust against outliers, they are not removed automatically.
+Overall, aggregation with Standartox allows to reduce the between-study variability when toxicity data are aggregated.
 
@@
     \includegraphics[width=1.0\textwidth]{article/figures/results_variability.png}
-    \caption{Variability of test results (EC\textsubscript{50}) in Standartox due to (A) the susceptibility of different taxa (i.e.
-\textit{Xenopus laevis}, \textit{Pseudokirchneriella subcapitata}, \textit{Oncorhynchus mykiss}, \textit{Lemna minor}) to the chemical atrazine tested for 96h, due to (B) different durations of tests of zinc sulfate on \textit{Daphnia magna} and due to (C) un-accountable variability of cupric sulfate tested on \textit{Pimephales promelas} for 96h.
-Yellow dots depict Standartox approximations and black and gray dots show un-aggregated values for each group.}
+    \caption{Violin plots of the variability of test results (EC\textsubscript{50}) in Standartox associated with (A) different species (i.e.
+\textit{Xenopus laevis} - Amphibia, \textit{Raphidocelis subcapitata} - Algae, \textit{Oncorhynchus mykiss} - fish, \textit{Lemna minor} - invertebrate) to the chemical atrazine tested for 96 h, (B) diﬀerent durations of toxicity tests with zinc sulfate and \textit{Daphnia magna} and (C) unaccountable variability of cupric sulfate tested on \textit{Pimephales promelas} for 96 h.
+Purple dots depict Standartox approximations and black dots show un-aggregated values for each group.
+To facilitate readability, data points are randomly scattered along a hypothetical y-axis and are greyed out within the violins.}
     \label{fig:stx-variability}
@@
 \subsection{Accuracy Assessment}
-In order to validate Standartox results we compared the outputs to values from other databases.
-Since 2006, the PPDB provides validated ecotoxicological information on pesticides \citep{lewis_international_2016}.
-91 \% of the Standartox geometric mean aggregations lie within one order of magnitude to the manually quality checked PPDB (n=3601) values.
-Considering only Standartox geometric means for which at least five tests were aggregated, the value increases to 92.7 \%.
+To validate Standartox results we compared the aggregated results using the geometric mean to values from other databases.
+The PPDB provides ecotoxicity data on a few selected species from chemical risk assessment that have been manually quality controlled through expert judgement \citep{lewis_international_2016}.
+The vast majority of aggregated values (91 \%) of Standartox lie within one order of magnitude of the corresponding PPDB values (n = 3601).
+This would increase to 92.7 \%, when restricting the comparison to Standartox values where data from at least five experiments were available.
 Similarly, we compared Standartox to ecotoxicity values for \textit{D.
-magna} from the ChemProp \citep{ufzdepartmentofecologicalchemistry_chemprop_2019} software, which estimates LC\textsubscript{50} values via quantitative structure-activity relationship (QSAR) models \citep{schuurmann_quantitative_2011} and found that (n=179) 95 \% lie within a factor of below one order of magnitude in comparison to the Standartox aggregations.
-However, the difference is not necessarily an indication of lower quality of these data in Standartox but may also be attributed to the inclusion of consideration of a wider range of test data and conditions as well as variance in the QSAR models, respectively (Figure \ref{fig:standartox_ppdb_diff}).
+magna} from the ChemProp \citep{ufzdepartmentofecologicalchemistry_chemprop_2019} software, which estimates LC\textsubscript{50} values via quantitative structure-activity relationship (QSAR) models \citep{schuurmann_quantitative_2011} We found that  95 \% of Standartox values lie within one order of magnitude of Chemprop (n = 179).
+However, the diﬀerence is not necessarily an indication of lower quality of Standartox estimates but may also reflect the wider range of experimental conditions for which data are available in the database underlying Standartox as well as inaccurate predictions for QSAR models, respectively (Figure \ref{fig:standartox_ppdb_diff}).
 
@@
     \includegraphics[width=1\linewidth]{article/figures/gg_ppdb_stan_compare_continous.png}
-    \caption{Comparison between Standartox and PPDB (A) and read across (B) values.
-The black lines mark exact coherence and red lines mark a divergence of a factor of 10.
-Compared organism groups are color coded.}
+    \caption{Comparison between Standartox and (A) PPDB and (B) ChemProp values. The black lines indicate identity and red lines mark a divergence of a factor of 10. Compared species are color coded.}
     \label{fig:standartox_ppdb_diff}
@@
 \subsection{Perspectives}
-Other initiatives harmonizing the large number of ecotoxicological test results were also developed recently.
-They partly aim for overlapping goals, yet have limitations or objectives that differentiate them from Standartox.
+Novel predictive frameworks incorporating chemical mode of action and species traits emphasize the need for holistic and automated analyses of large-scale ecotoxicological data \citep{malaj_evolutionary_2016, vandenberg_modeling_2019}.
+Indeed, the increasing amount of data from ecotoxicological tests and experiments that is becoming available has elicited several initiatives to  harmonize these data.
+These initiatives partly aim for overlapping goals, yet have limitations or objectives that discriminate them from Standartox.
 The Network of reference laboratories, research centres and related organisations for monitoring of emerging environmental substances (NORMAN) focuses on assembling river basin specific pollutants from studies and databases \citep{von_der_ohe_new_2011}.
-The EnviroTox database \href{https://envirotoxdatabase.org/} which also uses, amongst others the ECOTOX database as an input was recently published \citep{healthandenvironmentalsciencesinstitutehesi_envirotox_2019, connors_creation_2019}.
-In comparison to Standartox, EnviroTox focuses only on aquatic organisms and uses a qualitative algorithm to derive single ecotoxicity values.
-For example, they restrict the ecotoxicity data to fish, amphibian, invertebrate and algae taxa only and test durations to be at least above 24 hours.
-In contrast, Standartox doesn't refine to specific test durations, but leaves it up to the user to decide on such parameters.
-Besides, they add additional information on toxicity endpoints, such as acute or chronic classifications and mode of action assignments.
+The EnviroTox database \href{https://envirotoxdatabase.org/} which also uses, amongst others the ECOTOX database as an input has recently been published \citep{healthandenvironmentalsciencesinstitutehesi_envirotox_2019, connors_creation_2019}.
+In contrast to Standartox, EnviroTox  is restricted to selected aquatic organisms (i.e.
+fish, amphibians, invertebrates and algae) and experimental durations (at least 24 h) and uses a qualitative algorithm to derive single ecotoxicity values.
+Besides, EnviroTox provide additional information on toxicity endpoints, such as acute or chronic classifications and mode of action assignments.
 An action that can also be allocated to the user, since there are several different ways to classify such parameters, especially those mentioned, guaranteeing user flexibility.
-The EnvirTox database also allows for an aggregation of test results to derive single toxicity values for individual taxa whereas Standartox performs this aggregation for individual chemicals taxa combinations.
+The EnviroTox database allows for an aggregation into single toxicity values for individual taxa,  whereas Standartox performs this aggregation for individual taxa-chemical combinations.
+However, the Standartox results for individual taxa-chemical combinations could easily be aggregated across chemicals in a second step based on chemical parameters.
 Comptox, is a web tool published by the EPA which, similar to Standartox allows for filtering test results, the retrieval of additional chemical information and predicting toxicity properties, such as 48 h \textit{Daphnia magna} LC\textsubscript{50} values.
-However, toxicity predictions are limited to some standard organisms (e.g.
-\textit{Daphnia magna}), and the tool lacks the possibility to retrieve predicted values in an automated way \citep{williams_comptox_2017}.
-\citet{petschick_modeling_2019} modeled risk threshold level equivalents for aquatic organisms by using ecotoxicological effect data from the ECOTOX database to support regulation.
-Along with newly created ecotoxicological databases, methods of how to efficiently store ecotoxicological data are also proposed as can be seen in the MAGIC Knowledge Base \citep{bub_graphing_2019}.
-Likewise, ideas to create new predictive frameworks in ecotoxicology, incorporating chemical mode of actions and species traits \citep{vandenberg_modeling_2019} together with the recent increase in efforts to compile ecotoxicological test databases clearly emphasizes the need for holistic and automated analyses of large-scale ecotoxicological data.
-In summary, none of the above mentioned approaches aim for a holistic standardized aggregation method of exposure endpoints for individual chemicals and also lack the possibility for automated and scriptable user requests from common high level programming languages, such as R or Python \citep{pythoncoreteam_python_2020} (Table \ref{tab:database-differences}).
-Another important feature to improve toxicity estimates is to include a large number of test parameters, such as pH, temperature or conductivity amongst others, since toxicity test results are highly influenced by those \citep{rosenkrantz_influence_2013, li_temperature_2011}.
-Such information is not included in the aggregation performed by Standartox, since the ECOTOX database only provides sparse records on test parameters.
-Most frequently provided test parameters (and their percentage) are medium temperature (77\%), pH:56\%, hardness:27\%, dissolved oxygen:18\%, Alkalinity:15\% and salinity:9\%.
-Others are provided for less than 5\% of the tests.
-A text-mining approach, iterating through the individual publications could potentially increase this number.
+However, toxicity predictions are limited to standard organisms (e.g.
+\textit{Daphnia magna}), and the tool lacks the possibility for automatic retrieval of predicted \citep{williams_comptox_2017}.
 
-\begin{sidewaystable}
-% \begin{table}
+In summary, none of the above mentioned initiatives aim for a holistic and automated standardized aggregation method of exposure endpoints for individual chemicals.
+In addition, they lack the possibility to access the databases through common high level programming languages, such as R.
+As outlined above, toxicity estimates from different studies can be highly variable due to a wide range of experimental conditions such as pH, temperature or conductivity \citep{rosenkrantz_influence_2013, li_temperature_2011}.
+Integrating these factors into the aggregated estimates would improve toxicity estimates .
+However, the current implementation of Standartox does not consider these factors, because the ECOTOX database only provides sparse records on experimental conditions.
+The most frequently provided experimental conditions are temperature (77 \%), pH (56 \%), hardness (27 \%), dissolved oxygen (18 \%), Alkalinity (15 \%) and salinity (9 \%), and all others for less than 5 \%.
+A text-mining approach, where a literature reference is associated with  ecotoxicity raw data, iterating through the individual publications could potentially increase this number.
+
 \input{article/tables/database-differences.tex}
-    \caption{Differences of databases that provide ecotoxicological data.}
-    \label{tab:database-differences}
-% \end{table}
-\end{sidewaystable}

### 2020-03-01 article/parts/03_methods.tex
Updates from Overleaf

@@
 This data set is accessible via a web application (APP) and an application programming interface (API).
-An API provides the means for machine communication between a host and a client and thus allowing scriptable data queries.
+An API provides the means for machine communication between a host and a client and thus allows scriptable data queries.
 To facilitate the API access, the R \citep{rcoreteam_language_2017} package \textit{standartox} is built.
@@
 \subsection{Processing}
-Standartox downloads the quarterly released EPA ECOTOX database whenever a new version is published and builds it into a local PostgreSQL database \citep{szocs_build_2019}.
-Subsequently Structured Query Language (SQL) functions for further processing the data are developed.
+Standartox downloads the quarterly released EPA ECOTOX database and builds it into a local PostgreSQL database \citep{szocs_build_2019}.
+Subsequently, Structured Query Language (SQL) functions for further processing the data are implemented.
 In addition lookup tables (Supplement \ref{sup:conv-concentration}, \ref{sup:conv-duration}) that enable duration and concentration unit conversions of the data are created.
@@
 Units such as per day (mg/kg/day) are generally excluded.
-Test endpoints are restricted to three groups, named NOEX, LOEX and XX50.
+Experimental responses are restricted to three groups, namely NOEX, LOEX and XX50.
 The former two represent no-observed-adverse-effect and lowest-observed-adverse-effect concentrations or levels, respectively.
-The latter includes various sub-groups of the half maximal effective concentration, where half of the tested individuals show an effect (e.g.
+The latter includes various sub-groups of the half maximal effective concentration, where half of the individuals in an experiment display an effect (e.g.
 EC50, LC50, LD50), which is a common measure in ecotoxicology \citep{malaj_organic_2014}.
-Other endpoints, such as Bioconcentration factors (BCF), non-half maximal effective concentrations (e.g.
+Other responses, such as Bioconcentration factors (BCF), non-half maximal effective concentrations (e.g.
 IC10, EC25, LD99) or maximum acceptable toxicant (MATC) concentrations are removed.
@@
 mortality, growth) are created.
-In order to detect possible outliers among the test results, we flag values that exceed 1.5 times the interquartile range (IQR) within groups with identical test parameters (e.g.
-chemical, taxon, duraiton).
+To detect possible outliers among the test results, we flag values that exceed 1.5 times the interquartile range (IQR) within groups with identical test parameters (e.g.
+chemical, taxon, duration).
 Along with the compiled Standartox data set, a catalog, listing all distinct entries and value ranges, for categorical and continuous variables, respectively is created.
@@
 The API is reachable via the URL \url{http://139.14.20.252} and port 8000.
-Four API-endpoints (\textit{/catalog}, \textit{/filter}, \textit{/aggregate} and \textit{/meta}) can be queried from the client side.
+Four API-endpoints (\textit{/catalog}, \textit{/filter}, \textit{/aggregate} and \textit{/meta}) can be queried by users.
 The \textit{/catalog} API-endpoint returns a JavaScript Object Notation (JSON) file containing a catalog of possible filter parameters to choose from.
@@
 Lastly, the \textit{/meta} API-endpoint returns a JSON file with meta information, such as the timestamp of the request and the used Standartox version.
-The API is designed to be used with the R-package \textit{standartox} and therefore uses serialization methods specific to R (rds() from the R package base and fst() from the package fst).
+The API is designed to be used with the R-package \textit{standartox} and therefore uses serialization methods specific to R (rds() from the R package \pkg{base} and fst() from the package fst).
 In order to facilitate the API usage the R-package \textit{standartox} is created.
@@
 \subsection{Technical Validity}
-To guarantee appropriate unit conversion and harmonisation, we compare for each of the 1229 distinct concentration and for each of the 129 duration units the automatically converted units to one manually calculated one.
+To guarantee appropriate unit conversion and harmonisation, we compared for each of the 1229 distinct concentration and for each of the 129 duration units the automatically converted units to one manually calculated one.

### 2020-03-01 article/parts/05_conclusion.tex
Updates from Overleaf

@@
 
-Due to the steady incorporation of new ecotoxicity data, ecotoxicity values produced by Standartox will be subject to change and most certainly produce slightly different results in its aggregations in the future.
+Due to the steady incorporation of new ecotoxicity data, its values produced by Standartox will be subject to change and most certainly produce slightly different results in its aggregations in the future.
 However, we regard this as an advantage, since other published works that aim in similar directions often constitute a singular effort or require manual work for each update.

### 2020-03-01 article/parts/10_supplement.tex
Updates from Overleaf

--- 2019-11-05 article/parts/06_supplement.tex
@@
 \section*{Supplement}
-\subsection*{R-package list} %% TODO put in appropriate list
+
+%% Endpoints conflate
+\begin{table}
+    \label{tab:endpoints-conflate}
+    \catption{Table of Standartox endpoints and their corresponding Ecotox endpoints and a description.}
+    \input{article/tables/endpoint-conflate.tex}
+\end{table}
+
+%% Additional databases
+\begin{table}[ht]
+    \label{tab:data-base-additional}
+    \caption{Table of additionally queried publicly available data bases and their URLs.}
+    \input{article/tables/data-base-tables.tex}
+\end{table}
+
+%% App parameters
+(See Table \ref{tab:app-parameters})
+
+\pagebreak
+\section{CHECK THIS}
+
+It can either be accessed through the APP (Figure \ref{fig:app}) or the API \ref{tab:api-endpoints} together with the R-package standartox (Code example \ref{listing:standartox-example}).
+
+\subsection*{R-package list}
 \label{list:r-packages}
-%% TODO how to refernce this?
 \input{article/tables/r_package_list.tex}
-\pagebreak
 
-\subsection*{Data base tables}
-\input{article/tables/data_base_tables.tex}
-
-\subsection*{Unit conversion}
-
+\subsection*{Unit conversions}
+ TODO
 \begin{table}
@@
     \caption{Concentration unit conversion}
-    \label{tab:conv-concentration}
+    \label{sup:conv-concentration}
 \end{table}
@@
     \caption{Duration unit conversion}
-    \label{tab:conv-duration}
+    \label{sup:conv-duration}
 \end{table}
 
-\subsection*{Application parameters}
-\input{article/tables/app_parameters.tex}
+%\input{article/tables/chck-units-concentration.tex}
 
-\input{article/tables/chck-units-concentration.tex}
+%\input{article/tables/chck-units-duration.tex}
 
-\input{article/tables/chck-units-duration.tex}
+\subsection*{Scripts}
+\begin{figure}
+    % TODO make better graphic
+    \includegraphics[width=0.5\linewidth]{article/figures/scripts_dendrogram.png}
+    \caption{Processing pipeline}
+    \label{fig:pipeline-tree}
+\end{figure}

### 2020-03-01 article/tables/database-differences.tex
Updates from Overleaf

@@
 
-\begin{tabular}{|m{3cm}|m{3cm}|m{1cm}|m{2cm}|m{1cm}|m{2cm}|l|}
+\begin{sidewaystable}
+% \begin{table}
+\caption{Different databases that provide ecotoxicological data.
+Abbreviations: \textbf{ALL:} Most important parameters, such as chemical, Organism, Duration for filtering ecotoxicological data are incorporated.
+\textbf{WEB:} Accessible via a web application and a graphical user interface, not providing programmatic access.}
+\label{tab:database-differences}
+\begin{tabular}{|m{3cm}|m{3cm}|m{2cm}|m{2cm}|m{1cm}|l|}
 \hline
-Database & Publisher & Filter & Aggregation, Selection & Scriptable & Access & website \\
+Database & Publisher & Filter & Aggregation, Selection & Access & website \\
 \hline
-Comptox & Environmental Protection Agency & yes & no & no & web & https://comptox.epa.gov/dashboard \\
+Comptox & Environmental Protection Agency & Chemical & no & WEB, file & https://comptox.epa.gov/dashboard \\
 \hline
-Ecotox & Environmental Protection Agency & yes & no & no & web & https://webetox.uba.de/webETOX/index.do \\
+Ecotox & Environmental Protection Agency & ALL & no & WEB, file & https://webetox.uba.de/webETOX/index.do \\
 \hline
-EnviroTox & Health and Environmental Sciences Institute & yes & per species & no & web & https://envirotoxdatabase.org \\
+EnviroTox & Health and Environmental Sciences Institute & ALL & chemical, organism & WEB & https://envirotoxdatabase.org \\
 \hline
-Etox & Umweltbundesamt & yes & no & no & web & https://webetox.uba.de/webETOX/index.do \\
+Etox & Umweltbundesamt & ALL & no & WEB & https://webetox.uba.de/webETOX/index.do \\
 \hline
-Pesticide Property Data Base (PPDB) & University of Hertfordshire & fixed values & manual selection & no & web, file & https://sitem.herts.ac.uk/aeru/ppdb/index.htm \\
+Pesticide Property Data Base (PPDB) & University of Hertfordshire & fixed values & manual selection & WEB, file & https://sitem.herts.ac.uk/aeru/ppdb/index.htm \\
 \hline
-\textbf{Standartox} & \textbf{this article} & \textbf{yes} & \textbf{per chemical and organism} & \textbf{yes} & \textbf{web, API} & \textbf{http://standartox.uni-landau.de} \\
+\textbf{Standartox} & \textbf{this article} & \textbf{ALL} & \textbf{chemical, organism} & \textbf{API, WEB} & \textbf{http://standartox.uni-landau.de} \\
 \hline
 \end{tabular}
+% \end{table}
+\end{sidewaystable}

### 2020-03-01 article/code/standartox-example.tex

\begin{lstlisting}[
    caption = {Sample code to access the Standartox data base through the API with the help of the R-package. \textit{stx\_catalog()} returns a catalog of possible filter and aggregation parameters. \textit{stx\_query()} returns the Standartox object, a list of the filtered and the aggregated data as well as a meta data entry.},
    label = {lis:standartox-example}]
# install
install.packages('remotes')
remotes::install_github('andschar/standartox') # package not yet on CRAN
# retrieve catalog    
require(standartox)
catal = stx_catalog()
# retrieve data
l = stx_query(cas = '1071-83-6',
                endpoint = 'XX50',
                taxa = 'Oncorhynchus',
                duration = c(24, 120))
\end{lstlisting}

### 2020-03-02 article/parts/01_summary.tex
Updates from Overleaf

@@
 Similarly, permissible environmental concentrations are often derived from these test data, typically by a combination with safety factors to account for uncertainties.
-The test data mainly relate to a few, well tested standard organisms, such as the brown rat \textit{Rattus norvegicus}, the fathead minnow \textit{Pimephales promelas} and the microalga \textit{Raphidocelis subcapitata}.
+The test data mainly relate to a few, well tested standard organisms, such as the brown rat \textit{Rattus norvegicus}, the water flea \textit{Daphnia magna} and the microalga \textit{Raphidocelis subcapitata}.
 Nevertheless, a much greater variety of organisms has been used in ecotoxicological experiments.
+
 To date, only few initiatives exist that aim to create a public resource of ecotoxicological data, such as the United States Environmental Protection Agency ECOTOXicology knowledgebase (ECOTOX) (ca.
 1,000,000 test results, 13,000 taxa, 12,000 chemicals) \citep{elonen_ecotoxicology_2018}, the German Environmental Agency's Information System Ecotoxicology and Environmental Quality Targets (ETOX) \citep{umweltbundesamt_etox_2019}, the Pesticide Properties Data Base (PPDB) (ca.
-2000 chemicals) \citep{lewis_international_2016} or the Envirotox database \citep{healthandenvironmentalsciencesinstitutehesi_envirotox_2019, connors_creation_2019}.
+2000 pesticides) \citep{lewis_international_2016} or the Envirotox database \citep{healthandenvironmentalsciencesinstitutehesi_envirotox_2019, connors_creation_2019}.
 The former two compile all available results from experiments into a database.
@@
 chemical properties, habitat of species), which calls for more automated procedures.
-We therefore developed Standartox, a tool and database that aims to overcome the limitations of other databases by continuously incorporating the ever-growing number of test results in an automated process workflow that ultimately leads to a cleaned and harmonised ecotoxicity collection and single aggregated ecotoxicity values for a specific chemical-organism test combination.
-Standartox makes use of the publicly available and quarterly released ECOTOX database \citep{usepa_ecotox_2019} and restricts the data to three commonly used endpoints in ecotoxicology (i.e.
-XX50, LOEX \& NOEX), leading to about 600,000 ecotoxicological test results including 8000 chemicals, tested on about 8000 taxa.
-Standartox allows users to filter test results according to several parameters, e.g.
+
+We therefore developed Standartox, a tool and database that aims to overcome the limitations of other databases by continuously incorporating the ever-growing number of test results in an automated process workflow that ultimately leads to a cleaned and harmonised ecotoxicity data collection and provides methods to derive single aggregated ecotoxicity values for a specific chemical-organism test combination.
+Standartox makes use of the publicly available and quarterly updated ECOTOX database \citep{usepa_ecotox_2019} and restricts the data to commonly used endpoints in ecotoxicology, such as half maximal effective concentration (EC\textsubscript{50}) or no-observed-adverse-effect concentrations (NOEC), leading to about 600,000 ecotoxicological test results including 8000 chemicals, tested on about 8000 taxa.
+Users can filter test results according to several parameters, e.g.
 refining a search for ecotoxicity data on organisms occurring in specific habitats or regions of the world.
 Above that, Standartox aggregates ecotoxicological test results in a standardized way, by calculating the minimum, the geometric mean and the maximum of the results for each chemical and the associated, user-defined test parameters.
-This relaxes the commonly occurring problems of the variability of multiple ecotoxicological test results and which values to choose for chemical risk assessment (CRA).
+This tackles the commonly occurring problem of the variability of multiple ecotoxicological test results and which values to choose for assessing the effects of chemicals.
 Hence, Standartox provides the basis for reproducible science and combines information from different sources to simplify the derivation of risk indicators such as Species Sensitivity Distributions (SSD) and Toxic Units (TU), which represent two prominent concepts to assess effects on organisms in ecotoxicology \citep{posthuma_species_2002, kefford_definition_2011, schafer_effects_2011}.
-Besides aggregating ecotoxicological test results, Standartox provides a concise overview of the status of tested chemicals, allowing the identification of potential knowledge gaps.
-Moreover, Standartox could help in reducing the millions of animals used for toxicity testing each year \citep{hartung_chemical_2009}.
-Standartox comes with two front-ends, a web application (APP): \url{standartox.uni-landau.de} and a R \citep{rcoreteam_language_2017} package \textit{standartox}, providing convenience structures and thereby largely reducing processing time for users.
+Besides aggregating ecotoxicological test results, Standartox provides a concise overview of the tested chemicals, allowing the identification of potential knowledge gaps.
+Moreover, Standartox could help in reducing the millions of animals used for toxicity testing each year due to facilitated ecotoxicity data access \citep{hartung_chemical_2009}.
+Standartox comes with two front-ends, a web application: \url{standartox.uni-landau.de} and a R \citep{rcoreteam_language_2017} package \textit{standartox}, providing convenience structures and thereby largely reducing processing time for users.

### 2020-03-02 article/parts/02_data_description.tex

@@
 Subsequently, filter and aggregation methods are provided to allow for the retrieval of single toxicity equivalents for specific experimental conditions.
-The underlying ECOTOX database is released quarterly and provides on average \input{article/values/publish_rate.txt} (2014 - 2019) new toxicity entries with each release.
-Standartox is build directly after each new ECOTOX release in an automated way and therefore incorporates new data steadily.
-To guarantee reproducibility, Standartox comes with a version control and users can choose which Standartox version to access.
-Users can reach Standartox via a web app and an R-package (standartox) that accesses an application programming interface (API) providing the possibility of scriptable requests.
+The underlying ECOTOX database is updated quarterly and provides on average 5228 (2014 - 2019) new toxicity entries with each update.
+As soon as a new update of ECOTOX is available, Standartox is directly built in an automated way and therefore incorporates new data steadily.
+Standartox comes with a version control to guarantee reproducibility, meaning that once derived Standartox estimates can be retraced at a later stage, although new data might have updated the estimates already.
+Users can reach Standartox via a web application and an R-package (standartox) that accesses an application programming interface (API) providing the possibility of scriptable requests.
 
 \subsection{Filters}
-The data can be restricted to three endpoint groups XX50, LOEX and NOEX, conflating commonly used endpoints such as half maximal effective/lethal concentrations (E/LC\textsubscript{50}), no observed effect levels/concentrations (NOEC/L), as well as lowest observed effect levels/concentrations (LOEC/L), respectively (Table \ref{tab:endpoints-conflate}) (Fig.
-\ref{fig:stx-parameters}A).
+The data can be restricted to the three endpoint groups, namely half maximal effective/lethal concentration/dose values (e.g.
+EC50, LD50), henceforth abbreviated as XX50, lowest observed effect levels/concentrations (LOEC/L), henceforth abbreviated as LOEX and no observed effect levels/concentrations (NOEL/C), henceforth abbreviated as NOEX (Table \ref{tab:endpoints-conflate}).
 Standartox allows the ecotoxicity data to be filtered by 23 distinct effect groups (e.g.
@@
     \includegraphics[width=1.0\textwidth]{article/figures/standartox_parameters.png}
-    \caption{Share of 10 most frequent entries for the parameters endpoint (A), effect group (B), chemical class (C), taxonomic order (D), organism habitat (E) and organism occurrence region (F) in Standartox.}
+    \caption{Share of 10 most frequent entries for the parameters effect group (A), chemical role (B), chemical class (C), taxonomic order (D), organism habitat (E) and organism occurrence region (F) in Standartox.}
     \label{fig:stx-parameters}
@@
     \includegraphics[width=1.0\textwidth]{article/figures/results_variability.png}
-    \caption{Violin plots of the variability of test results (EC\textsubscript{50}) in Standartox associated with (A) different species (i.e.
-\textit{Xenopus laevis} - Amphibia, \textit{Raphidocelis subcapitata} - Algae, \textit{Oncorhynchus mykiss} - fish, \textit{Lemna minor} - invertebrate) to the chemical atrazine tested for 96 h, (B) diﬀerent durations of toxicity tests with zinc sulfate and \textit{Daphnia magna} and (C) unaccountable variability of cupric sulfate tested on \textit{Pimephales promelas} for 96 h.
-Purple dots depict Standartox approximations and black dots show un-aggregated values for each group.
+    \caption{Violin plots of the variability of test results (EC\textsubscript{50}) in Standartox because of different species (i.e.
+\textit{Xenopus laevis} - Amphibia, \textit{Raphidocelis subcapitata} - Algae, \textit{Oncorhynchus mykiss} - fish, \textit{Lemna minor} - invertebrate) to the chemical atrazine tested for 96 h (A), diﬀerent durations of toxicity tests with zinc sulfate and \textit{Daphnia magna} (B) and unaccountable variability of cupric sulfate tested on \textit{Pimephales promelas} for 96 h (C).
+Purple dots depict Standartox geometric mean estimates and black dots show un-aggregated values for each group.
 To facilitate readability, data points are randomly scattered along a hypothetical y-axis and are greyed out within the violins.}

### 2020-03-13 article/00_article2.tex
Updates from Overleaf

%  LaTeX support: latex@mdpi.com 
%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).

%=================================================================
\documentclass[journal,article,submit,moreauthors,pdftex]{Definitions/mdpi}

% If you would like to post an early version of this manuscript as a preprint, you may use preprint as the journal and change 'submit' to 'accept'. The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

%--------------------
% Class Options:
%--------------------
%----------
% journal
%----------
% Choose between the following MDPI journals:
% acoustics, actuators, addictions, admsci, aerospace, agriculture, agriengineering, agronomy, ai, algorithms, animals, antibiotics, antibodies, antioxidants, applmech, applsci, arts, asc, asi, atmosphere, atoms, axioms, batteries, bdcc, behavsci , beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsci , buildings, cancers, carbon , catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, children, civileng, cleantechnol, climate, clockssleep, cmd, coatings, colloids, computation, computers, condensedmatter, cosmetics, cryptography, crystals, dairy, data, dentistry, designs , diagnostics, diseases, diversity, drones, econometrics, economies, education, ejbc, ejihpe, electrochem, electronics, endocrines, energies, entropy, environments, epigenomes, est, fermentation, fibers, fire, fishes, fluids, foods, forecasting, forests, fractalfract, futureinternet, futurephys, galaxies, games, gastrointestdisord, gels, genealogy, genes, geohazards, geosciences, geriatrics, hazardousmatters, healthcare, hearts, heritage, highthroughput, horticulturae, humanities, hydrology, ijerph, ijfs, ijgi, ijms, ijtpp, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jcdd, jce, jcm, jcp, jcs, jdb, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmse, jne, jnt, jof, joitmc, jpm, jrfm, jsan, land, languages, laws, life, literature, logistics, lubricants, machines, magnetochemistry, make, marinedrugs, materials, mathematics, mca, medicina, medicines, medsci, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, modelling, molbank, molecules, mps, mti, nanomaterials, ncrna, ijns, neurosci, neuroglia, nitrogen, notspecified, nutrients, oceans, ohbm, optics, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, physics, plants, plasma, pollutants, polymers, polysaccharides, preprints , proceedings, processes, prosthesis, proteomes, psych, publications, quantumrep, quaternary, qubs, reactions, recycling, religions, remotesensing, reprodmed, reports, resources, risks, robotics, safety, sci, scipharm, sensors, separations, sexes, signals, sinusitis, smartcities, sna, societies, socsci, soilsystems, sports, standards, stats, surfaces, surgeries, sustainability, sustainableworld, symmetry, systems, technologies, telecom, test, tourismhosp, toxics, toxins, transplantology, tropicalmed, universe, urbansci, vaccines, vehicles, vetsci, vibration, viruses, vision, water, wem, wevj

%---------
% article
%---------
% The default type of manuscript is "article", but can be replaced by: 
% abstract, addendum, article, benchmark, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, conferenceproceedings, correction, conferencereport, expressionofconcern, extendedabstract, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimages, letter, meetingreport, newbookreceived, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, supfile, technicalnote, viewpoint
% supfile = supplementary materials

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.

%=================================================================
\firstpage{1} 
\makeatletter 
\setcounter{page}{\@firstpage} 
\makeatother
\pubvolume{xx}
\issuenum{1}
\articlenumber{5}
\pubyear{2020}
\copyrightyear{2020}
%\externaleditor{Academic Editor: name}
\history{Received: date; Accepted: date; Published: date}
%\updates{yes} % If there is an update available, un-comment this line

%% MDPI internal command: uncomment if new journal that already uses continuous page numbers 
%\continuouspages{yes}

%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx,epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Title}

% Author Orchid ID: enter ID or remove command
\newcommand{\orcidauthorA}{0000-0000-000-000X} % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-000-000X} % Add \orcidB{} behind the author's name

% Authors, for the paper (add full first names)
\Author{Firstname Lastname $^{1,\dagger,\ddagger}$\orcidA{}, Firstname Lastname $^{1,\ddagger}$ and Firstname Lastname $^{2,}$*}

% Authors, for metadata in PDF
\AuthorNames{Firstname Lastname, Firstname Lastname and Firstname Lastname}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Affiliation 1; e-mail@e-mail.com\\
$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}

% Contact information of the corresponding author
\corres{Correspondence: e-mail@e-mail.com; Tel.: (optional; include country code; if there are multiple corresponding authors, add author initials) +xx-xxxx-xxx-xxxx (F.L.)}

% Current address and/or shared authorship
\firstnote{Current address: Affiliation 3} 
\secondnote{These authors contributed equally to this work.}
% The commands \thirdnote{} till \eighthnote{} are available for further notes

%\simplesumm{} % Simple summary

%\conference{} % An extended version of a conference paper

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{A single paragraph of about 200 words maximum. For research articles, abstracts should give a pertinent overview of the work. We strongly encourage authors to use the following style of structured abstracts, but without headings: (1) Background: Place the question addressed in a broad context and highlight the purpose of the study; (2) Methods: Describe briefly the main methods or treatments applied; (3) Results: Summarize the article's main findings; and (4) Conclusion: Indicate the main conclusions or interpretations. The abstract should be an objective representation of the article, it must not contain results which are not presented and substantiated in the main text and should not exaggerate the main conclusions.}

% Keywords
\keyword{keyword 1; keyword 2; keyword 3 (list three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline.)}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Diversity
%\LSID{\url{http://}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences:
%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data:
%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately. If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal. In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}

%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Toxins
%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}

%\setcounter{secnumdepth}{4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setcounter{section}{-1} %% Remove this when starting to work on the template.
\section{How to Use this Template}
The template details the sections that can be used in a manuscript. Note that the order and names of article sections may differ from the requirements of the journal (e.g., the positioning of the Materials and Methods section). Please check the instructions for authors page of the journal to verify the correct order and names. For any questions, please contact the editorial office of the journal or support@mdpi.com. For LaTeX related questions please contact latex@mdpi.com.
%The order of the section titles is: Introduction, Materials and Methods, Results, Discussion, Conclusions for these journals: aerospace,algorithms,antibodies,antioxidants,atmosphere,axioms,biomedicines,carbon,crystals,designs,diagnostics,environments,fermentation,fluids,forests,fractalfract,informatics,information,inventions,jfmk,jrfm,lubricants,neonatalscreening,neuroglia,particles,pharmaceutics,polymers,processes,technologies,viruses,vision

\section{Introduction}
The introduction should briefly place the study in a broad context and highlight why it is important. It should define the purpose of the work and its significance. The current state of the research field should be reviewed carefully and key publications cited. Please highlight controversial and diverging hypotheses when necessary. Finally, briefly mention the main aim of the work and highlight the principal conclusions. As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research. Citing a journal paper \cite{ref-journal}. And now citing a book reference \cite{ref-book}. Please use the command \citep{ref-journal} for the following MDPI journals, which use author-date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Languages, Laws, Religions, Risks, Social Sciences.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

This section may be divided by subheadings. It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
\begin{quote}
This section may be divided by subheadings. It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
\end{quote}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Subsection}
\unskip
\subsubsection{Subsubsection}

Bulleted lists look like this:
\begin{itemize}[leftmargin=*,labelsep=5.8mm]
\item	First bullet
\item	Second bullet
\item	Third bullet
\end{itemize}

Numbered lists can be added as follows:
\begin{enumerate}[leftmargin=*,labelsep=4.9mm]
\item	First item 
\item	Second item
\item	Third item
\end{enumerate}

The text continues here.

\subsection{Figures, Tables and Schemes}

All figures and tables should be cited in the main text as Figure 1, Table 1, etc.

\begin{figure}[H]
\centering
\includegraphics[width=2 cm]{Definitions/logo-mdpi}
\caption{This is a figure, Schemes follow the same formatting. If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel. (\textbf{b}) Description of what is contained in the second panel. Figures should be placed in the main text near to the first time they are cited. A caption on a single line should be centered.}
\end{figure}

Text

Text

\begin{table}[H]
\caption{This is a table caption. Tables should be placed in the main text near to the first time they are cited.}
\centering
%% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{ccc}
\toprule
\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
\midrule
entry 1		& data			& data\\
entry 2		& data			& data\\
\bottomrule
\end{tabular}
\end{table}

Text

Text

%\begin{listing}[H]
%\caption{Title of the listing}
%\rule{\textwidth}{1pt}
%\raggedright Text of the listing. In font size footnotesize, small, or normalsize. Preferred format: left aligned and single spaced. Preferred border format: top border line and bottom border line.
%\rule{\textwidth}{1pt}
%\end{listing}

\subsection{Formatting of Mathematical Components}

This is an example of an equation:

\begin{equation}
a + b = c
\end{equation}
%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments). This ensures correct linenumbering. The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.

Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
%% Example of a theorem:
\begin{Theorem}
Example text of a theorem.
\end{Theorem}

The text continues here. Proofs must be formatted as follows:

%% Example of a proof:
\begin{proof}[Proof of Theorem 1]
Text of the proof. Note that the phrase `of Theorem 1' is optional if it is clear which theorem is being referred to.
\end{proof}
The text continues here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

Authors should discuss the results and how they can be interpreted in perspective of previous studies and of the working hypotheses. The findings and their implications should be discussed in the broadest context possible. Future research directions may also be highlighted.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and Methods}

Materials and Methods should be described with sufficient details to allow others to replicate and build on published results. Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers. Please disclose at the submission stage any restrictions on the availability of materials or information. New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.

Research manuscripts reporting large datasets that are deposited in a publicly available database should specify where the data have been deposited and provide the relevant accession numbers. If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review. They must be provided prior to publication.

Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Patents}
This section is not mandatory, but may be added if there are patents resulting from the work reported in this manuscript.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\supplementary{The following are available online at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title.}

% Only for the journal Methods and Protocols:
% If you wish to submit a video article, please do so with any other supplementary material.
% \supplementary{The following are available at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title. A supporting video article is available at doi: link.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided. The following statements should be used ``Conceptualization, X.X. and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y. and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing--original draft preparation, X.X.; writing--review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y. All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation. Authorship must be limited to those who have contributed substantially to the work reported.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''. Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections. This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\conflictsofinterest{Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results. Any role of the funders in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript, or in the decision to publish the results must be declared in this section. If there is no role, please state ``The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, or in the decision to publish the results''.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\abbreviations{The following abbreviations are used in this manuscript:\\

\noindent 
\begin{tabular}{@{}ll}
MDPI & Multidisciplinary Digital Publishing Institute\\
DOAJ & Directory of open access journals\\
TLA & Three letter acronym\\
LD & linear dichroism
\end{tabular}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
\appendix
\section{}
\unskip
\subsection{}
The appendix is an optional section that can contain details and data supplemental to the main text. For example, explanations of experimental details that would disrupt the flow of the main text, but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data is shown in the main text can be added here if brief, or as Supplementary data. Mathematical proofs of results not central to the paper can be added as an appendix.

\section{}
All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\reftitle{References}

% Please provide either the correct journal abbreviation (e.g. according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.

%=====================================
% References, variant A: external bibliography
%=====================================
%\externalbibliography{yes}
%\bibliography{your_external_BibTeX_file}

%=====================================
% References, variant B: internal bibliography
%=====================================
\begin{thebibliography}{999}
% Reference 1
\bibitem[Author1(year)]{ref-journal}
Author1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
% Reference 2
\bibitem[Author2(year)]{ref-book}
Author2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
\end{thebibliography}

% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences. For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p. 328; \citeyear{ref-journal-3b}, p.475). This produces: Wong (1999, p. 328; 2000, p. 475)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\sampleavailability{Samples of the compounds ...... are available from the authors.}

%% for journal Sci
%\reviewreports{\\
%Reviewer 1 comments and authors’ response\\
%Reviewer 2 comments and authors’ response\\
%Reviewer 3 comments and authors’ response
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}

### 2020-03-13 article/parts/02_data_description.tex
Updates from Overleaf

@@
 EC50, LD50), henceforth abbreviated as XX50, lowest observed effect levels/concentrations (LOEC/L), henceforth abbreviated as LOEX and no observed effect levels/concentrations (NOEL/C), henceforth abbreviated as NOEX (Table \ref{tab:endpoints-conflate}).
-Standartox allows the ecotoxicity data to be filtered by 23 distinct effect groups (e.g.
+Standartox allows the ecotoxicity data to be filtered by effect groups (e.g.
 mortality, population, growth) (Fig.
 \ref{fig:stx-parameters}B) and concentration types (e.g.
-formulation, active ingredient).
-In addition to these test-specific parameters, Standartox data entries can be filtered by chemical-specific parameters such as the CAS number and chemical classes (e.g.
+formulation, active ingredient) as well as test durations (in hours).
+In addition to these test-specific parameters, Standartox data entries can be filtered by chemical-specific parameters such as the CAS number and chemical roles (e.g.
 pesticides, metals, drugs) (Fig.
+\ref{fig:stx-parameters}B) and classes (e.g.
+organochlorine, triazine) (Fig.
 \ref{fig:stx-parameters}C).
@@
 \ref{fig:stx-parameters}F) as well as different taxonomic levels.
-Additionally users can refine the results to specific test durations (in hours).
 
@@
 \ref{fig:stx-variability}A).
-Moreover, multiple ecotoxicity values are available for individual species-chemical combinations and these can exhibit high variability due to several factors such as durations of ecotoxicity tests (Fig.
+Moreover, multiple ecotoxicity values are available for individual species-chemical combinations and these can also exhibit high variability due to several factors such as durations of ecotoxicity tests (Fig.
 \ref{fig:stx-variability}B), experimental conditions and physiology and genetic strain of the test individuals or populations.
@@
 for an individual species, across species of a genus or family), and chemical grouping (e.g.
-across all pesticides), Standartox provides several aggregation methods including the minimum, the maximum and the geometric mean as aggregates from the filtered data set for each chemical.
+across all pesticides), Standartox provides several aggregation methods including the minimum, the maximum and the geometric mean allowing to aggregate the filtered data set.
 The geometric mean is preferred in comparison to the arithmetic mean, because it is considerably less influenced by outliers and is suitable for skewed data.
@@
 In the course of the aggregation process, outliers that exceed 1.5 times the interquartile range are flagged to caution Standartox users.
-However, since the geometric mean is relatively robust against outliers, they are not removed automatically.
+However, since the geometric mean is relatively robust against outliers, they are not removed automatically from the aggregation.
 Overall, aggregation with Standartox allows to reduce the between-study variability when toxicity data are aggregated.
@@
 Similarly, we compared Standartox to ecotoxicity values for \textit{D.
-magna} from the ChemProp \citep{ufzdepartmentofecologicalchemistry_chemprop_2019} software, which estimates LC\textsubscript{50} values via quantitative structure-activity relationship (QSAR) models \citep{schuurmann_quantitative_2011} We found that  95 \% of Standartox values lie within one order of magnitude of Chemprop (n = 179).
+magna} from the ChemProp \citep{ufzdepartmentofecologicalchemistry_chemprop_2019} software, which estimates LC\textsubscript{50} values via quantitative structure-activity relationship (QSAR) models \citep{schuurmann_quantitative_2011} We found that  95 \% of Standartox values lie within one order of magnitude of the Chemprop (n = 179) vlaues.
 However, the diﬀerence is not necessarily an indication of lower quality of Standartox estimates but may also reflect the wider range of experimental conditions for which data are available in the database underlying Standartox as well as inaccurate predictions for QSAR models, respectively (Figure \ref{fig:standartox_ppdb_diff}).
@@
 Indeed, the increasing amount of data from ecotoxicological tests and experiments that is becoming available has elicited several initiatives to harmonize these data.
-These initiatives partly aim for overlapping goals, yet have limitations or objectives that discriminate them from Standartox.
-The Network of reference laboratories, research centres and related organisations for monitoring of emerging environmental substances (NORMAN) focuses on assembling river basin specific pollutants from studies and databases \citep{von_der_ohe_new_2011}.
-The EnviroTox database \href{https://envirotoxdatabase.org/} which also uses, amongst others the ECOTOX database as an input has recently been published \citep{healthandenvironmentalsciencesinstitutehesi_envirotox_2019, connors_creation_2019}.
+These initiatives partly aim for overlapping goals, yet have limitations or objectives that discriminate them from Standartox:
+
+Comptox, is a web tool published by the EPA which, similar to Standartox allows for filtering test results, the retrieval of additional chemical information and the estimation of toxicity equivalents, such as 48 h \textit{Daphnia magna} LC\textsubscript{50} values.
+However, toxicity estimations are limited to standard organisms (e.g.
+\textit{Daphnia magna}), and the tool lacks the possibility for automated retrieval \citep{williams_comptox_2017}.
+
+The EnviroTox database which also uses, amongst others the ECOTOX database as an input has recently been published \citep{healthandenvironmentalsciencesinstitutehesi_envirotox_2019, connors_creation_2019}.
 In contrast to Standartox, EnviroTox  is restricted to selected aquatic organisms (i.e.
-fish, amphibians, invertebrates and algae) and experimental durations (at least 24 h) and uses a qualitative algorithm to derive single ecotoxicity values.
+fish, amphibians, invertebrates and algae) and experimental durations (at least 24 h) and uses a rule-based algorithm to derive single ecotoxicity values.
 Besides, EnviroTox provide additional information on toxicity endpoints, such as acute or chronic classifications and mode of action assignments.
-An action that can also be allocated to the user, since there are several different ways to classify such parameters, especially those mentioned, guaranteeing user flexibility.
+We think, an action that can also be allocated to the user, since there are several different ways to classify such parameters, especially those mentioned, guaranteeing user flexibility.
 The EnviroTox database allows for an aggregation into single toxicity values for individual taxa, whereas Standartox performs this aggregation for individual taxa-chemical combinations.
-However, the Standartox results for individual taxa-chemical combinations could easily be aggregated across chemicals in a second step based on chemical parameters.
-Comptox, is a web tool published by the EPA which, similar to Standartox allows for filtering test results, the retrieval of additional chemical information and predicting toxicity properties, such as 48 h \textit{Daphnia magna} LC\textsubscript{50} values.
-However, toxicity predictions are limited to standard organisms (e.g.
-\textit{Daphnia magna}), and the tool lacks the possibility for automatic retrieval of predicted \citep{williams_comptox_2017}.
+However, the Standartox results for individual taxa-chemical combinations could easily be aggregated across chemicals in a second step to achieve aggregations performed in EnviroTox.
+
+The Etox database, like the ECOTOX database collects and provides filter methods for ecotoxicological test information.
+
+The PPDB provides data only on pesticides and as mentioned before provides single quality controlled values only for commonly used taxa, e.g. \textit{D. magna} or \textit{R. subcapitata}.
 
 In summary, none of the above mentioned initiatives aim for a holistic and automated standardized aggregation method of exposure endpoints for individual chemicals.
-In addition, they lack the possibility to access the databases through common high level programming languages, such as R.
+In addition, they lack the possibility to access the databases through a common high level programming language, such as R.
 As outlined above, toxicity estimates from different studies can be highly variable due to a wide range of experimental conditions such as pH, temperature or conductivity \citep{rosenkrantz_influence_2013, li_temperature_2011}.
-Integrating these factors into the aggregated estimates would improve toxicity estimates .
+Integrating these factors into the aggregated estimates would improve toxicity estimates.
 However, the current implementation of Standartox does not consider these factors, because the ECOTOX database only provides sparse records on experimental conditions.

### 2020-03-13 article/parts/03_methods.tex
Updates from Overleaf

@@
 An automated processing pipeline downloads the quarterly released ECOTOX database, performs several preparation steps on it and exports a final Standartox data set.
-This data set is accessible via a web application (APP) and an application programming interface (API).
+This data set is accessible via a web application and an application programming interface (API).
 An API provides the means for machine communication between a host and a client and thus allows scriptable data queries.
 To facilitate the API access, the R \citep{rcoreteam_language_2017} package \textit{standartox} is built.
+All data presented in this paper are derived from the Standartox build based on the ECOTOX release from the 12.12.2019.
 
@@
 In addition lookup tables (Supplement \ref{sup:conv-concentration}, \ref{sup:conv-duration}) that enable duration and concentration unit conversions of the data are created.
-A meta table providing information such as the release version of the ECOTOX database is then added.
-In the next step, Chemical Abstracts Service Numbers (CAS) and International Chemical Identifier (InChI) as well as taxonomic names are used to query additional information from publicly available databases on chemicals and organisms, respectively.
-This includes the Compendium of Pesticide Common Names \citep{wood_compendium_2019}, the Chemical Entities of Biological Interest (ChEBI) database \citep{hastings_chebi_2016}, the Chemical Identifier Resolver (CIR) service \citep{nationalinstitutesofhealthnih_chemical_2019}, the PHYSPROP database, the Pubchem database \citep{kim_pubchem_2016}, Eurostat \citep{europeancommission_eurostat_2019} and Wikidata \citep{vrandecic_wikidata_2014} for chemicals and the World Register of Marine Species (WoRMS) \citep{wormseditorialboard_world_2018} and the Global Biodiversity Information Facility (GBIF) \citep{_gbif_2019} (Table \ref{tab:data-base-additional}) for habitat and spatial distribution of organisms.
-This information is added to Standartox to allow filtering for specific classes of chemicals as well as spatial distribution (i.e.
+A meta table providing information such as the release version of the ECOTOX database is added.
+Then, provided Chemical Abstracts Service Numbers (CAS) and taxonomic names are used to query additional information from publicly available databases on chemicals and organisms, respectively.
+This includes the Compendium of Pesticide Common Names \citep{wood_compendium_2019}, the Chemical Entities of Biological Interest (ChEBI) database \citep{hastings_chebi_2016}, the Chemical Identifier Resolver (CIR) service \citep{nationalinstitutesofhealthnih_chemical_2019}, the Pubchem database \citep{kim_pubchem_2016}, Eurostat \citep{europeancommission_eurostat_2019} and Wikidata \citep{vrandecic_wikidata_2014} for chemicals and the World Register of Marine Species (WoRMS) \citep{wormseditorialboard_world_2018} and the Global Biodiversity Information Facility (GBIF) \citep{gbif_gbif_2019} (Table \ref{tab:data-base-additional}) for habitat and spatial distribution of organisms.
+Given that CAS and taxonomic names are not always unambiguous, e.g.
+the genus \textit{Eisenia} can refer to an algae and a worm at once, we first match them against the database specific identifiers and subsequently use them to retrieve the actual data.
+In a next step the data is added to Standartox to enable filtering for specific chemical roles (e.g.
+drug, pesticide, personal care product) and classes (e.g.
+pyrethroid, metal, carbamate) as well as spatial distribution (i.e.
 continents) and habitat preferences (e.g.
@@
 Finally, the Standartox data set is compiled, which includes the harmonisation of data, e.g.
-through conversion of units related to the concentrations and test duration.
-Out of the unique 1229 concentration units in the ECOTOX database, Standartox retains only those (n = X) that are unambiguously convertible to one of the following units: $\mu$g/l, g/m2, ppb, mg/kg, \% and flags, the remaining as \textit{other}, resulting in 873473 out of 952625 test results.
+through conversion of result concentration and test duration units.
+1237 distinct concentration units are converted to X harmonised ones, as are the 126 distinct duration units to X.
+Standartox retains only duration units, that can unambiguously be converted to hours, excluding duration units such as harvest or lifetime.
+To guarantee appropriate unit conversion and harmonisation, we compared one automated conversion to a manually done conversion for each of the distinct concentration and duration units.
 Furthermore, the units are cleaned, for example through removing additional information in the field such as \textit{food}, \textit{soil}, \textit{ai} that are also coded in other variables and hinder the processing of units.
-Concentrations that are given as rates such as per day (mg/kg/day) are generally excluded.
-Likewise, out of the 125 test duration units in the EPA ECOTOX, Standartox retains only those (n = X) that can be converted to hours, excluding ambiguous duration units such as harvest or lifetime and keeping 905110 out of 952625 test results.
-Thereby concentration unit additions such as \textit{food}, \textit{soil}, \textit{ai (active ingredient)} etc.
-are neglected since they are also coded in other variables and only hinder the processing of units.
-Units such as per day (mg/kg/day) are generally excluded.
-Experimental responses are restricted to three groups, namely NOEX, LOEX and XX50.
-The former two represent no-observed-adverse-effect and lowest-observed-adverse-effect concentrations or levels, respectively.
-The latter includes various sub-groups of the half maximal effective concentration, where half of the individuals in an experiment display an effect (e.g.
-EC50, LC50, LD50), which is a common measure in ecotoxicology \citep{malaj_organic_2014}.
-Other responses, such as Bioconcentration factors (BCF), non-half maximal effective concentrations (e.g.
+Concentrations that are given as rates such as per day (e.g.
+mg/kg/day) are multiplied by the days and the converted.
+Experimental endpoints are restricted to three groups, namely NOEX, LOEX and XX50.
+Other endpoints, such as Bioconcentration factors, non-half maximal effective concentrations (e.g.
 IC10, EC25, LD99) or maximum acceptable toxicant (MATC) concentrations are removed.
-Beyond that, filters for the CAS number, the concentration type (e.g.
-active ingredient, formulation), the chemical class (e.g.
-fungicides, metals), taxon, organism habitat (e.g freshwater, terrestrial) and region (e.g.
-Europe, Asia), test duration as well as effect type (e.g.
-mortality, growth) are created.
-To detect possible outliers among the test results, we flag values that exceed 1.5 times the interquartile range (IQR) within groups with identical test parameters (e.g.
-chemical, taxon, duration).
-Along with the compiled Standartox data set, a catalog, listing all distinct entries and value ranges, for categorical and continuous variables, respectively is created.
+Overall this reduces the ECOTOX data set from 971,430 to 611,248 harmonised test results that are included in Standartox.
+Along with that, a catalog, listing all distinct entries and value ranges, for categorical and continuous variables, respectively is created.
 Finally, we run quality control scripts that check the accuracy of the data.
 Single processing scripts are listed in Figure \ref{fig:pipeline-tree}.
-The final Standartox table together with the catalog is exported and accessible via web application and the API.
+The final Standartox table together with the catalog is exported and accessible via the web application and the API, through the R package.
 
@@
 
-\subsection{Access}
-The APP and the API load the compressed serialized Standartox data into memory and allow a client to interact with it.
+\subsection{Application methods}
+The web application and the API load the compressed serialized Standartox data into memory and allow a client to interact with it.
 The client thereby calls the functions stx\_filter() and stx\_aggregate() that filter and aggregate the data according to specific parameters (Table \ref{tab:app-parameters}).
-The interactive APP is built in R using the shiny web application framework, which runs with the help of a shiny server \citep{chang_shiny_2018} on \url{standartox.uni-landau.de}.
-The API is built by using the R package \textit{plumber} \citep{trestletechnologyllc_plumber_2018}, which allows for the creation of Representational state transfer (REST) APIs from R.
+The interactive web application is built in R using the shiny framework, which runs with the help of a shiny server \citep{chang_shiny_2018} on \url{standartox.uni-landau.de}.
+The API is built by using the R package \textit{plumber} \citep{trestletechnologyllc_plumber_2018}, which allows for the creation of Representational State Transfer (REST) APIs from R.
 REST is a software standard that defines web service communication rules.
@@
 In order to facilitate the API usage the R-package \textit{standartox} is created.
-
-\subsection{Technical Validity}
-To guarantee appropriate unit conversion and harmonisation, we compared for each of the 1229 distinct concentration and for each of the 129 duration units the automatically converted units to one manually calculated one.

### 2020-03-13 article/tables/database-differences.tex
Updates from Overleaf

@@
 \caption{Different databases that provide ecotoxicological data.
-Abbreviations: \textbf{ALL:} Most important parameters, such as chemical, Organism, Duration for filtering ecotoxicological data are incorporated.
-\textbf{WEB:} Accessible via a web application and a graphical user interface, not providing programmatic access.}
+Abbreviations: \textbf{ALL:} Most important test parameters, including chemical, taxon, duration for filtering ecotoxicological data are incorporated.
+\textbf{WEB:} Accessible via a web application through a graphical user interface.
+\textbf{API:} Accessible via an application programming interface.}
 \label{tab:database-differences}

### 2020-03-13 article/parts/04_user_notes.tex

@@
 \section{User notes}
-By accessing the APP users can filter and download the resulting data sets as a comma-separated values (csv) file.
+By accessing the web application users can filter and download the resulting data sets as a comma-separated values (csv) file.
 Users of the R-package can directly load the data in R.
 The R-package provides the two functions \textit{stx\_catalog()} and \textit{stx\_query()}.
-The first queries a catalog of possible Standartox parameters into a R list object.
+The first command queries a catalog of possible Standartox parameters into a R list object.
 The latter allows users to set the Standartox filter parameters and to fetch the actual data.

### 2020-03-14 article/article.tex
cleaned article folder

@@
-\documentclass[english]{article}
-\usepackage[T1]{fontenc}
-\usepackage{natbib} % citations
-\usepackage{csvsimple}
-\usepackage{pgfplotstable}
-\usepackage{rotating} % to rotate tables
-%% tikz packages
-\usepackage{tikz}
-\usepackage{forest}
-\usetikzlibrary{trees, arrows, shapes.geometric, positioning}
+%  LaTeX support: latex@mdpi.com 
+%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).
+
+%=================================================================
+\documentclass[journal,datadescriptor,accept,moreauthors,pdftex]{Definitions/mdpi}
+
+% If you would like to post an early version of this manuscript as a preprint, you may use preprint as the journal and change 'submit' to 'accept'.
+The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}.
+This is especially recommended for submission to arXiv, where line numbers should be removed before posting.
+For preprints.org, the editorial staff will make this change immediately prior to posting.
+
+%--------------------
+% Class Options:
+%--------------------
+%----------
+% journal
+%----------
+% Choose between the following MDPI journals:
+% acoustics, actuators, addictions, admsci, aerospace, agriculture, agriengineering, agronomy, algorithms, animals, antibiotics, antibodies, antioxidants, applsci, arts, asc, asi, atmosphere, atoms, axioms, batteries, bdcc, behavsci , beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsci , buildings, cancers, carbon , catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, children, cleantechnol, climate, clockssleep, cmd, coatings, colloids, computation, computers, condensedmatter, cosmetics, cryptography, crystals, dairy, data, dentistry, designs , diagnostics, diseases, diversity, drones, econometrics, economies, education, ejihpe, electrochem, electronics, energies, entropy, environments, epigenomes, est, fermentation, fibers, fire, fishes, fluids, foods, forecasting, forests, fractalfract, futureinternet, futurephys, galaxies, games, gastrointestdisord, gels, genealogy, genes, geohazards, geosciences, geriatrics, hazardousmatters, healthcare, heritage, highthroughput, horticulturae, humanities, hydrology, ijerph, ijfs, ijgi, ijms, ijns, ijtpp, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jcdd, jcm, jcp, jcs, jdb, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmse, jnt, jof, joitmc, jpm, jrfm, jsan, land, languages, laws, life, literature, logistics, lubricants, machines, magnetochemistry, make, marinedrugs, materials, mathematics, mca, medicina, medicines, medsci, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, modelling, molbank, molecules, mps, mti, nanomaterials, ncrna, neuroglia, nitrogen, notspecified, nutrients, ohbm, optics, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, physics, plants, plasma, polymers, polysaccharides, preprints , proceedings, processes, proteomes, psych, publications, quantumrep, quaternary, qubs, reactions, recycling, religions, remotesensing, reports, resources, risks, robotics, safety, sci, scipharm, sensors, separations, sexes, signals, sinusitis, smartcities, sna, societies, socsci, soilsystems, sports, standards, stats, surfaces, surgeries, sustainability, symmetry, systems, technologies, test, toxics, toxins, tropicalmed, universe, urbansci, vaccines, vehicles, vetsci, vibration, viruses, vision, water, wem, wevj
+
+%---------
+% article
+%---------
+% The default type of manuscript is "article", but can be replaced by: 
+% abstract, addendum, article, benchmark, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, conferenceproceedings, correction, conferencereport, expressionofconcern, extendedabstract, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimages, letter, meetingreport, newbookreceived, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, supfile, technicalnote, viewpoint
+% supfile = supplementary materials
+
+%----------
+% submit
+%----------
+% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted.
+This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information.
+Also, line numbering will be removed.
+Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
+
+%------------------
+% moreauthors
+%------------------
+% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
+
+%---------
+% pdftex
+%---------
+% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.
+
+%=================================================================
+\firstpage{1} 
+\makeatletter 
+\setcounter{page}{\@firstpage} 
+\makeatother
+\pubvolume{xx}
+\issuenum{1}
+\articlenumber{5}
+\pubyear{2019}
+\copyrightyear{2019}
+%\externaleditor{Academic Editor: name}
+\history{Received: date; Accepted: date; Published: date}
+\updates{yes} % If there is an update available, un-comment this line
+
+%% MDPI internal command: uncomment if new journal that already uses continuous page numbers 
+%\continuouspages{yes}
+
+%------------------------------------------------------------------
+% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
+%\pdfoutput=1
+
+%=================================================================
+% Add packages and commands here.
+The following packages are loaded in our class file: fontenc, calc, indentfirst, fancyhdr, graphicx, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, amsthm, hyphenat, natbib, hyperref, footmisc, geometry, caption, url, mdframed, tabto, soul, multirow, microtype, tikz
+%% Additional packages
 \usepackage{standalone}
-%% text packages
-\usepackage{siunitx} % fake ° symbol \ang{}
-\usepackage[super]{nth} % for 1st, 2nd etc. \nth{}
-\usepackage{hyperref}
+% code
+\usepackage{listings}
+% tikz
+\usepackage{tikzpeople}
+\usetikzlibrary{trees,
+                arrows,
+                shapes.geometric,
+                positioning,
+                calc,
+                backgrounds,
+                fit}
+\usepackage{rotating}
 
-\newcommand{\standartox}{STANDARTOX}
-\newcommand{\epa}{EPA ECOTOX data base}
-\newcommand{\app}{http://139.14.20.252:3838/etox-base-shiny/}
-\newcommand{\git}{https://github.com/andreasLD/etox-base}
-\newcommand{\gitapp}{https://github.com/andreasLD/etox-base-shiny}
-\newcommand{\ecfifty}{EC\textsubscript{50}}
+%=================================================================
+%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
+%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).
 
-\listfiles
+%=================================================================
+% Full title of the paper (Capitalized)
+\Title{Standartox: Standardising toxicity data}
 
+% Author Orchid ID: enter ID or remove command
+\newcommand{\orcidauthorA}{0000-0002-9290-3965} % Add \orcidA{} behind the author's name
+\newcommand{\orcidauthorB}{0000-0001-8732-3766} % Add \orcidB{} behind the author's name
+\newcommand{\orcidauthorC}{0000-0003-3510-1701} % Add \orcidB{} behind the author's name
+
+% Authors, for the paper (add full first names)
+\Author{Andreas Scharmüller $^{1}$*\orcidA{}, Verena C. Schreiner$^{1}$\orcidB and Ralf B. Schäfer $^{1}$\orcidC{}}
+
+% Authors, for metadata in PDF
+\AuthorNames{Andreas Scharmüller, Verena C. Schreiner and Ralf B. Schäfer}
+
+% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
+\address{%
+$^{1}$ \quad iES Landau, Institute for Environmental Sciences, University of Koblenz-Landau,
+D-76829 Landau, Germany; andschar@protonmail.com (A.S.), schreiner-verena@uni-landau.de (V.C.S.), schaefer-ralf@uni-landau.de (R.B.S.)}
+
+% Contact information of the corresponding author
+\corres{Correspondence: andschar@protonmail.com}
+
+% Current address and/or shared authorship
+% \firstnote{Current address: Affiliation 3} 
+% \secondnote{These authors contributed equally to this work.}
+% The commands \thirdnote{} till \eighthnote{} are available for further notes
+
+%\simplesumm{} % Simple summary
+
+%\conference{} % An extended version of a conference paper
+
+% Abstract (Do not insert blank lines, i.e. \\) 
+\abstract{An increasing number of chemicals such as pharmaceuticals, pesticides and synthetic hormones are in daily use all over the world.
+In the environment, chemicals can adversely affect populations and communities and in turn related ecosystem functions.
+To evaluate the risks from chemicals for ecosystems data on their toxicity, which is typically produced in standardized ecotoxicological laboratory tests, is required.
+However, for many chemicals, multiple ecotoxicity values are available for the same test organism, which can vary strongly.
+Current (Meta-) databases, including the  United States Environmental Protection Agency ECOTOXicology knowledgebase only collect ecotoxicological test data, without allowing for further aggregation steps.
+Others, such as the  Pesticide Property Data Base (PPDB) are confined to specific chemicals, here pesticides, thereby ignoring vast amounts of ecotoxicity data.
+We therefore developed Standartox, a tool and database that aims to overcome the limitations of other databases by continuously incorporating the ever-growing number of test results in an automated process workflow that ultimately leads to a single aggregated data point for a specific chemical-organism test combination, representing the toxicity of a chemical.
+Standartox comes with two front-ends, a web application (standartox.uni-landau.de) and an R package (standartox).}
+
+% Keywords
+\keyword{ecotoxicology; standardisation; environmental risk assessment; effect database; database; software; r; api;} % up to 10
+
+% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
+%\PACS{J0101}
+%\MSC{}
+%\JEL{}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Diversity
+%\LSID{\url{http://}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Applied Sciences:
+%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Data:
+\dataset{https://cfpub.epa.gov/ecotox/}
+%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately.
+If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal.
+In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}
+
+%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}
+\datasetlicense{MIT}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Toxins
+%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}
+
+%\setcounter{secnumdepth}{4}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{document}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\title{Etox-Base: A tool to derive exposure endpoints from chemical test results}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%The order of the section titles is: Introduction, Materials and Methods, Results, Discussion, Conclusions for these journals: aerospace,algorithms,antibodies,antioxidants,atmosphere,axioms,biomedicines,carbon,crystals,designs,diagnostics,environments,fermentation,fluids,forests,fractalfract,informatics,information,inventions,jfmk,jrfm,lubricants,neonatalscreening,neuroglia,particles,pharmaceutics,polymers,processes,technologies,viruses,vision
 
-\author{Andreas Scharm{\"u}ller\textsuperscript{1{*}},
-        Verena Schreiner\textsuperscript{1},
-        Ralf B. Sch{\"a}fer\textsuperscript{1}}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Summary - required}
+% The introduction should briefly place the study in a broad context and highlight why it is important.
+It should define the purpose of the work and its significance.
+The current state of the research field should be reviewed carefully and key publications cited.
+Please highlight controversial and diverging hypotheses when necessary.
+Finally, briefly mention the main aim of the work and highlight the principal conclusions.
+As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research.
+Citing a journal paper \cite{ref-journal}.
+And now citing a book reference \cite{ref-book}.
+Please use the command \citep{ref-journal} for the following MDPI journals, which use author-date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Languages, Laws, Religions, Risks, Social Sciences.
+An increasing number of chemicals such as pharmaceuticals, pesticides and synthetic hormones are in daily use all over the world.
+In Europe alone, some 100,000 chemicals are estimated to be in current use, whereof 30,000 are produced in quantities larger than one ton per year \citep{breithaupt_costs_2006}.
+Except for pesticides that are released into the environment deliberately, most chemicals enter the environment as a result of their use through different paths (e.g.
+atmospheric emission and deposition or discharge through wastewater) \citep{schwarzenbach_challenge_2006}.
+In the environment, chemicals can adversely affect populations and communities and in turn related ecosystem functions \citep{schafer_thresholds_2012, malaj_organic_2014, hallmann_declines_2014, barracaracciolo_pharmaceuticals_2015, johnston_review_2015}.
+Ultimately, this may compromise natures contribution to human well-being, for example the ecosystem services clean drinking and irrigation water as well as food production \citep{peters_review_2013, vandersluijs_neonicotinoids_2013, yamamuro_neonicotinoids_2019}.
+Pollution with man-made chemicals has been identified as one of three major environmental problems for which research gaps hamper the derivation of planetary boundaries, i.e.
+thresholds beyond which irreversible state shifts may occur \citep{steffen_anthropocene_2007, steffen_planetary_2015}.
+\citet{bernhardt_synthetic_2017} argue that the knowledge gap how chemicals affect populations, communities and in turn ecosystem functions and services, may impede the accomplishment of the Sustainable Development Goals \citep{rosa_transforming_2017} of the United Nations.
+Even highly regulated chemical compounds such as pesticides have been shown to cause strong adverse effects on non-target organisms, such as birds \citep{hallmann_declines_2014} aquatic insects \citep{beketov_pesticides_2013} or fish \citep{yamamuro_neonicotinoids_2019}, questioning the current regulation efforts \citep{schafer_future_2019}.
 
-\maketitle
-\thispagestyle{fancy}
+To evaluate the risks from chemicals for ecosystems, data on their toxicity is required, which is typically produced in standardised ecotoxicological laboratory tests.
+For example, \citet{morrissey_neonicotinoid_2015} used ecotoxicological test results from 49 insects and crustaceans to evaluate the effect of neonicotinoid insecticides in the aquatic ecosystem.
+Also, \citet{malaj_organic_2014} compiled experimental toxicity test results for 223 chemicals to assess the health of freshwater ecosystems in Europe.
+Similarly, permissible environmental concentrations are often derived from these test data, typically by a combination with safety factors to account for uncertainties.
+The test data mainly relate to a few, well tested standard organisms, such as the brown rat \textit{Rattus norvegicus}, the water flea \textit{Daphnia magna} and the microalga \textit{Raphidocelis subcapitata}.
+Nevertheless, a much greater variety of organisms has been used in ecotoxicological experiments.
 
-1. Institute of Environmental Sciences, University of Koblenz-Landau Fortstraße 7, 76829 Landau, Germany {*}corresponding author(s):
-Andreas Scharm{\"u}ller (scharmueller@uni-landau.de)
+To date, only few initiatives exist that aim to create a public resource of ecotoxicological data, such as the United States Environmental Protection Agency ECOTOXicology knowledgebase (ECOTOX) (ca.
+1,000,000 test results, 13,000 taxa, 12,000 chemicals) \citep{elonen_ecotoxicology_2018}, the German Environmental Agency's Information System Ecotoxicology and Environmental Quality Targets (ETOX) \citep{umweltbundesamt_etox_2019}, the Pesticide Properties Data Base (PPDB) (ca.
+2000 pesticides) \citep{lewis_international_2016} or the Envirotox database \citep{healthandenvironmentalsciencesinstitutehesi_envirotox_2019, connors_creation_2019}.
+The former two compile all available results from experiments into a database.
+However, for many chemicals, multiple ecotoxicity values are available for the same test organism, which can vary strongly and the lack of associated quality information and heterogeneous units hamper reproducible science.
+The PPDB database, in contrast, provides single ecotoxicity values only for pesticides and a few selected test organisms, thereby covering only a minor fraction of the vast amount of ecotoxicological data.
+The Envirotox is limited to aquatic organisms.
+Moreover, data analyses often require links to additional data resources, for example to append additional chemical and species information (e.g.
+chemical properties, habitat of species), which calls for more automated procedures.
 
-\begin{abstract}
-%%5This is a manuscript template for Data Descriptor submissions to \emph{Scientific
-%%Data} (http://www.nature.com/scientificdata). The abstract must be
-%%no longer than 170 words, and should succinctly describe the study,
-%%the assay(s) performed, the resulting data, and the reuse potential,
-%%but should not make any claims regarding new scientific findings.
-%%No references are allowed in this section.
+We therefore developed Standartox, a tool and database that aims to overcome the limitations of other databases by continuously incorporating the ever-growing number of test results in an automated process workflow that ultimately leads to a cleaned and harmonised ecotoxicity data collection and provides methods to derive single aggregated ecotoxicity values for a specific chemical-organism test combination.
+Standartox makes use of the publicly available and quarterly updated ECOTOX database \citep{usepa_ecotox_2019} and restricts the data to commonly used endpoints in ecotoxicology, such as half maximal effective concentration (EC\textsubscript{50}) or no-observed-adverse-effect concentrations (NOEC), leading to about 600,000 ecotoxicological test results including 8000 chemicals, tested on about 8000 taxa.
+Users can filter test results according to several parameters, e.g.
+refining a search for ecotoxicity data on organisms occurring in specific habitats or regions of the world.
+Above that, Standartox aggregates ecotoxicological test results in a standardized way, by calculating the minimum, the geometric mean and the maximum of the results for each chemical and the associated, user-defined test parameters.
+This tackles the commonly occurring problem of the variability of multiple ecotoxicological test results and which values to choose for assessing the effects of chemicals.
+Hence, Standartox provides the basis for reproducible science and combines information from different sources to simplify the derivation of risk indicators such as Species Sensitivity Distributions (SSD) and Toxic Units (TU), which represent two prominent concepts to assess effects on organisms in ecotoxicology \citep{posthuma_species_2002, kefford_definition_2011, schafer_effects_2011}.
+Besides aggregating ecotoxicological test results, Standartox provides a concise overview of the tested chemicals, allowing the identification of potential knowledge gaps.
+Moreover, Standartox could help in reducing the millions of animals used for toxicity testing each year due to facilitated ecotoxicity data access \citep{hartung_chemical_2009}.
+Standartox comes with two front-ends, a web application: \url{standartox.uni-landau.de} and a R \citep{rcoreteam_language_2017} package \textit{standartox}, providing convenience structures and thereby largely reducing processing time for users.
 
-<WRITE ABSTRACT>
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Data Description - required}
 
-\end{abstract}
-\pagebreak
+Standartox constitutes a collection of 600,000 quality checked ecotoxicological test results comprising roughly 8000 distinct chemicals and 8000 taxa that allows for the retrieval of single toxicity equivalents.
+It is build on the ECOTOX database \citep{usepa_ecotox_2019} and the data are processed to clean and harmonise entries to retrieve comparable toxicity endpoints.
+Subsequently, filter and aggregation methods are provided to allow for the retrieval of single toxicity equivalents for specific experimental conditions.
+The underlying ECOTOX database is updated quarterly and provides on average 5228 (2014 - 2019) new toxicity entries with each update.
+As soon as a new update of ECOTOX is available, Standartox is directly built in an automated way and therefore incorporates new data steadily.
+Standartox comes with a version control to guarantee reproducibility, meaning that once derived Standartox estimates can be retraced at a later stage, although new data might have updated the estimates already.
+Users can reach Standartox via a web application and an R-package (standartox) that accesses an application programming interface (API) providing the possibility of scriptable requests.
 
-%%%%%%%%%%%%%%%%%%%%%% 01 Background
-\input{article/parts/01_background.tex}
-\pagebreak
+\subsection{Filters}
+The data can be restricted to the three endpoint groups, namely half maximal effective/lethal concentration/dose values (e.g.
+EC50, LD50), henceforth abbreviated as XX50, lowest observed effect levels/concentrations (LOEC/L), henceforth abbreviated as LOEX and no observed effect levels/concentrations (NOEL/C), henceforth abbreviated as NOEX (Table \ref{tab:endpoints-conflate}).
+Standartox allows the ecotoxicity data to be filtered by effect groups (e.g.
+mortality, population, growth) (Fig.
+\ref{fig:stx-parameters}B) and concentration types (e.g.
+formulation, active ingredient) as well as test durations (in hours).
+In addition to these test-specific parameters, Standartox data entries can be filtered by chemical-specific parameters such as the CAS number and chemical roles (e.g.
+pesticides, metals, drugs) (Fig.
+\ref{fig:stx-parameters}B) and classes (e.g.
+organochlorine, triazine) (Fig.
+\ref{fig:stx-parameters}C).
+Furthermore, the Standartox data can be refined to organism-specific parameters, such as the organism habitat (e.g.
+freshwater, marine, terrestrial) (Fig.
+\ref{fig:stx-parameters}E), their occurrence regions (e.g.
+Europe, South America) (Fig.
+\ref{fig:stx-parameters}F) as well as different taxonomic levels.
 
-%%%%%%%%%%%%%%%%%%%%%% 02 Methods
-\input{article/parts/02_methods.tex}
-\pagebreak
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{article/figures/standartox_parameters.png}
+    \caption{Share of 10 most frequent entries for the parameters (\textbf{A}) effect group, (\textbf{B}) chemical role, (\textbf{C}) chemical class, (\textbf{D}) taxonomic order, (\textbf{E}) organism habitat and (\textbf{F}) organism occurrence region in Standartox.}
+    \label{fig:stx-parameters}
+\end{figure}
 
-%%%%%%%%%%%%%%%%%%%%%% 03 Data records
-\input{article/parts/03_data_records.tex}
-\pagebreak
+\subsection{Aggregation}
+Typically, species exhibit a differential sensitivity towards chemicals (Fig.
+\ref{fig:stx-variability}A).
+Moreover, multiple ecotoxicity values are available for individual species-chemical combinations and these can also exhibit high variability due to several factors such as durations of ecotoxicity tests (Fig.
+\ref{fig:stx-variability}B), experimental conditions and physiology and genetic strain of the test individuals or populations.
+To aggregate multiple ecotoxicity values into a single value on the desired taxonomic level (e.g.
+for an individual species, across species of a genus or family), and chemical grouping (e.g.
+across all pesticides), Standartox provides several aggregation methods including the minimum, the maximum and the geometric mean allowing to aggregate the filtered data set.
+The geometric mean is preferred in comparison to the arithmetic mean, because it is considerably less influenced by outliers and is suitable for skewed data.
+Also, the geometric mean is preferable over the median, because the median completely ignores the influence of large or small values, making it unreliable for small data sets.
+In the course of the aggregation process, outliers that exceed 1.5 times the interquartile range are flagged to caution Standartox users.
+However, since the geometric mean is relatively robust against outliers, they are not removed automatically from the aggregation.
+Overall, aggregation with Standartox allows to reduce the between-study variability when toxicity data are aggregated.
 
-%%%%%%%%%%%%%%%%%%%%%% 04 Technical validation
-\input{article/parts/04_technical_validation.tex}
-\pagebreak
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{article/figures/results_variability.png}
+    \caption{Violin plots of (\textbf{A}) the variability of test results (EC\textsubscript{50}) in Standartox because of different species (i.e.
+\textit{Xenopus laevis} - Amphibia, \textit{Raphidocelis subcapitata} - Algae, \textit{Oncorhynchus mykiss} - fish, \textit{Lemna minor} - invertebrate) to the chemical atrazine tested for 96 h, (\textbf{B}) diﬀerent durations of toxicity tests with zinc sulfate and \textit{Daphnia magna} and (\textbf{C}) unaccountable variability of cupric sulfate tested on \textit{Pimephales promelas} for 96 h.
+Purple dots depict Standartox geometric mean estimates and black dots show un-aggregated values for each group.
+To facilitate readability, data points are randomly scattered along a hypothetical y-axis and are greyed out within the violins.}
+    \label{fig:stx-variability}
+\end{figure}
 
-%%%%%%%%%%%%%%%%%%%%%% 05 Usage notes
-\input{article/parts/05_usage_notes.tex}
+\subsection{Accuracy Assessment}
+To validate Standartox results we compared the aggregated results using the geometric mean to values from other databases.
+The PPDB provides ecotoxicity data on a few selected species from chemical risk assessment that have been manually quality controlled through expert judgement \citep{lewis_international_2016}.
+The vast majority of aggregated values (91 \%) of Standartox lie within one order of magnitude of the corresponding PPDB values (n = 3601).
+This would increase to 92.7 \%, when restricting the comparison to Standartox values where data from at least five experiments were available.
+Similarly, we compared Standartox to ecotoxicity values for \textit{D.
+magna} from the ChemProp \citep{ufzdepartmentofecologicalchemistry_chemprop_2019} software, which estimates LC\textsubscript{50} values via quantitative structure-activity relationship (QSAR) models \citep{schuurmann_quantitative_2011} We found that  95 \% of Standartox values lie within one order of magnitude of the Chemprop (n = 179) vlaues.
+However, the diﬀerence is not necessarily an indication of lower quality of Standartox estimates but may also reflect the wider range of experimental conditions for which data are available in the database underlying Standartox as well as inaccurate predictions for QSAR models, respectively (Figure \ref{fig:standartox_ppdb_diff}).
 
-\section*{Acknowledgements}
-%%Text acknowledging non-author contributors. Acknowledgements should
-%%be brief, and should not include thanks to anonymous referees and
-%%editors, or effusive comments. Grant or contribution numbers may be
-%%acknowledged. Author contributions Please describe briefly the contributions
-%%of each author to this work on a separate line. 
-%%
-%%AK did this and that. 
-%%
-%%BG did this and that and the other.
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=1\linewidth]{article/figures/gg_ppdb_stan_compare_continous.png}
+    \caption{Comparison between Standartox and (\textbf{A}) PPDB and (\textbf{B}) ChemProp values.
+The black lines indicate identity and red lines mark a divergence of a factor of 10.
+Compared species are color coded.}
+    \label{fig:standartox_ppdb_diff}
+\end{figure}
 
-\section*{Competing financial interests}
-%%A competing financial interests statement is required for all accepted
-%%papers published in \emph{Scientific Data}. If none exist simply write,
-%%``The author(s) declare no competing financial interests''.
+\subsection{Perspectives}
+Novel predictive frameworks incorporating chemical mode of action and species traits emphasize the need for holistic and automated analyses of large-scale ecotoxicological data \citep{malaj_evolutionary_2016, vandenberg_modeling_2019}.
+Indeed, the increasing amount of data from ecotoxicological tests and experiments that is becoming available has elicited several initiatives to harmonize these data.
+These initiatives partly aim for overlapping goals, yet have limitations or objectives that discriminate them from Standartox:
 
-\section*{Figures Legends}
-%%Figure should be referred to using a consistent numbering scheme through
-%%the entire Data Descriptor. For initial submissions, authors may choose
-%%to supply this document as a single PDF with embedded figures, but
-%%separate figure image files must be provided for revisions and accepted
-%%manuscripts. In most cases, a Data Descriptor should not contain more
-%%than three figures, but more may be allowed when needed. We discourage
-%%the inclusion of figures in the Supplementary Information \textendash{}
-%%all key figures should be included here in the main Figure section. 
-%%
-%%Figure legends begin with a brief title sentence for the whole figure
-%%and continue with a short description of what is shown in each panel,
-%%as well as explaining any symbols used. Legend must total no more
-%%than 350 words, and may contain literature references.
+Comptox, is a web tool published by the EPA which, similar to Standartox allows for filtering test results, the retrieval of additional chemical information and the estimation of toxicity equivalents, such as 48 h \textit{Daphnia magna} LC\textsubscript{50} values.
+However, toxicity estimations are limited to standard organisms (e.g.
+\textit{Daphnia magna}), and the tool lacks the possibility for automated retrieval \citep{williams_comptox_2017}.
 
-\section*{Tables}
-%%Tables supporting the Data Descriptor. These can provide summary information
-%%(sample numbers, demographics, etc.), but they should generally not
-%%be used to present primary data (i.e. measurements). Tables containing
-%%primary data should be submitted to an appropriate data repository. 
-%%
-%%Tables may be provided within the \LaTeX{} document or as separate
-%%files (tab-delimited text or Excel files). Legends, where needed,
-%%should be included here. Generally, a Data Descriptor should have
-%%fewer than ten Tables, but more may be allowed when needed. Tables
-%%may be of any size, but only Tables which fit onto a single printed
-%%page will be included in the PDF version of the article (up to a maximum
-%%of three).
+The EnviroTox database which also uses, amongst others the ECOTOX database as an input has recently been published \citep{healthandenvironmentalsciencesinstitutehesi_envirotox_2019, connors_creation_2019}.
+In contrast to Standartox, EnviroTox  is restricted to selected aquatic organisms (i.e.
+fish, amphibians, invertebrates and algae) and experimental durations (at least 24 h) and uses a rule-based algorithm to derive single ecotoxicity values.
+Besides, EnviroTox provide additional information on toxicity endpoints, such as acute or chronic classifications and mode of action assignments.
+We think, an action that can also be allocated to the user, since there are several different ways to classify such parameters, especially those mentioned, guaranteeing user flexibility.
+The EnviroTox database allows for an aggregation into single toxicity values for individual taxa, whereas Standartox performs this aggregation for individual taxa-chemical combinations.
+However, the Standartox results for individual taxa-chemical combinations could easily be aggregated across chemicals in a second step to achieve aggregations performed in EnviroTox.
 
-%\begin{sidewaystable}
-%    \csvautotabular[respect all]{data/meta.csv};
-%    \caption{Meta variables table};
-%    \label{table:meta-variables};
-%\end{sidewaystable}
+The Etox database, like the ECOTOX database collects and provides filter methods for ecotoxicological test information.
 
-\begin{table}[h!]
-\begin{tabular}{|l|l|}
-\hline
-Script                      & Description                                      \\ \hline
-run\_build.R                 & runs all sub scripts                             \\ \hline
-setup.R                     & sets up the environment                          \\ \hline
-bd\_epa\_download.R         & download EPA Ecotox data base                    \\ \hline
-bd\_epa\_postgres.R         & build EPA Ecotox data base locally               \\ \hline
-da\_epa1.R                  & raw EPA Ecotox data export                       \\ \hline
-da\_epa2.R                  & process EPA Ecotox data                          \\ \hline
-da\_epa3.R                  & reduce EPA Ecotox data                           \\ \hline
-qu\_aw.R                    & query Compendium of Pesticide Common Names       \\ \hline
-qu\_chemspider\_scrape.R    & query Chemspider data base                       \\ \hline
-qu\_eurostat\_chem\_class.R & query Eurostat data set                          \\ \hline
-qu\_pc.R                    & query PubChem data base                          \\ \hline
-qu\_pp.R                    & query Physprop data base                         \\ \hline
-qu\_worms2.R                & query WORMS data base                            \\ \hline
-qu\_gbif.R                  & query GBIF data base                             \\ \hline
-re\_merge.R                 & merge all data                                   \\ \hline
-re\_combine.R               & combine information                              \\ \hline
-re\_analyses.R              & ???                                              \\ \hline
-re\_checks\_internal.R      & ???                                              \\ \hline
-re\_final.R                 & create and write final table                     \\ \hline
-wr\_dbase.R                 & write final table to data base                   \\ \hline
-wr\_app.R                   & write final table and variables to app directory \\ \hline
-wr\_meta.R                  & ???                                              \\ \hline
-wr\_stats.R                 & ???                                              \\ \hline
+The PPDB provides data only on pesticides and as mentioned before provides single quality controlled values only for commonly used taxa, e.g. \textit{D. magna} or \textit{R. subcapitata}.
+
+In summary, none of the above mentioned initiatives aim for a holistic and automated standardized aggregation method of exposure endpoints for individual chemicals.
+In addition, they lack the possibility to access the databases through a common high level programming language, such as R.
+As outlined above, toxicity estimates from different studies can be highly variable due to a wide range of experimental conditions such as pH, temperature or conductivity \citep{rosenkrantz_influence_2013, li_temperature_2011}.
+Integrating these factors into the aggregated estimates would improve toxicity estimates.
+However, the current implementation of Standartox does not consider these factors, because the ECOTOX database only provides sparse records on experimental conditions.
+The most frequently provided experimental conditions are temperature (77 \%), pH (56 \%), hardness (27 \%), dissolved oxygen (18 \%), Alkalinity (15 \%) and salinity (9 \%), and all others for less than 5 \%.
+A text-mining approach, where a literature reference is associated with  ecotoxicity raw data, iterating through the individual publications could potentially increase this number.
+
+\begin{table}[H]
+    \caption{Different databases that provide ecotoxicological data.
+Abbreviations: \textbf{ALL:} Most important test parameters, including chemical, taxon, duration for filtering ecotoxicological data are incorporated.
+\textbf{WEB:} Accessible via a web application through a graphical user interface.
+\textbf{API:} Accessible via an application programming interface.}
+    \label{tab:database-differences}
+    \centering
+    \begin{tabular}{m{3cm}m{3cm}m{2cm}m{2cm}m{1cm}}
+    \toprule
+    \textbf{Database} & \textbf{Publisher} & \textbf{Filter} & \textbf{Aggregation, Selection} & \textbf{Access} \\
+    \midrule
+    Comptox & Environmental Protection Agency & Chemical & no & WEB, file \\
+    Ecotox & Environmental Protection Agency & ALL & no & WEB, file \\
+    EnviroTox & Health and Environmental Sciences Institute & ALL & chemical, organism \\
+    Etox & Umweltbundesamt & ALL & no & WEB \\
+    Pesticide Property Data Base (PPDB) & University of Hertfordshire & fixed values & manual selection & WEB, file \\
+    Standartox & this article & ALL & chemical, organism & API, WEB \\
+    \bottomrule
 \end{tabular}
-\caption{Scripts that are used to process the data}
-\label{tab:scripts-pipeline}
 \end{table}
 
-\begin{table}[h!]
-\begin{tabular}{|l|l|}
-\hline
-Script                     & Description                                               \\ \hline
-setup.R                    & sets up the environment                                   \\ \hline
-ui.R                       & creates the graphical user interface of the application   \\ \hline
-server.R                   & handles functions and data input and output               \\ \hline
-fun\_filter.R              & function that filters the data                            \\ \hline
-fun\_aggregation.R         & function that performs the aggregation                    \\ \hline
-fun\_filagg\_plot\_ly.R    & creates interactive plot                                  \\ \hline
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Methods - required}
+An automated processing pipeline downloads the quarterly released ECOTOX database, performs several preparation steps on it and exports a final Standartox data set.
+This data set is accessible via a web application and an application programming interface (API).
+An API provides the means for machine communication between a host and a client and thus allows scriptable data queries.
+To facilitate the API access, the R \citep{rcoreteam_language_2017} package \textit{standartox} is built.
+All data presented in this paper are derived from the Standartox build based on the ECOTOX release from the 12.12.2019.
+
+\subsection{Processing}
+Standartox downloads the quarterly released ECOTOX database and builds it into a local PostgreSQL database \citep{szocs_build_2019}.
+Subsequently, Structured Query Language (SQL) functions for further processing the data are implemented.
+In addition lookup tables that enable duration and concentration unit conversions of the data are created.
+A meta table providing information such as the release version of the ECOTOX database is added.
+Then, provided Chemical Abstracts Service Numbers (CAS) and taxonomic names are used to query additional information from publicly available databases on chemicals and organisms, respectively.
+This includes the Compendium of Pesticide Common Names \citep{wood_compendium_2019}, the Chemical Entities of Biological Interest (ChEBI) database \citep{hastings_chebi_2016}, the Chemical Identifier Resolver (CIR) service \citep{nationalinstitutesofhealthnih_chemical_2019}, the Pubchem database \citep{kim_pubchem_2016}, Eurostat \citep{europeancommission_eurostat_2019} and Wikidata \citep{vrandecic_wikidata_2014} for chemicals and the World Register of Marine Species (WoRMS) \citep{wormseditorialboard_world_2018} and the Global Biodiversity Information Facility (GBIF) \citep{gbif_gbif_2019} (Table \ref{tab:data-base-additional}) for habitat and spatial distribution of organisms.
+Given that CAS and taxonomic names are not always unambiguous, e.g.
+the genus \textit{Eisenia} can refer to an algae and a worm at once, we first match them against the database specific identifiers and subsequently use them to retrieve the actual data.
+In a next step the data is added to Standartox to enable filtering for specific chemical roles (e.g.
+drug, pesticide, personal care product) and classes (e.g.
+pyrethroid, metal, carbamate) as well as spatial distribution (i.e.
+continents) and habitat preferences (e.g.
+freshwater) of individual taxa.
+Taxa that where not identified to at least genus level are excluded.
+Finally, the Standartox data set is compiled, which includes the harmonisation of data, e.g.
+through conversion of result concentration and test duration units.
+1237 distinct concentration units are converted to X harmonised ones, as are the 126 distinct duration units to X.
+Standartox retains only duration units, that can unambiguously be converted to hours, excluding duration units such as harvest or lifetime.
+To guarantee appropriate unit conversion and harmonisation, we compared one automated conversion to a manually done conversion for each of the distinct concentration and duration units.
+Furthermore, the units are cleaned, for example through removing additional information in the field such as \textit{food}, \textit{soil}, \textit{ai} that are also coded in other variables and hinder the processing of units.
+Concentrations that are given as rates such as per day (e.g.
+mg/kg/day) are multiplied by the days and the converted.
+Experimental endpoints are restricted to three groups, namely NOEX, LOEX and XX50.
+Other endpoints, such as Bioconcentration factors, non-half maximal effective concentrations (e.g.
+IC10, EC25, LD99) or maximum acceptable toxicant (MATC) concentrations are removed.
+Overall this reduces the ECOTOX data set from 971,430 to 611,248 harmonised test results that are included in Standartox.
+Along with that, a catalog, listing all distinct entries and value ranges, for categorical and continuous variables, respectively is created.
+Finally, we run quality control scripts that check the accuracy of the data.
+The compiled Standartox data set together with the catalog is exported and accessible via the web application and the API, through the R package.
+
+\begin{figure}
+    \centering
+    \includestandalone[scale=0.75]{article/tikz/standartox-organigram}
+    \caption{Organigram of Standartox.
+The ECOTOX knowledge base is downloaded quarterly and processed (i.e.
+query additional information with CAS and Taxa name and cleaning concentration and duration units.
+Subsequently a Standartox data set is compiled together with filter and aggregation methods.
+Thus, users can access the Standartox data set and filter and aggregate through a web application and an R-package.}
+    \label{fig:stx-organigram}
+\end{figure}
+
+\subsection{Application methods}
+The web application and the API load the compressed serialized Standartox data into memory and allow a client to interact with it.
+The client thereby calls the functions stx\_filter() and stx\_aggregate() that filter and aggregate the data according to specific parameters (Table \ref{tab:app-parameters}).
+The interactive web application is built in R using the shiny framework, which runs with the help of a shiny server \citep{chang_shiny_2018} on \url{standartox.uni-landau.de}.
+The API is built by using the R package \textit{plumber} \citep{trestletechnologyllc_plumber_2018}, which allows for the creation of Representational State Transfer (REST) APIs from R.
+REST is a software standard that defines web service communication rules.
+The API is reachable via the URL \url{http://139.14.20.252} and port 8000.
+Four API-endpoints (\textit{/catalog}, \textit{/filter}, \textit{/aggregate} and \textit{/meta}) can be queried by users.
+The \textit{/catalog} API-endpoint returns a JavaScript Object Notation (JSON) file containing a catalog of possible filter parameters to choose from.
+The \textit{/filter} returns the filtered Standartox table as a compressed serialized binary file created by the R \textit{fst} package \citep{klik_fst_2019}, to reduce size and allow for fast user queries.
+The \textit{/aggregate} API-endpoint returns an R function allowing the client to aggregate the filtered data.
+Lastly, the \textit{/meta} API-endpoint returns a JSON file with meta information, such as the timestamp of the request and the used Standartox version.
+The API is designed to be used with the R-package \textit{standartox} and therefore uses serialization methods specific to R (rds() from the R package \pkg{base} and fst() from the package fst).
+In order to facilitate the API usage the R-package \textit{standartox} is created.
+
+\begin{table}[H]
+    \caption{Input parameters for the Standartox web application and the R-package standartox (CAS - Chemical Abstracts Service Registry Number, NOEX and XX50 - Standartox endpoints \ref{tab:endpoints-conflate}), vers - Standartox version.}
+    \label{tab:app-parameters}
+    \centering
+    \begin{tabular}{cc}
+    \toprule
+    \textbf{parameter} & \textbf{example} \\ 
+    \midrule
+    cas & 7758987, 2921-88-2, 1912-24-9 \\
+    concentration\_type & Active ingredient, Formulation \\
+    chemical\_role & Antibiotic, Fungicide, Drug \\
+    chemical\_class & Conazole, Neonicotinoid, Triazine \\
+    taxa & Oncorhynchus mykiss, Rattus norvegicus, Daphnia magna \\
+    habitat & Marine, Brackish, Freshwater \\
+    region & Europe, Africa, Asia \\
+    duration & 24, 96 \\
+    effect & Mortality, Population, Growth \\
+    endpoint & NOEX, XX50 \\
+    vers & 20191212 \\
+    \bottomrule
 \end{tabular}
-\caption{Scripts that used for the web application}
-\label{tab:scripts-app}
 \end{table}
 
-\pagebreak
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{User Notes - optional}
+% \input{article/parts/04_user_notes}
+By accessing the web application users can filter and download the resulting data sets as a comma-separated values (csv) file.
+Users of the R-package can directly load the data in R.
+The R-package provides the two functions \textit{stx\_catalog()} and \textit{stx\_query()}.
+The first command queries a catalog of possible Standartox parameters into a R list object.
+The latter allows users to set the Standartox filter parameters and to fetch the actual data.
+It returns a R list of three tables (i.e.
+R data.frames) containing the filtered data set, the aggregated data set and a table with the meta information retrieved from the API endpoints \textit{/filter} and \textit{/meta}.
+A short R-code example is given below \ref{lis:standartox-example} and a detailed description on the usage of the R package is provided on its Github page: \url{https://github.com/andschar/standartox}.
 
-\bibliographystyle{apalike}
-\bibliography{refs/references-etox-base.bib,refs/references-etox-base-rpackages.bib,refs/references-etox-base-rshinypackages.bib}
-% multiple refs have to be written w/o space
+\begin{lstlisting}[
+    caption = {Sample code to access the Standartox data base through the API with the help of the R-package.
+\textit{stx\_catalog()} returns a catalog of possible filter and aggregation parameters.
+\textit{stx\_query()} returns the Standartox object, a list of the filtered and the aggregated data as well as a meta data entry.},
+    label = {lis:standartox-example}]
+# install
+install.packages('remotes')
+remotes::install_github('andschar/standartox') # package not yet on CRAN
+# retrieve catalog    
+require(standartox)
+catal = stx_catalog()
+# retrieve data
+l = stx_query(cas = '1071-83-6',
+                endpoint = 'XX50',
+                taxa = 'Oncorhynchus',
+                duration = c(24, 120))
+\end{lstlisting}
 
-\section*{Data Citations}
-
-\pagebreak
-\section*{TODO}
+Standartox's code is distributed on three Github repositories below.
+All code can be freely accessed under the MIT License <CITATION>.
+Most of the code is written in R 3.6.1 and associated packages (compare Supplement: \ref{list:r-packages}) as well as some parts in SQL for PostgreSQL 9.6.1.
 
@@
 
-\item compare Etox-BAse output with PPDB
+\item \url{https://github.com/andschar/standartox-build} \newline
+Contains the code for the data acquisition and processing pipeline.
 
-\item plot example SSD and TU
+\item \url{https://github.com/andschar/standartox-app} \newline
+Contains the code for the web application and the API.
 
-\item acute vs chronic
-
-\item OECD guidlines are not machine readable
-
-\item discuss flaws in column types (e.g some concentration entries contain '+' or '~' and are therefore not convertable to numeric) - This hampers large scale analysis greatly!
-
-\pagebreak
-\section*{MISC}
-
-\begin{figure}
-    \includestandalone[scale=0.5]{tikz/tree_source}
-    \caption{Alternative processing pipeline}
-    \label{fig:pipeline-tree}
-\end{figure}
-
-\section*{not included citations}
-
-\item Put in citation \citep{hartung_chemical_2009} ?
-which claims that REACH won't meet their assumptions and say 54 million vertebrate animals are needed for tests that would cost €9.5 billion over the next ten years.
-I.e 20 times more animals, 6 times the costs in comparison to the official estimates
-
-\item "comparison of rate of change in the production and variety of pesticides, pharmaceuticals and other chemicals over the past four decades to CO2 nutrient pollution and other global change factors" \citep{bernhardt_synthetic_2017}
-
-\item worldwide decline in entomofauna: \citep{sanchez-bayo_worldwide_2019}
+\item \url{https://github.com/andschar/standartox} \newline
+Contains the code for the R-package.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusion - optional}
+% This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.
+Due to the steady incorporation of new ecotoxicity data, its values produced by Standartox will be subject to change and most certainly produce slightly different results in its aggregations in the future.
+However, we regard this as an advantage, since other published works that aim in similar directions often constitute a singular effort or require manual work for each update.
+Standartox, in contrast automates these processes and as already described would provide means to access older versions anyhow, to assure reproducibility.
+In comparison to rule-based approaches, such as the PPDB for the derivation of single ecotoxicity values, Standartox has the advantage to not be prone to the subjectivity, the exhaustiveness and the incompleteness of a human induced set of rules.
+Above all, Standartox provides quick access through its design to be queried via the R language.
+In general, Standartox is designed to foster chemical risk assessment (CRA).
+Due to an increased amount of available ecotoxicological test data, it becomes fundamental to provide and distribute ecotoxicity information in adequate formats, both easily accessible for humans and easily processable for machines.
+Standartox is a tool that does this in a generalizing way and puts its focus on the aggregation of toxicity data, thereby adding a piece to the puzzle of modern ecotoxicological analyses.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vspace{6pt}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% optional
+%\supplementary{The following are available online at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title.}
+
+% Only for the journal Methods and Protocols:
+% If you wish to submit a video article, please do so with any other supplementary material.
+% \supplementary{The following are available at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title. A supporting video article is available at doi: link.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\authorcontributions{conceptualization, A.S.
+and V.C.S.; methodology, A.S.; software, A.S.; validation, A.S.; formal analysis, A.S.; investigation, A.S.; resources, A.S.; data curation, A.S.; writing--original draft preparation, A.S.; writing--review and editing, A.S., V.C.S.
+and R.B.S.; visualization, A.S.; supervision, R.B.S.; project administration, A.S.
+and R.B.S.; funding acquisition, Y.Y.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''.
+Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections.
+This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\conflictsofinterest{The authors declare no conflict of interest.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\abbreviations{The following abbreviations are used in this manuscript:\\
+\noindent 
+\begin{tabular}{@{}ll}
+API & Application Programming Interface \\
+CAS & Chemical Abstracts Service Registry Number \\
+ChEBI & Chemical Entities of Biological Interest database \\
+CIR & Chemical Identifier Resolver service \\
+E/LC/D\textsubscript{50} &  half maximal effective/lethal concentration/dose \\
+ECOTOX & US EPA Ecotox Knowledge Base \\
+GBIF & Global Biodiversity193Information Facility \\
+JSON & JavaScript Object Notation file format \\
+LOEC/L & Lowest Observed Effect Levels/Concentration \\
+NOEC/L & No Observed Effect Levels/Concentration \\
+PPDB & Pesticides Property Data Base \\
+QSAR & Quantitative structure–activity relationship \\
+REST & Representational State Transfer software architectural style \\
+SSD & Species Sensitivity Distribuion \\
+TU & Toxic Unit \\
+WoRMS & World Register of Marine Species \\
+\end{tabular}
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% optional
+\appendixtitles{yes} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
+\appendix
+\section{}
+\begin{table}[H]
+    \caption{Table of additionally queried publicly available data bases and their URLs.}
+    \label{tab:data-base-additional}
+    \centering
+\begin{tabular}{cc}
+    \toprule
+    \textbf{Data Base} & \textbf{URL} \\ 
+    \midrule
+    \makecell{Chemical Entities of \\ Biological Interest (ChEBI)} & \url{https://www.ebi.ac.uk/chebi}  \\
+    Chemical Identifier Resolver & \url{https://cactus.nci.nih.gov/chemical/structure}  \\[0.5cm]
+    ChemSpider & \url{http://www.chemspider.com}    \\[0.5cm]
+    Eurostat & \url{https://ec.europa.eu/eurostat/home} \\[0.5cm]
+    PubChem & \url{https://pubchem.ncbi.nlm.nih.gov} \\[0.5cm]
+    Wikidata & \url{https://www.wikidata.org/wiki/Wikidata:Main_Page} \\[0.5cm]
+    \makecell{Global Biodiversity \\ Information Facility (GBIF)} & \url{https://www.gbif.org} \\[0.5cm]
+    \makecell{World Register of \\ Marine Species (WoRMS)} & \url{http://marinespecies.org} \\
+    \bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}[H]
+    \caption{Table of how Standartox endpoints are derived from EPA ECOTOX endpoints.}
+    \label{tab:endpoints-conflate}
+    \centering
+\begin{tabular}{ccc}
+    \toprule
+    \textbf{Standartox endpoint} & \textbf{Ecotox endpoint} & \textbf{Ecotox endpoint description} \\
+    \midrule
+    XX50 & LC50 & Lethal concentration to 50\% of test organisms \\
+    XX50 & LD50 & Lethal dose to 50\% of test organisms \\
+    XX50 & EC50 & Effective concentration to 50\% of test organisms \\
+    XX50 & ED50 & Effective dose to 50\% of test organisms \\
+    XX50 & IC50 & Inhibition concentration to 50\% of test organisms \\
+    XX50 & ID50 & Inhibition dose to 50\% of test organisms \\
+    XX50 & ET50 & Effective response time to 50\% of test organisms \\
+    XX50 & LT50 & Time to 50\% mortality of organisms \\
+    NOEX & NOEC & No-observable-effect-concentration \\
+    NOEX & NOEL & No-observable-effect-level \\
+    LOEX & LOEC & Lowest observable effect concentration\\
+    LOEX & LOEL & Lowest-observable-effect-level \\
+    \bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}[H]
+    \caption{Application programming interface (API) endpoints, Http methods,
+    Requests and Response objects. Object is abbreviated with obj.}
+    \label{tab:api-endpoints}
+    \centering
+\begin{tabular}{cccc}
+    \toprule
+    \textbf{Endpoint} & \textbf{Http Method} & \textbf{Request} & \textbf{Response} \\ 
+    \midrule
+    /catalog & POST & \makecell{Standartox \\ version string} & \makecell{Catalog object of \\ possible filter parameters \\ JSON obj.}    \\[0.5cm]
+    /filter & POST & \makecell{Standartox filter \\ parameters (Table \ref{tab:app-parameters})} & \makecell{Filtered Standartox data \\ serialized R data.frame \\ R fst:: \citep{klik_fst_2019} obj.}   \\[0.5cm]
+    /aggregate & GET & / & \makecell{Aggregate function \\ serialized R function \\ R .rds \citep{rcoreteam_language_2017} obj.}    \\[0.5cm]
+    /meta & POST & \makecell{Standartox \\ version string} & \makecell{Meta data on request \\ JSON obj.}    \\
+    \bottomrule
+\end{tabular}
+\end{table}
+
+\section{}
+
+Rinker TW, Kurkiewicz D (2018).
+\emph{pacman: Package Management for R}.
+version 0.5.0, \url{http://github.com/trinker/pacman}.
+\newline R Core Team (2019).
+\emph{R: A Language and Environment for Statistical Computing}.
+R Foundation for Statistical Computing, Vienna, Austria.
+\url{https://www.R-project.org/}.
+\newline Wickham H, Hester J, Chang W (2019).
+\emph{devtools: Tools to Make Developing R Packages Easier}.
+R package version 2.2.1, \url{https://CRAN.R-project.org/package=devtools}.
+\newline Lang DT, team tC (2019).
+\emph{RCurl: General Network (HTTP/FTP/...) Client Interface for R}.
+R package version 1.95-4.12, \url{https://CRAN.R-project.org/package=RCurl}.
+\newline Wickham H (2019).
+\emph{stringr: Simple, Consistent Wrappers for Common String Operations}.
+R package version 1.4.0, \url{https://CRAN.R-project.org/package=stringr}.
+\newline Bengtsson H (2019).
+\emph{R.utils: Various Programming Utilities}.
+R package version 2.9.0, \url{https://CRAN.R-project.org/package=R.utils}.
+\newline Hiebert J (2016).
+\emph{udunits2: Udunits-2 Bindings for R}.
+R package version 0.13, \url{https://CRAN.R-project.org/package=udunits2}.
+\newline Wickham H (2019).
+\emph{rvest: Easily Harvest (Scrape) Web Pages}.
+R package version 0.3.4, \url{https://CRAN.R-project.org/package=rvest}.
+\newline Wickham H (2019).
+\emph{httr: Tools for Working with URLs and HTTP}.
+R package version 1.4.1, \url{https://CRAN.R-project.org/package=httr}.
+\newline Ooms J (2014).
+``The jsonlite Package: A Practical and Consistent Mapping Between JSON Data and R Objects.''
+\emph{arXiv:1403.2805 [stat.CO]}.
+\url{https://arxiv.org/abs/1403.2805}.
+\newline Wickham H, Bryan J (2019).
+\emph{readxl: Read Excel Files}.
+R package version 1.3.1, \url{https://CRAN.R-project.org/package=readxl}.
+\newline Walker A (2019).
+\emph{openxlsx: Read, Write and Edit XLSX Files}.
+R package version 4.1.0.1, \url{https://CRAN.R-project.org/package=openxlsx}.
+\newline Henry L, Wickham H (2019).
+\emph{purrr: Functional Programming Tools}.
+R package version 0.3.2, \url{https://CRAN.R-project.org/package=purrr}.
+\newline Dowle M, Srinivasan A (2019).
+\emph{data.table: Extension of `data.frame`}.
+R package version 1.12.4, \url{https://CRAN.R-project.org/package=data.table}.
+\newline Conway J, Eddelbuettel D, Nishiyama T, Prayaga SK, Tiffin N (2017).
+\emph{RPostgreSQL: R Interface to the 'PostgreSQL' Database System}.
+R package version 0.6-2, \url{https://CRAN.R-project.org/package=RPostgreSQL}.
+\newline R Special Interest Group on Databases (R-SIG-DB), Wickham H, Müller K (2018).
+\emph{DBI: R Database Interface}.
+R package version 1.0.0, \url{https://CRAN.R-project.org/package=DBI}.
+\newline Oksanen J, Blanchet FG, Friendly M, Kindt R, Legendre P, McGlinn D, Minchin PR, O'Hara RB, Simpson GL, Solymos P, Stevens MHH, Szoecs E, Wagner H (2019).
+\emph{vegan: Community Ecology Package}.
+R package version 2.5-5, \url{https://CRAN.R-project.org/package=vegan}.
+\newline Wickham H (2011).
+``The Split-Apply-Combine Strategy for Data Analysis.''
+\emph{Journal of Statistical Software}, \bold{40}(1), 1--29.
+\url{http://www.jstatsoft.org/v40/i01/}.
+\newline Komsta L (2011).
+\emph{outliers: Tests for outliers}.
+R package version 0.14, \url{https://CRAN.R-project.org/package=outliers}.
+\newline Wickham H (2019).
+\emph{feather: R Bindings to the Feather 'API'}.
+R package version 0.3.3, \url{https://CRAN.R-project.org/package=feather}.
+\newline Wickham H (2016).
+\emph{ggplot2: Elegant Graphics for Data Analysis}.
+Springer-Verlag New York.
+ISBN 978-3-319-24277-4, \url{https://ggplot2.tidyverse.org}.
+\newline Slowikowski K (2019).
+\emph{ggrepel: Automatically Position Non-Overlapping Text Labels with
+'ggplot2'}.
+R package version 0.8.1, \url{https://CRAN.R-project.org/package=ggrepel}.
+\newline Wilke C (2019).
+\emph{cowplot: Streamlined Plot Theme and Plot Annotations for 'ggplot2'}.
+R package version 0.9.4, \url{https://CRAN.R-project.org/package=cowplot}.
+\newline Neuwirth E (2014).
+\emph{RColorBrewer: ColorBrewer Palettes}.
+R package version 1.1-2, \url{https://CRAN.R-project.org/package=RColorBrewer}.
+\newline Tennekes M (2017).
+\emph{treemap: Treemap Visualization}.
+R package version 2.4-2, \url{https://CRAN.R-project.org/package=treemap}.
+\newline Chamberlain S, Barve V, Mcglinn D, Oldoni D, Desmet P, Geffert L, Ram K (2019).
+\emph{rgbif: Interface to the Global Biodiversity Information Facility API}.
+R package version 1.3.0, \url{https://CRAN.R-project.org/package=rgbif}.
+
+Chamberlain S, Boettiger C (2017).
+``R Python, and Ruby clients for GBIF species occurrence data.''
+\emph{PeerJ PrePrints}.
+\url{https://doi.org/10.7287/peerj.preprints.3304v1}.
+\newline Scott Chamberlain, Eduard Szocs (2013).
+``taxize - taxonomic search and retrieval in R.''
+\emph{F1000Research}.
+\url{http://f1000research.com/articles/2-191/v2}.
+
+Chamberlain S, Szoecs E, Foster Z, Arendsee Z, Boettiger C, Ram K, Bartomeus I, Baumgartner J, O'Donnell J, Oksanen J, Tzovaras BG, Marchand P, Tran V, Salmon M, Li G, Grenié M (2019).
+\emph{taxize: Taxonomic information from around the web}.
+R package version 0.9.7, \url{https://github.com/ropensci/taxize}.
+\newline Arel-Bundock V, Enevoldsen N, Yetman C (2018).
+``countrycode: An R package to convert country names and country codes.''
+\emph{Journal of Open Source Software}, \bold{3}(28), 848.
+\url{https://doi.org/10.21105/joss.00848}.
+\newline Microsoft, Weston S (2017).
+\emph{foreach: Provides Foreach Looping Construct for R}.
+R package version 1.4.4, \url{https://CRAN.R-project.org/package=foreach}.
+\newline Corporation M, Weston S (2018).
+\emph{doParallel: Foreach Parallel Adaptor for the 'parallel' Package}.
+R package version 1.0.14, \url{https://CRAN.R-project.org/package=doParallel}.
+\newline Klik M (2019).
+\emph{fst: Lightning Fast Serialization of Data Frames for R}.
+R package version 0.9.0, \url{https://CRAN.R-project.org/package=fst}.
+\newline Xie Y, Cheng J, Tan X (2019).
+\emph{DT: A Wrapper of the JavaScript Library 'DataTables'}.
+R package version 0.7, \url{https://CRAN.R-project.org/package=DT}.
+\newline Tennekes M (2017).
+\emph{treemap: Treemap Visualization}.
+R package version 2.4-2, \url{https://CRAN.R-project.org/package=treemap}.
+\newline
+
+% \unskip
+% \subsection{}
+% The appendix is an optional section that can contain details and data supplemental to the main text.
+For example, explanations of experimental details that would disrupt the flow of the main text, but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data is shown in the main text can be added here if brief, or as Supplementary data.
+Mathematical proofs of results not central to the paper can be added as an appendix.
+
+% \section{}
+% All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.
+
+%=====================================
+% References, variant A: internal bibliography
+%=====================================
+% \begin{thebibliography}{999}
+% Reference 1
+% \bibitem[Author1(year)]{ref-journal}
+% Author1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
+% Reference 2
+% \bibitem[Author2(year)]{ref-book}
+% Author2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
+% \end{thebibliography}
+
+% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences.
+For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
+% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
+% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p.
+328; \citeyear{ref-journal-3b}, p.475).
+This produces: Wong (1999, p.
+328; 2000, p.
+475)
+
+%=====================================
+% References, variant B: external bibliography
+%=====================================
+\reftitle{References}
+\externalbibliography{yes}
+\bibliography{refs/references-standartox.bib}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% optional
+\sampleavailability{Samples of the compounds ...... are available from the authors.}
+
+%% for journal Sci
+%\reviewreports{\\
+%Reviewer 1 comments and authors’ response\\
+%Reviewer 2 comments and authors’ response\\
+%Reviewer 3 comments and authors’ response
+%}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \end{document}

### 2020-03-14 article/article_misc.tex
cleaned article folder

--- 2020-01-27 article/parts/11_misc.tex
@@
+%%%%%%%%%%%%% REMOVED PERSPECTIVE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% deleted by Ralf
+For example, they restrict the ecotoxicity data to freshwater fish, amphibian, invertebrate and algae taxa and test durations of at least 24 hours only.
+In contrast, Standartox doesn't refine to specific test durations, but leaves it up to the user to decide on such parameters.
+% no raw ecotox data bases
+\citet{petschick_modeling_2019} modeled risk threshold level equivalents for aquatic organisms by using ecotoxicological effect data from the ECOTOX database to support regulation.
+Along with newly created ecotoxicological databases, methods of how to efficiently store ecotoxicological data are also proposed as can be seen in the MAGIC Knowledge Base \citep{bub_graphing_2019}.
+Likewise, ideas to create new predictive frameworks in ecotoxicology, incorporating chemical mode of actions and species traits \citep{vandenberg_modeling_2019} together with the recent increase in efforts to compile ecotoxicological test databases clearly emphasizes the need for holistic and automated analyses of large-scale ecotoxicological data.
+
 %%%%%%%%%%%%% OLD SUMMARY TEXT %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-03-14 article/template_mdpi.tex
cleaned article folder

--- 2020-03-13 article/00_article2.tex
@@
 % Choose between the following MDPI journals:
-% acoustics, actuators, addictions, admsci, aerospace, agriculture, agriengineering, agronomy, ai, algorithms, animals, antibiotics, antibodies, antioxidants, applmech, applsci, arts, asc, asi, atmosphere, atoms, axioms, batteries, bdcc, behavsci , beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsci , buildings, cancers, carbon , catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, children, civileng, cleantechnol, climate, clockssleep, cmd, coatings, colloids, computation, computers, condensedmatter, cosmetics, cryptography, crystals, dairy, data, dentistry, designs , diagnostics, diseases, diversity, drones, econometrics, economies, education, ejbc, ejihpe, electrochem, electronics, endocrines, energies, entropy, environments, epigenomes, est, fermentation, fibers, fire, fishes, fluids, foods, forecasting, forests, fractalfract, futureinternet, futurephys, galaxies, games, gastrointestdisord, gels, genealogy, genes, geohazards, geosciences, geriatrics, hazardousmatters, healthcare, hearts, heritage, highthroughput, horticulturae, humanities, hydrology, ijerph, ijfs, ijgi, ijms, ijtpp, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jcdd, jce, jcm, jcp, jcs, jdb, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmse, jne, jnt, jof, joitmc, jpm, jrfm, jsan, land, languages, laws, life, literature, logistics, lubricants, machines, magnetochemistry, make, marinedrugs, materials, mathematics, mca, medicina, medicines, medsci, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, modelling, molbank, molecules, mps, mti, nanomaterials, ncrna, ijns, neurosci, neuroglia, nitrogen, notspecified, nutrients, oceans, ohbm, optics, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, physics, plants, plasma, pollutants, polymers, polysaccharides, preprints , proceedings, processes, prosthesis, proteomes, psych, publications, quantumrep, quaternary, qubs, reactions, recycling, religions, remotesensing, reprodmed, reports, resources, risks, robotics, safety, sci, scipharm, sensors, separations, sexes, signals, sinusitis, smartcities, sna, societies, socsci, soilsystems, sports, standards, stats, surfaces, surgeries, sustainability, sustainableworld, symmetry, systems, technologies, telecom, test, tourismhosp, toxics, toxins, transplantology, tropicalmed, universe, urbansci, vaccines, vehicles, vetsci, vibration, viruses, vision, water, wem, wevj
+% acoustics, actuators, addictions, admsci, aerospace, agriculture, agriengineering, agronomy, algorithms, animals, antibiotics, antibodies, antioxidants, applsci, arts, asc, asi, atmosphere, atoms, axioms, batteries, bdcc, behavsci , beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsci , buildings, cancers, carbon , catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, children, cleantechnol, climate, clockssleep, cmd, coatings, colloids, computation, computers, condensedmatter, cosmetics, cryptography, crystals, dairy, data, dentistry, designs , diagnostics, diseases, diversity, drones, econometrics, economies, education, ejihpe, electrochem, electronics, energies, entropy, environments, epigenomes, est, fermentation, fibers, fire, fishes, fluids, foods, forecasting, forests, fractalfract, futureinternet, futurephys, galaxies, games, gastrointestdisord, gels, genealogy, genes, geohazards, geosciences, geriatrics, hazardousmatters, healthcare, heritage, highthroughput, horticulturae, humanities, hydrology, ijerph, ijfs, ijgi, ijms, ijns, ijtpp, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jcdd, jcm, jcp, jcs, jdb, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmse, jnt, jof, joitmc, jpm, jrfm, jsan, land, languages, laws, life, literature, logistics, lubricants, machines, magnetochemistry, make, marinedrugs, materials, mathematics, mca, medicina, medicines, medsci, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, modelling, molbank, molecules, mps, mti, nanomaterials, ncrna, neuroglia, nitrogen, notspecified, nutrients, ohbm, optics, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, physics, plants, plasma, polymers, polysaccharides, preprints , proceedings, processes, proteomes, psych, publications, quantumrep, quaternary, qubs, reactions, recycling, religions, remotesensing, reports, resources, risks, robotics, safety, sci, scipharm, sensors, separations, sexes, signals, sinusitis, smartcities, sna, societies, socsci, soilsystems, sports, standards, stats, surfaces, surgeries, sustainability, symmetry, systems, technologies, test, toxics, toxins, tropicalmed, universe, urbansci, vaccines, vehicles, vetsci, vibration, viruses, vision, water, wem, wevj
 
@@
 \articlenumber{5}
-\pubyear{2020}
-\copyrightyear{2020}
+\pubyear{2019}
+\copyrightyear{2019}
 %\externaleditor{Academic Editor: name}
@@
 % Add packages and commands here.
-The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx,epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption
+The following packages are loaded in our class file: fontenc, calc, indentfirst, fancyhdr, graphicx, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, amsthm, hyphenat, natbib, hyperref, footmisc, geometry, caption, url, mdframed, tabto, soul, multirow, microtype, tikz
 
@@
 
-%\begin{listing}[H]
-%\caption{Title of the listing}
-%\rule{\textwidth}{1pt}
-%\raggedright Text of the listing. In font size footnotesize, small, or normalsize. Preferred format: left aligned and single spaced. Preferred border format: top border line and bottom border line.
-%\rule{\textwidth}{1pt}
-%\end{listing}
+\begin{listing}[H]
+\caption{Title of the listing}
+\rule{\textwidth}{1pt}
+\raggedright Text of the listing. In font size footnotesize, small, or normalsize. Preferred format: left aligned and single spaced. Preferred border format: top border line and bottom border line.
+\rule{\textwidth}{1pt}
+\end{listing}
 
@@
 \authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
-The following statements should be used ``Conceptualization, X.X.
+The following statements should be used ``conceptualization, X.X.
 and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y.
-and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing--original draft preparation, X.X.; writing--review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y.
-All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation.
+and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing--original draft preparation, X.X.; writing--review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation.
 Authorship must be limited to those who have contributed substantially to the work reported.}
@@
 %% optional
-\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
+\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
 \appendix
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\reftitle{References}
-
-% Please provide either the correct journal abbreviation (e.g.
-according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
 % Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.
@@
 %=====================================
-% References, variant A: external bibliography
+% References, variant A: internal bibliography
 %=====================================
-%\externalbibliography{yes}
-%\bibliography{your_external_BibTeX_file}
-
-%=====================================
-% References, variant B: internal bibliography
-%=====================================
+\reftitle{References}
 \begin{thebibliography}{999}
@@
 
+%=====================================
+% References, variant B: external bibliography
+%=====================================
+%\externalbibliography{yes}
+%\bibliography{your_external_BibTeX_file}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-03-14 article/tikz/standartox-organigram.tex
cleaned article folder

\documentclass[border = 1.5cm]{standalone}
\usepackage{tikzpeople}
\usetikzlibrary{trees,
                arrows,
                shapes.geometric,
                positioning,
                calc,
                backgrounds,
                fit}
%% document
\begin{document}
%% layers
\pgfdeclarelayer{bg1}    % declare background layer
\pgfsetlayers{bg1, main}  % set the order of the layers (main is the standard layer)
%% styles
\tikzset{
	font = {\sffamily},
	ecotox/.style = {rectangle, minimum width = 4cm, minimum height = 3cm, text centered, rounded corners, draw = black, fill = yellow!10},
	standartox/.style = {rectangle, minimum width = 4cm, minimum height = 2cm, text centered, rounded corners, draw = black, fill = purple!10},
	app/.style = {rectangle, minimum width = 4cm, minimum height = 2cm, text centered, rounded corners, draw = black, fill = red!10},
	frame/.style = {draw, rectangle, rounded corners, dashed, blue, line width = 1pt, inner sep = 4mm},
	data/.style = {rectangle, minimum width = 7cm, minimum height = 3cm, text centered, rounded corners, draw = black},
	ids/.style = {diamond, minimum width = 3cm, minimum height = 2cm, text centered, draw = black, fill = green!10},
	db/.style = {rectangle, text centered, draw = black, rounded corners, minimum width = 3cm, minimum height = 1.25cm},
	arrow1/.style = {->, > = stealth, draw = black, color = gray!80, line width = 3pt}
}
\begin{tikzpicture}
%% processing
\node (epa-ecotox) [ecotox, above of = ecotox-id, align=center, xshift = -3cm, yshift = 4cm] {\Large EPA ECOTOX\\\Large raw data};
\node (ecotox-id) [standartox, align = center] at (-3cm,0cm) {\Large Identifiers};
\node (ecotox-unit) [standartox, below of = ecotox-id, align = center, yshift = -2cm] {\Large Unit\\\Large harmonisation};
\node (ecotox-chck) [standartox, below of = ecotox-unit, align = center, yshift = -2cm] {\Large Check\\\Large Conversions};
\node (cas) [ids, below of = ecotox-id, xshift = 6cm, yshift = 3cm] {\Large CAS};
\node (taxa) [ids, below of = cas, xshift = 0cm, yshift = -2.5cm] {\Large Taxa};
\node (db-chemical) [db, right of = cas, xshift = 4cm, align = center] {\Large Chemical\\\Large databases};
\node (db-habitat) [db, right of = taxa, xshift = 4cm, yshift = 1cm, align = center] {\Large Habitat\\\Large databases};
\node (db-region) [db, right of = taxa, xshift = 4cm, yshift = -1cm, align = center] {\Large Region\\\Large databases};
%% standartox
\node (stx-data) [standartox, below of = ecotox-chck, xshift = 3cm, yshift = -2.5cm, align = center] {\Large Standartox\\\Large data set};
\node (stx-filter) [standartox, below of = stx-data, xshift = 0cm, yshift = -2cm, align = center] {\Large Filter\\\Large methods};
\node (stx-agg) [standartox, below of = stx-filter, xshift = 0cm, yshift = -2cm, align = center] {\Large Aggregation\\\Large methods};
\node (stx-app) [app, below of = stx-data, yshift = 1cm, xshift = 6cm, align = center] {\Large Web\\\Large application};
\node (stx-rpackage) [app, below of = stx-data, yshift = -4cm, xshift = 6cm, minimum height = 5cm] {\Large R-package};
\node (stx-api) [app, below of = stx-rpackage, yshift = 2.25cm, xshift = 0cm, minimum width = 3cm, minimum height = 1.5cm] {\Large API};
%% arrows
\begin{pgfonlayer}{bg1}
    \draw [arrow1] (epa-ecotox.south) to [out=75,in=180] (stx-data.west);
    \draw [arrow1] (stx-data) -- (stx-filter);
    \draw [arrow1] (stx-filter) -- (stx-agg);
    \draw [arrow1] (ecotox-id.east) -- (cas.west);
    \draw [arrow1] (ecotox-id.east) -- (taxa.west);
    \draw [arrow1] (cas.east)-- (db-chemical.west);
    \draw [arrow1] (taxa.east) -- (db-habitat.west);
    \draw [arrow1] (taxa.east) -- (db-region.west);
\end{pgfonlayer}
%% frames
\node (frame-stx) [frame, fit = (stx-data) (stx-rpackage) (stx-app) ] {};
\node (frame-stx-text) [above, blue, inner sep = 2mm] at (frame-stx.north) {\Large Standartox};
%% people
\node (person) [bob, right of = stx-data, xshift = 9cm, yshift = -8cm, minimum size = 2cm] {};
\end{tikzpicture}
\end{document}

### 2020-04-03 article/article.tex

@@
 \usepackage{listings}
+\lstset{language=R,
+    basicstyle=\small\ttfamily,
+    stringstyle=\color{DarkGreen},
+    otherkeywords={0,1,2,3,4,5,6,7,8,9},
+    morekeywords={TRUE,FALSE},
+    deletekeywords={data,frame,length,as,character},
+    keywordstyle=\color{blue},
+    commentstyle=\color{DarkGreen},
+}
 % tikz
@@
 In the environment, chemicals can adversely affect populations and communities and in turn related ecosystem functions.
-To evaluate the risks from chemicals for ecosystems data on their toxicity, which is typically produced in standardized ecotoxicological laboratory tests, is required.
-However, for many chemicals, multiple ecotoxicity values are available for the same test organism, which can vary strongly.
-Current (Meta-) databases, including the  United States Environmental Protection Agency ECOTOXicology knowledgebase only collect ecotoxicological test data, without allowing for further aggregation steps.
-Others, such as the  Pesticide Property Data Base (PPDB) are confined to specific chemicals, here pesticides, thereby ignoring vast amounts of ecotoxicity data.
-We therefore developed Standartox, a tool and database that aims to overcome the limitations of other databases by continuously incorporating the ever-growing number of test results in an automated process workflow that ultimately leads to a single aggregated data point for a specific chemical-organism test combination, representing the toxicity of a chemical.
+To evaluate the risks from chemicals for ecosystems, data on their toxicity, which is typically produced in standardized ecotoxicological laboratory tests, is required.
+The results from ecotoxicological tests are compiled in (Meta-)databases such as the  United States Environmental Protection Agency ECOTOXicology knowledgebase.
+However, for many chemicals, multiple ecotoxicity data are available for the same test organism.
+These  can vary strongly, thereby causing uncertainty of related analyses.
+Given that most current databases lack aggregation steps or are confined to specific chemicals, we developed Standartox, a tool and database that continuously incorporates the ever-growing number of test results in an automated process workflow that ultimately leads to a single aggregated data point for a specific chemical-organism test combination, representing the toxicity of a chemical.
 Standartox comes with two front-ends, a web application (standartox.uni-landau.de) and an R package (standartox).}
@@
 % Keywords
-\keyword{ecotoxicology; standardisation; environmental risk assessment; effect database; database; software; r; api;} % up to 10
+\keyword{ecotoxicology; standardisation; environmental risk assessment; effect database; database; software; R; api;} % up to 10
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Summary - required}
+\section{Summary}
 % The introduction should briefly place the study in a broad context and highlight why it is important.
@@
 \citet{bernhardt_synthetic_2017} argue that the knowledge gap how chemicals affect populations, communities and in turn ecosystem functions and services, may impede the accomplishment of the Sustainable Development Goals \citep{rosa_transforming_2017} of the United Nations.
-Even highly regulated chemical compounds such as pesticides have been shown to cause strong adverse effects on non-target organisms, such as birds \citep{hallmann_declines_2014} aquatic insects \citep{beketov_pesticides_2013} or fish \citep{yamamuro_neonicotinoids_2019}, questioning the current regulation efforts \citep{schafer_future_2019}.
+Even highly regulated chemical compounds such as pesticides have been shown to cause strong adverse effects on non-target organisms, such as birds \citep{hallmann_declines_2014}, aquatic insects \citep{beketov_pesticides_2013} or fish \citep{yamamuro_neonicotinoids_2019}, questioning the current regulation efforts \citep{schafer_future_2019}.
 
-To evaluate the risks from chemicals for ecosystems, data on their toxicity is required, which is typically produced in standardised ecotoxicological laboratory tests.
+To evaluate the risks from chemicals to ecosystems, data on their toxicity is required, which is typically produced in standardised ecotoxicological laboratory tests.
 For example, \citet{morrissey_neonicotinoid_2015} used ecotoxicological test results from 49 insects and crustaceans to evaluate the effect of neonicotinoid insecticides in the aquatic ecosystem.
-Also, \citet{malaj_organic_2014} compiled experimental toxicity test results for 223 chemicals to assess the health of freshwater ecosystems in Europe.
+Also, \citet{malaj_organic_2014} compiled experimental toxicity test results for 223 chemicals to assess the risk from chemicals to freshwater ecosystems in Europe.
 Similarly, permissible environmental concentrations are often derived from these test data, typically by a combination with safety factors to account for uncertainties.
-The test data mainly relate to a few, well tested standard organisms, such as the brown rat \textit{Rattus norvegicus}, the water flea \textit{Daphnia magna} and the microalga \textit{Raphidocelis subcapitata}.
+The test data mainly relate to a few, well tested, standard organisms, such as the brown rat \textit{Rattus norvegicus}, the water flea \textit{Daphnia magna} and the microalga \textit{Raphidocelis subcapitata}.
 Nevertheless, a much greater variety of organisms has been used in ecotoxicological experiments.
@@
 To date, only few initiatives exist that aim to create a public resource of ecotoxicological data, such as the United States Environmental Protection Agency ECOTOXicology knowledgebase (ECOTOX) (ca.
-1,000,000 test results, 13,000 taxa, 12,000 chemicals) \citep{elonen_ecotoxicology_2018}, the German Environmental Agency's Information System Ecotoxicology and Environmental Quality Targets (ETOX) \citep{umweltbundesamt_etox_2019}, the Pesticide Properties Data Base (PPDB) (ca.
+1,000,000 test results, 13,000 taxa, 12,000 chemicals) \citep{elonen_ecotoxicology_2018}, the German Environmental Agency's Information System Ecotoxicology and Environmental Quality Targets (ETOX) \citep{umweltbundesamt_etox_2019}, the Pesticides Properties DataBase (PPDB) (ca.
 2000 pesticides) \citep{lewis_international_2016} or the Envirotox database \citep{healthandenvironmentalsciencesinstitutehesi_envirotox_2019, connors_creation_2019}.
 The former two compile all available results from experiments into a database.
-However, for many chemicals, multiple ecotoxicity values are available for the same test organism, which can vary strongly and the lack of associated quality information and heterogeneous units hamper reproducible science.
+However, for many chemicals, multiple ecotoxicity values are available for the same test organism.
+These  can vary strongly, thereby causing uncertainty of related analyses \citep{mark_analysis_1998, malaj_physiological_2012}.
+Moreover, the lack of associated quality information and heterogeneous units hamper reproducible science.
 The PPDB database, in contrast, provides single ecotoxicity values only for pesticides and a few selected test organisms, thereby covering only a minor fraction of the vast amount of ecotoxicological data.
-The Envirotox is limited to aquatic organisms.
+The Envirotox database is limited to aquatic organisms.
 Moreover, data analyses often require links to additional data resources, for example to append additional chemical and species information (e.g.
@@
 
-We therefore developed Standartox, a tool and database that aims to overcome the limitations of other databases by continuously incorporating the ever-growing number of test results in an automated process workflow that ultimately leads to a cleaned and harmonised ecotoxicity data collection and provides methods to derive single aggregated ecotoxicity values for a specific chemical-organism test combination.
+We therefore developed Standartox, a tool and database that aims to overcome the limitations of other databases by continuously incorporating the ever-growing number of test results in an automated process workflow that ultimately leads to a harmonised ecotoxicity data collection and provides methods to derive single aggregated ecotoxicity values for a specific chemical-organism test combination.
 Standartox makes use of the publicly available and quarterly updated ECOTOX database \citep{usepa_ecotox_2019} and restricts the data to commonly used endpoints in ecotoxicology, such as half maximal effective concentration (EC\textsubscript{50}) or no-observed-adverse-effect concentrations (NOEC), leading to about 600,000 ecotoxicological test results including 8000 chemicals, tested on about 8000 taxa.
-Users can filter test results according to several parameters, e.g.
+Standartox users can filter test results according to several parameters, e.g.
 refining a search for ecotoxicity data on organisms occurring in specific habitats or regions of the world.
-Above that, Standartox aggregates ecotoxicological test results in a standardized way, by calculating the minimum, the geometric mean and the maximum of the results for each chemical and the associated, user-defined test parameters.
-This tackles the commonly occurring problem of the variability of multiple ecotoxicological test results and which values to choose for assessing the effects of chemicals.
-Hence, Standartox provides the basis for reproducible science and combines information from different sources to simplify the derivation of risk indicators such as Species Sensitivity Distributions (SSD) and Toxic Units (TU), which represent two prominent concepts to assess effects on organisms in ecotoxicology \citep{posthuma_species_2002, kefford_definition_2011, schafer_effects_2011}.
+Above all, Standartox aggregates ecotoxicological test results in a standardized way, by calculating the minimum, the geometric mean and the maximum of the results for each chemical and the associated, user-defined test parameters.
+Hence, this reduces the variability between risk assessments that are due to the selection of different ecotoxicological test data \citep{mark_analysis_1998}.
+Thereby, Standartox provides the basis for reproducible science and combines information from different sources to simplify the derivation of risk indicators such as Species Sensitivity Distributions (SSD) and Toxic Units (TU), which represent two prominent concepts to assess effects on organisms in ecotoxicology \citep{posthuma_species_2002, kefford_definition_2011, schafer_effects_2011}.
 Besides aggregating ecotoxicological test results, Standartox provides a concise overview of the tested chemicals, allowing the identification of potential knowledge gaps.
-Moreover, Standartox could help in reducing the millions of animals used for toxicity testing each year due to facilitated ecotoxicity data access \citep{hartung_chemical_2009}.
-Standartox comes with two front-ends, a web application: \url{standartox.uni-landau.de} and a R \citep{rcoreteam_language_2017} package \textit{standartox}, providing convenience structures and thereby largely reducing processing time for users.
+Moreover, Standartox could help in reducing the millions of animals used for toxicity testing each year by facilitating access to ecotoxicity data \citep{hartung_chemical_2009}.
+Standartox comes with two front-ends, a web application: \url{standartox.uni-landau.de} and an R \citep{rcoreteam_language_2017} package \textit{standartox}, providing convenience structures and thereby largely reducing processing time for users.
 
@@
 \section{Data Description - required}
-
-Standartox constitutes a collection of 600,000 quality checked ecotoxicological test results comprising roughly 8000 distinct chemicals and 8000 taxa that allows for the retrieval of single toxicity equivalents.
-It is build on the ECOTOX database \citep{usepa_ecotox_2019} and the data are processed to clean and harmonise entries to retrieve comparable toxicity endpoints.
-Subsequently, filter and aggregation methods are provided to allow for the retrieval of single toxicity equivalents for specific experimental conditions.
-The underlying ECOTOX database is updated quarterly and provides on average 5228 (2014 - 2019) new toxicity entries with each update.
-As soon as a new update of ECOTOX is available, Standartox is directly built in an automated way and therefore incorporates new data steadily.
-Standartox comes with a version control to guarantee reproducibility, meaning that once derived Standartox estimates can be retraced at a later stage, although new data might have updated the estimates already.
-Users can reach Standartox via a web application and an R-package (standartox) that accesses an application programming interface (API) providing the possibility of scriptable requests.
+Standartox constitutes a collection of quality checked ecotoxicological test results.
+It is build on the ECOTOX database \citep{usepa_ecotox_2019} whose data are processed, cleaned and harmonised to retrieve comparable toxicity endpoints.
+Subsequently, filter and aggregation methods are created to allow for the retrieval of single toxicity equivalents for specific experimental conditions.
+The ECOTOX database is updated quarterly, providing on average 5228 (2014 - 2019) new toxicity entries.
+These are included in Standartox with each update.
 
@@
 The data can be restricted to the three endpoint groups, namely half maximal effective/lethal concentration/dose values (e.g.
-EC50, LD50), henceforth abbreviated as XX50, lowest observed effect levels/concentrations (LOEC/L), henceforth abbreviated as LOEX and no observed effect levels/concentrations (NOEL/C), henceforth abbreviated as NOEX (Table \ref{tab:endpoints-conflate}).
+EC\textsubscript{50}, LD\textsubscript{50}), henceforth abbreviated as XX50, lowest observed effect levels/concentrations (LOEC/L), henceforth abbreviated as LOEX and no observed effect levels/concentrations (NOEL/C), henceforth abbreviated as NOEX (Table~\ref{tab:endpoints-conflate}).
 Standartox allows the ecotoxicity data to be filtered by effect groups (e.g.
-mortality, population, growth) (Fig.
-\ref{fig:stx-parameters}B) and concentration types (e.g.
+mortality, population, growth) (Figure~\ref{fig:stx-parameters}A) and concentration types (e.g.
 formulation, active ingredient) as well as test durations (in hours).
 In addition to these test-specific parameters, Standartox data entries can be filtered by chemical-specific parameters such as the CAS number and chemical roles (e.g.
-pesticides, metals, drugs) (Fig.
-\ref{fig:stx-parameters}B) and classes (e.g.
-organochlorine, triazine) (Fig.
-\ref{fig:stx-parameters}C).
-Furthermore, the Standartox data can be refined to organism-specific parameters, such as the organism habitat (e.g.
-freshwater, marine, terrestrial) (Fig.
-\ref{fig:stx-parameters}E), their occurrence regions (e.g.
-Europe, South America) (Fig.
-\ref{fig:stx-parameters}F) as well as different taxonomic levels.
+pesticides, metals, drugs) (Figure~\ref{fig:stx-parameters}B) and classes (e.g.
+organochlorine, triazine) (Figure~\ref{fig:stx-parameters}C).
+Furthermore, the Standartox data can be refined to certain taxonomic groups (Figure~\ref{fig:stx-parameters}D) as well as organism-specific parameters, such as the organisms' habitat (e.g.
+freshwater, marine, terrestrial) (Figure~\ref{fig:stx-parameters}E) and distribution (e.g.
+Europe, South America) (Figure~\ref{fig:stx-parameters}F).
 
@@
     \includegraphics[width=0.8\textwidth]{article/figures/standartox_parameters.png}
-    \caption{Share of 10 most frequent entries for the parameters (\textbf{A}) effect group, (\textbf{B}) chemical role, (\textbf{C}) chemical class, (\textbf{D}) taxonomic order, (\textbf{E}) organism habitat and (\textbf{F}) organism occurrence region in Standartox.}
+    \caption{Share of 10 most frequent entries for the parameters (\textbf{A}) effect group, (\textbf{B}) chemical role, (\textbf{C}) chemical class, (\textbf{D}) taxonomic order, (\textbf{E}) organism habitat and (\textbf{F}) organism distribution in Standartox.
+Multiple classifications are possible (e.g.
+a chemical can be a fungicide and a pesticide).}
     \label{fig:stx-parameters}
@@
 \subsection{Aggregation}
-Typically, species exhibit a differential sensitivity towards chemicals (Fig.
-\ref{fig:stx-variability}A).
-Moreover, multiple ecotoxicity values are available for individual species-chemical combinations and these can also exhibit high variability due to several factors such as durations of ecotoxicity tests (Fig.
-\ref{fig:stx-variability}B), experimental conditions and physiology and genetic strain of the test individuals or populations.
+Typically, species exhibit a differential sensitivity towards chemicals (Figure~\ref{fig:stx-variability}A).
+Moreover, multiple ecotoxicity values are available for individual species-chemical combinations and these can also exhibit high variability due to several factors such as durations of ecotoxicity tests (Figure~\ref{fig:stx-variability}B), experimental conditions and physiological or genetic fitness differences between test individuals or populations.
+Not every factor is recorded though, leading to unexplainable variability (Figure~\ref{fig:stx-variability}C).
 To aggregate multiple ecotoxicity values into a single value on the desired taxonomic level (e.g.
@@
 across all pesticides), Standartox provides several aggregation methods including the minimum, the maximum and the geometric mean allowing to aggregate the filtered data set.
-The geometric mean is preferred in comparison to the arithmetic mean, because it is considerably less influenced by outliers and is suitable for skewed data.
-Also, the geometric mean is preferable over the median, because the median completely ignores the influence of large or small values, making it unreliable for small data sets.
+The geometric mean is preferred in comparison to the arithmetic mean, because it is less influenced by outliers and is suitable for skewed data.
+Also, the geometric mean is preferable over the median, because the median completely ignores the tails of the data distribution, making it unreliable for small data sets \citep{leith_comparison_2010}.
 In the course of the aggregation process, outliers that exceed 1.5 times the interquartile range are flagged to caution Standartox users.
-However, since the geometric mean is relatively robust against outliers, they are not removed automatically from the aggregation.
-Overall, aggregation with Standartox allows to reduce the between-study variability when toxicity data are aggregated.
+However, they are considered in aggregation, given that the geometric mean is relatively robust against outliers.
+Overall, Standartox provides a harmonised and reproducible approach to data aggregation that may contribute to reduce between-study variability.
 
@@
     \includegraphics[width=0.8\textwidth]{article/figures/results_variability.png}
-    \caption{Violin plots of (\textbf{A}) the variability of test results (EC\textsubscript{50}) in Standartox because of different species (i.e.
-\textit{Xenopus laevis} - Amphibia, \textit{Raphidocelis subcapitata} - Algae, \textit{Oncorhynchus mykiss} - fish, \textit{Lemna minor} - invertebrate) to the chemical atrazine tested for 96 h, (\textbf{B}) diﬀerent durations of toxicity tests with zinc sulfate and \textit{Daphnia magna} and (\textbf{C}) unaccountable variability of cupric sulfate tested on \textit{Pimephales promelas} for 96 h.
-Purple dots depict Standartox geometric mean estimates and black dots show un-aggregated values for each group.
-To facilitate readability, data points are randomly scattered along a hypothetical y-axis and are greyed out within the violins.}
+    \caption{Violin plots of of test results (EC\textsubscript{50}) in Standartox illustrating (\textbf{A}) differential variability and data distribution between species (i.e.
+\textit{Xenopus laevis} - Amphibia, \textit{Raphidocelis subcapitata} - Algae, \textit{Oncorhynchus mykiss} - fish, \textit{Lemna minor} - invertebrate) for the chemical atrazine in 96 h tests, (\textbf{B}) how the variability in toxicity tests with zinc sulfate and \textit{Daphnia magna} varies with test duration and (\textbf{C}) high variability that is not explained by the available test characteristics in the case of cupric sulfate tested on \textit{Pimephales promelas} for 96 h.
+Purple dots depict Standartox geometric mean estimates and black dots show raw data for each group.
+To facilitate readability, data points are randomly scattered along a hypothetical y-axis and are greyed out if within the violins.}
     \label{fig:stx-variability}
@@
 \subsection{Accuracy Assessment}
-To validate Standartox results we compared the aggregated results using the geometric mean to values from other databases.
+To validate Standartox results we compared geometric means resulting from the aggregation in Standartox to the corresponding values from other databases, for chemicals where data was available in both resources.
 The PPDB provides ecotoxicity data on a few selected species from chemical risk assessment that have been manually quality controlled through expert judgement \citep{lewis_international_2016}.
-The vast majority of aggregated values (91 \%) of Standartox lie within one order of magnitude of the corresponding PPDB values (n = 3601).
-This would increase to 92.7 \%, when restricting the comparison to Standartox values where data from at least five experiments were available.
+The vast majority of aggregated values (91~\%) of Standartox lie within one order of magnitude of the corresponding PPDB values (n = 3601).
+This would increase to 92.7~\%, when restricting the comparison to Standartox values where data from at least five experiments are available.
 Similarly, we compared Standartox to ecotoxicity values for \textit{D.
-magna} from the ChemProp \citep{ufzdepartmentofecologicalchemistry_chemprop_2019} software, which estimates LC\textsubscript{50} values via quantitative structure-activity relationship (QSAR) models \citep{schuurmann_quantitative_2011} We found that  95 \% of Standartox values lie within one order of magnitude of the Chemprop (n = 179) vlaues.
-However, the diﬀerence is not necessarily an indication of lower quality of Standartox estimates but may also reflect the wider range of experimental conditions for which data are available in the database underlying Standartox as well as inaccurate predictions for QSAR models, respectively (Figure \ref{fig:standartox_ppdb_diff}).
+magna} from the ChemProp \citep{ufzdepartmentofecologicalchemistry_chemprop_2019} software, which estimates LC\textsubscript{50} values via quantitative structure-activity relationship (QSAR) models \citep{schuurmann_quantitative_2011}.
+We found that 95~\% of Standartox values lie within one order of magnitude of the Chemprop (n = 179) values.
+However, the diﬀerence is not necessarily an indication of lower quality of Standartox estimates but may also reflect the wider range of experimental conditions for which data are available in the database underlying Standartox as well as inaccurate predictions for QSAR models, respectively (Figure~\ref{fig:standartox_ppdb_diff}).
 
@@
 Indeed, the increasing amount of data from ecotoxicological tests and experiments that is becoming available has elicited several initiatives to harmonize these data.
-These initiatives partly aim for overlapping goals, yet have limitations or objectives that discriminate them from Standartox:
+These initiatives partly aim for overlapping goals, yet have limitations or objectives that distinguish them from Standartox:
 
-Comptox, is a web tool published by the EPA which, similar to Standartox allows for filtering test results, the retrieval of additional chemical information and the estimation of toxicity equivalents, such as 48 h \textit{Daphnia magna} LC\textsubscript{50} values.
-However, toxicity estimations are limited to standard organisms (e.g.
+\textbf{Comptox}, is a web tool published by the EPA which, similar to Standartox allows for filtering test results, the retrieval of additional chemical information and the estimation of toxicity data, such as 48 h \textit{Daphnia magna} LC\textsubscript{50} values.
+However, toxicity estimations are limited to standard test organisms (e.g.
 \textit{Daphnia magna}), and the tool lacks the possibility for automated retrieval \citep{williams_comptox_2017}.
 
-The EnviroTox database which also uses, amongst others the ECOTOX database as an input has recently been published \citep{healthandenvironmentalsciencesinstitutehesi_envirotox_2019, connors_creation_2019}.
+The \textbf{EnviroTox} database which also uses, amongst others the ECOTOX database as an input has recently been published \citep{healthandenvironmentalsciencesinstitutehesi_envirotox_2019, connors_creation_2019}.
 In contrast to Standartox, EnviroTox  is restricted to selected aquatic organisms (i.e.
 fish, amphibians, invertebrates and algae) and experimental durations (at least 24 h) and uses a rule-based algorithm to derive single ecotoxicity values.
-Besides, EnviroTox provide additional information on toxicity endpoints, such as acute or chronic classifications and mode of action assignments.
-We think, an action that can also be allocated to the user, since there are several different ways to classify such parameters, especially those mentioned, guaranteeing user flexibility.
+Besides, EnviroTox provides additional information on toxicity endpoints, such as acute or chronic classifications and mode of action assignments.
+We intentionally omitted such classifications given that the approach to classification may vary with the purpose of the study.
 The EnviroTox database allows for an aggregation into single toxicity values for individual taxa, whereas Standartox performs this aggregation for individual taxa-chemical combinations.
-However, the Standartox results for individual taxa-chemical combinations could easily be aggregated across chemicals in a second step to achieve aggregations performed in EnviroTox.
+However, the Standartox results for individual taxa-chemical combinations could easily be aggregated across chemicals in a second step to provide a similar aggregation as that performed in EnviroTox.
 
-The Etox database, like the ECOTOX database collects and provides filter methods for ecotoxicological test information.
+The \textbf{Etox} database, like the ECOTOX database collects and provides filter methods for ecotoxicological test information.
 
-The PPDB provides data only on pesticides and as mentioned before provides single quality controlled values only for commonly used taxa, e.g. \textit{D. magna} or \textit{R. subcapitata}.
-
-In summary, none of the above mentioned initiatives aim for a holistic and automated standardized aggregation method of exposure endpoints for individual chemicals.
-In addition, they lack the possibility to access the databases through a common high level programming language, such as R.
-As outlined above, toxicity estimates from different studies can be highly variable due to a wide range of experimental conditions such as pH, temperature or conductivity \citep{rosenkrantz_influence_2013, li_temperature_2011}.
-Integrating these factors into the aggregated estimates would improve toxicity estimates.
-However, the current implementation of Standartox does not consider these factors, because the ECOTOX database only provides sparse records on experimental conditions.
-The most frequently provided experimental conditions are temperature (77 \%), pH (56 \%), hardness (27 \%), dissolved oxygen (18 \%), Alkalinity (15 \%) and salinity (9 \%), and all others for less than 5 \%.
-A text-mining approach, where a literature reference is associated with  ecotoxicity raw data, iterating through the individual publications could potentially increase this number.
+The \textbf{PPDB} provides data only on pesticides and as mentioned before provides single quality controlled values only for commonly used taxa, e.g. \textit{D. magna} or \textit{R. subcapitata}.
 
 \begin{table}[H]
-    \caption{Different databases that provide ecotoxicological data.
+    \caption{Overview on databases that provide ecotoxicological data.
 Abbreviations: \textbf{ALL:} Most important test parameters, including chemical, taxon, duration for filtering ecotoxicological data are incorporated.
@@
     Etox & Umweltbundesamt & ALL & no & WEB \\
-    Pesticide Property Data Base (PPDB) & University of Hertfordshire & fixed values & manual selection & WEB, file \\
+    Pesticides Properties DataBase (PPDB) & University of Hertfordshire & fixed values & manual selection & WEB, file \\
     Standartox & this article & ALL & chemical, organism & API, WEB \\
@@
 
+In summary, none of the above mentioned initiatives aim for an automated and standardized aggregation method of exposure endpoints for individual chemicals.
+In addition, they lack the possibility to access the databases through a common high level programming language, such as R.
+As outlined above, toxicity estimates from different studies can vary strongly due to a wide range of experimental conditions such as pH, temperature and conductivity \citep{rosenkrantz_influence_2013, li_temperature_2011}.
+Integrating these conditions into the aggregated estimates would certainly improve toxicity estimates.
+However, the current implementation of Standartox omits these conditions, because the ECOTOX database only provides sparse records on experimental conditions.
+The most frequently provided experimental conditions are temperature (77~\%), pH (56~\%), hardness (27~\%), dissolved oxygen (18~\%), Alkalinity (15~\%) and salinity (9~\%).
+For all other conditions less than 5~\% of data entries are available.
+A text-mining approach, where a literature reference is associated with ecotoxicity raw data, iterating through the individual publications could potentially increase this number.
+E.g.
+\citet{compson_linking_2018} successfully applied text-mining techniques to retrieve species trait data.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Methods - required}
+\section{Methods}
 An automated processing pipeline downloads the quarterly released ECOTOX database, performs several preparation steps on it and exports a final Standartox data set.
@@
 All data presented in this paper are derived from the Standartox build based on the ECOTOX release from the 12.12.2019.
+The code for Standartox is located in two Github repositories \url{https://github.com/andschar/standartox-build} and \url{https://github.com/andschar/standartox}.
+The former contains code to process the data and to build the web application and the API, the latter contains code to build the R-package.
+Most of the code is written in R 3.6.1 and associated packages (List: \ref{lis:r-packages}) and SQL for PostgreSQL 9.6.1.
 
@@
 Subsequently, Structured Query Language (SQL) functions for further processing the data are implemented.
-In addition lookup tables that enable duration and concentration unit conversions of the data are created.
-A meta table providing information such as the release version of the ECOTOX database is added.
+In addition lookup tables that enable the conversion of units such as duration and concentration are created.
+A meta-table providing information such as the release version of the ECOTOX database is added.
 Then, provided Chemical Abstracts Service Numbers (CAS) and taxonomic names are used to query additional information from publicly available databases on chemicals and organisms, respectively.
-This includes the Compendium of Pesticide Common Names \citep{wood_compendium_2019}, the Chemical Entities of Biological Interest (ChEBI) database \citep{hastings_chebi_2016}, the Chemical Identifier Resolver (CIR) service \citep{nationalinstitutesofhealthnih_chemical_2019}, the Pubchem database \citep{kim_pubchem_2016}, Eurostat \citep{europeancommission_eurostat_2019} and Wikidata \citep{vrandecic_wikidata_2014} for chemicals and the World Register of Marine Species (WoRMS) \citep{wormseditorialboard_world_2018} and the Global Biodiversity Information Facility (GBIF) \citep{gbif_gbif_2019} (Table \ref{tab:data-base-additional}) for habitat and spatial distribution of organisms.
-Given that CAS and taxonomic names are not always unambiguous, e.g.
-the genus \textit{Eisenia} can refer to an algae and a worm at once, we first match them against the database specific identifiers and subsequently use them to retrieve the actual data.
+This includes the Compendium of Pesticide Common Names \citep{wood_compendium_2019}, the Chemical Entities of Biological Interest (ChEBI) database \citep{hastings_chebi_2016}, the Chemical Identifier Resolver (CIR) service \citep{nationalinstitutesofhealthnih_chemical_2019}, the Pubchem database \citep{kim_pubchem_2016}, Eurostat \citep{europeancommission_eurostat_2019} and Wikidata \citep{vrandecic_wikidata_2014} for chemicals and the World Register of Marine Species (WoRMS) \citep{wormseditorialboard_world_2018} and the Global Biodiversity Information Facility (GBIF) \citep{gbif_gbif_2019} (Table~\ref{tab:data-base-additional}) for habitat and spatial distribution of organisms.
+Given that CAS and taxonomic names can be ambiguous, e.g.
+the genus \textit{Eisenia} can refer to an algae and a worm, we first match them against the database specific identifiers and subsequently use them to retrieve the actual data.
 In a next step the data is added to Standartox to enable filtering for specific chemical roles (e.g.
-drug, pesticide, personal care product) and classes (e.g.
-pyrethroid, metal, carbamate) as well as spatial distribution (i.e.
+drug, metal, pesticide, personal care product) and classes (e.g.
+pyrethroid, carbamate) as well as spatial distribution (i.e.
 continents) and habitat preferences (e.g.
 freshwater) of individual taxa.
-Taxa that where not identified to at least genus level are excluded.
+Taxa that were not identified to at least genus level are excluded, because relative toxicity comparisons have been shown to be not meaningful for higher taxonomic levels \citep{rainbow_trace_2002, buchwalter_differences_2005, malaj_physiological_2012}.
 Finally, the Standartox data set is compiled, which includes the harmonisation of data, e.g.
-through conversion of result concentration and test duration units.
+through conversion of test concentration and duration units.
 1237 distinct concentration units are converted to X harmonised ones, as are the 126 distinct duration units to X.
-Standartox retains only duration units, that can unambiguously be converted to hours, excluding duration units such as harvest or lifetime.
-To guarantee appropriate unit conversion and harmonisation, we compared one automated conversion to a manually done conversion for each of the distinct concentration and duration units.
+Standartox retains only duration units that can unambiguously be converted to hours.
+To guarantee appropriate unit conversion and harmonisation, we compared an automated to a manual conversion for each of the distinct concentration and duration units.
 Furthermore, the units are cleaned, for example through removing additional information in the field such as \textit{food}, \textit{soil}, \textit{ai} that are also coded in other variables and hinder the processing of units.
 Concentrations that are given as rates such as per day (e.g.
-mg/kg/day) are multiplied by the days and the converted.
+mg/kg/day) are multiplied by the days of the test and then converted.
 Experimental endpoints are restricted to three groups, namely NOEX, LOEX and XX50.
@@
 IC10, EC25, LD99) or maximum acceptable toxicant (MATC) concentrations are removed.
-Overall this reduces the ECOTOX data set from 971,430 to 611,248 harmonised test results that are included in Standartox.
-Along with that, a catalog, listing all distinct entries and value ranges, for categorical and continuous variables, respectively is created.
+Along with that, a catalog, listing all distinct entries and value ranges, for categorical and continuous variables, respectively, is created.
 Finally, we run quality control scripts that check the accuracy of the data.
@@
 
-\begin{figure}
+\begin{figure}[H]
     \centering
@@
 The ECOTOX knowledge base is downloaded quarterly and processed (i.e.
-query additional information with CAS and Taxa name and cleaning concentration and duration units.
-Subsequently a Standartox data set is compiled together with filter and aggregation methods.
+query additional information with CAS and taxa name and conversion of concentration and duration units.
+Subsequently, a Standartox data set is compiled together with filter and aggregation methods.
 Thus, users can access the Standartox data set and filter and aggregate through a web application and an R-package.}
@@
 \subsection{Application methods}
-The web application and the API load the compressed serialized Standartox data into memory and allow a client to interact with it.
-The client thereby calls the functions stx\_filter() and stx\_aggregate() that filter and aggregate the data according to specific parameters (Table \ref{tab:app-parameters}).
+When accessed, the web application and the API load the compressed serialized Standartox data into memory and allow the user to interact with it.
+The user can then call the functions stx\_filter() and stx\_aggregate() that filter and aggregate the data according to specific parameters (Table~\ref{tab:app-parameters}).
 The interactive web application is built in R using the shiny framework, which runs with the help of a shiny server \citep{chang_shiny_2018} on \url{standartox.uni-landau.de}.
 The API is built by using the R package \textit{plumber} \citep{trestletechnologyllc_plumber_2018}, which allows for the creation of Representational State Transfer (REST) APIs from R.
-REST is a software standard that defines web service communication rules.
+REST is a software architectural style that defines web service communication rules.
 The API is reachable via the URL \url{http://139.14.20.252} and port 8000.
-Four API-endpoints (\textit{/catalog}, \textit{/filter}, \textit{/aggregate} and \textit{/meta}) can be queried by users.
+Four API-endpoints (\textit{/catalog}, \textit{/filter}, \textit{/aggregate} and \textit{/meta}) can be queried (Table~\ref{tab:api-endpoints}).
 The \textit{/catalog} API-endpoint returns a JavaScript Object Notation (JSON) file containing a catalog of possible filter parameters to choose from.
 The \textit{/filter} returns the filtered Standartox table as a compressed serialized binary file created by the R \textit{fst} package \citep{klik_fst_2019}, to reduce size and allow for fast user queries.
-The \textit{/aggregate} API-endpoint returns an R function allowing the client to aggregate the filtered data.
+The \textit{/aggregate} API-endpoint returns an R function allowing the client to aggregate the filtered data locally.
 Lastly, the \textit{/meta} API-endpoint returns a JSON file with meta information, such as the timestamp of the request and the used Standartox version.
-The API is designed to be used with the R-package \textit{standartox} and therefore uses serialization methods specific to R (rds() from the R package \pkg{base} and fst() from the package fst).
-In order to facilitate the API usage the R-package \textit{standartox} is created.
+The API is designed to be used with the R-package \textit{standartox} and therefore uses serialization methods specific to R (rds() from the R package \textit{base} and fst() from the package \textit{fst}).
+To facilitate the API usage the R-package \textit{standartox} is created.
 
 \begin{table}[H]
-    \caption{Input parameters for the Standartox web application and the R-package standartox (CAS - Chemical Abstracts Service Registry Number, NOEX and XX50 - Standartox endpoints \ref{tab:endpoints-conflate}), vers - Standartox version.}
+    \caption{Input parameters for the Standartox web application and the R-package standartox (CAS - Chemical Abstracts Service Registry Number, NOEX and XX50 - Standartox endpoints \ref{tab:endpoints-conflate}), vers - Standartox version).}
     \label{tab:app-parameters}
@@
     endpoint & NOEX, XX50 \\
+    exposure & auquatic, diet \\
     vers & 20191212 \\
@@
 \section{User Notes - optional}
-% \input{article/parts/04_user_notes}
-By accessing the web application users can filter and download the resulting data sets as a comma-separated values (csv) file.
-Users of the R-package can directly load the data in R.
+Users can access Standartox either via the web application \url{standartox.uni-landau.de} or via the R-package \textit{standartox}.
+By accessing the web application, users can filter and download the resulting data sets as a comma-separated values (csv) file.
+Users of the R-package can directly load the data within R.
 The R-package provides the two functions \textit{stx\_catalog()} and \textit{stx\_query()}.
-The first command queries a catalog of possible Standartox parameters into a R list object.
+The first command queries a catalog of possible Standartox parameters into an R list object.
 The latter allows users to set the Standartox filter parameters and to fetch the actual data.
-It returns a R list of three tables (i.e.
-R data.frames) containing the filtered data set, the aggregated data set and a table with the meta information retrieved from the API endpoints \textit{/filter} and \textit{/meta}.
-A short R-code example is given below \ref{lis:standartox-example} and a detailed description on the usage of the R package is provided on its Github page: \url{https://github.com/andschar/standartox}.
+It returns an R list of three tables (i.e.
+R data.frames) containing the filtered data set, the aggregated data set and a table with the meta information retrieved from the API endpoints.
+A short R-code example is given below (Listing \ref{lis:standartox-example}) and a detailed description on the usage of the R package is provided on its Github page: \url{https://github.com/andschar/standartox}.
 
@@
     label = {lis:standartox-example}]
+
 # install
@@
 
-Standartox's code is distributed on three Github repositories below.
-All code can be freely accessed under the MIT License <CITATION>.
-Most of the code is written in R 3.6.1 and associated packages (compare Supplement: \ref{list:r-packages}) as well as some parts in SQL for PostgreSQL 9.6.1.
-
-\begin{itemize}
-
-\item \url{https://github.com/andschar/standartox-build} \newline
-Contains the code for the data acquisition and processing pipeline.
-
-\item \url{https://github.com/andschar/standartox-app} \newline
-Contains the code for the web application and the API.
-
-\item \url{https://github.com/andschar/standartox} \newline
-Contains the code for the R-package.
-
-\end{itemize}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.
-Due to the steady incorporation of new ecotoxicity data, its values produced by Standartox will be subject to change and most certainly produce slightly different results in its aggregations in the future.
-However, we regard this as an advantage, since other published works that aim in similar directions often constitute a singular effort or require manual work for each update.
-Standartox, in contrast automates these processes and as already described would provide means to access older versions anyhow, to assure reproducibility.
-In comparison to rule-based approaches, such as the PPDB for the derivation of single ecotoxicity values, Standartox has the advantage to not be prone to the subjectivity, the exhaustiveness and the incompleteness of a human induced set of rules.
+Due to the steady incorporation of new ecotoxicity data, the aggregated values produced by Standartox can be subject to change with future updates.
+We regard this as an advantage rather than a drawback because other published works that aim in a similar direction often constitute a singular effort or require manual work for each update.
+Standartox, in contrast automates the update process, yet still provides access to its older versions, assuring reproducability and version control.
+In comparison to rule-based approaches for the derivation of single ecotoxicity values, Standartox has the advantage to be free from the subjectivity of a set of human-induced rules.
 Above all, Standartox provides quick access through its design to be queried via the R language.
-In general, Standartox is designed to foster chemical risk assessment (CRA).
 Due to an increased amount of available ecotoxicological test data, it becomes fundamental to provide and distribute ecotoxicity information in adequate formats, both easily accessible for humans and easily processable for machines.
-Standartox is a tool that does this in a generalizing way and puts its focus on the aggregation of toxicity data, thereby adding a piece to the puzzle of modern ecotoxicological analyses.
+Standartox meets these requirements and puts its focus on the aggregation of toxicity data, thereby adding a piece to the puzzle of modern ecotoxicological data analyses.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{conceptualization, A.S.
-and V.C.S.; methodology, A.S.; software, A.S.; validation, A.S.; formal analysis, A.S.; investigation, A.S.; resources, A.S.; data curation, A.S.; writing--original draft preparation, A.S.; writing--review and editing, A.S., V.C.S.
-and R.B.S.; visualization, A.S.; supervision, R.B.S.; project administration, A.S.
-and R.B.S.; funding acquisition, Y.Y.}
+\authorcontributions{conceptualization and methods, A.S.,R.B.S.
+and V.C.S.; software development and validation, A.S.; data curation and analysis, A.S.; writing--original draft preparation, A.S.; writing--review and editing, A.S., V.C.S.
+and R.B.S.; supervision, R.B.S.}
 
@@
 \begin{tabular}{@{}ll}
-API & Application Programming Interface \\
-CAS & Chemical Abstracts Service Registry Number \\
+API & Application programming interface \\
+CAS & Chemical abstracts service registry Number \\
 ChEBI & Chemical Entities of Biological Interest database \\
 CIR & Chemical Identifier Resolver service \\
-E/LC/D\textsubscript{50} &  half maximal effective/lethal concentration/dose \\
-ECOTOX & US EPA Ecotox Knowledge Base \\
-GBIF & Global Biodiversity193Information Facility \\
+E/LC/D\textsubscript{50} &  Half maximal effective/lethal concentration/dose \\
+ECOTOX & US EPA Ecotox Knowledgebase \\
+GBIF & Global Biodiversity Information Facility \\
 JSON & JavaScript Object Notation file format \\
-LOEC/L & Lowest Observed Effect Levels/Concentration \\
-NOEC/L & No Observed Effect Levels/Concentration \\
-PPDB & Pesticides Property Data Base \\
+LOEC/L & Lowest observed effect levels/concentration \\
+NOEC/L & No observed effect levels/concentration \\
+PPDB & Pesticides Properties DataBase \\
 QSAR & Quantitative structure–activity relationship \\
-REST & Representational State Transfer software architectural style \\
-SSD & Species Sensitivity Distribuion \\
-TU & Toxic Unit \\
+REST & Representational state transfer software architectural style \\
+SSD & Species sensitivity distribution \\
+TU & Toxic unit \\
 WoRMS & World Register of Marine Species \\
@@
     XX50 & ET50 & Effective response time to 50\% of test organisms \\
-    XX50 & LT50 & Time to 50\% mortality of organisms \\
+    XX50 & LT50 & Time to 50\% mortality of test organisms \\
     NOEX & NOEC & No-observable-effect-concentration \\
@@
 \begin{table}[H]
-    \caption{Application programming interface (API) endpoints, Http methods,
-    Requests and Response objects. Object is abbreviated with obj.}
+    \caption{Application programming interface (API) endpoints, Http methods, Requests and Response objects. JSON - Javascript opject notation file}
     \label{tab:api-endpoints}
@@
     \midrule
-    /catalog & POST & \makecell{Standartox \\ version string} & \makecell{Catalog object of \\ possible filter parameters \\ JSON obj.}    \\[0.5cm]
-    /filter & POST & \makecell{Standartox filter \\ parameters (Table \ref{tab:app-parameters})} & \makecell{Filtered Standartox data \\ serialized R data.frame \\ R fst:: \citep{klik_fst_2019} obj.}   \\[0.5cm]
-    /aggregate & GET & / & \makecell{Aggregate function \\ serialized R function \\ R .rds \citep{rcoreteam_language_2017} obj.}    \\[0.5cm]
-    /meta & POST & \makecell{Standartox \\ version string} & \makecell{Meta data on request \\ JSON obj.}    \\
+    /catalog & POST & Standartox version string & Catalog object (JSON) \\
+    /filter & POST & Standartox filter parameters & Filtered Standartox data (serialized) \\
+    /aggregate & GET & / & Aggregate function (serialized) \\
+    /meta & POST & Standartox version string & Meta data on request (JSON) \\
     \bottomrule
@@
 
-Rinker TW, Kurkiewicz D (2018).
-\emph{pacman: Package Management for R}.
-version 0.5.0, \url{http://github.com/trinker/pacman}.
-\newline R Core Team (2019).
-\emph{R: A Language and Environment for Statistical Computing}.
-R Foundation for Statistical Computing, Vienna, Austria.
-\url{https://www.R-project.org/}.
-\newline Wickham H, Hester J, Chang W (2019).
-\emph{devtools: Tools to Make Developing R Packages Easier}.
-R package version 2.2.1, \url{https://CRAN.R-project.org/package=devtools}.
-\newline Lang DT, team tC (2019).
-\emph{RCurl: General Network (HTTP/FTP/...) Client Interface for R}.
-R package version 1.95-4.12, \url{https://CRAN.R-project.org/package=RCurl}.
-\newline Wickham H (2019).
-\emph{stringr: Simple, Consistent Wrappers for Common String Operations}.
-R package version 1.4.0, \url{https://CRAN.R-project.org/package=stringr}.
-\newline Bengtsson H (2019).
-\emph{R.utils: Various Programming Utilities}.
-R package version 2.9.0, \url{https://CRAN.R-project.org/package=R.utils}.
-\newline Hiebert J (2016).
-\emph{udunits2: Udunits-2 Bindings for R}.
-R package version 0.13, \url{https://CRAN.R-project.org/package=udunits2}.
-\newline Wickham H (2019).
-\emph{rvest: Easily Harvest (Scrape) Web Pages}.
-R package version 0.3.4, \url{https://CRAN.R-project.org/package=rvest}.
-\newline Wickham H (2019).
-\emph{httr: Tools for Working with URLs and HTTP}.
-R package version 1.4.1, \url{https://CRAN.R-project.org/package=httr}.
-\newline Ooms J (2014).
-``The jsonlite Package: A Practical and Consistent Mapping Between JSON Data and R Objects.''
-\emph{arXiv:1403.2805 [stat.CO]}.
-\url{https://arxiv.org/abs/1403.2805}.
-\newline Wickham H, Bryan J (2019).
-\emph{readxl: Read Excel Files}.
-R package version 1.3.1, \url{https://CRAN.R-project.org/package=readxl}.
-\newline Walker A (2019).
-\emph{openxlsx: Read, Write and Edit XLSX Files}.
-R package version 4.1.0.1, \url{https://CRAN.R-project.org/package=openxlsx}.
-\newline Henry L, Wickham H (2019).
-\emph{purrr: Functional Programming Tools}.
-R package version 0.3.2, \url{https://CRAN.R-project.org/package=purrr}.
-\newline Dowle M, Srinivasan A (2019).
-\emph{data.table: Extension of `data.frame`}.
-R package version 1.12.4, \url{https://CRAN.R-project.org/package=data.table}.
-\newline Conway J, Eddelbuettel D, Nishiyama T, Prayaga SK, Tiffin N (2017).
-\emph{RPostgreSQL: R Interface to the 'PostgreSQL' Database System}.
-R package version 0.6-2, \url{https://CRAN.R-project.org/package=RPostgreSQL}.
-\newline R Special Interest Group on Databases (R-SIG-DB), Wickham H, Müller K (2018).
-\emph{DBI: R Database Interface}.
-R package version 1.0.0, \url{https://CRAN.R-project.org/package=DBI}.
-\newline Oksanen J, Blanchet FG, Friendly M, Kindt R, Legendre P, McGlinn D, Minchin PR, O'Hara RB, Simpson GL, Solymos P, Stevens MHH, Szoecs E, Wagner H (2019).
-\emph{vegan: Community Ecology Package}.
-R package version 2.5-5, \url{https://CRAN.R-project.org/package=vegan}.
-\newline Wickham H (2011).
-``The Split-Apply-Combine Strategy for Data Analysis.''
-\emph{Journal of Statistical Software}, \bold{40}(1), 1--29.
-\url{http://www.jstatsoft.org/v40/i01/}.
-\newline Komsta L (2011).
-\emph{outliers: Tests for outliers}.
-R package version 0.14, \url{https://CRAN.R-project.org/package=outliers}.
-\newline Wickham H (2019).
-\emph{feather: R Bindings to the Feather 'API'}.
-R package version 0.3.3, \url{https://CRAN.R-project.org/package=feather}.
-\newline Wickham H (2016).
-\emph{ggplot2: Elegant Graphics for Data Analysis}.
-Springer-Verlag New York.
-ISBN 978-3-319-24277-4, \url{https://ggplot2.tidyverse.org}.
-\newline Slowikowski K (2019).
-\emph{ggrepel: Automatically Position Non-Overlapping Text Labels with
-'ggplot2'}.
-R package version 0.8.1, \url{https://CRAN.R-project.org/package=ggrepel}.
-\newline Wilke C (2019).
-\emph{cowplot: Streamlined Plot Theme and Plot Annotations for 'ggplot2'}.
-R package version 0.9.4, \url{https://CRAN.R-project.org/package=cowplot}.
-\newline Neuwirth E (2014).
-\emph{RColorBrewer: ColorBrewer Palettes}.
-R package version 1.1-2, \url{https://CRAN.R-project.org/package=RColorBrewer}.
-\newline Tennekes M (2017).
-\emph{treemap: Treemap Visualization}.
-R package version 2.4-2, \url{https://CRAN.R-project.org/package=treemap}.
-\newline Chamberlain S, Barve V, Mcglinn D, Oldoni D, Desmet P, Geffert L, Ram K (2019).
-\emph{rgbif: Interface to the Global Biodiversity Information Facility API}.
-R package version 1.3.0, \url{https://CRAN.R-project.org/package=rgbif}.
-
-Chamberlain S, Boettiger C (2017).
-``R Python, and Ruby clients for GBIF species occurrence data.''
-\emph{PeerJ PrePrints}.
-\url{https://doi.org/10.7287/peerj.preprints.3304v1}.
-\newline Scott Chamberlain, Eduard Szocs (2013).
-``taxize - taxonomic search and retrieval in R.''
-\emph{F1000Research}.
-\url{http://f1000research.com/articles/2-191/v2}.
-
-Chamberlain S, Szoecs E, Foster Z, Arendsee Z, Boettiger C, Ram K, Bartomeus I, Baumgartner J, O'Donnell J, Oksanen J, Tzovaras BG, Marchand P, Tran V, Salmon M, Li G, Grenié M (2019).
-\emph{taxize: Taxonomic information from around the web}.
-R package version 0.9.7, \url{https://github.com/ropensci/taxize}.
-\newline Arel-Bundock V, Enevoldsen N, Yetman C (2018).
-``countrycode: An R package to convert country names and country codes.''
-\emph{Journal of Open Source Software}, \bold{3}(28), 848.
-\url{https://doi.org/10.21105/joss.00848}.
-\newline Microsoft, Weston S (2017).
-\emph{foreach: Provides Foreach Looping Construct for R}.
-R package version 1.4.4, \url{https://CRAN.R-project.org/package=foreach}.
-\newline Corporation M, Weston S (2018).
-\emph{doParallel: Foreach Parallel Adaptor for the 'parallel' Package}.
-R package version 1.0.14, \url{https://CRAN.R-project.org/package=doParallel}.
-\newline Klik M (2019).
-\emph{fst: Lightning Fast Serialization of Data Frames for R}.
-R package version 0.9.0, \url{https://CRAN.R-project.org/package=fst}.
-\newline Xie Y, Cheng J, Tan X (2019).
-\emph{DT: A Wrapper of the JavaScript Library 'DataTables'}.
-R package version 0.7, \url{https://CRAN.R-project.org/package=DT}.
-\newline Tennekes M (2017).
-\emph{treemap: Treemap Visualization}.
-R package version 2.4-2, \url{https://CRAN.R-project.org/package=treemap}.
-\newline
+\begin{table}[H]
+    \caption{R-packages used for the compilation of the Standartox database.}
+    \label{tab:rpackages}
+    \centering
+\begin{tabular}{cc}
+    \toprule
+    \textbf{Title} & \textbf{Citation} \\ 
+    \midrule
+    devtools: Tools to Make Developing R Packages Easier & \citep{wickham_devtools_2019} \\
+    RCurl: General Network HTTP FTP Client Interface for R & \citep{lang_rcurl_2020} \\
+    stringr: Simple Consistent Wrappers for Common String Operations & \citep{wickham_stringr_2019} \\
+    R: Various Programming Utilities & \citep{bengtsson_r_2019} \\
+    udunits: Udunits 2 Bindings for R & \citep{hiebert_udunits_2016} \\
+    rvest: Easily Harvest Scrape Web Pages & \citep{wickham_rvest_2019} \\
+    httr: Tools for Working with URLs and HTTP & \citep{wickham_httr_2019} \\
+    The: A Practical and Consistent Mapping Between JSON Data and R Objects & \citep{ooms_the_2014} \\
+    readxl: Read Excel Files & \citep{wickham_readxl_2019} \\
+    openxlsx: Read Write and Edit xlsx Files & \citep{schauberger_openxlsx_2019} \\
+    purrr: Functional Programming Tools & \citep{henry_purrr_2019} \\ data: Extension of data frame & \citep{dowle_data_2019} \\
+    RPostgreSQL: R Interface to the PostgreSQL Database System & \citep{conway_rpostgresql_2017} \\ DBI: R Database Interface & \citep{b_dbi_2019} \\
+    vegan: Community Ecology Package & \citep{oksanen_vegan_2019} \\ The: NA & \citep{wickham_the_2011} \\
+    outliers: Tests for outliers & \citep{komsta_outliers_2011} \\ feather: R Bindings to the Feather API & \citep{wickham_feather_2019} \\
+    ggplot: Elegant Graphics for Data Analysis & \citep{wickham_ggplot_2016} \\ ggrepel: Automatically Position Non Overlapping Text Labels with & \citep{slowikowski_ggrepel_2019} \\
+    ggridges: Ridgeline Plots in ggplot2 & \citep{wilke_ggridges_2018} \\
+    scales: Scale Functions for Visualization & \citep{wickham_scales_2019} \\ treemapify: Draw Treemaps in ggplot2 & \citep{wilkins_treemapify_2019} \\ cowplot: Streamlined Plot Theme and Plot Annotations for ggplot2 & \citep{wilke_cowplot_2019} \\
+    RColorBrewer: ColorBrewer Palettes & \citep{neuwirth_rcolorbrewer_2014} \\ treemap: Treemap Visualization & \citep{tennekes_treemap_2017} \\ rgbif: Interface to the Global Biodiversity Information Facility API & \citep{chamberlain_rgbif_2020} \\
+    taxize: NA & \citep{chamberlain_taxize_2013} \\ countrycode: An R package to convert country names and country codes & \citep{bundock_countrycode_2018} \\ foreach: Provides Foreach Looping Construct for R & \citep{microsoft_foreach_2017} \\
+    doParallel: Foreach Parallel Adaptor for the parallel Package & \citep{corporation_doparallel_2018} \\
+    fst: Lightning Fast Serialization of Data Frames for R & \citep{klik_fst_2019} \\
+    DT: A Wrapper of the JavaScript Library DataTables & \citep{xie_dt_2019} \\
+    knitr: A General Purpose Package for Dynamic Report Generation in R & \citep{xie_knitr_2020} \\
+    \bottomrule
+\end{tabular}
+\end{table}
 

### 2020-04-06 article/article.tex
Updates from Overleaf

@@
 \usepackage{rotating}
+% check references
+\usepackage{refcheck}
 
@@
 \midrule
-    devtools: Tools to Make Developing R Packages Easier & \citep{wickham_devtools_2019} \\
-    RCurl: General Network HTTP FTP Client Interface for R & \citep{lang_rcurl_2020} \\
-    stringr: Simple Consistent Wrappers for Common String Operations & \citep{wickham_stringr_2019} \\
-    R: Various Programming Utilities & \citep{bengtsson_r_2019} \\
-    udunits: Udunits 2 Bindings for R & \citep{hiebert_udunits_2016} \\
-    rvest: Easily Harvest Scrape Web Pages & \citep{wickham_rvest_2019} \\
-    httr: Tools for Working with URLs and HTTP & \citep{wickham_httr_2019} \\
-    The: A Practical and Consistent Mapping Between JSON Data and R Objects & \citep{ooms_the_2014} \\
-    readxl: Read Excel Files & \citep{wickham_readxl_2019} \\
-    openxlsx: Read Write and Edit xlsx Files & \citep{schauberger_openxlsx_2019} \\
-    purrr: Functional Programming Tools & \citep{henry_purrr_2019} \\ data: Extension of data frame & \citep{dowle_data_2019} \\
-    RPostgreSQL: R Interface to the PostgreSQL Database System & \citep{conway_rpostgresql_2017} \\ DBI: R Database Interface & \citep{b_dbi_2019} \\
-    vegan: Community Ecology Package & \citep{oksanen_vegan_2019} \\ The: NA & \citep{wickham_the_2011} \\
-    outliers: Tests for outliers & \citep{komsta_outliers_2011} \\ feather: R Bindings to the Feather API & \citep{wickham_feather_2019} \\
-    ggplot: Elegant Graphics for Data Analysis & \citep{wickham_ggplot_2016} \\ ggrepel: Automatically Position Non Overlapping Text Labels with & \citep{slowikowski_ggrepel_2019} \\
-    ggridges: Ridgeline Plots in ggplot2 & \citep{wilke_ggridges_2018} \\
-    scales: Scale Functions for Visualization & \citep{wickham_scales_2019} \\ treemapify: Draw Treemaps in ggplot2 & \citep{wilkins_treemapify_2019} \\ cowplot: Streamlined Plot Theme and Plot Annotations for ggplot2 & \citep{wilke_cowplot_2019} \\
-    RColorBrewer: ColorBrewer Palettes & \citep{neuwirth_rcolorbrewer_2014} \\ treemap: Treemap Visualization & \citep{tennekes_treemap_2017} \\ rgbif: Interface to the Global Biodiversity Information Facility API & \citep{chamberlain_rgbif_2020} \\
-    taxize: NA & \citep{chamberlain_taxize_2013} \\ countrycode: An R package to convert country names and country codes & \citep{bundock_countrycode_2018} \\ foreach: Provides Foreach Looping Construct for R & \citep{microsoft_foreach_2017} \\
-    doParallel: Foreach Parallel Adaptor for the parallel Package & \citep{corporation_doparallel_2018} \\
-    fst: Lightning Fast Serialization of Data Frames for R & \citep{klik_fst_2019} \\
-    DT: A Wrapper of the JavaScript Library DataTables & \citep{xie_dt_2019} \\
-    knitr: A General Purpose Package for Dynamic Report Generation in R & \citep{xie_knitr_2020} \\
+bib2df: Parse a BibTeX File to a Data Frame & \citep{R-bib2df} \\
+countrycode: Convert Country Names and Country Codes & \citep{R-countrycode} \\
+cowplot: Streamlined Plot Theme and Plot Annotations for 'ggplot2' & \citep{R-cowplot} \\
+data.table: Extension of `data.frame` & \citep{R-data.table} \\
+DBI: R Database Interface & \citep{R-DBI} \\
+dbreport: Automated reports from tables & \citep{R-dbreport} \\
+devtools: Tools to Make Developing R Packages Easier & \citep{R-devtools} \\
+doParallel: Foreach Parallel Adaptor for the 'parallel' Package & \citep{R-doParallel} \\
+DT: A Wrapper of the JavaScript Library 'DataTables' & \citep{R-DT} \\
+foreach: Provides Foreach Looping Construct for R & \citep{R-foreach} \\
+fst: Lightning Fast Serialization of Data Frames for R & \citep{R-fst} \\
+ggplot2: Create Elegant Data Visualisations Using the Grammar of Graphics & \citep{R-ggplot2} \\ httr: Tools for Working with URLs and HTTP & \citep{R-httr} \\
+jsonlite: A Robust, High Performance JSON Parser and Generator for R & \citep{R-jsonlite} \\
+knitr: A General-Purpose Package for Dynamic Report Generation in R & \citep{R-knitr} \\
+openxlsx: Read, Write and Edit xlsx Files & \citep{R-openxlsx} \\ plotly: Create Interactive Web Graphics via 'plotly.js' & \citep{R-plotly} \\ plumber: An API Generator for R & \citep{R-plumber} \\
+R.utils: Various Programming Utilities & \citep{R-R.utils} \\
+RColorBrewer: ColorBrewer Palettes & \citep{R-RColorBrewer} \\
+reactlog: Reactivity Visualizer for 'shiny' & \citep{R-reactlog} \\
+readxl: Read Excel Files & \citep{R-readxl} \\ rgbif: Interface to the Global 'Biodiversity' Information Facility API & \citep{R-rgbif} \\ RPostgreSQL: R Interface to the 'PostgreSQL' Database System & \citep{R-RPostgreSQL} \\ rvest: Easily Harvest (Scrape) Web Pages & \citep{R-rvest} \\ scales: Scale Functions for Visualization & \citep{R-scales} \\ shiny: Web Application Framework for R & \citep{R-shiny} \\ shinydashboard: Create Dashboards with 'Shiny' & \citep{R-shinydashboard} \\ shinydashboardPlus: Add More 'AdminLTE2' Components to 'shinydashboard' & \citep{R-shinydashboardPlus} \\ shinyjs: Easily Improve the User Experience of Your Shiny Apps in Seconds & \citep{R-shinyjs} \\ shinyWidgets: Custom Inputs Widgets for Shiny & \citep{R-shinyWidgets} \\ ssdtools: Species Sensitivity Distributions & \citep{R-ssdtools} \\ stringi: Character String Processing Facilities & \citep{R-stringi} \\ stringr: Simple, Consistent Wrappers for Common String Operations & \citep{R-stringr} \\ taxize: Taxonomic Information from Around the Web & \citep{R-taxize} \\ treemap: Treemap Visualization & \citep{R-treemap} \\ treemapify: Draw Treemaps in 'ggplot2' & \citep{R-treemapify} \\ udunits2: Udunits-2 Bindings for R & \citep{R-udunits2} \\ webchem: Chemical Information from the Web & \citep{R-webchem} \\ countrycode: An R package to convert country names and country codes & \citep{R-countrycode} \\ ggplot2: Elegant Graphics for Data Analysis & \citep{R-ggplot2} \\ The jsonlite Package: A Practical and Consistent Mapping Between JSON Data and R Objects & \citep{R-The jsonlite Package} \\ Dynamic Documents with {R} and knitr & \citep{R-Dynamic Documents with {R} and knitr} \\ knitr: A Comprehensive Tool for Reproducible Research in R & \citep{R-knitr} \\ plotly for R & \citep{R-plotly for R} \\ R Python, and Ruby clients for GBIF species occurrence data & \citep{R-R Python, and Ruby clients for GBIF species occurrence data} \\ R package stringi: Character string processing facilities & \citep{R-R package stringi} \\ taxize - taxonomic search and retrieval in R & \citep{R-taxize - taxonomic search and retrieval in R} \\ webchem:  retrieve chemical information from the web & \citep{R-webchem}
+\\
 \bottomrule
@@
 \reftitle{References}
+\nocite{*}
 \externalbibliography{yes}
-\bibliography{refs/references-standartox.bib}
+\bibliography{refs/references-standartox.bib,
+              refs/references-standartox-rpackages.bib}
 
@@
 %% optional
-\sampleavailability{Samples of the compounds ...... are available from the authors.}
+% \sampleavailability{Samples of the compounds ...... are available from the authors.}
 

### 2020-04-06 article/article.tex
Updates from Overleaf

--- 2020-04-03 article/article.tex
@@
     \centering
-\begin{tabular}{cc}
+\begin{tabular}{ccc}
 \toprule
-    \textbf{Title} & \textbf{Citation} \\ 
+\textbf{Package} & \textbf{Description} & \textbf{Citation} \\ 
 \midrule
-    devtools: Tools to Make Developing R Packages Easier & \citep{wickham_devtools_2019} \\
-    RCurl: General Network HTTP FTP Client Interface for R & \citep{lang_rcurl_2020} \\
-    stringr: Simple Consistent Wrappers for Common String Operations & \citep{wickham_stringr_2019} \\
-    R: Various Programming Utilities & \citep{bengtsson_r_2019} \\
-    udunits: Udunits 2 Bindings for R & \citep{hiebert_udunits_2016} \\
-    rvest: Easily Harvest Scrape Web Pages & \citep{wickham_rvest_2019} \\
-    httr: Tools for Working with URLs and HTTP & \citep{wickham_httr_2019} \\
-    The: A Practical and Consistent Mapping Between JSON Data and R Objects & \citep{ooms_the_2014} \\
-    readxl: Read Excel Files & \citep{wickham_readxl_2019} \\
-    openxlsx: Read Write and Edit xlsx Files & \citep{schauberger_openxlsx_2019} \\
-    purrr: Functional Programming Tools & \citep{henry_purrr_2019} \\ data: Extension of data frame & \citep{dowle_data_2019} \\
-    RPostgreSQL: R Interface to the PostgreSQL Database System & \citep{conway_rpostgresql_2017} \\ DBI: R Database Interface & \citep{b_dbi_2019} \\
-    vegan: Community Ecology Package & \citep{oksanen_vegan_2019} \\ The: NA & \citep{wickham_the_2011} \\
-    outliers: Tests for outliers & \citep{komsta_outliers_2011} \\ feather: R Bindings to the Feather API & \citep{wickham_feather_2019} \\
-    ggplot: Elegant Graphics for Data Analysis & \citep{wickham_ggplot_2016} \\ ggrepel: Automatically Position Non Overlapping Text Labels with & \citep{slowikowski_ggrepel_2019} \\
-    ggridges: Ridgeline Plots in ggplot2 & \citep{wilke_ggridges_2018} \\
-    scales: Scale Functions for Visualization & \citep{wickham_scales_2019} \\ treemapify: Draw Treemaps in ggplot2 & \citep{wilkins_treemapify_2019} \\ cowplot: Streamlined Plot Theme and Plot Annotations for ggplot2 & \citep{wilke_cowplot_2019} \\
-    RColorBrewer: ColorBrewer Palettes & \citep{neuwirth_rcolorbrewer_2014} \\ treemap: Treemap Visualization & \citep{tennekes_treemap_2017} \\ rgbif: Interface to the Global Biodiversity Information Facility API & \citep{chamberlain_rgbif_2020} \\
-    taxize: NA & \citep{chamberlain_taxize_2013} \\ countrycode: An R package to convert country names and country codes & \citep{bundock_countrycode_2018} \\ foreach: Provides Foreach Looping Construct for R & \citep{microsoft_foreach_2017} \\
-    doParallel: Foreach Parallel Adaptor for the parallel Package & \citep{corporation_doparallel_2018} \\
-    fst: Lightning Fast Serialization of Data Frames for R & \citep{klik_fst_2019} \\
-    DT: A Wrapper of the JavaScript Library DataTables & \citep{xie_dt_2019} \\
-    knitr: A General Purpose Package for Dynamic Report Generation in R & \citep{xie_knitr_2020} \\
+bib2df & Parse a BibTeX File to a Data Frame & \citep{R-bib2df} \\
+countrycode & Convert Country Names and Country Codes & \citep{R-countrycode} \\
+cowplot & Streamlined Plot Theme and Plot Annotations for 'ggplot2' & \citep{R-cowplot} \\
+data.table & Extension of `data.frame` & \citep{R-data.table} \\
+DBI & R Database Interface & \citep{R-DBI} \\
+dbreport & Automated reports from tables & \citep{R-dbreport} \\
+devtools & Tools to Make Developing R Packages Easier & \citep{R-devtools} \\
+doParallel & Foreach Parallel Adaptor for the 'parallel' Package & \citep{R-doParallel} \\
+DT & A Wrapper of the JavaScript Library 'DataTables' & \citep{R-DT} \\
+foreach & Provides Foreach Looping Construct for R & \citep{R-foreach} \\
+fst & Lightning Fast Serialization of Data Frames for R & \citep{R-fst} \\
+ggplot2 & Create Elegant Data Visualisations Using the Grammar of Graphics & \citep{R-ggplot2} \\ httr & Tools for Working with URLs and HTTP & \citep{R-httr} \\
+jsonlite & A Robust, High Performance JSON Parser and Generator for R & \citep{R-jsonlite} \\
+knitr & A General-Purpose Package for Dynamic Report Generation in R & \citep{R-knitr} \\
+openxlsx & Read, Write and Edit xlsx Files & \citep{R-openxlsx} \\
+plotly & Create Interactive Web Graphics via 'plotly.js' & \citep{R-plotly} \\
+plumber & An API Generator for R & \citep{R-plumber} \\
+R.utils & Various Programming Utilities & \citep{R-R.utils} \\
+RColorBrewer & ColorBrewer Palettes & \citep{R-RColorBrewer} \\
+reactlog & Reactivity Visualizer for 'shiny' & \citep{R-reactlog} \\
+readxl & Read Excel Files & \citep{R-readxl} \\
+rgbif & Interface to the Global 'Biodiversity' Information Facility API & \citep{R-rgbif} \\ RPostgreSQL & R Interface to the 'PostgreSQL' Database System & \citep{R-RPostgreSQL} \\
+rvest & Easily Harvest (Scrape) Web Pages & \citep{R-rvest} \\
+scales & Scale Functions for Visualization & \citep{R-scales} \\
+shiny & Web Application Framework for R & \citep{R-shiny} \\
+shinydashboard & Create Dashboards with 'Shiny' & \citep{R-shinydashboard} \\
+shinydashboardPlus & Add More 'AdminLTE2' Components to 'shinydashboard' & \citep{R-shinydashboardPlus} \\
+shinyjs & Easily Improve the User Experience of Your Shiny Apps in Seconds & \citep{R-shinyjs} \\ shinyWidgets & Custom Inputs Widgets for Shiny & \citep{R-shinyWidgets} \\
+stringi & Character String Processing Facilities & \citep{R-stringi} \\
+stringr & Simple, Consistent Wrappers for Common String Operations & \citep{R-stringr} \\
+taxize & Taxonomic Information from Around the Web & \citep{R-taxize} \\
+treemap & Treemap Visualization & \citep{R-treemap} \\
+treemapify & Draw Treemaps in 'ggplot2' & \citep{R-treemapify} \\
+udunits2 & Udunits-2 Bindings for R & \citep{R-udunits2} \\
+webchem & Chemical Information from the Web & \citep{R-webchem} \\
 \bottomrule

### 2020-04-06 article/article.tex
Updates from Overleaf

@@
 % Full title of the paper (Capitalized)
-\Title{Standartox: Standardising toxicity data}
+\Title{Standartox: Standardizing toxicity data}
 
@@
 % Keywords
-\keyword{ecotoxicology; standardisation; environmental risk assessment; effect database; database; software; R; api;} % up to 10
+\keyword{ecotoxicology; standardization; environmental risk assessment; effect database; database; software; R; api;} % up to 10
 
@@
 
-To evaluate the risks from chemicals to ecosystems, data on their toxicity is required, which is typically produced in standardised ecotoxicological laboratory tests.
+To evaluate the risks from chemicals to ecosystems, data on their toxicity is required, which is typically produced in standardized ecotoxicological laboratory tests.
 For example, \citet{morrissey_neonicotinoid_2015} used ecotoxicological test results from 49 insects and crustaceans to evaluate the effect of neonicotinoid insecticides in the aquatic ecosystem.
@@
 
-We therefore developed Standartox, a tool and database that aims to overcome the limitations of other databases by continuously incorporating the ever-growing number of test results in an automated process workflow that ultimately leads to a harmonised ecotoxicity data collection and provides methods to derive single aggregated ecotoxicity values for a specific chemical-organism test combination.
+We therefore developed Standartox, a tool and database that aims to overcome the limitations of other databases by continuously incorporating the ever-growing number of test results in an automated process workflow that ultimately leads to a harmonized ecotoxicity data collection and provides methods to derive single aggregated ecotoxicity values for a specific chemical-organism test combination.
 Standartox makes use of the publicly available and quarterly updated ECOTOX database \citep{usepa_ecotox_2019} and restricts the data to commonly used endpoints in ecotoxicology, such as half maximal effective concentration (EC\textsubscript{50}) or no-observed-adverse-effect concentrations (NOEC), leading to about 600,000 ecotoxicological test results including 8000 chemicals, tested on about 8000 taxa.
@@
 Standartox constitutes a collection of quality checked ecotoxicological test results.
-It is build on the ECOTOX database \citep{usepa_ecotox_2019} whose data are processed, cleaned and harmonised to retrieve comparable toxicity endpoints.
+It is build on the ECOTOX database \citep{usepa_ecotox_2019} whose data are processed, cleaned and harmonized to retrieve comparable toxicity endpoints.
 Subsequently, filter and aggregation methods are created to allow for the retrieval of single toxicity equivalents for specific experimental conditions.
@@
 However, they are considered in aggregation, given that the geometric mean is relatively robust against outliers.
-Overall, Standartox provides a harmonised and reproducible approach to data aggregation that may contribute to reduce between-study variability.
+Overall, Standartox provides a harmonized and reproducible approach to data aggregation that may contribute to reduce between-study variability.
 
@@
 through conversion of test concentration and duration units.
-1237 distinct concentration units are converted to X harmonised ones, as are the 126 distinct duration units to X.
+1237 distinct concentration units are converted to X harmonized ones, as are the 126 distinct duration units to X.
 Standartox retains only duration units that can unambiguously be converted to hours.

### 2020-04-07 article/article.tex
update

@@
 % Only for the journal Data:
-\dataset{https://cfpub.epa.gov/ecotox/}
 %\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately.
@@
 We therefore developed Standartox, a tool and database that aims to overcome the limitations of other databases by continuously incorporating the ever-growing number of test results in an automated process workflow that ultimately leads to a harmonized ecotoxicity data collection and provides methods to derive single aggregated ecotoxicity values for a specific chemical-organism test combination.
-Standartox makes use of the publicly available and quarterly updated ECOTOX database \citep{usepa_ecotox_2019} and restricts the data to commonly used endpoints in ecotoxicology, such as half maximal effective concentration (EC\textsubscript{50}) or no-observed-adverse-effect concentrations (NOEC), leading to about 600,000 ecotoxicological test results including 8000 chemicals, tested on about 8000 taxa.
+Standartox makes use of the publicly available and quarterly updated ECOTOX database \citep{usepa_ecotox_2019} and restricts the data to commonly used endpoints in ecotoxicology, such as half maximal effective concentration (EC\textsubscript{50}) or no-observed-adverse-effect concentrations (NOEC), leading to about 600,000 ecotoxicological test results including about 8000 chemicals, tested on about 10000 taxa in the current version.
 Standartox users can filter test results according to several parameters, e.g.
@@
 Moreover, Standartox could help in reducing the millions of animals used for toxicity testing each year by facilitating access to ecotoxicity data \citep{hartung_chemical_2009}.
-Standartox comes with two front-ends, a web application: \url{standartox.uni-landau.de} and an R \citep{rcoreteam_language_2017} package \textit{standartox}, providing convenience structures and thereby largely reducing processing time for users.
+Standartox comes with two front-ends, a web application: \url{standartox.uni-landau.de} and the R \citep{rcoreteam_language_2017} package \textit{standartox}, providing convenience structures and thereby largely reducing processing time for users.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Data Description - required}
+\section{Data Description}
 Standartox constitutes a collection of quality checked ecotoxicological test results.
@@
 These initiatives partly aim for overlapping goals, yet have limitations or objectives that distinguish them from Standartox:
-
-\textbf{Comptox}, is a web tool published by the EPA which, similar to Standartox allows for filtering test results, the retrieval of additional chemical information and the estimation of toxicity data, such as 48 h \textit{Daphnia magna} LC\textsubscript{50} values.
+\par
+\textbf{Comptox}, is a web tool published by the EPA which, similar to Standartox allows for filtering test results, the retrieval of additional chemical information as well as predicted toxicity data \citep{martin_prediction_2001}, such as 48 h \textit{Daphnia magna} LC\textsubscript{50} values.
 However, toxicity estimations are limited to standard test organisms (e.g.
-\textit{Daphnia magna}), and the tool lacks the possibility for automated retrieval \citep{williams_comptox_2017}.
-
+\textit{Daphnia magna}), and the tool lacks the possibility for automated data retrieval \citep{williams_comptox_2017}.
+\par
 The \textbf{EnviroTox} database which also uses, amongst others the ECOTOX database as an input has recently been published \citep{healthandenvironmentalsciencesinstitutehesi_envirotox_2019, connors_creation_2019}.
@@
 However, the Standartox results for individual taxa-chemical combinations could easily be aggregated across chemicals in a second step to provide a similar aggregation as that performed in EnviroTox.
-
-The \textbf{Etox} database, like the ECOTOX database collects and provides filter methods for ecotoxicological test information.
-
-The \textbf{PPDB} provides data only on pesticides and as mentioned before provides single quality controlled values only for commonly used taxa, e.g. \textit{D. magna} or \textit{R. subcapitata}.
+Since Standartox aims to provide toxicity estimations for specific taxa and chemicals we refrain from aggregating further.
+\par
+The \textbf{Etox} database collects ecotoxicity test information and provides methods to filter those.
+Like the ECOTOX database, it also doesn't provide methods to perform aggregations of the ecotoxicity data and the database is not accessible in automated way.
+In contrast to the ECOTOX database it is not possible to download the whole database.
+\par
+The \textbf{PPDB} provides data only on pesticides, and as mentioned before, it provides single quality controlled values only for commonly used taxa, e.g.
+\textit{D.
+magna} or \textit{R.
+subcapitata}.
+\par
+In summary, none of the above mentioned initiatives aim for an automated and standardized aggregation method of exposure endpoints for individual chemicals.
+In addition, they lack the possibility to access the databases through a common high level programming language, such as R.
 
@@
     \caption{Overview on databases that provide ecotoxicological data.
-Abbreviations: \textbf{ALL:} Most important test parameters, including chemical, taxon, duration for filtering ecotoxicological data are incorporated.
-\textbf{WEB:} Accessible via a web application through a graphical user interface.
-\textbf{API:} Accessible via an application programming interface.}
+Abbreviations: ALL: Most important test parameters, including chemical, taxon, duration for filtering ecotoxicological data are incorporated.
+Web: Accessible via a web application through a graphical user interface.
+API: Accessible via an application programming interface.}
     \label{tab:database-differences}
     \centering
-    \begin{tabular}{m{3cm}m{3cm}m{2cm}m{2cm}m{1cm}}
+    \begin{tabular}{ccccc}
     \toprule
-    \textbf{Database} & \textbf{Publisher} & \textbf{Filter} & \textbf{Aggregation, Selection} & \textbf{Access} \\
+    \textbf{Database} & \textbf{Filter} & \textbf{Aggregation, Selection} & \textbf{Access} \\
     \midrule
-    Comptox & Environmental Protection Agency & Chemical & no & WEB, file \\
-    Ecotox & Environmental Protection Agency & ALL & no & WEB, file \\
-    EnviroTox & Health and Environmental Sciences Institute & ALL & chemical, organism \\
-    Etox & Umweltbundesamt & ALL & no & WEB \\
-    Pesticides Properties DataBase (PPDB) & University of Hertfordshire & fixed values & manual selection & WEB, file \\
-    Standartox & this article & ALL & chemical, organism & API, WEB \\
+    Comptox \citep{williams_comptox_2017} & Chemical & no & Web, file \\
+    Ecotox \citep{usepa_ecotox_2019} & ALL & no & Web, file \\
+    EnviroTox \citep{healthandenvironmentalsciencesinstitutehesi_envirotox_2019} & ALL & chemical, organism & Web \\
+    Etox \citep{umweltbundesamt_etox_2019} & ALL & no & Web \\
+    Pesticides Properties DataBase (PPDB) \citep{lewis_international_2016} & fixed values & manual selection & Web, file \\
+    Standartox & ALL & chemical, organism & API, Web \\
     \bottomrule
@@
 
-In summary, none of the above mentioned initiatives aim for an automated and standardized aggregation method of exposure endpoints for individual chemicals.
-In addition, they lack the possibility to access the databases through a common high level programming language, such as R.
+\par
 As outlined above, toxicity estimates from different studies can vary strongly due to a wide range of experimental conditions such as pH, temperature and conductivity \citep{rosenkrantz_influence_2013, li_temperature_2011}.
@@
 The former contains code to process the data and to build the web application and the API, the latter contains code to build the R-package.
-Most of the code is written in R 3.6.1 and associated packages (List: \ref{lis:r-packages}) and SQL for PostgreSQL 9.6.1.
+Most of the code is written in R 3.6.1 and associated packages (List: \ref{tab:rpackages}) and SQL for PostgreSQL 9.6.1.
 
 \subsection{Processing}
-Standartox downloads the quarterly released ECOTOX database and builds it into a local PostgreSQL database \citep{szocs_build_2019}.
+Standartox downloads the quarterly released ECOTOX database and builds it into a local PostgreSQL database.
 Subsequently, Structured Query Language (SQL) functions for further processing the data are implemented.
@@
 This includes the Compendium of Pesticide Common Names \citep{wood_compendium_2019}, the Chemical Entities of Biological Interest (ChEBI) database \citep{hastings_chebi_2016}, the Chemical Identifier Resolver (CIR) service \citep{nationalinstitutesofhealthnih_chemical_2019}, the Pubchem database \citep{kim_pubchem_2016}, Eurostat \citep{europeancommission_eurostat_2019} and Wikidata \citep{vrandecic_wikidata_2014} for chemicals and the World Register of Marine Species (WoRMS) \citep{wormseditorialboard_world_2018} and the Global Biodiversity Information Facility (GBIF) \citep{gbif_gbif_2019} (Table~\ref{tab:data-base-additional}) for habitat and spatial distribution of organisms.
-Given that CAS and taxonomic names can be ambiguous, e.g.
-the genus \textit{Eisenia} can refer to an algae and a worm, we first match them against the database specific identifiers and subsequently use them to retrieve the actual data.
+Given that taxonomic names can be ambiguous, e.g.
+the genus \textit{Eisenia} can refer to an algae and a worm, we first match the taxa names against specific database identifiers (e.g.
+WoRMS \citep{wormseditorialboard_world_2020}) and subsequently check their accordance with the underlying ECOTOX data taxonomy.
+Then we query the actual data by using the identifiers.
 In a next step the data is added to Standartox to enable filtering for specific chemical roles (e.g.
@@
 Standartox retains only duration units that can unambiguously be converted to hours.
-To guarantee appropriate unit conversion and harmonisation, we compared an automated to a manual conversion for each of the distinct concentration and duration units.
+To guarantee appropriate unit conversion and harmonisation, we compared the results of an automated unit conversion to a manual one for each of the distinct concentration and duration units.
 Furthermore, the units are cleaned, for example through removing additional information in the field such as \textit{food}, \textit{soil}, \textit{ai} that are also coded in other variables and hinder the processing of units.
@@
 Other endpoints, such as Bioconcentration factors, non-half maximal effective concentrations (e.g.
-IC10, EC25, LD99) or maximum acceptable toxicant (MATC) concentrations are removed.
+IC10, EC25, LD99) or maximum acceptable toxicant concentrations are removed.
 Along with that, a catalog, listing all distinct entries and value ranges, for categorical and continuous variables, respectively, is created.
@@
 The ECOTOX knowledge base is downloaded quarterly and processed (i.e.
-query additional information with CAS and taxa name and conversion of concentration and duration units.
+query additional information with CAS and taxa names and conversion of concentration and duration units.
 Subsequently, a Standartox data set is compiled together with filter and aggregation methods.
@@
 When accessed, the web application and the API load the compressed serialized Standartox data into memory and allow the user to interact with it.
-The user can then call the functions stx\_filter() and stx\_aggregate() that filter and aggregate the data according to specific parameters (Table~\ref{tab:app-parameters}).
+The user can then call the functions \textit{stx\_filter()} and \textit{stx\_aggregate()} that filter and aggregate the data according to specific parameters (Table~\ref{tab:app-parameters}).
 The interactive web application is built in R using the shiny framework, which runs with the help of a shiny server \citep{chang_shiny_2018} on \url{standartox.uni-landau.de}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{User Notes - optional}
+\section{User Notes}
 Users can access Standartox either via the web application \url{standartox.uni-landau.de} or via the R-package \textit{standartox}.
@@
 It returns an R list of three tables (i.e.
-R data.frames) containing the filtered data set, the aggregated data set and a table with the meta information retrieved from the API endpoints.
+R \textit{data.frame}s) containing the filtered data set, the aggregated data set and a table with the meta information retrieved from the API endpoints.
 A short R-code example is given below (Listing \ref{lis:standartox-example}) and a detailed description on the usage of the R package is provided on its Github page: \url{https://github.com/andschar/standartox}.
@@
 \begin{lstlisting}[
-    caption = {Sample code to access the Standartox data base through the API with the help of the R-package.
+    caption = {Sample code to access the Standartox database through the API and the the R-package \textit{standartox}.
 \textit{stx\_catalog()} returns a catalog of possible filter and aggregation parameters.
-\textit{stx\_query()} returns the Standartox object, a list of the filtered and the aggregated data as well as a meta data entry.},
+\textit{stx\_query()} returns the Standartox object, a list of the filtered and the aggregated data as well as a meta data entry.
+Example for XX50 tests on the compound glyphosate (CAS: 1071-83-6) and the taxon \textit{Oncorhynchus} lasting 24 h to 120 h.},
     label = {lis:standartox-example}]
-
 # install
 install.packages('remotes')
-remotes::install_github('andschar/standartox') # package not yet on CRAN
+remotes::install_github('andschar/standartox')
 # retrieve catalog    
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusion - optional}
+\section{Conclusion}
 % This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.
@@
 We regard this as an advantage rather than a drawback because other published works that aim in a similar direction often constitute a singular effort or require manual work for each update.
-Standartox, in contrast automates the update process, yet still provides access to its older versions, assuring reproducability and version control.
+Standartox, in contrast automates the update process, yet still provides access to its older versions, assuring reproducibility and version control.
 In comparison to rule-based approaches for the derivation of single ecotoxicity values, Standartox has the advantage to be free from the subjectivity of a set of human-induced rules.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{conceptualization and methods, A.S.,R.B.S.
+\authorcontributions{conceptualization and methods, A.S., R.B.S.
 and V.C.S.; software development and validation, A.S.; data curation and analysis, A.S.; writing--original draft preparation, A.S.; writing--review and editing, A.S., V.C.S.
 and R.B.S.; supervision, R.B.S.}
+% \authorcontributions{conceptualization, A.S., R.B.S.
+and V.C.S.; methodology, A.S.; software, A.S.; validation, A.S., Y.Y.
+and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing--original draft preparation, X.X.; writing--review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y.''}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections.
-This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}
+\acknowledgments{The authors want to thank Eduard Szöcs whose blog post (https://edild.github.io/localecotox/) on how to build a local version of the EPA ECOTOX database was the starting point for this project.}
 
@@
 \conflictsofinterest{The authors declare no conflict of interest.} 
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{table}[H]
-    \caption{Table of additionally queried publicly available data bases and their URLs.}
+    \caption{Table of additionally queried publicly available databases and their URLs.}
     \label{tab:data-base-additional}
@@
     \toprule
-    \textbf{Data Base} & \textbf{URL} \\ 
+    \textbf{Database} & \textbf{URL} \\ 
     \midrule
@@
 \begin{table}[H]
-    \caption{Application programming interface (API) endpoints, Http methods, Requests and Response objects. JSON - Javascript opject notation file}
+    \caption{Application programming interface (API) endpoints, HTTP methods, Requests and Response objects. JSON - Javascript opject notation file}
     \label{tab:api-endpoints}

### 2020-04-07 article/tikz/standartox-organigram.tex
update

@@
 \node (cas) [ids, below of = ecotox-id, xshift = 6cm, yshift = 3cm] {\Large CAS};
-\node (taxa) [ids, below of = cas, xshift = 0cm, yshift = -2.5cm] {\Large Taxa};
+\node (taxa) [ids, below of = cas, xshift = 0cm, yshift = -2.5cm] {\Large Taxon};
 \node (db-chemical) [db, right of = cas, xshift = 4cm, align = center] {\Large Chemical\\\Large databases};
@@
 %% standartox
-\node (stx-data) [standartox, below of = ecotox-chck, xshift = 3cm, yshift = -2.5cm, align = center] {\Large Standartox\\\Large data set};
+\node (stx-data) [standartox, below of = ecotox-chck, xshift = 3cm, yshift = -2.5cm, align = center] {\Large Standartox\\\Large database};
 \node (stx-filter) [standartox, below of = stx-data, xshift = 0cm, yshift = -2cm, align = center] {\Large Filter\\\Large methods};

### 2020-04-08 article/article.tex

@@
 Then, provided Chemical Abstracts Service Numbers (CAS) and taxonomic names are used to query additional information from publicly available databases on chemicals and organisms, respectively.
-This includes the Compendium of Pesticide Common Names \citep{wood_compendium_2019}, the Chemical Entities of Biological Interest (ChEBI) database \citep{hastings_chebi_2016}, the Chemical Identifier Resolver (CIR) service \citep{nationalinstitutesofhealthnih_chemical_2019}, the Pubchem database \citep{kim_pubchem_2016}, Eurostat \citep{europeancommission_eurostat_2019} and Wikidata \citep{vrandecic_wikidata_2014} for chemicals and the World Register of Marine Species (WoRMS) \citep{wormseditorialboard_world_2018} and the Global Biodiversity Information Facility (GBIF) \citep{gbif_gbif_2019} (Table~\ref{tab:data-base-additional}) for habitat and spatial distribution of organisms.
+This includes the Compendium of Pesticide Common Names \citep{wood_compendium_2019}, the Chemical Entities of Biological Interest (ChEBI) database \citep{hastings_chebi_2016}, the Chemical Identifier Resolver (CIR) service \citep{nationalinstitutesofhealthnih_chemical_2019}, the Pubchem database \citep{kim_pubchem_2016}, Eurostat \citep{europeancommission_eurostat_2019} and Wikidata \citep{vrandecic_wikidata_2014} for chemicals and the World Register of Marine Species (WoRMS) \citep{wormseditorialboard_world_2018} and the Global Biodiversity Information Facility (GBIF) \citep{gbif_gbif_2020} (Table~\ref{tab:data-base-additional}) for habitat and spatial distribution of organisms.
 Given that taxonomic names can be ambiguous, e.g.

### 2020-04-09 article/article.tex

@@
 \articlenumber{5}
-\pubyear{2019}
-\copyrightyear{2019}
+\pubyear{2020}
+\copyrightyear{2020}
 %\externaleditor{Academic Editor: name}
@@
 % check references
-\usepackage{refcheck}
+% \usepackage{refcheck} % debuging
 
@@
 % Keywords
-\keyword{ecotoxicology; standardization; environmental risk assessment; effect database; database; software; R; api;} % up to 10
+\keyword{ecotoxicology; standardization; environmental risk assessment; effect database; database; software; R; API;} % up to 10
 
@@
 The data can be restricted to the three endpoint groups, namely half maximal effective/lethal concentration/dose values (e.g.
-EC\textsubscript{50}, LD\textsubscript{50}), henceforth abbreviated as XX50, lowest observed effect levels/concentrations (LOEC/L), henceforth abbreviated as LOEX and no observed effect levels/concentrations (NOEL/C), henceforth abbreviated as NOEX (Table~\ref{tab:endpoints-conflate}).
+EC\textsubscript{50}, LD\textsubscript{50}), henceforth abbreviated as XX\textsubscript{50}, lowest observed effect levels/concentrations (LOEC/L), henceforth abbreviated as LOEX and no observed effect levels/concentrations (NOEL/C), henceforth abbreviated as NOEX (Table~\ref{tab:endpoints-conflate}).
 Standartox allows the ecotoxicity data to be filtered by effect groups (e.g.
@@
 \textit{Xenopus laevis} - Amphibia, \textit{Raphidocelis subcapitata} - Algae, \textit{Oncorhynchus mykiss} - fish, \textit{Lemna minor} - invertebrate) for the chemical atrazine in 96 h tests, (\textbf{B}) how the variability in toxicity tests with zinc sulfate and \textit{Daphnia magna} varies with test duration and (\textbf{C}) high variability that is not explained by the available test characteristics in the case of cupric sulfate tested on \textit{Pimephales promelas} for 96 h.
-Purple dots depict Standartox geometric mean estimates and black dots show raw data for each group.
+Red dots depict Standartox geometric mean estimates and black dots show raw data for each group.
 To facilitate readability, data points are randomly scattered along a hypothetical y-axis and are greyed out if within the violins.}
@@
 Then, provided Chemical Abstracts Service Numbers (CAS) and taxonomic names are used to query additional information from publicly available databases on chemicals and organisms, respectively.
-This includes the Compendium of Pesticide Common Names \citep{wood_compendium_2019}, the Chemical Entities of Biological Interest (ChEBI) database \citep{hastings_chebi_2016}, the Chemical Identifier Resolver (CIR) service \citep{nationalinstitutesofhealthnih_chemical_2019}, the Pubchem database \citep{kim_pubchem_2016}, Eurostat \citep{europeancommission_eurostat_2019} and Wikidata \citep{vrandecic_wikidata_2014} for chemicals and the World Register of Marine Species (WoRMS) \citep{wormseditorialboard_world_2018} and the Global Biodiversity Information Facility (GBIF) \citep{gbif_gbif_2020} (Table~\ref{tab:data-base-additional}) for habitat and spatial distribution of organisms.
+This includes the Compendium of Pesticide Common Names \citep{wood_compendium_2019}, the Chemical Entities of Biological Interest (ChEBI) database \citep{hastings_chebi_2016}, the Chemical Identifier Resolver (CIR) service \citep{nationalinstitutesofhealthnih_chemical_2019}, the Pubchem database \citep{kim_pubchem_2016}, Eurostat \citep{europeancommission_eurostat_2019} and Wikidata \citep{vrandecic_wikidata_2014} for chemicals and the World Register of Marine Species (WoRMS) \citep{wormseditorialboard_world_2018} and the Global Biodiversity Information Facility (GBIF) \citep{gbiftheglobalbiodiversityinformationfacility_what_2020} (Table~\ref{tab:data-base-additional}) for habitat and spatial distribution of organisms.
 Given that taxonomic names can be ambiguous, e.g.
@@
 mg/kg/day) are multiplied by the days of the test and then converted.
-Experimental endpoints are restricted to three groups, namely NOEX, LOEX and XX50.
+Experimental endpoints are restricted to three groups, namely NOEX, LOEX and XX\textsubscript{50}.
 Other endpoints, such as Bioconcentration factors, non-half maximal effective concentrations (e.g.
@@
 \begin{table}[H]
-    \caption{Input parameters for the Standartox web application and the R-package standartox (CAS - Chemical Abstracts Service Registry Number, NOEX and XX50 - Standartox endpoints \ref{tab:endpoints-conflate}), vers - Standartox version).}
+    \caption{Input parameters for the Standartox web application and the R-package standartox (CAS - Chemical Abstracts Service Registry Number, NOEX and XX\textsubscript{50} - Standartox endpoints \ref{tab:endpoints-conflate}), vers - Standartox version).}
     \label{tab:app-parameters}
@@
 \textit{stx\_query()} returns the Standartox object, a list of the filtered and the aggregated data as well as a meta data entry.
-Example for XX50 tests on the compound glyphosate (CAS: 1071-83-6) and the taxon \textit{Oncorhynchus} lasting 24 h to 120 h.},
+Example for XX\textsubscript{50} tests on the compound glyphosate (CAS: 1071-83-6) and the taxon \textit{Oncorhynchus} lasting 24 h to 120 h.},
     label = {lis:standartox-example}]
@@
 E/LC/D\textsubscript{50} &  Half maximal effective/lethal concentration/dose \\
+XX\textsubscript{50} & Summarizes E/LC/D\textsubscript{50} (Table \ref{tab:endpoints-conflate}) \\
 ECOTOX & US EPA Ecotox Knowledgebase \\
@@
 LOEC/L & Lowest observed effect levels/concentration \\
+LOEX & Summarizes LOEC/L (Table \ref{tab:endpoints-conflate}) \\
 NOEC/L & No observed effect levels/concentration \\
+NOEX & Summarizes NOEC/L (Table \ref{tab:endpoints-conflate}) \\
 PPDB & Pesticides Properties DataBase \\
@@
     \midrule
-    XX50 & LC50 & Lethal concentration to 50\% of test organisms \\
-    XX50 & LD50 & Lethal dose to 50\% of test organisms \\
-    XX50 & EC50 & Effective concentration to 50\% of test organisms \\
-    XX50 & ED50 & Effective dose to 50\% of test organisms \\
-    XX50 & IC50 & Inhibition concentration to 50\% of test organisms \\
-    XX50 & ID50 & Inhibition dose to 50\% of test organisms \\
-    XX50 & ET50 & Effective response time to 50\% of test organisms \\
-    XX50 & LT50 & Time to 50\% mortality of test organisms \\
+    XX\textsubscript{50} & LC\textsubscript{50} & Lethal concentration to 50\% of test organisms \\
+    XX\textsubscript{50} & LD\textsubscript{50} & Lethal dose to 50\% of test organisms \\
+    XX\textsubscript{50} & EC\textsubscript{50} & Effective concentration to 50\% of test organisms \\
+    XX\textsubscript{50} & ED\textsubscript{50} & Effective dose to 50\% of test organisms \\
+    XX\textsubscript{50} & IC\textsubscript{50} & Inhibition concentration to 50\% of test organisms \\
+    XX\textsubscript{50} & ID\textsubscript{50} & Inhibition dose to 50\% of test organisms \\
+    XX\textsubscript{50} & ET\textsubscript{50} & Effective response time to 50\% of test organisms \\
+    XX\textsubscript{50} & LT\textsubscript{50} & Time to 50\% mortality of test organisms \\
     NOEX & NOEC & No-observable-effect-concentration \\
@@
 \reftitle{References}
-\nocite{*}
+% \nocite{*} % debuging
 \externalbibliography{yes}

### 2020-04-09 README.Rmd
release 0.1

--- 2019-09-27 README.Rmd
@@
 
-Standartox is a data base tool standartizing the retrieval and aggregation of ecotoxicological test data.
-It is based on the [EPA ECOTOX data base](https://cfpub.epa.gov/ecotox/) as well as on data from several other chemical data bases and allows users to filter and aggregate ecotoxicological test data in an easy way.
+Standartox is a database and tool facilitating the retrieval of ecotoxicological test data.
+It is based on the [EPA ECOTOX data base](https://cfpub.epa.gov/ecotox/) as well as on data from several other chemical databases and allows users to filter and aggregate ecotoxicological test data in an easy way.
 It can either be accessed via a [web-app](http://standartox.uni-landau.de/) or via the R-package [standartox](https://github.com/andschar/standartox).
-Ecotoxicological test data are used in environmental risk assessment to calculate effect measures such as [TU - Toxic Units](https://en.wikipedia.org/wiki/Toxic_unit) or [SSD - Species Sensitivity Distributions](https://edild.github.io/ssd/) to asses toxicity of chemicals.
+Ecotoxicological test data is used in environmental risk assessment to calculate effect measures such as [TU - Toxic Units](https://en.wikipedia.org/wiki/Toxic_unit) or [SSD - Species Sensitivity Distributions](https://edild.github.io/ssd/) to asses environmental toxicity of chemicals.
 
-The project lives in these three repositories:
+The project lives in two repositories:
 
-- [standartox-build](https://github.com/andschar/standartox-build) - Compiles the tool
-- [standartox-app](https://github.com/andschar/standartox-app) - Serves the website and API
+  - [standartox-build](https://github.com/andschar/standartox-build) - Compiles and serves database
   - [standartox](https://github.com/andschar/standartox) - R-package

### 2020-04-09 R/conv_unit_result_duration.R

@@ -1,17 +1,14 @@
-# script aggregates physico-chemical data from 3rd party data bases
+# script to build application data
 
 # setup -------------------------------------------------------------------
 source(file.path(src, 'gn_setup.R'))
 
 # scripts -----------------------------------------------------------------
-q = read_char(file.path(sql, 'phch_fin.sql'))
+q = read_char(file.path(sql, 'conv_unit_result_duration.sql'))
 
 # query -------------------------------------------------------------------
 drv = dbDriver("PostgreSQL")
 con = dbConnect(drv, user = DBuser, dbname = DBetox, host = DBhost, port = DBport, password = DBpassword)
-
-dbSendQuery(con, "DROP SCHEMA IF EXISTS phch_fin CASCADE;")
-dbSendQuery(con, "CREATE SCHEMA phch_fin;")
 
 dbSendQuery(con, q)
 
@@ -19,9 +16,8 @@ dbDisconnect(con)
 dbUnloadDriver(drv)
 
 # log ---------------------------------------------------------------------
-log_msg('QUERY: phch final table created.')
+log_msg('CONV: Unit conversion scripts run.')
 
 # cleaning ----------------------------------------------------------------
 clean_workspace()
-  
-  
+

### 2020-04-09 R/fun_as_true.R

@@ -1,6 +1,6 @@
 # function that converts vector to logical and sets FLASE to NA (hence only TRUE left)
 
 as_true = function(x) {
-  x = as.logical(x)
-  ifelse(isTRUE(x), x, NA)
+  x = as.logical(as.numeric(x))
+  ifelse(x == TRUE, x, NA)
 }

### 2020-04-09 R/look_unit_duration.R

@@ -31,7 +31,7 @@ chck_dupl(units, 'obs_duration_unit')
 
 # write -------------------------------------------------------------------
 write_tbl(units, user = DBuser, host = DBhost, port = DBport, password = DBpassword,
-          dbname = DBetox, schema = 'lookup', tbl = 'duration_unit_lookup',
+          dbname = DBetox, schema = 'lookup', tbl = 'lookup_unit_duration',
           key = 'obs_duration_unit',
           comment = 'Lookup table for duration units')
 

### 2020-04-09 R/look_unit_result.R

@@ -231,16 +231,16 @@ chck_dupl(unit2, 'conc1_unit')
 
 # write -------------------------------------------------------------------
 ## csv
-fwrite(unit2, file.path(summdir, 'concentration_unit_lookup_summary.csv'))
+fwrite(unit2, file.path(summdir, 'lookup_result_summary.csv'))
 
 ## postgres
 write_tbl(unit2, user = DBuser, host = DBhost, port = DBport, password = DBpassword,
-          dbname = DBetox, schema = 'lookup', tbl = 'concentration_unit_lookup',
+          dbname = DBetox, schema = 'lookup', tbl = 'lookup_unit_result',
           key = 'conc1_unit',
           comment = 'Lookup table for concentration units')
 
 # log ---------------------------------------------------------------------
-log_msg('LOOK: Concentration lookup tables script run.')
+log_msg('LOOK: Result lookup tables script run.')
 
 # cleaning ----------------------------------------------------------------
 clean_workspace()

### 2020-04-09 R/qu_phch_compile.R

@@ -1,17 +1,14 @@
-# script aggregates organism data from 3rd party data bases
+# script aggregates physico-chemical data from 3rd party data bases
 
 # setup -------------------------------------------------------------------
 source(file.path(src, 'gn_setup.R'))
 
 # scripts -----------------------------------------------------------------
-q = read_char(file.path(sql, 'taxa_fin.sql'))
+q = read_char(file.path(sql, 'phch_compile.sql'))
 
 # query -------------------------------------------------------------------
 drv = dbDriver("PostgreSQL")
 con = dbConnect(drv, user = DBuser, dbname = DBetox, host = DBhost, port = DBport, password = DBpassword)
-
-dbSendQuery(con, "DROP SCHEMA IF EXISTS taxa_fin CASCADE;")
-dbSendQuery(con, "CREATE SCHEMA taxa_fin;")
 
 dbSendQuery(con, q)
 
@@ -19,7 +16,9 @@ dbDisconnect(con)
 dbUnloadDriver(drv)
 
 # log ---------------------------------------------------------------------
-log_msg('QUERY: taxa final table created.')
+log_msg('QUERY: phch final table created.')
 
 # cleaning ----------------------------------------------------------------
 clean_workspace()
+  
+  

### 2020-04-09 R/qu_taxa_compile.R

@@ -1,10 +1,10 @@
-# script to build application data
+# script aggregates organism data from 3rd party data bases
 
 # setup -------------------------------------------------------------------
 source(file.path(src, 'gn_setup.R'))
 
 # scripts -----------------------------------------------------------------
-q = read_char(file.path(sql, 'conv_unit_result.sql')) # WORK IN PROGRESS
+q = read_char(file.path(sql, 'taxa_compile.sql'))
 
 # query -------------------------------------------------------------------
 drv = dbDriver("PostgreSQL")
@@ -16,8 +16,7 @@ dbDisconnect(con)
 dbUnloadDriver(drv)
 
 # log ---------------------------------------------------------------------
-log_msg('CONV: Unit conversion scripts run.')
+log_msg('QUERY: taxa final table created.')
 
 # cleaning ----------------------------------------------------------------
 clean_workspace()
-

### 2020-04-09 article/R/fig_standartox_ppdb_rx_comparison_dwld.R

@@ -38,9 +38,9 @@ q_ppdb = "SELECT
 ppdb = read_query(user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
                   query = q_ppdb)
 # Read across
-q_rx = "select cas, lc50_dm_rx
-        from phch.phch_variables
-        left join phch.phch_variables_prop using (variable_id)"
+q_rx = "SELECT cas, lc50_dm_rx
+        FROM phch.phch_variables
+        LEFT join phch.phch_variables_prop using (variable_id)"
 rx = read_query(user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = 'bfg_monitoring',
                 query = q_rx)
 

### 2020-04-09 article/R/fig_standartox_ppdb_rx_comparison_prep.R

@@ -1,5 +1,4 @@
 # script to compare STANDARTOX and ppdb results (prepare)
-# TODO clean up script
 
 # setup -------------------------------------------------------------------
 source(file.path(src, 'gn_setup.R'))
@@ -26,16 +25,16 @@ stan2 = stan[ conc_unit_stan %in% c('ug/l', 'ppb'),
 stan2 = na.omit(stan2)
 # PPDB
 ppdb2 = ppdb[ endpoint %in% c('LC50', 'LD50', 'EC50') ]
-ppdb2[, cas := gsub('-', '', cas) ]
 # ECOSAR
 rx2 = rx[ cas != '' & !is.na(lc50_dm_rx) ]
-rx2[ , cas := gsub('-', '', cas) ]
 rx2[ , taxon := 'Daphnia magna' ]
 rx2[ , dur := 48 ]
 setnames(rx2, 'lc50_dm_rx', 'conc_rx')
 
 # merge -------------------------------------------------------------------
 dat = merge(stan2, ppdb2, by = c('cas', 'taxon', 'dur'), suffixes = c('_stan', '_ppdb'))
+taxa = dat[ , .N, taxon ][ order(-N) ][ 1:8, taxon ] # 8 most ocurring taxa
+dat = dat[ taxon %in% taxa ]
 dat[ , diff := abs(conc_stan_gm / conc_ppdb) ]
 dat_dm = Reduce(function(...) merge(..., by = c('cas', 'taxon', 'dur'), all = F),
                 list(stan2, ppdb2, rx2))
@@ -45,7 +44,7 @@ dat_dm[ , diff := abs(conc_stan_gm / conc_rx) ]
 ## continuous vs continuous
 # NOTE color palette taken from: https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/
 gg1 = ggplot(dat, aes(y = conc_stan_gm, x = conc_ppdb, col = taxon)) +
-  geom_point() +
+  geom_point(size = 2.5) +
   geom_abline() +
   geom_abline(intercept = 1, col = 'red') +
   geom_abline(intercept = -1, col = 'red') +
@@ -61,11 +60,11 @@ gg1 = ggplot(dat, aes(y = conc_stan_gm, x = conc_ppdb, col = taxon)) +
         legend.title = element_blank())
 
 gg2 = ggplot(dat_dm, aes(y = conc_stan_gm, x = conc_rx, col = taxon)) +
[…]

### 2020-04-09 run_build.R

@@ -70,8 +70,8 @@ if (report) {
 
 # merge tables ------------------------------------------------------------
 if (build) {
-  source(file.path(src, 'qu_phch_fin.R'), max.deparse.length = mdl)
-  source(file.path(src, 'qu_taxa_fin.R'), max.deparse.length = mdl)
+  source(file.path(src, 'qu_phch_compile.R'), max.deparse.length = mdl)
+  source(file.path(src, 'qu_taxa_compile.R'), max.deparse.length = mdl)
 }
 
 # Lookup ------------------------------------------------------------------
@@ -82,7 +82,7 @@ if (lookup) {
 }
 
 # unit conversion ---------------------------------------------------------
-source(file.path(src, 'conv_unit_result.R'), max.deparse.length = mdl) # converts result units
+source(file.path(src, 'conv_unit_result_duration.R'), max.deparse.length = mdl)
 
 # Standartox --------------------------------------------------------------
 if (build_standartox) {

### 2020-04-09 run_build_TEST.R

@@ -16,16 +16,16 @@ if (sink_console) {
 }
 
 # SCRIPT TO TEST ----------------------------------------------------------
-# source(file.path(src, 'rep_query_prep.R'), max.deparse.length = mdl)
-# source(file.path(src, 'qu_taxa_fin.R'), max.deparse.length = mdl) # TODO error  GREATEST(wo.brack, epa_habi.brackish)::boolean AS brack
-# source(file.path(src, 'qu_phch_fin.R'), max.deparse.length = mdl)
-# source(file.path(src, 'look_concentration_units.R'), max.deparse.length = mdl)
-# source(file.path(src, 'look_duration_units.R'), max.deparse.length = mdl)
+source(file.path(src, 'id_compile_table.R'), max.deparse.length = mdl)
+source(file.path(src, 'qu_taxa_compile.R'), max.deparse.length = mdl)
+source(file.path(src, 'qu_phch_compile.R'), max.deparse.length = mdl)
 source(file.path(src, 'look_unit_result.R'), max.deparse.length = mdl)
-source(file.path(src, 'conv_unit_result.R'), max.deparse.length = mdl) # converts result units
+source(file.path(src, 'look_unit_duration.R'), max.deparse.length = mdl)
+source(file.path(src, 'conv_unit_result_duration.R'), max.deparse.length = mdl) # converts result units
 source(file.path(src, 'bd_standartox.R'), max.deparse.length = mdl)
 source(file.path(src, 'exp_standartox.R'), max.deparse.length = mdl)
 source(file.path(src, 'exp_standartox_catalog.R'), max.deparse.length = mdl)
+source(file.path(src, 'cpy_standartox.R'), max.deparse.length = mdl)
 source(file.path(src, 'gn_setup.R'))
 con = DBI::dbConnect(RPostgreSQL::PostgreSQL(), #RPostgres::Postgres(),
                      dbname = DBetox,

### 2020-04-10 article/article.tex

@@
 %=================================================================
-\documentclass[journal,datadescriptor,accept,moreauthors,pdftex]{Definitions/mdpi}
+\documentclass[journal,datadescriptor,submit,moreauthors,pdftex]{Definitions/mdpi}
 
@@
 The PPDB provides ecotoxicity data on a few selected species from chemical risk assessment that have been manually quality controlled through expert judgement \citep{lewis_international_2016}.
-The vast majority of aggregated values (91~\%) of Standartox lie within one order of magnitude of the corresponding PPDB values (n = 3601).
-This would increase to 92.7~\%, when restricting the comparison to Standartox values where data from at least five experiments are available.
+The vast majority of aggregated values (91.9~\%) of Standartox lie within one order of magnitude of the corresponding PPDB values (n = 3601).
+This would increase to 92.6~\%, when restricting the comparison to Standartox values where data from at least five experiments are available.
 Similarly, we compared Standartox to ecotoxicity values for \textit{D.
@@
 However, the Standartox results for individual taxa-chemical combinations could easily be aggregated across chemicals in a second step to provide a similar aggregation as that performed in EnviroTox.
-Since Standartox aims to provide toxicity estimations for specific taxa and chemicals we refrain from aggregating further.
 \par
@@
 through conversion of test concentration and duration units.
-1237 distinct concentration units are converted to X harmonized ones, as are the 126 distinct duration units to X.
-Standartox retains only duration units that can unambiguously be converted to hours.
+1237 distinct concentration units are converted to six harmonized ones (i.e.
+g/l, g/m2, ppb, g/g, l/l, l/m2) when conversion is possible.
+Likewise, the 126 distinct duration units are converted to hours whenever this is unambiguously possible.
 To guarantee appropriate unit conversion and harmonisation, we compared the results of an automated unit conversion to a manual one for each of the distinct concentration and duration units.
@@
 Other endpoints, such as Bioconcentration factors, non-half maximal effective concentrations (e.g.
-IC10, EC25, LD99) or maximum acceptable toxicant concentrations are removed.
+IC\textsubscript{10}, EC\textsubscript{25}, LD\textsubscript{99}) or maximum acceptable toxicant concentrations are removed.
 Along with that, a catalog, listing all distinct entries and value ranges, for categorical and continuous variables, respectively, is created.

### 2020-04-10 article/article.tex

@@
 \usepackage{standalone}
+\usepackage{listings}
+\usepackage{xcolor}
+\usepackage{rotating}
+\usepackage{tikzpeople}
+\usepackage{makecell}
+\usepackage{refcheck} % debuging
 % code
-\usepackage{listings}
-\lstset{language=R,
-    basicstyle=\small\ttfamily,
-    stringstyle=\color{DarkGreen},
-    otherkeywords={0,1,2,3,4,5,6,7,8,9},
-    morekeywords={TRUE,FALSE},
-    deletekeywords={data,frame,length,as,character},
-    keywordstyle=\color{blue},
-    commentstyle=\color{DarkGreen},
+\lstset{
+    language = R,
+    basicstyle = \small\ttfamily,
+    stringstyle = \color{orange},
+    % key words
+    otherkeywords = {0,1,2,3,4,5,6,7,8,9},
+    morekeywords = {TRUE,FALSE},
+    keywordstyle = \color{blue},
+    % comments
+    commentstyle = \color{darkgray},
+    morecomment = [f][\color{red}][0]{\#},
+    % frame
+    frame = single,
+    backgroundcolor = \color{white}
 }
 % tikz
-\usepackage{tikzpeople}
 \usetikzlibrary{trees,
@@
                 fit}
-\usepackage{rotating}
-% check references
-% \usepackage{refcheck} % debuging
 
@@
 Given that most current databases lack aggregation steps or are confined to specific chemicals, we developed Standartox, a tool and database that continuously incorporates the ever-growing number of test results in an automated process workflow that ultimately leads to a single aggregated data point for a specific chemical-organism test combination, representing the toxicity of a chemical.
-Standartox comes with two front-ends, a web application (standartox.uni-landau.de) and an R package (standartox).}
+Standartox comes with two front-ends, a web application \url{standartox.uni-landau.de} and an R package \textit{standartox}.}
 
@@
 In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}
+\dataset{\url{ftp://newftp.epa.gov/ecotox/ecotox_ascii_12_12_2019.exe}}
 
@@
 The PPDB provides ecotoxicity data on a few selected species from chemical risk assessment that have been manually quality controlled through expert judgement \citep{lewis_international_2016}.
-The vast majority of aggregated values (91.9~\%) of Standartox lie within one order of magnitude of the corresponding PPDB values (n = 3601).
+The vast majority of aggregated values (91.9~\%) of Standartox lie within one order of magnitude of the corresponding PPDB values (n~=~3601).
 This would increase to 92.6~\%, when restricting the comparison to Standartox values where data from at least five experiments are available.
@@
 magna} from the ChemProp \citep{ufzdepartmentofecologicalchemistry_chemprop_2019} software, which estimates LC\textsubscript{50} values via quantitative structure-activity relationship (QSAR) models \citep{schuurmann_quantitative_2011}.
-We found that 95~\% of Standartox values lie within one order of magnitude of the Chemprop (n = 179) values.
+We found that 95~\% of Standartox values lie within one order of magnitude of the ChemProp (n~=~179) values.
 However, the diﬀerence is not necessarily an indication of lower quality of Standartox estimates but may also reflect the wider range of experimental conditions for which data are available in the database underlying Standartox as well as inaccurate predictions for QSAR models, respectively (Figure~\ref{fig:standartox_ppdb_diff}).
@@
 Besides, EnviroTox provides additional information on toxicity endpoints, such as acute or chronic classifications and mode of action assignments.
-We intentionally omitted such classifications given that the approach to classification may vary with the purpose of the study.
+We intentionally omitted such classifications given that the approach to classification may vary with the purpose of the study or because of different classification schemes \citep{kienzler_mode_2017}.
 The EnviroTox database allows for an aggregation into single toxicity values for individual taxa, whereas Standartox performs this aggregation for individual taxa-chemical combinations.
@@
 The \textbf{Etox} database collects ecotoxicity test information and provides methods to filter those.
-Like the ECOTOX database, it also doesn't provide methods to perform aggregations of the ecotoxicity data and the database is not accessible in automated way.
-In contrast to the ECOTOX database it is not possible to download the whole database.
+Like the ECOTOX database, it also lacks methods to perform aggregations of the ecotoxicity data and only provides manual (non-automated) access.
+In contrast to the ECOTOX database, the Etox database can not be downloaded.
 \par
@@
 Then, provided Chemical Abstracts Service Numbers (CAS) and taxonomic names are used to query additional information from publicly available databases on chemicals and organisms, respectively.
-This includes the Compendium of Pesticide Common Names \citep{wood_compendium_2019}, the Chemical Entities of Biological Interest (ChEBI) database \citep{hastings_chebi_2016}, the Chemical Identifier Resolver (CIR) service \citep{nationalinstitutesofhealthnih_chemical_2019}, the Pubchem database \citep{kim_pubchem_2016}, Eurostat \citep{europeancommission_eurostat_2019} and Wikidata \citep{vrandecic_wikidata_2014} for chemicals and the World Register of Marine Species (WoRMS) \citep{wormseditorialboard_world_2018} and the Global Biodiversity Information Facility (GBIF) \citep{gbiftheglobalbiodiversityinformationfacility_what_2020} (Table~\ref{tab:data-base-additional}) for habitat and spatial distribution of organisms.
+This includes the Compendium of Pesticide Common Names \citep{wood_compendium_2019}, the Chemical Entities of Biological Interest (ChEBI) database \citep{hastings_chebi_2016}, the Chemical Identifier Resolver (CIR) service \citep{nationalinstitutesofhealthnih_chemical_2019}, the Pubchem database \citep{kim_pubchem_2016}, Eurostat \citep{europeancommission_eurostat_2019} and Wikidata \citep{vrandecic_wikidata_2014} for chemicals and the World Register of Marine Species (WoRMS) \citep{wormseditorialboard_world_2020} and the Global Biodiversity Information Facility (GBIF) \citep{gbiftheglobalbiodiversityinformationfacility_what_2020} (Table~\ref{tab:data-base-additional}) for habitat and spatial distribution of organisms.
 Given that taxonomic names can be ambiguous, e.g.
@@
 The user can then call the functions \textit{stx\_filter()} and \textit{stx\_aggregate()} that filter and aggregate the data according to specific parameters (Table~\ref{tab:app-parameters}).
-The interactive web application is built in R using the shiny framework, which runs with the help of a shiny server \citep{chang_shiny_2018} on \url{standartox.uni-landau.de}.
-The API is built by using the R package \textit{plumber} \citep{trestletechnologyllc_plumber_2018}, which allows for the creation of Representational State Transfer (REST) APIs from R.
+The interactive web application is built in R using the shiny framework, which runs with the help of a shiny server \citep{R-shiny} on \url{standartox.uni-landau.de}.
+The API is built by using the R package \textit{plumber} \citep{R-plumber}, which allows for the creation of Representational State Transfer (REST) APIs from R.
 REST is a software architectural style that defines web service communication rules.
@@
 The \textit{/catalog} API-endpoint returns a JavaScript Object Notation (JSON) file containing a catalog of possible filter parameters to choose from.
-The \textit{/filter} returns the filtered Standartox table as a compressed serialized binary file created by the R \textit{fst} package \citep{klik_fst_2019}, to reduce size and allow for fast user queries.
+The \textit{/filter} returns the filtered Standartox table as a compressed serialized binary file created by the R \textit{fst} package \citep{R-fst}, to reduce size and allow for fast user queries.
 The \textit{/aggregate} API-endpoint returns an R function allowing the client to aggregate the filtered data locally.
@@
 \begin{table}[H]
-    \caption{Input parameters for the Standartox web application and the R-package standartox (CAS - Chemical Abstracts Service Registry Number, NOEX and XX\textsubscript{50} - Standartox endpoints \ref{tab:endpoints-conflate}), vers - Standartox version).}
+    \caption{Input parameters for the Standartox web application and the R-package \textit{standartox} (CAS - Chemical Abstracts Service Registry Number, NOEX and XX\textsubscript{50} - Standartox endpoints \ref{tab:endpoints-conflate}), vers - Standartox version).}
     \label{tab:app-parameters}
@@
 \begin{lstlisting}[
+    language = R,
     caption = {Sample code to access the Standartox database through the API and the the R-package \textit{standartox}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''.
-Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}
+\funding{This research was funded by the German Environment Agency (UBA) grant number 3714 67 4040/2.}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\acknowledgments{The authors want to thank Eduard Szöcs whose blog post (https://edild.github.io/localecotox/) on how to build a local version of the EPA ECOTOX database was the starting point for this project.}
+\acknowledgments{The authors thank Eduard Szöcs for inspiration through a blog post on ho to build a local EPA Ecotox database.}
 
@@
 PPDB & Pesticides Properties DataBase \\
-QSAR & Quantitative structure–activity relationship \\
+QSAR & Quantitative structure-activity relationship \\
 REST & Representational state transfer software architectural style \\
@@
     \toprule
-    \textbf{Endpoint} & \textbf{Http Method} & \textbf{Request} & \textbf{Response} \\
+    \textbf{Endpoint} & \textbf{HTTP Method} & \textbf{Request} & \textbf{Response} \\
     \midrule
@@
 \reftitle{References}
-% \nocite{*} % debuging
+\nocite{*} % debuging
 \externalbibliography{yes}

### 2020-04-10 article/tikz/standartox-organigram.tex
submission

@@
 %% processing
-\node (epa-ecotox) [ecotox, above of = ecotox-id, align=center, xshift = -3cm, yshift = 4cm] {\Large EPA ECOTOX\\\Large raw data};
-\node (ecotox-id) [standartox, align = center] at (-3cm,0cm) {\Large Identifiers};
+\node (ecotox-id) [standartox, align = center] at (0cm,0cm) {\Large Identifiers};
+\node (epa-ecotox) [ecotox, above of = ecotox-id, align=center, xshift = 0cm, yshift = 4cm] {\Large EPA ECOTOX\\\Large raw data};
 \node (ecotox-unit) [standartox, below of = ecotox-id, align = center, yshift = -2cm] {\Large Unit\\\Large harmonisation};

### 2020-04-10 article/article.tex
Updates from Overleaf

@@
 \usepackage{makecell}
-\usepackage{refcheck} % debuging
+% \usepackage{refcheck} % debuging
 % code
@@
 We found that 95~\% of Standartox values lie within one order of magnitude of the ChemProp (n~=~179) values.
-However, the diﬀerence is not necessarily an indication of lower quality of Standartox estimates but may also reflect the wider range of experimental conditions for which data are available in the database underlying Standartox as well as inaccurate predictions for QSAR models, respectively (Figure~\ref{fig:standartox_ppdb_diff}).
+However, the difference is not necessarily an indication of lower quality of Standartox estimates but may also reflect the wider range of experimental conditions for which data are available in the database underlying Standartox as well as inaccurate predictions for QSAR models, respectively (Figure~\ref{fig:standartox_ppdb_diff}).
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\acknowledgments{The authors thank Eduard Szöcs for inspiration through a blog post on ho to build a local EPA Ecotox database.}
+\acknowledgments{The authors thank Eduard Szöcs for inspiration through a blog post on how to build a local version of the EPA Ecotox database.}
 
@@
 \reftitle{References}
-\nocite{*} % debuging
+% \nocite{*} % debuging
 \externalbibliography{yes}

### 2020-05-02 article/article.tex
submission
Updates from Overleaf

@@
 % Authors, for the paper (add full first names)
-\Author{Andreas Scharmüller $^{1}$*\orcidA{}, Verena C. Schreiner$^{1}$\orcidB and Ralf B. Schäfer $^{1}$\orcidC{}}
+\Author{Andreas Scharmüller $^{1,2}$*\orcidA{}, Verena C. Schreiner$^{1}$\orcidB and Ralf B. Schäfer $^{1}$\orcidC{}}
 
@@
 $^{1}$ \quad iES Landau, Institute for Environmental Sciences, University of Koblenz-Landau,
-D-76829 Landau, Germany; andschar@protonmail.com (A.S.), schreiner-verena@uni-landau.de (V.C.S.), schaefer-ralf@uni-landau.de (R.B.S.)}
+D-76829 Landau, Germany; andschar@protonmail.com (A.S.), schreiner-verena@uni-landau.de (V.C.S.), schaefer-ralf@uni-landau.de (R.B.S.)\\
+$^{2}$ \quad Laboratoire d’Hydrologie et de Géochimie de Strasbourg, UMR 7517 CNRS-Université de Strasbourg, 1 rue Blessig, 67084 Strasbourg Cedex, France}
 
@@
 To evaluate the risks from chemicals for ecosystems, data on their toxicity, which is typically produced in standardized ecotoxicological laboratory tests, is required.
-The results from ecotoxicological tests are compiled in (Meta-)databases such as the  United States Environmental Protection Agency ECOTOXicology knowledgebase.
+The results from ecotoxicological tests are compiled in (Meta-)databases such as the United States Environmental Protection Agency (EPA) ECOTOXicology knowledgebase (ECOTOX).
 However, for many chemicals, multiple ecotoxicity data are available for the same test organism.
@@
 Given that most current databases lack aggregation steps or are confined to specific chemicals, we developed Standartox, a tool and database that continuously incorporates the ever-growing number of test results in an automated process workflow that ultimately leads to a single aggregated data point for a specific chemical-organism test combination, representing the toxicity of a chemical.
-Standartox comes with two front-ends, a web application \url{standartox.uni-landau.de} and an R package \textit{standartox}.}
+Standartox comes with two front-ends, the web application \url{standartox.uni-landau.de} and the R package \textit{standartox}.}
 
@@
 
-To date, only few initiatives exist that aim to create a public resource of ecotoxicological data, such as the United States Environmental Protection Agency ECOTOXicology knowledgebase (ECOTOX) (ca.
+To date, only few initiatives exist that aim to create a public resource of ecotoxicological data, such as the United States Environmental (EPA) Protection Agency ECOTOXicology knowledgebase (ECOTOX) (ca.
 1,000,000 test results, 13,000 taxa, 12,000 chemicals) \citep{elonen_ecotoxicology_2018}, the German Environmental Agency's Information System Ecotoxicology and Environmental Quality Targets (ETOX) \citep{umweltbundesamt_etox_2019}, the Pesticides Properties DataBase (PPDB) (ca.
@@
 Besides aggregating ecotoxicological test results, Standartox provides a concise overview of the tested chemicals, allowing the identification of potential knowledge gaps.
-Moreover, Standartox could help in reducing the millions of animals used for toxicity testing each year by facilitating access to ecotoxicity data \citep{hartung_chemical_2009}.
+Moreover, Standartox could help in reducing the millions of animals used for toxicity testing each year by facilitating access to ecotoxicity data, which is in favor of for example the guidelines by the Organisation for Economic Co-operation and Development (OECD) \citep{oecd_oecd_2020, hartung_chemical_2009}.
 Standartox comes with two front-ends, a web application: \url{standartox.uni-landau.de} and the R \citep{rcoreteam_language_2017} package \textit{standartox}, providing convenience structures and thereby largely reducing processing time for users.
@@
 Also, the geometric mean is preferable over the median, because the median completely ignores the tails of the data distribution, making it unreliable for small data sets \citep{leith_comparison_2010}.
+Posthuma et al.
+\citet{posthuma_species_2019} showed the usefulness of SSDs and its underlying geometric mean aggregations when assessing environmental effects of chemicals.
 In the course of the aggregation process, outliers that exceed 1.5 times the interquartile range are flagged to caution Standartox users.
-However, they are considered in aggregation, given that the geometric mean is relatively robust against outliers.
+However, they are considered in the aggregation, given that the geometric mean is relatively robust against outliers.
 Overall, Standartox provides a harmonized and reproducible approach to data aggregation that may contribute to reduce between-study variability.
@@
 \textit{Daphnia magna}), and the tool lacks the possibility for automated data retrieval \citep{williams_comptox_2017}.
+\textbf{Comptox} is built on the \textbf{Aggregated Computational Toxicology Resource (ACToR)} database, which constitutes the basis for several applications published by the EPA.
+It collects physicochemical and toxicological data on more than 500,000 environmental chemicals and pharmaceutical compounds from various resources and presents them in a curated list on the web \citep{judson_actor_2008, judson_aggregating_2012}.
+However, no filter mechanisms or aggregation methods are provided in \textbf{ACToR} per se.
 \par
@@
 To guarantee appropriate unit conversion and harmonisation, we compared the results of an automated unit conversion to a manual one for each of the distinct concentration and duration units.
+This assures that 652 of the 1237 units  (95.3 \% of the data) are converted correctly.
+The remainder could not be converted.
 Furthermore, the units are cleaned, for example through removing additional information in the field such as \textit{food}, \textit{soil}, \textit{ai} that are also coded in other variables and hinder the processing of units.
@@
 Finally, we run quality control scripts that check the accuracy of the data.
+
+CONTINUE HERE
+
 The compiled Standartox data set together with the catalog is exported and accessible via the web application and the API, through the R package.
@@
 Standartox meets these requirements and puts its focus on the aggregation of toxicity data, thereby adding a piece to the puzzle of modern ecotoxicological data analyses.
+The Standartox database will be maintained by the first author and the Institute of Environmental Sciences, Landau, Germany.
 

### 2020-05-04 Rmd/rep_conv_unit_result_duration.Rmd
release 0.1

---
title: "Unit conversion report"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      fig.width = 15, fig.height = 8)
```

```{r data, echo=FALSE, include=FALSE}
## converted units
q = "SELECT *
     FROM ecotox.results2"
units = read_query(user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
                   query = q)
## conversion check
chck_unit_result_conversion = readRDS(file.path(cachedir, 'chck_unit_result_conversion.rds'))
chck_unit_duration_conversion = readRDS(file.path(cachedir, 'chck_unit_duration_conversion.rds'))
```

## Result units

### Distinct units reduction
```{r}
res_n = units[ , lapply(.SD, uniqueN), .SDcols = c('conc1_unit', 'conc1_unit2', 'conc1_unit3', 'conc1_unit4')]
ggplot(melt(res_n), aes(y = value, x = variable)) +
  geom_bar(stat = 'identity')
```

Number of tests removed b/c of removed units: `r tab = table(units$conc1_remove); round(tab[2] / nrow(units) * 100, 1)`

### Number of tests per kept original units
Area is on `log10` scale.

```{r}
res_dist = units[ conc1_remove == FALSE, .N, conc1_unit ]
res_dist[ , label := paste0(conc1_unit, '\n', N) ]
ggplot(res_dist, aes(area = log10(N), label = label)) +
  geom_treemap() +
  geom_treemap_text(col = 'white')
```

### Number of tests per removed original units
Area is on `log10` scale.

```{r}
res_dist = units[ conc1_remove == TRUE, .N, conc1_unit ]
res_dist[ , label := paste0(conc1_unit, '\n', N) ]
ggplot(res_dist, aes(area = log10(N), label = label)) +
  geom_treemap() +
  geom_treemap_text(col = 'white')
```

### Number of tests per kept converted units
Area is on `log10` scale.

```{r}
res_dist = units[ conc1_remove == FALSE, .N, conc1_unit4 ]
res_dist[ , label := paste0(conc1_unit4, '\n', N) ]
ggplot(res_dist, aes(area = log10(N), label = label)) +
  geom_treemap() +
  geom_treemap_text(col = 'white')
```

### Number of tests per removed converted units
Area is on `log10` scale.

```{r}
res_dist_rm = units[ conc1_remove == TRUE, .N, conc1_unit4 ]
res_dist_rm[ , label := paste0(conc1_unit4, '\n', N) ]
ggplot(res_dist_rm, aes(area = log10(N), label = label)) +
  geom_treemap() +
  geom_treemap_text(col = 'white')
```

### Conversion check

#### All units
```{r}
chck_res = chck_unit_result_conversion[ , .N, .(chck_mean2, chck_unit2, chck_mean4, chck_unit4) ]
m = melt(chck_res, measure.vars = c('chck_mean2', 'chck_unit2', 'chck_mean4', 'chck_unit4'))
m[ , chck := fifelse(value == TRUE, 'correct', fifelse(value == FALSE, 'wrong', NA_character_)) ]
m[ is.na(chck), chck := 'no chck\nvalue' ]
# plot
ggplot(m, aes(y = N, x = variable, group = chck, fill = chck, label = N)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  geom_text(position = position_dodge(width = 0.9),
            hjust = 1.2, col = 'white', size = 4) +
  scale_fill_manual(values = c('springgreen1', 'salmon1', 'firebrick4')) +
    labs(title = 'Standartox unit conversion check',
         subtitle = 'All units',
         x = NULL) +
  coord_flip()
```

#### Only not removed units
```{r}
chck_res_nrm = chck_unit_result_conversion[ conc1_remove != TRUE,
                                            .N,
                                            .(chck_mean2, chck_unit2, chck_mean4, chck_unit4) ]
m = melt(chck_res_nrm,
         measure.vars = c('chck_mean2', 'chck_unit2', 'chck_mean4', 'chck_unit4'))
m[ , chck := fifelse(value == TRUE, 'correct', fifelse(value == FALSE, 'wrong', NA_character_)) ]
m[ is.na(chck), chck := 'no chck\nvalue' ]
# plot
ggplot(m, aes(y = N, x = variable, group = chck, fill = chck, label = N)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  geom_text(position = position_dodge(width = 0.9),
            hjust = 1.2, col = 'white', size = 4) +
  scale_fill_manual(values = c('springgreen1', 'salmon1', 'firebrick4')) +
  labs(title = 'Standartox unit conversion check',
       subtitle = 'Only kept units',
       x = NULL) +
  coord_flip()
```

```{r paper-review, eval=FALSE}
result_id_excl = chck_unit_result_conversion[ conc1_unit4 %in% c('g/l', 'g/g', 'g/m2', 'ppb', 'l/l', 'l/m2') ][ chck_unit2 == FALSE | is.na(chck_unit2), result_id ]
chck_paper = chck_unit_result_conversion[ conc1_unit4 %in% c('g/l', 'g/g', 'g/m2', 'ppb', 'l/l', 'l/m2') &
                                            !result_id %in% result_id_excl ,
                                          .N,
                                          .(chck_unit2, chck_unit4) ]
conc1_unit_ = chck_unit_result_conversion[ conc1_unit4 %in% c('g/l', 'g/g', 'g/m2', 'ppb', 'l/l', 'l/m2') &
                                             !result_id %in% result_id_excl,
                                           conc1_unit ]

q = "SELECT conc1_unit, count(*) n
 FROM ecotox.results2
 GROUP BY conc1_unit
 ORDER BY n DESC"
dat = read_query(user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
                 query = q)
dat[ , kept := fifelse(conc1_unit %in% conc1_unit_, TRUE, FALSE) ]
dat2 = dat[ , .(sum(n)), kept]
dat2$V1[1] / sum(dat2$V1)

m = melt(chck_paper,
         measure.vars = c('chck_unit2', 'chck_unit4'))
m[ , chck := fifelse(value == TRUE, 'correct', fifelse(value == FALSE, 'wrong', NA_character_)) ]
m[ is.na(chck), chck := 'no chck\nvalue' ]
# plot
ggplot(m, aes(y = N, x = variable, group = chck, fill = chck, label = N)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  geom_text(position = position_dodge(width = 0.9),
            hjust = 1.2, col = 'white', size = 4) +
  scale_fill_manual(values = c('springgreen1', 'salmon1', 'firebrick4')) +
  labs(title = 'Standartox unit conversion check',
       subtitle = 'Only kept units',
       x = NULL) +
  coord_flip()
```

## Duration units

```{r}
dur_n = units[ , lapply(.SD, uniqueN), .SDcols = c('obs_duration_unit', 'obs_duration_unit2') ]
ggplot(melt(dur_n), aes(y = value, x = variable)) +
  geom_bar(stat = 'identity')
```

### Number of tests per kept original units
Area is on `log10` scale.

```{r}
dur_dist = units[ conc1_remove == FALSE, .N, obs_duration_unit ]
dur_dist[ , label := paste0(obs_duration_unit, '\n', N) ]
ggplot(dur_dist, aes(area = log10(N), label = label)) +
  geom_treemap() +
  geom_treemap_text(col = 'white')
```

### Number of tests per removed original units
Area is on `log10` scale.

```{r}
dur_dist = units[ conc1_remove == TRUE, .N, obs_duration_unit ]
dur_dist[ , label := paste0(obs_duration_unit, '\n', N) ]
ggplot(dur_dist, aes(area = log10(N), label = label)) +
  geom_treemap() +
  geom_treemap_text(col = 'white')
```

### Number of tests per kept converted units
Area is on `log10` scale.

```{r}
dur_dist = units[ conc1_remove == FALSE, .N, obs_duration_unit2 ]
dur_dist[ , label := paste0(obs_duration_unit2, '\n', N) ]
ggplot(dur_dist, aes(area = log10(N), label = label)) +
  geom_treemap() +
  geom_treemap_text(col = 'white')
```

### Number of tests per removed converted units
Area is on `log10` scale.

```{r}
dur_dist_rm = units[ conc1_remove == TRUE, .N, obs_duration_unit2 ]
dur_dist_rm[ , label := paste0(obs_duration_unit2, '\n', N) ]
ggplot(dur_dist_rm, aes(area = log10(N), label = label)) +
  geom_treemap() +
  geom_treemap_text(col = 'white')
```

### Conversion check

#### All units
```{r}
chck_dur = chck_unit_duration_conversion[ , .N, .(chck_mean2, chck_unit2) ]
m = melt(chck_dur, measure.vars = c('chck_mean2', 'chck_unit2'))
m[ , chck := fifelse(value == TRUE, 'correct', fifelse(value == FALSE, 'wrong', NA_character_)) ]
m[ is.na(chck), chck := 'no chck\nvalue' ]
# plot
ggplot(m, aes(y = N, x = variable, fill = chck, label = N)) +
  geom_bar(stat = 'identity', position = position_dodge()) +
  geom_text(position = position_dodge(width = 0.9), hjust = 1.2, col = 'white', size = 4) +
  scale_fill_manual(values = c('springgreen1', 'salmon1', 'firebrick4')) +
  labs(x = NULL) +
  coord_flip()
```

```{r plot, eval=FALSE, fig.width=10, fig.height=30}
# TODO plot is way to large
# ## 1
# test = units[ , .(n = .N), conc1_unit4 ]
# ggplot(test, aes(area = n, label = conc1_unit4)) +
#   geom_treemap() +
#   geom_treemap_text(col = 'white')
# 
# 
# # ## 2
# require(ggalluvial)
# 
# dat = units[ conc1_remove == TRUE, .N, .(conc1_unit, conc1_unit2, conc1_unit3, conc1_unit4, conc1_remove) ]
# setorder(dat, -N)
# 
# gg_alluvium = ggplot(dat,
#        aes(axis1 = reorder(conc1_unit, -N),
#            axis2 = reorder(conc1_unit2, -N), # axis3 = conc1_unit3, axis4 = conc1_unit4,
#            y = N, fill = conc1_remove)) +
#   geom_alluvium() +
#   geom_stratum(width = 1/12, fill = "black", color = "grey") +
#   geom_label(stat = "stratum", infer.label = TRUE)
#   
# ggsave(gg_alluvium, filename = file.path('/tmp/alluvium.svg'),
#        width = 20, height = 60, dpi = 90, limitsize = FALSE)

```

### 2020-05-04 R/chck_habitat.R

@@ -1,0 +1,83 @@
+# script that check taxa for appropriate habitat classifications
+# TODO elaborate on this
+# TODO maybe use testthat:: ?
+
+# setup -------------------------------------------------------------------
+source(file.path(src, 'gn_setup.R'))
+
+# freshwater --------------------------------------------------------------
+tax_fresh = c('Astacus astacus', 'Elodea canadensis', 'Danio rerio', 'Gammarus fossarum', 'Gammarus pulex')
+
+# marine ------------------------------------------------------------------
+tax_marin = c('Cerastoderma edule', 'Homarus americanus', 'Posidonie oceanica')
+fam_marin = c('Triglidae')
+phylum_division_marin = c('Echinodermata')
+
+# terrestrial -------------------------------------------------------------
+tax_terre = c('Apis mellifera', 'Rattus norvegicus', 'Zea mays')
+
+# query -------------------------------------------------------------------
+# fresh
+q = paste0("SELECT *
+            FROM standartox.taxa
+            WHERE tax_taxon IN ('",
+           paste0(tax_fresh, collapse = "', '"),
+           "');")
+fresh = read_query(user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
+                   query = q)
+# marin
+q = paste0("SELECT *
+            FROM standartox.taxa
+            WHERE tax_taxon IN ('",
+           paste0(tax_marin, collapse = "', '"),
+           "')
+           OR tax_family IN ('",
+           paste0(fam_marin, collapse = "', '"),
+           "')
+           OR tax_phylum_division IN ('",
+           paste0(phylum_division_marin, collapse = "', '"),
+           "');")
+marin = read_query(user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
+                   query = q)
+# terre
+q = paste0("SELECT *
+            FROM standartox.taxa
+            WHERE tax_taxon IN ('",
+           paste0(tax_terre, collapse = "', '"),
+           "');")
+terre = read_query(user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
+                   query = q)
+
+# chck --------------------------------------------------------------------
+# fresh
+chck_equals(all(fresh$hab_freshwater), TRUE)
+chck_equals(any(fresh$hab_terrestrial), TRUE)
+# marin
+chck_equals(all(marin$hab_marine), TRUE)
+# terre
[…]

### 2020-05-04 R/chck_unit_result_conversion_DOINEEDTHIS.R

@@ -1,0 +1,132 @@
+# script to check concentration and duraiton unit converisons
+# NOTE use package testthat in the future
+
+# setup -------------------------------------------------------------------
+source(file.path(src, 'gn_setup.R'))
+
+## CONTINUE HERE!!!!!!!!!!!!!!!!!!!!!!!!!
+
+# data --------------------------------------------------------------------
+# units_chck = fread(file.path(lookupdir, 'concentration_unit_lookup_chck.csv'),
+#                    na.strings = '')
+# cols = c('result_id', 'conc1_mean', 'conc1_unit', 'conc1_mean_manual', 'conc1_unit_manual')
+# units_chck = units_chck[ , .SD, .SDcols = cols ]
+# units_chck2 = copy(units_chck)
+# todo = c('ug/l' = 'g/l',
+#          'mg/kg' = 'g/g',
+#          'g/m2' = 'g/m2',
+#          'ul/l' = 'l/l',
+#          'ml/g' = 'l/g',
+#          'g/cm' = 'g/m',
+#          'Bq/ml' = 'Bq/l',
+#          'Bq/mg' = 'Bq/g',
+#          'ml/m2' = 'l/m2',
+#          'g/d' = 'g/h',
+#          'g/d' = 'g/h')
+# for (i in seq_along(todo)) {
+#   from = names(todo)[i]
+#   to = todo[i]
+#   units_chck2[ conc1_unit_manual == from, `:=`
+#                (conc1_mean_manual = ud.convert(conc1_mean_manual, from, to),
+#                  conc1_unit_manual = to) ]
+# }
+# units_chck2[ , .N, conc1_unit_manual ][ order(-N) ]
+# # units_chck2[ , .N, conc1_unit_manual2 ]
+# 
+# fwrite(units_chck2, file.path(lookupdir, 'concentration_unit_lookup_chck2.csv'))
+# 
+
+units_chck = fread(file.path(lookupdir, 'concentration_unit_lookup_chck.csv'),
+                   na.strings = '')
+units_chck = units_chck[ , .SD,
+                         .SDcols = c('result_id',
+                                     # 'conc1_mean', 'conc1_unit', 'conc1_mean2', 'conc1_unit2',
+                                     'conc1_mean2_manual', 'conc1_unit2_manual',
[…]

### 2020-05-04 article/article.tex

@@
 Red dots depict Standartox geometric mean estimates and black dots show raw data for each group.
+Red error bars show the standard deviation of the geometric mean.
 To facilitate readability, data points are randomly scattered along a hypothetical y-axis and are greyed out if within the violins.}
@@
 Then, provided Chemical Abstracts Service Numbers (CAS) and taxonomic names are used to query additional information from publicly available databases on chemicals and organisms, respectively.
-This includes the Compendium of Pesticide Common Names \citep{wood_compendium_2019}, the Chemical Entities of Biological Interest (ChEBI) database \citep{hastings_chebi_2016}, the Chemical Identifier Resolver (CIR) service \citep{nationalinstitutesofhealthnih_chemical_2019}, the Pubchem database \citep{kim_pubchem_2016}, Eurostat \citep{europeancommission_eurostat_2019} and Wikidata \citep{vrandecic_wikidata_2014} for chemicals and the World Register of Marine Species (WoRMS) \citep{wormseditorialboard_world_2020} and the Global Biodiversity Information Facility (GBIF) \citep{gbiftheglobalbiodiversityinformationfacility_what_2020} (Table~\ref{tab:data-base-additional}) for habitat and spatial distribution of organisms.
+This includes the Compendium of Pesticide Common Names \citep{wood_compendium_2019}, the Chemical Entities of Biological Interest (ChEBI) database \citep{hastings_chebi_2016}, the Chemical Identifier Resolver (CIR) service \citep{nationalinstitutesofhealthnih_chemical_2019}, the Pubchem database \citep{kim_pubchem_2016}, Eurostat \citep{europeancommission_eurostat_2019} and Wikidata \citep{vrandecic_wikidata_2014} for chemicals and the World Register of Marine Species (WoRMS) \citep{wormseditorialboard_world_2020}, the Global Biodiversity Information Facility (GBIF) \citep{gbiftheglobalbiodiversityinformationfacility_what_2020} and the freshwaterecology.info database \citep{schmidt-kloiber_www_2015} (Table~\ref{tab:data-base-additional}) for habitat and spatial distribution of organisms.
 Given that taxonomic names can be ambiguous, e.g.
@@
 This assures that 652 of the 1237 units (95.3 \% of the data) are converted correctly.
-The remainder could not be converted.
+The remainder could not be converted and is removed.
 Furthermore, the units are cleaned, for example through removing additional information in the field such as \textit{food}, \textit{soil}, \textit{ai} that are also coded in other variables and hinder the processing of units.
@@
 Along with that, a catalog, listing all distinct entries and value ranges, for categorical and continuous variables, respectively, is created.
-Finally, we run quality control scripts that check the accuracy of the data.
-
-CONTINUE HERE
-
 The compiled Standartox data set together with the catalog is exported and accessible via the web application and the API, through the R package.
@@
 The API is reachable via the URL \url{http://139.14.20.252} and port 8000.
-Four API-endpoints (\textit{/catalog}, \textit{/filter}, \textit{/aggregate} and \textit{/meta}) can be queried (Table~\ref{tab:api-endpoints}).
+Three API-endpoints (\textit{/catalog}, \textit{/filter}, and \textit{/meta}) can be queried (Table~\ref{tab:api-endpoints}).
 The \textit{/catalog} API-endpoint returns a JavaScript Object Notation (JSON) file containing a catalog of possible filter parameters to choose from.
 The \textit{/filter} returns the filtered Standartox table as a compressed serialized binary file created by the R \textit{fst} package \citep{R-fst}, to reduce size and allow for fast user queries.
-The \textit{/aggregate} API-endpoint returns an R function allowing the client to aggregate the filtered data locally.
 Lastly, the \textit{/meta} API-endpoint returns a JSON file with meta information, such as the timestamp of the request and the used Standartox version.
@@
     \makecell{World Register of \\ Marine Species (WoRMS)} & \url{http://marinespecies.org} \\
+    freshwaterecology.info & \url{www.freshwaterecology.info} \\
     \bottomrule
@@
     /filter & POST & Standartox filter parameters & Filtered Standartox data (serialized) \\
-    /aggregate & GET & / & Aggregate function (serialized) \\
     /meta & POST & Standartox version string & Meta data on request (JSON) \\
